---
title: "CD4_misc"
output: html_document
date: "2025-05-28"
author: "David Zemmour"
editor_options: 
  chunk_output_type: console
---

**Initialize: load libraries and custom functions**

```{bash}
sinteractive --exclusive --time=12:00:00 --account=pi-zemmour --partition=jfkfloor2 --qos=jfkfloor2 --nodes=1 
sinteractive --exclusive --time=12:00:00 --account=pi-zemmour --partition=beagle3-bigmem --nodes=1 
sinteractive --exclusive --time=12:00:00 --account=pi-zemmour --partition=zemmour-hm --qos=zemmour --nodes=1 
sinteractive --time=03:00:00 --account=pi-zemmour --partition=caslake --nodes=1 --mem=96GB

module load python/anaconda-2022.05 
module load hdf5/1.12.0 #for BPCells but also others
module load openblas/0.3.13
source activate /project/zemmour/david/envs/scvi_20240315
#source activate /project/zemmour/david/envs/r443
module load openblas/0.3.13

cd /project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/CD4
```


```{r}
setwd("~/google_drive_uchicago/ImmgenT/analysis/data_integration/IGT1_96/CD4")
libs = c("ZemmourLib","forcats", "gridExtra","patchwork", "Matrix", "Seurat", "ggplot2", "dplyr", "reshape2", "ggrastr","scattermore", "RColorBrewer", "pals", "scales", "pheatmap", "rafalib") 
sapply(libs, function(x) suppressMessages(suppressWarnings(library(x, character.only = TRUE, quietly = T, warn.conflicts  = F))))
#source("/project/jfkfloor2/zemmourlab/david/immgent/immgent_integration_git/custom_functions_david.R")

ColorRamp = rev(colorRampPalette(brewer.pal(n = 7,name = "RdYlBu"))(100))
ColorRamp = rev(colorRampPalette(c('red','white','blue'))(20))
ColorRamp = rev(rainbow(10, end = 4/6))
ColorRamp = rev(viridis(100))
ColorRamp = brewer.pal(9, "Blues")

#mycol = c("black","blue","red","darkgreen", "orange","purple", "cyan", "greenyellow",    "salmon", "magenta","pink", "tan", "brown") # c("magenta", "red",  "darkgreen", "cyan", "blue", "blue", "blue", "black", "blue", "blue", "blue", "blue", "orange")

mypal_level2 = ZemmourLib::immgent_colors$level2
mypal_level2["not classified"] = "black"
# names(mypal_level2) = gsub("nonconv", replacement = "Tz",names(mypal_level2) )
# mypal_level2 = immgent_colors$level2_option2
# mypal_level2["not classified"] = "black"
mypal_level1 = immgent_colors$level1
mypal_level1["not classified"] = "black"
# names(mypal_level1) = gsub("nonconv", replacement = "Tz",names(mypal_level1) )
mypal_organ = immgent_colors$organ_simplified
mypal_level2group = c("resting" = "blue", "activated" = "red", "miniverse" = "darkgreen", "proliferating" = "black", "other" = "grey", "preT" = "grey" )

#Large Color palette
library("pals")
n = 70
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
mypal1 = unique(unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals))))
mypal1 = mypal1[-4]
mypal = c(glasbey(), polychrome(), mypal1)
names(mypal) = NULL
#pal.bands(mypal, main="Colormap suggestions")
# mypal_organ = setNames(mypal, unique(so$organ_simplified))
# mypal_organ = mypal_organ[!is.na(names(mypal_organ))]
# mypal_level1 = setNames(mypal, unique(so$annotation_level1))
# mypal_level1 = mypal_level1[!is.na(names(mypal_level1))]
# mypal_level2 = setNames(mypal, unique(so_orig$annotation_level2))
# mypal_level2 = mypal_level2[!is.na(names(mypal_level2))]
# mypal_level2["CD4_cl18"] = "blue"
mypal_igt = setNames(mypal, unique(so_orig$IGT))
mypal_igt = mypal_igt[!is.na(names(mypal_igt))]

```

All mapped datasets are here: '/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/20250109_freeze/odhran/CD4'
```{r}
# fig_dir = "~/google_drive_uchicago/ImmgenT/20250109_freeze/paper_cosmology/figures/bits"
# fig_dir = "/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/paper_cosmology/figures/bits"

out_dir = "~/google_drive_uchicago/ImmgenT/analysis/data_integration/IGT1_96/CD4/TRBI" 
out_dir = "/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/CD4/TRBI"

out_dir2 = "~/google_drive_uchicago/ImmgenT/analysis/data_integration/IGT1_96/CD4/misc_plots" 
out_dir2 = "/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/CD4/misc_plots"
```


**Most up-to-date data**
on Midway
```{r}
so_orig = readRDS("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/igt1_96_withtotalvi20260206_clean.Rds") 

so = so_orig[,so_orig@meta.data %>% filter(organ_simplified != "thymus") %>% row.names()]
so = NormalizeData(so, assay = "RNA", normalization.method = "LogNormalize")


```

on Desktop
```{r}
# setwd("~/google_drive_uchicago/ImmgenT/analysis/data_integration/IGT1_96/")
so_orig = readRDS("../igt1_96_withtotalvi20260206_clean_ADTonly.Rds")

so = so_orig[,so_orig@meta.data %>% filter(annotation_level1 == "CD4" & organ_simplified != "thymus") %>% row.names()]
so$annotation_level2
# so = NormalizeData(so, assay = "RNA", normalization.method = "LogNormalize")
# so = NormalizeData(so, assay = "ADT", normalization.method = "LogNormalize")

```

**T-RBI Inspecting individual query studies**
see here for https://docs.google.com/spreadsheets/d/10icsN1kaUNOgMYi-oVTFx8AH9CI_52SlQb3p798MfQY/edit?usp=sharing

Background CD4 immgenT
(part not needed anymore using the CD4 immgent object)
```{r}
immgent_colors = readRDS('/Users/david/Library/CloudStorage/GoogleDrive-david.puyraimond@gmail.com/My Drive/ImmgenT/analysis/immgent_integration_git/immgent_color_palettes/20250709/immgent_color_palettes_list_20250709.Rds')
mypal_level2 = immgent_colors$CD4

so_orig = readRDS("igt1_96_withtotalvi20250710_clean_ADTonly.Rds")
so_orig[["mde_incremental_level2"]] = so_orig[["mde_incremental"]] 
so_orig[["mde_incremental_allT"]] = so_orig[["mde2_totalvi_20241006"]] 

so = so_orig[,so_orig$annotation_level1 == "CD4"]
so@meta.data %>% count(annotation_level2)

# sub = so@meta.data %>% group_by(annotation_level2) %>% slice_sample(n = 10000) %>% ungroup() %>% pull(cellID)
# so = so[,sub]
so$sample = "immgent"

# DimPlot(so, reduction = "mde_incremental_level2", group.by = "annotation_level2", cols = mypal_level2)
# MyDimPlotHighlight(so, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "annotation_level2", mycols = mypal_level2, print_plot1 = T, print_plot2 = T,labelclusters = T)

```

mypal_level2 for old annotation
```{r}
# immgent_colors = readRDS('/Users/david/Library/CloudStorage/GoogleDrive-david.puyraimond@gmail.com/My Drive/ImmgenT/analysis/immgent_integration_git/immgent_color_palettes/20250709/immgent_color_palettes_list_20250709.Rds')
# mypal_level2 = immgent_colors$CD4

```

https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE281547 - Kevin/Jenkins - Listeria, spleen

```{r}
query_dir = '/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/20250109_freeze/odhran/CD4/GSE281547_Jenkins_b6_aLm/Trial_Final/'
so_query = readRDS('/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/20250109_freeze/odhran/CD4/GSE281547_Jenkins_b6_aLm/Trial_Final/Final_Seurat_object.Rds') 

so_query$sample_name[so_query$sample_name == "Sample-2_D06"] = "Lm_D06"
so_query$sample_name[so_query$sample_name == "Sample-3_D21"] = "Lm_D21"
so_query$sample_name[so_query$sample_name == "Sample-1_D60"] = "Lm_D60"

so_query$cellID = colnames(so_query)
sub = so_query@meta.data %>% group_by(sample_name) %>% slice_sample(n = 4500) %>% ungroup() %>% pull(cellID)
so_query = so_query[,sub]

mypal_annotation = mypal
mypal_annotation = setNames(mypal_annotation, unique(so_query$annotations))

# umap_mat <- as.matrix( so_query@meta.data[, c("umaprna_1", "umaprna_2")])
# colnames(umap_mat) <- c("umappaper_1", "umappaper_2")
# rownames(umap_mat) <- rownames(so_query@meta.data)
# so_query[["umap_paper"]] <- CreateDimReducObject(
#   embeddings = umap_mat,
#   key = "UMAPPAPER_",
#   assay = DefaultAssay(so_query)
# )

head(so_query@meta.data)
# table(so_query$cluster)
# table(so_query$level2_plot)
# table(so_query$batch)
# table(so_query$tissue)
# table(so_query$annotation, so_query$tissue)

# table(so_query$batch_sample)
# table(so_query$level2_plot, so_query$cluster)
# table(so_query$level2_plot, so_query$level1_final)
# so_query$annotation = paste("cl", so_query$annotation, sep = "_")

so_query@meta.data %>% count(sample_name)
so_query@meta.data %>% count(level2_final)

# so_query$sample = as.integer(sub('.*-(\\d+)$', '\\1', colnames(so_query)))

sub = so_query@meta.data %>% filter(!(level2_final %in% c("not classified"))) %>% row.names()

DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "annotations", cols = mypal_annotation)

p1 = DimPlot(so_query[,sub], reduction = "mde_incremental_level2", group.by = "annotations", cols = mypal_annotation) 
p$layers <- c(list(immgenTCD4MDE()), p1$layers)
p

pdf(sprintf("%s/GSE281547_umap_dataset.pdf",query_dir),14, 7, useDingbats = F)
p1 = DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "annotations", cols = mypal_annotation)
p2 = DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "level2_final", cols = mypal_level2)
p1 + p2
DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "annotations", cols = mypal_annotation, split.by = "sample_name")
DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "level2_final", cols = mypal_level2, split.by = "sample_name")
# FeaturePlot(so_query[,sub], reduction = "mde_incremental_level2", features = "Pdcd1")
dev.off()

so2 = merge(so, so_query[,sub], merge.dr = T)

# mypal_annotation = setNames(c("#5D7AA5", "#BED4EB", "#E59339", "#F5CA8F", "#6FA649", "#B9E185", "#B29A39", "#EFD374", "#649B9F", "#A9CBC2", "#D2635A", "#F3B0B0"), c("cl_0", "cl_1", "cl_2", "cl_3", "cl_4", "cl_5", "cl_6", "cl_7", "cl_8", "cl_9", "cl_10", "cl_11"))
# mypal_annotation = mypal

pdf(sprintf("%s/GSE281547_RBI.pdf",query_dir),7, 7, useDingbats = F)
MyDimPlotHighlight(so2, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "level2_final", mycols = mypal_level2, print_plot1 = T, print_plot2 = T,labelclusters = F, xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = T, highlight_size = 1)
MyDimPlotHighlight(so2, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "annotations", mycols = mypal_annotation, print_plot1 = T, print_plot2 = T,labelclusters = F, xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = T, highlight_size = 1)

# MyDimPlotHighlight(so2, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "annotation", mycols = mypal_annotation, print_plot1 = T, print_plot2 = T,labelclusters = F)
for (s in unique(so_query$sample_name)) {
    print(s)
    tmp = so2[,so2$sample_name %in% c(s) | so2$sample %in% c("immgent")]
    print(MyDimPlotHighlight(tmp, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "level2_final", mycols = mypal_level2, print_plot1 = T, print_plot2 = T, labelclusters = T, xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = F, highlight_size = 0.5, title = sprintf("sample %s", s)))
    print(MyDimPlotHighlight(tmp, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "annotations", mycols = mypal_annotation, print_plot1 = T, print_plot2 = T,labelclusters = T, , xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = F, highlight_size = 0.5, title = sprintf("sample %s", s)))
    # print(MyDimPlotHighlight(so2[,so2$sample %in% c("immgent", s)], umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "annotation", mycols = mypal_annotation, print_plot1 = T, print_plot2 = T,labelclusters = T, title = sprintf("sample: %s", s)))
}

# p1 = DimPlot(so_query[,sub], reduction = "mde_incremental_level2", group.by = "annotation", cols = mypal_annotation)
# p2 = DimPlot(so_query[,sub], reduction = "mde_incremental_level2", group.by = "level2_plot", cols = mypal_level2)
# p1 + p2
dev.off()


##DotPlot

library(dplyr)
library(ggplot2)
library(scales)

df_dot <- so_query@meta.data %>%
  count(annotations, level2_final, name = "n") %>%
  group_by(annotations) %>%
  mutate(
    prop = n / sum(n),
    logn = log10(n + 1)
  ) %>%
  ungroup() %>%
  mutate(
    # keep the order as they appear in the data
    annotations  = factor(annotations,  levels = unique(annotations)),
    level2_final = factor(level2_final, levels = unique(level2_final))
  )
df_dot$level2_final = factor(df_dot$level2_final, levels = c(levels(so$annotation_level2), "not classified"))

p = ggplot(df_dot, aes(x = level2_final, y = annotations)) +
  geom_point(aes(size = prop, color = n)) +
  scale_size_continuous(labels = percent_format(accuracy = 1)) +
  scale_color_viridis_c(option = "magma", name = "log10(n+1)", direction =  -1) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 60, hjust = 1),
    panel.grid.major = element_line(color = "grey90")
  ) +
  labs(
    x = "level2_final",
    y = "annotations",
    size = "% within annotation"
  )
pdf(sprintf("%s/DotPlot_PaperAnnot_immgenTAnnot.pdf",query_dir),7, 5, useDingbats = F)
p
dev.off()


```

https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE182320 - Andreatta/Carmona - LCMV arm and cl13, spleen

```{r}
query_dir = '/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/20250109_freeze/odhran/CD4/GSE182320_IAb_Tet_WT_Zbtb7bKO_LCMV/Trial_Final/'
so_query = readRDS('/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/20250109_freeze/odhran/CD4/GSE182320_IAb_Tet_WT_Zbtb7bKO_LCMV/Trial_Final/Final_Seurat_object.Rds') 

so_query@meta.data %>% colnames()
so_query@meta.data %>% count(Sample, sample_name)

so_query$sample_name = so_query$Sample
so_query$annotations = so_query$functional.cluster
so_query@meta.data %>% count(annotations)
so_query@meta.data %>% count(level2_final)

so_query$sample_type = NA
so_query$sample_type[grepl(x = so_query$sample_name, pattern = "Arm_d7")] = "Arm_d7"
so_query$sample_type[grepl(x = so_query$sample_name, pattern = "Arm_d21")] = "Arm_d21"
so_query$sample_type[grepl(x = so_query$sample_name, pattern = "Arm_d60")] = "Arm_d60"
so_query$sample_type[grepl(x = so_query$sample_name, pattern = "Cl13_d7")] = "Cl13_d7"
so_query$sample_type[grepl(x = so_query$sample_name, pattern = "Cl13_d21")] = "Cl13_d21"
so_query@meta.data %>% count(sample_name, sample_type)
so_query@meta.data %>% count(sample_type)

so_query$cellID = colnames(so_query)
sub = so_query@meta.data %>% group_by(sample_type) %>% slice_sample(n = 3913) %>% ungroup() %>% pull(cellID)
so_query = so_query[,sub]

mypal_annotation = mypal
mypal_annotation = setNames(mypal_annotation, unique(so_query$annotations))

# umap_mat <- as.matrix( so_query@meta.data[, c("umaprna_1", "umaprna_2")])
# colnames(umap_mat) <- c("umappaper_1", "umappaper_2")
# rownames(umap_mat) <- rownames(so_query@meta.data)
# so_query[["umap_paper"]] <- CreateDimReducObject(
#   embeddings = umap_mat,
#   key = "UMAPPAPER_",
#   assay = DefaultAssay(so_query)
# )

head(so_query@meta.data)
# table(so_query$cluster)
# table(so_query$level2_plot)
# table(so_query$batch)
# table(so_query$tissue)
# table(so_query$annotation, so_query$tissue)

# table(so_query$batch_sample)
# table(so_query$level2_plot, so_query$cluster)
# table(so_query$level2_plot, so_query$level1_final)
# so_query$annotation = paste("cl", so_query$annotation, sep = "_")

so_query@meta.data %>% count(sample_name)
so_query@meta.data %>% count(level2_final)

# so_query$sample = as.integer(sub('.*-(\\d+)$', '\\1', colnames(so_query)))

sub = so_query@meta.data %>% filter(!(level2_final %in% c("not classified"))) %>% row.names()

pdf(sprintf("%s/GSE182320_umap_dataset.pdf",query_dir),14, 7, useDingbats = F)
p1 = DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "annotations", cols = mypal_annotation)
p2 = DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "level2_final", cols = mypal_level2)
p3 = DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "sample_name", cols = mypal)
p1 + p2 + p3
DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "annotations", cols = mypal_annotation, split.by = "sample_type")
DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "level2_final", cols = mypal_level2, split.by = "sample_type")
# FeaturePlot(so_query[,sub], reduction = "mde_incremental_level2", features = "Pdcd1")
dev.off()

so2 = merge(so, so_query[,sub], merge.dr = T)

# mypal_annotation = setNames(c("#5D7AA5", "#BED4EB", "#E59339", "#F5CA8F", "#6FA649", "#B9E185", "#B29A39", "#EFD374", "#649B9F", "#A9CBC2", "#D2635A", "#F3B0B0"), c("cl_0", "cl_1", "cl_2", "cl_3", "cl_4", "cl_5", "cl_6", "cl_7", "cl_8", "cl_9", "cl_10", "cl_11"))
# mypal_annotation = mypal

pdf(sprintf("%s/GSE182320_RBI.pdf",query_dir),7, 7, useDingbats = F)
MyDimPlotHighlight(so2, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "level2_final", mycols = mypal_level2, print_plot1 = T, print_plot2 = T,labelclusters = F, xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = T, highlight_size = 1)
MyDimPlotHighlight(so2, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "annotations", mycols = mypal_annotation, print_plot1 = T, print_plot2 = T,labelclusters = F, xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = T, highlight_size = 1)
MyDimPlotHighlight(so2, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "sample_type", mycols = mypal, print_plot1 = T, print_plot2 = T,labelclusters = F, xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = T, highlight_size = 1)


# MyDimPlotHighlight(so2, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "annotation", mycols = mypal_annotation, print_plot1 = T, print_plot2 = T,labelclusters = F)
for (s in unique(so_query$sample_type)) {
    print(s)
    tmp = so2[,so2$sample_type %in% c(s) | so2$sample %in% c("immgent")]
    print(MyDimPlotHighlight(tmp, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "level2_final", mycols = mypal_level2, print_plot1 = T, print_plot2 = T, labelclusters = T, xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = F, highlight_size = 0.5, title = sprintf("sample %s", s)))
    print(MyDimPlotHighlight(tmp, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "annotations", mycols = mypal_annotation, print_plot1 = T, print_plot2 = T,labelclusters = T, , xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = F, highlight_size = 0.5, title = sprintf("sample %s", s)))
    # print(MyDimPlotHighlight(so2[,so2$sample %in% c("immgent", s)], umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "annotation", mycols = mypal_annotation, print_plot1 = T, print_plot2 = T,labelclusters = T, title = sprintf("sample: %s", s)))
}

# p1 = DimPlot(so_query[,sub], reduction = "mde_incremental_level2", group.by = "annotation", cols = mypal_annotation)
# p2 = DimPlot(so_query[,sub], reduction = "mde_incremental_level2", group.by = "level2_plot", cols = mypal_level2)
# p1 + p2
dev.off()

##DotPlot

library(dplyr)
library(ggplot2)
library(scales)

df_dot <- so_query@meta.data %>%
  count(annotations, level2_final, name = "n") %>%
  group_by(annotations) %>%
  mutate(
    prop = n / sum(n),
    logn = log10(n + 1)
  ) %>%
  ungroup() %>%
  mutate(
    # keep the order as they appear in the data
    annotations  = factor(annotations,  levels = unique(annotations)),
    level2_final = factor(level2_final, levels = unique(level2_final))
  )
df_dot$level2_final = factor(df_dot$level2_final, levels = c(levels(so$annotation_level2), "not classified"))

p = ggplot(df_dot, aes(x = level2_final, y = annotations)) +
  geom_point(aes(size = prop, color = n)) +
  scale_size_continuous(labels = percent_format(accuracy = 1)) +
  scale_color_viridis_c(option = "magma", name = "log10(n+1)", direction =  -1) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 60, hjust = 1),
    panel.grid.major = element_line(color = "grey90")
  ) +
  labs(
    x = "level2_final",
    y = "annotations",
    size = "% within annotation"
  )
pdf(sprintf("%s/DotPlot_PaperAnnot_immgenTAnnot.pdf",query_dir),7, 5, useDingbats = F)
p
dev.off()

```

https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE121002 - LCMV arm D7 and D30 (in WT and Zbtb7bKO), spleen - Ciucci/Bosselut paper
No paper annotation

```{r}
query_dir = '/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/20250109_freeze/odhran/CD4/GSE121002_IAb_Tet_binding_Listeria/Not_classified_Full_Trial/Not_classified_loop_fixed/'
so_query = readRDS('/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/20250109_freeze/odhran/CD4/GSE121002_IAb_Tet_binding_Listeria/Not_classified_Full_Trial/Not_classified_loop_fixed/Final_Seurat_object.Rds') 

so_query@meta.data %>% colnames()
head(so_query@meta.data)
so_query@meta.data %>% count(sample_name, level1_final)

so_query@meta.data %>% filter(sample_name %in% c("Batch1_GSM3423796_d7_GP66", "Batch1_GSM3423797_d7_WT", "Batch1_GSM3423799_d30_WT") & level1_final == "CD4") %>% count(sample_name)

so_query$sample_type = NA
so_query$sample_type[grepl(x = so_query$sample_name, pattern = "d7")] = "Arm_d7"
so_query$sample_type[grepl(x = so_query$sample_name, pattern = "d30")] = "Arm_d30"
so_query@meta.data %>% count(sample_name, sample_type)
so_query@meta.data %>% count(sample_type)

so_query$cellID = colnames(so_query)
sub = so_query@meta.data %>% filter(sample_name %in% c("Batch1_GSM3423796_d7_GP66", "Batch1_GSM3423797_d7_WT", "Batch1_GSM3423799_d30_WT") & level1_final == "CD4") %>% group_by(sample_type) %>% slice_sample(n = 471) %>% ungroup() %>% pull(cellID)
so_query = so_query[,sub]
so_query@meta.data %>% count(sample_name, sample_type)

# so_query$sample_name = so_query$Sample
# so_query$annotations = so_query$functional.cluster
# so_query@meta.data %>% count(annotations)
# so_query@meta.data %>% count(level2_final)

mypal_annotation = mypal
# mypal_annotation = setNames(mypal_annotation, unique(so_query$annotations))

# umap_mat <- as.matrix( so_query@meta.data[, c("umaprna_1", "umaprna_2")])
# colnames(umap_mat) <- c("umappaper_1", "umappaper_2")
# rownames(umap_mat) <- rownames(so_query@meta.data)
# so_query[["umap_paper"]] <- CreateDimReducObject(
#   embeddings = umap_mat,
#   key = "UMAPPAPER_",
#   assay = DefaultAssay(so_query)
# )

head(so_query@meta.data)
# table(so_query$cluster)
# table(so_query$level2_plot)
# table(so_query$batch)
# table(so_query$tissue)
# table(so_query$annotation, so_query$tissue)

# table(so_query$batch_sample)
# table(so_query$level2_plot, so_query$cluster)
# table(so_query$level2_plot, so_query$level1_final)
# so_query$annotation = paste("cl", so_query$annotation, sep = "_")

so_query@meta.data %>% count(sample_name)
so_query@meta.data %>% count(level2_final)

# so_query$sample = as.integer(sub('.*-(\\d+)$', '\\1', colnames(so_query)))

sub = so_query@meta.data %>% filter(!(level2_final %in% c("CD4"))) %>% row.names()

pdf(sprintf("%s/GSE121002_umap_dataset.pdf",query_dir),14, 7, useDingbats = F)
# p1 = DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "annotations", cols = mypal_annotation)
p2 = DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "level2_final", cols = mypal_level2)
p3 = DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "sample_name", cols = mypal)
# p1 + p2 + p3
p2 + p3
# DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "annotations", cols = mypal_annotation, split.by = "sample_type")
DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "level2_final", cols = mypal_level2, split.by = "sample_name")
# FeaturePlot(so_query[,sub], reduction = "mde_incremental_level2", features = "Pdcd1")
dev.off()

so2 = merge(so, so_query[,sub], merge.dr = T)

# mypal_annotation = setNames(c("#5D7AA5", "#BED4EB", "#E59339", "#F5CA8F", "#6FA649", "#B9E185", "#B29A39", "#EFD374", "#649B9F", "#A9CBC2", "#D2635A", "#F3B0B0"), c("cl_0", "cl_1", "cl_2", "cl_3", "cl_4", "cl_5", "cl_6", "cl_7", "cl_8", "cl_9", "cl_10", "cl_11"))
# mypal_annotation = mypal

pdf(sprintf("%s/GSE121002_RBI.pdf",query_dir),7, 7, useDingbats = F)
MyDimPlotHighlight(so2, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "level2_final", mycols = mypal_level2, print_plot1 = T, print_plot2 = T,labelclusters = F, xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = F, highlight_size = 1)
# MyDimPlotHighlight(so2, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "annotations", mycols = mypal_annotation, print_plot1 = T, print_plot2 = T,labelclusters = F, xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = T, highlight_size = 1)
MyDimPlotHighlight(so2, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "sample_type", mycols = mypal, print_plot1 = T, print_plot2 = T,labelclusters = F, xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = F, highlight_size = 1)


# MyDimPlotHighlight(so2, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "annotation", mycols = mypal_annotation, print_plot1 = T, print_plot2 = T,labelclusters = F)
for (s in unique(so_query$sample_type)) {
    print(s)
    tmp = so2[,so2$sample_type %in% c(s) | so2$sample %in% c("immgent")]
    print(MyDimPlotHighlight(tmp, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "level2_final", mycols = mypal_level2, print_plot1 = T, print_plot2 = T, labelclusters = T, xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = F, highlight_size = 0.5, title = sprintf("sample %s", s)))
    # print(MyDimPlotHighlight(tmp, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "annotations", mycols = mypal_annotation, print_plot1 = T, print_plot2 = T,labelclusters = T, , xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = F, highlight_size = 0.5, title = sprintf("sample %s", s)))
    # print(MyDimPlotHighlight(so2[,so2$sample %in% c("immgent", s)], umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "annotation", mycols = mypal_annotation, print_plot1 = T, print_plot2 = T,labelclusters = T, title = sprintf("sample: %s", s)))
}

# p1 = DimPlot(so_query[,sub], reduction = "mde_incremental_level2", group.by = "annotation", cols = mypal_annotation)
# p2 = DimPlot(so_query[,sub], reduction = "mde_incremental_level2", group.by = "level2_plot", cols = mypal_level2)
# p1 + p2
dev.off()



```

https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE158896 - LCMV arm D10 spleen, CD4+CD44+GP66+
No paper annotation

```{r}
query_dir = '/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/20250109_freeze/odhran/CD4/GSE158896_IAb_Tet_LCMV_d35/Trial_Final/'
so_query = readRDS('/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/20250109_freeze/odhran/CD4/GSE158896_IAb_Tet_LCMV_d35/Trial_Final/Final_Seurat_object.Rds') 

so_query@meta.data %>% colnames()
head(so_query@meta.data)
so_query@meta.data %>% count(sample_name, level1_final)

so_query$sample_name = so_query$sample
# so_query$annotations = so_query$functional.cluster
# so_query@meta.data %>% count(annotations)
# so_query@meta.data %>% count(level2_final)

mypal_annotation = mypal
# mypal_annotation = setNames(mypal_annotation, unique(so_query$annotations))head(so_query)

# umap_mat <- as.matrix( so_query@meta.data[, c("umaprna_1", "umaprna_2")])
# colnames(umap_mat) <- c("umappaper_1", "umappaper_2")
# rownames(umap_mat) <- rownames(so_query@meta.data)
# so_query[["umap_paper"]] <- CreateDimReducObject(
#   embeddings = umap_mat,
#   key = "UMAPPAPER_",
#   assay = DefaultAssay(so_query)
# )

head(so_query@meta.data)
# table(so_query$cluster)
# table(so_query$level2_plot)
# table(so_query$batch)
# table(so_query$tissue)
# table(so_query$annotation, so_query$tissue)

# table(so_query$batch_sample)
# table(so_query$level2_plot, so_query$cluster)
# table(so_query$level2_plot, so_query$level1_final)
# so_query$annotation = paste("cl", so_query$annotation, sep = "_")

so_query@meta.data %>% count(sample_name)
so_query@meta.data %>% count(level2_final)

# so_query$sample = as.integer(sub('.*-(\\d+)$', '\\1', colnames(so_query)))

sub = so_query@meta.data %>% filter(!(level2_final %in% c("not classified"))) %>% row.names()

pdf(sprintf("%s/GSE158896_umap_dataset.pdf",query_dir),14, 7, useDingbats = F)
# p1 = DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "annotations", cols = mypal_annotation)
p2 = DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "level2_final", cols = mypal_level2)
p3 = DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "sample_name", cols = mypal)
# p1 + p2 + p3
p2 + p3
# DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "annotations", cols = mypal_annotation, split.by = "sample_type")
DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "level2_final", cols = mypal_level2, split.by = "sample_name")
# FeaturePlot(so_query[,sub], reduction = "mde_incremental_level2", features = "Pdcd1")
dev.off()

so2 = merge(so, so_query[,sub], merge.dr = T)

# mypal_annotation = setNames(c("#5D7AA5", "#BED4EB", "#E59339", "#F5CA8F", "#6FA649", "#B9E185", "#B29A39", "#EFD374", "#649B9F", "#A9CBC2", "#D2635A", "#F3B0B0"), c("cl_0", "cl_1", "cl_2", "cl_3", "cl_4", "cl_5", "cl_6", "cl_7", "cl_8", "cl_9", "cl_10", "cl_11"))
# mypal_annotation = mypal

pdf(sprintf("%s/GSE121002_RBI.pdf",query_dir),7, 7, useDingbats = F)
MyDimPlotHighlight(so2, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "level2_final", mycols = mypal_level2, print_plot1 = T, print_plot2 = T,labelclusters = F, xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = T, highlight_size = 1)
# MyDimPlotHighlight(so2, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "annotations", mycols = mypal_annotation, print_plot1 = T, print_plot2 = T,labelclusters = F, xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = T, highlight_size = 1)
# MyDimPlotHighlight(so2, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "sample_type", mycols = mypal, print_plot1 = T, print_plot2 = T,labelclusters = F, xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = F, highlight_size = 1)


# MyDimPlotHighlight(so2, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "annotation", mycols = mypal_annotation, print_plot1 = T, print_plot2 = T,labelclusters = F)
# for (s in unique(so_query$sample_type)) {
#     print(s)
#     tmp = so2[,so2$sample_type %in% c(s) | so2$sample %in% c("immgent")]
#     print(MyDimPlotHighlight(tmp, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "level2_final", mycols = mypal_level2, print_plot1 = T, print_plot2 = T, labelclusters = T, xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = F, highlight_size = 0.5, title = sprintf("sample %s", s)))
    # print(MyDimPlotHighlight(tmp, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "annotations", mycols = mypal_annotation, print_plot1 = T, print_plot2 = T,labelclusters = T, , xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = F, highlight_size = 0.5, title = sprintf("sample %s", s)))
    # print(MyDimPlotHighlight(so2[,so2$sample %in% c("immgent", s)], umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "annotation", mycols = mypal_annotation, print_plot1 = T, print_plot2 = T,labelclusters = T, title = sprintf("sample: %s", s)))
# }

# p1 = DimPlot(so_query[,sub], reduction = "mde_incremental_level2", group.by = "annotation", cols = mypal_annotation)
# p2 = DimPlot(so_query[,sub], reduction = "mde_incremental_level2", group.by = "level2_plot", cols = mypal_level2)
# p1 + p2
dev.off()



```

https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE198466 - lung CD4+ tet+ against Blastomyces
No paper annotation
Mice were inoculated either intranasally (IN) or subcutaneously (SC) with Blastomyces dermatitidis endoglucanase-2 (Bl-Eng2) six weeks prior to pathogen challenge. Tetramer-positive CD4+ T cells were FACS sorted and sequenced 3 days after the pulmonary challenge with B. dermatitidis

```{r}
query_dir = '/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/20250109_freeze/odhran/CD4/GSE198466_Vaccinated/Trial_Final/'
so_query = readRDS('/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/20250109_freeze/odhran/CD4/GSE198466_Vaccinated/Trial_Final/Final_Seurat_object.Rds') 

so_query@meta.data %>% colnames()
head(so_query@meta.data)
so_query$sample_name = so_query$original
so_query@meta.data %>% count(sample_name)

# so_query$sample_type = NA
# so_query$sample_type[grepl(x = so_query$sample_name, pattern = "d7")] = "Arm_d7"
# so_query$sample_type[grepl(x = so_query$sample_name, pattern = "d30")] = "Arm_d30"
# so_query@meta.data %>% count(sample_name, sample_type)
# so_query@meta.data %>% count(sample_type)

so_query$cellID = colnames(so_query)
sub = so_query@meta.data %>% group_by(sample_name) %>% slice_sample(n = 10689) %>% ungroup() %>% pull(cellID)
so_query = so_query[,sub]
so_query@meta.data %>% count(sample_name, sample_type)

# so_query$sample_name = so_query$Sample
# so_query$annotations = so_query$functional.cluster
# so_query@meta.data %>% count(annotations)
# so_query@meta.data %>% count(level2_final)

mypal_annotation = mypal
# mypal_annotation = setNames(mypal_annotation, unique(so_query$annotations))

# umap_mat <- as.matrix( so_query@meta.data[, c("umaprna_1", "umaprna_2")])
# colnames(umap_mat) <- c("umappaper_1", "umappaper_2")
# rownames(umap_mat) <- rownames(so_query@meta.data)
# so_query[["umap_paper"]] <- CreateDimReducObject(
#   embeddings = umap_mat,
#   key = "UMAPPAPER_",
#   assay = DefaultAssay(so_query)
# )

head(so_query@meta.data)
# table(so_query$cluster)
# table(so_query$level2_plot)
# table(so_query$batch)
# table(so_query$tissue)
# table(so_query$annotation, so_query$tissue)

# table(so_query$batch_sample)
# table(so_query$level2_plot, so_query$cluster)
# table(so_query$level2_plot, so_query$level1_final)
# so_query$annotation = paste("cl", so_query$annotation, sep = "_")

so_query@meta.data %>% count(sample_name)
so_query@meta.data %>% count(level2_final)

# so_query$sample = as.integer(sub('.*-(\\d+)$', '\\1', colnames(so_query)))

sub = so_query@meta.data %>% filter(!(level2_final %in% c("not classified"))) %>% row.names()

pdf(sprintf("%s/umap_dataset.pdf",query_dir),14, 7, useDingbats = F)
# p1 = DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "annotations", cols = mypal_annotation)
p2 = DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "level2_final", cols = mypal_level2)
p3 = DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "sample_name", cols = mypal)
# p1 + p2 + p3
p2 + p3
# DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "annotations", cols = mypal_annotation, split.by = "sample_type")
DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "level2_final", cols = mypal_level2, split.by = "sample_name")
# FeaturePlot(so_query[,sub], reduction = "mde_incremental_level2", features = "Pdcd1")
dev.off()

so2 = merge(so, so_query[,sub], merge.dr = T)

# mypal_annotation = setNames(c("#5D7AA5", "#BED4EB", "#E59339", "#F5CA8F", "#6FA649", "#B9E185", "#B29A39", "#EFD374", "#649B9F", "#A9CBC2", "#D2635A", "#F3B0B0"), c("cl_0", "cl_1", "cl_2", "cl_3", "cl_4", "cl_5", "cl_6", "cl_7", "cl_8", "cl_9", "cl_10", "cl_11"))
# mypal_annotation = mypal

pdf(sprintf("%s/RBI.pdf",query_dir),7, 7, useDingbats = F)
MyDimPlotHighlight(so2, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "level2_final", mycols = mypal_level2, print_plot1 = T, print_plot2 = T,labelclusters = F, xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = F, highlight_size = 1)
# MyDimPlotHighlight(so2, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "annotations", mycols = mypal_annotation, print_plot1 = T, print_plot2 = T,labelclusters = F, xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = T, highlight_size = 1)
MyDimPlotHighlight(so2, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "sample_name", mycols = mypal, print_plot1 = T, print_plot2 = T,labelclusters = F, xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = F, highlight_size = 1)


# MyDimPlotHighlight(so2, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "annotation", mycols = mypal_annotation, print_plot1 = T, print_plot2 = T,labelclusters = F)
for (s in unique(so_query$sample_name)) {
    print(s)
    tmp = so2[,so2$sample_name %in% c(s) | so2$sample %in% c("immgent")]
    print(MyDimPlotHighlight(tmp, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "level2_final", mycols = mypal_level2, print_plot1 = T, print_plot2 = T, labelclusters = T, xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = F, highlight_size = 0.5, title = sprintf("sample %s", s)))
    # print(MyDimPlotHighlight(tmp, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "annotations", mycols = mypal_annotation, print_plot1 = T, print_plot2 = T,labelclusters = T, , xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = F, highlight_size = 0.5, title = sprintf("sample %s", s)))
    # print(MyDimPlotHighlight(so2[,so2$sample %in% c("immgent", s)], umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "annotation", mycols = mypal_annotation, print_plot1 = T, print_plot2 = T,labelclusters = T, title = sprintf("sample: %s", s)))
}

# p1 = DimPlot(so_query[,sub], reduction = "mde_incremental_level2", group.by = "annotation", cols = mypal_annotation)
# p2 = DimPlot(so_query[,sub], reduction = "mde_incremental_level2", group.by = "level2_plot", cols = mypal_level2)
# p1 + p2
dev.off()


```

https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE200635 - tet+ CD4+ from draining lymph node and tumor, after subcutaneous injection of MC38-GP tumor cells, 13 days post-inoculation
No paper annotation
GP66:I-Ab+ CD4+ from Draining lymph node and Tumor
```{r}
query_dir = '/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/20250109_freeze/odhran/CD4/GSE200635_IAb_Tet_LCMV/Trial_Final/'
so_query = readRDS(sprintf('%s/Final_Seurat_object.Rds', query_dir))
so_query$barcode = sub(".*_", "", colnames(so_query)) 

demultiplexed = readRDS(sprintf("%s/GSE200635_demux.rds", query_dir))
colnames(demultiplexed)
so_query$Sample_demux = demultiplexed$Sample_demux[match(so_query$barcode, colnames(demultiplexed))] %>% as.character()

so_query$cellID = colnames(so_query)
so_query@meta.data %>% count(Sample_demux, IGTHT)

##Update metadata
# so_query$sample_name[grepl(x = so_query$Sample_demux, pattern = "CD44-2")] = "MC38GP_GSE200635_2"
so_query$sample_name[grepl(x = so_query$Sample_demux, pattern = "LN-1")] = "MC38GP_dLN_GSE200635_1"
so_query$sample_name[grepl(x = so_query$Sample_demux, pattern = "LN-2")] = "MC38GP_dLN_GSE200635_2"
so_query$sample_name[grepl(x = so_query$Sample_demux, pattern = "TIL-1")] = "MC38GP_Tumor_GSE200635_1"
so_query$sample_name[grepl(x = so_query$Sample_demux, pattern = "TIL-2")] = "MC38GP_Tumor_GSE200635_2"

so_query$sample_type = NA
so_query$sample_type[grepl(x = so_query$sample_name, pattern = "MC38GP_dLN_GSE200635")] = "MC38GP_dLN"
so_query$sample_type[grepl(x = so_query$sample_name, pattern = "MC38GP_Tumor_GSE200635")] = "MC38GP_Tumor"
# so_query$sample_type[grepl(x = so_query$sample_name, pattern = "MC38GP_GSE200635_2")] = "MC38GP_GSE200635_2"

so_query$annotations = NA

so_query$organ_simplified = NA
so_query$organ_simplified[grepl(x = so_query$sample_name, pattern = "dLN")] = "SLO"
so_query$organ_simplified[grepl(x = so_query$sample_name, pattern = "Tumor")] = "MC38GP"

so_query$target_cells_simplified = "CD4pGP66Tetp"
so_query$Ag_spe = "GP66:I-Ab"
so_query$infection_agent = NA
so_query$tumor = "MC38.GP"
# so_query$infection_agent[grepl(x = so_query$sample_name, pattern = "Arm")] = "LCMV Armstrong"
so_query$infection_phase = NA
# so_query$infection_phase[grepl(x = so_query$sample_name, pattern = "d7")] = "acute"
# so_query$infection_phase[grepl(x = so_query$sample_name, pattern = "d21")] = "memory"
# so_query$infection_phase[grepl(x = so_query$sample_name, pattern = "d44")] = "memory"

so_query@meta.data %>% count(sample_name, IGTHT)

so_query@meta.data %>% count(sample_name, sample_type, organ_simplified, target_cells_simplified, Ag_spe, infection_agent, infection_phase, tumor)

#update annotation
so_query = so_query[,sub]
so_query$level2_final.v1 = so_query$level2_final
so_query$annotation_level2 = newann$level2.new[match(so_query$level2_final.v1, newann$level2)]
any(is.na(so_query$annotation_level2))
so_query@meta.data %>% count(annotation_level2, level2_final.v1)

##Save a clean version
so_query@meta.data %>% count(level2_final)
sub = so_query@meta.data %>% filter(!(level2_final %in% c("not classified")), !(sample_name %in% "Batch1_GSM6040471")) %>% row.names()
so_query = so_query[,sub]
so_query@meta.data %>% count(sample_name, sample_type, organ_simplified, target_cells_simplified, Ag_spe, infection_agent, infection_phase, tumor)
saveRDS(so_query, sprintf("%s/Final_Seurat_object_clean_david.Rds", query_dir))

#####Plots

# sub = so_query@meta.data %>% group_by(sample_name) %>% slice_sample(n = 307) %>% ungroup() %>% pull(cellID)
# so_query = so_query[,sub]
# so_query@meta.data %>% count(sample_name)

# sub = so_query@meta.data %>% filter(!(level2_final %in% c("not classified"))) %>% row.names() 
# so_query = so_query[,sub]

# DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "sample_name", cols = mypal)
# p1 = DimPlot(so_query[,sub], reduction = "mde_incremental_level2", group.by = "annotations", cols = mypal_annotation) 
# p$layers <- c(list(immgenTCD4MDE()), p1$layers)
# p

pdf(sprintf("%s/umap_dataset.pdf",query_dir),20, 10, useDingbats = F)
p1 = DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "sample_name", cols = mypal)
p2 = DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "level2_final", cols = mypal_level2)
p1 + p2
DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "annotations", cols = mypal_annotation, split.by = "sample_name", ncol = 4)
DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "annotation_level2", cols = mypal_level2, split.by = "sample_name", ncol = 4)
# FeaturePlot(so_query[,sub], reduction = "mde_incremental_level2", features = "Pdcd1")
dev.off()


# mypal_annotation = setNames(c("#5D7AA5", "#BED4EB", "#E59339", "#F5CA8F", "#6FA649", "#B9E185", "#B29A39", "#EFD374", "#649B9F", "#A9CBC2", "#D2635A", "#F3B0B0"), c("cl_0", "cl_1", "cl_2", "cl_3", "cl_4", "cl_5", "cl_6", "cl_7", "cl_8", "cl_9", "cl_10", "cl_11"))
# mypal_annotation = mypal

pdf(sprintf("%s/RBI.pdf",query_dir),7, 7, useDingbats = F)
# MyDimPlotHighlight(so2, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "level2_final", mycols = mypal_level2, print_plot1 = T, print_plot2 = T,labelclusters = F, xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = T, highlight_size = 1)
# MyDimPlotHighlight(so2, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "annotations", mycols = mypal_annotation, print_plot1 = T, print_plot2 = T,labelclusters = F, xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = T, highlight_size = 1)

p1 = DimPlot(so_query[,sub], reduction = "mde_incremental_level2", group.by = "annotation_level2", cols = mypal_level2)
bg = AddEmbeddingRasterBackground("CD4")
p1$layers = c(list(bg), p1$layers)
p1 + xlim(c(-2.5,4)) + ylim(c(-2.5,2.5)) + theme(legend.position = "none") + labs(title = "annotation_level2")

p2 = DimPlot(so_query[,sub], reduction = "mde_incremental_level2", group.by = "annotations", cols = mypal)
bg = AddEmbeddingRasterBackground("CD4")
p2$layers = c(list(bg), p2$layers)
p2 + xlim(c(-2.5,4)) + ylim(c(-2.5,2.5)) + theme(legend.position = "none") + labs(title = "annotations")

# MyDimPlotHighlight(so2, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "annotation", mycols = mypal_annotation, print_plot1 = T, print_plot2 = T,labelclusters = F)
for (s in unique(so_query$sample_name)) {
    print(s)
    bg = AddEmbeddingRasterBackground("CD4")
    p1 = DimPlot(so_query[,so_query$sample_name %in% s], reduction = "mde_incremental_level2", group.by = "annotation_level2", cols = mypal_level2)
    p1$layers = c(list(bg), p1$layers)
    print(p1 + xlim(c(-2.5,4)) + ylim(c(-2.5,2.5)) + theme(legend.position = "none") + labs(title = sprintf("sample %s", s)))
    
    # p2 = DimPlot(so_query[,so_query$sample_name %in% s], reduction = "mde_incremental_level2", group.by = "annotations", cols = mypal_level2)
    # p2$layers = c(list(bg), p2$layers)
    # print(p2 + xlim(c(-2.5,4)) + ylim(c(-2.5,2.5)) + theme(legend.position = "none") + labs(title = "level2_final"))

}

for (s in unique(so_query$sample_type)) {
    print(s)
    bg = AddEmbeddingRasterBackground("CD4")
    p1 = DimPlot(so_query[,so_query$sample_type %in% s], reduction = "mde_incremental_level2", group.by = "annotation_level2", cols = mypal_level2)
    p1$layers = c(list(bg), p1$layers)
    print(p1 + xlim(c(-2.5,4)) + ylim(c(-2.5,2.5)) + theme(legend.position = "none") + labs(title = sprintf("sample %s", s)))
    
    # p2 = DimPlot(so_query[,so_query$sample_name %in% s], reduction = "mde_incremental_level2", group.by = "annotations", cols = mypal_level2)
    # p2$layers = c(list(bg), p2$layers)
    # print(p2 + xlim(c(-2.5,4)) + ylim(c(-2.5,2.5)) + theme(legend.position = "none") + labs(title = "level2_final"))

}

# p1 = DimPlot(so_query[,sub], reduction = "mde_incremental_level2", group.by = "annotation", cols = mypal_annotation)
# p2 = DimPlot(so_query[,sub], reduction = "mde_incremental_level2", group.by = "level2_plot", cols = mypal_level2)
# p1 + p2
dev.off()

```

https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE184259 - SMARTA CD4+ T cell subsets during LCMV infection D7 and D21, spleen, SI LP and IEL, WT and Bcl6 KO.
No paper annotation
```{r}
query_dir = '/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/20250109_freeze/odhran/CD4/GSE184259_LCMV_CD4_WT_Bcl6KO_D7_D21/Trial_Final/'
so_query = readRDS('/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/20250109_freeze/odhran/CD4/GSE184259_LCMV_CD4_WT_Bcl6KO_D7_D21/Trial_Final/Final_Seurat_object.Rds') 

so_query$cellID = colnames(so_query)
so_query@meta.data %>% count(sample_name, IGTHT)

##Update metadata
so_query$sample_name[grepl(x = so_query$IGTHT, pattern = "Batch1_GSM5582632_D21_SPL_1")] = "Arm_d21_SLO_GSE184259_1"
so_query$sample_name[grepl(x = so_query$IGTHT, pattern = "Batch1_GSM5582633_D21_LPL_1")] = "Arm_d21_SILP_GSE184259_1"
so_query$sample_name[grepl(x = so_query$IGTHT, pattern = "Batch1_GSM5582634_D21_SPL_2")] = "Arm_d21_SLO_GSE184259_2"
so_query$sample_name[grepl(x = so_query$IGTHT, pattern = "Batch1_GSM5582635_D21_LPL_2")] = "Arm_d21_SILP_GSE184259_2"
so_query$sample_name[grepl(x = so_query$IGTHT, pattern = "Batch1_GSM5582636_D7_SPL_1")] = "Arm_d7_SLO_GSE184259_1"
so_query$sample_name[grepl(x = so_query$IGTHT, pattern = "Batch1_GSM5582637_D7_IEL_1")] = "Arm_d7_SIIEL_GSE184259_1"
so_query$sample_name[grepl(x = so_query$IGTHT, pattern = "Batch1_GSM5582638_D7_LPL_1")] = "Arm_d7_SILP_GSE184259_1"
so_query$sample_name[grepl(x = so_query$IGTHT, pattern = "Batch1_GSM5582639_D7_SPL_2")] = "Arm_d7_SLO_GSE184259_2"
so_query$sample_name[grepl(x = so_query$IGTHT, pattern = "Batch1_GSM5582640_D7_IEL_2")] = "Arm_d7_SIIEL_GSE184259_2"
so_query$sample_name[grepl(x = so_query$IGTHT, pattern = "Batch1_GSM5582641_D7_LPL_2")] = "Arm_d7_SILP_GSE184259_2"
so_query$sample_name[grepl(x = so_query$IGTHT, pattern = "Batch1_GSM5582642_D7_Bcl6KO_SPL")] = "Arm_d7_SLO_Bcl6KO_GSE184259"
so_query$sample_name[grepl(x = so_query$IGTHT, pattern = "Batch1_GSM5582643_D7_Bcl6KO_IEL")] = "Arm_d7_SIIEL_Bcl6KO_GSE184259"
so_query$sample_name[grepl(x = so_query$IGTHT, pattern = "Batch1_GSM5582644_D7_Bcl6KO_LPL")] = "Arm_d7_SILP_Bcl6KO_GSE184259"

so_query$sample_type = NA
so_query$sample_type[grepl(x = so_query$sample_name, pattern = "Arm_d21_SLO_GSE184259")] = "Arm_d21_SLO"
so_query$sample_type[grepl(x = so_query$sample_name, pattern = "Arm_d21_SILP_GSE184259")] = "Arm_d21_SILP"
so_query$sample_type[grepl(x = so_query$sample_name, pattern = "Arm_d7_SLO_GSE184259")] = "Arm_d7_SLO"
so_query$sample_type[grepl(x = so_query$sample_name, pattern = "Arm_d7_SILP_GSE184259")] = "Arm_d7_SILP"
so_query$sample_type[grepl(x = so_query$sample_name, pattern = "Arm_d7_SIIEL_GSE184259")] = "Arm_d7_SIIEL"
so_query$sample_type[grepl(x = so_query$sample_name, pattern = "Arm_d7_SLO_Bcl6KO_GSE184259")] = "Arm_d7_SLO_Bcl6KO"
so_query$sample_type[grepl(x = so_query$sample_name, pattern = "Arm_d7_SILP_Bcl6KO_GSE184259")] = "Arm_d7_SILP_Bcl6KO"
so_query$sample_type[grepl(x = so_query$sample_name, pattern = "Arm_d7_SIIEL_Bcl6KO_GSE184259")] = "Arm_d7_SIIEL_Bcl6KO"

so_query$annotations = NA

so_query$organ_simplified = NA
so_query$organ_simplified[grepl(x = so_query$sample_name, pattern = "SLO")] = "SLO"
so_query$organ_simplified[grepl(x = so_query$sample_name, pattern = "SILP")] = "SILP"
so_query$organ_simplified[grepl(x = so_query$sample_name, pattern = "SIIEL")] = "SIIEL"

so_query$target_cells_simplified = "CD4pSmarta"
so_query$Ag_spe = "Smarta"
so_query$infection_agent = "LCMV Armstrong"
# so_query$infection_agent[grepl(x = so_query$sample_name, pattern = "Arm")] = "LCMV Armstrong"
so_query$infection_phase = NA
so_query$infection_phase[grepl(x = so_query$sample_name, pattern = "d7")] = "acute"
so_query$infection_phase[grepl(x = so_query$sample_name, pattern = "d21")] = "memory"

so_query@meta.data %>% count(IGTHT, sample_name)

so_query$annotation_level2 = so_query$level2_final
so_query@meta.data %>% count(annotation_level2)

##Save a clean version
sub = so_query@meta.data %>% filter(!(level2_final %in% c("not classified"))) %>% row.names()
so_query = so_query[,sub]
# saveRDS(so_query, sprintf("%s/Final_Seurat_object_clean_david.Rds", query_dir))

#####Plots

# sub = so_query@meta.data %>% group_by(sample_name) %>% slice_sample(n = 307) %>% ungroup() %>% pull(cellID)
# so_query = so_query[,sub]
# so_query@meta.data %>% count(sample_name)

sub = so_query@meta.data %>% filter(!(level2_final %in% c("not classified"))) %>% row.names() 
so_query = so_query[,sub]

# DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "sample_name", cols = mypal)
# p1 = DimPlot(so_query[,sub], reduction = "mde_incremental_level2", group.by = "annotations", cols = mypal_annotation) 
# p$layers <- c(list(immgenTCD4MDE()), p1$layers)
# p

pdf(sprintf("%s/umap_dataset.pdf",query_dir),14, 10, useDingbats = F)
p1 = DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "annotations", cols = mypal_annotation)
p2 = DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "level2_final", cols = mypal_level2)
p1 + p2
DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "annotations", cols = mypal_annotation, split.by = "sample_name", ncol = 4)
DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "level2_final", cols = mypal_level2, split.by = "sample_name", ncol = 4)
# FeaturePlot(so_query[,sub], reduction = "mde_incremental_level2", features = "Pdcd1")
dev.off()


# mypal_annotation = setNames(c("#5D7AA5", "#BED4EB", "#E59339", "#F5CA8F", "#6FA649", "#B9E185", "#B29A39", "#EFD374", "#649B9F", "#A9CBC2", "#D2635A", "#F3B0B0"), c("cl_0", "cl_1", "cl_2", "cl_3", "cl_4", "cl_5", "cl_6", "cl_7", "cl_8", "cl_9", "cl_10", "cl_11"))
# mypal_annotation = mypal

pdf(sprintf("%s/RBI.pdf",query_dir),7, 7, useDingbats = F)
# MyDimPlotHighlight(so2, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "level2_final", mycols = mypal_level2, print_plot1 = T, print_plot2 = T,labelclusters = F, xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = T, highlight_size = 1)
# MyDimPlotHighlight(so2, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "annotations", mycols = mypal_annotation, print_plot1 = T, print_plot2 = T,labelclusters = F, xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = T, highlight_size = 1)

p1 = DimPlot(so_query[,sub], reduction = "mde_incremental_level2", group.by = "level2_final", cols = mypal_level2)
bg = AddEmbeddingRasterBackground("CD4")
p1$layers = c(list(bg), p1$layers)
p1 + xlim(c(-2.5,4)) + ylim(c(-2.5,2.5)) + theme(legend.position = "none") + labs(title = "level2_final")

p2 = DimPlot(so_query[,sub], reduction = "mde_incremental_level2", group.by = "annotations", cols = mypal)
bg = AddEmbeddingRasterBackground("CD4")
p2$layers = c(list(bg), p2$layers)
p2 + xlim(c(-2.5,4)) + ylim(c(-2.5,2.5)) + theme(legend.position = "none") + labs(title = "annotations")

# MyDimPlotHighlight(so2, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "annotation", mycols = mypal_annotation, print_plot1 = T, print_plot2 = T,labelclusters = F)
for (s in unique(so_query$sample_name)) {
    print(s)
    bg = AddEmbeddingRasterBackground("CD4")
    p1 = DimPlot(so_query[,so_query$sample_name %in% s], reduction = "mde_incremental_level2", group.by = "level2_final", cols = mypal_level2)
    p1$layers = c(list(bg), p1$layers)
    print(p1 + xlim(c(-2.5,4)) + ylim(c(-2.5,2.5)) + theme(legend.position = "none") + labs(title = sprintf("sample %s", s)))
    
    # p2 = DimPlot(so_query[,so_query$sample_name %in% s], reduction = "mde_incremental_level2", group.by = "annotations", cols = mypal_level2)
    # p2$layers = c(list(bg), p2$layers)
    # print(p2 + xlim(c(-2.5,4)) + ylim(c(-2.5,2.5)) + theme(legend.position = "none") + labs(title = "level2_final"))

}

for (s in unique(so_query$sample_type)) {
    print(s)
    bg = AddEmbeddingRasterBackground("CD4")
    p1 = DimPlot(so_query[,so_query$sample_type %in% s], reduction = "mde_incremental_level2", group.by = "level2_final", cols = mypal_level2)
    p1$layers = c(list(bg), p1$layers)
    print(p1 + xlim(c(-2.5,4)) + ylim(c(-2.5,2.5)) + theme(legend.position = "none") + labs(title = sprintf("sample %s", s)))
    
    # p2 = DimPlot(so_query[,so_query$sample_name %in% s], reduction = "mde_incremental_level2", group.by = "annotations", cols = mypal_level2)
    # p2$layers = c(list(bg), p2$layers)
    # print(p2 + xlim(c(-2.5,4)) + ylim(c(-2.5,2.5)) + theme(legend.position = "none") + labs(title = "level2_final"))

}

# p1 = DimPlot(so_query[,sub], reduction = "mde_incremental_level2", group.by = "annotation", cols = mypal_annotation)
# p2 = DimPlot(so_query[,sub], reduction = "mde_incremental_level2", group.by = "level2_plot", cols = mypal_level2)
# p1 + p2
dev.off()



```

https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE175742 - SMARTA CD4+ T cell subsets during LCMV infection D7 and D21 and D44, spleen
No paper annotation
```{r}
query_dir = '/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/20250109_freeze/odhran/CD4/GSE175742_LCMV_CD4_D7_D21_D44/Trial_Final/'
so_query = readRDS('/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/20250109_freeze/odhran/CD4/GSE175742_LCMV_CD4_D7_D21_D44/Trial_Final/Final_Seurat_object.Rds') 

so_query$cellID = colnames(so_query)
so_query@meta.data %>% count(sample_name, IGTHT)

##Update metadata
so_query$sample_name[grepl(x = so_query$IGTHT, pattern = "Batch1_GSM5345277_D7_SPL_1")] = "Arm_d7_SLO_GSE175742_1"
so_query$sample_name[grepl(x = so_query$IGTHT, pattern = "Batch1_GSM5345278_D7_SPL_2")] = "Arm_d7_SLO_GSE175742_2"
so_query$sample_name[grepl(x = so_query$IGTHT, pattern = "Batch1_GSM5345279_D21_SPL_1")] = "Arm_d21_SLO_GSE175742_1"
so_query$sample_name[grepl(x = so_query$IGTHT, pattern = "Batch1_GSM5345280_D21_SPL_2")] = "Arm_d21_SLO_GSE175742_2"
so_query$sample_name[grepl(x = so_query$IGTHT, pattern = "Batch1_GSM5345281_D44_SPL_1")] = "Arm_d44_SLO_GSE175742_1"
so_query$sample_name[grepl(x = so_query$IGTHT, pattern = "Batch1_GSM5345282_D44_SPL_2")] = "Arm_d44_SLO_GSE175742_2"

so_query$sample_type = NA
so_query$sample_type[grepl(x = so_query$sample_name, pattern = "Arm_d7_SLO_GSE175742")] = "Arm_d7_SLO"
so_query$sample_type[grepl(x = so_query$sample_name, pattern = "Arm_d21_SLO_GSE175742")] = "Arm_d21_SLO"
so_query$sample_type[grepl(x = so_query$sample_name, pattern = "Arm_d44_SLO_GSE175742")] = "Arm_d44_SLO"

so_query$annotations = NA

so_query$organ_simplified = "SLO"
# so_query$organ_simplified[grepl(x = so_query$sample_name, pattern = "SLO")] = "SLO"

so_query$target_cells_simplified = "CD4pSmarta"
so_query$Ag_spe = "Smarta"
so_query$infection_agent = "LCMV Armstrong"
# so_query$infection_agent[grepl(x = so_query$sample_name, pattern = "Arm")] = "LCMV Armstrong"
so_query$infection_phase = NA
so_query$infection_phase[grepl(x = so_query$sample_name, pattern = "d7")] = "acute"
so_query$infection_phase[grepl(x = so_query$sample_name, pattern = "d21")] = "memory"
so_query$infection_phase[grepl(x = so_query$sample_name, pattern = "d44")] = "memory"

so_query@meta.data %>% count(sample_name, IGTHT)

so_query@meta.data %>% count(sample_name, sample_type, organ_simplified, target_cells_simplified, Ag_spe, infection_agent, infection_phase)

so_query$annotation_level2 = so_query$level2_final
so_query@meta.data %>% count(annotation_level2)

##Save a clean version
so_query@meta.data %>% count(level2_final)
sub = so_query@meta.data %>% filter(!(level2_final %in% c("not classified"))) %>% row.names()
so_query = so_query[,sub]
saveRDS(so_query, sprintf("%s/Final_Seurat_object_clean_david.Rds", query_dir))

#####Plots

# sub = so_query@meta.data %>% group_by(sample_name) %>% slice_sample(n = 307) %>% ungroup() %>% pull(cellID)
# so_query = so_query[,sub]
# so_query@meta.data %>% count(sample_name)

sub = so_query@meta.data %>% filter(!(level2_final %in% c("not classified"))) %>% row.names() 
so_query = so_query[,sub]

# DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "sample_name", cols = mypal)
# p1 = DimPlot(so_query[,sub], reduction = "mde_incremental_level2", group.by = "annotations", cols = mypal_annotation) 
# p$layers <- c(list(immgenTCD4MDE()), p1$layers)
# p

pdf(sprintf("%s/umap_dataset.pdf",query_dir),20, 10, useDingbats = F)
p1 = DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "sample_name", cols = mypal)
p2 = DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "level2_final", cols = mypal_level2)
p1 + p2
DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "annotations", cols = mypal_annotation, split.by = "sample_name", ncol = 4)
DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "level2_final", cols = mypal_level2, split.by = "sample_name", ncol = 4)
# FeaturePlot(so_query[,sub], reduction = "mde_incremental_level2", features = "Pdcd1")
dev.off()


# mypal_annotation = setNames(c("#5D7AA5", "#BED4EB", "#E59339", "#F5CA8F", "#6FA649", "#B9E185", "#B29A39", "#EFD374", "#649B9F", "#A9CBC2", "#D2635A", "#F3B0B0"), c("cl_0", "cl_1", "cl_2", "cl_3", "cl_4", "cl_5", "cl_6", "cl_7", "cl_8", "cl_9", "cl_10", "cl_11"))
# mypal_annotation = mypal

pdf(sprintf("%s/RBI.pdf",query_dir),7, 7, useDingbats = F)
# MyDimPlotHighlight(so2, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "level2_final", mycols = mypal_level2, print_plot1 = T, print_plot2 = T,labelclusters = F, xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = T, highlight_size = 1)
# MyDimPlotHighlight(so2, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "annotations", mycols = mypal_annotation, print_plot1 = T, print_plot2 = T,labelclusters = F, xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = T, highlight_size = 1)

p1 = DimPlot(so_query[,sub], reduction = "mde_incremental_level2", group.by = "level2_final", cols = mypal_level2)
bg = AddEmbeddingRasterBackground("CD4")
p1$layers = c(list(bg), p1$layers)
p1 + xlim(c(-2.5,4)) + ylim(c(-2.5,2.5)) + theme(legend.position = "none") + labs(title = "level2_final")

p2 = DimPlot(so_query[,sub], reduction = "mde_incremental_level2", group.by = "annotations", cols = mypal)
bg = AddEmbeddingRasterBackground("CD4")
p2$layers = c(list(bg), p2$layers)
p2 + xlim(c(-2.5,4)) + ylim(c(-2.5,2.5)) + theme(legend.position = "none") + labs(title = "annotations")

# MyDimPlotHighlight(so2, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "annotation", mycols = mypal_annotation, print_plot1 = T, print_plot2 = T,labelclusters = F)
for (s in unique(so_query$sample_name)) {
    print(s)
    bg = AddEmbeddingRasterBackground("CD4")
    p1 = DimPlot(so_query[,so_query$sample_name %in% s], reduction = "mde_incremental_level2", group.by = "level2_final", cols = mypal_level2)
    p1$layers = c(list(bg), p1$layers)
    print(p1 + xlim(c(-2.5,4)) + ylim(c(-2.5,2.5)) + theme(legend.position = "none") + labs(title = sprintf("sample %s", s)))
    
    # p2 = DimPlot(so_query[,so_query$sample_name %in% s], reduction = "mde_incremental_level2", group.by = "annotations", cols = mypal_level2)
    # p2$layers = c(list(bg), p2$layers)
    # print(p2 + xlim(c(-2.5,4)) + ylim(c(-2.5,2.5)) + theme(legend.position = "none") + labs(title = "level2_final"))

}

for (s in unique(so_query$sample_type)) {
    print(s)
    bg = AddEmbeddingRasterBackground("CD4")
    p1 = DimPlot(so_query[,so_query$sample_type %in% s], reduction = "mde_incremental_level2", group.by = "level2_final", cols = mypal_level2)
    p1$layers = c(list(bg), p1$layers)
    print(p1 + xlim(c(-2.5,4)) + ylim(c(-2.5,2.5)) + theme(legend.position = "none") + labs(title = sprintf("sample %s", s)))
    
    # p2 = DimPlot(so_query[,so_query$sample_name %in% s], reduction = "mde_incremental_level2", group.by = "annotations", cols = mypal_level2)
    # p2$layers = c(list(bg), p2$layers)
    # print(p2 + xlim(c(-2.5,4)) + ylim(c(-2.5,2.5)) + theme(legend.position = "none") + labs(title = "level2_final"))

}

# p1 = DimPlot(so_query[,sub], reduction = "mde_incremental_level2", group.by = "annotation", cols = mypal_annotation)
# p2 = DimPlot(so_query[,sub], reduction = "mde_incremental_level2", group.by = "level2_plot", cols = mypal_level2)
# p1 + p2
dev.off()



```

**T-RBI Merge query seurat objects**
--> trbi_8studies_20260311.Rds
```{r}
newann = read.table('/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/20260206_freeze/annotation_table_20260201_conversion_forOdhran.csv', header = T, sep = ",")

queries = list()

query_dir = '/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/20250109_freeze/odhran/CD4/GSE281547_Jenkins_b6_aLm/Trial_Final/'
so_query = readRDS(sprintf("%s/Final_Seurat_object.Rds", query_dir)) 
so_query$cellID = colnames(so_query)
sub = so_query@meta.data %>% filter(!(level2_final %in% c("not classified"))) %>% row.names()
so_query = so_query[,sub]
so_query$level2_final.v1 = so_query$level2_final
so_query$annotation_level2 = newann$level2.new[match(so_query$level2_final.v1, newann$level2)]
any(is.na(so_query$annotation_level2))
so_query@meta.data %>% count(annotation_level2, level2_final.v1)
so_query@meta.data %>% count(sample_name, IGTHT)
so_query$sample_name[so_query$IGTHT == "Sample-2_D06"] = "Lm_D06_SLO_GSE281547"
so_query$sample_name[so_query$IGTHT == "Sample-3_D21"] = "Lm_D21_SLO_GSE281547"
so_query$sample_name[so_query$IGTHT == "Sample-1_D60"] = "Lm_D60_SLO_GSE281547"
so_query$sample_type = NA
so_query$sample_type[so_query$IGTHT == "Sample-2_D06"] = "Lm_D06_SLO"
so_query$sample_type[so_query$IGTHT == "Sample-3_D21"] = "Lm_D21_SLO"
so_query$sample_type[so_query$IGTHT == "Sample-1_D60"] = "Lm_D60_SLO"
so_query$annotations = so_query$annotations
so_query$organ_simplified = "SLO"
so_query$target_cells_simplified = "CD4p2WTetp"
so_query$Ag_spe = "2W:I-Ab"
so_query$infection_agent = NA
so_query$infection_agent[grepl(x = so_query$sample_name, pattern = "Lm")] = "Listeria monocytogenes"
so_query$infection_phase = NA
so_query$infection_phase[grepl(x = so_query$sample_name, pattern = "D06")] = "acute"
so_query$infection_phase[grepl(x = so_query$sample_name, pattern = "D21")] = "memory"
so_query$infection_phase[grepl(x = so_query$sample_name, pattern = "D60")] = "memory"
so_query@meta.data %>% count(sample_name, sample_type, organ_simplified, target_cells_simplified, Ag_spe, infection_agent, infection_phase)
# saveRDS(so_query, sprintf("%s/Final_Seurat_object_clean_david.Rds", query_dir))
queries[["GSE281547_Osum_Lm"]] = so_query

query_dir = '/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/20250109_freeze/odhran/CD4/GSE182320_IAb_Tet_WT_Zbtb7bKO_LCMV/Trial_Final/'
so_query = readRDS(sprintf("%s/Final_Seurat_object.Rds", query_dir)) 
so_query$cellID = colnames(so_query)
sub = so_query@meta.data %>% filter(!(level2_final %in% c("not classified"))) %>% row.names()
so_query = so_query[,sub]
so_query$level2_final.v1 = so_query$level2_final
so_query$annotation_level2 = newann$level2.new[match(so_query$level2_final.v1, newann$level2)]
any(is.na(so_query$annotation_level2))
so_query@meta.data %>% count(annotation_level2, level2_final.v1)
so_query$sample_name = sprintf("%s_%s_%s", so_query$Sample, "SLO", "GSE182320")
so_query@meta.data %>% count(sample_name, IGTHT)
so_query$sample_type = NA
so_query$sample_type[grepl(x = so_query$IGTHT, pattern = "Arm_d7")] = "Arm_d7_SLO"
so_query$sample_type[grepl(x = so_query$IGTHT, pattern = "Arm_d21")] = "Arm_d21_SLO"
so_query$sample_type[grepl(x = so_query$IGTHT, pattern = "Arm_d60")] = "Arm_d60_SLO"
so_query$sample_type[grepl(x = so_query$IGTHT, pattern = "Cl13_d7")] = "Cl13_d7_SLO"
so_query$sample_type[grepl(x = so_query$IGTHT, pattern = "Cl13_d21")] = "Cl13_d21_SLO"
so_query$annotations = so_query$functional.cluster
so_query$organ_simplified = "SLO"
so_query$target_cells_simplified = "CD4pGP66Tetp"
so_query$Ag_spe = "GP66:I-Ab"
so_query$infection_agent = NA
so_query$infection_agent[grepl(x = so_query$sample_name, pattern = "Arm")] = "LCMV Armstrong"
so_query$infection_agent[grepl(x = so_query$sample_name, pattern = "Cl13")] = "LCMV Clone 13"
so_query$infection_phase = NA
so_query$infection_phase[grepl(x = so_query$sample_name, pattern = "d7")] = "acute"
so_query$infection_phase[grepl(x = so_query$sample_name, pattern = "Arm_d21")] = "memory"
so_query$infection_phase[grepl(x = so_query$sample_name, pattern = "Cl13_d21")] = "chronic"
so_query$infection_phase[grepl(x = so_query$sample_name, pattern = "d60")] = "memory"
so_query@meta.data %>% count(sample_name, sample_type, organ_simplified, target_cells_simplified, Ag_spe, infection_agent, infection_phase)
# saveRDS(so_query, sprintf("%s/Final_Seurat_object_clean_david.Rds", query_dir))
queries[["GSE182320_Andreatta_LCMVArmCl13"]] = so_query

query_dir = '/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/20250109_freeze/odhran/CD4/GSE121002_IAb_Tet_binding_Listeria/Not_classified_Full_Trial/Not_classified_loop_fixed/'
so_query = readRDS(sprintf("%s/Final_Seurat_object.Rds", query_dir)) 
so_query$cellID = colnames(so_query)
sub = so_query@meta.data %>% filter(sample_name %in% c("Batch1_GSM3423796_d7_GP66", "Batch1_GSM3423797_d7_WT", "Batch1_GSM3423799_d30_WT") & level1_final == "CD4" & level2_final != "CD4") %>% pull(cellID)
so_query = so_query[,sub]
so_query$level2_final.v1 = so_query$level2_final
so_query$annotation_level2 = newann$level2.new[match(so_query$level2_final.v1, newann$level2)]
any(is.na(so_query$annotation_level2))
so_query@meta.data %>% count(annotation_level2, level2_final.v1)
so_query@meta.data %>% count(sample_name, IGTHT)
sub = so_query@meta.data %>% filter(sample_name %in% c("Batch1_GSM3423796_d7_GP66", "Batch1_GSM3423797_d7_WT", "Batch1_GSM3423799_d30_WT") & level1_final == "CD4") %>% pull(cellID)
so_query = so_query[,sub]
so_query$sample_name[grepl(x = so_query$sample_name, pattern = "d7")] = "Arm_d7_SLO_GSE121002"
so_query$sample_name[grepl(x = so_query$sample_name, pattern = "d30")] = "Arm_d30_SLO_GSE121002"
sub = so_query@meta.data %>% filter(!(level2_final %in% c("CD4"))) %>% row.names()
so_query$sample_type = NA
so_query$sample_type[grepl(x = so_query$sample_name, pattern = "d7")] = "Arm_d7_SLO"
so_query$sample_type[grepl(x = so_query$sample_name, pattern = "d30")] = "Arm_d30_SLO"
so_query$annotations = NA
so_query$organ_simplified = "SLO"
so_query$target_cells_simplified = "CD4pCD44pGP66Tetp"
so_query$Ag_spe = "GP66:I-Ab"
so_query$infection_agent = NA
so_query$infection_agent[grepl(x = so_query$sample_name, pattern = "Arm")] = "LCMV Armstrong"
so_query$infection_phase = NA
so_query$infection_phase[grepl(x = so_query$sample_name, pattern = "d7")] = "acute"
so_query$infection_phase[grepl(x = so_query$sample_name, pattern = "d30")] = "memory"
so_query@meta.data %>% count(sample_name, sample_type, organ_simplified, target_cells_simplified, Ag_spe, infection_agent, infection_phase)
# saveRDS(so_query, sprintf("%s/Final_Seurat_object_clean_david.Rds", query_dir))
queries[["GSE121002_Ciucci_LCMVArm"]] = so_query

query_dir = '/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/20250109_freeze/odhran/CD4/GSE158896_IAb_Tet_LCMV_d35/Trial_Final/'
so_query = readRDS(sprintf("%s/Final_Seurat_object.Rds", query_dir)) 
so_query$cellID = colnames(so_query)
sub = so_query@meta.data %>% filter(!(level2_final %in% c("not classified"))) %>% row.names()
so_query = so_query[,sub]
so_query$level2_final.v1 = so_query$level2_final
so_query$annotation_level2 = newann$level2.new[match(so_query$level2_final.v1, newann$level2)]
any(is.na(so_query$annotation_level2))
so_query@meta.data %>% count(annotation_level2, level2_final.v1)
so_query@meta.data %>% count(sample_name)
so_query$sample_name = so_query$sample
so_query$sample_name[so_query$IGTHT == "Batch1_GSM4815363_M1_GEX"] = "Arm_d10_SLO_GSE158896_1"
so_query$sample_name[so_query$IGTHT == "Batch1_GSM4815364_M2_GEX"] = "Arm_d10_SLO_GSE158896_2"
so_query$sample_name[so_query$IGTHT == "Batch1_GSM4815365_M3_GEX"] = "Arm_d10_SLO_GSE158896_3"
so_query$sample_name[so_query$IGTHT == "Batch1_GSM4815366_M4_GEX"] = "Arm_d10_SLO_GSE158896_4"
so_query$sample_name[so_query$IGTHT == "Batch1_GSM4815367_M5_GEX"] = "Arm_d10_SLO_GSE158896_5"
so_query$sample_type = "Arm_d10_SLO"
so_query$organ_simplified = "SLO"
so_query$target_cells_simplified = "CD4pCD44pGP66Tetp"
so_query$Ag_spe = "GP66:I-Ab"
so_query$infection_agent = NA
so_query$infection_agent[grepl(x = so_query$sample_name, pattern = "Arm")] = "LCMV Armstrong"
so_query$infection_phase = "acute"
so_query@meta.data %>% count(sample_name, sample_type, organ_simplified, target_cells_simplified, Ag_spe, infection_agent, infection_phase)
# saveRDS(so_query, sprintf("%s/Final_Seurat_object_clean_david.Rds", query_dir))
queries[["GSE158896_Khatun_LCMVArm"]] = so_query

query_dir = '/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/20250109_freeze/odhran/CD4/GSE198466_Vaccinated/Trial_Final/'
so_query = readRDS(sprintf("%s/Final_Seurat_object.Rds", query_dir)) 
so_query$cellID = colnames(so_query)
sub = so_query@meta.data %>% filter(!(level2_final %in% c("not classified"))) %>% row.names()
so_query = so_query[,sub]
so_query$level2_final.v1 = so_query$level2_final
so_query$annotation_level2 = newann$level2.new[match(so_query$level2_final.v1, newann$level2)]
any(is.na(so_query$annotation_level2))
so_query@meta.data %>% count(annotation_level2, level2_final.v1)
so_query@meta.data %>% count(sample_name)
so_query$sample_name[so_query$sample_name == "GSM5949203_IN_GSM5949203_IN"] = "Blasto_lung_secondaryD3_intranasalvaccine_GSE198466"
so_query$sample_name[so_query$sample_name == "GSM5949204_SC_GSM5949204_SC"] = "Blasto_lung_secondaryD3_subcutvaccine_GSE198466"
so_query$sample_type = NA
so_query$sample_type[so_query$sample_name == "Blasto_lung_secondaryD3_intranasalvaccine_GSE198466"] = "Blasto_lung_secondaryD3_intranasalvaccine"
so_query$sample_type[so_query$sample_name == "Blasto_lung_secondaryD3_subcutvaccine_GSE198466"] = "Blasto_lung_secondaryD3_subcutvaccine"
so_query$annotations = NA
so_query$organ_simplified = "lung"
so_query$target_cells_simplified = "CD4pEng2Tetp"
so_query$Ag_spe = "Bl-Eng2:classII"
so_query$infection_agent = NA
so_query$infection_agent[grepl(x = so_query$sample_name, pattern = "Blasto")] = "Blastomyces dermatitidis"
so_query$infection_phase = "recall"
so_query@meta.data %>% count(sample_name, sample_type, organ_simplified, target_cells_simplified, Ag_spe, infection_agent, infection_phase)
# saveRDS(so_query, sprintf("%s/Final_Seurat_object_clean_david.Rds", query_dir))
queries[["GSE198466_Woodring_Blasto"]] = so_query

query_dir = '/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/20250109_freeze/odhran/CD4/GSE200635_IAb_Tet_LCMV/Trial_Final/'
so_query = readRDS(sprintf("%s/Final_Seurat_object_clean_david.Rds", query_dir))
queries[["GSE200635_Andreatta_Tumor"]] = so_query

query_dir = '/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/20250109_freeze/odhran/CD4/GSE184259_LCMV_CD4_WT_Bcl6KO_D7_D21/Trial_Final/'
so_query = readRDS(sprintf("%s/Final_Seurat_object_clean_david.Rds", query_dir))
queries[["GSE184259_Nguyen_LCMV"]] = so_query

query_dir = '/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/20250109_freeze/odhran/CD4/GSE175742_LCMV_CD4_D7_D21_D44/Trial_Final/'
so_query = readRDS(sprintf("%s/Final_Seurat_object_clean_david.Rds", query_dir))
queries[["GSE175742_Shaw_LCMV"]] = so_query

lapply(queries, function(x) x@meta.data %>% count(sample_name, sample_type, organ_simplified, target_cells_simplified, Ag_spe, infection_agent, infection_phase))

saveRDS(queries, file = sprintf("%s/trbi_8studies_20260311.Rds",out_dir))


```

Merge on Midway3 if needed (rest of the code can be done on Laptop)
--> trbi_8studies_20260311_merged.Rds, trbi_8studies_20260311_merged_diet.Rds
```{r}
#Merge and keep only overlapping genes
library(Seurat)
library(purrr)

so_list = readRDS(sprintf("%s/trbi_8studies_20260311.Rds",out_dir))

common_features <- Reduce(
  intersect,
  lapply(so_list, rownames)
)
length(common_features)
head(common_features)

so_list_filt = list()
for (i in 1:length(so_list)) {
    print(i)
    so_list_filt[[i]] = subset(so_list[[i]], features = common_features)
    
    so_list_filt[[i]]$dataset = names(so_list)[i]
    
    message("Rename cells...")
    so_list_filt[[i]] = RenameCells(so_list_filt[[i]], new.names = paste0(names(so_list)[i], "_", colnames(so_list_filt[[i]])))
}

rm(so_list)
so_merged = merge(so_list_filt[[1]],so_list_filt[2:length(so_list_filt)],  collapse = T, merge.data = T, merge.dr = T)

so_merged[["mde_incremental"]] <- so_merged[["mde_incremental_level2"]]
so_merged[["mde_incremental_level2"]] <- NULL

so_merged[["mde2_totalvi_20241006"]] <- so_merged[["mde_incremental_allT"]]
so_merged[["mde_incremental_allT"]] <- NULL

so_merged$IGTHT_old = so_merged$IGTHT
so_merged$IGTHT = so_merged$sample_name

so_merged@meta.data %>% filter(is.na(annotation_level2 )) %>% count(dataset )

# newann = read.table('/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/20260206_freeze/annotation_table_20260201_conversion_forOdhran.csv', header = T, sep = ",")
# so_merged$level2_final.v1 = so_merged$level2_final
# so_merged$level2_final = newann$level2.new[match(so_merged$level2_final.v1, newann$level2)]
# any(is.na(so_merged$level2_final))
# so_merged@meta.data %>% count(level2_final, level2_final.v1)

# saveRDS(so_merged, file = "trbi_6studies_20260306_merged.Rds")
saveRDS(queries, file = sprintf("%s/trbi_8studies_20260311_merged.Rds",out_dir))

## merge and Diet Seurat
features_keep = intersect(c("Cd4", "Cd8b1", "Cd8a", "Trdc", "Foxp3", "Zbtb16", "Trdc", 
                            "Ifng", "Il4", "Il13", "Il17a", "Il17f", "Il22", "Tbx21", "Rorc", "Gata3", "Tox", "Tox2", "Pdcd1", "Cxcr5", "Sell", "Cd44", "Ccr7", "Ly6c1", "Ly6c2"), rownames(obj[["RNA"]]@counts))
so_merged_slim = DietSeurat(
    so_merged,
    assays = "RNA",
    layers = "counts",
    features = features_keep,
    dimreducs  = c("mde_incremental", "mde2_totalvi_20241006"),
    misc       = FALSE
)
saveRDS(queries, file = sprintf("%s/trbi_8studies_20260311_merged_diet.Rds",out_dir))

########### merge and Diet Seurat from so_list, if needed
# so_list = readRDS("trbi_6studies_20260306.Rds")
so_list = readRDS("trbi_8studies_20260311.Rds")

so_list_slim = list()
for (i in 1:length(so_list)) {
    print(i)
    obj = so_list[[i]]
    dim_keep <- intersect(c("mde_incremental_allT", "mde_incremental_level2"), names(obj@reductions))
    features_keep = intersect(c("Cd4", "Cd8b1", "Cd8a", "Trdc", "Foxp3", "Zbtb16", "Trdc", 
                                "Ifng", "Il4", "Il13", "Il17a", "Il17f", "Il22", "Tbx21", "Rorc", "Gata3", "Tox", "Tox2", "Pdcd1", "Cxcr5", "Sell", "Cd44", "Ccr7", "Ly6c1", "Ly6c2"), rownames(obj[["RNA"]]@counts))
    
    message("DietSeurat...")
    obj_slim = DietSeurat(
        obj,
        assays = "RNA",
        layers = "counts",
        features = features_keep,
        dimreducs  = dim_keep,
        misc       = FALSE
    )
    
    obj_slim$dataset = names(so_list)[i]
    
    message("Rename cells...")
    obj_slim = RenameCells(obj_slim, new.names = paste0(names(so_list)[i], "_", colnames(obj_slim)))
    
    so_list_slim[[names(so_list)[i]]] = obj_slim
    
}

# Merge all slimmed objects
so_merged = merge(so_list_slim[[1]],so_list_slim[2:length(so_list_slim)],  collapse = T, merge.data = T, merge.dr = T)
# so_merged = Reduce(function(x, y) merge(x, y), so_list_slim)
so_merged

so_merged[["mde_incremental"]] <- so_merged[["mde_incremental_level2"]]
so_merged[["mde_incremental_level2"]] <- NULL

so_merged[["mde2_totalvi_20241006"]] <- so_merged[["mde_incremental_allT"]]
so_merged[["mde_incremental_allT"]] <- NULL

newann = read.table('/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/20260206_freeze/annotation_table_20260201_conversion_forOdhran.csv', header = T, sep = ",")
so_merged$level2_final.v1 = so_merged$level2_final
so_merged$level2_final = newann$level2.new[match(so_merged$level2_final.v1, newann$level2)]
any(is.na(so_merged$level2_final))
so_merged@meta.data %>% count(level2_final, level2_final.v1)

saveRDS(so_merged, file = "trbi_6studies_20260306_merged_diet.Rds")

```

Table with query datasets and cell counts: 
138685 cells
```{r}
# so_merged_orig = readRDS("trbi_6studies_20260306_merged_diet.Rds")
so_merged_orig = readRDS(sprintf("%s/trbi_8studies_20260311_merged_diet.Rds",out_dir))

# so_merged_orig@meta.data %>% count(dataset,sample_name, name = "n_cells") %>% write.table(file = "trbi_6studies_20260306_table.txt", row.names = F, col.names = T, sep = "\t", quote = F)
so_merged_orig@meta.data %>% count(dataset,sample_name, name = "n_cells") %>% write.table(file = saveRDS(queries, file = sprintf("%s/trbi_8studies_20260311_table.tsv",out_dir)), row.names = F, col.names = T, sep = "\t", quote = F)

so_merged = so_merged_orig[,!so_merged_orig$level1_final %in% c("nonT", "unclear")]

so_merged@meta.data %>% count(dataset, sample_name, organ_simplified, target_cells_simplified, Ag_spe, infection_phase) 

```

Convert to new annotation if needed (done before now)
```{r}
# newann = read.table('/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/20260206_freeze/annotation_table_20260201_conversion_forOdhran.csv', header = T, sep = ",")
# so_merged$level2_final.v1 = so_merged$level2_final
# so_merged$level2_final = newann$level2.new[match(so_merged$level2_final.v1, newann$level2)]
# any(is.na(so_merged$level2_final))

# so_merged@meta.data %>% count(sample_name)
# so_orig@meta.data %>% count(infection_agent)
# so_merged$infection_agent = NA
# so_merged$infection_agent[grepl(x = so_merged$sample_name, pattern = "Lm")] = "Listeria monocytogenes"
# so_merged$infection_agent[grepl(x = so_merged$sample_name, pattern = "Arm")] = "LCMV Armstrong"
# so_merged$infection_agent[grepl(x = so_merged$sample_name, pattern = "Cl13")] = "LCMV clone 13"
# so_merged$infection_agent[grepl(x = so_merged$sample_name, pattern = "Blasto")] = "Blastomyces dermatitidis"


```

**T-RBI MDE plots**

MDE plot for all sample types
```{r}
so_merged@meta.data %>% count(sample_name, sample_type)
so_merged@meta.data %>% count(sample_type)
# i = "Arm_d10"

pdf(sprintf("%s/TRBI_MDEsample_type.pdf",out_dir),7, 7, useDingbats = F)
for (i in unique(so_merged$sample_type)) {
    print(i)
    tmp = so_merged[,so_merged@meta.data %>% filter(sample_type == i) %>% rownames()]
    
    p = DimPlot(tmp, reduction = "mde_incremental", group.by = "annotation_level2", cols = mypal_level2) + ggtitle(i)
    bg = AddEmbeddingRasterBackground(lineage = "CD4")
    p$layers = c(list(bg), p$layers)
    p = p + ggplot2::coord_fixed() + xlim(c(-2.5,4)) + ylim(c(-2.5,2.5)) +theme_void() + NoLegend()
    print(p)
}
dev.off()

```

Listeria Osum et al.
```{r}
so_merged_orig = readRDS(sprintf("%s/trbi_8studies_20260311_merged.Rds",out_dir))
so_merged_orig$cellID = colnames(so_merged_orig)
so_merged = so_merged_orig

so_merged@meta.data %>% count(dataset)
so_merged@meta.data %>% filter(dataset == "GSE281547_Osum_Lm") %>% count(sample_name)
sub = so_merged@meta.data %>% filter(dataset == "GSE281547_Osum_Lm") %>% group_by(sample_name) %>% slice_sample(n = 4500) %>% ungroup() %>% pull(cellID)
tmp = so_merged[,sub]
tmp@meta.data %>% count(sample_name)
    
pdf(sprintf("%s/TRBI_MDE_ListeriaOsum.pdf",out_dir),7, 7, useDingbats = F)
for (i in unique(tmp$sample_type)) {
    print(i)
    tmp = so_merged[,so_merged@meta.data %>% filter(sample_type == i) %>% rownames()]
    
    p = DimPlot(tmp, reduction = "mde_incremental", group.by = "annotation_level2", cols = mypal_level2) + ggtitle(i)
    bg = AddEmbeddingRasterBackground(lineage = "CD4")
    p$layers = c(list(bg), p$layers)
    p = p + ggplot2::coord_fixed() + xlim(c(-2.5,4)) + ylim(c(-2.5,2.5)) +theme_void() + NoLegend()
    print(p)
}
dev.off()

```

LCMV GP66 tet in SLO: GSE121002_Ciucci_LCMVArm, GSE158896_Khatun_LCMVArm, GSE182320_Andreatta_LCMVArmCl13
```{r}
# so_merged_orig = readRDS(sprintf("%s/trbi_8studies_20260311_merged.Rds",out_dir))
# so_merged_orig$cellID = colnames(so_merged_orig)
# so_merged = so_merged_orig

so_merged@meta.data %>% count(dataset, sample_type, Ag_spe, infection_agent)
so_merged@meta.data %>% filter(dataset %in% c("GSE121002_Ciucci_LCMVArm", "GSE158896_Khatun_LCMVArm", "GSE182320_Andreatta_LCMVArmCl13"), infection_agent == "LCMV Armstrong") %>% count(sample_type)
sub = so_merged@meta.data %>% filter(dataset %in% c("GSE121002_Ciucci_LCMVArm", "GSE158896_Khatun_LCMVArm", "GSE182320_Andreatta_LCMVArmCl13"), infection_agent == "LCMV Armstrong", !(sample_type %in% c("Arm_d30_SLO"))) %>% group_by(sample_type) %>% slice_sample(n = 6400) %>% ungroup() %>% pull(cellID)
tmp = so_merged[,sub]
tmp@meta.data %>% count(sample_name)
    
pdf(sprintf("%s/TRBI_MDE_LCMV_SLO_GP66tet.pdf",out_dir),7, 7, useDingbats = F)
for (i in unique(tmp$sample_type)) {
    print(i)
    tmp = so_merged[,so_merged@meta.data %>% filter(sample_type == i) %>% rownames()]
    
    p = DimPlot(tmp, reduction = "mde_incremental", group.by = "annotation_level2", cols = mypal_level2) + ggtitle(i)
    bg = AddEmbeddingRasterBackground(lineage = "CD4")
    p$layers = c(list(bg), p$layers)
    p = p + ggplot2::coord_fixed() + xlim(c(-2.5,4)) + ylim(c(-2.5,2.5)) +theme_void() + NoLegend()
    print(p)
}
dev.off()

```

LCMV Smarta in SLO: GSE175742_Shaw_LCMV, GSE184259_Nguyen_LCMV
```{r}
# so_merged_orig = readRDS(sprintf("%s/trbi_8studies_20260311_merged.Rds",out_dir))
# so_merged_orig$cellID = colnames(so_merged_orig)
# so_merged = so_merged_orig

so_merged@meta.data %>% count(dataset, sample_type, Ag_spe, infection_agent)
so_merged@meta.data %>% filter(dataset %in% c("GSE175742_Shaw_LCMV", "GSE184259_Nguyen_LCMV"), infection_agent == "LCMV Armstrong") %>% count(sample_type, organ_simplified)
sub = so_merged@meta.data %>% filter(dataset %in% c("GSE175742_Shaw_LCMV", "GSE184259_Nguyen_LCMV"), infection_agent == "LCMV Armstrong", organ_simplified == "SLO", !(sample_name %in% c("Arm_d7_SLO_Bcl6KO_GSE184259"))) %>% group_by(sample_type) %>% slice_sample(n = 2900) %>% ungroup() %>% pull(cellID)
tmp = so_merged[,sub]
tmp@meta.data %>% count(sample_name)
    
pdf(sprintf("%s/TRBI_MDE_LCMV_SLO_Smarta.pdf",out_dir),7, 7, useDingbats = F)
for (i in unique(tmp$sample_type)) {
    print(i)
    tmp = so_merged[,so_merged@meta.data %>% filter(sample_type == i) %>% rownames()]
    
    p = DimPlot(tmp, reduction = "mde_incremental", group.by = "annotation_level2", cols = mypal_level2) + ggtitle(i)
    bg = AddEmbeddingRasterBackground(lineage = "CD4")
    p$layers = c(list(bg), p$layers)
    p = p + ggplot2::coord_fixed() + xlim(c(-2.5,4)) + ylim(c(-2.5,2.5)) +theme_void() + NoLegend()
    print(p)
}
dev.off()

```

**T-RBI Plot % cluster by sample, infection phase etc**

Calculate proportions, LCMV and Listeria, SLO only
```{r}
library(dplyr)
library(ggplot2)
library(forcats)
library(scales)
library(stringr)

cluster_col <- "annotation_level2"   # change if needed

so_merged@meta.data %>% filter(infection_phase %in% c("acute", "memory"), organ_simplified %in% "SLO", infection_agent %in% c("LCMV Armstrong", "Listeria monocytogenes"), !(sample_name %in% c("Arm_d7_SLO_Bcl6KO_GSE184259"))) %>% count(dataset, sample_name, infection_agent, infection_phase, organ_simplified, Ag_spe)

df_plot <- so_merged@meta.data %>%
  filter(infection_phase %in% c("acute", "memory"), organ_simplified %in% "SLO", infection_agent %in% c("LCMV Armstrong", "Listeria monocytogenes"), !(sample_name %in% c("Arm_d7_SLO_Bcl6KO_GSE184259"))) %>%
  count(dataset, sample_name, infection_agent, infection_phase, .data[[cluster_col]], name = "n") %>%
  group_by(dataset, sample_name, infection_phase) %>%
  mutate(
    sample_total = sum(n),
    pct = 100 * n / sample_total
  ) %>%
  ungroup() %>%
  rename(cluster = all_of(cluster_col))
df_plot$cluster = factor(df_plot$cluster, levels = levels(so$annotation_level2))

df_plot2 <- df_plot %>%
  filter(infection_phase %in% c("acute","memory")) %>%
  mutate(
    cluster_phase = paste(cluster, infection_phase)
  )
clusters <- levels(so$annotation_level2)[grepl("^CD4", levels(so$annotation_level2))]
x_levels <- as.vector(
  rbind(
    paste(clusters, "acute"),
    paste(clusters, "memory")
  )
)
df_plot2$cluster_phase <- factor(df_plot2$cluster_phase, levels = x_levels)

```

LCMV only, SLO
```{r}

df_plot3 = df_plot2 %>% filter(!(cluster %in% c("CD4.wM", "CD4.wN", "CD4.wZ")), infection_agent %in% c("LCMV Armstrong"))
df_plot3$cluster = factor(df_plot3$cluster, levels = levels(so$annotation_level2))


p = ggplot(df_plot3, aes(cluster_phase, pct)) +
    # geom_point(
    #     aes(color = dataset),
    #     size = 1,
    #     alpha = 1
    # ) +
    geom_jitter(
        aes(color = dataset),
        width = 0.15,
        size = 1
    )+
    stat_summary(
        fun = mean,
        geom = "crossbar",
        width = 0.5,
        color = "black"
    )+
    facet_wrap(~ infection_phase, scales = "free_x") +
    scale_color_manual(values = mypal) +
    theme_bw() +
    theme(
        axis.text.x = element_text(angle = 60, hjust = 1),
        panel.grid.major.x = element_blank()
    ) +
    labs(
        x = "Cluster / infection phase",
        y = "% cells"
    )
p

# p2 = ggplot(df_plot3, aes(cluster_phase, pct)) +
#     # geom_point(
#     #     aes(color = infection_phase),
#     #     size = 1,
#     #     alpha = 1
#     # ) +
#     geom_jitter(
#         aes(color = infection_phase),
#         width = 0.15,
#         size = 1
#     )+
#     stat_summary(
#         fun = mean,
#         geom = "crossbar",
#         width = 0.5,
#         color = "black"
#     )+
#     facet_wrap(~ dataset, scales = "free_x") +
#     scale_color_manual(values = c("acute" = "blue", "memory" = "red")) +
#     theme_bw() +
#     theme(
#         axis.text.x = element_text(angle = 60, hjust = 1),
#         panel.grid.major.x = element_blank()
#     ) +
#     labs(
#         x = "Cluster / infection phase",
#         y = "% cells"
#     )
# p2

p2 = ggplot(df_plot3, aes(cluster, pct, color = infection_phase, shape = dataset)) +
    geom_jitter(
        position = position_jitterdodge(
            jitter.width = 0.1,
            dodge.width = 0.6
        ),
        size = 1
    ) +
    stat_summary(
        aes(group = infection_phase),
        fun = mean,
        geom = "crossbar",
        width = 0.5,
        position = position_dodge(width = 0.6),
        color = "black"
    ) +
    facet_wrap(~ dataset, scales = "free_x") +
    scale_color_manual(values = c("acute" = "blue", "memory" = "red")) +
    theme_bw() +
    theme(
        axis.text.x = element_text(angle = 60, hjust = 1),
        panel.grid.major.x = element_blank()
    ) +
    labs(
        x = "Cluster",
        y = "% cells"
    )
p2

p3 = ggplot(df_plot3, aes(cluster, pct, color = infection_phase, shape = dataset)) +
    geom_jitter(
        position = position_jitterdodge(
            jitter.width = 0.1,
            dodge.width = 0.6
        ),
        size = 2
    ) +
    stat_summary(
        aes(group = infection_phase),
        fun = mean,
        geom = "crossbar",
        width = 0.5,
        position = position_dodge(width = 0.6),
        color = "black"
    ) +
    scale_color_manual(values = c("acute" = "blue", "memory" = "red")) +
    theme_bw() +
    theme(
        axis.text.x = element_text(angle = 60, hjust = 1),
        panel.grid.major.x = element_blank()
    ) +
    labs(
        x = "Cluster",
        y = "% cells"
    )
p3

pdf(sprintf("%s/ClusterProp_LCMV_acute_memory.pdf", out_dir), 10,5, useDingbats = F)
p + NoGrid()
p2 + NoGrid()
p3 + NoGrid()
dev.off()

```

Listeria only
```{r}
df_plot3 = df_plot2 %>% filter(!(cluster %in% c("CD4.wM", "CD4.wN", "CD4.wZ")), infection_agent %in% c("Listeria monocytogenes"))
df_plot3$cluster = factor(df_plot3$cluster, levels = levels(so$annotation_level2))

p3 = ggplot(df_plot3, aes(cluster, pct, color = infection_phase, shape = dataset)) +
    geom_jitter(
        position = position_jitterdodge(
            jitter.width = 0.1,
            dodge.width = 0.6
        ),
        size = 3
    ) +
    stat_summary(
        aes(group = infection_phase),
        fun = mean,
        geom = "crossbar",
        width = 0.5,
        position = position_dodge(width = 0.6),
        color = "black"
    ) +
    scale_color_manual(values = c("acute" = "blue", "memory" = "red")) +
    theme_bw() +
    theme(
        axis.text.x = element_text(angle = 60, hjust = 1),
        panel.grid.major.x = element_blank()
    ) +
    labs(
        x = "Cluster",
        y = "% cells"
    )
p3

pdf(sprintf("%s/ClusterProp_Listeria_acute_memory.pdf", out_dir), 7,5, useDingbats = F)
p3 + NoGrid()
dev.off()

```

**Pseudobulk of T-RBI studies**
```{r}
so_merged = readRDS("trbi_8studies_20260311_merged.Rds")
head(so_merged@meta.data)

so_merged$pb_group <- paste(
    so_merged$sample_name,
    so_merged$annotation_level2,
    sep = "."
) %>% gsub("\\_", ".", .)


pb = AggregateExpression(
    object = so_merged,
    assay = "RNA",
    group.by = "pb_group",
    # features = tf,
    normalization.method = "LogNormalize",
    scale.factor = 10000,
    verbose = T,
    return.seurat = TRUE
)

so_merged@meta.data %>% colnames()
pb_meta <- so_merged@meta.data %>%
    dplyr::select(dataset, sample_name, sample_type, annotation_level2, organ_simplified, infection_agent, infection_phase, Ag_spe, tumor, pb_group) %>% group_by(pb_group) %>% mutate(ncells = n()) %>% ungroup() %>% distinct(pb_group, .keep_all = TRUE) %>% as.data.frame()

rownames(pb_meta) <- pb_meta$pb_group
pb_meta <- pb_meta[colnames(pb), ]

pb@meta.data = pb_meta

# saveRDS(pb, file = "trbi_6studies_20260306_pseudobulk.Rds")
saveRDS(pb, file = sprintf("%s/trbi_8studies_20260311_pseudobulk.Rds", out_dir) )

```

**Pseudobulk DGE memory vs acute analysis**
DGE limma - genes that are consistently different between memory and acute CD4 T cells in SLO, across all clusters, infections (LCMV arm and Listeria), and datasets.
```{r}
library(edgeR)

pb_orig = readRDS(sprintf("%s/trbi_8studies_20260311_pseudobulk.Rds", out_dir))

pb_orig@meta.data %>% count(dataset, sample_name, infection_phase)
pb = pb_orig[,pb_orig@meta.data %>% filter(infection_agent %in% c("LCMV Armstrong", "Listeria monocytogenes"), infection_phase %in% c("acute", "memory"), dataset %in% c("GSE121002_Ciucci_LCMVArm", "GSE182320_Andreatta_LCMVArmCl13", "GSE175742_Shaw_LCMV", "GSE281547_Osum_Lm"), !(annotation_level2 %in% c("CD4.wM", "CD4.wZ", "CD4.wN", "CD4.P", "CD4.U", "CD4.V", "CD4.X", "CD4.Y", "CD4.E", "CD4.F")), ncells > 10) %>% rownames()]
dim(pb)

#check that there are at least 2 replicates per cluster timepoint 
pb@meta.data %>% count(infection_phase, annotation_level2)

dge = DGEList(counts = pb@assays$RNA$counts)
dge = calcNormFactors(dge)

meta = pb@meta.data
meta %>% count(dataset, sample_name, annotation_level2,  infection_agent, infection_phase)
meta$annotation_level2 = factor(meta$annotation_level2, levels = unique(meta$annotation_level2))
meta$infection_phase = factor(meta$infection_phase, levels = c("acute", "memory"))
meta$infection_agent = make.names(meta$infection_agent)
meta$infection_agent = factor(meta$infection_agent, levels= c("LCMV.Armstrong", "Listeria.monocytogenes"))
meta$sample_name = factor(meta$sample_name)
meta$dataset = factor(meta$dataset)

# design <- model.matrix(
#   ~ 0 + level2_final*infection_phase + infection_agent + dataset,
#   data = meta
# )
# colnames(design) = gsub(colnames(design), pattern = "level2_final", replacement = "")
# colnames(design) = gsub(colnames(design), pattern = "dataset", replacement = "")
# colnames(design) = gsub(colnames(design), pattern = "infection_phase", replacement = "")
# colnames(design) = gsub(colnames(design), pattern = "infection_agent", replacement = "")
# colnames(design) = make.names(colnames(design))
# colnames(design) 

design <- model.matrix(
  ~ 0 + annotation_level2*infection_phase + infection_agent + sample_name,
  data = meta
)
colnames(design) = gsub(colnames(design), pattern = "annotation_level2", replacement = "")
colnames(design) = gsub(colnames(design), pattern = "sample_name", replacement = "")
colnames(design) = gsub(colnames(design), pattern = "infection_phase", replacement = "")
colnames(design) = gsub(colnames(design), pattern = "infection_agent", replacement = "")
colnames(design) = make.names(colnames(design))
colnames(design) 

#Filter genes using the design-aware approach:
keep_genes <- filterByExpr(dge, design = design)
dge <- dge[keep_genes, , keep.lib.sizes = FALSE]
dge <- calcNormFactors(dge)

library(limma)
v = voom(dge, design, plot = TRUE)
fit = lmFit(v, design)

tt_list = list()

#acute vs memory across all clusters and both infection agents
contrast <- makeContrasts(
    memory,
    levels = design
)
fit2 <- contrasts.fit(fit, contrast)
fit2 <- eBayes(fit2)
tt = topTable(fit2, number = Inf, adjust.method = "BH", sort.by = "none")
head(tt)
i = "memory_vs_acute"
tt_list[[i]] = tt
tt_list[[i]]$SYMBOL = rownames(tt)

##acute vs memory in CD4.Q
# contrast <- makeContrasts(
#     CD4.Q.memory+memory,
#     levels = design
# )
# fit2 <- contrasts.fit(fit, contrast)
# fit2 <- eBayes(fit2)
# tt = topTable(fit2, number = Inf, adjust.method = "BH", sort.by = "none")
# head(tt)
# tt %>% arrange(desc(logFC)) %>% head(100)
# i = "CD4.Q.memory_vs_acute"
# tt_list[[i]] = tt
# tt_list[[i]]$SYMBOL = rownames(tt)

##acute vs memory in each cluster
clusters <- make.names(levels(meta$annotation_level2))
contrast_formulas <- sapply(clusters, function(cl) {
  if (cl == clusters[1]) {
    "memory"
  } else {
    paste0("memory + ", cl, ".memory")
  }
})
names(contrast_formulas) <- paste0(clusters, "_memory_vs_acute")
contrast_formulas

contrast_matrix <- makeContrasts(
  contrasts = contrast_formulas,
  levels = design
)

fit2 <- contrasts.fit(fit, contrast_matrix)
fit2 <- eBayes(fit2)

tt_list <- lapply(colnames(contrast_matrix), function(coef_name) {
  tt <- topTable(
    fit2,
    coef = coef_name,
    number = Inf,
    adjust.method = "BH",
    sort.by = "none"
  )
  tt$SYMBOL <- rownames(tt)
  tt$contrast <- coef_name
  tt$cluster <- sub("_memory_vs_acute$", "", coef_name)
  tt
})

names(contrast_formulas)
colnames(contrast_matrix)

names(tt_list) <- names(contrast_formulas)
names(tt_list)[1] = "memory_vs_acute"

# vplot = data.frame(fc = 2^(tt$logFC), pval = tt$adj.P.Val, SYMBOL = rownames(tt))
# ZemmourLib::Vplot(vplot, ylimits = c(10^-50, 1), xlimits = NULL) + ggtitle(label = "memory_vs_acute_all")

saveRDS(tt_list, sprintf("%s/ttlist_memory_vs_acute_SLO_LCMVListeriaCombined.Rds", out_dir))

prefix_file_name = "memory_vs_acute_SLO_LCMVListeriaCombined"
tmp = tt_list
pdf(sprintf("%s/DiffExpression_volcano_%s.pdf", out_dir, prefix_file_name), width = 10, height = 10)
# i = "memory_vs_acute_all2"
for (i in 1:length(tmp)) {
    print(i)
   
     vplot = na.omit(data.frame(SYMBOL = tt_list[[i]]$SYMBOL, fc = 2^tt_list[[i]]$logFC, pval = tt_list[[i]]$P.Value + 10^-300, AveExpr = 2^(tt_list[[i]]$AveExpr))) #dealing with NA
    # vplot = data.frame(SYMBOL = tmp[[i]]$SYMBOL, fc = 2^tmp[[i]]$logFC, pval = tmp[[i]]$P.Value+10^-300, AveExpr = 2^(tmp[[i]]$AveExpr))
    p = Vplot(vplot = vplot, xlab = names(tmp)[i], xlimits = c(min(vplot$fc[vplot$fc != 0]), max(vplot$fc)), ylimits = c(min(vplot$pval), 1)) #c(min(vplot$pval[vplot$pval != 0]),1)
    sub = vplot %>%
        filter(abs(log2(fc)) > 0.5 & pval < 0.05) %>%
        arrange(desc(log2(fc))) %>%
        pull(SYMBOL)
    highlight_genes = unique(c(head(sub, 50), tail(sub, 50)))
    # sub = (log2(vplot$fc) > 0.5 | log2(vplot$fc) < -0.5)  & vplot$pval < 0.05
    # sub = vplot[sub,] %>% arrange(desc(log2(fc)))
    # sub1 = sub %>% pull(SYMBOL) %>% head(25) #vplot$SYMBOL %in% as.character(sub2$SYMBOL)[1:200]
    # sub2 = sub %>% pull(SYMBOL) %>% tail(25)
    highlight_filter = vplot$SYMBOL %in% highlight_genes
    # sub = vplot$SYMBOL %in% c(sub1, sub2)#as.character(sub2$SYMBOL)[1:50]
    library(ggrepel)
    print(p + geom_text_repel(data = vplot[highlight_filter,], aes(x = fc, y = pval, label = SYMBOL)))
    # print(p + geom_text(data = vplot[highlight_filter,], aes(x = fc, y = pval, label = SYMBOL)))
}
dev.off()

#LCMV Arm SLO - acute vs memory
#Listeria SLO - acute vs memory
#LCMV Arm and Listeria - acute vs memory
#LCMV Arm SLO - acute vs memory in each cluster
#Listeria SLO - acute vs memory


```

DGE limma - genes that are consistently different between memory and acute CD4 T cells in SLO, across all clusters, LCMV only
```{r}
library(edgeR)

pb_orig = readRDS(sprintf("%s/trbi_8studies_20260311_pseudobulk.Rds", out_dir))

pb_orig@meta.data %>% count(dataset, sample_name)
pb = pb_orig[,pb_orig@meta.data %>% filter(infection_agent %in% c("LCMV Armstrong"), infection_phase %in% c("acute", "memory"), dataset %in% c("GSE121002_Ciucci_LCMVArm", "GSE182320_Andreatta_LCMVArmCl13", "GSE175742_Shaw_LCMV"), !(annotation_level2 %in% c("CD4.wM", "CD4.wZ", "CD4.wN", "CD4.P", "CD4.U", "CD4.V", "CD4.X", "CD4.Y", "CD4.E", "CD4.F")), ncells > 10) %>% rownames()]
dim(pb)
pb@meta.data %>% count(dataset, sample_name)

#check that there are at least 2 replicates per cluster timepoint 
pb@meta.data %>% count(infection_phase, annotation_level2)

dge = DGEList(counts = pb@assays$RNA$counts)
dge = calcNormFactors(dge)

meta = pb@meta.data
meta %>% count(dataset, sample_name, annotation_level2,  infection_agent, infection_phase)
meta$annotation_level2 = factor(meta$annotation_level2, levels = unique(meta$annotation_level2))
meta$infection_phase = factor(meta$infection_phase, levels = c("acute", "memory"))
meta$infection_agent = make.names(meta$infection_agent)
meta$infection_agent = factor(meta$infection_agent, levels= c("LCMV.Armstrong"))
meta$sample_name = factor(meta$sample_name)
meta$dataset = factor(meta$dataset)

# design <- model.matrix(
#   ~ 0 + level2_final*infection_phase + infection_agent + dataset,
#   data = meta
# )
# colnames(design) = gsub(colnames(design), pattern = "level2_final", replacement = "")
# colnames(design) = gsub(colnames(design), pattern = "dataset", replacement = "")
# colnames(design) = gsub(colnames(design), pattern = "infection_phase", replacement = "")
# colnames(design) = gsub(colnames(design), pattern = "infection_agent", replacement = "")
# colnames(design) = make.names(colnames(design))
# colnames(design) 

design <- model.matrix(
  ~ 0 + annotation_level2*infection_phase + sample_name,
  data = meta
)
colnames(design) = gsub(colnames(design), pattern = "annotation_level2", replacement = "")
colnames(design) = gsub(colnames(design), pattern = "sample_name", replacement = "")
colnames(design) = gsub(colnames(design), pattern = "infection_phase", replacement = "")
colnames(design) = make.names(colnames(design))
colnames(design) 

#Filter genes using the design-aware approach:
keep_genes <- filterByExpr(dge, design = design)
dge <- dge[keep_genes, , keep.lib.sizes = FALSE]
dge <- calcNormFactors(dge)

library(limma)
v = voom(dge, design, plot = TRUE)
fit = lmFit(v, design)

tt_list = list()

#acute vs memory across all clusters and both infection agents
# contrast <- makeContrasts(
#     memory,
#     levels = design
# )
# fit2 <- contrasts.fit(fit, contrast)
# fit2 <- eBayes(fit2)
# tt = topTable(fit2, number = Inf, adjust.method = "BH", sort.by = "none")
# head(tt)
# i = "memory_vs_acute"
# tt_list[[i]] = tt
# tt_list[[i]]$SYMBOL = rownames(tt)

##acute vs memory in CD4.Q
# contrast <- makeContrasts(
#     CD4.Q.memory+memory,
#     levels = design
# )
# fit2 <- contrasts.fit(fit, contrast)
# fit2 <- eBayes(fit2)
# tt = topTable(fit2, number = Inf, adjust.method = "BH", sort.by = "none")
# head(tt)
# tt %>% arrange(desc(logFC)) %>% head(100)
# i = "CD4.Q.memory_vs_acute"
# tt_list[[i]] = tt
# tt_list[[i]]$SYMBOL = rownames(tt)

##acute vs memory in each cluster
clusters <- make.names(levels(meta$annotation_level2))
contrast_formulas <- sapply(clusters, function(cl) {
  if (cl == clusters[1]) {
    "memory"
  } else {
    paste0("memory + ", cl, ".memory")
  }
})
names(contrast_formulas) <- paste0(clusters, "_memory_vs_acute")
contrast_formulas

contrast_matrix <- makeContrasts(
  contrasts = contrast_formulas,
  levels = design
)

fit2 <- contrasts.fit(fit, contrast_matrix)
fit2 <- eBayes(fit2)

tt_list <- lapply(colnames(contrast_matrix), function(coef_name) {
  tt <- topTable(
    fit2,
    coef = coef_name,
    number = Inf,
    adjust.method = "BH",
    sort.by = "none"
  )
  tt$SYMBOL <- rownames(tt)
  tt$contrast <- coef_name
  tt$cluster <- sub("_memory_vs_acute$", "", coef_name)
  tt
})

names(contrast_formulas)
colnames(contrast_matrix)

names(tt_list) <- names(contrast_formulas)
names(tt_list)[1] = "memory_vs_acute"

saveRDS(tt_list, sprintf("%s/ttlist_memory_vs_acute_SLO_LCMV.Rds", out_dir))

# vplot = data.frame(fc = 2^(tt$logFC), pval = tt$adj.P.Val, SYMBOL = rownames(tt))
# ZemmourLib::Vplot(vplot, ylimits = c(10^-50, 1), xlimits = NULL) + ggtitle(label = "memory_vs_acute_all")

prefix_file_name = "memory_vs_acute_SLO_LCMV"
tmp = tt_list
pdf(sprintf("%s/DiffExpression_volcano_%s.pdf", out_dir, prefix_file_name), width = 10, height = 10)
# i = "memory_vs_acute_all2"
for (i in 1:length(tmp)) {
    print(i)
   
     vplot = na.omit(data.frame(SYMBOL = tt_list[[i]]$SYMBOL, fc = 2^tt_list[[i]]$logFC, pval = tt_list[[i]]$P.Value + 10^-300, AveExpr = 2^(tt_list[[i]]$AveExpr))) #dealing with NA
    # vplot = data.frame(SYMBOL = tmp[[i]]$SYMBOL, fc = 2^tmp[[i]]$logFC, pval = tmp[[i]]$P.Value+10^-300, AveExpr = 2^(tmp[[i]]$AveExpr))
    p = Vplot(vplot = vplot, xlab = names(tmp)[i], xlimits = c(min(vplot$fc[vplot$fc != 0]), max(vplot$fc)), ylimits = c(min(vplot$pval), 1)) #c(min(vplot$pval[vplot$pval != 0]),1)
    sub = vplot %>%
        filter(abs(log2(fc)) > 0.5 & pval < 0.05) %>%
        arrange(desc(log2(fc))) %>%
        pull(SYMBOL)
    highlight_genes = unique(c(head(sub, 50), tail(sub, 50)))
    # sub = (log2(vplot$fc) > 0.5 | log2(vplot$fc) < -0.5)  & vplot$pval < 0.05
    # sub = vplot[sub,] %>% arrange(desc(log2(fc)))
    # sub1 = sub %>% pull(SYMBOL) %>% head(25) #vplot$SYMBOL %in% as.character(sub2$SYMBOL)[1:200]
    # sub2 = sub %>% pull(SYMBOL) %>% tail(25)
    highlight_filter = vplot$SYMBOL %in% highlight_genes
    # sub = vplot$SYMBOL %in% c(sub1, sub2)#as.character(sub2$SYMBOL)[1:50]
    library(ggrepel)
    print(p + geom_text_repel(data = vplot[highlight_filter,], aes(x = fc, y = pval, label = SYMBOL)))
    # print(p + geom_text(data = vplot[highlight_filter,], aes(x = fc, y = pval, label = SYMBOL)))
}
dev.off()

#LCMV Arm SLO - acute vs memory
#Listeria SLO - acute vs memory
#LCMV Arm and Listeria - acute vs memory
#LCMV Arm SLO - acute vs memory in each cluster
#Listeria SLO - acute vs memory


```

DGE limma - genes that are consistently different between memory and acute CD4 T cells in SLO, across all clusters, Listeria only
```{r}
library(edgeR)

pb_orig = readRDS(sprintf("%s/trbi_8studies_20260311_pseudobulk.Rds", out_dir))

pb_orig@meta.data %>% count(dataset, sample_name)
pb = pb_orig[,pb_orig@meta.data %>% filter(infection_agent %in% c("Listeria monocytogenes"), infection_phase %in% c("acute", "memory"), dataset %in% c("GSE281547_Osum_Lm"), !(annotation_level2 %in% c("CD4.wM", "CD4.wZ", "CD4.wN", "CD4.P", "CD4.U", "CD4.V", "CD4.X", "CD4.Y", "CD4.E", "CD4.F")), ncells > 10) %>% rownames()]
dim(pb)
pb@meta.data %>% count(dataset, sample_name)

#check that there are at least 2 replicates per cluster timepoint 
pb@meta.data %>% count(infection_phase, annotation_level2)

dge = DGEList(counts = pb@assays$RNA$counts)
dge = calcNormFactors(dge)

meta = pb@meta.data
meta %>% count(dataset, sample_name, annotation_level2,  infection_agent, infection_phase)
meta$annotation_level2 = factor(meta$annotation_level2, levels = unique(meta$annotation_level2))
meta$infection_phase = factor(meta$infection_phase, levels = c("acute", "memory"))
# meta$infection_agent = make.names(meta$infection_agent)
# meta$infection_agent = factor(meta$infection_agent, levels= c("LCMV.Armstrong"))
meta$sample_name = factor(meta$sample_name)
meta$dataset = factor(meta$dataset)

# design <- model.matrix(
#   ~ 0 + level2_final*infection_phase + infection_agent + dataset,
#   data = meta
# )
# colnames(design) = gsub(colnames(design), pattern = "level2_final", replacement = "")
# colnames(design) = gsub(colnames(design), pattern = "dataset", replacement = "")
# colnames(design) = gsub(colnames(design), pattern = "infection_phase", replacement = "")
# colnames(design) = gsub(colnames(design), pattern = "infection_agent", replacement = "")
# colnames(design) = make.names(colnames(design))
# colnames(design) 

design <- model.matrix(
  ~ 0 + annotation_level2*infection_phase + sample_name,
  data = meta
)
colnames(design) = gsub(colnames(design), pattern = "annotation_level2", replacement = "")
colnames(design) = gsub(colnames(design), pattern = "sample_name", replacement = "")
colnames(design) = gsub(colnames(design), pattern = "infection_phase", replacement = "")
colnames(design) = make.names(colnames(design))
colnames(design) 
limma::nonEstimable(design)

#Filter genes using the design-aware approach:
keep_genes <- filterByExpr(dge, design = design)
dge <- dge[keep_genes, , keep.lib.sizes = FALSE]
dge <- calcNormFactors(dge)

library(limma)
v = voom(dge, design, plot = TRUE)
fit = lmFit(v, design)

tt_list = list()

#acute vs memory across all clusters and both infection agents
# contrast <- makeContrasts(
#     memory,
#     levels = design
# )
# fit2 <- contrasts.fit(fit, contrast)
# fit2 <- eBayes(fit2)
# tt = topTable(fit2, number = Inf, adjust.method = "BH", sort.by = "none")
# head(tt)
# i = "memory_vs_acute"
# tt_list[[i]] = tt
# tt_list[[i]]$SYMBOL = rownames(tt)

##acute vs memory in CD4.Q
# contrast <- makeContrasts(
#     CD4.Q.memory+memory,
#     levels = design
# )
# fit2 <- contrasts.fit(fit, contrast)
# fit2 <- eBayes(fit2)
# tt = topTable(fit2, number = Inf, adjust.method = "BH", sort.by = "none")
# head(tt)
# tt %>% arrange(desc(logFC)) %>% head(100)
# i = "CD4.Q.memory_vs_acute"
# tt_list[[i]] = tt
# tt_list[[i]]$SYMBOL = rownames(tt)

##acute vs memory in each cluster
clusters <- make.names(levels(meta$annotation_level2))
contrast_formulas <- sapply(clusters, function(cl) {
  if (cl == clusters[1]) {
    "memory"
  } else {
    paste0("memory + ", cl, ".memory")
  }
})
names(contrast_formulas) <- paste0(clusters, "_memory_vs_acute")
contrast_formulas

contrast_matrix <- makeContrasts(
  contrasts = contrast_formulas,
  levels = design
)

fit2 <- contrasts.fit(fit, contrast_matrix)

fit2 <- eBayes(fit2)

tt_list <- lapply(colnames(contrast_matrix), function(coef_name) {
  tt <- topTable(
    fit2,
    coef = coef_name,
    number = Inf,
    adjust.method = "BH",
    sort.by = "none"
  )
  tt$SYMBOL <- rownames(tt)
  tt$contrast <- coef_name
  tt$cluster <- sub("_memory_vs_acute$", "", coef_name)
  tt
})

names(contrast_formulas)
colnames(contrast_matrix)

names(tt_list) <- names(contrast_formulas)
names(tt_list)[1] = "memory_vs_acute"
tt_list[[1]]["Il7r",]

saveRDS(tt_list, sprintf("%s/ttlist_memory_vs_acute_SLO_Listeria.Rds", out_dir))

# vplot = data.frame(fc = 2^(tt$logFC), pval = tt$adj.P.Val, SYMBOL = rownames(tt))
# ZemmourLib::Vplot(vplot, ylimits = c(10^-50, 1), xlimits = NULL) + ggtitle(label = "memory_vs_acute_all")

prefix_file_name = "memory_vs_acute_SLO_Listeria"
tmp = tt_list
pdf(sprintf("%s/DiffExpression_volcano_%s.pdf", out_dir, prefix_file_name), width = 10, height = 10)
# i = "memory_vs_acute_all2"
for (i in 1:length(tmp)) {
    print(i)
   
     vplot = na.omit(data.frame(SYMBOL = tt_list[[i]]$SYMBOL, fc = 2^tt_list[[i]]$logFC, pval = tt_list[[i]]$P.Value + 10^-300, AveExpr = 2^(tt_list[[i]]$AveExpr))) #dealing with NA
    # vplot = data.frame(SYMBOL = tmp[[i]]$SYMBOL, fc = 2^tmp[[i]]$logFC, pval = tmp[[i]]$P.Value+10^-300, AveExpr = 2^(tmp[[i]]$AveExpr))
    p = Vplot(vplot = vplot, xlab = names(tmp)[i], xlimits = c(min(vplot$fc[vplot$fc != 0]), max(vplot$fc)), ylimits = c(min(vplot$pval), 1)) #c(min(vplot$pval[vplot$pval != 0]),1)
    sub = vplot %>%
        filter(abs(log2(fc)) > 0.5 & pval < 0.05) %>%
        arrange(desc(log2(fc))) %>%
        pull(SYMBOL)
    highlight_genes = unique(c(head(sub, 50), tail(sub, 50)))
    # sub = (log2(vplot$fc) > 0.5 | log2(vplot$fc) < -0.5)  & vplot$pval < 0.05
    # sub = vplot[sub,] %>% arrange(desc(log2(fc)))
    # sub1 = sub %>% pull(SYMBOL) %>% head(25) #vplot$SYMBOL %in% as.character(sub2$SYMBOL)[1:200]
    # sub2 = sub %>% pull(SYMBOL) %>% tail(25)
    highlight_filter = vplot$SYMBOL %in% highlight_genes
    # sub = vplot$SYMBOL %in% c(sub1, sub2)#as.character(sub2$SYMBOL)[1:50]
    library(ggrepel)
    print(p + geom_text_repel(data = vplot[highlight_filter,], aes(x = fc, y = pval, label = SYMBOL)))
    # print(p + geom_text(data = vplot[highlight_filter,], aes(x = fc, y = pval, label = SYMBOL)))
}
dev.off()

#LCMV Arm SLO - acute vs memory
#Listeria SLO - acute vs memory
#LCMV Arm and Listeria - acute vs memory
#LCMV Arm SLO - acute vs memory in each cluster
#Listeria SLO - acute vs memory


```

FCFC plot LCMV vs Listeria - acute vs memory
```{r}

tt_list_LCMV = readRDS(sprintf("%s/ttlist_memory_vs_acute_SLO_LCMV.Rds", out_dir))
tt_list_Lm = readRDS(sprintf("%s/ttlist_memory_vs_acute_SLO_Listeria.Rds", out_dir))

features = intersect(tt_list_LCMV[["memory_vs_acute"]]$SYMBOL, tt_list_Lm[["memory_vs_acute"]]$SYMBOL)
vplot = data.frame(fc1 = 2^tt_list_LCMV[["memory_vs_acute"]][features,"logFC"], fc2 = 2^tt_list_Lm[["memory_vs_acute"]][features,"logFC"], SYMBOL = features)
rownames(vplot) = vplot$SYMBOL
genes_to_label = vplot$SYMBOL[abs(log2(vplot$fc1)) > 2 | abs(log2(vplot$fc2)) > 2]
FCFCplot(vplot, genes_to_label = genes_to_label)

```

(Heatmap using pseudobulk, not working very well (too strong batch effects) )
```{r}
pb_orig = readRDS(sprintf("%s/trbi_8studies_20260311_pseudobulk.Rds", out_dir))
pb_orig@meta.data %>% count(dataset, sample_name)
pb = pb_orig[,pb_orig@meta.data %>% filter(infection_agent %in% c("LCMV Armstrong", "Listeria monocytogenes"), infection_phase %in% c("acute", "memory"), dataset %in% c("GSE121002_Ciucci_LCMVArm", "GSE182320_Andreatta_LCMVArmCl13", "GSE175742_Shaw_LCMV", "GSE281547_Osum_Lm"), !(annotation_level2 %in% c("CD4.wM", "CD4.wZ", "CD4.wN", "CD4.P", "CD4.U", "CD4.V", "CD4.X", "CD4.Y", "CD4.E", "CD4.F")), ncells > 10) %>% rownames()]
dim(pb)
pb_0 = pb

tt_list_LCMV = readRDS(sprintf("%s/ttlist_memory_vs_acute_SLO_LCMV.Rds", out_dir))
tt_list_Lm = readRDS(sprintf("%s/ttlist_memory_vs_acute_SLO_Listeria.Rds", out_dir))
tt_list_LCMVLm = readRDS(sprintf("%s/ttlist_memory_vs_acute_SLO_LCMVListeriaCombined.Rds", out_dir))

genes_to_label1 = tt_list_LCMV$memory_vs_acute %>% filter( abs(logFC > 2) & adj.P.Val < 0.05) %>% pull(SYMBOL) %>% head(50)
genes_to_label2 = tt_list_Lm$memory_vs_acute %>% filter( abs(logFC > 1) & adj.P.Val < 0.05) %>% pull(SYMBOL) #%>% head(20)
# genes_to_label = intersect(genes_to_label1, genes_to_label2)
genes_to_label = genes_to_label1


# genes_to_label = tt_list_LCMVLm$memory_vs_acute %>% filter( abs(logFC > 1) & adj.P.Val < 0.05) %>% pull(SYMBOL) #%>% head(20)



# DoHeatmap(
#   pb,
#   features = genes_to_label,
#   cells = NULL,
#   group.by = "annotation_level2",
#   group.bar = TRUE,
#   group.colors = NULL,
#   disp.min = -2.5,
#   disp.max = NULL,
#   slot = "scale.data",
#   assay = NULL,
#   label = TRUE,
#   size = 5.5,
#   hjust = 0,
#   vjust = 0,
#   angle = 45,
#   raster = TRUE,
#   draw.lines = TRUE,
#   lines.width = NULL,
#   group.bar.height = 0.02,
#   combine = TRUE
# ) + NoLegend()
# 
# library(ComplexHeatmap)
# library(circlize)

pb = pb_0[,pb_0@meta.data %>% filter(infection_agent %in% c("LCMV Armstrong")) %>% rownames()]
pb = ScaleData(pb, do.scale = F, do.center = T)
mat <- GetAssayData(pb, slot = "scale.data")[genes_to_label, ]
meta <- pb@meta.data
meta$infection_phase <- factor(meta$infection_phase, levels = c("acute", "memory"))
meta$annotation_level2 <- factor(meta$annotation_level2, levels = unique(meta$annotation_level2))
meta$cluster_phase <- paste(meta$annotation_level2, meta$infection_phase, sep = ".")
meta <- meta %>%
  mutate(col_id = rownames(.)) %>%
  arrange(annotation_level2, infection_phase)
mat <- mat[, meta$col_id, drop = FALSE]

cluster_cols <- mypal_level2
phase_cols <- c(
  acute = "blue",
  memory = "red"
)

agent_cols <- c(
  "LCMV Armstrong" = "#1b9e77",
  "Listeria monocytogenes" = "#d95f02",
  "Blastomyces" = "#7570b3"
)

ha <- HeatmapAnnotation(
  cluster = meta$annotation_level2,
  phase = meta$infection_phase,
  infection_agent = meta$infection_agent,
  col = list(
    cluster = cluster_cols,
    phase = phase_cols,
    infection_agent = agent_cols
  ),
  annotation_name_side = "left"
)

ht <- Heatmap(
  mat,
  name = "Z-score",
  top_annotation = ha,
  cluster_rows = TRUE,
  cluster_columns = FALSE,
  show_column_names = FALSE,
  column_split = meta$cluster_phase,
  use_raster = TRUE,
  row_names_gp = gpar(fontsize = 6)
)

draw(ht)


```

Heatmap using limma coefficients - LCMV only
```{r}
#LCMV or Listeria
tt_list_LCMV = readRDS(sprintf("%s/ttlist_memory_vs_acute_SLO_LCMV.Rds", out_dir))
# genes_to_label = tt_list_LCMV$memory_vs_acute %>% filter( abs(logFC) > 2 & adj.P.Val < 0.05) %>% pull(SYMBOL) %>% head(50)
tt = tt_list_LCMV

# genes_to_label = intersect(genes_to_label1, genes_to_label2)
# genes_to_label = genes_to_label1

#filter gene names
genes_to_label <- c()
for (i in names(tt)) {
  new_genes <- tt[[i]] %>%
    arrange(desc(abs(logFC))) %>%
    filter(adj.P.Val < 0.05) %>%
    pull(SYMBOL)
  new_genes <- setdiff(new_genes, genes_to_label)
  genes_to_label <- c(genes_to_label, head(new_genes, 10))
}

genes_to_label <- genes_to_label[
  !grepl(
    "^Gm|^mt-|^Rpl|^Rps|^Snor|^Mir|^Linc|^Rp[sl]|-ps|Rik$|^Hbb",
    genes_to_label,
    ignore.case = TRUE
  )
]

#plot heatmap
mat_coef <- sapply(tt, function(tt) tt$logFC)
colnames(mat_coef)[colnames(mat_coef) %in% "memory_vs_acute" ] = "CD4.A_memory_vs_acute"
rownames(mat_coef) <- tt[[1]]$SYMBOL
mat_plot = mat_coef[genes_to_label, , drop = FALSE]

library(ComplexHeatmap)
library(circlize)

col_fun <- circlize::colorRamp2(
  c(-7, 0, 7),
  c("#4575B4", "white", "#D73027")
)

pdf(sprintf("%s/heatmap_memory_vs_acute_SLO_LCMV.pdf", out_dir), width = 3, 8, useDingbats = F)
Heatmap(
  mat_plot,
  name = "log2FC",
  col = col_fun,
  cluster_columns = FALSE,
  cluster_rows = T,
    row_names_gp = gpar(fontsize = 6),
      column_names_gp = gpar(fontsize = 6),
  column_title = "Memory vs Acute DGE per cluster"
)
dev.off()

```

Heatmap using limma coefficients - Listeria only
```{r}
tt_list_Lm = readRDS(sprintf("%s/ttlist_memory_vs_acute_SLO_Listeria.Rds", out_dir))
# genes_to_label = tt_list_LCMV$memory_vs_acute %>% filter( abs(logFC) > 2 & adj.P.Val < 0.05) %>% pull(SYMBOL) %>% head(50)
tt = tt_list_Lm

# genes_to_label = intersect(genes_to_label1, genes_to_label2)
# genes_to_label = genes_to_label1

#filter gene names
genes_to_label <- c()
for (i in names(tt)) {
  new_genes <- tt[[i]] %>%
    arrange(desc(abs(logFC))) %>%
    filter(adj.P.Val < 0.05) %>%
    pull(SYMBOL)
  new_genes <- setdiff(new_genes, genes_to_label)
  genes_to_label <- c(genes_to_label, head(new_genes, 10))
}

genes_to_label <- genes_to_label[
  !grepl(
    "^Gm|^mt-|^Rpl|^Rps|^Snor|^Mir|^Linc|^Rp[sl]|-ps|Rik$|^Hbb",
    genes_to_label,
    ignore.case = TRUE
  )
]

#plot heatmap
mat_coef <- sapply(tt, function(tt) tt$logFC)
colnames(mat_coef)[colnames(mat_coef) %in% "memory_vs_acute" ] = "CD4.A_memory_vs_acute"
rownames(mat_coef) <- tt[[1]]$SYMBOL
mat_plot = mat_coef[genes_to_label, , drop = FALSE]

library(ComplexHeatmap)
library(circlize)

col_fun <- circlize::colorRamp2(
  c(-7, 0, 7),
  c("#4575B4", "white", "#D73027")
)

# pdf(sprintf("%s/heatmap_memory_vs_acute_SLO_LCMV.pdf", out_dir), width = 3, 8, useDingbats = F)
pdf(sprintf("%s/heatmap_memory_vs_acute_SLO_Listeria.pdf", out_dir), width = 3, 8, useDingbats = F)
Heatmap(
  mat_plot,
  name = "log2FC",
  col = col_fun,
  cluster_columns = FALSE,
  cluster_rows = T,
    row_names_gp = gpar(fontsize = 6),
      column_names_gp = gpar(fontsize = 6),
  column_title = "Memory vs Acute DGE per cluster"
)
dev.off()

```


**Pseudobulk DGE memory timepoint: DGE cluster differences at memory timepoint in LCMV only**
DGE limma - genes that are different between A,B,C vs G,H,I,Q,R,S CD4 T cells in SLO at memory timepoints, LCMV only
```{r}
library(edgeR)

pb_orig = readRDS(sprintf("%s/trbi_8studies_20260311_pseudobulk.Rds", out_dir))

pb_orig@meta.data %>% count(dataset, sample_name)
pb = pb_orig[,pb_orig@meta.data %>% filter(infection_agent %in% c("LCMV Armstrong"), infection_phase %in% c("memory"), dataset %in% c("GSE121002_Ciucci_LCMVArm", "GSE182320_Andreatta_LCMVArmCl13", "GSE175742_Shaw_LCMV"), !(annotation_level2 %in% c("CD4.wM", "CD4.wZ", "CD4.wN", "CD4.P", "CD4.T", "CD4.U", "CD4.X", "CD4.Y", "CD4.E", "CD4.F")), ncells > 10) %>% rownames()]
dim(pb)
pb@meta.data %>% count(dataset, sample_name)
pb@meta.data %>% count(dataset, sample_name)

#check that there are at least 2 replicates per cluster timepoint 
pb@meta.data %>% count(infection_phase, annotation_level2)

dge = DGEList(counts = pb@assays$RNA$counts)
dge = calcNormFactors(dge)

meta = pb@meta.data
meta %>% count(dataset, sample_name, annotation_level2,  infection_agent, infection_phase)
meta$annotation_level2 = factor(meta$annotation_level2, levels = unique(meta$annotation_level2))
meta$infection_phase = factor(meta$infection_phase, levels = c("acute", "memory"))
meta$infection_agent = make.names(meta$infection_agent)
meta$infection_agent = factor(meta$infection_agent, levels= c("LCMV.Armstrong"))
meta$sample_name = factor(meta$sample_name)
meta$dataset = factor(meta$dataset)

# design <- model.matrix(
#   ~ 0 + level2_final*infection_phase + infection_agent + dataset,
#   data = meta
# )
# colnames(design) = gsub(colnames(design), pattern = "level2_final", replacement = "")
# colnames(design) = gsub(colnames(design), pattern = "dataset", replacement = "")
# colnames(design) = gsub(colnames(design), pattern = "infection_phase", replacement = "")
# colnames(design) = gsub(colnames(design), pattern = "infection_agent", replacement = "")
# colnames(design) = make.names(colnames(design))
# colnames(design) 

design <- model.matrix(
  ~ 0 + annotation_level2 + sample_name,
  data = meta
)
colnames(design) = gsub(colnames(design), pattern = "annotation_level2", replacement = "")
colnames(design) = gsub(colnames(design), pattern = "sample_name", replacement = "")
colnames(design) = gsub(colnames(design), pattern = "infection_phase", replacement = "")
colnames(design) = make.names(colnames(design))
colnames(design) 

#Filter genes using the design-aware approach:
keep_genes <- filterByExpr(dge, design = design)
dge <- dge[keep_genes, , keep.lib.sizes = FALSE]
dge <- calcNormFactors(dge)

library(limma)
v = voom(dge, design, plot = TRUE)
fit = lmFit(v, design)

tt_list = list()

#Contrast for A,B,C vs G,H,IQ,R,S at memory
clusters <- grep("^CD4", colnames(design), value = TRUE)
contrast_formulas <- sapply(clusters, function(cl) {
  others <- setdiff(clusters, cl)
  paste0(
    cl,
    " - (",
    paste(others, collapse = " + "),
    ")/", length(others)
  )
})
names(contrast_formulas) <- paste0(clusters, "_vs_others")

contrast_formulas["CD4ABCD_vs_GHIQRS"] = "(CD4.A + CD4.B + CD4.C)/3 -(CD4.G + CD4.H + CD4.I + CD4.Q + CD4.R + CD4.S)/6"

contrast_matrix <- makeContrasts(
  contrasts = contrast_formulas,
  levels = design
)

fit2 <- contrasts.fit(fit, contrast_matrix)
fit2 <- eBayes(fit2)

tt_list <- lapply(colnames(contrast_matrix), function(coef_name) {
  tt <- topTable(
    fit2,
    coef = coef_name,
    number = Inf,
    adjust.method = "BH",
    sort.by = "none"
  )
  tt$SYMBOL <- rownames(tt)
  tt$contrast <- coef_name
  tt$cluster <- sub("_vs_others$", "", coef_name)
  tt
})

names(contrast_formulas)
colnames(contrast_matrix)

names(tt_list) <- names(contrast_formulas)

saveRDS(tt_list, sprintf("%s/ttlist_LCMV_SLO_memory_OneClusterVsAllOthers.Rds", out_dir))

# vplot = data.frame(fc = 2^(tt$logFC), pval = tt$adj.P.Val, SYMBOL = rownames(tt))
# ZemmourLib::Vplot(vplot, ylimits = c(10^-50, 1), xlimits = NULL) + ggtitle(label = "memory_vs_acute_all")

prefix_file_name = "LCMV_SLO_memory_OneClusterVsAllOthers"
tmp = tt_list
pdf(sprintf("%s/DiffExpression_volcano_%s.pdf", out_dir, prefix_file_name), width = 10, height = 10)
# i = "memory_vs_acute_all2"
for (i in 1:length(tmp)) {
    print(i)
   
     vplot = na.omit(data.frame(SYMBOL = tt_list[[i]]$SYMBOL, fc = 2^tt_list[[i]]$logFC, pval = tt_list[[i]]$P.Value + 10^-300, AveExpr = 2^(tt_list[[i]]$AveExpr))) #dealing with NA
    # vplot = data.frame(SYMBOL = tmp[[i]]$SYMBOL, fc = 2^tmp[[i]]$logFC, pval = tmp[[i]]$P.Value+10^-300, AveExpr = 2^(tmp[[i]]$AveExpr))
    p = Vplot(vplot = vplot, xlab = names(tmp)[i], xlimits = c(min(vplot$fc[vplot$fc != 0]), max(vplot$fc)), ylimits = c(min(vplot$pval), 1)) #c(min(vplot$pval[vplot$pval != 0]),1)
    sub = vplot %>%
        filter(abs(log2(fc)) > 0.5 & pval < 0.05) %>%
        arrange(desc(log2(fc))) %>%
        pull(SYMBOL)
    highlight_genes = unique(c(head(sub, 50), tail(sub, 50)))
    # sub = (log2(vplot$fc) > 0.5 | log2(vplot$fc) < -0.5)  & vplot$pval < 0.05
    # sub = vplot[sub,] %>% arrange(desc(log2(fc)))
    # sub1 = sub %>% pull(SYMBOL) %>% head(25) #vplot$SYMBOL %in% as.character(sub2$SYMBOL)[1:200]
    # sub2 = sub %>% pull(SYMBOL) %>% tail(25)
    highlight_filter = vplot$SYMBOL %in% highlight_genes
    # sub = vplot$SYMBOL %in% c(sub1, sub2)#as.character(sub2$SYMBOL)[1:50]
    library(ggrepel)
    print(p + geom_text_repel(data = vplot[highlight_filter,], aes(x = fc, y = pval, label = SYMBOL)))
    # print(p + geom_text(data = vplot[highlight_filter,], aes(x = fc, y = pval, label = SYMBOL)))
}
dev.off()

#LCMV Arm SLO - acute vs memory
#Listeria SLO - acute vs memory
#LCMV Arm and Listeria - acute vs memory
#LCMV Arm SLO - acute vs memory in each cluster
#Listeria SLO - acute vs memory

```

Heatmap using limma coefficients 
```{r}
#LCMV or Listeria
tt_list_LCMVmem = readRDS(sprintf("%s/ttlist_LCMV_SLO_memory_OneClusterVsAllOthers.Rds", out_dir))
# genes_to_label = tt_list_LCMVmem$memory_vs_acute %>% filter( abs(logFC) > 2 & adj.P.Val < 0.05) %>% pull(SYMBOL) %>% head(50)
tt = tt_list_LCMVmem 

comp_filt = names(tt)
comp_filt = comp_filt[grepl("_vs_others", comp_filt)]

#filter gene names
genes_to_label <- c()
for (i in comp_filt) {
  new_genes <- tt[[i]] %>%
    arrange(desc(abs(logFC))) %>%
    filter(adj.P.Val < 0.05) %>%
    pull(SYMBOL)
  new_genes <- setdiff(new_genes, genes_to_label)
  genes_to_label <- c(genes_to_label, head(new_genes, 10))
}

genes_to_label <- genes_to_label[
  !grepl(
    "^Gm|^mt-|^Rpl|^Rps|^Snor|^Mir|^Linc|^Rp[sl]|-ps|Rik$|^Hbb",
    genes_to_label,
    ignore.case = TRUE
  )
]

#plot heatmap
mat_coef <- sapply(tt, function(tt) tt$logFC)
rownames(mat_coef) <- tt[[1]]$SYMBOL
mat_plot = mat_coef[genes_to_label, comp_filt, drop = FALSE]

library(ComplexHeatmap)
library(circlize)

col_fun <- circlize::colorRamp2(
  c(-4, 0, 4),
  c("#4575B4", "white", "#D73027")
)

# pdf(sprintf("%s/heatmap_memory_vs_acute_SLO_LCMV.pdf", out_dir), width = 3, 8, useDingbats = F)
pdf(sprintf("%s/Heatmap_LCMV_SLO_memory_OneClusterVsAllOthers.pdf", out_dir), width = 3, 8, useDingbats = F)
Heatmap(
  mat_plot,
  name = "log2FC",
  col = col_fun,
  cluster_columns = FALSE,
  cluster_rows = F,
    row_names_gp = gpar(fontsize = 6),
      column_names_gp = gpar(fontsize = 6),
  column_title = "one cluster vs all others at memory timepoint"
)
dev.off()

```

**Pseudobulk DGE memory timepoint: DGE cluster differences at memory timepoint in Listeria only**
DGE limma - genes that are different between A,B,C vs G,H,I,Q,R,S CD4 T cells in SLO at memory timepoints, LCMV only
```{r}
library(edgeR)

pb_orig = readRDS(sprintf("%s/trbi_8studies_20260311_pseudobulk.Rds", out_dir))

pb_orig@meta.data %>% count(dataset, sample_name)
pb = pb_orig[,pb_orig@meta.data %>% filter(infection_agent %in% c("Listeria monocytogenes"), infection_phase %in% c("memory"), dataset %in% c("GSE281547_Osum_Lm"), !(annotation_level2 %in% c("CD4.wM", "CD4.wZ", "CD4.wN", "CD4.P", "CD4.T", "CD4.U", "CD4.X", "CD4.Y", "CD4.E", "CD4.F")), ncells > 10) %>% rownames()]
dim(pb)
pb@meta.data %>% count(dataset, sample_name)

#check that there are at least 2 replicates per cluster timepoint 
pb@meta.data %>% count(infection_phase, annotation_level2)

dge = DGEList(counts = pb@assays$RNA$counts)
dge = calcNormFactors(dge)

meta = pb@meta.data
meta %>% count(dataset, sample_name, annotation_level2,  infection_agent, infection_phase)
meta$annotation_level2 = factor(meta$annotation_level2, levels = unique(meta$annotation_level2))
meta$infection_phase = factor(meta$infection_phase, levels = c("acute", "memory"))
meta$infection_agent = make.names(meta$infection_agent)
meta$infection_agent = factor(meta$infection_agent, levels= c("Listeria.monocytogenes"))
meta$sample_name = factor(meta$sample_name)
meta$dataset = factor(meta$dataset)

# design <- model.matrix(
#   ~ 0 + level2_final*infection_phase + infection_agent + dataset,
#   data = meta
# )
# colnames(design) = gsub(colnames(design), pattern = "level2_final", replacement = "")
# colnames(design) = gsub(colnames(design), pattern = "dataset", replacement = "")
# colnames(design) = gsub(colnames(design), pattern = "infection_phase", replacement = "")
# colnames(design) = gsub(colnames(design), pattern = "infection_agent", replacement = "")
# colnames(design) = make.names(colnames(design))
# colnames(design) 

design <- model.matrix(
  ~ 0 + annotation_level2 + sample_name,
  data = meta
)
colnames(design) = gsub(colnames(design), pattern = "annotation_level2", replacement = "")
colnames(design) = gsub(colnames(design), pattern = "sample_name", replacement = "")
colnames(design) = gsub(colnames(design), pattern = "infection_phase", replacement = "")
colnames(design) = make.names(colnames(design))
colnames(design) 

#Filter genes using the design-aware approach:
keep_genes <- filterByExpr(dge, design = design)
dge <- dge[keep_genes, , keep.lib.sizes = FALSE]
dge <- calcNormFactors(dge)

library(limma)
v = voom(dge, design, plot = TRUE)
fit = lmFit(v, design)

tt_list = list()

#Contrast for A,B,C vs G,H,IQ,R,S at memory
clusters <- grep("^CD4", colnames(design), value = TRUE)
contrast_formulas <- sapply(clusters, function(cl) {
  others <- setdiff(clusters, cl)
  paste0(
    cl,
    " - (",
    paste(others, collapse = " + "),
    ")/", length(others)
  )
})
names(contrast_formulas) <- paste0(clusters, "_vs_others")

contrast_formulas["CD4ABCD_vs_GHIQRS"] = "(CD4.A + CD4.B + CD4.C)/3 -(CD4.G + CD4.H + CD4.I + CD4.Q + CD4.R + CD4.S)/6"

contrast_matrix <- makeContrasts(
  contrasts = contrast_formulas,
  levels = design
)

fit2 <- contrasts.fit(fit, contrast_matrix)
fit2 <- eBayes(fit2)

tt_list <- lapply(colnames(contrast_matrix), function(coef_name) {
  tt <- topTable(
    fit2,
    coef = coef_name,
    number = Inf,
    adjust.method = "BH",
    sort.by = "none"
  )
  tt$SYMBOL <- rownames(tt)
  tt$contrast <- coef_name
  tt$cluster <- sub("_vs_others$", "", coef_name)
  tt
})

names(contrast_formulas)
colnames(contrast_matrix)

names(tt_list) <- names(contrast_formulas)

saveRDS(tt_list, sprintf("%s/ttlist_Listeria_SLO_memory_OneClusterVsAllOthers.Rds", out_dir))

# vplot = data.frame(fc = 2^(tt$logFC), pval = tt$adj.P.Val, SYMBOL = rownames(tt))
# ZemmourLib::Vplot(vplot, ylimits = c(10^-50, 1), xlimits = NULL) + ggtitle(label = "memory_vs_acute_all")

prefix_file_name = "Listeria_SLO_memory_OneClusterVsAllOthers"
tmp = tt_list
pdf(sprintf("%s/DiffExpression_volcano_%s.pdf", out_dir, prefix_file_name), width = 10, height = 10)
# i = "memory_vs_acute_all2"
for (i in 1:length(tmp)) {
    print(i)
   
     vplot = na.omit(data.frame(SYMBOL = tt_list[[i]]$SYMBOL, fc = 2^tt_list[[i]]$logFC, pval = tt_list[[i]]$P.Value + 10^-300, AveExpr = 2^(tt_list[[i]]$AveExpr))) #dealing with NA
    # vplot = data.frame(SYMBOL = tmp[[i]]$SYMBOL, fc = 2^tmp[[i]]$logFC, pval = tmp[[i]]$P.Value+10^-300, AveExpr = 2^(tmp[[i]]$AveExpr))
    p = Vplot(vplot = vplot, xlab = names(tmp)[i], xlimits = c(min(vplot$fc[vplot$fc != 0]), max(vplot$fc)), ylimits = c(min(vplot$pval), 1)) #c(min(vplot$pval[vplot$pval != 0]),1)
    sub = vplot %>%
        filter(abs(log2(fc)) > 0.5 & pval < 0.05) %>%
        arrange(desc(log2(fc))) %>%
        pull(SYMBOL)
    highlight_genes = unique(c(head(sub, 50), tail(sub, 50)))
    # sub = (log2(vplot$fc) > 0.5 | log2(vplot$fc) < -0.5)  & vplot$pval < 0.05
    # sub = vplot[sub,] %>% arrange(desc(log2(fc)))
    # sub1 = sub %>% pull(SYMBOL) %>% head(25) #vplot$SYMBOL %in% as.character(sub2$SYMBOL)[1:200]
    # sub2 = sub %>% pull(SYMBOL) %>% tail(25)
    highlight_filter = vplot$SYMBOL %in% highlight_genes
    # sub = vplot$SYMBOL %in% c(sub1, sub2)#as.character(sub2$SYMBOL)[1:50]
    library(ggrepel)
    print(p + geom_text_repel(data = vplot[highlight_filter,], aes(x = fc, y = pval, label = SYMBOL)))
    # print(p + geom_text(data = vplot[highlight_filter,], aes(x = fc, y = pval, label = SYMBOL)))
}
dev.off()

```

Heatmap using limma coefficients 
```{r}
tt_list_Lm = readRDS(sprintf("%s/ttlist_Listeria_SLO_memory_OneClusterVsAllOthers.Rds", out_dir))
# genes_to_label = tt_list_Lm$memory_vs_acute %>% filter( abs(logFC) > 2 & adj.P.Val < 0.05) %>% pull(SYMBOL) %>% head(50)
tt = tt_list_Lm 

comp_filt = names(tt)
comp_filt = comp_filt[grepl("_vs_others", comp_filt)]

#filter gene names
genes_to_label <- c()
for (i in comp_filt) {
  new_genes <- tt[[i]] %>%
    arrange(desc(abs(logFC))) %>%
    filter(adj.P.Val < 0.05) %>%
    pull(SYMBOL)
  new_genes <- setdiff(new_genes, genes_to_label)
  genes_to_label <- c(genes_to_label, head(new_genes, 10))
}

genes_to_label <- genes_to_label[
  !grepl(
    "^Gm|^mt-|^Rpl|^Rps|^Snor|^Mir|^Linc|^Rp[sl]|-ps|Rik$|^Hbb",
    genes_to_label,
    ignore.case = TRUE
  )
]

#plot heatmap
mat_coef <- sapply(tt, function(tt) tt$logFC)
rownames(mat_coef) <- tt[[1]]$SYMBOL
mat_plot = mat_coef[genes_to_label, comp_filt, drop = FALSE]

library(ComplexHeatmap)
library(circlize)

col_fun <- circlize::colorRamp2(
  c(-5, 0, 5),
  c("#4575B4", "white", "#D73027")
)

# pdf(sprintf("%s/heatmap_memory_vs_acute_SLO_LCMV.pdf", out_dir), width = 3, 8, useDingbats = F)
pdf(sprintf("%s/Heatmap_Listeria_SLO_memory_OneClusterVsAllOthers.pdf", out_dir), width = 3, 8, useDingbats = F)
Heatmap(
  mat_plot,
  name = "log2FC",
  col = col_fun,
  cluster_columns = FALSE,
  cluster_rows = F,
    row_names_gp = gpar(fontsize = 6),
      column_names_gp = gpar(fontsize = 6),
  column_title = "one cluster vs all others at memory timepoint"
)
dev.off()

```

**GSE184259_Nguyen: Trm and BCL6 KO**

```{r}
# pdf(sprintf("%s/TRBI_MDEsample_type_GSE184259_Nguyen.pdf",out_dir),7, 7, useDingbats = F)
# so_merged@meta.data %>% filter(dataset == "GSE184259_Nguyen_LCMV") %>% count(sample_name)
# so_merged$cellID = colnames(so_merged)
# sub = so_merged@meta.data %>% filter(dataset == "GSE184259_Nguyen_LCMV", sample_type %in% c("Arm_d21_SILP", "Arm_d7_SILP")) %>% group_by(sample_type) %>% slice_sample(n = 3602) %>% ungroup() %>% pull(cellID)
# all(sub %in% colnames(so_merged))
# tmp0 = so_merged[,sub]
# for (i in unique(tmp$sample_type)) {
#     print(i)
#     tmp = so_merged[,so_merged@meta.data %>% filter(sample_type == i) %>% rownames()]
#     
#     p = DimPlot(tmp, reduction = "mde_incremental", group.by = "annotation_level2", cols = mypal_level2) + ggtitle(i)
#     bg = AddEmbeddingRasterBackground(lineage = "CD4")
#     p$layers = c(list(bg), p$layers)
#     p = p + ggplot2::coord_fixed() + xlim(c(-2.5,4)) + ylim(c(-2.5,2.5)) +theme_void() + NoLegend()
#     print(p)
# }
# dev.off()


so_merged_orig = readRDS(sprintf("%s/trbi_8studies_20260311_merged.Rds",out_dir))
so_merged_orig$cellID = colnames(so_merged_orig)
so_merged = so_merged_orig

so_merged@meta.data %>% count(dataset, sample_type, Ag_spe, infection_agent)
so_merged@meta.data %>% filter(dataset %in% c("GSE184259_Nguyen_LCMV"), infection_agent == "LCMV Armstrong") %>% count(sample_type, organ_simplified)
sub = so_merged@meta.data %>% filter(dataset %in% c("GSE184259_Nguyen_LCMV"), infection_agent == "LCMV Armstrong", organ_simplified %in% c("SLO", "SILP", "SIIEL"), !(sample_name %in% c("Arm_d7_SIIEL_Bcl6KO_GSE184259", "Arm_d7_SILP_Bcl6KO_GSE184259", "Arm_d7_SLO_Bcl6KO_GSE184259"))) %>% group_by(sample_type) %>% slice_sample(n = 1800) %>% ungroup() %>% pull(cellID)
tmp = so_merged[,sub]
tmp@meta.data %>% count(sample_type)
    
pdf(sprintf("%s/TRBI_MDE_LCMV_SLO_NLT_Smarta_Nguyen.pdf",out_dir),7, 7, useDingbats = F)
for (i in unique(tmp$sample_type)) {
    print(i)
    tmp = so_merged[,so_merged@meta.data %>% filter(sample_type == i) %>% rownames()]
    
    p = DimPlot(tmp, reduction = "mde_incremental", group.by = "annotation_level2", cols = mypal_level2) + ggtitle(i)
    bg = AddEmbeddingRasterBackground(lineage = "CD4")
    p$layers = c(list(bg), p$layers)
    p = p + ggplot2::coord_fixed() + xlim(c(-2.5,4)) + ylim(c(-2.5,2.5)) +theme_void() + NoLegend()
    print(p)
}
dev.off()

```

Calculate proportions, TRBI_memory_acute_ClusterProp_LCMV_NLT.pdf
```{r}
library(dplyr)
library(ggplot2)
library(forcats)
library(scales)
library(stringr)

# so_merged@meta.data %>% filter(dataset == "GSE184259_Nguyen_LCMV") %>% count(sample_name)
# sub = so_merged@meta.data %>% filter(dataset == "GSE184259_Nguyen_LCMV") %>% pull(cellID)
# tmp0 = so_merged[,sub]

cluster_col <- "annotation_level2"   # change if needed

so_merged@meta.data %>% filter(!is.na(dataset), !is.na(sample_name), !is.na(infection_phase), infection_phase %in% c("acute", "memory"), organ_simplified %in% "SLO", !(dataset %in% c("GSE184259_Nguyen_LCMV"))) %>% count(dataset, sample_name, infection_phase)

df_plot <- so_merged@meta.data %>% filter(dataset == "GSE184259_Nguyen_LCMV") %>% 
  count(sample_type, sample_name, infection_phase, organ_simplified, .data[[cluster_col]], name = "n") %>%
  group_by(sample_name) %>%
  mutate(
    sample_total = sum(n),
    pct = 100 * n / sample_total
  ) %>%
  ungroup() %>%
  rename(cluster = all_of(cluster_col))

df_plot2 <- df_plot %>%
  filter(infection_phase %in% c("acute","memory")) %>%
  mutate(
    cluster_phase = paste(cluster, infection_phase)
  )
clusters <- levels(so$annotation_level2)[grepl("^CD4", levels(so$annotation_level2))]
x_levels <- as.vector(
  rbind(
    paste(clusters, "acute"),
    paste(clusters, "memory")
  )
)
df_plot2$cluster_phase <- factor(df_plot2$cluster_phase, levels = x_levels)
df_plot2$organ_simplified = factor(df_plot2$organ_simplified, levels = c("SLO", "SILP", "SIIEL"))

##NLT D7 vs D21
df_plot2 %>% count(sample_type, sample_name)
df_plot3 = df_plot2 %>% filter(!(cluster %in% c("CD4.wM", "CD4.wN", "CD4.wZ")), organ_simplified %in% c("SILP", "SIIEL"), !(sample_type %in% c("Arm_d7_SIIEL_Bcl6KO", "Arm_d7_SILP_Bcl6KO", "Arm_d7_SLO_Bcl6KO")) )
df_plot3 %>% count(sample_name)
df_plot3$cluster = factor(df_plot3$cluster, levels = intersect(clusters, df_plot3$cluster))

p1 = ggplot(df_plot3, aes(cluster, pct, color = infection_phase, shape = sample_type)) +
    geom_jitter(
        position = position_jitterdodge(
            jitter.width = 0.1,
            dodge.width = 0.6
        ),
        size = 2
    ) +
    stat_summary(
        aes(group = infection_phase),
        fun = mean,
        geom = "crossbar",
        width = 0.5,
        position = position_dodge(width = 0.6),
        color = "black"
    ) +
    scale_color_manual(values = c("acute" = "blue", "memory" = "red")) +
    theme_bw() +
    ggtitle ("LCMV NLT D7 vs D21") +
    theme(
        axis.text.x = element_text(angle = 60, hjust = 1),
        panel.grid.major.x = element_blank()
    ) +
    labs(
        x = "Cluster",
        y = "% cells"
    )
p1 + NoGrid()


##LCMV SLO vs NLT D7
df_plot2 %>% count(sample_type, sample_name)
df_plot3 = df_plot2 %>% filter(!(cluster %in% c("CD4.wM", "CD4.wN", "CD4.wZ")), infection_phase == "acute", organ_simplified %in% c("SLO","SILP", "SIIEL"), !(sample_type %in% c("Arm_d7_SIIEL_Bcl6KO", "Arm_d7_SILP_Bcl6KO", "Arm_d7_SLO_Bcl6KO")) )
df_plot3$cluster = factor(df_plot3$cluster, levels = intersect(clusters, df_plot3$cluster))

p2 = ggplot(df_plot3, aes(cluster, pct, color = organ_simplified)) +
    geom_jitter(
        position = position_jitterdodge(
            jitter.width = 0.1,
            dodge.width = 0.6
        ),
        size = 2
    ) +
    stat_summary(
        aes(group = organ_simplified),
        fun = mean,
        geom = "crossbar",
        width = 0.5,
        position = position_dodge(width = 0.6),
        color = "black"
    ) +
    scale_color_manual(values = c("SLO" = "blue", "SILP" = "red", "SIIEL" = "orange")) +
    theme_bw() +
    ggtitle ("LCMV SLO vs NLT D7") +
    theme(
        axis.text.x = element_text(angle = 60, hjust = 1),
        panel.grid.major.x = element_blank()
    ) +
    labs(
        x = "Cluster",
        y = "% cells"
    )
p2 + NoGrid()

##LCMV SLO vs NLT D21
df_plot2 %>% count(sample_type, sample_name)
df_plot3 = df_plot2 %>% filter(!(cluster %in% c("CD4.wM", "CD4.wN", "CD4.wZ")), infection_phase == "memory", organ_simplified %in% c("SLO","SILP", "SIIEL"), !(sample_type %in% c("Arm_d7_SIIEL_Bcl6KO", "Arm_d7_SILP_Bcl6KO", "Arm_d7_SLO_Bcl6KO")) )
df_plot3$cluster = factor(df_plot3$cluster, levels = intersect(clusters, df_plot3$cluster))

p3 = ggplot(df_plot3, aes(cluster, pct, color = organ_simplified)) +
    geom_jitter(
        position = position_jitterdodge(
            jitter.width = 0.1,
            dodge.width = 0.6
        ),
        size = 2
    ) +
    stat_summary(
        aes(group = organ_simplified),
        fun = mean,
        geom = "crossbar",
        width = 0.5,
        position = position_dodge(width = 0.6),
        color = "black"
    ) +
    scale_color_manual(values = c("SLO" = "blue", "SILP" = "red", "SIIEL" = "orange")) +
    theme_bw() +
    ggtitle ("LCMV SLO vs NLT D21") +
    theme(
        axis.text.x = element_text(angle = 60, hjust = 1),
        panel.grid.major.x = element_blank()
    ) +
    labs(
        x = "Cluster",
        y = "% cells"
    )
p3 + NoGrid()

pdf(sprintf("%s/ClusterProp_LCMV_NLT_memory_acute.pdf", out_dir), 8,5, useDingbats = F)
p1 + NoGrid()
p2 + NoGrid()
p3 + NoGrid()
dev.off()

```

DGE limma - genes that are different between A,B,C vs G,H,I,Q,R,S CD4 T cells in SLO at memory timepoints, LCMV only
```{r}
library(edgeR)

pb_orig = readRDS(sprintf("%s/trbi_8studies_20260311_pseudobulk.Rds", out_dir))

pb_orig@meta.data %>% count(dataset, sample_name)
pb = pb_orig[,pb_orig@meta.data %>% filter(dataset == "GSE184259_Nguyen_LCMV", infection_phase %in% c("memory"), organ_simplified %in% c("SLO", "SILP"), annotation_level2 %in% c("CD4.G", "CD4.Q"), !(sample_name %in% c("Arm_d7_SILP_Bcl6KO_GSE184259", "Arm_d7_SLO_Bcl6KO_GSE184259")), ncells > 10) %>% rownames()] #!(annotation_level2 %in% c("CD4.wM", "CD4.wZ", "CD4.wN", "CD4.P", "CD4.T", "CD4.U", "CD4.X", "CD4.Y", "CD4.E", "CD4.F"))
dim(pb)
pb@meta.data %>% count(dataset, sample_name)

#check that there are at least 2 replicates per cluster timepoint 
pb@meta.data %>% count(infection_phase, annotation_level2)

dge = DGEList(counts = pb@assays$RNA$counts)
dge = calcNormFactors(dge)

meta = pb@meta.data
meta %>% count(dataset, sample_name, annotation_level2,  infection_agent, infection_phase, organ_simplified)
meta$organ_simplified = factor(meta$organ_simplified, levels = c("SLO", "SILP"))
meta$annotation_level2 = factor(meta$annotation_level2, levels = unique(meta$annotation_level2))
meta$infection_phase = factor(meta$infection_phase, levels = c("acute", "memory"))
meta$infection_agent = make.names(meta$infection_agent)
# meta$infection_agent = factor(meta$infection_agent, levels= c("Listeria.monocytogenes"))
meta$sample_name = factor(meta$sample_name)
meta$dataset = factor(meta$dataset)

# design <- model.matrix(
#   ~ 0 + level2_final*infection_phase + infection_agent + dataset,
#   data = meta
# )
# colnames(design) = gsub(colnames(design), pattern = "level2_final", replacement = "")
# colnames(design) = gsub(colnames(design), pattern = "dataset", replacement = "")
# colnames(design) = gsub(colnames(design), pattern = "infection_phase", replacement = "")
# colnames(design) = gsub(colnames(design), pattern = "infection_agent", replacement = "")
# colnames(design) = make.names(colnames(design))
# colnames(design) 

# design <- model.matrix(
#   ~ 0 + annotation_level2*infection_phase + organ_simplified + sample_name,
#   data = meta
# )
design <- model.matrix(
  ~ 0 + annotation_level2*organ_simplified + sample_name,
  data = meta
)
colnames(design) = gsub(colnames(design), pattern = "annotation_level2", replacement = "")
colnames(design) = gsub(colnames(design), pattern = "sample_name", replacement = "")
colnames(design) = gsub(colnames(design), pattern = "infection_phase", replacement = "")
colnames(design) = gsub(colnames(design), pattern = "organ_simplified", replacement = "")
colnames(design) = make.names(colnames(design))
colnames(design) 
limma::nonEstimable(design)

#Filter genes using the design-aware approach:
keep_genes <- filterByExpr(dge, design = design)
dge <- dge[keep_genes, , keep.lib.sizes = FALSE]
dge <- calcNormFactors(dge)

library(limma)
v = voom(dge, design, plot = TRUE)
fit = lmFit(v, design)

tt_list = list()

#Contrast for A,B,C vs G,H,IQ,R,S at memory
# clusters <- grep("^CD4", colnames(design), value = TRUE)
# contrast_formulas <- sapply(clusters, function(cl) {
#   others <- setdiff(clusters, cl)
#   paste0(
#     cl,
#     " - (",
#     paste(others, collapse = " + "),
#     ")/", length(others)
#   )
# })
# names(contrast_formulas) <- paste0(clusters, "_vs_others")

contrast_formulas = c()
contrast_formulas["CD4G_SILP_vs_SLO"] = "SILP"
contrast_formulas["CD4Q_SILP_vs_SLO"] = "CD4.Q + SILP"

contrast_matrix <- makeContrasts(
  contrasts = contrast_formulas,
  levels = design
)

fit2 <- contrasts.fit(fit, contrast_matrix)
fit2 <- eBayes(fit2)

tt_list <- lapply(colnames(contrast_matrix), function(coef_name) {
  tt <- topTable(
    fit2,
    coef = coef_name,
    number = Inf,
    adjust.method = "BH",
    sort.by = "none"
  )
  tt$SYMBOL <- rownames(tt)
  tt$contrast <- coef_name
  tt$cluster <- sub("_vs_others$", "", coef_name)
  tt
})

names(contrast_formulas)
colnames(contrast_matrix)

names(tt_list) <- names(contrast_formulas)

saveRDS(tt_list, sprintf("%s/ttlist_Nguyen_CD4Trm.Rds", out_dir))

# vplot = data.frame(fc = 2^(tt$logFC), pval = tt$adj.P.Val, SYMBOL = rownames(tt))
# ZemmourLib::Vplot(vplot, ylimits = c(10^-50, 1), xlimits = NULL) + ggtitle(label = "memory_vs_acute_all")

prefix_file_name = "Nguyen_CD4Trm"
tmp = tt_list
pdf(sprintf("%s/DiffExpression_volcano_%s.pdf", out_dir, prefix_file_name), width = 10, height = 10)
# i = "memory_vs_acute_all2"
for (i in 1:length(tmp)) {
    print(i)
   
     vplot = na.omit(data.frame(SYMBOL = tt_list[[i]]$SYMBOL, fc = 2^tt_list[[i]]$logFC, pval = tt_list[[i]]$P.Value + 10^-300, AveExpr = 2^(tt_list[[i]]$AveExpr))) #dealing with NA
    # vplot = data.frame(SYMBOL = tmp[[i]]$SYMBOL, fc = 2^tmp[[i]]$logFC, pval = tmp[[i]]$P.Value+10^-300, AveExpr = 2^(tmp[[i]]$AveExpr))
    p = Vplot(vplot = vplot, xlab = names(tmp)[i], xlimits = c(min(vplot$fc[vplot$fc != 0]), max(vplot$fc)), ylimits = c(min(vplot$pval), 1)) #c(min(vplot$pval[vplot$pval != 0]),1)
    sub = vplot %>%
        filter(abs(log2(fc)) > 0.5 & pval < 0.05) %>%
        arrange(desc(log2(fc))) %>%
        pull(SYMBOL)
    highlight_genes = unique(c(head(sub, 50), tail(sub, 50)))
    # sub = (log2(vplot$fc) > 0.5 | log2(vplot$fc) < -0.5)  & vplot$pval < 0.05
    # sub = vplot[sub,] %>% arrange(desc(log2(fc)))
    # sub1 = sub %>% pull(SYMBOL) %>% head(25) #vplot$SYMBOL %in% as.character(sub2$SYMBOL)[1:200]
    # sub2 = sub %>% pull(SYMBOL) %>% tail(25)
    highlight_filter = vplot$SYMBOL %in% highlight_genes
    # sub = vplot$SYMBOL %in% c(sub1, sub2)#as.character(sub2$SYMBOL)[1:50]
    library(ggrepel)
    print(p + geom_text_repel(data = vplot[highlight_filter,], aes(x = fc, y = pval, label = SYMBOL)))
    # print(p + geom_text(data = vplot[highlight_filter,], aes(x = fc, y = pval, label = SYMBOL)))
}
dev.off()

```

Heatmap using limma coefficients 
```{r}
tt_list_Lm = readRDS(sprintf("%s/ttlist_Listeria_SLO_memory_OneClusterVsAllOthers.Rds", out_dir))
# genes_to_label = tt_list_Lm$memory_vs_acute %>% filter( abs(logFC) > 2 & adj.P.Val < 0.05) %>% pull(SYMBOL) %>% head(50)
tt = tt_list_Lm 

comp_filt = names(tt)
comp_filt = comp_filt[grepl("_vs_others", comp_filt)]

#filter gene names
genes_to_label <- c()
for (i in comp_filt) {
  new_genes <- tt[[i]] %>%
    arrange(desc(abs(logFC))) %>%
    filter(adj.P.Val < 0.05) %>%
    pull(SYMBOL)
  new_genes <- setdiff(new_genes, genes_to_label)
  genes_to_label <- c(genes_to_label, head(new_genes, 10))
}

genes_to_label <- genes_to_label[
  !grepl(
    "^Gm|^mt-|^Rpl|^Rps|^Snor|^Mir|^Linc|^Rp[sl]|-ps|Rik$|^Hbb",
    genes_to_label,
    ignore.case = TRUE
  )
]

#plot heatmap
mat_coef <- sapply(tt, function(tt) tt$logFC)
rownames(mat_coef) <- tt[[1]]$SYMBOL
mat_plot = mat_coef[genes_to_label, comp_filt, drop = FALSE]

library(ComplexHeatmap)
library(circlize)

col_fun <- circlize::colorRamp2(
  c(-5, 0, 5),
  c("#4575B4", "white", "#D73027")
)

# pdf(sprintf("%s/heatmap_memory_vs_acute_SLO_LCMV.pdf", out_dir), width = 3, 8, useDingbats = F)
pdf(sprintf("%s/Heatmap_Listeria_SLO_memory_OneClusterVsAllOthers.pdf", out_dir), width = 3, 8, useDingbats = F)
Heatmap(
  mat_plot,
  name = "log2FC",
  col = col_fun,
  cluster_columns = FALSE,
  cluster_rows = F,
    row_names_gp = gpar(fontsize = 6),
      column_names_gp = gpar(fontsize = 6),
  column_title = "one cluster vs all others at memory timepoint"
)
dev.off()

```

**GSE184259_Nguyen: BCL6 KO**

```{r}
pdf(sprintf("%s/TRBI_MDEsample_type_GSE184259_Nguyen_Bcl6KO.pdf",out_dir),7, 7, useDingbats = F)
so_merged@meta.data %>% filter(dataset == "GSE184259_Nguyen_LCMV") %>% count(sample_type)
so_merged$cellID = colnames(so_merged)
sub = so_merged@meta.data %>% filter(dataset == "GSE184259_Nguyen_LCMV", sample_type %in% c("Arm_d7_SLO", "Arm_d7_SLO_Bcl6KO", "Arm_d7_SIIEL", "Arm_d7_SIIEL_Bcl6KO", "Arm_d7_SILP", "Arm_d7_SILP_Bcl6KO")) %>% group_by(sample_type) %>% slice_sample(n = 1870) %>% ungroup() %>% pull(cellID)
all(sub %in% colnames(so_merged))
tmp0 = so_merged[,sub]
for (i in unique(tmp0$sample_type)) {
    print(i)
    tmp = tmp0[,tmp0@meta.data %>% filter(sample_type == i) %>% rownames()]
    
    p = DimPlot(tmp, reduction = "mde_incremental", group.by = "annotation_level2", cols = mypal_level2) + ggtitle(i)
    bg = AddEmbeddingRasterBackground(lineage = "CD4")
    p$layers = c(list(bg), p$layers)
    p = p + ggplot2::coord_fixed() + xlim(c(-2.5,4)) + ylim(c(-2.5,2.5)) +theme_void() + NoLegend()
    print(p)
}
dev.off()

```

Calculate proportions, TRBI_memory_acute_ClusterProp_LCMV_NLT.pdf
```{r}
library(dplyr)
library(ggplot2)
library(forcats)
library(scales)
library(stringr)

# so_merged@meta.data %>% filter(dataset == "GSE184259_Nguyen_LCMV") %>% count(sample_name)
# sub = so_merged@meta.data %>% filter(dataset == "GSE184259_Nguyen_LCMV") %>% pull(cellID)
# tmp0 = so_merged[,sub]

cluster_col <- "annotation_level2"   # change if needed

so_merged@meta.data %>% filter(!is.na(dataset), !is.na(sample_name), !is.na(infection_phase), infection_phase %in% c("acute"), dataset %in% c("GSE184259_Nguyen_LCMV")) %>% count(dataset, sample_name, infection_phase)

df_plot <- so_merged@meta.data %>% filter(dataset == "GSE184259_Nguyen_LCMV") %>% 
  count(sample_type, sample_name, infection_phase, organ_simplified, .data[[cluster_col]], name = "n") %>%
  group_by(sample_name) %>%
  mutate(
    sample_total = sum(n),
    pct = 100 * n / sample_total
  ) %>%
  ungroup() %>%
  rename(cluster = all_of(cluster_col))

df_plot2 <- df_plot %>%
  filter(infection_phase %in% c("acute","memory")) %>%
  mutate(
    cluster_phase = paste(cluster, infection_phase)
  )
clusters <- levels(so$annotation_level2)[grepl("^CD4", levels(so$annotation_level2))]
x_levels <- as.vector(
  rbind(
    paste(clusters, "acute"),
    paste(clusters, "memory")
  )
)
df_plot2$cluster_phase <- factor(df_plot2$cluster_phase, levels = x_levels)
df_plot2$organ_simplified = factor(df_plot2$organ_simplified, levels = c("SLO", "SILP", "SIIEL"))

##LCMV SLO D7: WT vs BCL6KO
df_plot2 %>% count(sample_type, sample_name)
df_plot3 = df_plot2 %>% filter(!(cluster %in% c("CD4.wM", "CD4.wN", "CD4.wZ")), infection_phase == "acute", organ_simplified %in% c("SLO"), sample_type %in% c("Arm_d7_SLO", "Arm_d7_SLO_Bcl6KO") )
df_plot3$cluster = factor(df_plot3$cluster, levels = levels(so$annotation_level2))

p2 = ggplot(df_plot3, aes(cluster, pct, color = sample_type)) +
    geom_jitter(
        position = position_jitterdodge(
            jitter.width = 0.1,
            dodge.width = 0.6
        ),
        size = 2
    ) +
    stat_summary(
        aes(group = sample_type),
        fun = mean,
        geom = "crossbar",
        width = 0.5,
        position = position_dodge(width = 0.6),
        color = "black"
    ) +
    scale_color_manual(values = c("Arm_d7_SLO" = "blue", "Arm_d7_SLO_Bcl6KO" = "red")) +
    theme_bw() +
    ggtitle ("LCMV WT vs BCL6 KO, SLO D7") +
    theme(
        axis.text.x = element_text(angle = 60, hjust = 1),
        panel.grid.major.x = element_blank()
    ) +
    labs(
        x = "Cluster",
        y = "% cells"
    )
p2 + NoGrid()

##LCMV SI D7: WT vs BCL6KO
df_plot2 %>% count(sample_type, sample_name)
df_plot3 = df_plot2 %>% filter(!(cluster %in% c("CD4.wM", "CD4.wN", "CD4.wZ")), infection_phase == "acute", sample_type %in% c("Arm_d7_SIIEL","Arm_d7_SILP", "Arm_d7_SIIEL_Bcl6KO", "Arm_d7_SILP_Bcl6KO") )
df_plot3$cluster = factor(df_plot3$cluster, levels = levels(so$annotation_level2))
df_plot3$sample_type = factor(df_plot3$sample_type, levels = c("Arm_d7_SIIEL","Arm_d7_SILP", "Arm_d7_SIIEL_Bcl6KO", "Arm_d7_SILP_Bcl6KO") )

p3 = ggplot(df_plot3, aes(cluster, pct, color = sample_type, shape = sample_type)) +
    geom_jitter(
        position = position_jitterdodge(
            jitter.width = 0.1,
            dodge.width = 0.6
        ),
        size = 2
    ) +
    stat_summary(
        aes(group = sample_type),
        fun = mean,
        geom = "crossbar",
        width = 0.5,
        position = position_dodge(width = 0.6),
        color = "black"
    ) +
    scale_color_manual(values = c("Arm_d7_SIIEL" = "blue", "Arm_d7_SILP" = "blue", "Arm_d7_SIIEL_Bcl6KO" = "red", "Arm_d7_SILP_Bcl6KO" = "red")) +
    theme_bw() +
    ggtitle ("LCMV WT vs BCL6 KO, SI LP and IEL D7") +
    theme(
        axis.text.x = element_text(angle = 60, hjust = 1),
        panel.grid.major.x = element_blank()
    ) +
    labs(
        x = "Cluster",
        y = "% cells"
    )
p3 + NoGrid()

pdf(sprintf("%s/TRBI_memory_acute_ClusterProp_LCMV_BCL6KO.pdf", out_dir), 8,5, useDingbats = F)
p2 + NoGrid()
p3 + NoGrid()
dev.off()


```


**Analysis within ImmgenT data**

Plot MDE immgenT - SLO in LCMV arm and Listeria
```{r}
sub = so@meta.data %>% filter(infection_agent %in% c("ActA-deficient Listeria monocytogenes (DPL1942)-2W", "LCMV Armstrong"), organ_simplified %in% c("SLO", "LN", "spleen"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43")) %>% count(sample_code, sample_type, organ_simplified)
tmp = so[,so@meta.data %>% filter(infection_agent %in% c("ActA-deficient Listeria monocytogenes (DPL1942)-2W", "LCMV Armstrong"), organ_simplified %in% c("SLO", "LN", "spleen"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43")) %>% rownames()]
tmp$organ_simplified0 = "SLO"

tmp@meta.data %>% count(infection_agent, timepoint, Ag_spe)

tmp@meta.data %>% filter(infection_agent == "LCMV Armstrong" & Ag_spe == "SMARTA") %>% count(timepoint)
sub = tmp@meta.data %>% filter(infection_agent == "LCMV Armstrong" & Ag_spe == "SMARTA") %>% group_by(timepoint) %>% slice_sample(n = 121) %>% ungroup() %>% pull(cellID)
tmp2 = tmp[,sub]
pdf(sprintf("%s/MDE_LCMV_timecourse_IGT363840.pdf.pdf",out_dir2),7, 7, useDingbats = F)
for (i in unique(tmp2$timepoint)) {
    print(i)
    tmp3 = tmp2[,tmp2@meta.data %>% filter(timepoint == i) %>% rownames()]
    
    p = DimPlot(tmp3, reduction = "mde_incremental", group.by = "annotation_level2", cols = mypal_level2, pt.size = 2) + ggtitle(i)
    bg = AddEmbeddingRasterBackground(lineage = "CD4")
    p$layers = c(list(bg), p$layers)
    p = p + ggplot2::coord_fixed() + xlim(c(-2.5,4)) + ylim(c(-2.5,2.5)) +theme_void() + NoLegend()
    print(p)
}
dev.off()

tmp@meta.data %>% filter(infection_agent == "ActA-deficient Listeria monocytogenes (DPL1942)-2W") %>% count(timepoint)
sub = tmp@meta.data %>% filter(infection_agent == "ActA-deficient Listeria monocytogenes (DPL1942)-2W") %>% group_by(timepoint) %>% slice_sample(n = 500) %>% ungroup() %>% pull(cellID)
tmp2 = tmp[,sub]
pdf(sprintf("%s/MDE_Listeria_timecourse_IGT43.pdf",out_dir2),7, 7, useDingbats = F)
for (i in unique(tmp2$timepoint)) {
    print(i)
    tmp3 = tmp2[,tmp2@meta.data %>% filter(timepoint == i) %>% rownames()]
    
    p = DimPlot(tmp3, reduction = "mde_incremental", group.by = "annotation_level2", cols = mypal_level2, pt.size = 2) + ggtitle(i)
    bg = AddEmbeddingRasterBackground(lineage = "CD4")
    p$layers = c(list(bg), p$layers)
    p = p + ggplot2::coord_fixed() + xlim(c(-2.5,4)) + ylim(c(-2.5,2.5)) +theme_void() + NoLegend()
    print(p)
}
dev.off()
```

Plot MDE immgenT CD4 Trm
```{r}

so@meta.data %>% filter(condition_broad %in% c("bacteria", "virus"), !(organ_simplified %in% c("SLO", "LN", "spleen"))) %>% count(sample_code, sample_type, organ_simplified, Ag_spe)

so@meta.data %>% filter(infection_agent %in% c("LCMV Armstrong"), !(organ_simplified %in% c("SLO", "LN", "spleen")), Ag_spe == "SMARTA", IGT %in% c("IGT36", "IGT38", "IGT40")) %>% count(sample_code, sample_type, organ_simplified)
tmp = so[,so@meta.data %>% filter(infection_agent %in% c("LCMV Armstrong"), !(organ_simplified %in% c("SLO", "LN", "spleen")), Ag_spe == "SMARTA", IGT %in% c("IGT36", "IGT38", "IGT40"))  %>% rownames()]
tmp$organ_simplified0 = "NLT"

tmp@meta.data %>% count(infection_agent, timepoint, Ag_spe)

tmp@meta.data %>% filter(infection_agent == "LCMV Armstrong" & Ag_spe == "SMARTA") %>% count(timepoint)
sub = tmp@meta.data %>% filter(infection_agent == "LCMV Armstrong" & Ag_spe == "SMARTA") %>% group_by(timepoint) %>% ungroup() %>% pull(cellID) # %>% slice_sample(n = 121)
tmp2 = tmp[,sub]
pdf(sprintf("%s/MDE_LCMV_timecourse_NLT_IGT363840.pdf",out_dir2),7, 7, useDingbats = F)
for (i in unique(tmp2$timepoint)) {
    print(i)
    tmp3 = tmp2[,tmp2@meta.data %>% filter(timepoint == i) %>% rownames()]
    
    p = DimPlot(tmp3, reduction = "mde_incremental", group.by = "annotation_level2", cols = mypal_level2, pt.size = 2) + ggtitle(i)
    bg = AddEmbeddingRasterBackground(lineage = "CD4")
    p$layers = c(list(bg), p$layers)
    p = p + ggplot2::coord_fixed() + xlim(c(-2.5,4)) + ylim(c(-2.5,2.5)) +theme_void() + NoLegend()
    print(p)
}
dev.off()

```

TopicDGE LCMV
```{r}
mdata_orig = so@meta.data

L1 = readRDS(sprintf('/Users/david/google_drive_uchicago/ImmgenT/analysis/data_integration/IGT1_96/topic/flashier20250215_alldata_backfit/L1.Rds')) #cells x factor
F1 = readRDS(sprintf("/Users/david/google_drive_uchicago/ImmgenT/analysis/data_integration/IGT1_96/topic/flashier20250215_alldata_backfit/F1.Rds")) #gene x factor

mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43"), Ag_spe %in% c("SMARTA", "2WTetp"), organ_simplified %in% c("spleen", "LN", "SLO") ) %>% count(infection_phase, annotation_level2)
mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43")) %>% count(Ag_spe)

group_list = list(
    
        list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43"), Ag_spe %in% c("SMARTA"), organ_simplified %in% c("spleen", "LN", "SLO"), infection_phase == "cleared" )  %>% pull(cellID),
         group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43"), Ag_spe %in% c("SMARTA"), organ_simplified %in% c("spleen", "LN", "SLO"), infection_phase == "acute" )  %>% pull(cellID),
         title = "CD4 memory vs acute in LCMV"),

    # list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43"), Ag_spe %in% c("SMARTA"), organ_simplified %in% c("spleen", "LN", "SLO"), annotation_level2 %in% c("CD4.A","CD4.B", "CD4.C"), infection_phase == "cleared" )  %>% pull(cellID),
    #      group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43"), Ag_spe %in% c("SMARTA"), organ_simplified %in% c("spleen", "LN", "SLO"), annotation_level2 %in% c("CD4.A","CD4.B", "CD4.C"), infection_phase == "acute" )  %>% pull(cellID),
    #      title = "CD4 ABC memory vs acute in LCMV"),
    
    list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43"), Ag_spe %in% c("SMARTA"), organ_simplified %in% c("spleen", "LN", "SLO"), annotation_level2 %in% c("CD4.G","CD4.H", "CD4.I"), infection_phase == "cleared" )  %>% pull(cellID),
         group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43"), Ag_spe %in% c("SMARTA"), organ_simplified %in% c("spleen", "LN", "SLO"), annotation_level2 %in% c("CD4.G","CD4.H", "CD4.I"), infection_phase == "acute" )  %>% pull(cellID),
         title = "CD4 GHI memory vs acute in LCMV"),
    
    list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43"), Ag_spe %in% c("SMARTA"), organ_simplified %in% c("spleen", "LN", "SLO"), annotation_level2 %in% c("CD4.Q","CD4.R", "CD4.S"), infection_phase == "cleared" )  %>% pull(cellID),
         group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43"), Ag_spe %in% c("SMARTA"), organ_simplified %in% c("spleen", "LN", "SLO"), annotation_level2 %in% c("CD4.Q","CD4.R", "CD4.S"), infection_phase == "acute" )  %>% pull(cellID),
         title = "CD4 QRS memory vs acute in LCMV")

)

results_list = list()

for (i in 1:length(group_list)) {
    print(group_list[[i]]$title)
    results_list[[group_list[[i]]$title]] = FlashierDGE(F1, L1, group_list[[i]]$group1, group_list[[i]]$group2, title_plot = group_list[[i]]$title)
}


# results$p1  # First plot (MA plot of factors)
# results$p2  # Second plot (MA plot of genes)
# results$diff_factors  %>% arrange(desc(log2FC)) %>% head(10)
# results$diff_genes %>% arrange(desc(log2FC))  %>% head(50)# table for fold changes of genes

results_list_LCMV = results_list

#Heatmap with the results
tt = results_list 

comp_filt = names(tt)
# comp_filt = comp_filt[grepl("_vs_others", comp_filt)]

# #filter gene names
# genes_to_label <- c()
# for (i in comp_filt) {
#   new_genes <- tt[[i]]$diff_genes %>%
#     arrange(desc(log2FC)) %>%
#     # filter(adj.P.Val < 0.05) %>%
#     pull(SYMBOL)
#   new_genes <- setdiff(new_genes, genes_to_label)
#   genes_to_label <- c(genes_to_label, head(new_genes, 20))
# }

genes_to_label <- c()
for (i in comp_filt) {
  df <- tt[[i]]$diff_genes #%>% filter(adj.P.Val < 0.05)
  # top up
  up_genes <- df %>%
    arrange(desc(log2FC)) %>%
    pull(SYMBOL)
  # top down
  down_genes <- df %>%
    arrange(log2FC) %>%
    pull(SYMBOL)
  # remove genes already selected
  up_genes <- setdiff(up_genes, genes_to_label)
  down_genes <- setdiff(down_genes, genes_to_label)

  # add top 20 each
  genes_to_label <- c(
    genes_to_label,
    head(up_genes, 10),
    head(down_genes, 10)
  )
}

genes_to_label <- genes_to_label[
  !grepl(
    "^Gm|^mt-|^Rpl|^Rps|^Snor|^Mir|^Linc|^Rp[sl]|-ps|Rik$|^Hbb",
    genes_to_label,
    ignore.case = TRUE
  )
]

genes_to_label_LCMV = genes_to_label

#plot heatmap
# mat_coef <- sapply(tt, function(tt) tt$logFC)
mat_coef <- Reduce(cbind, lapply(tt, function(x) x$diff_genes$log2FC))
rownames(mat_coef) <- tt[[1]]$diff_genes$SYMBOL
colnames(mat_coef) = make.names(names(tt))
mat_coef_LCMV = mat_coef
mat_plot = mat_coef[genes_to_label, ]
range(mat_plot)

library(ComplexHeatmap)
library(circlize)

col_fun <- circlize::colorRamp2(
  c(-1, 0, 1),
  c("#4575B4", "white", "#D73027")
)

# pdf(sprintf("%s/heatmap_memory_vs_acute_SLO_LCMV.pdf", out_dir), width = 3, 8, useDingbats = F)
pdf(sprintf("%s/Heatmap_LCMV_SLO_memory_vs_acute_IGT363840.pdf", out_dir2), width = 3, 8, useDingbats = F)
Heatmap(
  mat_plot,
  name = "log2FC",
  col = col_fun,
  cluster_columns = FALSE,
  cluster_rows = T,
    row_names_gp = gpar(fontsize = 6),
      column_names_gp = gpar(fontsize = 6),
  column_title = "memory vs acute across clusters in LCMV SLO"
)
dev.off()




```

TopicDGE Listeria
```{r}
# mdata_orig = so@meta.data
# L1 = readRDS(sprintf('/Users/david/google_drive_uchicago/ImmgenT/analysis/data_integration/IGT1_96/topic/flashier20250215_alldata_backfit/L1.Rds')) #cells x factor
# F1 = readRDS(sprintf("/Users/david/google_drive_uchicago/ImmgenT/analysis/data_integration/IGT1_96/topic/flashier20250215_alldata_backfit/F1.Rds")) #gene x factor

mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43"), Ag_spe %in% c("2WTetp"), organ_simplified %in% c("spleen", "LN", "SLO") ) %>% count(infection_phase, annotation_level2)
mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43")) %>% count(Ag_spe)

group_list = list(
    
        list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43"), Ag_spe %in% c("2WTetp"), organ_simplified %in% c("spleen", "LN", "SLO"), infection_phase == "cleared" )  %>% pull(cellID),
         group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43"), Ag_spe %in% c("2WTetp"), organ_simplified %in% c("spleen", "LN", "SLO"), infection_phase == "acute" )  %>% pull(cellID),
         title = "CD4 memory vs acute in Listeria"),

    # list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43"), Ag_spe %in% c("2WTetp"), organ_simplified %in% c("spleen", "LN", "SLO"), annotation_level2 %in% c("CD4.A","CD4.B", "CD4.C"), infection_phase == "cleared" )  %>% pull(cellID),
    #      group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43"), Ag_spe %in% c("2WTetp"), organ_simplified %in% c("spleen", "LN", "SLO"), annotation_level2 %in% c("CD4.A","CD4.B", "CD4.C"), infection_phase == "acute" )  %>% pull(cellID),
    #      title = "CD4 ABC memory vs acute in Listeria"),
    
    list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43"), Ag_spe %in% c("2WTetp"), organ_simplified %in% c("spleen", "LN", "SLO"), annotation_level2 %in% c("CD4.G","CD4.H", "CD4.I"), infection_phase == "cleared" )  %>% pull(cellID),
         group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43"), Ag_spe %in% c("2WTetp"), organ_simplified %in% c("spleen", "LN", "SLO"), annotation_level2 %in% c("CD4.G","CD4.H", "CD4.I"), infection_phase == "acute" )  %>% pull(cellID),
         title = "CD4 GHI memory vs acute in Listeria"),
    
    list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43"), Ag_spe %in% c("2WTetp"), organ_simplified %in% c("spleen", "LN", "SLO"), annotation_level2 %in% c("CD4.Q","CD4.R", "CD4.S"), infection_phase == "cleared" )  %>% pull(cellID),
         group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43"), Ag_spe %in% c("2WTetp"), organ_simplified %in% c("spleen", "LN", "SLO"), annotation_level2 %in% c("CD4.Q","CD4.R", "CD4.S"), infection_phase == "acute" )  %>% pull(cellID),
         title = "CD4 QRS memory vs acute in Listeria")
)

results_list = list()

for (i in 1:length(group_list)) {
    print(group_list[[i]]$title)
    results_list[[group_list[[i]]$title]] = FlashierDGE(F1, L1, group_list[[i]]$group1, group_list[[i]]$group2, title_plot = group_list[[i]]$title)
}


# results$p1  # First plot (MA plot of factors)
# results$p2  # Second plot (MA plot of genes)
# results$diff_factors  %>% arrange(desc(log2FC)) %>% head(10)
# results$diff_genes %>% arrange(desc(log2FC))  %>% head(50)# table for fold changes of genes

results_list_Lm = results_list

#Heatmap with the results
tt = results_list 

comp_filt = names(tt)
# comp_filt = comp_filt[grepl("_vs_others", comp_filt)]

#filter gene names
# genes_to_label <- c()
# for (i in comp_filt) {
#   new_genes <- tt[[i]]$diff_genes %>%
#     arrange(desc(abs(log2FC))) %>%
#     # filter(adj.P.Val < 0.05) %>%
#     pull(SYMBOL)
#   new_genes <- setdiff(new_genes, genes_to_label)
#   genes_to_label <- c(genes_to_label, head(new_genes, 20))
# }

genes_to_label <- c()
for (i in comp_filt) {
  df <- tt[[i]]$diff_genes #%>% filter(adj.P.Val < 0.05)
  # top up
  up_genes <- df %>%
    arrange(desc(log2FC)) %>%
    pull(SYMBOL)
  # top down
  down_genes <- df %>%
    arrange(log2FC) %>%
    pull(SYMBOL)
  # remove genes already selected
  up_genes <- setdiff(up_genes, genes_to_label)
  down_genes <- setdiff(down_genes, genes_to_label)

  # add top 20 each
  genes_to_label <- c(
    genes_to_label,
    head(up_genes, 10),
    head(down_genes, 10)
  )
}

genes_to_label <- genes_to_label[
  !grepl(
    "^Gm|^mt-|^Rpl|^Rps|^Snor|^Mir|^Linc|^Rp[sl]|-ps|Rik$|^Hbb",
    genes_to_label,
    ignore.case = TRUE
  )
]

genes_to_label_Lm = genes_to_label

#plot heatmap
# mat_coef <- sapply(tt, function(tt) tt$logFC)
mat_coef <- Reduce(cbind, lapply(tt, function(x) x$diff_genes$log2FC))
rownames(mat_coef) <- tt[[1]]$diff_genes$SYMBOL
colnames(mat_coef) = make.names(names(tt))
mat_coef_Lm = mat_coef
mat_plot = mat_coef[genes_to_label, ]
range(mat_plot)

library(ComplexHeatmap)
library(circlize)

col_fun <- circlize::colorRamp2(
  c(-0.5, 0, 0.5),
  c("#4575B4", "white", "#D73027")
)

# pdf(sprintf("%s/heatmap_memory_vs_acute_SLO_LCMV.pdf", out_dir), width = 3, 8, useDingbats = F)
pdf(sprintf("%s/Heatmap_Listeria_SLO_memory_vs_acute_IGT363840.pdf", out_dir2), width = 3, 8, useDingbats = F)
Heatmap(
  mat_plot,
  name = "log2FC",
  col = col_fun,
  cluster_columns = FALSE,
  cluster_rows = T,
    row_names_gp = gpar(fontsize = 6),
      column_names_gp = gpar(fontsize = 6),
  column_title = "memory vs acute across clusters in Listeria SLO"
)
dev.off()

```

Heatmap with LCMV and Listeria together
```{r}
tt = c(results_list_Lm, results_list_LCMV)
genes_to_label = unique(genes_to_label_LCMV, genes_to_label_Lm)

all(rownames(mat_coef_LCMV) == rownames(mat_coef_Lm))
mat_coef =  cbind(mat_coef_LCMV, mat_coef_Lm)
mat_plot = mat_coef[genes_to_label, ]
mat_plot_LCMV = mat_coef_LCMV[genes_to_label, ]
mat_plot_Lm = mat_coef_Lm[genes_to_label, ]

range(mat_plot)

library(ComplexHeatmap)
library(circlize)

pdf(sprintf("%s/Heatmap_ListeriaLCMV_SLO_memory_vs_acute_IGT36384043.pdf", out_dir2), width = 3, 8, useDingbats = F)
col_fun <- circlize::colorRamp2(
  c(-1, 0, 1),
  c("#4575B4", "white", "#D73027")
)
Heatmap(
  mat_plot_LCMV,
  name = "log2FC",
  col = col_fun,
  cluster_columns = FALSE,
  cluster_rows = F,
    row_names_gp = gpar(fontsize = 6),
      column_names_gp = gpar(fontsize = 6),
  column_title = "memory vs acute across clusters in Listeria SLO"
)

col_fun <- circlize::colorRamp2(
  c(-0.5, 0, 0.5),
  c("#4575B4", "white", "#D73027")
)
Heatmap(
  mat_plot_Lm,
  name = "log2FC",
  col = col_fun,
  cluster_columns = FALSE,
  cluster_rows = F,
    row_names_gp = gpar(fontsize = 6),
      column_names_gp = gpar(fontsize = 6),
  column_title = "memory vs acute across clusters in Listeria SLO"
)
dev.off()

```

TopicDGE LCMV NLT vs SLO
```{r}
mdata_orig = so@meta.data

L1 = readRDS(sprintf('/Users/david/google_drive_uchicago/ImmgenT/analysis/data_integration/IGT1_96/topic/flashier20250215_alldata_backfit/L1.Rds')) #cells x factor
F1 = readRDS(sprintf("/Users/david/google_drive_uchicago/ImmgenT/analysis/data_integration/IGT1_96/topic/flashier20250215_alldata_backfit/F1.Rds")) #gene x factor

mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43"), Ag_spe %in% c("SMARTA"), organ_simplified %in% c("spleen", "LN", "SLO") ) %>% count(infection_phase, organ_simplified)
mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43"), Ag_spe %in% c("SMARTA"), !(organ_simplified %in% c("spleen", "LN", "SLO")) ) %>% count(infection_phase, organ_simplified)
mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43")) %>% count(Ag_spe)

group_list = list(
    
    list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43"), Ag_spe %in% c("SMARTA"), organ_simplified %in% c("spleen", "LN", "SLO"), annotation_level2 %in% c("CD4.G","CD4.Q"), infection_phase == "cleared" ) %>% pull(cellID),
         group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43"), Ag_spe %in% c("SMARTA"), !(organ_simplified %in% c("spleen", "LN", "SLO")), annotation_level2 %in% c("CD4.G","CD4.Q"), infection_phase == "cleared" )  %>% pull(cellID),
         title = "SLO vs NLT in GQ, in LCMV at memory timepoint"),
    
    list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43"), Ag_spe %in% c("SMARTA"), organ_simplified %in% c("spleen", "LN", "SLO"), annotation_level2 %in% c("CD4.G"), infection_phase == "cleared" ) %>% pull(cellID),
         group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43"), Ag_spe %in% c("SMARTA"), !(organ_simplified %in% c("spleen", "LN", "SLO")), annotation_level2 %in% c("CD4.G"), infection_phase == "cleared" )  %>% pull(cellID),
         title = "SLO vs NLT in G, in LCMV at memory timepoint"),
    
    list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43"), Ag_spe %in% c("SMARTA"), organ_simplified %in% c("spleen", "LN", "SLO"), annotation_level2 %in% c("CD4.Q"), infection_phase == "cleared" ) %>% pull(cellID),
         group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43"), Ag_spe %in% c("SMARTA"), !(organ_simplified %in% c("spleen", "LN", "SLO")), annotation_level2 %in% c("CD4.Q"), infection_phase == "cleared" )  %>% pull(cellID),
         title = "SLO vs NLT in Q, in LCMV at memory timepoint")
    
)

results_list = list()

for (i in 1:length(group_list)) {
    print(group_list[[i]]$title)
    results_list[[group_list[[i]]$title]] = FlashierDGE(F1, L1, group_list[[i]]$group1, group_list[[i]]$group2, title_plot = group_list[[i]]$title)
}

# results = results_list[[3]]
# results$p1  # First plot (MA plot of factors)
# results$p2  # Second plot (MA plot of genes)
# results$diff_factors  %>% arrange(desc(log2FC)) %>% head(10)
# results$diff_genes %>% arrange(desc(log2FC))  %>% head(50)# table for fold changes of genes

pdf(sprintf("%s/Volcano_LCMV_SLO_vs_NLT_IGT363840.pdf", out_dir2), width = 5, 5, useDingbats = F)
results = results_list[["SLO vs NLT in GQ, in LCMV at memory timepoint"]]
results$p1  # First plot (MA plot of factors)
results$p2  # Second plot (MA plot of genes)
# results$diff_factors  %>% arrange(desc(log2FC)) %>% head(10)
# results$diff_genes %>% arrange(desc(log2FC))  %>% head(50)# table for fold changes of genes
dev.off()

#Heatmap with the results
tt = results_list 

comp_filt = names(tt)
# comp_filt = comp_filt[grepl("_vs_others", comp_filt)]

# #filter gene names
# genes_to_label <- c()
# for (i in comp_filt) {
#   new_genes <- tt[[i]]$diff_genes %>%
#     arrange(desc(log2FC)) %>%
#     # filter(adj.P.Val < 0.05) %>%
#     pull(SYMBOL)
#   new_genes <- setdiff(new_genes, genes_to_label)
#   genes_to_label <- c(genes_to_label, head(new_genes, 20))
# }

genes_to_label <- c()
for (i in comp_filt) {
  df <- tt[[i]]$diff_genes #%>% filter(adj.P.Val < 0.05)
  # top up
  up_genes <- df %>%
    arrange(desc(log2FC)) %>%
    pull(SYMBOL)
  # top down
  down_genes <- df %>%
    arrange(log2FC) %>%
    pull(SYMBOL)
  # remove genes already selected
  up_genes <- setdiff(up_genes, genes_to_label)
  down_genes <- setdiff(down_genes, genes_to_label)

  # add top 20 each
  genes_to_label <- c(
    genes_to_label,
    head(up_genes, 10),
    head(down_genes, 10)
  )
}

genes_to_label <- genes_to_label[
  !grepl(
    "^Gm|^mt-|^Rpl|^Rps|^Snor|^Mir|^Linc|^Rp[sl]|-ps|Rik$|^Hbb",
    genes_to_label,
    ignore.case = TRUE
  )
]


#plot heatmap
# mat_coef <- sapply(tt, function(tt) tt$logFC)
mat_coef <- Reduce(cbind, lapply(tt, function(x) x$diff_genes$log2FC))
rownames(mat_coef) <- tt[[1]]$diff_genes$SYMBOL
colnames(mat_coef) = make.names(names(tt))
mat_coef_LCMV = mat_coef
mat_plot = mat_coef[genes_to_label, ]
range(mat_plot)

library(ComplexHeatmap)
library(circlize)

col_fun <- circlize::colorRamp2(
  c(-1, 0, 1),
  c("#4575B4", "white", "#D73027")
)

# pdf(sprintf("%s/heatmap_memory_vs_acute_SLO_LCMV.pdf", out_dir), width = 3, 8, useDingbats = F)
pdf(sprintf("%s/Heatmap_LCMV_SLO_vs_NLT_IGT363840.pdf", out_dir2), width = 3, 8, useDingbats = F)
Heatmap(
  mat_plot,
  name = "log2FC",
  col = col_fun,
  cluster_columns = FALSE,
  cluster_rows = T,
    row_names_gp = gpar(fontsize = 6),
      column_names_gp = gpar(fontsize = 6),
  column_title = "SLO vs NLT in GQ, in LCMV"
)
dev.off()


```



```{r}
vplot = data.frame(fc1 = 2^results_list_Lm$`CD4 memory vs acute in Listeria`$diff_genes$log2FC, fc2 = 2^results_list_LCMV$`CD4 memory vs acute in LCMV`$diff_genes$log2FC, SYMBOL = results_list_LCMV$`CD4 memory vs acute in LCMV`$diff_genes$SYMBOL)
FCFCplot(vplot, xlimits = c(1/1.1, 1.1), ylimits = c(1/2, 2), genes_to_label = c(genes_to_label_Lm, genes_to_label_LCMV))

```








