---
title: "CD4_misc"
output: html_document
date: "2025-05-28"
author: "David Zemmour"
editor_options: 
  chunk_output_type: console
---

**Initialize: load libraries and custom functions**

```{bash}
sinteractive --exclusive --time=12:00:00 --account=pi-zemmour --partition=jfkfloor2 --qos=jfkfloor2 --nodes=1 
sinteractive --exclusive --time=12:00:00 --account=pi-zemmour --partition=beagle3-bigmem --nodes=1 
sinteractive --exclusive --time=12:00:00 --account=pi-zemmour --partition=zemmour-hm --qos=zemmour --nodes=1 
sinteractive --time=03:00:00 --account=pi-zemmour --partition=caslake --nodes=1 --mem=96GB

module load python/anaconda-2022.05 
module load hdf5/1.12.0 #for BPCells but also others
module load openblas/0.3.13
module load ImageMagick/7.1.1-7
source activate /project/zemmour/david/envs/scvi_20240315
#source activate /project/zemmour/david/envs/r443
module load openblas/0.3.13

cd /project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/CD4
R
```


```{r}
setwd("~/google_drive_uchicago/ImmgenT/analysis/data_integration/IGT1_96/CD4")
libs = c("ZemmourLib","forcats", "gridExtra","patchwork", "Matrix", "Seurat", "ggplot2", "dplyr", "reshape2", "ggrastr","scattermore", "RColorBrewer", "pals", "scales", "pheatmap", "rafalib") 
sapply(libs, function(x) suppressMessages(suppressWarnings(library(x, character.only = TRUE, quietly = T, warn.conflicts  = F))))
#source("/project/jfkfloor2/zemmourlab/david/immgent/immgent_integration_git/custom_functions_david.R")

ColorRamp = rev(colorRampPalette(brewer.pal(n = 7,name = "RdYlBu"))(100))
ColorRamp = rev(colorRampPalette(c('red','white','blue'))(20))
ColorRamp = rev(rainbow(10, end = 4/6))
ColorRamp = rev(viridis(100))
ColorRamp = brewer.pal(9, "Blues")

#mycol = c("black","blue","red","darkgreen", "orange","purple", "cyan", "greenyellow",    "salmon", "magenta","pink", "tan", "brown") # c("magenta", "red",  "darkgreen", "cyan", "blue", "blue", "blue", "black", "blue", "blue", "blue", "blue", "orange")

mypal_level2 = ZemmourLib::immgent_colors$level2
mypal_level2["not classified"] = "black"
# names(mypal_level2) = gsub("nonconv", replacement = "Tz",names(mypal_level2) )
# mypal_level2 = immgent_colors$level2_option2
# mypal_level2["not classified"] = "black"
mypal_level1 = immgent_colors$level1
mypal_level1["not classified"] = "black"
# names(mypal_level1) = gsub("nonconv", replacement = "Tz",names(mypal_level1) )
mypal_organ = immgent_colors$organ_simplified
mypal_level2group = c("resting" = "blue", "activated" = "red", "miniverse" = "darkgreen", "proliferating" = "black", "other" = "grey", "preT" = "grey" )

#Large Color palette
library("pals")
n = 70
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
mypal1 = unique(unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals))))
mypal1 = mypal1[-4]
mypal = c(glasbey(), polychrome(), mypal1)
names(mypal) = NULL
#pal.bands(mypal, main="Colormap suggestions")
# mypal_organ = setNames(mypal, unique(so$organ_simplified))
# mypal_organ = mypal_organ[!is.na(names(mypal_organ))]
# mypal_level1 = setNames(mypal, unique(so$annotation_level1))
# mypal_level1 = mypal_level1[!is.na(names(mypal_level1))]
# mypal_level2 = setNames(mypal, unique(so_orig$annotation_level2))
# mypal_level2 = mypal_level2[!is.na(names(mypal_level2))]
# mypal_level2["CD4_cl18"] = "blue"
mypal_igt = setNames(mypal, unique(so_orig$IGT))
mypal_igt = mypal_igt[!is.na(names(mypal_igt))]

```

```{r}
# fig_dir = "~/google_drive_uchicago/ImmgenT/20250109_freeze/paper_cosmology/figures/bits"
# fig_dir = "/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/paper_cosmology/figures/bits"

out_dir = "~/google_drive_uchicago/ImmgenT/analysis/data_integration/IGT1_96/CD4/misc_plots" 
out_dir = "/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/CD4/misc_plots"

```


**Most up-to-date data**
on Midway
```{r}
so_orig = readRDS("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/igt1_96_withtotalvi20260206_clean.Rds") 

so = so_orig[,so_orig@meta.data %>% filter(organ_simplified != "thymus") %>% row.names()]
so = NormalizeData(so, assay = "RNA", normalization.method = "LogNormalize")
```

on Desktop
```{r}
setwd("~/google_drive_uchicago/ImmgenT/analysis/data_integration/IGT1_96/")
so_orig = readRDS("igt1_96_withtotalvi20260206_clean_ADTonly.Rds")

so = so_orig[,so_orig@meta.data %>% filter(annotation_level1 == "CD4" & organ_simplified != "thymus") %>% row.names()]
# so = NormalizeData(so, assay = "RNA", normalization.method = "LogNormalize")
# so = NormalizeData(so, assay = "ADT", normalization.method = "LogNormalize")

```

**Stats for the paper**

```{r}
so@meta.data %>% filter(annotation_level2_group %in% c("activated")) %>% group_by(annotation_level2) %>% summarize(n = n()) %>% mutate(percent = n/sum(n)*100) 

so@meta.data %>% count(condition_broad)
so@meta.data %>% filter(organ_simplified %in% c("spleen", "SLO", "LN")) %>% count()


```


**Figure 4 FCFC plots**
Tfh/Tex
```{r}
tt_list = readRDS('/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/analysis/data_integration/IGT1_96/CD4/DGE_limma/20260206/ttlist_OneVsAll.Rds')
tt_list = readRDS('/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/analysis/data_integration/IGT1_96/CD4/DGE_limma/20260206/ttlist_in_Activated.Rds')

names(tt_list)


# Create the vplot dataframe with fold changes
# vplot = data.frame(fc1 = exp(tt_list$CD4.H_vs_All$logFC), fc2 = exp(tt_list$CD4.I_vs_All$logFC), SYMBOL = tt_list$CD4.H_vs_All$SYMBOL)
vplot = data.frame(fc1 = exp(tt_list$CD4.I_vs_AllActivated$logFC), fc2 = exp(tt_list$CD4.H_vs_AllActivated$logFC), SYMBOL = tt_list$CD4.H_vs_AllActivated$SYMBOL)

# Identify genes with large fold changes (>2)
# genes_to_plot = abs(tt_list$CD4.H_vs_All$logFC) > 0.69 | abs(tt_list$CD4.I_vs_All$logFC) > 0.69
genes_to_plot = abs(tt_list$CD4.H_vs_AllActivated$logFC) > 0.5 | abs(tt_list$CD4.I_vs_AllActivated$logFC) > 0.5
# genes_to_label = vplot[genes_to_plot, ]
genes_to_label = vplot$SYMBOL[genes_to_plot]

# Create the plot
pdf(sprintf("%s/FCFCplot_H_vs_I_.pdf", out_dir), 6,6, useDingbats = F)
FCFCplot(vplot, 
         xlab = "FC for CD4.I", 
         ylab = "FC for CD4.H", 
         main = "Fold Change vs Fold Change", 
         genes_to_label = NULL)
dev.off()
pdf(sprintf("%s/FCFCplot_H_vs_I_labels.pdf", out_dir), 12,12, useDingbats = F)
FCFCplot(vplot, 
         xlab = "FC for CD4.I", 
         ylab = "FC for CD4.H", 
         main = "Fold Change vs Fold Change", 
         genes_to_label = genes_to_label, 
            )
dev.off()

```


Top10 in H or I
```{r}
prop = readRDS('/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/analysis/data_integration/IGT1_96/SampleComp/20260206/annotation_level2_PropPerSamplePerLevel1_withMetadata.Rds')

library(dplyr)
library(tidyr)
library(ggplot2)

# your starting point
prop_sub <- prop %>% 
  select(
    sample_type, sample_code,
    CD4.H.ncells, CD4.H.prop,
    CD4.I.ncells, CD4.I.prop
  )

# ---- compute averages ---- 
prop_avg <- prop_sub %>% 
  group_by(sample_type) %>%
  summarise(
    CD4.H.prop = mean(CD4.H.prop, na.rm = TRUE),
    CD4.I.prop = mean(CD4.I.prop, na.rm = TRUE),
    CD4.H.ncells = mean(CD4.H.ncells, na.rm = TRUE),
    CD4.I.ncells = mean(CD4.I.ncells, na.rm = TRUE),
    n_samples = n(),
    .groups = "drop"
  ) %>%
  filter(n_samples > 2, (CD4.H.ncells > 20 | CD4.I.ncells > 20)) %>%
  arrange(desc(CD4.H.prop)) %>%
  slice_head(n = 20)

# keep only top sample_types
top_types <- prop_avg$sample_type

# ---- reshape mean ----
mean_long <- prop_avg %>%
  select(sample_type, CD4.H.prop, CD4.I.prop) %>%
  pivot_longer(
    cols = c(CD4.H.prop, CD4.I.prop),
    names_to = "cluster",
    values_to = "prop"
  ) %>%
  mutate(
    cluster = recode(cluster,
                     "CD4.H.prop" = "CD4.H",
                     "CD4.I.prop" = "CD4.I"),
    prop_plot = ifelse(cluster == "CD4.H", -prop, prop)
  )

# ---- reshape individual samples ----
points_long <- prop_sub %>%
  filter(sample_type %in% top_types) %>%
  pivot_longer(
    cols = c(CD4.H.prop, CD4.I.prop),
    names_to = "cluster",
    values_to = "prop"
  ) %>%
  mutate(
    cluster = recode(cluster,
                     "CD4.H.prop" = "CD4.H",
                     "CD4.I.prop" = "CD4.I"),
    prop_plot = ifelse(cluster == "CD4.H", -prop, prop)
  )

# keep order
mean_long$sample_type <- factor(mean_long$sample_type, levels = rev(prop_avg$sample_type))
points_long$sample_type <- factor(points_long$sample_type, levels = rev(prop_avg$sample_type))

# ---- plot ----
p = ggplot() +
  geom_col(
    data = mean_long,
    aes(x = sample_type, y = prop_plot, fill = cluster),
    width = 0.7,
    alpha = 0.8
  ) +
  geom_point(
    data = points_long,
    aes(x = sample_type, y = prop_plot), color = "black",
    position = position_jitter(width = 0.15),
    size = 2,
    alpha = 1
  ) +
  geom_hline(yintercept = 0) +
  coord_flip() +
  scale_y_continuous(labels = abs) +
  scale_fill_manual(values = mypal_level2) +
  # scale_color_manual(values = mypal_level2) +
  labs(
    x = NULL,
    y = "Proportion (H left, I right)"
  ) +
  theme_classic(base_size = 14) +
  theme(
    axis.text.y = element_text(size = 12),
    legend.position = "top"
  )

pdf(sprintf("%s/Top10_samples_HI.pdf", out_dir), 10,5, useDingbats = F)
print(p)
dev.off()


```



**MDE plots**
```{r}
pdf("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/CD4/MDE_clusters_20250527.pdf", width = 10, height = 10, useDingbats = F)
p = DimPlot(so, reduction = "mde_incremental", group.by = "annotation_level2", split.by = NULL, raster = T, raster.dpi = c(512,512), label = T, pt.size = 1) + scale_color_manual(values = mypal_level2)
p
p + NoLegend()

p = DimPlot(so, reduction = "mde_incremental", group.by = "annotation_level2", split.by = NULL, raster = T, raster.dpi = c(512,512), label = F, pt.size = 1) + scale_color_manual(values = mypal_level2)
p + NoLegend()

p = MyDimPlotHighlight(seurat_object = so, umap_to_plot = "mde_incremental", cells_to_highlight = colnames(so)[which(so$annotation_level2 %in% c("CD4_cl1","CD4_cl2","CD4_cl3", "CD4_cl4", "CD4_cl5", "CD4_cl7", "CD4_cl12"))], highlight_column_name = "annotation_level2", pixels = c(512, 512), mycols = mypal_level2, title = "", highlight_size = 0.1, highlight_alpha = 0.1, print_plot1 = T, print_plot2 = T, labelclusters = T)
p

p = MyDimPlotHighlight(seurat_object = so, umap_to_plot = "mde_incremental", cells_to_highlight = colnames(so)[which(so$annotation_level2 %in% c("CD4_cl2"))], highlight_column_name = "annotation_level2", pixels = c(512, 512), mycols = mypal_level2, title = "", highlight_size = 0.1, highlight_alpha = 0.1, print_plot1 = T, print_plot2 = T, labelclusters = T)
p

p = MyDimPlotHighlight(seurat_object = so, umap_to_plot = "mde_incremental", cells_to_highlight = colnames(so)[which(so$annotation_level2 %in% c("CD4_cl5", "CD4_cl7"))], highlight_column_name = "annotation_level2", pixels = c(512, 512), mycols = mypal_level2, title = "", highlight_size = 0.1, highlight_alpha = 0.1, print_plot1 = T, print_plot2 = T, labelclusters = T)
p

p = MyDimPlotHighlight(seurat_object = so, umap_to_plot = "mde_incremental", cells_to_highlight = colnames(so)[which(so$annotation_level2 %in% c("CD4_cl9", "CD4_cl10", "CD4_cl13", "CD4_cl14"))], highlight_column_name = "annotation_level2", pixels = c(512, 512), mycols = mypal_level2, title = "", highlight_size = 0.1, highlight_alpha = 0.1, print_plot1 = T, print_plot2 = T, labelclusters = T)
p

p = MyDimPlotHighlight(seurat_object = so, umap_to_plot = "mde_incremental", cells_to_highlight = colnames(so)[which(so$annotation_level2 %in% c("CD4_cl9","CD4_cl21","CD4_cl10","CD4_cl13", "CD4_cl14", "CD4_cl15", "CD4_cl16", "CD4_cl17", "CD4_cl18", "CD4_cl22", "CD4_cl23"))], highlight_column_name = "annotation_level2", pixels = c(512, 512), mycols = mypal_level2, title = "", highlight_size = 0.1, highlight_alpha = 0.1, print_plot1 = T, print_plot2 = T, labelclusters = T)
p

p = MyDimPlotHighlight(seurat_object = so, umap_to_plot = "mde_incremental", cells_to_highlight = colnames(so)[which(so$annotation_level2 %in% c("CD4_cl15", "CD4_cl16", "CD4_cl17", "CD4_cl18"))], highlight_column_name = "annotation_level2", pixels = c(512, 512), mycols = mypal_level2, title = "", highlight_size = 0.1, highlight_alpha = 0.1, print_plot1 = T, print_plot2 = T, labelclusters = T)
p

p = MyDimPlotHighlight(seurat_object = so, umap_to_plot = "mde_incremental", cells_to_highlight = colnames(so)[which(so$annotation_level2 %in% c("CD4_cl21", "CD4_cl22", "CD4_cl23"))], highlight_column_name = "annotation_level2", pixels = c(512, 512), mycols = mypal_level2, title = "", highlight_size = 0.1, highlight_alpha = 0.1, print_plot1 = T, print_plot2 = T, labelclusters = T)
p


p = MyDimPlotHighlight(seurat_object = so, umap_to_plot = "mde_incremental", cells_to_highlight = colnames(so)[which(so$annotation_level2 %in% c("CD4_cl9","CD4_cl21"))], highlight_column_name = "annotation_level2", pixels = c(512, 512), mycols = mypal_level2, title = "", highlight_size = 0.1, highlight_alpha = 0.1, print_plot1 = T, print_plot2 = T, labelclusters = T)
p

p = MyDimPlotHighlight(seurat_object = so, umap_to_plot = "mde_incremental", cells_to_highlight = colnames(so)[which(so$annotation_level2 %in% c("CD4_cl10","CD4_cl13", "CD4_cl14", "CD4_cl15", "CD4_cl16", "CD4_cl17", "CD4_cl18", "CD4_cl22", "CD4_cl23"))], highlight_column_name = "annotation_level2", pixels = c(512, 512), mycols = mypal_level2, title = "", highlight_size = 0.1, highlight_alpha = 0.1, print_plot1 = T, print_plot2 = T, labelclusters = T)
p
dev.off()
```

**Feature plots**
```{r}
so = NormalizeData(so,assay = "RNA", normalization.method = "LogNormalize",verbose = T)

FeaturePlot(so, reduction = "mde_incremental", slot = "data", features = "Tox", order = T, raster = T, raster.dpi = c(1024,1024))
FeaturePlot(so, reduction = "mde_incremental", slot = "data", features = "Gata3", order = T, raster = T, raster.dpi = c(1024,1024))
FeaturePlot(so, reduction = "mde_incremental", slot = "data", features = "Tbx21", order = T, raster = T, raster.dpi = c(1024,1024))
FeaturePlot(so, reduction = "mde_incremental", slot = "data", features = "Rorc", order = T, raster = T, raster.dpi = c(1024,1024))
FeaturePlot(so, reduction = "mde_incremental", slot = "data", features = "Tox2", order = T, raster = T, raster.dpi = c(1024,1024))
```

**Downsample seurat object**
```{r}
sample_cells = so@meta.data %>%
  group_by(!!sym("IGTHT")) %>%
  sample_n(size = min(500, n()), replace = FALSE) %>% 
    pull(IGT_cellID)

so = so[,sample_cells]
saveRDS(so, "/Users/david/Desktop/Treg_App/igt1_96_CD4_20250212_downsampled.Rds")
```

**Figure Tip/BAS**

2a
```{r}
pdf(sprintf("%s/MDE_level2.pdf", out_dir), 5,5, useDingbats = F)
DimPlot(so, reduction = "mde_incremental", group.by = "annotation_level2", cols = mypal_level2, raster = T, raster.dpi = c(1024,1024), pt.size = 1, alpha = 1, label = T, label.size = 2, label.box = T) + NoGrid()  + NoLegend()
p = DimPlot(so, reduction = "mde_incremental", group.by = "annotation_level2", cols = mypal_level2, raster = T, raster.dpi = c(1024,1024), pt.size = 1, alpha = 1) + xlim(c(-2.5,4)) + ylim(c(-2.5,2.5))
p
p + theme_void() + NoLegend() + NoGrid()
dev.off()

MyFeatureScatter(so,feature1 = "CD62L", feature2 = "CD44")

```

**Figure 2**
tips percent
```{r}
so@meta.data %>% count(annotation_level2, annotation_level2_group)
mdata = so@meta.data %>% filter(annotation_level2_group == "activated", !(annotation_level2 %in% c("CD4.wN", "CD4.wZ"))) 

mdata %>% count(mdata)
mdata$tip_mid = "tip"
mdata$tip_mid[mdata$annotation_level2 %in% c("CD4.G", "CD4.Q")] = "midland"
mdata$tip_mid[mdata$annotation_level2 %in% c("CD4.R", "CD4.S")] = "th1_tip"
mdata$tip_mid[mdata$annotation_level2 %in% c("CD4.T", "CD4.U")] = "th2_tip"
mdata$tip_mid[mdata$annotation_level2 %in% c("CD4.X", "CD4.Y")] = "th17_tip"
mdata$tip_mid[mdata$annotation_level2 %in% c("CD4.H", "CD4.I")] = "tfhchron_tip"
mdata %>% count(annotation_level2, tip_mid)

library(dplyr)

summary_df <- mdata %>%
    group_by(IGTHT, tip_mid) %>%
    summarise(n = n(), .groups = "drop") %>%
    # total per sample
    group_by(IGTHT) %>%
    mutate(
        total_cells = sum(n),
        percent = n / total_cells * 100
    ) %>%
    ungroup()

n_categories_gt1 <- summary_df %>% 
    filter(tip_mid != "midland") %>%
    group_by(IGTHT) %>%
    summarise(
        n_categories_gt1 = sum(percent > 1),
        .groups = "drop"
    )

n_categories_gt1 %>% count(n_categories_gt1)

# library(tidyr)
# wide_counts <- summary_df %>%
#   select(IGTHT, tip_mid, n) %>%
#   pivot_wider(names_from = tip_mid, values_from = n, values_fill = 0)
# wide_percent <- summary_df %>%
#   select(IGTHT, tip_mid, percent) %>%
#   pivot_wider(names_from = tip_mid, values_from = percent, values_fill = 0)

n_categories_gt1 %>%
  count(n_categories_gt1) %>%
  ggplot(aes(x = factor(n_categories_gt1), y = n)) +
  geom_col(fill = "grey70", color = "black") +
  labs(
    x = "# tip_mid categories (>1%)",
    y = "# IGTHT samples"
  ) +
  theme_classic()

library(dplyr)
library(ggplot2)

plot_df <- n_categories_gt1 %>%
  count(n_categories_gt1) %>%
  mutate(percent = n / sum(n) * 100)

pdf(sprintf("%s/barplot_number_tips_per_sample.pdf", out_dir), 3,6, useDingbats = F)
ggplot(plot_df, aes(x = factor(n_categories_gt1), y = percent)) +
  geom_col(fill = "grey70", color = "black") +
  geom_text(
    aes(label = sprintf("%.1f%%", percent)),
    vjust = -0.3,
    size = 5
  ) +
  labs(
    x = "# tip (>1%)",
    y = "% of samples"
  ) +
  theme_classic(base_size = 16) +
  theme(
    axis.text = element_text(size = 14)
  ) +
  expand_limits(y = max(plot_df$percent) * 1.1)
dev.off()
```

**Figure 3 - tip signature**
--> limma_wrapper_20260318_CD4_tips.sh
limma_contrasts_20260318_CD4_tips.R
```{r}
ttlist = readRDS('/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/analysis/data_integration/IGT1_96/CD4/DGE_limma/20260318_tips/ttlist_tips.Rds')

# ttlist = read.table(file = '/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/analysis/data_integration/IGT1_96/CD4/DGE_limma/20260318_tips/ttlist_tips.txt', sep = "\t", header = T)
# ttlist["Ifng",]

names(ttlist)
tmp = ttlist

# i = "RS_vs_TUXYHI"

gene_sig = list()

pdf("/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/analysis/data_integration/IGT1_96/CD4/DGE_limma/20260318_tips/volcanos_more_labels.pdf", 15, 15, useDingbats = F)
for (i in names(ttlist)) {
    print(i)
    vplot = na.omit(data.frame(SYMBOL = tmp[[i]]$SYMBOL, fc = 2^tmp[[i]]$logFC, pval = tmp[[i]]$P.Value + 10^-300, AveExpr = 2^(tmp[[i]]$AveExpr))) #dealing with NA
    # vplot = data.frame(SYMBOL = tmp[[i]]$SYMBOL, fc = 2^tmp[[i]]$logFC, pval = tmp[[i]]$P.Value+10^-300, AveExpr = 2^(tmp[[i]]$AveExpr))
    p = Vplot(vplot = vplot, xlab = names(tmp)[i], xlimits = c(min(vplot$fc[vplot$fc != 0]), max(vplot$fc)), ylimits = c(min(vplot$pval), 1))  #c(min(vplot$pval[vplot$pval != 0]),1)
    sub = vplot %>%
        filter(log2(fc) > 0 & pval < 10^-60) %>%
        arrange(desc(log2(fc))) %>%
        pull(SYMBOL)
    sub
    gene_sig[[sprintf("%s_up", i)]] = sub
    highlight_genes = sub #unique(c(head(sub, 25), tail(sub, 25)))
    # sub = (log2(vplot$fc) > 0.5 | log2(vplot$fc) < -0.5)  & vplot$pval < 0.05
    # sub = vplot[sub,] %>% arrange(desc(log2(fc)))
    # sub1 = sub %>% pull(SYMBOL) %>% head(25) #vplot$SYMBOL %in% as.character(sub2$SYMBOL)[1:200]
    # sub2 = sub %>% pull(SYMBOL) %>% tail(25)
    highlight_filter = vplot$SYMBOL %in% highlight_genes
    # sub = vplot$SYMBOL %in% c(sub1, sub2)#as.character(sub2$SYMBOL)[1:50]
    library(ggrepel)
    print(p + geom_text_repel(data = vplot[highlight_filter,], aes(x = fc, y = pval, label = SYMBOL)) + ggtitle(i) + xlab(sprintf("FC %s", i)))
}
dev.off()


```

use pseudobulk and AUC so that keep only genes for which the third is at least 1.5 and mean expression in > 0.25 log1pCK10K
```{r}
pseudobulk_orig = readRDS(sprintf("%s/igt1_96_pseudobulk_byclusterannotationlevel2_log1pCP10K.Rds", "/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/analysis/data_integration/IGT1_96/TF_analysis/20260319"))

# pseudobulk_orig@meta.data %>% filter(grepl("CD4",annotation_level2), annotation_level2 %in% c())
pseudobulk = pseudobulk_orig[,pseudobulk_orig@meta.data %>% filter(annotation_level2 %in% c("CD4.R", "CD4.S", "CD4.T", "CD4.U", "CD4.X", "CD4.Y", "CD4.H", "CD4.I")) %>% rownames()]

library(dplyr)
library(ggplot2)
library(tibble)

gene = "Cd48"
plot_df <- pseudobulk[["RNA"]]$data[gene, , drop = FALSE] %>%
  as.data.frame() %>%
  rownames_to_column("gene") %>%
  tidyr::pivot_longer(
    cols = -gene,
    names_to = "cluster",
    values_to = "expression"
  )
ggplot(plot_df, aes(x = cluster, y = expression, fill = cluster)) +
  geom_col(color = "black") +
  scale_fill_manual(values = mypal_level2) +
  theme_classic() +
  labs(
    x = NULL,
    y = "pseudobulk expression"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) + ggtitle(gene)
 
names(gene_sig)
gene_sig_curated = list()
min_expr = 0.25

# gene_sig["RS_vs_TUXYHI_up"] %>% unlist() %>% as.character() %>% sort()
mat <- pseudobulk[["RNA"]]$data[gene_sig["RS_vs_TUXYHI_up"] %>% unlist() %>% as.character(),]
target_clusters <- c("CD4.R", "CD4.S")
other_clusters  <- setdiff(colnames(mat), target_clusters)
max_other <- apply(mat[, other_clusters, drop = FALSE], 1, max)
enriched_genes <- rownames(mat)[
  (mat[, target_clusters[1]] >= 1.5 * max_other) & mat[, target_clusters[1]] > min_expr &
  (mat[, target_clusters[2]] >= 1.5 * max_other) & mat[, target_clusters[2]] > min_expr
]
enriched_genes %>% sort()
gene_sig_curated[["RS_vs_TUXYHI_up"]] = enriched_genes

# gene_sig["TU_vs_RSXYHI_up"] %>% unlist() %>% as.character() %>% sort()
mat <- pseudobulk[["RNA"]]$data[gene_sig["TU_vs_RSXYHI_up"] %>% unlist() %>% as.character(),]
target_clusters <- c("CD4.T", "CD4.U")
other_clusters  <- setdiff(colnames(mat), target_clusters)
max_other <- apply(mat[, other_clusters, drop = FALSE], 1, max)
enriched_genes <- rownames(mat)[
  (mat[, target_clusters[1]] >= 1.5 * max_other) & mat[, target_clusters[1]] > min_expr &
  (mat[, target_clusters[2]] >= 1.5 * max_other) & mat[, target_clusters[2]] > min_expr
]
enriched_genes %>% sort()
gene_sig_curated[["TU_vs_RSXYHI_up"]] = enriched_genes

# gene_sig["XY_vs_RSTUHI_up"] %>% unlist() %>% as.character() %>% sort()
mat <- pseudobulk[["RNA"]]$data[gene_sig["XY_vs_RSTUHI_up"] %>% unlist() %>% as.character(),]
target_clusters <- c("CD4.X", "CD4.Y")
other_clusters  <- setdiff(colnames(mat), target_clusters)
max_other <- apply(mat[, other_clusters, drop = FALSE], 1, max)
enriched_genes <- rownames(mat)[
  (mat[, target_clusters[1]] >= 1.5 * max_other) & mat[, target_clusters[1]] > min_expr &
  (mat[, target_clusters[2]] >= 1.5 * max_other) & mat[, target_clusters[2]] > min_expr
]
enriched_genes %>% sort()
gene_sig_curated[["XY_vs_RSTUHI_up"]] = enriched_genes

# gene_sig["HI_vs_RSTUXY_up"] %>% unlist() %>% as.character() %>% sort()
mat <- pseudobulk[["RNA"]]$data[gene_sig["HI_vs_RSTUXY_up"] %>% unlist() %>% as.character(),]
target_clusters <- c("CD4.H", "CD4.I")
other_clusters  <- setdiff(colnames(mat), target_clusters)
max_other <- apply(mat[, other_clusters, drop = FALSE], 1, max)
enriched_genes <- rownames(mat)[
  (mat[, target_clusters[1]] >= 1.5 * max_other) & mat[, target_clusters[1]] > min_expr &
  (mat[, target_clusters[2]] >= 1.5 * max_other) & mat[, target_clusters[2]] > min_expr
]
enriched_genes %>% sort()
gene_sig_curated[["HI_vs_RSTUXY_up"]] = enriched_genes

#ensure_directory(sprintf("%s/tip_signatures", out_dir))
for (i in names(gene_sig_curated)) {
    write.table(gene_sig_curated[[i]], file = sprintf("%s/tip_signatures/%s.txt", out_dir, i), sep = "\t",quote = F, row.names = F, col.names = F)
}
```

Export Fig S3
```{r}
ttlist = readRDS('/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/analysis/data_integration/IGT1_96/CD4/DGE_limma/20260318_tips/ttlist_tips.Rds')

names(ttlist)

i = "RS_vs_TUXYHI"
sig = read.table(file = sprintf("%s/tip_signatures/%s_up.txt", "~/google_drive_uchicago/ImmgenT/analysis/data_integration/IGT1_96/CD4/", i))[,1]
ttlist[[i]][sig,]
write.table(ttlist[[i]][sig,c("SYMBOL", "logFC", "adj.P.Val")], file = sprintf("%s/tip_signatures/%s_table.txt", "~/google_drive_uchicago/ImmgenT/analysis/data_integration/IGT1_96/CD4/", i), sep = "\t",quote = F, row.names = F, col.names = T)


i = "TU_vs_RSXYHI"
sig = read.table(file = sprintf("%s/tip_signatures/%s_up.txt", "~/google_drive_uchicago/ImmgenT/analysis/data_integration/IGT1_96/CD4/", i))[,1]
ttlist[[i]][sig,]
write.table(ttlist[[i]][sig,c("SYMBOL", "logFC", "adj.P.Val")], file = sprintf("%s/tip_signatures/%s_table.txt", "~/google_drive_uchicago/ImmgenT/analysis/data_integration/IGT1_96/CD4/", i), sep = "\t",quote = F, row.names = F, col.names = T)

i = "XY_vs_RSTUHI"
sig = read.table(file = sprintf("%s/tip_signatures/%s_up.txt", "~/google_drive_uchicago/ImmgenT/analysis/data_integration/IGT1_96/CD4/", i))[,1]
ttlist[[i]][sig,]
write.table(ttlist[[i]][sig,c("SYMBOL", "logFC", "adj.P.Val")], file = sprintf("%s/tip_signatures/%s_table.txt", "~/google_drive_uchicago/ImmgenT/analysis/data_integration/IGT1_96/CD4/", i), sep = "\t",quote = F, row.names = F, col.names = T)

i = "HI_vs_RSTUXY"
sig = read.table(file = sprintf("%s/tip_signatures/%s_up.txt", "~/google_drive_uchicago/ImmgenT/analysis/data_integration/IGT1_96/CD4/", i))[,1]
ttlist[[i]][sig,]
write.table(ttlist[[i]][sig,c("SYMBOL", "logFC", "adj.P.Val")], file = sprintf("%s/tip_signatures/%s_table.txt", "~/google_drive_uchicago/ImmgenT/analysis/data_integration/IGT1_96/CD4/", i), sep = "\t",quote = F, row.names = F, col.names = T)


```


Plotting on the volcano plot the tip signatures
```{r}
Vplot = function (vplot, xlab = "FC", xlimits = c(0.1, 10), ylimits = c(10^-300, 1), color = "grey", size = 1, xlab = "FC", ylab = "p value") 
{
    requireNamespace("ggplot2", quietly = TRUE)
    requireNamespace("ggrastr", quietly = TRUE)
    requireNamespace("scales", quietly = TRUE)
    p = ggplot(data = vplot) + ggrastr::geom_point_rast(aes(x = .data$fc, 
        y = .data$pval), colour = color, alpha = I(1), size = size, 
        raster.dpi = 100) + scale_x_continuous(trans = scales::log_trans(10), 
        limits = xlimits) + scale_y_continuous(trans = scales::log_trans(10), 
        breaks = scales::trans_breaks("log10", function(x) 10^x), 
        labels = scales::trans_format("log10", scales::math_format(10^.x)), 
        limits = ylimits) + annotation_logticks(sides = "b") + 
        xlab(xlab) + theme_bw() + ylab(ylab) + theme(axis.text.x = element_text(size = 20, 
        angle = 0, hjust = 0.5), axis.text.y = element_text(size = 20), 
        legend.text = element_text(size = 20), axis.title.x = element_text(size = 20), 
        axis.title.y = element_text(size = 20))
    return(p)
}


comp_to_plot = c("RS_vs_TUXYHI", "TU_vs_RSXYHI", "XY_vs_RSTUHI", "HI_vs_RSTUXY")
cols = c("darkorchid", "violetred2", "burlywood2","darkseagreen2" )
names(cols) = comp_to_plot
tmp = ttlist
pdf(sprintf("%s/Volcanos_Tips_signature.pdf", out_dir), 10, 10, useDingbats = F)
for (i in comp_to_plot) {
    print(i)
    vplot = na.omit(data.frame(SYMBOL = tmp[[i]]$SYMBOL, fc = 2^tmp[[i]]$logFC, pval = tmp[[i]]$P.Value + 10^-300, AveExpr = 2^(tmp[[i]]$AveExpr))) #dealing with NA
    # vplot = data.frame(SYMBOL = tmp[[i]]$SYMBOL, fc = 2^tmp[[i]]$logFC, pval = tmp[[i]]$P.Value+10^-300, AveExpr = 2^(tmp[[i]]$AveExpr))
    p = Vplot(vplot = vplot, xlab = names(tmp)[i], xlimits = c(min(vplot$fc[vplot$fc != 0]), max(vplot$fc)), ylimits = c(min(vplot$pval), 1), color = cols[i], size = 2)  #c(min(vplot$pval[vplot$pval != 0]),1)
    highlight_genes = gene_sig_curated[[sprintf("%s_up", i)]]
    highlight_filter = vplot$SYMBOL %in% highlight_genes
    # sub = vplot$SYMBOL %in% c(sub1, sub2)#as.character(sub2$SYMBOL)[1:50]
    library(ggrepel)
    print(p)
    print(p + geom_text_repel(data = vplot[highlight_filter,], aes(x = fc, y = pval, label = SYMBOL)) + ggtitle(i) + xlab(sprintf("FC %s", i)))
}
dev.off()

```


**Difference between Q.Th1 and RS tip, Q.Th2 and TU tip etc?**
--> th_modulescores_withannot.Rds
```{r}
so_orig = readRDS("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/igt1_96_withtotalvi20260206_clean.Rds") 
so = so_orig[,so_orig@meta.data %>% filter(, annotation_level1 == "CD4", organ_simplified != "thymus") %>% row.names()]

th_scores = readRDS("module_score_matrix.Rds") %>% t()
th_scores[1:2,]
thr_th1 = 0
thr_th2 = 0.25
thr_th17 = 0.25
thr_tfhchron = 0

# correct binarization
th_scores_bin <- data.frame(
  row.names = rownames(th_scores),
  Th1 = th_scores[, "Th1 program"] > thr_th1,
  Th2 = th_scores[, "Th2 program"] > thr_th2,
  Th17 = th_scores[, "Th17 program"] > thr_th17,
  Tfhchron = th_scores[, "Tfh/Tchron program"] > thr_tfhchron
)

cell_class <- apply(th_scores_bin, 1, function(x) {
  pos <- colnames(th_scores_bin)[as.logical(x)]
  if (length(pos) == 0) "None" else paste(pos, collapse = "")
})

all(rownames(th_scores_bin) == names(cell_class))
th_scores_bin$program = cell_class

th_scores = as.data.frame(th_scores)
all(rownames(th_scores) == names(cell_class))
th_scores$program = cell_class

th_scores_bin %>% group_by(program) %>% summarize(n = n()) %>% mutate(percent = n/sum(n)*100)
th_scores %>% group_by(program) %>% summarize(n = n()) %>% mutate(percent = n/sum(n)*100)

# saveRDS(th_scores, sprintf("%s/th_modulescores_withannot.Rds", "~/google_drive_uchicago/ImmgenT/analysis/data_integration/IGT1_96/CD4/"))

```

Barplot of distribution of Th programs
```{r}

# so <- readRDS(sprintf("%s/CD4_GHIQRSTUXY_igt1_96_withtotalvi20260206_clean_withThPrograms.Rds", out_dir))

th_scores = readRDS(sprintf("%s/th_modulescores_withannot.Rds", "~/google_drive_uchicago/ImmgenT/analysis/data_integration/IGT1_96/CD4/"))
th_scores_bin = th_scores

library(dplyr)
library(tidyr)
library(ggplot2)
library(patchwork)

plot_df <- th_scores_bin %>%
  group_by(program) %>%
  summarise(n = n(), .groups = "drop") %>%
  mutate(percent = n / sum(n) * 100)

# program_order <- plot_df %>%
#   arrange(desc(percent)) %>%
#   pull(program)

program_order <- c(
  "None",
  "Th1", "Th2", "Th17", "Tfhchron",
  "Th1Th2", "Th1Th17", "Th1Tfhchron", "Th2Th17", "Th2Tfhchron", "Th17Tfhchron",
  "Th1Th2Th17", "Th1Th2Tfhchron", "Th1Th17Tfhchron", "Th2Th17Tfhchron",
  "Th1Th2Th17Tfhchron"
)

plot_df$program <- factor(plot_df$program, levels = program_order)

library(stringr)
heatmap_df <- expand.grid(
  row = c("Th1", "Th2", "Th17", "Tfhchron"),
  program = program_order,
  stringsAsFactors = FALSE
) %>%
  rowwise() %>%
  mutate(
    members = list(str_extract_all(program, "Tfhchron|Th17|Th2|Th1")[[1]]),
    filled = as.integer(row %in% members)
  ) %>%
  ungroup() %>%
  select(-members)

heatmap_df$program <- factor(heatmap_df$program, levels = program_order)
heatmap_df$row <- factor(heatmap_df$row, levels = rev(c("Th1", "Th2", "Th17", "Tfhchron")))

p_bar <- ggplot(plot_df, aes(x = program, y = percent)) +
  geom_col(fill = "grey70", color = "black") +
  geom_text(
    aes(label = sprintf("%.1f", percent)),
    vjust = -0.25,
    size = 2.8
  ) +
  labs(
    x = NULL,
    y = "% of cells"
  ) +
  theme_classic(base_size = 16) +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.y = element_text(size = 16),
    axis.text.y = element_text(size = 14)
  ) +
  expand_limits(y = max(plot_df$percent) * 1.08)

p_heat <- ggplot(heatmap_df, aes(x = program, y = row, fill = factor(filled))) +
  geom_tile(color = "black", linewidth = 0.4) +
  scale_fill_manual(values = c("0" = "white", "1" = "black"), guide = "none") +
  labs(
    x = NULL,
    y = NULL
  ) +
  theme_classic(base_size = 16) +
  theme(
    axis.line = element_blank(),
    axis.ticks = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 12),
    axis.text.y = element_text(size = 14)
  )

p = p_bar / p_heat + plot_layout(heights = c(4, 1.2))

p

pdf(sprintf("%s/barplot_polyfunctional.pdf", out_dir), 8,8)
p
dev.off()

```

Barplot of distribution of Th programs in G and Q only
```{r}

so <- readRDS(sprintf("%s/CD4_GHIQRSTUXY_igt1_96_withtotalvi20260206_clean_withThPrograms.Rds", out_dir))

th_scores = readRDS(sprintf("%s/th_modulescores_withannot.Rds", "~/google_drive_uchicago/ImmgenT/analysis/data_integration/IGT1_96/CD4/"))
th_scores_bin = th_scores[so@meta.data %>% filter(annotation_level2 %in% c("CD4.G", "CD4.Q")) %>% row.names(),]


library(dplyr)
library(tidyr)
library(ggplot2)
library(patchwork)

plot_df <- th_scores_bin %>%
  group_by(program) %>%
  summarise(n = n(), .groups = "drop") %>%
  mutate(percent = n / sum(n) * 100)

# program_order <- plot_df %>%
#   arrange(desc(percent)) %>%
#   pull(program)

program_order <- c(
  "None",
  "Th1", "Th2", "Th17", "Tfhchron",
  "Th1Th2", "Th1Th17", "Th1Tfhchron", "Th2Th17", "Th2Tfhchron", "Th17Tfhchron",
  "Th1Th2Th17", "Th1Th2Tfhchron", "Th1Th17Tfhchron", "Th2Th17Tfhchron",
  "Th1Th2Th17Tfhchron"
)

plot_df$program <- factor(plot_df$program, levels = program_order)

# heatmap_df <- expand.grid(
#   row = c("Th1", "Th2", "Th17", "Tfhchron"),
#   program = program_order,
#   stringsAsFactors = FALSE
# ) %>%
#   mutate(
#     filled = case_when(
#       row == "Th1"      & grepl("Th1", program) ~ 1,
#       row == "Th2"      & grepl("Th2", program) ~ 1,
#       row == "Th17"     & grepl("Th17", program) ~ 1,
#       row == "Tfhchron" & grepl("Tfhchron", program) ~ 1,
#       TRUE ~ 0
#     )
#   )

# heatmap_df <- expand.grid(
#   row = c("Th1", "Th2", "Th17", "Tfhchron"),
#   program = program_order,
#   stringsAsFactors = FALSE
# ) %>%
#   rowwise() %>%
#   mutate(
#     members = list(c(
#       if (grepl("Tfhchron", program)) "Tfhchron" else NULL,
#       if (grepl("Th1", program)) "Th1" else NULL,
#       if (grepl("Th2", program)) "Th2" else NULL,
#       if (grepl("Th17", program)) "Th17" else NULL
#     )),
#     filled = as.integer(row %in% members)
#   ) %>%
#   ungroup() %>%
#   select(-members)

library(stringr)
heatmap_df <- expand.grid(
  row = c("Th1", "Th2", "Th17", "Tfhchron"),
  program = program_order,
  stringsAsFactors = FALSE
) %>%
  rowwise() %>%
  mutate(
    members = list(str_extract_all(program, "Tfhchron|Th17|Th2|Th1")[[1]]),
    filled = as.integer(row %in% members)
  ) %>%
  ungroup() %>%
  select(-members)

heatmap_df$program <- factor(heatmap_df$program, levels = program_order)
heatmap_df$row <- factor(heatmap_df$row, levels = rev(c("Th1", "Th2", "Th17", "Tfhchron")))

p_bar <- ggplot(plot_df, aes(x = program, y = percent)) +
  geom_col(fill = "grey70", color = "black") +
  geom_text(
    aes(label = sprintf("%.1f", percent)),
    vjust = -0.25,
    size = 2.8
  ) +
  labs(
    x = NULL,
    y = "% of cells"
  ) +
  theme_classic(base_size = 16) +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.y = element_text(size = 16),
    axis.text.y = element_text(size = 14)
  ) +
  expand_limits(y = max(plot_df$percent) * 1.08)

p_heat <- ggplot(heatmap_df, aes(x = program, y = row, fill = factor(filled))) +
  geom_tile(color = "black", linewidth = 0.4) +
  scale_fill_manual(values = c("0" = "white", "1" = "black"), guide = "none") +
  labs(
    x = NULL,
    y = NULL
  ) +
  theme_classic(base_size = 16) +
  theme(
    axis.line = element_blank(),
    axis.ticks = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 12),
    axis.text.y = element_text(size = 14)
  )

p = p_bar / p_heat + plot_layout(heights = c(4, 1.2))

p

pdf(sprintf("%s/barplot_polyfunctional_GQOnly.pdf", out_dir), 8,8)
p
dev.off()


```


Barplot polyfunctional 2
```{r}
so <- readRDS(sprintf("%s/CD4_GHIQRSTUXY_igt1_96_withtotalvi20260206_clean_withThPrograms.Rds", out_dir))

so@meta.data %>% count(program)
so$is_polyfunctional <- !so@meta.data$program %in% c("None", "Tfhchron", "Th1", "Th17", "Th2")

pf <- so@meta.data %>%
  as.data.frame() %>%
  mutate(
    cluster_group = case_when(
      annotation_level2 == "CD4.G" ~ "G",
      annotation_level2 == "CD4.H" ~ "HI",
      annotation_level2 == "CD4.I" ~ "HI",
      annotation_level2 == "CD4.Q" ~ "Q",
      annotation_level2 %in% c("CD4.R","CD4.S") ~ "R/S",
      annotation_level2 %in% c("CD4.T","CD4.U") ~ "T/U",
      annotation_level2 %in% c("CD4.X","CD4.Y") ~ "X/Y",
      TRUE ~ NA_character_
    )
  ) %>%
  filter(!is.na(cluster_group))

pf_summary <- pf %>%
  group_by(cluster_group) %>%
  summarise(
    percent_polyfunctional = mean(is_polyfunctional) * 100,
    n_cells = n()
  )

p1 <- ggplot(pf_summary, aes(x = cluster_group, y = percent_polyfunctional)) +
  geom_col(fill = "black") +
  geom_text(
    aes(label = sprintf("%.1f%%", percent_polyfunctional)),
    vjust = -0.5,
    size = 4
  ) +
  theme_classic() +
  labs(
    x = "Cluster",
    y = "% Polyfunctional cells"
  )  + coord_cartesian(ylim = c(0,20))

# pdf("polyfunctional_barplot.pdf", width = 8, height = 6, useDingbats = FALSE)
# print(p)
# dev.off()

#prop of polyfunctional, coming from each
pf_sub <- subset(pf, subset = is_polyfunctional== "TRUE")

pf_summary <- pf_sub %>%
  group_by(cluster_group) %>%
  summarise(
    n_cells = n()
  ) %>%
  mutate(
    percent_of_polyfunctional = n_cells / sum(n_cells) * 100
  )


p2 <- ggplot(pf_summary, aes(x = cluster_group, y = percent_of_polyfunctional)) +
  geom_col(fill = "black") +
  geom_text(
    aes(label = sprintf("%.1f%%", percent_of_polyfunctional)),
    vjust = -0.5,
    size = 4
  ) +
  theme_classic() +
  labs(
    x = "Cluster",
    y = "% of All Polyfunctional cells"
  ) + coord_cartesian(ylim = c(0,45))

pdf(sprintf("%s/barplot_polyfunctional_percent.pdf", out_dir), width = 4, height = 8, useDingbats = FALSE)
print(p1)
print(p2)
dev.off()
```

Barplot Non committed
```{r}
so <- readRDS(sprintf("%s/CD4_GHIQRSTUXY_igt1_96_withtotalvi20260206_clean_withThPrograms.Rds", out_dir))

so@meta.data %>% count(program)
so$is_noncom <- so@meta.data$program %in% c("None")

pf <- so@meta.data %>%
  as.data.frame() %>%
  mutate(
    cluster_group = case_when(
      annotation_level2 == "CD4.G" ~ "G",
      annotation_level2 == "CD4.H" ~ "HI",
      annotation_level2 == "CD4.I" ~ "HI",
      annotation_level2 == "CD4.Q" ~ "Q",
      annotation_level2 %in% c("CD4.R","CD4.S") ~ "R/S",
      annotation_level2 %in% c("CD4.T","CD4.U") ~ "T/U",
      annotation_level2 %in% c("CD4.X","CD4.Y") ~ "X/Y",
      TRUE ~ NA_character_
    )
  ) %>%
  filter(!is.na(cluster_group))

pf_summary <- pf %>%
  group_by(cluster_group) %>%
  summarise(
    percent_noncom = mean(is_noncom) * 100,
    n_cells = n()
  )

p1 <- ggplot(pf_summary, aes(x = cluster_group, y = percent_noncom)) +
  geom_col(fill = "black") +
  geom_text(
    aes(label = sprintf("%.1f%%", percent_noncom)),
    vjust = -0.5,
    size = 4
  ) +
  theme_classic() +
  labs(
    x = "Cluster",
    y = "% Non Committed cells"
  )  + coord_cartesian(ylim = c(0,100))

# pdf("polyfunctional_barplot.pdf", width = 8, height = 6, useDingbats = FALSE)
# print(p)
# dev.off()

#prop of polyfunctional, coming from each
pf_sub <- subset(pf, subset = is_noncom == "TRUE")

pf_summary <- pf_sub %>%
  group_by(cluster_group) %>%
  summarise(
    n_cells = n()
  ) %>%
  mutate(
    percent_of_noncom = n_cells / sum(n_cells) * 100
  )


p2 <- ggplot(pf_summary, aes(x = cluster_group, y = percent_of_noncom)) +
  geom_col(fill = "black") +
  geom_text(
    aes(label = sprintf("%.1f%%", percent_of_noncom)),
    vjust = -0.5,
    size = 4
  ) +
  theme_classic() +
  labs(
    x = "Cluster",
    y = "% of All Non Committed cells"
  ) + coord_cartesian(ylim = c(0,50))

pdf(sprintf("%s/barplot_noncommitted_percent.pdf", out_dir), width = 4, height = 5, useDingbats = FALSE)
print(p1)
print(p2)
dev.off()
```


**Make seurat for DGE limma with program annotation to compare Q_th with tips and G-Th with tips**
```{r}
th_scores = readRDS(sprintf("%s/th_modulescores_withannot.Rds", "~/google_drive_uchicago/ImmgenT/analysis/data_integration/IGT1_96/CD4"))

so
so2 = so[,rownames(th_scores)]
so2@meta.data = data.frame(so2@meta.data, th_scores)
so2@meta.data %>% count(annotation_level2, program)
so2$annotation_level2_program = paste(so2$annotation_level2, so2$program, sep = ".")
so2@meta.data %>% count(annotation_level2, annotation_level2_program)

# saveRDS(so2, file = sprintf("%s/CD4_GHIQRSTUXY_igt1_96_withtotalvi20260206_clean_withThPrograms.Rds", out_dir))

so = readRDS(file = sprintf("%s/CD4_GHIQRSTUXY_igt1_96_withtotalvi20260206_clean_withThPrograms.Rds", out_dir))
so$annotation_level2_program[so$annotation_level2 %in% c("CD4.H")] = "CD4.H"
so$annotation_level2_program[so$annotation_level2 %in% c("CD4.I")] = "CD4.I"
so$annotation_level2_program[so$annotation_level2 %in% c("CD4.R")] = "CD4.R"
so$annotation_level2_program[so$annotation_level2 %in% c("CD4.S")] = "CD4.S"
so$annotation_level2_program[so$annotation_level2 %in% c("CD4.T")] = "CD4.T"
so$annotation_level2_program[so$annotation_level2 %in% c("CD4.U")] = "CD4.U"
so$annotation_level2_program[so$annotation_level2 %in% c("CD4.X")] = "CD4.X"
so$annotation_level2_program[so$annotation_level2 %in% c("CD4.Y")] = "CD4.Y"

```

For Q programs vs tips
-->limma_wrapper_20260320_CD4_ThprogramsQ.sh
```{r}
#For Q programs vs tips
so@meta.data %>% count(annotation_level2, annotation_level2_program)
so@meta.data %>% filter(annotation_level2_program %in% c("CD4.R", "CD4.S", "CD4.T", "CD4.U", "CD4.X", "CD4.Y", "CD4.Q.Th1", "CD4.Q.Th2", "CD4.Q.Th17")) %>% count(annotation_level2, annotation_level2_program)
so2 = so[,so@meta.data %>% filter(annotation_level2_program %in% c("CD4.R", "CD4.S", "CD4.T", "CD4.U", "CD4.X", "CD4.Y", "CD4.Q.Th1", "CD4.Q.Th2", "CD4.Q.Th17")) %>% rownames()]
so2$annotation_level2_program.IGTHT = paste(so2$annotation_level2_program, so2$IGTHT, sep = ".")
# saveRDS(so2, file = sprintf("%s/CD4_QRSTUXY_igt1_96_withtotalvi20260206_clean_annotation_level2_program.Rds", out_dir))

```

on desktop
```{r}
ttlist = readRDS('/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/analysis/data_integration/IGT1_96/CD4/DGE_limma/20260320_CD4_QprogramsvsTips/ttlist_QprogramsvsTips.Rds')
tmp = ttlist
gene_sig = list()

cols = c("darkorchid", "violetred2", "burlywood2" )
names(cols) = names(tmp)

pdf(sprintf("%s/Volcano_QprogramsvsTips.pdf", out_dir), 15, 15, useDingbats = F)
for (i in names(ttlist)) {
    print(i)
    vplot = na.omit(data.frame(SYMBOL = tmp[[i]]$SYMBOL, fc = 2^tmp[[i]]$logFC, pval = tmp[[i]]$P.Value + 10^-80, AveExpr = 2^(tmp[[i]]$AveExpr))) #dealing with NA
    # vplot = data.frame(SYMBOL = tmp[[i]]$SYMBOL, fc = 2^tmp[[i]]$logFC, pval = tmp[[i]]$P.Value+10^-300, AveExpr = 2^(tmp[[i]]$AveExpr))
    p = Vplot(vplot = vplot, xlab = names(tmp)[i], xlimits = c(min(vplot$fc[vplot$fc != 0]), max(vplot$fc)), ylimits = c(min(vplot$pval), 1),size = 2, color = cols[i])  #c(min(vplot$pval[vplot$pval != 0]),1)
    sub = vplot %>%
        filter(abs(log2(fc)) > 0 & pval < 10^-5) %>%
        arrange(desc(log2(fc))) %>%
        pull(SYMBOL)
    sub
    # gene_sig[[sprintf("%s_up", i)]] = sub
    highlight_genes = sub #unique(c(head(sub, 25), tail(sub, 25)))
    # sub = (log2(vplot$fc) > 0.5 | log2(vplot$fc) < -0.5)  & vplot$pval < 0.05
    # sub = vplot[sub,] %>% arrange(desc(log2(fc)))
    # sub1 = sub %>% pull(SYMBOL) %>% head(25) #vplot$SYMBOL %in% as.character(sub2$SYMBOL)[1:200]
    # sub2 = sub %>% pull(SYMBOL) %>% tail(25)
    highlight_filter = vplot$SYMBOL %in% highlight_genes
    # sub = vplot$SYMBOL %in% c(sub1, sub2)#as.character(sub2$SYMBOL)[1:50]
    library(ggrepel)
    print(p)
    print(p + geom_text_repel(data = vplot[highlight_filter,], aes(x = fc, y = pval, label = SYMBOL)) + ggtitle(i) + xlab(sprintf("FC %s", i)))
}
dev.off()




```

For G programs vs tips
-->limma_wrapper_20260320_CD4_ThprogramsG.sh
```{r}
#For G programs vs tips
so2 = so[,so@meta.data %>% filter(annotation_level2_program %in% c("CD4.R", "CD4.S", "CD4.T", "CD4.U", "CD4.X", "CD4.Y", "CD4.H", "CD4.I", "CD4.G.Th1", "CD4.G.Th2", "CD4.G.Th17", "CD4.G.Tfhchron")) %>% rownames()]
so2@meta.data %>% count(annotation_level2, annotation_level2_program)
so2$annotation_level2_program.IGTHT = paste(so2$annotation_level2_program, so2$IGTHT, sep = ".")
saveRDS(so2, file = sprintf("%s/CD4_GHIRSTUXY_igt1_96_withtotalvi20260206_clean_annotation_level2_program.Rds", out_dir))

```

on desktop
```{r}
ttlist = readRDS('/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/analysis/data_integration/IGT1_96/CD4/DGE_limma/20260320_CD4_GprogramsvsTips/ttlist_GprogramsvsTips.Rds')
tmp = ttlist
gene_sig = list()

cols = c("darkorchid", "violetred2", "burlywood2","darkseagreen2" )
names(cols) = names(tmp)

pdf(sprintf("%s/Volcano_GprogramsvsTips.pdf", out_dir), 15, 15, useDingbats = F)
for (i in names(ttlist)) {
    print(i)
    vplot = na.omit(data.frame(SYMBOL = tmp[[i]]$SYMBOL, fc = 2^tmp[[i]]$logFC, pval = tmp[[i]]$P.Value + 10^-20, AveExpr = 2^(tmp[[i]]$AveExpr))) #dealing with NA
    # vplot = data.frame(SYMBOL = tmp[[i]]$SYMBOL, fc = 2^tmp[[i]]$logFC, pval = tmp[[i]]$P.Value+10^-300, AveExpr = 2^(tmp[[i]]$AveExpr))
    p = Vplot(vplot = vplot, xlab = names(tmp)[i], xlimits = c(min(vplot$fc[vplot$fc != 0]), max(vplot$fc)), ylimits = c(min(vplot$pval), 1),size = 2, color = cols[i])  #c(min(vplot$pval[vplot$pval != 0]),1)
    sub = vplot %>%
        filter(abs(log2(fc)) > 0 & pval < 10^-3) %>%
        arrange(desc(log2(fc))) %>%
        pull(SYMBOL)
    sub
    # gene_sig[[sprintf("%s_up", i)]] = sub
    highlight_genes = sub #unique(c(head(sub, 25), tail(sub, 25)))
    # sub = (log2(vplot$fc) > 0.5 | log2(vplot$fc) < -0.5)  & vplot$pval < 0.05
    # sub = vplot[sub,] %>% arrange(desc(log2(fc)))
    # sub1 = sub %>% pull(SYMBOL) %>% head(25) #vplot$SYMBOL %in% as.character(sub2$SYMBOL)[1:200]
    # sub2 = sub %>% pull(SYMBOL) %>% tail(25)
    highlight_filter = vplot$SYMBOL %in% highlight_genes
    # sub = vplot$SYMBOL %in% c(sub1, sub2)#as.character(sub2$SYMBOL)[1:50]
    library(ggrepel)
    print(p)
    print(p + geom_text_repel(data = vplot[highlight_filter,], aes(x = fc, y = pval, label = SYMBOL)) + ggtitle(i) + xlab(sprintf("FC %s", i)))
}
dev.off()


```

**Heatmap with DGE between different Th clusters**
Heatmaps show top 10 genes up and top 10 down with abs(log2FC) > 0.5 and p < 10^-10 in at least one comparison

Load signatures
```{r}
ttlist1 = readRDS('/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/analysis/data_integration/IGT1_96/CD4/DGE_limma/20260318_tips/ttlist_tips.Rds')
ttlist1 = ttlist1[8:10]
names(ttlist1)
ttlist2 = readRDS('/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/analysis/data_integration/IGT1_96/CD4/DGE_limma/20260320_CD4_QprogramsvsTips/ttlist_QprogramsvsTips.Rds')
names(ttlist2)
ttlist3 = readRDS('/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/analysis/data_integration/IGT1_96/CD4/DGE_limma/20260320_CD4_GprogramsvsTips/ttlist_GprogramsvsTips.Rds')
names(ttlist3)
ttlist3 = ttlist3[1:3]
names(ttlist3)

```

Genes present in at least 1 comparison
```{r}
tmp = list(ttlist2[["RS_vs_Qth1"]], ttlist2[["TU_vs_Qth2"]], ttlist2[["XY_vs_Qth17"]])
names(tmp) = c( "RS_vs_Qth1", "TU_vs_Qth2", "XY_vs_Qth17")

tmp_list = list()

names(tmp) 
gene_sig = list()

cols = c("grey", "grey", "grey", "grey" )
names(cols) = names(tmp)

comp_filt = names(tmp)

genes_to_label <- c()
for (i in comp_filt) {
    df <- tmp[[i]] #%>% filter(adj.P.Val < 0.05)
    # top up
    up_genes <- df %>% filter(logFC > 0.5,  P.Value < 10^-10) %>% 
        arrange(desc(logFC)) %>%
        pull(SYMBOL)
    # top down
    down_genes <- df %>% filter(logFC < -0.5,  P.Value < 10^-10) %>% 
        arrange(logFC) %>%
        pull(SYMBOL)
    # remove genes already selected
    # up_genes <- setdiff(up_genes, genes_to_label)
    # down_genes <- setdiff(down_genes, genes_to_label)
    
    # add top 20 each
    genes_to_label <- c(
        genes_to_label,
        head(up_genes, 1000), head(down_genes, 1000)
        # head(up_genes, 100), head(down_genes, 100)
    )
}

genes_to_label <- genes_to_label[
    !grepl(
        "^Gm|^mt-|^Rpl|^Rps|^Snor|^Mir|^Linc|^Rp[sl]|-ps|Rik$|^Hbb",
        genes_to_label,
        ignore.case = TRUE
    )
] %>% unique() %>% sort()
genes_to_label

# genes_to_label_top100 = list()
# genes_to_label_top100[["Th1"]] = genes_to_label

#plot heatmap
# mat_coef <- sapply(tmp, function(tmp) tmp$logFC)
lapply(tmp, function(x) all(x$SYMBOL == tmp[[1]]$SYMBOL)) ##check that rownames match!!!

common_genes <- Reduce(intersect, lapply(tmp, function(x) x$SYMBOL))
tmp <- lapply(tmp, function(x) {
  x %>%
    dplyr::filter(SYMBOL %in% common_genes) %>%
    dplyr::arrange(match(SYMBOL, common_genes))
})
lapply(tmp, function(x) identical(x$SYMBOL, tmp[[1]]$SYMBOL))

mat_coef <- Reduce(cbind, lapply(tmp, function(x) x[,"logFC"]))
rownames(mat_coef) = tmp[[1]]$SYMBOL
colnames(mat_coef) = make.names(names(tmp))
mat_plot = mat_coef[genes_to_label, ]
range(mat_plot)

# tmp[["R_vs_S"]][genes_to_label,1]
# mat_coef[genes_to_label,"R_vs_S"]

library(ComplexHeatmap)
library(circlize)

col_fun <- circlize::colorRamp2(
    c(-2, 0, 2),
    c("#4575B4", "white", "#D73027")
)

# # pdf(sprintf("%s/heatmap_memory_vs_acute_SLO_LCMV.pdf", out_dir), width = 3, 8, useDingbats = F)
# pdf(sprintf("%s/Heatmap_Tips_vsQTh_AllGenes.pdf", out_dir), width = 3, 15, useDingbats = F)
# # pdf(sprintf("%s/Heatmap_Th1Clusters_Top10Genes.pdf", out_dir), width = 3, 5, useDingbats = F)
# Heatmap(
#     mat_plot,
#     name = "log2FC",
#     col = col_fun,
#     cluster_columns = FALSE,
#     cluster_rows = T,
#     row_names_gp = gpar(fontsize = 6),
#     column_names_gp = gpar(fontsize = 6),
#     column_title = "Tips vs QTh clusters"
# )
# dev.off()

pdf(sprintf("%s/Heatmap_Tips_vsQTh_AllGenes.pdf", out_dir), width = 3, 5, useDingbats = F)
# pdf(sprintf("%s/Heatmap_Th1Clusters_Top10Genes.pdf", out_dir), width = 3, 5, useDingbats = F)
Heatmap(
    mat_plot,
    name = "log2FC",
    col = col_fun,
    cluster_columns = FALSE,
    cluster_rows = T,
    row_names_gp = gpar(fontsize =2),
    column_names_gp = gpar(fontsize = 6),
    column_title = "Tips vs QTh clusters"
)
dev.off()


###Suppl Table
mat_coef <- Reduce(cbind, lapply(tmp, function(x) x[,"logFC"]))
rownames(mat_coef) = tmp[[1]]$SYMBOL
colnames(mat_coef) = make.names(paste(names(tmp), "_log2FC", sep = ""))
mat_coef = mat_coef[genes_to_label, ]
mat_pv <- Reduce(cbind, lapply(tmp, function(x) x[,"adj.P.Val"]))
rownames(mat_pv) = tmp[[1]]$SYMBOL
colnames(mat_pv) = make.names(paste(names(tmp), "_adj.P.Value", sep = ""))
mat_pv = mat_pv[genes_to_label, ]
all(rownames(mat_coef) == rownames(mat_pv))

# write.table(data.frame(SYMBOL = rownames(mat_coef), mat_coef, mat_pv), file = sprintf("%s/Tips_vsQTh_table.txt", out_dir), sep = "\t", quote = F, row.names = F)


###FCFC plots
pdf(sprintf("%s/FCFC_Tips_vsQTh1Th2.pdf", out_dir), width = 30, 30, useDingbats = F)
# pdf(sprintf("%s/Heatmap_Th1Clusters_Top10Genes.pdf", out_dir), width = 3, 5, useDingbats = F)

vplot = data.frame(fc1 = 2^tmp$RS_vs_Qth1$logFC, fc2 = 2^(tmp$TU_vs_Qth2$logFC), SYMBOL = tmp$RS_vs_Qth1$SYMBOL)
genes_to_label = c(tmp$RS_vs_Qth1 %>% filter(abs(logFC) > 0.5,  P.Value < 10^-10) %>% pull(SYMBOL) %>% head(100),
                   tmp$TU_vs_Qth2 %>% filter(abs(logFC) > 0.5,  P.Value < 10^-10) %>% pull(SYMBOL) %>% head(100)) %>% unique()
                   #     
                   #     arrange(desc(logFC)) %>% pull(SYMBOL) %>% head(20),
                   # tmp$RS_vs_Qth1 %>% arrange(desc(logFC)) %>% pull(SYMBOL) %>% tail(20),
                   # tmp$TU_vs_Qth2 %>% arrange(desc(logFC)) %>% pull(SYMBOL) %>% head(20),
                   # tmp$TU_vs_Qth2 %>% arrange(desc(logFC)) %>% pull(SYMBOL) %>% tail(20)) %>% unique()
FCFCplot(vplot = vplot, genes_to_label =  genes_to_label, xlimits = c(1/5,5), ylimits = c(1/5,5) )
dev.off()

pdf(sprintf("%s/FCFC_Tips_vsQTh1Th17.pdf", out_dir), width = 30, 30, useDingbats = F)
# pdf(sprintf("%s/Heatmap_Th1Clusters_Top10Genes.pdf", out_dir), width = 3, 5, useDingbats = F)

vplot = data.frame(fc1 = 2^tmp$RS_vs_Qth1$logFC, fc2 = 2^(tmp$XY_vs_Qth17$logFC), SYMBOL = tmp$RS_vs_Qth1$SYMBOL)
genes_to_label = c(tmp$RS_vs_Qth1 %>% filter(abs(logFC) > 0.5,  P.Value < 10^-10) %>% pull(SYMBOL) %>% head(100),
                   tmp$XY_vs_Qth17 %>% filter(abs(logFC) > 0.5,  P.Value < 10^-10) %>% pull(SYMBOL) %>% head(100)) %>% unique()
                   #     
                   #     arrange(desc(logFC)) %>% pull(SYMBOL) %>% head(20),
                   # tmp$RS_vs_Qth1 %>% arrange(desc(logFC)) %>% pull(SYMBOL) %>% tail(20),
                   # tmp$TU_vs_Qth2 %>% arrange(desc(logFC)) %>% pull(SYMBOL) %>% head(20),
                   # tmp$TU_vs_Qth2 %>% arrange(desc(logFC)) %>% pull(SYMBOL) %>% tail(20)) %>% unique()
FCFCplot(vplot = vplot, genes_to_label =  genes_to_label, xlimits = c(1/5,5), ylimits = c(1/5,5) )
dev.off()

```

Genes present in at least 2 comparisons with FDR < 1% and log2FC > 0.5
```{r}
library(dplyr)
library(purrr)

tmp = list(
  RS_vs_Qth1 = ttlist2[["RS_vs_Qth1"]],
  TU_vs_Qth2 = ttlist2[["TU_vs_Qth2"]],
  XY_vs_Qth17 = ttlist2[["XY_vs_Qth17"]]
)

# combine all comparisons into one long table
tmp_long <- imap_dfr(tmp, ~{
  .x %>%
    select(SYMBOL, logFC, adj.P.Val) %>%
    mutate(comparison = .y)
})

# genes strongly up in at least one comparison
strong_genes_up <- tmp_long %>%
  filter(logFC > 0.5, adj.P.Val < 10^-2) %>%
  distinct(SYMBOL) %>%
  pull(SYMBOL)

# among those, require mild up in at least 2 comparisons
genes_filt_up <- tmp_long %>%
  filter(SYMBOL %in% strong_genes_up) %>%
  mutate(pass_mild = logFC > 0 & adj.P.Val < 10^-2) %>%
  group_by(SYMBOL) %>%
  summarise(
    n_mild = sum(pass_mild),
    .groups = "drop"
  ) %>%
  filter(n_mild >= 2) %>%
  pull(SYMBOL)

# genes strongly down in at least one comparison
strong_genes_down <- tmp_long %>%
  filter(logFC < -0.5, adj.P.Val < 10^-2) %>%
  distinct(SYMBOL) %>%
  pull(SYMBOL)

# among those, require mild up in at least 2 comparisons
genes_filt_down <- tmp_long %>%
  filter(SYMBOL %in% strong_genes_down) %>%
  mutate(pass_mild = logFC < 0 & adj.P.Val < 10^-2) %>%
  group_by(SYMBOL) %>%
  summarise(
    n_mild = sum(pass_mild),
    .groups = "drop"
  ) %>%
  filter(n_mild >= 2) %>%
  pull(SYMBOL)

genes_filt = c(genes_filt_up, genes_filt_down)
length(genes_filt)

genes_to_label <- genes_filt[
    !grepl(
        "^Gm|^mt-|^Rpl|^Rps|^Snor|^Mir|^Linc|^Rp[sl]|-ps|Rik$|^Hbb",
        genes_filt,
        ignore.case = TRUE
    )
] %>% unique() %>% sort()
genes_filt


mat_plot2 = mat_coef[genes_to_label, ]
range(mat_plot2)

library(ComplexHeatmap)
library(circlize)

col_fun <- circlize::colorRamp2(
    c(-1.5, 0, 1.5),
    c("#4575B4", "white", "#D73027")
)

pdf(sprintf("%s/Heatmap_Tips_vsQTh_SharedGenes.pdf", out_dir), width = 3, 6, useDingbats = F)
Heatmap(
    mat_plot2,
    name = "log2FC",
    col = col_fun,
    cluster_columns = T,
    cluster_rows = T,
    row_names_gp = gpar(fontsize = 6),
    column_names_gp = gpar(fontsize = 6),
    column_title = "Shared Th1 Th2 Th17 tips"
)
dev.off()



```



Th1
```{r}
# tmp = list(ttlist3[["RS_vs_Gth1"]], ttlist2[["RS_vs_Qth1"]], ttlist1[["R_vs_S"]])
# names(tmp) = c( "RS_vs_Gth1", "RS_vs_Qth1", "R_vs_S")
tmp = list(ttlist3[["RS_vs_Gth1"]], ttlist2[["RS_vs_Qth1"]], ttlist1[["R_vs_S"]])
names(tmp) = c( "RS_vs_Gth1", "RS_vs_Qth1", "R_vs_S")
tmp[["S_vs_R"]] = tmp[["R_vs_S"]]
tmp[["S_vs_R"]]$logFC = -tmp[["R_vs_S"]]$logFC
# tmp[["R_vs_S"]] = NULL

tmp[["Gth1_vs_RS"]] = tmp[["RS_vs_Gth1"]]
tmp[["Gth1_vs_RS"]]$logFC = -tmp[["RS_vs_Gth1"]]$logFC
tmp[["RS_vs_Gth1"]] = NULL

tmp[["Qth1_vs_RS"]] = tmp[["RS_vs_Qth1"]]
tmp[["Qth1_vs_RS"]]$logFC = -tmp[["RS_vs_Qth1"]]$logFC
tmp[["RS_vs_Qth1"]] = NULL

tmp = tmp[c("Gth1_vs_RS", "Qth1_vs_RS","R_vs_S", "S_vs_R")]

tmp_list = list()
tmp_list[["Th1"]] = tmp

names(tmp) 
gene_sig = list()

cols = c("grey", "grey", "grey", "grey" )
names(cols) = names(tmp)

# pdf(sprintf("%s/Volcano_Th1_clusters.pdf", out_dir), 15, 15, useDingbats = F)
for (i in names(tmp)) {
    print(i)
    vplot = na.omit(data.frame(SYMBOL = tmp[[i]]$SYMBOL, fc = 2^tmp[[i]]$logFC, pval = tmp[[i]]$P.Value + 10^-80, AveExpr = 2^(tmp[[i]]$AveExpr))) #dealing with NA
    # vplot = data.frame(SYMBOL = tmp[[i]]$SYMBOL, fc = 2^tmp[[i]]$logFC, pval = tmp[[i]]$P.Value+10^-300, AveExpr = 2^(tmp[[i]]$AveExpr))
    p = Vplot(vplot = vplot, xlab = names(tmp)[i], xlimits = c(min(vplot$fc[vplot$fc != 0]), max(vplot$fc)), ylimits = c(min(vplot$pval), 1),size = 2, color = cols[i])  #c(min(vplot$pval[vplot$pval != 0]),1)
    sub = vplot %>%
        filter(abs(log2(fc)) > 0 & pval < 10^-5) %>%
        arrange(desc(log2(fc))) %>%
        pull(SYMBOL)
    sub
    # gene_sig[[sprintf("%s_up", i)]] = sub
    highlight_genes = sub #unique(c(head(sub, 25), tail(sub, 25)))
    # sub = (log2(vplot$fc) > 0.5 | log2(vplot$fc) < -0.5)  & vplot$pval < 0.05
    # sub = vplot[sub,] %>% arrange(desc(log2(fc)))
    # sub1 = sub %>% pull(SYMBOL) %>% head(25) #vplot$SYMBOL %in% as.character(sub2$SYMBOL)[1:200]
    # sub2 = sub %>% pull(SYMBOL) %>% tail(25)
    highlight_filter = vplot$SYMBOL %in% highlight_genes
    # sub = vplot$SYMBOL %in% c(sub1, sub2)#as.character(sub2$SYMBOL)[1:50]
    library(ggrepel)
    # print(p)
    print(p + geom_text_repel(data = vplot[highlight_filter,], aes(x = fc, y = pval, label = SYMBOL)) + ggtitle(i) + xlab(sprintf("FC %s", i)))
}
# dev.off()

comp_filt = names(tmp)

genes_to_label <- c()
for (i in comp_filt) {
    df <- tmp[[i]] #%>% filter(adj.P.Val < 0.05)
    # top up
    up_genes <- df %>% filter(logFC > 0.5,  P.Value < 10^-10) %>% 
        arrange(desc(logFC)) %>%
        pull(SYMBOL)
    # top down
    down_genes <- df %>% filter(logFC < -0.5,  P.Value < 10^-10) %>% 
        arrange(logFC) %>%
        pull(SYMBOL)
    # remove genes already selected
    # up_genes <- setdiff(up_genes, genes_to_label)
    # down_genes <- setdiff(down_genes, genes_to_label)
    
    # add top 20 each
    genes_to_label <- c(
        genes_to_label,
        head(up_genes, 10), head(down_genes, 10)
        # head(up_genes, 100), head(down_genes, 100)
    )
}

genes_to_label <- genes_to_label[
    !grepl(
        "^Gm|^mt-|^Rpl|^Rps|^Snor|^Mir|^Linc|^Rp[sl]|-ps|Rik$|^Hbb",
        genes_to_label,
        ignore.case = TRUE
    )
] %>% unique()

# genes_to_label_top100 = list()
# genes_to_label_top100[["Th1"]] = genes_to_label

genes_to_label_top10 = list()
genes_to_label_top10[["Th1"]] = genes_to_label

#plot heatmap
# mat_coef <- sapply(tmp, function(tmp) tmp$logFC)
lapply(tmp, function(x) all(x$SYMBOL == tmp[[1]]$SYMBOL)) ##check that rownames match!!!

common_genes <- Reduce(intersect, lapply(tmp, function(x) x$SYMBOL))
tmp <- lapply(tmp, function(x) {
  x %>%
    dplyr::filter(SYMBOL %in% common_genes) %>%
    dplyr::arrange(match(SYMBOL, common_genes))
})
lapply(tmp, function(x) identical(x$SYMBOL, tmp[[1]]$SYMBOL))

mat_coef <- Reduce(cbind, lapply(tmp, function(x) x[,"logFC"]))
rownames(mat_coef) = tmp[[1]]$SYMBOL
colnames(mat_coef) = make.names(names(tmp))
mat_plot = mat_coef[genes_to_label, ]
range(mat_plot)

# tmp[["R_vs_S"]][genes_to_label,1]
# mat_coef[genes_to_label,"R_vs_S"]

library(ComplexHeatmap)
library(circlize)

col_fun <- circlize::colorRamp2(
    c(-3, 0, 3),
    c("#4575B4", "white", "#D73027")
)

# pdf(sprintf("%s/heatmap_memory_vs_acute_SLO_LCMV.pdf", out_dir), width = 3, 8, useDingbats = F)
pdf(sprintf("%s/Heatmap_Th1Clusters_Top100Genes.pdf", out_dir), width = 3, 25, useDingbats = F)
# pdf(sprintf("%s/Heatmap_Th1Clusters_Top10Genes.pdf", out_dir), width = 3, 5, useDingbats = F)
Heatmap(
    mat_plot,
    name = "log2FC",
    col = col_fun,
    cluster_columns = FALSE,
    cluster_rows = T,
    row_names_gp = gpar(fontsize = 6),
    column_names_gp = gpar(fontsize = 6),
    column_title = "Th1 clusters"
)
dev.off()



```

Th2
```{r}
tmp = list(ttlist3[["TU_vs_Gth2"]], ttlist2[["TU_vs_Qth2"]], ttlist1[["T_vs_U"]])
names(tmp) = c( "TU_vs_Gth2", "TU_vs_Qth2", "T_vs_U")
tmp[["U_vs_T"]] = tmp[["T_vs_U"]]
tmp[["U_vs_T"]]$logFC = -tmp[["T_vs_U"]]$logFC
# tmp[["T_vs_U"]] = NULL

tmp[["Gth2_vs_TU"]] = tmp[["TU_vs_Gth2"]]
tmp[["Gth2_vs_TU"]]$logFC = -tmp[["TU_vs_Gth2"]]$logFC
tmp[["TU_vs_Gth2"]] = NULL

tmp[["Qth2_vs_TU"]] = tmp[["TU_vs_Qth2"]]
tmp[["Qth2_vs_TU"]]$logFC = -tmp[["TU_vs_Qth2"]]$logFC
tmp[["TU_vs_Qth2"]] = NULL

tmp = tmp[c("Gth2_vs_TU", "Qth2_vs_TU","T_vs_U", "U_vs_T")]

tmp_list[["Th2"]] = tmp

names(tmp) 
gene_sig = list()

cols = c("grey", "grey", "grey", "grey" )
names(cols) = names(tmp)

# pdf(sprintf("%s/Volcano_Th2_clusters.pdf", out_dir), 15, 15, useDingbats = F)
for (i in names(tmp)) {
    print(i)
    vplot = na.omit(data.frame(SYMBOL = tmp[[i]]$SYMBOL, fc = 2^tmp[[i]]$logFC, pval = tmp[[i]]$P.Value + 10^-80, AveExpr = 2^(tmp[[i]]$AveExpr))) #dealing with NA
    # vplot = data.frame(SYMBOL = tmp[[i]]$SYMBOL, fc = 2^tmp[[i]]$logFC, pval = tmp[[i]]$P.Value+10^-300, AveExpr = 2^(tmp[[i]]$AveExpr))
    p = Vplot(vplot = vplot, xlab = names(tmp)[i], xlimits = c(min(vplot$fc[vplot$fc != 0]), max(vplot$fc)), ylimits = c(min(vplot$pval), 1),size = 2, color = cols[i])  #c(min(vplot$pval[vplot$pval != 0]),1)
    sub = vplot %>%
        filter(abs(log2(fc)) > 0 & pval < 10^-5) %>%
        arrange(desc(log2(fc))) %>%
        pull(SYMBOL)
    sub
    # gene_sig[[sprintf("%s_up", i)]] = sub
    highlight_genes = sub #unique(c(head(sub, 25), tail(sub, 25)))
    # sub = (log2(vplot$fc) > 0.5 | log2(vplot$fc) < -0.5)  & vplot$pval < 0.05
    # sub = vplot[sub,] %>% arrange(desc(log2(fc)))
    # sub1 = sub %>% pull(SYMBOL) %>% head(25) #vplot$SYMBOL %in% as.character(sub2$SYMBOL)[1:200]
    # sub2 = sub %>% pull(SYMBOL) %>% tail(25)
    highlight_filter = vplot$SYMBOL %in% highlight_genes
    # sub = vplot$SYMBOL %in% c(sub1, sub2)#as.character(sub2$SYMBOL)[1:50]
    library(ggrepel)
    # print(p)
    print(p + geom_text_repel(data = vplot[highlight_filter,], aes(x = fc, y = pval, label = SYMBOL)) + ggtitle(i) + xlab(sprintf("FC %s", i)))
}
# dev.off()

comp_filt = names(tmp)

genes_to_label <- c()
for (i in comp_filt) {
    df <- tmp[[i]] #%>% filter(adj.P.Val < 0.05)
    # top up
    up_genes <- df %>% filter(logFC > 0.5,  P.Value < 10^-10) %>% 
        arrange(desc(logFC)) %>%
        pull(SYMBOL)
    # top down
    down_genes <- df %>% filter(logFC < -0.5,  P.Value < 10^-10) %>% 
        arrange(logFC) %>%
        pull(SYMBOL)
    # remove genes already selected
    # up_genes <- setdiff(up_genes, genes_to_label)
    # down_genes <- setdiff(down_genes, genes_to_label)
    
    # add top 20 each
    genes_to_label <- c(
        genes_to_label,
        head(up_genes, 10), head(down_genes, 10)
        # head(up_genes, 100), head(down_genes, 100)
    )
}

genes_to_label <- genes_to_label[
    !grepl(
        "^Gm|^mt-|^Tpl|^Rps|^Snor|^Mir|^Linc|^Rp[sl]|-ps|Rik$|^Hbb",
        genes_to_label,
        ignore.case = TRUE
    )
] %>% unique()

genes_to_label_top10[["Th2"]] = genes_to_label
# genes_to_label_top100[["Th2"]] = genes_to_label

#plot heatmap
# mat_coef <- sapply(tmp, function(tmp) tmp$logFC)
lapply(tmp, function(x) all(x$SYMBOL == tmp[[1]]$SYMBOL)) ##check that rownames match!!!

common_genes <- Reduce(intersect, lapply(tmp, function(x) x$SYMBOL))
tmp <- lapply(tmp, function(x) {
  x %>%
    dplyr::filter(SYMBOL %in% common_genes) %>%
    dplyr::arrange(match(SYMBOL, common_genes))
})
lapply(tmp, function(x) identical(x$SYMBOL, tmp[[1]]$SYMBOL))

mat_coef <- Reduce(cbind, lapply(tmp, function(x) x[,"logFC"]))
rownames(mat_coef) = tmp[[1]]$SYMBOL
colnames(mat_coef) = make.names(names(tmp))
mat_plot = mat_coef[genes_to_label, ]
range(mat_plot)

# tmp[["R_vs_S"]][genes_to_label,1]
# mat_coef[genes_to_label,"R_vs_S"]

library(ComplexHeatmap)
library(circlize)

col_fun <- circlize::colorRamp2(
    c(-2, 0, 2),
    c("#4575B4", "white", "#D73027")
)

# pdf(sprintf("%s/Heatmap_Th2Clusters_Top100Genes.pdf", out_dir), width = 3, 8, useDingbats = F)
pdf(sprintf("%s/Heatmap_Th2Clusters_Top10Genes.pdf", out_dir), width = 3, 4, useDingbats = F)
Heatmap(
    mat_plot,
    name = "log2FC",
    col = col_fun,
    cluster_columns = FALSE,
    cluster_rows = T,
    row_names_gp = gpar(fontsize = 6),
    column_names_gp = gpar(fontsize = 6),
    column_title = "Th2 clusters"
)
dev.off()

```

Th17
```{r}
tmp = list(ttlist3[["XY_vs_Gth17"]], ttlist2[["XY_vs_Qth17"]], ttlist1[["X_vs_Y"]])
names(tmp) = c( "XY_vs_Gth17", "XY_vs_Qth17", "X_vs_Y")
tmp[["Y_vs_X"]] = tmp[["X_vs_Y"]]
tmp[["Y_vs_X"]]$logFC = -tmp[["X_vs_Y"]]$logFC
# tmp[["X_vs_Y"]] = NULL

tmp[["Gth17_vs_XY"]] = tmp[["XY_vs_Gth17"]]
tmp[["Gth17_vs_XY"]]$logFC = -tmp[["XY_vs_Gth17"]]$logFC
tmp[["XY_vs_Gth17"]] = NULL

tmp[["Qth17_vs_XY"]] = tmp[["XY_vs_Qth17"]]
tmp[["Qth17_vs_XY"]]$logFC = -tmp[["XY_vs_Qth17"]]$logFC
tmp[["XY_vs_Qth17"]] = NULL

tmp = tmp[c("Gth17_vs_XY", "Qth17_vs_XY","X_vs_Y", "Y_vs_X")]

tmp_list[["Th17"]] = tmp

names(tmp) 
gene_sig = list()

cols = c("grey", "grey", "grey", "grey" )
names(cols) = names(tmp)

# pdf(sprintf("%s/Volcano_Th17_clusters.pdf", out_dir), 15, 15, useDingbats = F)
for (i in names(tmp)) {
    print(i)
    vplot = na.omit(data.frame(SYMBOL = tmp[[i]]$SYMBOL, fc = 2^tmp[[i]]$logFC, pval = tmp[[i]]$P.Value + 10^-80, AveExpr = 2^(tmp[[i]]$AveExpr))) #dealing with NA
    # vplot = data.frame(SYMBOL = tmp[[i]]$SYMBOL, fc = 2^tmp[[i]]$logFC, pval = tmp[[i]]$P.Value+10^-300, AveExpr = 2^(tmp[[i]]$AveExpr))
    p = Vplot(vplot = vplot, xlab = names(tmp)[i], xlimits = c(min(vplot$fc[vplot$fc != 0]), max(vplot$fc)), ylimits = c(min(vplot$pval), 1),size = 2, color = cols[i])  #c(min(vplot$pval[vplot$pval != 0]),1)
    sub = vplot %>%
        filter(abs(log2(fc)) > 0 & pval < 10^-5) %>%
        arrange(desc(log2(fc))) %>%
        pull(SYMBOL)
    sub
    # gene_sig[[sprintf("%s_up", i)]] = sub
    highlight_genes = sub #unique(c(head(sub, 25), tail(sub, 25)))
    # sub = (log2(vplot$fc) > 0.5 | log2(vplot$fc) < -0.5)  & vplot$pval < 0.05
    # sub = vplot[sub,] %>% arrange(desc(log2(fc)))
    # sub1 = sub %>% pull(SYMBOL) %>% head(25) #vplot$SYMBOL %in% as.character(sub2$SYMBOL)[1:200]
    # sub2 = sub %>% pull(SYMBOL) %>% tail(25)
    highlight_filter = vplot$SYMBOL %in% highlight_genes
    # sub = vplot$SYMBOL %in% c(sub1, sub2)#as.character(sub2$SYMBOL)[1:50]
    library(ggrepel)
    # print(p)
    print(p + geom_text_repel(data = vplot[highlight_filter,], aes(x = fc, y = pval, label = SYMBOL)) + ggtitle(i) + xlab(sprintf("FC %s", i)))
}
# dev.off()

comp_filt = names(tmp)

genes_to_label <- c()
for (i in comp_filt) {
    df <- tmp[[i]] #%>% filter(adj.P.Val < 0.05)
    # top up
    up_genes <- df %>% filter(logFC > 0.5,  P.Value < 10^-10) %>% 
        arrange(desc(logFC)) %>%
        pull(SYMBOL)
    # top down
    down_genes <- df %>% filter(logFC < -0.5,  P.Value < 10^-10) %>% 
        arrange(logFC) %>%
        pull(SYMBOL)
    # remove genes already selected
    # up_genes <- setdiff(up_genes, genes_to_label)
    # down_genes <- setdiff(down_genes, genes_to_label)
    
    # add top 20 each
    genes_to_label <- c(
        genes_to_label,
        head(up_genes, 10), head(down_genes, 10)
        # head(up_genes, 100), head(down_genes, 100)
    )
}

genes_to_label <- genes_to_label[
    !grepl(
        "^Gm|^mt-|^Tpl|^Rps|^Snor|^Mir|^Linc|^Rp[sl]|-ps|Rik$|^Hbb",
        genes_to_label,
        ignore.case = TRUE
    )
] %>% unique()

genes_to_label_top10[["Th17"]] = genes_to_label
# genes_to_label_top100[["Th17"]] = genes_to_label

#plot heatmap
# mat_coef <- sapply(tmp, function(tmp) tmp$logFC)
lapply(tmp, function(x) all(x$SYMBOL == tmp[[1]]$SYMBOL)) ##check that rownames match!!!

common_genes <- Reduce(intersect, lapply(tmp, function(x) x$SYMBOL))
tmp <- lapply(tmp, function(x) {
  x %>%
    dplyr::filter(SYMBOL %in% common_genes) %>%
    dplyr::arrange(match(SYMBOL, common_genes))
})
lapply(tmp, function(x) identical(x$SYMBOL, tmp[[1]]$SYMBOL))

mat_coef <- Reduce(cbind, lapply(tmp, function(x) x[,"logFC"]))
rownames(mat_coef) = tmp[[1]]$SYMBOL
colnames(mat_coef) = make.names(names(tmp))
mat_plot = mat_coef[genes_to_label, ]
range(mat_plot)

# tmp[["R_vs_S"]][genes_to_label,1]
# mat_coef[genes_to_label,"R_vs_S"]

library(ComplexHeatmap)
library(circlize)

col_fun <- circlize::colorRamp2(
    c(-3, 0, 3),
    c("#4575B4", "white", "#D73027")
)

# pdf(sprintf("%s/Heatmap_Th17Clusters_Top100Genes.pdf", out_dir), width = 3, 15, useDingbats = F)
pdf(sprintf("%s/Heatmap_Th17Clusters_Top10Genes.pdf", out_dir), width = 3, 6, useDingbats = F)
Heatmap(
    mat_plot,
    name = "log2FC",
    col = col_fun,
    cluster_columns = FALSE,
    cluster_rows = T,
    row_names_gp = gpar(fontsize = 6),
    column_names_gp = gpar(fontsize = 6),
    column_title = "Th17 clusters"
)
dev.off()



```

G and Q signature heatmap
```{r}
genes_to_label_top100
genes_to_label0 = genes_to_label_top100 %>% unlist() %>% unique()
# genes_to_label0 = genes_to_label_top10 %>% unlist() %>% unique()

tmp_all = c(tmp_list[[1]], tmp_list[[2]], tmp_list[[3]])
tmp = tmp_all
names(tmp)
tmp = tmp[c("Gth1_vs_RS", "Gth2_vs_TU", "Gth17_vs_XY", "Qth1_vs_RS", "Qth2_vs_TU", "Qth17_vs_XY")]

lapply(tmp, function(x) all(x$SYMBOL == tmp[[1]]$SYMBOL)) ##check that rownames match!!!

common_genes <- Reduce(intersect, lapply(tmp, function(x) x$SYMBOL))
tmp <- lapply(tmp, function(x) {
  x %>%
    dplyr::filter(SYMBOL %in% common_genes) %>%
    dplyr::arrange(match(SYMBOL, common_genes))
})
lapply(tmp, function(x) identical(x$SYMBOL, tmp[[1]]$SYMBOL))

mat_coef <- Reduce(cbind, lapply(tmp, function(x) x[,"logFC"]))
rownames(mat_coef) = tmp[[1]]$SYMBOL
colnames(mat_coef) = make.names(names(tmp))
mat_plot = mat_coef[genes_to_label0, ]
range(mat_plot)

# tmp[["R_vs_S"]][genes_to_label,1]
# mat_coef[genes_to_label,"R_vs_S"]

library(ComplexHeatmap)
library(circlize)

col_fun <- circlize::colorRamp2(
    c(-2, 0, 2),
    c("#4575B4", "white", "#D73027")
)

# pdf(sprintf("%s/Heatmap_G_Q_Clusters_Top10Genes.pdf", out_dir), width = 3, 4, useDingbats = F)
# pdf(sprintf("%s/Heatmap_G_Q_Clusters_Top100Genes.pdf", out_dir), width = 3, 25, useDingbats = F)
pdf(sprintf("%s/Heatmap_Th2Clusters_Top10Genes.pdf", out_dir), width = 3, 8, useDingbats = F)
Heatmap(
    mat_plot,
    name = "log2FC",
    col = col_fun,
    cluster_columns = F,
    cluster_rows = T,
    row_names_gp = gpar(fontsize = 6),
    column_names_gp = gpar(fontsize = 6),
    column_title = "Q and G sig"
)
dev.off()


#Filter genes just for G and Q
mat_plot2 = mat_coef[genes_to_label0, ]
genes_to_label = c(genes_to_label0[rowSums(mat_plot2[,c("Gth1_vs_RS", "Gth2_vs_TU", "Gth17_vs_XY")] > 0) ==3],
                   # genes_to_label[rowSums(mat_plot2[,c("Gth1_vs_RS", "Gth2_vs_TU", "Gth17_vs_XY")] < 0) ==3],
                   genes_to_label0[rowSums(mat_plot2[,c("Qth1_vs_RS", "Qth2_vs_TU", "Qth17_vs_XY")] > 0) ==3]
                   # genes_to_label[rowSums(mat_plot2[,c("Qth1_vs_RS", "Qth2_vs_TU", "Qth17_vs_XY")] < -0) ==3],
                   #genes_to_label0[rowSums(mat_plot2[,c("Qth1_vs_RS", "Qth2_vs_TU", "Qth17_vs_XY", "Gth1_vs_RS", "Gth2_vs_TU", "Gth17_vs_XY")] < -0.5) ==3]
                   ) %>% unique()

mat_plot = mat_coef[genes_to_label, ]
range(mat_plot)

# tmp[["R_vs_S"]][genes_to_label,1]
# mat_coef[genes_to_label,"R_vs_S"]

library(ComplexHeatmap)
library(circlize)

col_fun <- circlize::colorRamp2(
    c(-2.5, 0, 1.5),
    c("#4575B4", "white", "#D73027")
)

pdf(sprintf("%s/Heatmap_G_Q_SharedSig.pdf", out_dir), width = 5, 5, useDingbats = F)
Heatmap(
    mat_plot,
    name = "log2FC",
    col = col_fun,
    cluster_columns = T,
    cluster_rows = T,
    row_names_gp = gpar(fontsize = 6),
    column_names_gp = gpar(fontsize = 6),
    column_title = "G and Q signatures"
)
dev.off()

```

Q only signature heatmap
```{r}
genes_to_label_top100
genes_to_label0 = genes_to_label_top10 %>% unlist() %>% unique()

tmp_all = c(tmp_list[[1]], tmp_list[[2]], tmp_list[[3]])
tmp = tmp_all
names(tmp)
tmp = tmp[c("Qth1_vs_RS", "Qth2_vs_TU", "Qth17_vs_XY")]

lapply(tmp, function(x) all(x$SYMBOL == tmp[[1]]$SYMBOL)) ##check that rownames match!!!

common_genes <- Reduce(intersect, lapply(tmp, function(x) x$SYMBOL))
tmp <- lapply(tmp, function(x) {
  x %>%
    dplyr::filter(SYMBOL %in% common_genes) %>%
    dplyr::arrange(match(SYMBOL, common_genes))
})
lapply(tmp, function(x) identical(x$SYMBOL, tmp[[1]]$SYMBOL))

mat_coef <- Reduce(cbind, lapply(tmp, function(x) x[,"logFC"]))
rownames(mat_coef) = tmp[[1]]$SYMBOL
colnames(mat_coef) = make.names(names(tmp))
mat_plot = mat_coef[genes_to_label0, ]
range(mat_plot)

# tmp[["R_vs_S"]][genes_to_label,1]
# mat_coef[genes_to_label,"R_vs_S"]

library(ComplexHeatmap)
library(circlize)

col_fun <- circlize::colorRamp2(
    c(-2, 0, 2),
    c("#4575B4", "white", "#D73027")
)

# pdf(sprintf("%s/Heatmap_G_Q_Clusters_Top10Genes.pdf", out_dir), width = 3, 4, useDingbats = F)
# pdf(sprintf("%s/Heatmap_G_Q_Clusters_Top100Genes.pdf", out_dir), width = 3, 20, useDingbats = F)
pdf(sprintf("%s/Heatmap_Q_Clusters_Top10Genes2.pdf", out_dir), width = 3, 8, useDingbats = F)
# pdf(sprintf("%s/Heatmap_Th2Clusters_Top10Genes.pdf", out_dir), width = 3, 4, useDingbats = F)
mat_plot = -mat_plot
colnames(mat_plot)
colnames(mat_plot) = c("RS_vs_Qth1", "TU_vsQth2", "XU_vsQTh17")
Heatmap(
    mat_plot,
    name = "log2FC",
    col = col_fun,
    cluster_columns = F,
    cluster_rows = T,
    row_names_gp = gpar(fontsize = 6),
    column_names_gp = gpar(fontsize = 6),
    column_title = "Q.Th vs their tips"
    
)
dev.off()



```

G only signature heatmap
```{r}
genes_to_label_top100
genes_to_label0 = genes_to_label_top10 %>% unlist() %>% unique()

tmp_all = c(tmp_list[[1]], tmp_list[[2]], tmp_list[[3]])
tmp = tmp_all
names(tmp)
tmp = tmp[c("Gth1_vs_RS", "Gth2_vs_TU", "Gth17_vs_XY")]

lapply(tmp, function(x) all(x$SYMBOL == tmp[[1]]$SYMBOL)) ##check that rownames match!!!

common_genes <- Reduce(intersect, lapply(tmp, function(x) x$SYMBOL))
tmp <- lapply(tmp, function(x) {
  x %>%
    dplyr::filter(SYMBOL %in% common_genes) %>%
    dplyr::arrange(match(SYMBOL, common_genes))
})
lapply(tmp, function(x) identical(x$SYMBOL, tmp[[1]]$SYMBOL))

mat_coef <- Reduce(cbind, lapply(tmp, function(x) x[,"logFC"]))
rownames(mat_coef) = tmp[[1]]$SYMBOL
colnames(mat_coef) = make.names(names(tmp))
mat_plot = mat_coef[genes_to_label0, ]
range(mat_plot)

# tmp[["R_vs_S"]][genes_to_label,1]
# mat_coef[genes_to_label,"R_vs_S"]

library(ComplexHeatmap)
library(circlize)

col_fun <- circlize::colorRamp2(
    c(-2, 0, 2),
    c("#4575B4", "white", "#D73027")
)

# pdf(sprintf("%s/Heatmap_G_Q_Clusters_Top10Genes.pdf", out_dir), width = 3, 4, useDingbats = F)
# pdf(sprintf("%s/Heatmap_G_Q_Clusters_Top100Genes.pdf", out_dir), width = 3, 20, useDingbats = F)
pdf(sprintf("%s/Heatmap_G_Clusters_Top10Genes2.pdf", out_dir), width = 3, 8, useDingbats = F)
# pdf(sprintf("%s/Heatmap_Th2Clusters_Top10Genes.pdf", out_dir), width = 3, 4, useDingbats = F)
Heatmap(
    mat_plot,
    name = "log2FC",
    col = col_fun,
    cluster_columns = F,
    cluster_rows = T,
    row_names_gp = gpar(fontsize = 6),
    column_names_gp = gpar(fontsize = 6),
    column_title = "G.Th vs their tips"
)
dev.off()

```


Reccurrent genes across comparisons?
```{r}
genes_to_label_top100
genes_to_label0 = genes_to_label_top100 %>% unlist() %>% unique()

tmp_all = c(tmp_list[[1]], tmp_list[[2]], tmp_list[[3]])
tmp = tmp_all
names(tmp)

lapply(tmp, function(x) all(x$SYMBOL == tmp[[1]]$SYMBOL)) ##check that rownames match!!!

common_genes <- Reduce(intersect, lapply(tmp, function(x) x$SYMBOL))
tmp <- lapply(tmp, function(x) {
  x %>%
    dplyr::filter(SYMBOL %in% common_genes) %>%
    dplyr::arrange(match(SYMBOL, common_genes))
})
lapply(tmp, function(x) identical(x$SYMBOL, tmp2[[1]]$SYMBOL))

mat_coef <- Reduce(cbind, lapply(tmp, function(x) x[,"logFC"]))
rownames(mat_coef) = tmp[[1]]$SYMBOL
colnames(mat_coef) = make.names(names(tmp))
mat_plot = mat_coef[genes_to_label, ]
range(mat_plot)

# tmp[["R_vs_S"]][genes_to_label,1]
# mat_coef[genes_to_label,"R_vs_S"]

library(ComplexHeatmap)
library(circlize)

col_fun <- circlize::colorRamp2(
    c(-3, 0, 3),
    c("#4575B4", "white", "#D73027")
)

pdf(sprintf("%s/Heatmap_Th1Th2Th117Clusters_Top100Genes.pdf", out_dir), width = 5, 40, useDingbats = F)
# pdf(sprintf("%s/Heatmap_Th2Clusters_Top10Genes.pdf", out_dir), width = 3, 4, useDingbats = F)
Heatmap(
    mat_plot,
    name = "log2FC",
    col = col_fun,
    cluster_columns = FALSE,
    cluster_rows = T,
    row_names_gp = gpar(fontsize = 6),
    column_names_gp = gpar(fontsize = 6),
    column_title = "Th1 Th2 Th17 clusters"
)
dev.off()


#Filter genes
mat_plot2 = mat_coef[genes_to_label0, ]
genes_to_label = c(genes_to_label0[rowSums(mat_plot2[,c("Gth1_vs_RS", "Gth2_vs_TU", "Gth17_vs_XY")] > 0) ==3],
                   # genes_to_label[rowSums(mat_plot2[,c("Gth1_vs_RS", "Gth2_vs_TU", "Gth17_vs_XY")] < 0) ==3],
                   genes_to_label0[rowSums(mat_plot2[,c("Qth1_vs_RS", "Qth2_vs_TU", "Qth17_vs_XY")] > 0) ==3],
                   # genes_to_label[rowSums(mat_plot2[,c("Qth1_vs_RS", "Qth2_vs_TU", "Qth17_vs_XY")] < -0) ==3],
                   genes_to_label0[rowSums(mat_plot2[,c("Qth1_vs_RS", "Qth2_vs_TU", "Qth17_vs_XY", "Gth1_vs_RS", "Gth2_vs_TU", "Gth17_vs_XY")] < -0.5) >=3]) %>% unique()

mat_plot = mat_coef[genes_to_label, ]
range(mat_plot)

write.table(x = genes_to_label, file = sprintf("%s/genesig_fig3l.txt", out_dir), quote = F, row.names = F, col.names = F)

# tmp[["R_vs_S"]][genes_to_label,1]
# mat_coef[genes_to_label,"R_vs_S"]

library(ComplexHeatmap)
library(circlize)

col_fun <- circlize::colorRamp2(
    c(-2, 0, 2),
    c("#4575B4", "white", "#D73027")
)

pdf(sprintf("%s/Heatmap_Th1Th2Th117Clusters_CommonGQTips.pdf", out_dir), width = 5, 8, useDingbats = F)
Heatmap(
    mat_plot,
    name = "log2FC",
    col = col_fun,
    cluster_columns = T,
    cluster_rows = T,
    row_names_gp = gpar(fontsize = 6),
    column_names_gp = gpar(fontsize = 6),
    column_title = "Th1 Th2 Th17 clusters"
)
dev.off()

#Filter genes just for G and Q
mat_plot2 = mat_coef[genes_to_label0, ]
genes_to_label = c(genes_to_label0[rowSums(mat_plot2[,c("Gth1_vs_RS", "Gth2_vs_TU", "Gth17_vs_XY")] > -0) ==3],
                   # genes_to_label[rowSums(mat_plot2[,c("Gth1_vs_RS", "Gth2_vs_TU", "Gth17_vs_XY")] < 0) ==3],
                   genes_to_label0[rowSums(mat_plot2[,c("Qth1_vs_RS", "Qth2_vs_TU", "Qth17_vs_XY")] > 0) ==3]
                   # genes_to_label[rowSums(mat_plot2[,c("Qth1_vs_RS", "Qth2_vs_TU", "Qth17_vs_XY")] < -0) ==3],
                   #genes_to_label0[rowSums(mat_plot2[,c("Qth1_vs_RS", "Qth2_vs_TU", "Qth17_vs_XY", "Gth1_vs_RS", "Gth2_vs_TU", "Gth17_vs_XY")] < -0.5) ==3]
                   ) %>% unique()

mat_plot = mat_coef[genes_to_label, ]
range(mat_plot)

# tmp[["R_vs_S"]][genes_to_label,1]
# mat_coef[genes_to_label,"R_vs_S"]

library(ComplexHeatmap)
library(circlize)

col_fun <- circlize::colorRamp2(
    c(-3, 0, 3),
    c("#4575B4", "white", "#D73027")
)

pdf(sprintf("%s/Heatmap_Th1Th2Th117Clusters_CommonGQ.pdf", out_dir), width = 5, 5, useDingbats = F)
Heatmap(
    mat_plot,
    name = "log2FC",
    col = col_fun,
    cluster_columns = T,
    cluster_rows = T,
    row_names_gp = gpar(fontsize = 6),
    column_names_gp = gpar(fontsize = 6),
    column_title = "Th1 Th2 Th17 clusters"
)
dev.off()

```


```{r}
so_orig@meta.data %>% arrange(as.numeric(igt_number)) %>% count(igt_number,exp_name)

so_orig@meta.data %>% filter(IGT %in% c("IGT7", "IGT8", "IGT8"), annotation_level1 == "CD4") %>% count(annotation_level2, organ_simplified)

FindAllMarkers(so_orig[,], ident.1 = "CD4.G", ident.2 = "CD4.Q", group.by = "annotation_level3", test.use = "wilcox", logfc.threshold = 0.1, min.pct = 0.1)

```



**Heatmap with DGE between 2 tip clusters**

```{r}
ttlist1 = readRDS('/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/analysis/data_integration/IGT1_96/CD4/DGE_limma/20260318_tips/ttlist_tips.Rds')
ttlist1 = ttlist1[8:10]

tmp = ttlist1

names(tmp) 
gene_sig = list()

cols = c("darkorchid", "violetred2", "burlywood2" )
names(cols) = names(tmp)

pdf(sprintf("%s/Volcano_DGE_ClustersWithingEachTip.pdf", out_dir), 15, 15, useDingbats = F)
for (i in names(tmp)) {
    print(i)
    vplot = na.omit(data.frame(SYMBOL = tmp[[i]]$SYMBOL, fc = 2^tmp[[i]]$logFC, pval = tmp[[i]]$P.Value + 10^-80, AveExpr = 2^(tmp[[i]]$AveExpr))) #dealing with NA
    # vplot = data.frame(SYMBOL = tmp[[i]]$SYMBOL, fc = 2^tmp[[i]]$logFC, pval = tmp[[i]]$P.Value+10^-300, AveExpr = 2^(tmp[[i]]$AveExpr))
    p = Vplot(vplot = vplot, xlab = names(tmp)[i], xlimits = c(min(vplot$fc[vplot$fc != 0]), max(vplot$fc)), ylimits = c(min(vplot$pval), 1),size = 2, color = cols[i])  #c(min(vplot$pval[vplot$pval != 0]),1)
    sub = vplot %>%
        filter(abs(log2(fc)) > 0.1 & pval < 10^-5) %>%
        arrange(desc(log2(fc))) %>%
        pull(SYMBOL)
    sub
    # gene_sig[[sprintf("%s_up", i)]] = sub
    highlight_genes = sub #unique(c(head(sub, 25), tail(sub, 25)))
    # sub = (log2(vplot$fc) > 0.5 | log2(vplot$fc) < -0.5)  & vplot$pval < 0.05
    # sub = vplot[sub,] %>% arrange(desc(log2(fc)))
    # sub1 = sub %>% pull(SYMBOL) %>% head(25) #vplot$SYMBOL %in% as.character(sub2$SYMBOL)[1:200]
    # sub2 = sub %>% pull(SYMBOL) %>% tail(25)
    highlight_filter = vplot$SYMBOL %in% highlight_genes
    # sub = vplot$SYMBOL %in% c(sub1, sub2)#as.character(sub2$SYMBOL)[1:50]
    library(ggrepel)
    # print(p)
    print(p + geom_text_repel(data = vplot[highlight_filter,], aes(x = fc, y = pval, label = SYMBOL)) + ggtitle(i) + xlab(sprintf("FC %s", i)))
}
dev.off()

comp_filt = names(tmp)

genes_to_label <- c()
for (i in comp_filt) {
    df <- tmp[[i]] #%>% filter(adj.P.Val < 0.05)
    # top up
    up_genes <- df %>% filter(logFC > 0.1,  P.Value < 10^-5) %>% 
        arrange(desc(logFC)) %>%
        pull(SYMBOL) 
    # top down
    down_genes <- df %>% filter(logFC < -0.1,  P.Value < 10^-5) %>% 
        arrange(logFC) %>%
        pull(SYMBOL)
    # remove genes already selected
    # up_genes <- setdiff(up_genes, genes_to_label)
    # down_genes <- setdiff(down_genes, genes_to_label)
    
    # add top 20 each
    genes_to_label <- c(
        genes_to_label,
        head(up_genes, 10), head(down_genes, 10)
        # head(up_genes, 100), head(down_genes, 100)
    )
}

genes_to_label <- genes_to_label[
    !grepl(
        "^Gm|^mt-|^Tpl|^Rps|^Snor|^Mir|^Linc|^Rp[sl]|-ps|Rik$|^Hbb",
        genes_to_label,
        ignore.case = TRUE
    )
]

genes_to_label = unique(genes_to_label )

#plot heatmap
# mat_coef <- sapply(tmp, function(tmp) tmp$logFC)
lapply(tmp, function(x) all(x$SYMBOL == tmp[[1]]$SYMBOL)) ##check that rownames match!!!

common_genes <- Reduce(intersect, lapply(tmp, function(x) x$SYMBOL))
tmp <- lapply(tmp, function(x) {
  x %>%
    dplyr::filter(SYMBOL %in% common_genes) %>%
    dplyr::arrange(match(SYMBOL, common_genes))
})
lapply(tmp, function(x) identical(x$SYMBOL, tmp[[1]]$SYMBOL))

mat_coef <- Reduce(cbind, lapply(tmp, function(x) x[,"logFC"]))
rownames(mat_coef) = tmp[[1]]$SYMBOL
colnames(mat_coef) = make.names(names(tmp))
mat_plot = mat_coef[genes_to_label, ]
range(mat_plot)

# tmp[["R_vs_S"]][genes_to_label,1]
# mat_coef[genes_to_label,"R_vs_S"]

library(ComplexHeatmap)
library(circlize)

col_fun <- circlize::colorRamp2(
    c(-3, 0, 3),
    c("#4575B4", "white", "#D73027")
)

pdf(sprintf("%s/Heatmap_DGE_ClustersWithingEachTipTop10genes.pdf", out_dir), width = 5, 5, useDingbats = F)
# pdf(sprintf("%s/Heatmap_Th17Clusters_Top10Genes.pdf", out_dir), width = 3, 6, useDingbats = F)
Heatmap(
    mat_plot,
    name = "log2FC",
    col = col_fun,
    cluster_columns = FALSE,
    cluster_rows = T,
    row_names_gp = gpar(fontsize = 6),
    column_names_gp = gpar(fontsize = 6),
    column_title = "Top 10 genes between clusters in each tip"
)
dev.off()

```

```{r}
ttlist1 = readRDS('/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/analysis/data_integration/IGT1_96/CD4/DGE_limma/20260318_tips/ttlist_tips.Rds')
tmp = ttlist1[8:10]
names(tmp)
lapply(tmp, function(x) all(x$SYMBOL == tmp[[1]]$SYMBOL)) ##check that rownames match!!!
# common_genes <- Reduce(intersect, lapply(tmp, function(x) x$SYMBOL))
# tmp <- lapply(tmp, function(x) {
#   x %>%
#     dplyr::filter(SYMBOL %in% common_genes) %>%
#     dplyr::arrange(match(SYMBOL, common_genes))
# })
# lapply(tmp, function(x) identical(x$SYMBOL, tmp[[1]]$SYMBOL))

vplot = data.frame(fc1 = 2^tmp$R_vs_S$logFC, fc2 = 2^(-tmp$T_vs_U$logFC), SYMBOL = tmp$R_vs_S$SYMBOL)
genes_to_label = c(tmp$R_vs_S %>% arrange(desc(logFC)) %>% pull(SYMBOL) %>% head(20),
                   tmp$R_vs_S %>% arrange(desc(logFC)) %>% pull(SYMBOL) %>% tail(20),
                   tmp$T_vs_U %>% arrange(desc(logFC)) %>% pull(SYMBOL) %>% head(20),
                   tmp$T_vs_U %>% arrange(desc(logFC)) %>% pull(SYMBOL) %>% tail(20)) %>% unique()
FCFCplot(vplot = vplot, genes_to_label =  genes_to_label, xlimits = c(1/5,5), ylimits = c(1/5,5) )

vplot = data.frame(fc1 = 2^tmp$R_vs_S$logFC, fc2 = 2^(-tmp$X_vs_Y$logFC), SYMBOL = tmp$R_vs_S$SYMBOL)
genes_to_label = c(tmp$R_vs_S %>% arrange(desc(logFC)) %>% pull(SYMBOL) %>% head(20),
                   tmp$R_vs_S %>% arrange(desc(logFC)) %>% pull(SYMBOL) %>% tail(20),
                   tmp$X_vs_Y %>% arrange(desc(logFC)) %>% pull(SYMBOL) %>% head(20),
                   tmp$X_vs_Y %>% arrange(desc(logFC)) %>% pull(SYMBOL) %>% tail(20)) %>% unique()
FCFCplot(vplot = vplot, genes_to_label =  genes_to_label, xlimits = c(1/5,5), ylimits = c(1/5,5) )
vplot = data.frame(fc1 = 2^tmp$R_vs_S$logFC, fc2 = 2^(-tmp$X_vs_Y$logFC), SYMBOL = tmp$R_vs_S$SYMBOL)
FCFCplot(vplot = vplot, xlimits = c(1/5,5), ylimits = c(1/5,5) )

vplot = data.frame(fc1 = 2^tmp$T_vs_U$logFC, fc2 = 2^(tmp$X_vs_Y$logFC), SYMBOL = tmp$R_vs_S$SYMBOL)
FCFCplot(vplot = vplot )
```




**Figure 3 - CD4 GP analysis**
```{r}
mdata_orig = so@meta.data

#option 1 (best): import F1 and L1 from the fit object: 
L1 = readRDS(sprintf('/Users/david/google_drive_uchicago/ImmgenT/analysis/data_integration/IGT1_96/topic/flashier20250215_alldata_backfit/L1.Rds')) #cells x factor
F1 = readRDS(sprintf("/Users/david/google_drive_uchicago/ImmgenT/analysis/data_integration/IGT1_96/topic/flashier20250215_alldata_backfit/F1.Rds")) #gene x factor

mdata_orig %>% select(IGT, exp_name) %>% unique()

mdata_orig %>% count(organ_simplified)

#c("CD4.R", "CD4.S", "CD4.T", "CD4.U", "CD4.X", "CD4.Y", "CD4.H", "CD4.I")

group_list = list(
#1    
    list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.R", "CD4.S")) %>% pull(cellID), 
         group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.T", "CD4.U", "CD4.X", "CD4.Y", "CD4.H", "CD4.I")) %>% pull(cellID),  
         title = "CD4.RS vs CD4.TUXYHI"),
    
    list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.T", "CD4.U")) %>% pull(cellID), 
         group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.R", "CD4.S", "CD4.X", "CD4.Y", "CD4.H", "CD4.I")) %>% pull(cellID),  
         title = "CD4.TU vs CD4.RSXYHI"),
    
    list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.X", "CD4.Y")) %>% pull(cellID), 
         group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.R", "CD4.S", "CD4.T", "CD4.U", "CD4.H", "CD4.I")) %>% pull(cellID),  
         title = "CD4.XY vs CD4.RSTUHI"),
    
    list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.H", "CD4.I")) %>% pull(cellID), 
         group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.R", "CD4.S", "CD4.T", "CD4.U", "CD4.X", "CD4.Y")) %>% pull(cellID),  
         title = "CD4.HI vs CD4.RSTUXY"),
#2    
    list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.R", "CD4.S")) %>% pull(cellID), 
         group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.T", "CD4.U", "CD4.X", "CD4.Y")) %>% pull(cellID),  
         title = "CD4.RS vs CD4.TUXY"),
    
    list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.T", "CD4.U")) %>% pull(cellID), 
         group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.R", "CD4.S", "CD4.X", "CD4.Y")) %>% pull(cellID),  
         title = "CD4.TU vs CD4.RSXY"),
    
    list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.X", "CD4.Y")) %>% pull(cellID), 
         group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.R", "CD4.S", "CD4.T", "CD4.U")) %>% pull(cellID),  
         title = "CD4.XY vs CD4.RSTU"),
#3    
        list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.S")) %>% pull(cellID), 
         group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c( "CD4.U", "CD4.Y")) %>% pull(cellID),  
         title = "CD4.S vs CD4.UY"),
    
    list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.U")) %>% pull(cellID), 
         group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.S", "CD4.Y")) %>% pull(cellID),  
         title = "CD4.U vs CD4.SY"),
    
    list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c( "CD4.Y")) %>% pull(cellID), 
         group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.S", "CD4.U")) %>% pull(cellID),  
         title = "CD4.Y vs CD4.SU"),
#4    
    list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.R", "CD4.S") & organ_simplified %in% c("colon LP", "small intestine LP")) %>% pull(cellID), 
         group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.T", "CD4.U", "CD4.X", "CD4.Y", "CD4.H", "CD4.I") & organ_simplified %in% c("colon LP", "small intestine LP")) %>% pull(cellID),  
         title = "CD4.RS vs CD4.TUXYHI in gut"),
    
    list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.T", "CD4.U") & organ_simplified %in% c("colon LP", "small intestine LP")) %>% pull(cellID), 
         group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.R", "CD4.S", "CD4.X", "CD4.Y", "CD4.H", "CD4.I") & organ_simplified %in% c("colon LP", "small intestine LP")) %>% pull(cellID),  
         title = "CD4.TU vs CD4.RSXYHI in gut"),
    
    list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.X", "CD4.Y") & organ_simplified %in% c("colon LP", "small intestine LP")) %>% pull(cellID), 
         group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.R", "CD4.S", "CD4.T", "CD4.U", "CD4.H", "CD4.I") & organ_simplified %in% c("colon LP", "small intestine LP")) %>% pull(cellID),  
         title = "CD4.XY vs CD4.RSTUHI in gut"),
    
    list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.H", "CD4.I") & organ_simplified %in% c("colon LP", "small intestine LP")) %>% pull(cellID), 
         group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.R", "CD4.S", "CD4.T", "CD4.U", "CD4.X", "CD4.Y") & organ_simplified %in% c("colon LP", "small intestine LP")) %>% pull(cellID),  
         title = "CD4.HI vs CD4.RSTUXY in gut"),

#5    
    list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.R", "CD4.S") & organ_simplified %in% c("spleen", "LN")) %>% pull(cellID), 
         group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.T", "CD4.U", "CD4.X", "CD4.Y", "CD4.H", "CD4.I") & organ_simplified %in% c("spleen", "LN")) %>% pull(cellID),  
         title = "CD4.RS vs CD4.TUXYHI in SLO"),
    
    list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.T", "CD4.U") & organ_simplified %in% c("spleen", "LN")) %>% pull(cellID), 
         group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.R", "CD4.S", "CD4.X", "CD4.Y", "CD4.H", "CD4.I") & organ_simplified %in% c("spleen", "LN")) %>% pull(cellID),  
         title = "CD4.TU vs CD4.RSXYHI in SLO"),
    
    list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.X", "CD4.Y") & organ_simplified %in% c("spleen", "LN")) %>% pull(cellID), 
         group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.R", "CD4.S", "CD4.T", "CD4.U", "CD4.H", "CD4.I") & organ_simplified %in% c("spleen", "LN")) %>% pull(cellID),  
         title = "CD4.XY vs CD4.RSTUHI in SLO"),
    
    list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.H", "CD4.I") & organ_simplified %in% c("spleen", "LN")) %>% pull(cellID), 
         group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.R", "CD4.S", "CD4.T", "CD4.U", "CD4.X", "CD4.Y") & organ_simplified %in% c("spleen", "LN")) %>% pull(cellID),  
         title = "CD4.HI vs CD4.RSTUXY in SLO"),

#6   
    list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.G")) %>% pull(cellID), 
         group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.R", "CD4.S", "CD4.T", "CD4.U", "CD4.X", "CD4.Y", "CD4.H", "CD4.I") ) %>% pull(cellID),  
         title = "CD4.G vs CD4.RSTUXYHI"),
    
    list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.Q")) %>% pull(cellID), 
         group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.R", "CD4.S", "CD4.T", "CD4.U", "CD4.X", "CD4.Y", "CD4.H", "CD4.I")) %>% pull(cellID),  
         title = "CD4.Q vs CD4.RSTUXYHI"),

    list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.Q")) %>% pull(cellID), 
         group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.R", "CD4.S")) %>% pull(cellID),  
         title = "CD4.Q vs CD4.RS")
    
)

results_list = list()

for (i in 1:length(group_list)) {
    print(group_list[[i]]$title)
    results_list[[group_list[[i]]$title]] = FlashierDGE(F1, L1, group_list[[i]]$group1, group_list[[i]]$group2, title_plot = group_list[[i]]$title)
}

results_list[["CD4.RS vs CD4.TUXY"]]

```

GP in tips
```{r}

pdf(sprintf("%s/FlashierDGE_volcano_GP_tips1.pdf", out_dir), 15,15, useDingbats = F)
results_list[[1]]$p1
results_list[[2]]$p1
results_list[[3]]$p1
results_list[[4]]$p1
dev.off()

pdf(sprintf("%s/FlashierDGE_volcano_GP_tips2.pdf", out_dir), 15,15, useDingbats = F)
results_list[[5]]$p1
results_list[[6]]$p1
results_list[[7]]$p1
dev.off()

pdf(sprintf("%s/FlashierDGE_volcano_GP_tips3.pdf", out_dir), 15,15, useDingbats = F)
results_list[[8]]$p1
results_list[[9]]$p1
results_list[[10]]$p1
dev.off()

pdf(sprintf("%s/FlashierDGE_volcano_GP_tipsGut.pdf", out_dir), 15,15, useDingbats = F)
results_list[[11]]$p1
results_list[[12]]$p1
results_list[[13]]$p1
results_list[[14]]$p1
dev.off()

pdf(sprintf("%s/FlashierDGE_volcano_GP_tipsSLO.pdf", out_dir), 15,15, useDingbats = F)
results_list[[15]]$p1
results_list[[16]]$p1
results_list[[17]]$p1
results_list[[18]]$p1
dev.off()

pdf(sprintf("%s/FlashierDGE_volcano_GP_Midland.pdf", out_dir), 15,15, useDingbats = F)
results_list[[19]]$p1
results_list[[20]]$p1
results_list[[21]]$p1
dev.off()

results_list[[11]]$diff_genes["Tbx21",]
results_list[[12]]$diff_genes["Gata3",]
results_list[[13]]$diff_genes["Rorc",]
results_list[[14]]$diff_genes["Bcl6",]

results_list[[11]]$diff_genes["Ifng",]
results_list[[12]]$diff_genes["Il5",]
results_list[[13]]$diff_genes["Il17a",]
results_list[[14]]$diff_genes["Il21",]
```

```{r}

tt = results_list[5:7]

results_list[[1]]$diff_genes["Tbx21",]
results_list[[2]]$diff_genes["Gata3",]
results_list[[3]]$diff_genes["Rorc",]
results_list[[4]]$diff_genes["Bcl6",]

results_list[[5]]$diff_genes["Tbx21",]
results_list[[6]]$diff_genes["Gata3",]
results_list[[7]]$diff_genes["Rorc",]

results_list[[5]]$p1
results_list[[6]]$p1
results_list[[7]]$p1

results_list[[8]]$p1
results_list[[9]]$p1
results_list[[10]]$p1

plot(results_list[[8]]$diff_genes$log2FC, results_list[[9]]$diff_genes$log2FC)

comp_filt = names(tt)
# comp_filt = comp_filt[grepl("_vs_others", comp_filt)]

# #filter gene names
# genes_to_label <- c()
# for (i in comp_filt) {
#   new_genes <- tt[[i]]$diff_genes %>%
#     arrange(desc(log2FC)) %>%
#     # filter(adj.P.Val < 0.05) %>%
#     pull(SYMBOL)
#   new_genes <- setdiff(new_genes, genes_to_label)
#   genes_to_label <- c(genes_to_label, head(new_genes, 20))
# }

genes_to_label <- c()
for (i in comp_filt) {
  df <- tt[[i]]$diff_genes #%>% filter(adj.P.Val < 0.05)
  # top up
  up_genes <- df %>% filter(log2FC > 0.249) %>%
    arrange(desc(log2FC)) %>%
    pull(SYMBOL)
  # top down
  # down_genes <- df %>%
  #   arrange(log2FC) %>%
  #   pull(SYMBOL)
  # remove genes already selected
  up_genes <- setdiff(up_genes, genes_to_label)
  # down_genes <- setdiff(down_genes, genes_to_label)

  # add top 20 each
  genes_to_label <- c(
    genes_to_label,
    head(up_genes, 100)#, head(down_genes, 10)
  )
}

genes_to_label <- genes_to_label[
  !grepl(
    "^Gm|^mt-|^Rpl|^Rps|^Snor|^Mir|^Linc|^Rp[sl]|-ps|Rik$|^Hbb",
    genes_to_label,
    ignore.case = TRUE
  )
]

#plot heatmap
# mat_coef <- sapply(tt, function(tt) tt$logFC)
mat_coef <- Reduce(cbind, lapply(tt, function(x) x$diff_genes$log2FC))
rownames(mat_coef) <- tt[[1]]$diff_genes$SYMBOL
colnames(mat_coef) = make.names(names(tt))
mat_plot = mat_coef[genes_to_label, ]
range(mat_plot)

library(ComplexHeatmap)
library(circlize)

col_fun <- circlize::colorRamp2(
  c(-3, 0, 3),
  c("#4575B4", "white", "#D73027")
)

# pdf(sprintf("%s/heatmap_memory_vs_acute_SLO_LCMV.pdf", out_dir), width = 3, 8, useDingbats = F)
# pdf(sprintf("%s/Heatmap_LCMV_SLO_memory_vs_acute_IGT363840.pdf", out_dir2), width = 3, 8, useDingbats = F)
Heatmap(
  mat_plot,
  name = "log2FC",
  col = col_fun,
  cluster_columns = FALSE,
  cluster_rows = T,
    row_names_gp = gpar(fontsize = 6),
      column_names_gp = gpar(fontsize = 6),
  column_title = "memory vs acute across clusters in LCMV SLO"
)
# dev.off()

# Access the plots and data
results$p1  # First plot (MA plot of factors)
results$p2  # Second plot (MA plot of genes)
results$diff_factors  # table for fold changes of gene programs
results$diff_genes %>% arrange(desc(log2FC)) # table for fold changes of genes


```




**Figure 5 - memory **
--> see CD4_TRBI.Rmd for external datasets

Use new smarta definition!!!
```{r}
smarta_citeseq = read.table(sprintf("%s/smarta_citeseq_igt363840_2026.csv","/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/analysis/data_integration/IGT1_96/CD4/IGT363840_SmartaP14"), header = F)[,1]
so$smarta_2026 =  colnames(so) %in% smarta_citeseq
so@meta.data %>% count(Ag_spe, IGT)
so$Ag_spe[so$Ag_spe == "SMARTA"] = NA
so$Ag_spe[colnames(so) %in% smarta_citeseq] = "SMARTA"

so$infection_phase2 = NA
so$infection_phase2[so$IGT == "IGT43" & so$timepoint %in% c("6 dpi", "10 dpi")] = "acute"
so$infection_phase2[so$IGT == "IGT43" & so$timepoint %in% c("25 dpi", "30 dpi", "60 dpi")] = "cleared"
so$infection_phase2[so$IGT %in% c("IGT36", "IGT38", "IGT40")] = so$infection_phase[so$IGT %in% c("IGT36", "IGT38", "IGT40")]

so$organ_simplified2 = NA
so$organ_simplified2[so$organ_simplified %in% c("spleen", "LN", "SLO")] = "SLO"
so$organ_simplified2[!so$organ_simplified %in% c("spleen", "LN", "SLO")] = "NLT"

```

Plot MDE immgenT in LCMV arm and Listeria
```{r}
sub = so@meta.data %>% filter(infection_agent %in% c("ActA-deficient Listeria monocytogenes (DPL1942)-2W", "LCMV Armstrong"), organ_simplified %in% c("SLO", "LN", "spleen"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43")) %>% count(sample_code, sample_type, organ_simplified)
tmp = so[,so@meta.data %>% filter(infection_agent %in% c("ActA-deficient Listeria monocytogenes (DPL1942)-2W", "LCMV Armstrong"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43")) %>% rownames()] #organ_simplified %in% c("SLO", "LN", "spleen")
tmp$organ_simplified0 = "SLO"

tmp@meta.data %>% count(infection_agent, infection_phase, infection_phase2, timepoint, Ag_spe, organ_simplified2)
tmp@meta.data %>% filter(infection_agent == "ActA-deficient Listeria monocytogenes (DPL1942)-2W") %>% count(timepoint, infection_phase2, organ_simplified2)

tmp@meta.data %>% count(infection_agent, timepoint, Ag_spe)

###LCMV SLO
tmp@meta.data %>% filter(infection_agent == "LCMV Armstrong", Ag_spe == "SMARTA", organ_simplified2 == "SLO") %>% count(timepoint, infection_phase2, Ag_spe)
# sub = tmp@meta.data %>% filter(infection_agent == "LCMV Armstrong" & Ag_spe == "SMARTA") %>% group_by(timepoint) %>% slice_sample(n = 121) %>% ungroup() %>% pull(cellID)
sub = tmp@meta.data %>% filter(infection_agent == "LCMV Armstrong" & Ag_spe == "SMARTA", organ_simplified2 == "SLO") %>% pull(cellID)
tmp2 = tmp[,sub]
pdf(sprintf("%s/MDE_LCMV_timecourse_IGT363840_SLO.pdf",out_dir),7, 7, useDingbats = F)
for (i in unique(tmp2$timepoint)) {
    print(i)
    tmp3 = tmp2[,tmp2@meta.data %>% filter(timepoint == i) %>% rownames()]
    
    p = DimPlot(tmp3, reduction = "mde_incremental", group.by = "annotation_level2", cols = mypal_level2, pt.size = 2) + ggtitle(sprintf("%s, %s cells",i, ncol(tmp3)))
    bg = AddEmbeddingRasterBackground(lineage = "CD4")
    p$layers = c(list(bg), p$layers)
    p = p + ggplot2::coord_fixed() + xlim(c(-2.5,4)) + ylim(c(-2.5,2.5)) +theme_void() + NoLegend()
    print(p)
}
# dev.off()

# tmp@meta.data %>% filter(infection_agent == "LCMV Armstrong" & Ag_spe == "SMARTA") %>% count(infection_phase)
# sub = tmp@meta.data %>% filter(infection_agent == "LCMV Armstrong" & Ag_spe == "SMARTA", organ_simplified2 == "SLO") %>% group_by(infection_phase)  %>% ungroup() %>% pull(cellID) #%>% slice_sample(n = 80)
# tmp2 = tmp[,sub]
# pdf(sprintf("%s/MDE_LCMV_timecourse_IGT363840.pdf.pdf",out_dir),7, 7, useDingbats = F)
for (i in unique(tmp2$infection_phase2)) {
    print(i)
    tmp3 = tmp2[,tmp2@meta.data %>% filter(infection_phase2 == i) %>% rownames()]
    
    p = DimPlot(tmp3, reduction = "mde_incremental", group.by = "annotation_level2", cols = mypal_level2, pt.size = 2) + ggtitle(sprintf("%s, %s cells",i, ncol(tmp3)))
    bg = AddEmbeddingRasterBackground(lineage = "CD4")
    p$layers = c(list(bg), p$layers)
    p = p + ggplot2::coord_fixed() + xlim(c(-2.5,4)) + ylim(c(-2.5,2.5)) +theme_void() + NoLegend()
    print(p)
}
dev.off()


###Listeria
tmp@meta.data %>% filter(infection_agent == "ActA-deficient Listeria monocytogenes (DPL1942)-2W") %>% count(timepoint, infection_phase2)
tmp@meta.data %>% filter(infection_agent == "ActA-deficient Listeria monocytogenes (DPL1942)-2W", organ_simplified2 == "SLO") %>% count(infection_phase2, timepoint, organ_simplified2, Ag_spe)
# sub = tmp@meta.data %>% filter(infection_agent == "ActA-deficient Listeria monocytogenes (DPL1942)-2W") %>% group_by(infection_phase2) %>% slice_sample(n = 500)  %>% ungroup() %>% pull(cellID) 
sub = tmp@meta.data %>% filter(infection_agent == "ActA-deficient Listeria monocytogenes (DPL1942)-2W", organ_simplified2 == "SLO") %>% pull(cellID) #%>% slice_sample(n = 500) 
tmp2 = tmp[,sub]
pdf(sprintf("%s/MDE_Listeria_timecourse_IGT43_SLO.pdf",out_dir),7, 7, useDingbats = F)
for (i in unique(tmp2$timepoint)) {
    print(i)
    tmp3 = tmp2[,tmp2@meta.data %>% filter(timepoint == i) %>% rownames()]
    nrow(tmp3)
    
    p = DimPlot(tmp3, reduction = "mde_incremental", group.by = "annotation_level2", cols = mypal_level2, pt.size = 2) + ggtitle(sprintf("%s, %s cells",i, ncol(tmp3)))
    bg = AddEmbeddingRasterBackground(lineage = "CD4")
    p$layers = c(list(bg), p$layers)
    p = p + ggplot2::coord_fixed() + xlim(c(-2.5,4)) + ylim(c(-2.5,2.5)) +theme_void() + NoLegend()
    print(p)
}

#To check that the AddEmbeddingRasterBackground was working well
# for (i in unique(tmp2$timepoint)) {
#     print(i)
#     p = MyDimPlotHighlight(seurat_object = so, umap_to_plot = "mde_incremental", cells_to_highlight = tmp2@meta.data %>% filter(timepoint == i) %>% rownames(), highlight_column_name = "annotation_level2", pixels = c(512, 512), mycols = mypal_level2, title = "", highlight_size = 1, highlight_alpha = 1, print_plot1 = T, print_plot2 = F, labelclusters = F)
#     print(p)
# }

for (i in na.omit(unique(tmp2$infection_phase2))) {
    print(i)
    tmp3 = tmp2[,tmp2@meta.data %>% filter(infection_phase2 == i) %>% rownames()]
    
    p = DimPlot(tmp3, reduction = "mde_incremental", group.by = "annotation_level2", cols = mypal_level2, pt.size = 2) + ggtitle(sprintf("%s, %s cells",i, ncol(tmp3)))
    bg = AddEmbeddingRasterBackground(lineage = "CD4")
    p$layers = c(list(bg), p$layers)
    p = p + ggplot2::coord_fixed() + xlim(c(-2.5,4)) + ylim(c(-2.5,2.5)) +theme_void() + NoLegend()
    print(p)
}
dev.off()

###LCMV NLT
tmp@meta.data %>% filter(infection_agent == "LCMV Armstrong", Ag_spe == "SMARTA", organ_simplified2 == "NLT") %>% count(timepoint, infection_phase2, Ag_spe, organ_simplified2,organ_simplified)
# sub = tmp@meta.data %>% filter(infection_agent == "LCMV Armstrong" & Ag_spe == "SMARTA") %>% group_by(timepoint) %>% slice_sample(n = 121) %>% ungroup() %>% pull(cellID)
sub = tmp@meta.data %>% filter(infection_agent == "LCMV Armstrong" & Ag_spe == "SMARTA", organ_simplified2 == "NLT") %>% pull(cellID)
tmp2 = tmp[,sub]
pdf(sprintf("%s/MDE_LCMV_timecourse_IGT363840_NLT.pdf",out_dir),7, 7, useDingbats = F)
for (i in unique(tmp2$timepoint)) {
    print(i)
    tmp3 = tmp2[,tmp2@meta.data %>% filter(timepoint == i) %>% rownames()]
    
    p = DimPlot(tmp3, reduction = "mde_incremental", group.by = "annotation_level2", cols = mypal_level2, pt.size = 2) + ggtitle(sprintf("%s, %s cells",i, ncol(tmp3)))
    bg = AddEmbeddingRasterBackground(lineage = "CD4")
    p$layers = c(list(bg), p$layers)
    p = p + ggplot2::coord_fixed() + xlim(c(-2.5,4)) + ylim(c(-2.5,2.5)) +theme_void() + NoLegend()
    print(p)
}
# dev.off()

# tmp@meta.data %>% filter(infection_agent == "LCMV Armstrong" & Ag_spe == "SMARTA") %>% count(infection_phase)
# sub = tmp@meta.data %>% filter(infection_agent == "LCMV Armstrong" & Ag_spe == "SMARTA", organ_simplified2 == "SLO") %>% group_by(infection_phase)  %>% ungroup() %>% pull(cellID) #%>% slice_sample(n = 80)
# tmp2 = tmp[,sub]
for (i in unique(tmp2$infection_phase2)) {
    print(i)
    tmp3 = tmp2[,tmp2@meta.data %>% filter(infection_phase2 == i) %>% rownames()]
    
    p = DimPlot(tmp3, reduction = "mde_incremental", group.by = "annotation_level2", cols = mypal_level2, pt.size = 2) + ggtitle(sprintf("%s, %s cells",i, ncol(tmp3)))
    bg = AddEmbeddingRasterBackground(lineage = "CD4")
    p$layers = c(list(bg), p$layers)
    p = p + ggplot2::coord_fixed() + xlim(c(-2.5,4)) + ylim(c(-2.5,2.5)) +theme_void() + NoLegend()
    print(p)
}
dev.off()

###LCMV SILP_SMG_prostate
tmp@meta.data %>% filter(infection_agent == "LCMV Armstrong", Ag_spe == "SMARTA", organ_simplified2 == "NLT", organ_simplified %in% c("small intestine LP", "submandibular gland", "prostate")) %>% count(timepoint, infection_phase2, Ag_spe, organ_simplified2,organ_simplified)
# sub = tmp@meta.data %>% filter(infection_agent == "LCMV Armstrong" & Ag_spe == "SMARTA") %>% group_by(timepoint) %>% slice_sample(n = 121) %>% ungroup() %>% pull(cellID)
sub = tmp@meta.data %>% filter(infection_agent == "LCMV Armstrong" & Ag_spe == "SMARTA", organ_simplified2 == "NLT", organ_simplified %in% c("small intestine LP", "submandibular gland", "prostate")) %>% pull(cellID)
tmp2 = tmp[,sub]
pdf(sprintf("%s/MDE_LCMV_timecourse_IGT363840_NLT_SILP_SMG_prostate.pdf",out_dir),7, 7, useDingbats = F)
for (i in unique(tmp2$timepoint)) {
    print(i)
    tmp3 = tmp2[,tmp2@meta.data %>% filter(timepoint == i) %>% rownames()]
    
    p = DimPlot(tmp3, reduction = "mde_incremental", group.by = "annotation_level2", cols = mypal_level2, pt.size = 2) + ggtitle(sprintf("%s, %s cells",i, ncol(tmp3)))
    bg = AddEmbeddingRasterBackground(lineage = "CD4")
    p$layers = c(list(bg), p$layers)
    p = p + ggplot2::coord_fixed() + xlim(c(-2.5,4)) + ylim(c(-2.5,2.5)) +theme_void() + NoLegend()
    print(p)
}
# dev.off()

# tmp@meta.data %>% filter(infection_agent == "LCMV Armstrong" & Ag_spe == "SMARTA") %>% count(infection_phase)
# sub = tmp@meta.data %>% filter(infection_agent == "LCMV Armstrong" & Ag_spe == "SMARTA", organ_simplified2 == "SLO") %>% group_by(infection_phase)  %>% ungroup() %>% pull(cellID) #%>% slice_sample(n = 80)
# tmp2 = tmp[,sub]
for (i in unique(tmp2$infection_phase2)) {
    print(i)
    tmp3 = tmp2[,tmp2@meta.data %>% filter(infection_phase2 == i) %>% rownames()]
    
    p = DimPlot(tmp3, reduction = "mde_incremental", group.by = "annotation_level2", cols = mypal_level2, pt.size = 2) + ggtitle(sprintf("%s, %s cells",i, ncol(tmp3)))
    bg = AddEmbeddingRasterBackground(lineage = "CD4")
    p$layers = c(list(bg), p$layers)
    p = p + ggplot2::coord_fixed() + xlim(c(-2.5,4)) + ylim(c(-2.5,2.5)) +theme_void() + NoLegend()
    print(p)
}
dev.off()

###LCMV Lung bone marrow
tmp@meta.data %>% filter(infection_agent == "LCMV Armstrong", Ag_spe == "SMARTA", organ_simplified2 == "NLT", organ_simplified %in% c("lung", "bone marrow")) %>% count(timepoint, infection_phase2, Ag_spe, organ_simplified2,organ_simplified)
# sub = tmp@meta.data %>% filter(infection_agent == "LCMV Armstrong" & Ag_spe == "SMARTA") %>% group_by(timepoint) %>% slice_sample(n = 121) %>% ungroup() %>% pull(cellID)
sub = tmp@meta.data %>% filter(infection_agent == "LCMV Armstrong" & Ag_spe == "SMARTA", organ_simplified2 == "NLT", organ_simplified %in% c("lung", "bone marrow")) %>% pull(cellID)
tmp2 = tmp[,sub]
pdf(sprintf("%s/MDE_LCMV_timecourse_IGT363840_NLT_SILP_lung_bonemarrow.pdf",out_dir),7, 7, useDingbats = F)
for (i in unique(tmp2$timepoint)) {
    print(i)
    tmp3 = tmp2[,tmp2@meta.data %>% filter(timepoint == i) %>% rownames()]
    
    p = DimPlot(tmp3, reduction = "mde_incremental", group.by = "annotation_level2", cols = mypal_level2, pt.size = 2) + ggtitle(sprintf("%s, %s cells",i, ncol(tmp3)))
    bg = AddEmbeddingRasterBackground(lineage = "CD4")
    p$layers = c(list(bg), p$layers)
    p = p + ggplot2::coord_fixed() + xlim(c(-2.5,4)) + ylim(c(-2.5,2.5)) +theme_void() + NoLegend()
    print(p)
}
# dev.off()

# tmp@meta.data %>% filter(infection_agent == "LCMV Armstrong" & Ag_spe == "SMARTA") %>% count(infection_phase)
# sub = tmp@meta.data %>% filter(infection_agent == "LCMV Armstrong" & Ag_spe == "SMARTA", organ_simplified2 == "SLO") %>% group_by(infection_phase)  %>% ungroup() %>% pull(cellID) #%>% slice_sample(n = 80)
# tmp2 = tmp[,sub]
for (i in unique(tmp2$infection_phase2)) {
    print(i)
    tmp3 = tmp2[,tmp2@meta.data %>% filter(infection_phase2 == i) %>% rownames()]
    
    p = DimPlot(tmp3, reduction = "mde_incremental", group.by = "annotation_level2", cols = mypal_level2, pt.size = 2) + ggtitle(sprintf("%s, %s cells",i, ncol(tmp3)))
    bg = AddEmbeddingRasterBackground(lineage = "CD4")
    p$layers = c(list(bg), p$layers)
    p = p + ggplot2::coord_fixed() + xlim(c(-2.5,4)) + ylim(c(-2.5,2.5)) +theme_void() + NoLegend()
    print(p)
}
dev.off()

```

Cluster proportion LCMV and Listeria acute vs memory
```{r}
library(dplyr)
library(ggplot2)
library(forcats)
library(scales)
library(stringr)
library(tidyr)

cluster_col <- "annotation_level2"   # change if needed

# so@meta.data %>% filter(IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43")) %>% count(organ_simplified)
# so$organ_simplified2 = NA
# so$organ_simplified2[so$organ_simplified %in% c("spleen", "LN", "SLO")] = "SLO"
# so$organ_simplified2[!so$organ_simplified %in% c("spleen", "LN", "SLO")] = "NLT"

#checks
so@meta.data %>% filter(IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43"), infection_phase2 %in% c("acute", "cleared"), infection_agent %in% c("LCMV Armstrong", "ActA-deficient Listeria monocytogenes (DPL1942)-2W")) %>% count(infection_phase, infection_phase2, IGT, Ag_spe)
so@meta.data %>% filter(IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43"), infection_phase2 %in% c("acute", "cleared"), infection_agent %in% c("LCMV Armstrong", "ActA-deficient Listeria monocytogenes (DPL1942)-2W")) %>% count(IGT, organ_simplified, organ_simplified2)

so@meta.data %>% count(infection_agent, infection_phase, infection_phase, timepoint, Ag_spe, organ)
so@meta.data %>% filter(infection_agent == "ActA-deficient Listeria monocytogenes (DPL1942)-2W") %>% count(timepoint, infection_phase2)

#Make prop table
so@meta.data %>% filter(infection_phase %in% c("acute", "cleared"),infection_agent %in% c("LCMV Armstrong", "ActA-deficient Listeria monocytogenes (DPL1942)-2W"), Ag_spe %in% c("2WTetp", "SMARTA")) %>% count(sample_code, infection_agent, infection_phase2, organ_simplified2, Ag_spe)

clusters <- levels(so$annotation_level2)[grepl("^CD4", levels(so$annotation_level2))]

df_plot <- so@meta.data %>%
  filter(
    infection_phase2 %in% c("acute", "cleared"),
    infection_agent %in% c("LCMV Armstrong", "ActA-deficient Listeria monocytogenes (DPL1942)-2W"),
    Ag_spe %in% c("2WTetp", "SMARTA"),
  ) %>%
  count(
    infection_agent, infection_phase2, organ_simplified2, timepoint, #sample_code
    .data[[cluster_col]],
    name = "n"
  ) %>%
  rename(cluster = all_of(cluster_col)) %>%
  mutate(cluster = factor(cluster, levels = levels(so$annotation_level2))) %>%
  group_by(infection_agent, infection_phase2, organ_simplified2, timepoint) %>% #sample_code
  complete(cluster = clusters, fill = list(n = 0)) %>%
  mutate(
    sample_total = sum(n),
    pct = 100 * n / sample_total
  ) %>%
  ungroup()

df_plot2 <- df_plot %>%
  mutate(
    cluster_phase = paste(cluster, infection_phase2)
  )

x_levels <- as.vector(rbind(
  paste(clusters, "acute"),
  paste(clusters, "cleared")
))

df_plot2$cluster_phase <- factor(df_plot2$cluster_phase, levels = x_levels)

#make 
# df_plot <- so@meta.data %>%
#   filter(infection_phase2 %in% c("acute", "cleared"), infection_agent %in% c("LCMV Armstrong", "ActA-deficient Listeria monocytogenes (DPL1942)-2W"), Ag_spe %in% c("2WTetp", "SMARTA")) %>%
#   count(sample_code, infection_agent, infection_phase2,organ_simplified2, timepoint, .data[[cluster_col]], name = "n") %>%
#   group_by(sample_code, infection_agent, infection_phase2, organ_simplified2) %>%
#     complete(cluster = clusters, fill = list(n = 0)) %>%
#   mutate(
#     sample_total = sum(n),
#     pct = 100 * n / sample_total
#   ) %>%
#   ungroup() %>%
#   rename(cluster = all_of(cluster_col))
# df_plot$cluster = factor(df_plot$cluster, levels = levels(so$annotation_level2))
# 
# df_plot2 <- df_plot %>%
#   filter(infection_phase2 %in% c("acute","cleared")) %>%
#   mutate(
#     cluster_phase = paste(cluster, infection_phase2)
#   )
# clusters <- levels(so$annotation_level2)[grepl("^CD4", levels(so$annotation_level2))]
# x_levels <- as.vector(
#   rbind(
#     paste(clusters, "acute"),
#     paste(clusters, "cleared")
#   )
# )
# df_plot2$cluster_phase <- factor(df_plot2$cluster_phase, levels = x_levels)

# LCMV SLO
df_plot3 = df_plot2 %>% filter(!(cluster %in% c("CD4.wM", "CD4.wN", "CD4.wZ")), infection_agent %in% c("LCMV Armstrong"), organ_simplified2 %in% "SLO")
df_plot3$cluster = factor(df_plot3$cluster, levels = levels(so$annotation_level2))
# df_plot3 = df_plot3 %>% filter(sample_total > 10)

# df_plot3 %>% select(sample_code, infection_phase2) %>% unique() %>% count(sample_code, infection_phase2)

p3 = ggplot(df_plot3, aes(cluster, pct, color = infection_phase2, shape = timepoint)) +
    geom_jitter(
        position = position_jitterdodge(
            jitter.width = 0.1,
            dodge.width = 0.6
        ),
        size = 3
    ) +
    stat_summary(
        aes(group = infection_phase2),
        fun = mean,
        geom = "crossbar",
        width = 0.5,
        position = position_dodge(width = 0.6),
        color = "black"
    ) +
    scale_color_manual(values = c("acute" = "blue", "cleared" = "red")) +
    theme_bw() +
    theme(
        axis.text.x = element_text(angle = 60, hjust = 1),
        panel.grid.major.x = element_blank()
    ) +
    labs(
        x = "Cluster",
        y = "% cells"
    )
p3

pdf(sprintf("%s/ClusterProp_LCMV_acute_memory_SLO_IGT363840.pdf", out_dir), 8,5, useDingbats = F)
p3 + NoGrid()
dev.off()

# Listeria SLO
df_plot3 = df_plot2 %>% filter(!(cluster %in% c("CD4.wM", "CD4.wN", "CD4.wZ")), infection_agent %in% c("ActA-deficient Listeria monocytogenes (DPL1942)-2W"), organ_simplified2 %in% "SLO")
df_plot3$cluster = factor(df_plot3$cluster, levels = levels(so$annotation_level2))
# df_plot3 = df_plot3 %>% filter(sample_total > 10)

# df_plot3 %>% select(sample_code, infection_phase2) %>% unique() %>% count(sample_code, infection_phase2)

p3 = ggplot(df_plot3, aes(cluster, pct, color = infection_phase2)) +
    geom_jitter(
        position = position_jitterdodge(
            jitter.width = 0.1,
            dodge.width = 0.6
        ),
        size = 3
    ) +
    stat_summary(
        aes(group = infection_phase2),
        fun = mean,
        geom = "crossbar",
        width = 0.5,
        position = position_dodge(width = 0.6),
        color = "black"
    ) +
    scale_color_manual(values = c("acute" = "blue", "cleared" = "red")) +
    theme_bw() +
    theme(
        axis.text.x = element_text(angle = 60, hjust = 1),
        panel.grid.major.x = element_blank()
    ) +
    labs(
        x = "Cluster",
        y = "% cells"
    )
p3

pdf(sprintf("%s/ClusterProp_Listeria_acute_memory_SLO_IGT43.pdf", out_dir), 8,5, useDingbats = F)
p3 + NoGrid()
dev.off()

# LCMV NLT
df_plot3 = df_plot2 %>% filter(!(cluster %in% c("CD4.wM", "CD4.wN", "CD4.wZ")), infection_agent %in% c("LCMV Armstrong"), organ_simplified2 %in% "NLT")
df_plot3$cluster = factor(df_plot3$cluster, levels = levels(so$annotation_level2))
df_plot3 = df_plot3 %>% filter(sample_total > 10)

# df_plot3 %>% select(sample_code, infection_phase2) %>% unique() %>% count(sample_code, infection_phase2)

p3 = ggplot(df_plot3, aes(cluster, pct, color = infection_phase2, shape = timepoint)) +
    geom_jitter(
        position = position_jitterdodge(
             jitter.width = 0.1,
            dodge.width = 0.6
        ),
        size = 3
    ) +
    stat_summary(
        aes(group = infection_phase2),
        fun = mean,
        geom = "crossbar",
        width = 0.5,
        position = position_dodge(width = 0.6),
        color = "black"
    ) +
    scale_color_manual(values = c("acute" = "blue", "cleared" = "red")) +
    theme_bw() +
    theme(
        axis.text.x = element_text(angle = 60, hjust = 1),
        panel.grid.major.x = element_blank()
    ) +
    labs(
        x = "Cluster",
        y = "% cells"
    )
p3

pdf(sprintf("%s/ClusterProp_LCMV_acute_memory_NLT_IGT363840.pdf", out_dir), 8,5, useDingbats = F)
p3 + NoGrid()
dev.off()

```

TopicDGE LCMV
```{r}
mdata_orig = so@meta.data

L1 = readRDS(sprintf('/Users/david/google_drive_uchicago/ImmgenT/analysis/data_integration/IGT1_96/topic/flashier20250215_alldata_backfit/L1.Rds')) #cells x factor
F1 = readRDS(sprintf("/Users/david/google_drive_uchicago/ImmgenT/analysis/data_integration/IGT1_96/topic/flashier20250215_alldata_backfit/F1.Rds")) #gene x factor

mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43"), Ag_spe %in% c("SMARTA", "2WTetp"), organ_simplified2 %in% c("SLO"),infection_phase2 %in% c("acute", "cleared") ) %>% count(infection_agent, timepoint, infection_phase2, organ_simplified2, Ag_spe )
mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43")) %>% count(Ag_spe)

group_list = list(
    
    # list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40"), Ag_spe %in% c("SMARTA"), organ_simplified2 %in% c("SLO"), infection_phase == "cleared" )  %>% pull(cellID),
    #      group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40"), Ag_spe %in% c("SMARTA"), organ_simplified2 %in% c("SLO"), infection_phase == "acute" )  %>% pull(cellID),
    #      title = "CD4 memory vs acute in LCMV"),
    
    list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40"), Ag_spe %in% c("SMARTA"), organ_simplified2 %in% c("SLO"), infection_phase == "cleared", annotation_level2 %in% c("CD4.G"))  %>% pull(cellID),
         group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40"), Ag_spe %in% c("SMARTA"), organ_simplified2 %in% c("SLO"), infection_phase == "acute", annotation_level2 %in% c("CD4.G") )  %>% pull(cellID),
         title = "CD4.G memory vs acute in LCMV"),
    
    list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40"), Ag_spe %in% c("SMARTA"), organ_simplified2 %in% c("SLO"), infection_phase == "cleared", annotation_level2 == "CD4.H" )  %>% pull(cellID),
         group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40"), Ag_spe %in% c("SMARTA"), organ_simplified2 %in% c("SLO"), infection_phase == "acute", annotation_level2 == "CD4.H" )  %>% pull(cellID),
         title = "CD4.H memory vs acute in LCMV")
    
    # list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40"), Ag_spe %in% c("SMARTA"), organ_simplified2 %in% c("SLO"), infection_phase == "cleared", annotation_level2 %in% c("CD4.G", "CD4.H"))  %>% pull(cellID),
    #      group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40"), Ag_spe %in% c("SMARTA"), organ_simplified2 %in% c("SLO"), infection_phase == "acute", annotation_level2 %in% c("CD4.G", "CD4.H") )  %>% pull(cellID),
    #      title = "CD4.GH memory vs acute in LCMV"),
    
    # list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40"), Ag_spe %in% c("SMARTA"), organ_simplified2 %in% c("SLO"), infection_phase == "cleared", annotation_level2 %in% c("CD4.Q", "CD4.R", "CD4.S"))  %>% pull(cellID),
    #      group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40"), Ag_spe %in% c("SMARTA"), organ_simplified2 %in% c("SLO"), infection_phase == "acute", annotation_level2 %in% c("CD4.Q", "CD4.R", "CD4.S") )  %>% pull(cellID),
    #      title = "CD4.QRS memory vs acute in LCMV")
    
    # list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40"), Ag_spe %in% c("SMARTA"), organ_simplified2 %in% c("SLO"), infection_phase == "cleared", annotation_level2 %in% c("CD4.Q"))  %>% pull(cellID),
    #      group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40"), Ag_spe %in% c("SMARTA"), organ_simplified2 %in% c("SLO"), infection_phase == "acute", annotation_level2 %in% c("CD4.Q") )  %>% pull(cellID),
    #      title = "CD4.Q memory vs acute in LCMV")

    # list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43"), Ag_spe %in% c("SMARTA"), organ_simplified %in% c("spleen", "LN", "SLO"), annotation_level2 %in% c("CD4.A","CD4.B", "CD4.C"), infection_phase == "cleared" )  %>% pull(cellID),
    #      group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43"), Ag_spe %in% c("SMARTA"), organ_simplified %in% c("spleen", "LN", "SLO"), annotation_level2 %in% c("CD4.A","CD4.B", "CD4.C"), infection_phase == "acute" )  %>% pull(cellID),
    #      title = "CD4 ABC memory vs acute in LCMV"),
    
    # list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43"), Ag_spe %in% c("SMARTA"), organ_simplified %in% c("spleen", "LN", "SLO"), annotation_level2 %in% c("CD4.G","CD4.H", "CD4.I"), infection_phase == "cleared" )  %>% pull(cellID),
    #      group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43"), Ag_spe %in% c("SMARTA"), organ_simplified %in% c("spleen", "LN", "SLO"), annotation_level2 %in% c("CD4.G","CD4.H", "CD4.I"), infection_phase == "acute" )  %>% pull(cellID),
    #      title = "CD4 GHI memory vs acute in LCMV"),
    # 
    # list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43"), Ag_spe %in% c("SMARTA"), organ_simplified %in% c("spleen", "LN", "SLO"), annotation_level2 %in% c("CD4.Q","CD4.R", "CD4.S"), infection_phase == "cleared" )  %>% pull(cellID),
    #      group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43"), Ag_spe %in% c("SMARTA"), organ_simplified %in% c("spleen", "LN", "SLO"), annotation_level2 %in% c("CD4.Q","CD4.R", "CD4.S"), infection_phase == "acute" )  %>% pull(cellID),
    #      title = "CD4 QRS memory vs acute in LCMV")

)

results_list = list()

for (i in 1:length(group_list)) {
    print(group_list[[i]]$title)
    results_list[[group_list[[i]]$title]] = FlashierDGE(F1, L1, group_list[[i]]$group1, group_list[[i]]$group2, title_plot = group_list[[i]]$title)
}


# results$p1  # First plot (MA plot of factors)
# results$p2  # Second plot (MA plot of genes)
# results$diff_factors  %>% arrange(desc(log2FC)) %>% head(10)
# results$diff_genes %>% arrange(desc(log2FC))  %>% head(50)# table for fold changes of genes

results_list_LCMV = results_list

#Heatmap with the results
tt = results_list 

comp_filt = names(tt)
# comp_filt = comp_filt[grepl("_vs_others", comp_filt)]

# #filter gene names
# genes_to_label <- c()
# for (i in comp_filt) {
#   new_genes <- tt[[i]]$diff_genes %>%
#     arrange(desc(log2FC)) %>%
#     # filter(adj.P.Val < 0.05) %>%
#     pull(SYMBOL)
#   new_genes <- setdiff(new_genes, genes_to_label)
#   genes_to_label <- c(genes_to_label, head(new_genes, 20))
# }

genes_to_label <- c()
for (i in comp_filt) {
  df <- tt[[i]]$diff_genes #%>% filter(adj.P.Val < 0.05)
  # top up
  up_genes <- df %>%
    arrange(desc(log2FC)) %>%
    pull(SYMBOL)
  # top down
  down_genes <- df %>%
    arrange(log2FC) %>%
    pull(SYMBOL)
  # remove genes already selected
  # up_genes <- setdiff(up_genes, genes_to_label)
  # down_genes <- setdiff(down_genes, genes_to_label)

  # add top 20 each
  genes_to_label <- c(
    genes_to_label,
    head(up_genes, 20),
    head(down_genes, 20)
  )
}

genes_to_label <- genes_to_label[
  !grepl(
    "^Gm|^mt-|^Rpl|^Rps|^Snor|^Mir|^Linc|^Rp[sl]|-ps|Rik$|^Hbb",
    genes_to_label,
    ignore.case = TRUE
  )
] %>% unique()

genes_to_label_LCMV = genes_to_label

#plot heatmap
# mat_coef <- sapply(tt, function(tt) tt$logFC)
mat_coef <- Reduce(cbind, lapply(tt, function(x) x$diff_genes$log2FC))
rownames(mat_coef) <- tt[[1]]$diff_genes$SYMBOL
colnames(mat_coef) = make.names(names(tt))
mat_coef_LCMV = mat_coef
mat_plot = mat_coef[genes_to_label, ]
range(mat_plot)

library(ComplexHeatmap)
library(circlize)

col_fun <- circlize::colorRamp2(
  c(-1, 0, 1),
  c("#4575B4", "white", "#D73027")
)

# pdf(sprintf("%s/heatmap_memory_vs_acute_SLO_LCMV.pdf", out_dir), width = 3, 8, useDingbats = F)
pdf(sprintf("%s/Heatmap_LCMV_SLO_memory_vs_acute_IGT363840.pdf", out_dir), width = 3, 8, useDingbats = F)
Heatmap(
  mat_plot,
  name = "log2FC",
  col = col_fun,
  cluster_columns = FALSE,
  cluster_rows = T,
    row_names_gp = gpar(fontsize = 6),
      column_names_gp = gpar(fontsize = 6),
  column_title = "memory vs acute across clusters in LCMV SLO"
)
dev.off()




```

TopicDGE Listeria
```{r}
# mdata_orig = so@meta.data
# L1 = readRDS(sprintf('/Users/david/google_drive_uchicago/ImmgenT/analysis/data_integration/IGT1_96/topic/flashier20250215_alldata_backfit/L1.Rds')) #cells x factor
# F1 = readRDS(sprintf("/Users/david/google_drive_uchicago/ImmgenT/analysis/data_integration/IGT1_96/topic/flashier20250215_alldata_backfit/F1.Rds")) #gene x factor

mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43"), Ag_spe %in% c("2WTetp"), organ_simplified %in% c("spleen", "LN", "SLO") ) %>% count(infection_phase, annotation_level2)
mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43")) %>% count(Ag_spe)

group_list = list(
    
        # list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43"), Ag_spe %in% c("2WTetp"), organ_simplified %in% c("spleen", "LN", "SLO"), infection_phase == "cleared" )  %>% pull(cellID),
        #  group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43"), Ag_spe %in% c("2WTetp"), organ_simplified %in% c("spleen", "LN", "SLO"), infection_phase == "acute" )  %>% pull(cellID),
        #  title = "CD4 memory vs acute in Listeria"),

    # list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43"), Ag_spe %in% c("2WTetp"), organ_simplified %in% c("spleen", "LN", "SLO"), annotation_level2 %in% c("CD4.A","CD4.B", "CD4.C"), infection_phase == "cleared" )  %>% pull(cellID),
    #      group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43"), Ag_spe %in% c("2WTetp"), organ_simplified %in% c("spleen", "LN", "SLO"), annotation_level2 %in% c("CD4.A","CD4.B", "CD4.C"), infection_phase == "acute" )  %>% pull(cellID),
    #      title = "CD4 ABC memory vs acute in Listeria"),
    
    # list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43"), Ag_spe %in% c("2WTetp"), organ_simplified %in% c("spleen", "LN", "SLO"), annotation_level2 %in% c("CD4.G","CD4.H", "CD4.I"), infection_phase == "cleared" )  %>% pull(cellID),
    #      group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43"), Ag_spe %in% c("2WTetp"), organ_simplified %in% c("spleen", "LN", "SLO"), annotation_level2 %in% c("CD4.G","CD4.H", "CD4.I"), infection_phase == "acute" )  %>% pull(cellID),
    #      title = "CD4 GHI memory vs acute in Listeria"),
    
    list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43"), Ag_spe %in% c("2WTetp"), organ_simplified %in% c("spleen", "LN", "SLO"), annotation_level2 %in% c("CD4.G"), infection_phase == "cleared" )  %>% pull(cellID),
         group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43"), Ag_spe %in% c("2WTetp"), organ_simplified %in% c("spleen", "LN", "SLO"), annotation_level2 %in% c("CD4.G"), infection_phase == "acute" )  %>% pull(cellID),
         title = "CD4 G memory vs acute in Listeria"),
    
        list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43"), Ag_spe %in% c("2WTetp"), organ_simplified %in% c("spleen", "LN", "SLO"), annotation_level2 %in% c("CD4.H"), infection_phase == "cleared" )  %>% pull(cellID),
         group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43"), Ag_spe %in% c("2WTetp"), organ_simplified %in% c("spleen", "LN", "SLO"), annotation_level2 %in% c("CD4.H"), infection_phase == "acute" )  %>% pull(cellID),
         title = "CD4 H memory vs acute in Listeria"),
    
         #    list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43"), Ag_spe %in% c("2WTetp"), organ_simplified %in% c("spleen", "LN", "SLO"), annotation_level2 %in% c("CD4.I"), infection_phase == "cleared" )  %>% pull(cellID),
         # group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43"), Ag_spe %in% c("2WTetp"), organ_simplified %in% c("spleen", "LN", "SLO"), annotation_level2 %in% c("CD4.I"), infection_phase == "acute" )  %>% pull(cellID),
         # title = "CD4 I memory vs acute in Listeria"),
    
                list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43"), Ag_spe %in% c("2WTetp"), organ_simplified %in% c("spleen", "LN", "SLO"), annotation_level2 %in% c("CD4.Q"), infection_phase == "cleared" )  %>% pull(cellID),
         group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43"), Ag_spe %in% c("2WTetp"), organ_simplified %in% c("spleen", "LN", "SLO"), annotation_level2 %in% c("CD4.Q"), infection_phase == "acute" )  %>% pull(cellID),
         title = "CD4 Q memory vs acute in Listeria")
    
    # list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43"), Ag_spe %in% c("2WTetp"), organ_simplified %in% c("spleen", "LN", "SLO"), annotation_level2 %in% c("CD4.Q","CD4.R", "CD4.S"), infection_phase == "cleared" )  %>% pull(cellID),
    #      group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43"), Ag_spe %in% c("2WTetp"), organ_simplified %in% c("spleen", "LN", "SLO"), annotation_level2 %in% c("CD4.Q","CD4.R", "CD4.S"), infection_phase == "acute" )  %>% pull(cellID),
    #      title = "CD4 QRS memory vs acute in Listeria")
)

results_list = list()

for (i in 1:length(group_list)) {
    print(group_list[[i]]$title)
    results_list[[group_list[[i]]$title]] = FlashierDGE(F1, L1, group_list[[i]]$group1, group_list[[i]]$group2, title_plot = group_list[[i]]$title)
}


# results$p1  # First plot (MA plot of factors)
# results$p2  # Second plot (MA plot of genes)
# results$diff_factors  %>% arrange(desc(log2FC)) %>% head(10)
# results$diff_genes %>% arrange(desc(log2FC))  %>% head(50)# table for fold changes of genes

results_list_Lm = results_list

#Heatmap with the results
tt = results_list 

comp_filt = names(tt)
# comp_filt = comp_filt[grepl("_vs_others", comp_filt)]

#filter gene names
# genes_to_label <- c()
# for (i in comp_filt) {
#   new_genes <- tt[[i]]$diff_genes %>%
#     arrange(desc(abs(log2FC))) %>%
#     # filter(adj.P.Val < 0.05) %>%
#     pull(SYMBOL)
#   new_genes <- setdiff(new_genes, genes_to_label)
#   genes_to_label <- c(genes_to_label, head(new_genes, 20))
# }

genes_to_label <- c()
for (i in comp_filt) {
  df <- tt[[i]]$diff_genes #%>% filter(adj.P.Val < 0.05)
  # top up
  up_genes <- df %>%
    arrange(desc(log2FC)) %>%
    pull(SYMBOL)
  # top down
  down_genes <- df %>%
    arrange(log2FC) %>%
    pull(SYMBOL)
  # remove genes already selected
  # up_genes <- setdiff(up_genes, genes_to_label)
  # down_genes <- setdiff(down_genes, genes_to_label)

  # add top 20 each
  genes_to_label <- c(
    genes_to_label,
    head(up_genes, 20),
    head(down_genes, 20)
  )
}

genes_to_label <- genes_to_label[
  !grepl(
    "^Gm|^mt-|^Rpl|^Rps|^Snor|^Mir|^Linc|^Rp[sl]|-ps|Rik$|^Hbb",
    genes_to_label,
    ignore.case = TRUE
  )
] %>% unique()

genes_to_label_Lm = genes_to_label

#plot heatmap
# mat_coef <- sapply(tt, function(tt) tt$logFC)
mat_coef <- Reduce(cbind, lapply(tt, function(x) x$diff_genes$log2FC))
rownames(mat_coef) <- tt[[1]]$diff_genes$SYMBOL
colnames(mat_coef) = make.names(names(tt))
mat_coef_Lm = mat_coef
mat_plot = mat_coef[genes_to_label, ]
range(mat_plot)

library(ComplexHeatmap)
library(circlize)

col_fun <- circlize::colorRamp2(
  c(-0.5, 0, 0.5),
  c("#4575B4", "white", "#D73027")
)

# pdf(sprintf("%s/heatmap_memory_vs_acute_SLO_LCMV.pdf", out_dir), width = 3, 8, useDingbats = F)
pdf(sprintf("%s/Heatmap_Listeria_SLO_memory_vs_acute_IGT363840_2.pdf", out_dir), width = 3, 7, useDingbats = F)
Heatmap(
  mat_plot,
  name = "log2FC",
  col = col_fun,
  cluster_columns = FALSE,
  cluster_rows = T,
    row_names_gp = gpar(fontsize = 6),
      column_names_gp = gpar(fontsize = 6),
  column_title = "memory vs acute across clusters in Listeria SLO"
)
 dev.off()

```

Heatmap with LCMV and Listeria together
```{r}
tt = c(results_list_Lm, results_list_LCMV)
genes_to_label = setdiff(unique(genes_to_label_LCMV, genes_to_label_Lm), c("Cd8b1", "CT010467.1"))
intersect(genes_to_label_LCMV, genes_to_label_Lm)

all(rownames(mat_coef_LCMV) == rownames(mat_coef_Lm))
mat_coef =  cbind(mat_coef_LCMV, mat_coef_Lm)
mat_plot = mat_coef[genes_to_label, ]
mat_plot_LCMV = mat_coef_LCMV[genes_to_label, ]
mat_plot_Lm = mat_coef_Lm[genes_to_label, ]

range(mat_plot)

library(ComplexHeatmap)
library(circlize)

pdf(sprintf("%s/Heatmap_ListeriaLCMV_SLO_memory_vs_acute_IGT36384043_2.pdf", out_dir), width = 3, 8, useDingbats = F)
col_fun <- circlize::colorRamp2(
  c(-1, 0, 1),
  c("#4575B4", "white", "#D73027")
)
Heatmap(
  mat_plot_LCMV,
  name = "log2FC",
  col = col_fun,
  cluster_columns = FALSE,
  cluster_rows = T,
    row_names_gp = gpar(fontsize = 6),
      column_names_gp = gpar(fontsize = 6),
  column_title = "memory vs acute across clusters in Listeria SLO"
)

col_fun <- circlize::colorRamp2(
  c(-1, 0, 1),
  c("#4575B4", "white", "#D73027")
)
Heatmap(
  mat_plot_Lm,
  name = "log2FC",
  col = col_fun,
  cluster_columns = FALSE,
  cluster_rows = T,
    row_names_gp = gpar(fontsize = 6),
      column_names_gp = gpar(fontsize = 6),
  column_title = "memory vs acute across clusters in Listeria SLO"
)

col_fun <- circlize::colorRamp2(
  c(-0.75, 0, 0.75),
  c("#4575B4", "white", "#D73027")
)
Heatmap(
  cbind(mat_plot_Lm, mat_plot_LCMV),
  name = "log2FC",
  col = col_fun,
  cluster_columns = T,
  cluster_rows = T,
    row_names_gp = gpar(fontsize = 6),
      column_names_gp = gpar(fontsize = 6),
  column_title = "memory vs acute across clusters in Listeria + LCMV"
)
dev.off()

```

TopicDGE LCMV NLT vs SLO
```{r}
mdata_orig = so@meta.data

L1 = readRDS(sprintf('/Users/david/google_drive_uchicago/ImmgenT/analysis/data_integration/IGT1_96/topic/flashier20250215_alldata_backfit/L1.Rds')) #cells x factor
F1 = readRDS(sprintf("/Users/david/google_drive_uchicago/ImmgenT/analysis/data_integration/IGT1_96/topic/flashier20250215_alldata_backfit/F1.Rds")) #gene x factor

mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43"), Ag_spe %in% c("SMARTA"), organ_simplified %in% c("spleen", "LN", "SLO") ) %>% count(infection_phase, organ_simplified)
mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43"), Ag_spe %in% c("SMARTA"), !(organ_simplified %in% c("spleen", "LN", "SLO")) ) %>% count(infection_phase, organ_simplified)
mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43")) %>% count(Ag_spe)

group_list = list(
    
    list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43"), Ag_spe %in% c("SMARTA"), organ_simplified %in% c("spleen", "LN", "SLO"), annotation_level2 %in% c("CD4.G","CD4.Q"), infection_phase == "cleared" ) %>% pull(cellID),
         group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43"), Ag_spe %in% c("SMARTA"), !(organ_simplified %in% c("spleen", "LN", "SLO")), annotation_level2 %in% c("CD4.G","CD4.Q"), infection_phase == "cleared" )  %>% pull(cellID),
         title = "SLO vs NLT in GQ, in LCMV at memory timepoint"),
    
    list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43"), Ag_spe %in% c("SMARTA"), organ_simplified %in% c("spleen", "LN", "SLO"), annotation_level2 %in% c("CD4.G"), infection_phase == "cleared" ) %>% pull(cellID),
         group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43"), Ag_spe %in% c("SMARTA"), !(organ_simplified %in% c("spleen", "LN", "SLO")), annotation_level2 %in% c("CD4.G"), infection_phase == "cleared" )  %>% pull(cellID),
         title = "SLO vs NLT in G, in LCMV at memory timepoint"),
    
    list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43"), Ag_spe %in% c("SMARTA"), organ_simplified %in% c("spleen", "LN", "SLO"), annotation_level2 %in% c("CD4.Q"), infection_phase == "cleared" ) %>% pull(cellID),
         group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43"), Ag_spe %in% c("SMARTA"), !(organ_simplified %in% c("spleen", "LN", "SLO")), annotation_level2 %in% c("CD4.Q"), infection_phase == "cleared" )  %>% pull(cellID),
         title = "SLO vs NLT in Q, in LCMV at memory timepoint")
    
)

results_list = list()

for (i in 1:length(group_list)) {
    print(group_list[[i]]$title)
    results_list[[group_list[[i]]$title]] = FlashierDGE(F1, L1, group_list[[i]]$group1, group_list[[i]]$group2, title_plot = group_list[[i]]$title)
}

# results = results_list[[3]]
# results$p1  # First plot (MA plot of factors)
# results$p2  # Second plot (MA plot of genes)
# results$diff_factors  %>% arrange(desc(log2FC)) %>% head(10)
# results$diff_genes %>% arrange(desc(log2FC))  %>% head(50)# table for fold changes of genes

pdf(sprintf("%s/Volcano_LCMV_SLO_vs_NLT_IGT363840.pdf", out_dir2), width = 5, 5, useDingbats = F)
results = results_list[["SLO vs NLT in GQ, in LCMV at memory timepoint"]]
results$p1  # First plot (MA plot of factors)
results$p2  # Second plot (MA plot of genes)
# results$diff_factors  %>% arrange(desc(log2FC)) %>% head(10)
# results$diff_genes %>% arrange(desc(log2FC))  %>% head(50)# table for fold changes of genes
dev.off()

#Heatmap with the results
tt = results_list 

comp_filt = names(tt)
# comp_filt = comp_filt[grepl("_vs_others", comp_filt)]

# #filter gene names
# genes_to_label <- c()
# for (i in comp_filt) {
#   new_genes <- tt[[i]]$diff_genes %>%
#     arrange(desc(log2FC)) %>%
#     # filter(adj.P.Val < 0.05) %>%
#     pull(SYMBOL)
#   new_genes <- setdiff(new_genes, genes_to_label)
#   genes_to_label <- c(genes_to_label, head(new_genes, 20))
# }

genes_to_label <- c()
for (i in comp_filt) {
  df <- tt[[i]]$diff_genes #%>% filter(adj.P.Val < 0.05)
  # top up
  up_genes <- df %>%
    arrange(desc(log2FC)) %>%
    pull(SYMBOL)
  # top down
  down_genes <- df %>%
    arrange(log2FC) %>%
    pull(SYMBOL)
  # remove genes already selected
  up_genes <- setdiff(up_genes, genes_to_label)
  down_genes <- setdiff(down_genes, genes_to_label)

  # add top 20 each
  genes_to_label <- c(
    genes_to_label,
    head(up_genes, 10),
    head(down_genes, 10)
  )
}

genes_to_label <- genes_to_label[
  !grepl(
    "^Gm|^mt-|^Rpl|^Rps|^Snor|^Mir|^Linc|^Rp[sl]|-ps|Rik$|^Hbb",
    genes_to_label,
    ignore.case = TRUE
  )
]


#plot heatmap
# mat_coef <- sapply(tt, function(tt) tt$logFC)
mat_coef <- Reduce(cbind, lapply(tt, function(x) x$diff_genes$log2FC))
rownames(mat_coef) <- tt[[1]]$diff_genes$SYMBOL
colnames(mat_coef) = make.names(names(tt))
mat_coef_LCMV = mat_coef
mat_plot = mat_coef[genes_to_label, ]
range(mat_plot)

library(ComplexHeatmap)
library(circlize)

col_fun <- circlize::colorRamp2(
  c(-1, 0, 1),
  c("#4575B4", "white", "#D73027")
)

# pdf(sprintf("%s/heatmap_memory_vs_acute_SLO_LCMV.pdf", out_dir), width = 3, 8, useDingbats = F)
pdf(sprintf("%s/Heatmap_LCMV_SLO_vs_NLT_IGT363840.pdf", out_dir2), width = 3, 8, useDingbats = F)
Heatmap(
  mat_plot,
  name = "log2FC",
  col = col_fun,
  cluster_columns = FALSE,
  cluster_rows = T,
    row_names_gp = gpar(fontsize = 6),
      column_names_gp = gpar(fontsize = 6),
  column_title = "SLO vs NLT in GQ, in LCMV"
)
dev.off()


```


**Figure 6: resting vs activated**

```{r}
ttlist = readRDS('/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/20260206_freeze/CD4/DGE_limma/20260206/ttlist_Level2Group.Rds')

names(ttlist)
tmp = ttlist

i = "Activated_vs_Resting"

# gene_sig = list()

pdf(sprintf("%s/volcanos_Resting_vs_Activated_noLabels.pdf", out_dir), 10, 10, useDingbats = F)
pdf(sprintf("%s/volcanos_Resting_vs_Activated_Labels.pdf", out_dir), 25, 25, useDingbats = F)

# for (i in names(ttlist)) {
    print(i)
    vplot = na.omit(data.frame(SYMBOL = tmp[[i]]$SYMBOL, fc = 2^(-tmp[[i]]$logFC), pval = tmp[[i]]$P.Value + 10^-150, AveExpr = 2^(tmp[[i]]$AveExpr))) #dealing with NA
    # vplot = data.frame(SYMBOL = tmp[[i]]$SYMBOL, fc = 2^tmp[[i]]$logFC, pval = tmp[[i]]$P.Value+10^-300, AveExpr = 2^(tmp[[i]]$AveExpr))
    p = Vplot(vplot = vplot, xlab = names(tmp)[i], xlimits = c(min(vplot$fc[vplot$fc != 0]), max(vplot$fc)), ylimits = c(min(vplot$pval), 1), color = "black")  #c(min(vplot$pval[vplot$pval != 0]),1)
    sub = vplot %>%
        filter(abs(log2(fc)) > 1 & pval < 10^-100) %>%
        arrange(desc(log2(fc))) %>%
        pull(SYMBOL)
    sub
    gene_sig[[sprintf("%s_up", i)]] = sub
    highlight_genes = sub #unique(c(head(sub, 25), tail(sub, 25)))
    # sub = (log2(vplot$fc) > 0.5 | log2(vplot$fc) < -0.5)  & vplot$pval < 0.05
    # sub = vplot[sub,] %>% arrange(desc(log2(fc)))
    # sub1 = sub %>% pull(SYMBOL) %>% head(25) #vplot$SYMBOL %in% as.character(sub2$SYMBOL)[1:200]
    # sub2 = sub %>% pull(SYMBOL) %>% tail(25)
    highlight_filter = vplot$SYMBOL %in% highlight_genes
    # sub = vplot$SYMBOL %in% c(sub1, sub2)#as.character(sub2$SYMBOL)[1:50]
    library(ggrepel)
    print(p)
    # print(p + geom_text_repel(data = vplot[highlight_filter,], aes(x = fc, y = pval, label = SYMBOL)) + ggtitle(i) + xlab(sprintf("FC %s", i)))
# }
dev.off()
```

```{r}
ttlist = readRDS('/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/20260206_freeze/CD4/DGE_limma/20260206/ttlist_in_Resting.Rds')

tmp = ttlist
for (i in names(ttlist)) {
    print(i)
    vplot = na.omit(data.frame(SYMBOL = tmp[[i]]$SYMBOL, fc = 2^(tmp[[i]]$logFC), pval = tmp[[i]]$P.Value + 10^-100, AveExpr = 2^(tmp[[i]]$AveExpr))) #dealing with NA
    # vplot = data.frame(SYMBOL = tmp[[i]]$SYMBOL, fc = 2^tmp[[i]]$logFC, pval = tmp[[i]]$P.Value+10^-300, AveExpr = 2^(tmp[[i]]$AveExpr))
    p = Vplot(vplot = vplot, xlab = names(tmp)[i], xlimits = c(min(vplot$fc[vplot$fc != 0]), max(vplot$fc)), ylimits = c(min(vplot$pval), 1), color = "black")  #c(min(vplot$pval[vplot$pval != 0]),1)
    sub = vplot %>%
        filter(abs(log2(fc)) > 0.5 & pval < 10^-10) %>%
        arrange(desc(log2(fc))) %>%
        pull(SYMBOL)
    sub
    gene_sig[[sprintf("%s_up", i)]] = sub
    highlight_genes = sub #unique(c(head(sub, 25), tail(sub, 25)))
    # sub = (log2(vplot$fc) > 0.5 | log2(vplot$fc) < -0.5)  & vplot$pval < 0.05
    # sub = vplot[sub,] %>% arrange(desc(log2(fc)))
    # sub1 = sub %>% pull(SYMBOL) %>% head(25) #vplot$SYMBOL %in% as.character(sub2$SYMBOL)[1:200]
    # sub2 = sub %>% pull(SYMBOL) %>% tail(25)
    highlight_filter = vplot$SYMBOL %in% highlight_genes
    # sub = vplot$SYMBOL %in% c(sub1, sub2)#as.character(sub2$SYMBOL)[1:50]
    library(ggrepel)
    # print(p)
    print(p + geom_text_repel(data = vplot[highlight_filter,], aes(x = fc, y = pval, label = SYMBOL)) + ggtitle(i) + xlab(sprintf("FC %s", i)))
}
```


**Figure 7 flow analysis**

organ comparison in healthy 
```{r}
flow_data = read.table(file = '/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/analysis/data_integration/IGT1_96/CD4/misc_plots/flow_prop_sample_clusterpertconv.csv', header = T, sep = ",")

library(dplyr)
library(tidyr)
library(ggplot2)

flow_long <- flow_data %>%
  filter(condition_detailed == "healthy") %>%
  pivot_longer(
    cols = starts_with("CD4."),
    names_to = "cluster",
    values_to = "prop"
  )

flow_long$cluster <- factor(flow_long$cluster, levels = c("CD4.A","CD4.G","CD4.H","CD4.I","CD4.Q","CD4.RS","CD4.TU","CD4.XY"))
flow_long$organ_simplified = factor(flow_long$organ_simplified, levels = c("spleen", "lung", "colon"))

# mean per cluster and organ
mean_df <- flow_long %>%
  group_by(cluster, organ_simplified) %>%
  summarise(mean_prop = mean(prop), .groups = "drop")

# ANOVA p-value per cluster
anova_df <- flow_long %>%
  group_by(cluster) %>%
  summarise(
    p_value = summary(aov(prop ~ organ_simplified))[[1]][["Pr(>F)"]][1],
    y_pos = max(prop, na.rm = TRUE) * 1.18,
    label = paste0("p=", signif(p_value, 2)),
    .groups = "drop"
  )

p = ggplot(flow_long, aes(x = cluster, y = prop, color = organ_simplified)) +
  geom_point(
    position = position_jitterdodge(jitter.width = 0.12, dodge.width = 0.65),
    size = 3,
    alpha = 0.8
  ) +
  geom_point(
    data = mean_df,
    aes(x = cluster, y = mean_prop, color = organ_simplified),
    position = position_dodge(width = 0.65),
    shape = 95,
    size = 8
  ) +
  geom_text(
    data = mean_df,
    aes(x = cluster, y = mean_prop, label = sprintf("%.2f", mean_prop), color = organ_simplified),
    position = position_dodge(width = 0.65),
    vjust = -0.7,
    size = 3
  ) +
  geom_text(
    data = anova_df,
    aes(x = cluster, y = y_pos, label = label),
    inherit.aes = FALSE,
    size = 3.2,
    color = "black"
  ) +
    ggtitle("Healthy organ comparison") +
  labs(
    x = NULL,
    y = "Proportion of CD4 cluster",
    color = "Organ"
  ) +
    scale_color_manual(values = c("spleen" = "blue", "lung" = "darkgreen", "colon" = "red")) +
  theme_classic(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  expand_limits(y = max(anova_df$y_pos) * 1.05)

p

pdf(sprintf("%s/flow_organ_comparison_healthy.pdf", out_dir), width = 8, height = 5, useDingbats = F)
print(p)
dev.off()


```

colon healthy vs scurfy
```{r}
flow_data = read.table(file = '/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/analysis/data_integration/IGT1_96/CD4/misc_plots/flow_prop_sample_clusterpertconv.csv', header = T, sep = ",")

library(dplyr)
library(tidyr)
library(ggplot2)

flow_long <- flow_data %>%
  filter(organ_simplified == "colon", condition_detailed %in% c("healthy", "scurfy")) %>%
  pivot_longer(
    cols = starts_with("CD4."),
    names_to = "cluster",
    values_to = "prop"
  )

flow_long$cluster <- factor(
  flow_long$cluster,
  levels = c("CD4.A", "CD4.G", "CD4.H", "CD4.I", "CD4.Q", "CD4.RS", "CD4.TU", "CD4.XY")
)

flow_long$condition_detailed <- factor(
  flow_long$condition_detailed,
  levels = c("healthy", "scurfy")
)

# mean per cluster and condition
mean_df <- flow_long %>%
  group_by(cluster, condition_detailed) %>%
  summarise(mean_prop = mean(prop), .groups = "drop")

# p-value per cluster
test_df <- flow_long %>%
  group_by(cluster) %>%
  summarise(
    p_value = t.test(prop ~ condition_detailed)$p.value,
    y_pos = max(prop, na.rm = TRUE) * 1.18,
    label = paste0("p=", signif(p_value, 2)),
    .groups = "drop"
  )

p = ggplot(flow_long, aes(x = cluster, y = prop, color = condition_detailed)) +
  geom_point(
    position = position_jitterdodge(jitter.width = 0.12, dodge.width = 0.5),
    size = 3,
    alpha = 0.8
  ) +
  geom_point(
    data = mean_df,
    aes(x = cluster, y = mean_prop, color = condition_detailed),
    position = position_dodge(width = 0.5),
    shape = 95,
    size = 8
  ) +
  geom_text(
    data = mean_df,
    aes(x = cluster, y = mean_prop, label = sprintf("%.2f", mean_prop), color = condition_detailed),
    position = position_dodge(width = 0.5),
    vjust = -0.7,
    size = 3
  ) +
  geom_text(
    data = test_df,
    aes(x = cluster, y = y_pos, label = label),
    inherit.aes = FALSE,
    size = 3.2,
    color = "black"
  ) +
        ggtitle("Healthy vs scurfy in colon") +
  labs(
    x = NULL,
    y = "Proportion of CD4 cluster",
    color = "Condition"
  ) +
  scale_color_manual(values = c("healthy" = "black", "scurfy" = "red")) +
  theme_classic(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  expand_limits(y = max(test_df$y_pos) * 1.05)

p

pdf(sprintf("%s/flow_scurfy_vs_healthy_colon.pdf", out_dir), width = 8, height = 5, useDingbats = F)
print(p)
dev.off()

```

spleen healthy vs scurfy
```{r}
flow_data = read.table(file = '/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/analysis/data_integration/IGT1_96/CD4/misc_plots/flow_prop_sample_clusterpertconv.csv', header = T, sep = ",")

library(dplyr)
library(tidyr)
library(ggplot2)

flow_long <- flow_data %>%
  filter(organ_simplified == "spleen", condition_detailed %in% c("healthy", "scurfy")) %>%
  pivot_longer(
    cols = starts_with("CD4."),
    names_to = "cluster",
    values_to = "prop"
  )

flow_long$cluster <- factor(
  flow_long$cluster,
  levels = c("CD4.A", "CD4.G", "CD4.H", "CD4.I", "CD4.Q", "CD4.RS", "CD4.TU", "CD4.XY")
)

flow_long$condition_detailed <- factor(
  flow_long$condition_detailed,
  levels = c("healthy", "scurfy")
)

# mean per cluster and condition
mean_df <- flow_long %>%
  group_by(cluster, condition_detailed) %>%
  summarise(mean_prop = mean(prop), .groups = "drop")

# p-value per cluster
test_df <- flow_long %>%
  group_by(cluster) %>%
  summarise(
    p_value = t.test(prop ~ condition_detailed)$p.value,
    y_pos = max(prop, na.rm = TRUE) * 1.18,
    label = paste0("p=", signif(p_value, 2)),
    .groups = "drop"
  )

p = ggplot(flow_long, aes(x = cluster, y = prop, color = condition_detailed)) +
  geom_point(
    position = position_jitterdodge(jitter.width = 0.12, dodge.width = 0.5),
    size = 3,
    alpha = 0.8
  ) +
  geom_point(
    data = mean_df,
    aes(x = cluster, y = mean_prop, color = condition_detailed),
    position = position_dodge(width = 0.5),
    shape = 95,
    size = 8
  ) +
  geom_text(
    data = mean_df,
    aes(x = cluster, y = mean_prop, label = sprintf("%.2f", mean_prop), color = condition_detailed),
    position = position_dodge(width = 0.5),
    vjust = -0.7,
    size = 3
  ) +
  geom_text(
    data = test_df,
    aes(x = cluster, y = y_pos, label = label),
    inherit.aes = FALSE,
    size = 3.2,
    color = "black"
  ) +
            ggtitle("Healthy vs scurfy in spleen") +
  labs(
    x = NULL,
    y = "Proportion of CD4 cluster",
    color = "Condition"
  ) +
  scale_color_manual(values = c("healthy" = "black", "scurfy" = "red")) +
  theme_classic(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  expand_limits(y = max(test_df$y_pos) * 1.05)

p

pdf(sprintf("%s/flow_scurfy_vs_healthy_spleen.pdf", out_dir), width = 8, height = 5, useDingbats = F)
print(p)
dev.off()

```

Lung healthy vs scurfy
```{r}
flow_data = read.table(file = '/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/analysis/data_integration/IGT1_96/CD4/misc_plots/flow_prop_sample_clusterpertconv.csv', header = T, sep = ",")

library(dplyr)
library(tidyr)
library(ggplot2)

flow_long <- flow_data %>%
  filter(organ_simplified == "lung", condition_detailed %in% c("healthy", "scurfy")) %>%
  pivot_longer(
    cols = starts_with("CD4."),
    names_to = "cluster",
    values_to = "prop"
  )

flow_long$cluster <- factor(
  flow_long$cluster,
  levels = c("CD4.A", "CD4.G", "CD4.H", "CD4.I", "CD4.Q", "CD4.RS", "CD4.TU", "CD4.XY")
)

flow_long$condition_detailed <- factor(
  flow_long$condition_detailed,
  levels = c("healthy", "scurfy")
)

# mean per cluster and condition
mean_df <- flow_long %>%
  group_by(cluster, condition_detailed) %>%
  summarise(mean_prop = mean(prop), .groups = "drop")

# p-value per cluster
test_df <- flow_long %>%
  group_by(cluster) %>%
  summarise(
    p_value = t.test(prop ~ condition_detailed)$p.value,
    y_pos = max(prop, na.rm = TRUE) * 1.18,
    label = paste0("p=", signif(p_value, 2)),
    .groups = "drop"
  )

p = ggplot(flow_long, aes(x = cluster, y = prop, color = condition_detailed)) +
  geom_point(
    position = position_jitterdodge(jitter.width = 0.12, dodge.width = 0.5),
    size = 3,
    alpha = 0.8
  ) +
  geom_point(
    data = mean_df,
    aes(x = cluster, y = mean_prop, color = condition_detailed),
    position = position_dodge(width = 0.5),
    shape = 95,
    size = 8
  ) +
  geom_text(
    data = mean_df,
    aes(x = cluster, y = mean_prop, label = sprintf("%.2f", mean_prop), color = condition_detailed),
    position = position_dodge(width = 0.5),
    vjust = -0.7,
    size = 3
  ) +
  geom_text(
    data = test_df,
    aes(x = cluster, y = y_pos, label = label),
    inherit.aes = FALSE,
    size = 3.2,
    color = "black"
  ) +
            ggtitle("Healthy vs scurfy in lung") +
  labs(
    x = NULL,
    y = "Proportion of CD4 cluster",
    color = "Condition"
  ) +
  scale_color_manual(values = c("healthy" = "black", "scurfy" = "red")) +
  theme_classic(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  expand_limits(y = max(test_df$y_pos) * 1.05)

p

pdf(sprintf("%s/flow_scurfy_vs_healthy_lung.pdf", out_dir), width = 8, height = 5, useDingbats = F)
print(p)
dev.off()

```

spleen healthy vs LCMV
```{r}
flow_data = read.table(file = '/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/analysis/data_integration/IGT1_96/CD4/misc_plots/flow_prop_sample_clusterpertconv.csv', header = T, sep = ",")

library(dplyr)
library(tidyr)
library(ggplot2)

flow_long <- flow_data %>%
  filter(organ_simplified == "spleen", condition_detailed %in% c("healthy", "LCMV")) %>%
  pivot_longer(
    cols = starts_with("CD4."),
    names_to = "cluster",
    values_to = "prop"
  )

flow_long$cluster <- factor(
  flow_long$cluster,
  levels = c("CD4.A", "CD4.G", "CD4.H", "CD4.I", "CD4.Q", "CD4.RS", "CD4.TU", "CD4.XY")
)

flow_long$condition_detailed <- factor(
  flow_long$condition_detailed,
  levels = c("healthy", "LCMV")
)

# mean per cluster and condition
mean_df <- flow_long %>%
  group_by(cluster, condition_detailed) %>%
  summarise(mean_prop = mean(prop), .groups = "drop")

# p-value per cluster
test_df <- flow_long %>%
  group_by(cluster) %>%
  summarise(
    p_value = t.test(prop ~ condition_detailed)$p.value,
    y_pos = max(prop, na.rm = TRUE) * 1.18,
    label = paste0("p=", signif(p_value, 2)),
    .groups = "drop"
  )

p = ggplot(flow_long, aes(x = cluster, y = prop, color = condition_detailed)) +
  geom_point(
    position = position_jitterdodge(jitter.width = 0.12, dodge.width = 0.5),
    size = 3,
    alpha = 0.8
  ) +
  geom_point(
    data = mean_df,
    aes(x = cluster, y = mean_prop, color = condition_detailed),
    position = position_dodge(width = 0.5),
    shape = 95,
    size = 8
  ) +
  geom_text(
    data = mean_df,
    aes(x = cluster, y = mean_prop, label = sprintf("%.2f", mean_prop), color = condition_detailed),
    position = position_dodge(width = 0.5),
    vjust = -0.7,
    size = 3
  ) +
  geom_text(
    data = test_df,
    aes(x = cluster, y = y_pos, label = label),
    inherit.aes = FALSE,
    size = 3.2,
    color = "black"
  ) +
            ggtitle("Healthy vs LCMV in spleen") +
  labs(
    x = NULL,
    y = "Proportion of CD4 cluster",
    color = "Condition"
  ) +
  scale_color_manual(values = c("healthy" = "black", "LCMV" = "red")) +
  theme_classic(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  expand_limits(y = max(test_df$y_pos) * 1.05)

p

pdf(sprintf("%s/flow_LCMV_vs_healthy_spleen.pdf", out_dir), width = 8, height = 5, useDingbats = F)
print(p)
dev.off()

```

lung healthy vs LCMV
```{r}
flow_data = read.table(file = '/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/analysis/data_integration/IGT1_96/CD4/misc_plots/flow_prop_sample_clusterpertconv.csv', header = T, sep = ",")

library(dplyr)
library(tidyr)
library(ggplot2)

flow_long <- flow_data %>%
  filter(organ_simplified == "lung", condition_detailed %in% c("healthy", "LCMV")) %>%
  pivot_longer(
    cols = starts_with("CD4."),
    names_to = "cluster",
    values_to = "prop"
  )

flow_long$cluster <- factor(
  flow_long$cluster,
  levels = c("CD4.A", "CD4.G", "CD4.H", "CD4.I", "CD4.Q", "CD4.RS", "CD4.TU", "CD4.XY")
)

flow_long$condition_detailed <- factor(
  flow_long$condition_detailed,
  levels = c("healthy", "LCMV")
)

# mean per cluster and condition
mean_df <- flow_long %>%
  group_by(cluster, condition_detailed) %>%
  summarise(mean_prop = mean(prop), .groups = "drop")

# p-value per cluster
test_df <- flow_long %>%
  group_by(cluster) %>%
  summarise(
    p_value = t.test(prop ~ condition_detailed)$p.value,
    y_pos = max(prop, na.rm = TRUE) * 1.18,
    label = paste0("p=", signif(p_value, 2)),
    .groups = "drop"
  )

p = ggplot(flow_long, aes(x = cluster, y = prop, color = condition_detailed)) +
  geom_point(
    position = position_jitterdodge(jitter.width = 0.12, dodge.width = 0.5),
    size = 3,
    alpha = 0.8
  ) +
  geom_point(
    data = mean_df,
    aes(x = cluster, y = mean_prop, color = condition_detailed),
    position = position_dodge(width = 0.5),
    shape = 95,
    size = 8
  ) +
  geom_text(
    data = mean_df,
    aes(x = cluster, y = mean_prop, label = sprintf("%.2f", mean_prop), color = condition_detailed),
    position = position_dodge(width = 0.5),
    vjust = -0.7,
    size = 3
  ) +
  geom_text(
    data = test_df,
    aes(x = cluster, y = y_pos, label = label),
    inherit.aes = FALSE,
    size = 3.2,
    color = "black"
  ) +
            ggtitle("Healthy vs LCMV in lung") +
  labs(
    x = NULL,
    y = "Proportion of CD4 cluster",
    color = "Condition"
  ) +
  scale_color_manual(values = c("healthy" = "black", "LCMV" = "red")) +
  theme_classic(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  expand_limits(y = max(test_df$y_pos) * 1.05)

p

pdf(sprintf("%s/flow_LCMV_vs_healthy_lung.pdf", out_dir), width = 8, height = 5, useDingbats = F)
print(p)
dev.off()

```

Dotplot with p values
```{r}
library(dplyr)
library(tidyr)
library(ggplot2)

flow_long <- flow_data %>%
  filter(
    condition_detailed %in% c("healthy", "scurfy", "LCMV")
  ) %>%
  pivot_longer(
    cols = starts_with("CD4."),
    names_to = "cluster",
    values_to = "prop"
  ) %>%
  mutate(
    organ_simplified = factor(organ_simplified, levels = c("spleen", "lung", "colon")),
    condition_detailed = factor(condition_detailed, levels = c("healthy", "scurfy", "LCMV")),
    cluster = factor(cluster, levels = c("CD4.A", "CD4.G", "CD4.H", "CD4.I", "CD4.XY", "CD4.Q", "CD4.RS", "CD4.TU"))
  )

anova_df <- flow_long %>%
  group_by(cluster, organ_simplified) %>%
  summarise(
    pval = {
      fit <- aov(prop ~ condition_detailed)
      summary(fit)[[1]][["Pr(>F)"]][1]
    },
    .groups = "drop"
  ) %>%
  mutate(signif = pval < 0.05)

plot_df <- flow_long %>%
  group_by(organ_simplified, condition_detailed, cluster) %>%
  summarise(mean_prop = mean(prop, na.rm = TRUE), .groups = "drop") %>%
  left_join(anova_df, by = c("cluster", "organ_simplified"))

plot_df$organ_simplified = factor(plot_df$organ_simplified, levels = rev(c("spleen", "lung", "colon")))

cluster_cols <- c(
  "CD4.A"  = "#3f5ab5",
  "CD4.G"  = "#8cc63e",
  "CD4.H"  = "#b7d7a8",
  "CD4.I"  = "#ef4444",
  "CD4.XY" = "#ec4899",
  "CD4.Q"  = "#f59e0b",
  "CD4.RS" = "#8e44ad",
  "CD4.TU" = "#ef4444"
)

p = ggplot(plot_df, aes(x = condition_detailed, y = organ_simplified)) +
  geom_point(
    aes(
      size = mean_prop,
      fill = cluster,
      stroke = ifelse(signif, 1.5, 0.3)  # thick border if significant
    ),
    shape = 21,
    color = "black"
  ) +
  facet_wrap(~ cluster, nrow = 1) +
  scale_fill_manual(values = cluster_cols, guide = "none") +
  scale_size_continuous(
    name = "Proportion of Tconv",
    range = c(1, 16)
  ) +
  labs(
    x = NULL,
    y = "Flow-gated:"
  ) +
  theme_classic(base_size = 16) +
  theme(
    strip.background = element_blank(),
    strip.text = element_text(size = 16),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 12),
    axis.text.y = element_text(size = 16),
    panel.spacing.x = unit(0.8, "lines")
  )

p 

pdf(sprintf("%s/flow_dotplot_all.pdf", out_dir), width = 18, height = 3, useDingbats = F)
print(p)
dev.off()
```




