---
title: "CD4_misc"
output: html_document
date: "2025-05-28"
author: "David Zemmour"
editor_options: 
  chunk_output_type: console
---

**Initialize: load libraries and custom functions**

```{bash}
sinteractive --exclusive --time=12:00:00 --account=pi-zemmour --partition=jfkfloor2 --qos=jfkfloor2 --nodes=1 
sinteractive --exclusive --time=12:00:00 --account=pi-zemmour --partition=beagle3-bigmem --nodes=1 
sinteractive --exclusive --time=12:00:00 --account=pi-zemmour --partition=zemmour-hm --qos=zemmour --nodes=1 
sinteractive --time=03:00:00 --account=pi-zemmour --partition=caslake --nodes=1 --mem=96GB

module load python/anaconda-2022.05 
module load hdf5/1.12.0 #for BPCells but also others
module load openblas/0.3.13
module load ImageMagick/7.1.1-7
source activate /project/zemmour/david/envs/scvi_20240315
#source activate /project/zemmour/david/envs/r443
module load openblas/0.3.13

cd /project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/CD4
R
```


```{r}
setwd("~/google_drive_uchicago/ImmgenT/analysis/data_integration/IGT1_96/CD4")
libs = c("ZemmourLib","forcats", "gridExtra","patchwork", "Matrix", "Seurat", "ggplot2", "dplyr", "reshape2", "ggrastr","scattermore", "RColorBrewer", "pals", "scales", "pheatmap", "rafalib") 
sapply(libs, function(x) suppressMessages(suppressWarnings(library(x, character.only = TRUE, quietly = T, warn.conflicts  = F))))
#source("/project/jfkfloor2/zemmourlab/david/immgent/immgent_integration_git/custom_functions_david.R")

ColorRamp = rev(colorRampPalette(brewer.pal(n = 7,name = "RdYlBu"))(100))
ColorRamp = rev(colorRampPalette(c('red','white','blue'))(20))
ColorRamp = rev(rainbow(10, end = 4/6))
ColorRamp = rev(viridis(100))
ColorRamp = brewer.pal(9, "Blues")

#mycol = c("black","blue","red","darkgreen", "orange","purple", "cyan", "greenyellow",    "salmon", "magenta","pink", "tan", "brown") # c("magenta", "red",  "darkgreen", "cyan", "blue", "blue", "blue", "black", "blue", "blue", "blue", "blue", "orange")

mypal_level2 = ZemmourLib::immgent_colors$level2
mypal_level2["not classified"] = "black"
# names(mypal_level2) = gsub("nonconv", replacement = "Tz",names(mypal_level2) )
# mypal_level2 = immgent_colors$level2_option2
# mypal_level2["not classified"] = "black"
mypal_level1 = immgent_colors$level1
mypal_level1["not classified"] = "black"
# names(mypal_level1) = gsub("nonconv", replacement = "Tz",names(mypal_level1) )
mypal_organ = immgent_colors$organ_simplified
mypal_level2group = c("resting" = "blue", "activated" = "red", "miniverse" = "darkgreen", "proliferating" = "black", "other" = "grey", "preT" = "grey" )

#Large Color palette
library("pals")
n = 70
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
mypal1 = unique(unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals))))
mypal1 = mypal1[-4]
mypal = c(glasbey(), polychrome(), mypal1)
names(mypal) = NULL
#pal.bands(mypal, main="Colormap suggestions")
# mypal_organ = setNames(mypal, unique(so$organ_simplified))
# mypal_organ = mypal_organ[!is.na(names(mypal_organ))]
# mypal_level1 = setNames(mypal, unique(so$annotation_level1))
# mypal_level1 = mypal_level1[!is.na(names(mypal_level1))]
# mypal_level2 = setNames(mypal, unique(so_orig$annotation_level2))
# mypal_level2 = mypal_level2[!is.na(names(mypal_level2))]
# mypal_level2["CD4_cl18"] = "blue"
mypal_igt = setNames(mypal, unique(so_orig$IGT))
mypal_igt = mypal_igt[!is.na(names(mypal_igt))]

```

```{r}
# fig_dir = "~/google_drive_uchicago/ImmgenT/20250109_freeze/paper_cosmology/figures/bits"
# fig_dir = "/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/paper_cosmology/figures/bits"

out_dir = "~/google_drive_uchicago/ImmgenT/analysis/data_integration/IGT1_96/CD4/misc_plots" 
out_dir = "/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/CD4/misc_plots"

```


**Most up-to-date data**
on Midway
```{r}
so_orig = readRDS("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/igt1_96_withtotalvi20260206_clean.Rds") 

so = so_orig[,so_orig@meta.data %>% filter(organ_simplified != "thymus") %>% row.names()]
so = NormalizeData(so, assay = "RNA", normalization.method = "LogNormalize")
```

on Desktop
```{r}
setwd("~/google_drive_uchicago/ImmgenT/analysis/data_integration/IGT1_96/")
so_orig = readRDS("igt1_96_withtotalvi20260206_clean_ADTonly.Rds")

so = so_orig[,so_orig@meta.data %>% filter(annotation_level1 == "CD4" & organ_simplified != "thymus") %>% row.names()]
# so = NormalizeData(so, assay = "RNA", normalization.method = "LogNormalize")
# so = NormalizeData(so, assay = "ADT", normalization.method = "LogNormalize")

```

**FCFC plots**
Tfh/Tex
```{r}
source('/Users/david/Library/CloudStorage/GoogleDrive-david.puyraimond@gmail.com/My Drive/ImmgenT/analysis/immgent_integration_git/custom_functions_david.R')

tt_list = readRDS('/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/analysis/data_integration/IGT1_96/CD4/DGE_limma/20250109/ttlist_OneVsAll.Rds')

names(tt_list)


# Create the vplot dataframe with fold changes
vplot = data.frame(fc1 = exp(tt_list$CD4_cl13_vs_All$logFC), fc2 = exp(tt_list$CD4_cl14_vs_All$logFC), SYMBOL = tt_list$CD4_cl13_vs_All$SYMBOL)

# Identify genes with large fold changes (>2)
genes_to_plot = abs(tt_list$CD4_cl13_vs_All$logFC) > 0.69 | abs(tt_list$CD4_cl14_vs_All$logFC) > 0.69
# genes_to_label = vplot[genes_to_plot, ]
genes_to_label = vplot$SYMBOL[genes_to_plot]

# Create the plot
FCFCplot(vplot, 
         xlab = "FC for CD4_cl13", 
         ylab = "FC for CD4_cl14", 
         main = "Fold Change vs Fold Change", 
         genes_to_label = genes_to_label)

FCFCplot(vplot, 
         xlab = "FC for CD4_cl13", 
         ylab = "FC for CD4_cl14", 
         main = "Fold Change vs Fold Change", 
         genes_to_label = NULL)

vplot = data.frame(fc1 = exp(tt_list$CD8_cl10_vs_All$logFC), fc2 = exp(tt_list$CD8_cl7_vs_All$logFC), SYMBOL = tt_list$CD8_cl10_vs_All$SYMBOL)
# Identify genes with large fold changes (>2)
genes_to_plot = abs(tt_list$CD8_cl10_vs_All$logFC) > 0.69 | abs(tt_list$CD8_cl7_vs_All$logFC) > 0.69
genes_to_label = vplot[genes_to_plot, ]
# Create the plot
FCFCplot(vplot, 
         xlab = "FC for CD8_cl10", 
         ylab = "FC for CD8_cl7", 
         main = "Fold Change vs Fold Change", 
         printgeomtext = T, 
         genes_to_label = genes_to_label)



```



**MDE plots**
```{r}
pdf("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/CD4/MDE_clusters_20250527.pdf", width = 10, height = 10, useDingbats = F)
p = DimPlot(so, reduction = "mde_incremental", group.by = "annotation_level2", split.by = NULL, raster = T, raster.dpi = c(512,512), label = T, pt.size = 1) + scale_color_manual(values = mypal_level2)
p
p + NoLegend()

p = DimPlot(so, reduction = "mde_incremental", group.by = "annotation_level2", split.by = NULL, raster = T, raster.dpi = c(512,512), label = F, pt.size = 1) + scale_color_manual(values = mypal_level2)
p + NoLegend()

p = MyDimPlotHighlight(seurat_object = so, umap_to_plot = "mde_incremental", cells_to_highlight = colnames(so)[which(so$annotation_level2 %in% c("CD4_cl1","CD4_cl2","CD4_cl3", "CD4_cl4", "CD4_cl5", "CD4_cl7", "CD4_cl12"))], highlight_column_name = "annotation_level2", pixels = c(512, 512), mycols = mypal_level2, title = "", highlight_size = 0.1, highlight_alpha = 0.1, print_plot1 = T, print_plot2 = T, labelclusters = T)
p

p = MyDimPlotHighlight(seurat_object = so, umap_to_plot = "mde_incremental", cells_to_highlight = colnames(so)[which(so$annotation_level2 %in% c("CD4_cl2"))], highlight_column_name = "annotation_level2", pixels = c(512, 512), mycols = mypal_level2, title = "", highlight_size = 0.1, highlight_alpha = 0.1, print_plot1 = T, print_plot2 = T, labelclusters = T)
p

p = MyDimPlotHighlight(seurat_object = so, umap_to_plot = "mde_incremental", cells_to_highlight = colnames(so)[which(so$annotation_level2 %in% c("CD4_cl5", "CD4_cl7"))], highlight_column_name = "annotation_level2", pixels = c(512, 512), mycols = mypal_level2, title = "", highlight_size = 0.1, highlight_alpha = 0.1, print_plot1 = T, print_plot2 = T, labelclusters = T)
p

p = MyDimPlotHighlight(seurat_object = so, umap_to_plot = "mde_incremental", cells_to_highlight = colnames(so)[which(so$annotation_level2 %in% c("CD4_cl9", "CD4_cl10", "CD4_cl13", "CD4_cl14"))], highlight_column_name = "annotation_level2", pixels = c(512, 512), mycols = mypal_level2, title = "", highlight_size = 0.1, highlight_alpha = 0.1, print_plot1 = T, print_plot2 = T, labelclusters = T)
p

p = MyDimPlotHighlight(seurat_object = so, umap_to_plot = "mde_incremental", cells_to_highlight = colnames(so)[which(so$annotation_level2 %in% c("CD4_cl9","CD4_cl21","CD4_cl10","CD4_cl13", "CD4_cl14", "CD4_cl15", "CD4_cl16", "CD4_cl17", "CD4_cl18", "CD4_cl22", "CD4_cl23"))], highlight_column_name = "annotation_level2", pixels = c(512, 512), mycols = mypal_level2, title = "", highlight_size = 0.1, highlight_alpha = 0.1, print_plot1 = T, print_plot2 = T, labelclusters = T)
p

p = MyDimPlotHighlight(seurat_object = so, umap_to_plot = "mde_incremental", cells_to_highlight = colnames(so)[which(so$annotation_level2 %in% c("CD4_cl15", "CD4_cl16", "CD4_cl17", "CD4_cl18"))], highlight_column_name = "annotation_level2", pixels = c(512, 512), mycols = mypal_level2, title = "", highlight_size = 0.1, highlight_alpha = 0.1, print_plot1 = T, print_plot2 = T, labelclusters = T)
p

p = MyDimPlotHighlight(seurat_object = so, umap_to_plot = "mde_incremental", cells_to_highlight = colnames(so)[which(so$annotation_level2 %in% c("CD4_cl21", "CD4_cl22", "CD4_cl23"))], highlight_column_name = "annotation_level2", pixels = c(512, 512), mycols = mypal_level2, title = "", highlight_size = 0.1, highlight_alpha = 0.1, print_plot1 = T, print_plot2 = T, labelclusters = T)
p


p = MyDimPlotHighlight(seurat_object = so, umap_to_plot = "mde_incremental", cells_to_highlight = colnames(so)[which(so$annotation_level2 %in% c("CD4_cl9","CD4_cl21"))], highlight_column_name = "annotation_level2", pixels = c(512, 512), mycols = mypal_level2, title = "", highlight_size = 0.1, highlight_alpha = 0.1, print_plot1 = T, print_plot2 = T, labelclusters = T)
p

p = MyDimPlotHighlight(seurat_object = so, umap_to_plot = "mde_incremental", cells_to_highlight = colnames(so)[which(so$annotation_level2 %in% c("CD4_cl10","CD4_cl13", "CD4_cl14", "CD4_cl15", "CD4_cl16", "CD4_cl17", "CD4_cl18", "CD4_cl22", "CD4_cl23"))], highlight_column_name = "annotation_level2", pixels = c(512, 512), mycols = mypal_level2, title = "", highlight_size = 0.1, highlight_alpha = 0.1, print_plot1 = T, print_plot2 = T, labelclusters = T)
p
dev.off()
```

**Feature plots**
```{r}
so = NormalizeData(so,assay = "RNA", normalization.method = "LogNormalize",verbose = T)

FeaturePlot(so, reduction = "mde_incremental", slot = "data", features = "Tox", order = T, raster = T, raster.dpi = c(1024,1024))
FeaturePlot(so, reduction = "mde_incremental", slot = "data", features = "Gata3", order = T, raster = T, raster.dpi = c(1024,1024))
FeaturePlot(so, reduction = "mde_incremental", slot = "data", features = "Tbx21", order = T, raster = T, raster.dpi = c(1024,1024))
FeaturePlot(so, reduction = "mde_incremental", slot = "data", features = "Rorc", order = T, raster = T, raster.dpi = c(1024,1024))
FeaturePlot(so, reduction = "mde_incremental", slot = "data", features = "Tox2", order = T, raster = T, raster.dpi = c(1024,1024))
```

**Downsample seurat object**
```{r}
sample_cells = so@meta.data %>%
  group_by(!!sym("IGTHT")) %>%
  sample_n(size = min(500, n()), replace = FALSE) %>% 
    pull(IGT_cellID)

so = so[,sample_cells]
saveRDS(so, "/Users/david/Desktop/Treg_App/igt1_96_CD4_20250212_downsampled.Rds")
```

**Figure Tip/BAS**

2a
```{r}
pdf(sprintf("%s/MDE_level2.pdf", out_dir), 5,5, useDingbats = F)
DimPlot(so, reduction = "mde_incremental", group.by = "annotation_level2", cols = mypal_level2, raster = T, raster.dpi = c(1024,1024), pt.size = 1, alpha = 1, label = T, label.size = 2, label.box = T) + NoGrid()  + NoLegend()
p = DimPlot(so, reduction = "mde_incremental", group.by = "annotation_level2", cols = mypal_level2, raster = T, raster.dpi = c(1024,1024), pt.size = 1, alpha = 1) + xlim(c(-2.5,4)) + ylim(c(-2.5,2.5))
p
p + theme_void() + NoLegend() + NoGrid()
dev.off()

MyFeatureScatter(so,feature1 = "CD62L", feature2 = "CD44")

```

**Figure 2**
tips percent
```{r}
so@meta.data %>% count(annotation_level2, annotation_level2_group)
mdata = so@meta.data %>% filter(annotation_level2_group == "activated", !(annotation_level2 %in% c("CD4.wN", "CD4.wZ"))) 

mdata %>% count(mdata)
mdata$tip_mid = "tip"
mdata$tip_mid[mdata$annotation_level2 %in% c("CD4.G", "CD4.Q")] = "midland"
mdata$tip_mid[mdata$annotation_level2 %in% c("CD4.R", "CD4.S")] = "th1_tip"
mdata$tip_mid[mdata$annotation_level2 %in% c("CD4.T", "CD4.U")] = "th2_tip"
mdata$tip_mid[mdata$annotation_level2 %in% c("CD4.X", "CD4.Y")] = "th17_tip"
mdata$tip_mid[mdata$annotation_level2 %in% c("CD4.H", "CD4.I")] = "tfhchron_tip"
mdata %>% count(annotation_level2, tip_mid)

library(dplyr)

summary_df <- mdata %>%
    group_by(IGTHT, tip_mid) %>%
    summarise(n = n(), .groups = "drop") %>%
    # total per sample
    group_by(IGTHT) %>%
    mutate(
        total_cells = sum(n),
        percent = n / total_cells * 100
    ) %>%
    ungroup()

n_categories_gt1 <- summary_df %>% 
    filter(tip_mid != "midland") %>%
    group_by(IGTHT) %>%
    summarise(
        n_categories_gt1 = sum(percent > 1),
        .groups = "drop"
    )

n_categories_gt1 %>% count(n_categories_gt1)

# library(tidyr)
# wide_counts <- summary_df %>%
#   select(IGTHT, tip_mid, n) %>%
#   pivot_wider(names_from = tip_mid, values_from = n, values_fill = 0)
# wide_percent <- summary_df %>%
#   select(IGTHT, tip_mid, percent) %>%
#   pivot_wider(names_from = tip_mid, values_from = percent, values_fill = 0)

n_categories_gt1 %>%
  count(n_categories_gt1) %>%
  ggplot(aes(x = factor(n_categories_gt1), y = n)) +
  geom_col(fill = "grey70", color = "black") +
  labs(
    x = "# tip_mid categories (>1%)",
    y = "# IGTHT samples"
  ) +
  theme_classic()

library(dplyr)
library(ggplot2)

plot_df <- n_categories_gt1 %>%
  count(n_categories_gt1) %>%
  mutate(percent = n / sum(n) * 100)

pdf(sprintf("%s/barplot_number_tips_per_sample.pdf", out_dir), 3,6, useDingbats = F)
ggplot(plot_df, aes(x = factor(n_categories_gt1), y = percent)) +
  geom_col(fill = "grey70", color = "black") +
  geom_text(
    aes(label = sprintf("%.1f%%", percent)),
    vjust = -0.3,
    size = 5
  ) +
  labs(
    x = "# tip (>1%)",
    y = "% of samples"
  ) +
  theme_classic(base_size = 16) +
  theme(
    axis.text = element_text(size = 14)
  ) +
  expand_limits(y = max(plot_df$percent) * 1.1)
dev.off()
```

**Figure 3 - tip signature**
--> limma_wrapper_20260318_CD4_tips.sh
limma_contrasts_20260318_CD4_tips.R
```{r}
ttlist = readRDS('/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/analysis/data_integration/IGT1_96/CD4/DGE_limma/20260318_tips/ttlist_tips.Rds')

# ttlist = read.table(file = '/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/analysis/data_integration/IGT1_96/CD4/DGE_limma/20260318_tips/ttlist_tips.txt', sep = "\t", header = T)
# ttlist["Ifng",]

names(ttlist)
tmp = ttlist

# i = "RS_vs_TUXYHI"

gene_sig = list()

pdf("/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/analysis/data_integration/IGT1_96/CD4/DGE_limma/20260318_tips/volcanos_more_labels.pdf", 15, 15, useDingbats = F)
for (i in names(ttlist)) {
    print(i)
    vplot = na.omit(data.frame(SYMBOL = tmp[[i]]$SYMBOL, fc = 2^tmp[[i]]$logFC, pval = tmp[[i]]$P.Value + 10^-300, AveExpr = 2^(tmp[[i]]$AveExpr))) #dealing with NA
    # vplot = data.frame(SYMBOL = tmp[[i]]$SYMBOL, fc = 2^tmp[[i]]$logFC, pval = tmp[[i]]$P.Value+10^-300, AveExpr = 2^(tmp[[i]]$AveExpr))
    p = Vplot(vplot = vplot, xlab = names(tmp)[i], xlimits = c(min(vplot$fc[vplot$fc != 0]), max(vplot$fc)), ylimits = c(min(vplot$pval), 1))  #c(min(vplot$pval[vplot$pval != 0]),1)
    sub = vplot %>%
        filter(log2(fc) > 0 & pval < 10^-60) %>%
        arrange(desc(log2(fc))) %>%
        pull(SYMBOL)
    sub
    gene_sig[[sprintf("%s_up", i)]] = sub
    highlight_genes = sub #unique(c(head(sub, 25), tail(sub, 25)))
    # sub = (log2(vplot$fc) > 0.5 | log2(vplot$fc) < -0.5)  & vplot$pval < 0.05
    # sub = vplot[sub,] %>% arrange(desc(log2(fc)))
    # sub1 = sub %>% pull(SYMBOL) %>% head(25) #vplot$SYMBOL %in% as.character(sub2$SYMBOL)[1:200]
    # sub2 = sub %>% pull(SYMBOL) %>% tail(25)
    highlight_filter = vplot$SYMBOL %in% highlight_genes
    # sub = vplot$SYMBOL %in% c(sub1, sub2)#as.character(sub2$SYMBOL)[1:50]
    library(ggrepel)
    print(p + geom_text_repel(data = vplot[highlight_filter,], aes(x = fc, y = pval, label = SYMBOL)) + ggtitle(i) + xlab(sprintf("FC %s", i)))
}
dev.off()


```

use pseudobulk and AUC so that keep only genes for which the third is at least twice 
```{r}
pseudobulk_orig = readRDS(sprintf("%s/igt1_96_pseudobulk_byclusterannotationlevel2_log1pCP10K.Rds", "/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/analysis/data_integration/IGT1_96/TF_analysis/20260319"))

# pseudobulk_orig@meta.data %>% filter(grepl("CD4",annotation_level2), annotation_level2 %in% c())
pseudobulk = pseudobulk_orig[,pseudobulk_orig@meta.data %>% filter(annotation_level2 %in% c("CD4.R", "CD4.S", "CD4.T", "CD4.U", "CD4.X", "CD4.Y", "CD4.H", "CD4.I")) %>% rownames()]

library(dplyr)
library(ggplot2)
library(tibble)

gene = "Cd48"
plot_df <- pseudobulk[["RNA"]]$data[gene, , drop = FALSE] %>%
  as.data.frame() %>%
  rownames_to_column("gene") %>%
  tidyr::pivot_longer(
    cols = -gene,
    names_to = "cluster",
    values_to = "expression"
  )
ggplot(plot_df, aes(x = cluster, y = expression, fill = cluster)) +
  geom_col(color = "black") +
  scale_fill_manual(values = mypal_level2) +
  theme_classic() +
  labs(
    x = NULL,
    y = "pseudobulk expression"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) + ggtitle(gene)
 
names(gene_sig)
gene_sig_curated = list()
min_expr = 0.25

# gene_sig["RS_vs_TUXYHI_up"] %>% unlist() %>% as.character() %>% sort()
mat <- pseudobulk[["RNA"]]$data[gene_sig["RS_vs_TUXYHI_up"] %>% unlist() %>% as.character(),]
target_clusters <- c("CD4.R", "CD4.S")
other_clusters  <- setdiff(colnames(mat), target_clusters)
max_other <- apply(mat[, other_clusters, drop = FALSE], 1, max)
enriched_genes <- rownames(mat)[
  (mat[, target_clusters[1]] >= 1.5 * max_other) & mat[, target_clusters[1]] > min_expr &
  (mat[, target_clusters[2]] >= 1.5 * max_other) & mat[, target_clusters[2]] > min_expr
]
enriched_genes %>% sort()
gene_sig_curated[["RS_vs_TUXYHI_up"]] = enriched_genes

# gene_sig["TU_vs_RSXYHI_up"] %>% unlist() %>% as.character() %>% sort()
mat <- pseudobulk[["RNA"]]$data[gene_sig["TU_vs_RSXYHI_up"] %>% unlist() %>% as.character(),]
target_clusters <- c("CD4.T", "CD4.U")
other_clusters  <- setdiff(colnames(mat), target_clusters)
max_other <- apply(mat[, other_clusters, drop = FALSE], 1, max)
enriched_genes <- rownames(mat)[
  (mat[, target_clusters[1]] >= 1.5 * max_other) & mat[, target_clusters[1]] > min_expr &
  (mat[, target_clusters[2]] >= 1.5 * max_other) & mat[, target_clusters[2]] > min_expr
]
enriched_genes %>% sort()
gene_sig_curated[["TU_vs_RSXYHI_up"]] = enriched_genes

# gene_sig["XY_vs_RSTUHI_up"] %>% unlist() %>% as.character() %>% sort()
mat <- pseudobulk[["RNA"]]$data[gene_sig["XY_vs_RSTUHI_up"] %>% unlist() %>% as.character(),]
target_clusters <- c("CD4.X", "CD4.Y")
other_clusters  <- setdiff(colnames(mat), target_clusters)
max_other <- apply(mat[, other_clusters, drop = FALSE], 1, max)
enriched_genes <- rownames(mat)[
  (mat[, target_clusters[1]] >= 1.5 * max_other) & mat[, target_clusters[1]] > min_expr &
  (mat[, target_clusters[2]] >= 1.5 * max_other) & mat[, target_clusters[2]] > min_expr
]
enriched_genes %>% sort()
gene_sig_curated[["XY_vs_RSTUHI_up"]] = enriched_genes

# gene_sig["HI_vs_RSTUXY_up"] %>% unlist() %>% as.character() %>% sort()
mat <- pseudobulk[["RNA"]]$data[gene_sig["HI_vs_RSTUXY_up"] %>% unlist() %>% as.character(),]
target_clusters <- c("CD4.H", "CD4.I")
other_clusters  <- setdiff(colnames(mat), target_clusters)
max_other <- apply(mat[, other_clusters, drop = FALSE], 1, max)
enriched_genes <- rownames(mat)[
  (mat[, target_clusters[1]] >= 1.5 * max_other) & mat[, target_clusters[1]] > min_expr &
  (mat[, target_clusters[2]] >= 1.5 * max_other) & mat[, target_clusters[2]] > min_expr
]
enriched_genes %>% sort()
gene_sig_curated[["HI_vs_RSTUXY_up"]] = enriched_genes

#ensure_directory(sprintf("%s/tip_signatures", out_dir))
for (i in names(gene_sig_curated)) {
    write.table(gene_sig_curated[[i]], file = sprintf("%s/tip_signatures/%s.txt", out_dir, i), sep = "\t",quote = F, row.names = F, col.names = F)
}


```

Plotting on the volcano plot the tip signatures
```{r}
Vplot = function (vplot, xlab = "FC", xlimits = c(0.1, 10), ylimits = c(10^-300, 1), color = "grey", size = 1, xlab = "FC", ylab = "p value") 
{
    requireNamespace("ggplot2", quietly = TRUE)
    requireNamespace("ggrastr", quietly = TRUE)
    requireNamespace("scales", quietly = TRUE)
    p = ggplot(data = vplot) + ggrastr::geom_point_rast(aes(x = .data$fc, 
        y = .data$pval), colour = color, alpha = I(1), size = size, 
        raster.dpi = 100) + scale_x_continuous(trans = scales::log_trans(10), 
        limits = xlimits) + scale_y_continuous(trans = scales::log_trans(10), 
        breaks = scales::trans_breaks("log10", function(x) 10^x), 
        labels = scales::trans_format("log10", scales::math_format(10^.x)), 
        limits = ylimits) + annotation_logticks(sides = "b") + 
        xlab(xlab) + theme_bw() + ylab(ylab) + theme(axis.text.x = element_text(size = 20, 
        angle = 0, hjust = 0.5), axis.text.y = element_text(size = 20), 
        legend.text = element_text(size = 20), axis.title.x = element_text(size = 20), 
        axis.title.y = element_text(size = 20))
    return(p)
}


comp_to_plot = c("RS_vs_TUXYHI", "TU_vs_RSXYHI", "XY_vs_RSTUHI", "HI_vs_RSTUXY")
cols = c("darkorchid", "violetred2", "burlywood2","darkseagreen2" )
names(cols) = comp_to_plot
tmp = ttlist
pdf(sprintf("%s/Volcanos_Tips_signature.pdf", out_dir), 10, 10, useDingbats = F)
for (i in comp_to_plot) {
    print(i)
    vplot = na.omit(data.frame(SYMBOL = tmp[[i]]$SYMBOL, fc = 2^tmp[[i]]$logFC, pval = tmp[[i]]$P.Value + 10^-300, AveExpr = 2^(tmp[[i]]$AveExpr))) #dealing with NA
    # vplot = data.frame(SYMBOL = tmp[[i]]$SYMBOL, fc = 2^tmp[[i]]$logFC, pval = tmp[[i]]$P.Value+10^-300, AveExpr = 2^(tmp[[i]]$AveExpr))
    p = Vplot(vplot = vplot, xlab = names(tmp)[i], xlimits = c(min(vplot$fc[vplot$fc != 0]), max(vplot$fc)), ylimits = c(min(vplot$pval), 1), color = cols[i], size = 2)  #c(min(vplot$pval[vplot$pval != 0]),1)
    highlight_genes = gene_sig_curated[[sprintf("%s_up", i)]]
    highlight_filter = vplot$SYMBOL %in% highlight_genes
    # sub = vplot$SYMBOL %in% c(sub1, sub2)#as.character(sub2$SYMBOL)[1:50]
    library(ggrepel)
    print(p)
    print(p + geom_text_repel(data = vplot[highlight_filter,], aes(x = fc, y = pval, label = SYMBOL)) + ggtitle(i) + xlab(sprintf("FC %s", i)))
}
dev.off()

```

**Difference between Q.Th1 and RS tip, Q.Th2 and TU tip etc?**
--> th_modulescores_withannot.Rds
```{r}
so_orig = readRDS("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/igt1_96_withtotalvi20260206_clean.Rds") 
so = so_orig[,so_orig@meta.data %>% filter(, annotation_level1 == "CD4", organ_simplified != "thymus") %>% row.names()]

th_scores = readRDS("module_score_matrix.Rds") %>% t()
th_scores[1:2,]
thr_th1 = 0
thr_th2 = 0.25
thr_th17 = 0.25
thr_tfhchron = 0

# correct binarization
th_scores_bin <- data.frame(
  row.names = rownames(th_scores),
  Th1 = th_scores[, "Th1 program"] > thr_th1,
  Th2 = th_scores[, "Th2 program"] > thr_th2,
  Th17 = th_scores[, "Th17 program"] > thr_th17,
  Tfhchron = th_scores[, "Tfh/Tchron program"] > thr_tfhchron
)

cell_class <- apply(th_scores_bin, 1, function(x) {
  pos <- colnames(th_scores_bin)[as.logical(x)]
  if (length(pos) == 0) "None" else paste(pos, collapse = "")
})

all(rownames(th_scores_bin) == names(cell_class))
th_scores_bin$program = cell_class

th_scores = as.data.frame(th_scores)
all(rownames(th_scores) == names(cell_class))
th_scores$program = cell_class

th_scores_bin %>% group_by(program) %>% summarize(n = n()) %>% mutate(percent = n/sum(n)*100)
th_scores %>% group_by(program) %>% summarize(n = n()) %>% mutate(percent = n/sum(n)*100)

saveRDS(th_scores, sprintf("../%s/th_modulescores_withannot.Rds", out_dir))

```

Barplot of  polyfunctional
```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(patchwork)

plot_df <- th_scores_bin %>%
  group_by(program) %>%
  summarise(n = n(), .groups = "drop") %>%
  mutate(percent = n / sum(n) * 100)

# program_order <- plot_df %>%
#   arrange(desc(percent)) %>%
#   pull(program)

program_order <- c(
  "None",
  "Th1", "Th2", "Th17", "Tfhchron",
  "Th1Th2", "Th1Th17", "Th1Tfhchron", "Th2Th17", "Th2Tfhchron", "Th17Tfhchron",
  "Th1Th2Th17", "Th1Th2Tfhchron", "Th1Th17Tfhchron", "Th2Th17Tfhchron",
  "Th1Th2Th17Tfhchron"
)

plot_df$program <- factor(plot_df$program, levels = program_order)

# heatmap_df <- expand.grid(
#   row = c("Th1", "Th2", "Th17", "Tfhchron"),
#   program = program_order,
#   stringsAsFactors = FALSE
# ) %>%
#   mutate(
#     filled = case_when(
#       row == "Th1"      & grepl("Th1", program) ~ 1,
#       row == "Th2"      & grepl("Th2", program) ~ 1,
#       row == "Th17"     & grepl("Th17", program) ~ 1,
#       row == "Tfhchron" & grepl("Tfhchron", program) ~ 1,
#       TRUE ~ 0
#     )
#   )

heatmap_df <- expand.grid(
  row = c("Th1", "Th2", "Th17", "Tfhchron"),
  program = program_order,
  stringsAsFactors = FALSE
) %>%
  rowwise() %>%
  mutate(
    members = list(c(
      if (grepl("Tfhchron", program)) "Tfhchron" else NULL,
      if (grepl("Th17", program)) "Th17" else NULL,
      if (grepl("Th2", program)) "Th2" else NULL,
      if (grepl("Th1", program)) "Th1" else NULL
    )),
    filled = as.integer(row %in% members)
  ) %>%
  ungroup() %>%
  select(-members)

heatmap_df$program <- factor(heatmap_df$program, levels = program_order)
heatmap_df$row <- factor(heatmap_df$row, levels = rev(c("Th1", "Th2", "Th17", "Tfhchron")))

p_bar <- ggplot(plot_df, aes(x = program, y = percent)) +
  geom_col(fill = "grey70", color = "black") +
  geom_text(
    aes(label = sprintf("%.1f", percent)),
    vjust = -0.25,
    size = 2.8
  ) +
  labs(
    x = NULL,
    y = "% of cells"
  ) +
  theme_classic(base_size = 16) +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.y = element_text(size = 16),
    axis.text.y = element_text(size = 14)
  ) +
  expand_limits(y = max(plot_df$percent) * 1.08)

p_heat <- ggplot(heatmap_df, aes(x = program, y = row, fill = factor(filled))) +
  geom_tile(color = "black", linewidth = 0.4) +
  scale_fill_manual(values = c("0" = "white", "1" = "black"), guide = "none") +
  labs(
    x = NULL,
    y = NULL
  ) +
  theme_classic(base_size = 16) +
  theme(
    axis.line = element_blank(),
    axis.ticks = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 12),
    axis.text.y = element_text(size = 14)
  )

p = p_bar / p_heat + plot_layout(heights = c(4, 1.2))

pdf(sprintf("%s/barplot_polyfunctional.pdf", out_dir), 8,8)
p
dev.off()


###########

so@meta.data %>% count(annotation_level2, annotation_level2_group)
so@meta.data %>% count(annotation_level2_group, annotation_level2)

anndata = readRDS("CD4Q_CD4G_cell_ids_with_program_labels.rds")
anndata %>% count(annotation_level2)
anndata_g = anndata %>% filter(annotation_level2 == "CD4.G")
anndata_q = anndata %>% filter(annotation_level2 == "CD4.Q")

so$program = NA
so$program[anndata_g$cell_id] = anndata_g$g_program
so$program[anndata_q$cell_id] = anndata_q$q_program

so@meta.data %>% filter(annotation_level2 %in% c("CD4.G", "CD4.Q")) %>% count(annotation_level2, program)


```

Make seurat for DGE limma with program annotation
```{r}
th_scores = readRDS(sprintf("%s/th_modulescores_withannot.Rds", out_dir))

so
so2 = so[,rownames(th_scores)]
so2@meta.data = data.frame(so2@meta.data, th_scores)
so2@meta.data %>% count(annotation_level2, program)
so2$annotation_level2_program = paste(so2$annotation_level2, so2$program, sep = ".")
so2@meta.data %>% count(annotation_level2, annotation_level2_program)

# saveRDS(so2, file = sprintf("%s/CD4_GHIQRSTUXY_igt1_96_withtotalvi20260206_clean_withThPrograms.Rds", out_dir))

so = readRDS(file = sprintf("%s/CD4_GHIQRSTUXY_igt1_96_withtotalvi20260206_clean_withThPrograms.Rds", out_dir))
so$annotation_level2_program[so$annotation_level2 %in% c("CD4.H")] = "CD4.H"
so$annotation_level2_program[so$annotation_level2 %in% c("CD4.I")] = "CD4.I"
so$annotation_level2_program[so$annotation_level2 %in% c("CD4.R")] = "CD4.R"
so$annotation_level2_program[so$annotation_level2 %in% c("CD4.S")] = "CD4.S"
so$annotation_level2_program[so$annotation_level2 %in% c("CD4.T")] = "CD4.T"
so$annotation_level2_program[so$annotation_level2 %in% c("CD4.U")] = "CD4.U"
so$annotation_level2_program[so$annotation_level2 %in% c("CD4.X")] = "CD4.X"
so$annotation_level2_program[so$annotation_level2 %in% c("CD4.Y")] = "CD4.Y"


so@meta.data %>% count(annotation_level2, annotation_level2_program)
so@meta.data %>% filter(annotation_level2_program %in% c("CD4.R", "CD4.S", "CD4.T", "CD4.U", "CD4.X", "CD4.Y", "CD4.Q.Th1", "CD4.Q.Th2", "CD4.Q.Th17")) %>% count(annotation_level2, annotation_level2_program)
so2 = so[,so@meta.data %>% filter(annotation_level2_program %in% c("CD4.R", "CD4.S", "CD4.T", "CD4.U", "CD4.X", "CD4.Y", "CD4.Q.Th1", "CD4.Q.Th2", "CD4.Q.Th17")) %>% rownames()]
so2$annotation_level2_program.IGTHT = paste(so2$annotation_level2_program, so2$IGTHT, sep = ".")
saveRDS(so2, file = sprintf("%s/CD4_QRSTUXY_igt1_96_withtotalvi20260206_clean_annotation_level2_program.Rds", out_dir))

so2 = so[,so@meta.data %>% filter(annotation_level2_program %in% c("CD4.R", "CD4.S", "CD4.T", "CD4.U", "CD4.X", "CD4.Y", "CD4.H", "CD4.I", "CD4.G.Th1", "CD4.G.Th2", "CD4.G.Th17", "CD4.G.Tfhchron")) %>% rownames()]
so2@meta.data %>% count(annotation_level2, annotation_level2_program)
so2$annotation_level2_program.IGTHT = paste(so2$annotation_level2_program, so2$IGTHT, sep = ".")
saveRDS(so2, file = sprintf("%s/CD4_GHIRSTUXY_igt1_96_withtotalvi20260206_clean_annotation_level2_program.Rds", out_dir))

```


**Figure 3 - CD4 GP analysis**
```{r}
mdata_orig = so@meta.data

#option 1 (best): import F1 and L1 from the fit object: 
L1 = readRDS(sprintf('/Users/david/google_drive_uchicago/ImmgenT/analysis/data_integration/IGT1_96/topic/flashier20250215_alldata_backfit/L1.Rds')) #cells x factor
F1 = readRDS(sprintf("/Users/david/google_drive_uchicago/ImmgenT/analysis/data_integration/IGT1_96/topic/flashier20250215_alldata_backfit/F1.Rds")) #gene x factor

mdata_orig %>% select(IGT, exp_name) %>% unique()

mdata_orig %>% count(organ_simplified)

#c("CD4.R", "CD4.S", "CD4.T", "CD4.U", "CD4.X", "CD4.Y", "CD4.H", "CD4.I")

group_list = list(
#1    
    list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.R", "CD4.S")) %>% pull(cellID), 
         group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.T", "CD4.U", "CD4.X", "CD4.Y", "CD4.H", "CD4.I")) %>% pull(cellID),  
         title = "CD4.RS vs CD4.TUXYHI"),
    
    list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.T", "CD4.U")) %>% pull(cellID), 
         group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.R", "CD4.S", "CD4.X", "CD4.Y", "CD4.H", "CD4.I")) %>% pull(cellID),  
         title = "CD4.TU vs CD4.RSXYHI"),
    
    list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.X", "CD4.Y")) %>% pull(cellID), 
         group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.R", "CD4.S", "CD4.T", "CD4.U", "CD4.H", "CD4.I")) %>% pull(cellID),  
         title = "CD4.XY vs CD4.RSTUHI"),
    
    list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.H", "CD4.I")) %>% pull(cellID), 
         group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.R", "CD4.S", "CD4.T", "CD4.U", "CD4.X", "CD4.Y")) %>% pull(cellID),  
         title = "CD4.HI vs CD4.RSTUXY"),
#2    
    list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.R", "CD4.S")) %>% pull(cellID), 
         group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.T", "CD4.U", "CD4.X", "CD4.Y")) %>% pull(cellID),  
         title = "CD4.RS vs CD4.TUXY"),
    
    list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.T", "CD4.U")) %>% pull(cellID), 
         group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.R", "CD4.S", "CD4.X", "CD4.Y")) %>% pull(cellID),  
         title = "CD4.TU vs CD4.RSXY"),
    
    list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.X", "CD4.Y")) %>% pull(cellID), 
         group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.R", "CD4.S", "CD4.T", "CD4.U")) %>% pull(cellID),  
         title = "CD4.XY vs CD4.RSTU"),
#3    
        list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.S")) %>% pull(cellID), 
         group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c( "CD4.U", "CD4.Y")) %>% pull(cellID),  
         title = "CD4.S vs CD4.UY"),
    
    list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.U")) %>% pull(cellID), 
         group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.S", "CD4.Y")) %>% pull(cellID),  
         title = "CD4.U vs CD4.SY"),
    
    list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c( "CD4.Y")) %>% pull(cellID), 
         group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.S", "CD4.U")) %>% pull(cellID),  
         title = "CD4.Y vs CD4.SU"),
#4    
    list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.R", "CD4.S") & organ_simplified %in% c("colon LP", "small intestine LP")) %>% pull(cellID), 
         group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.T", "CD4.U", "CD4.X", "CD4.Y", "CD4.H", "CD4.I") & organ_simplified %in% c("colon LP", "small intestine LP")) %>% pull(cellID),  
         title = "CD4.RS vs CD4.TUXYHI in gut"),
    
    list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.T", "CD4.U") & organ_simplified %in% c("colon LP", "small intestine LP")) %>% pull(cellID), 
         group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.R", "CD4.S", "CD4.X", "CD4.Y", "CD4.H", "CD4.I") & organ_simplified %in% c("colon LP", "small intestine LP")) %>% pull(cellID),  
         title = "CD4.TU vs CD4.RSXYHI in gut"),
    
    list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.X", "CD4.Y") & organ_simplified %in% c("colon LP", "small intestine LP")) %>% pull(cellID), 
         group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.R", "CD4.S", "CD4.T", "CD4.U", "CD4.H", "CD4.I") & organ_simplified %in% c("colon LP", "small intestine LP")) %>% pull(cellID),  
         title = "CD4.XY vs CD4.RSTUHI in gut"),
    
    list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.H", "CD4.I") & organ_simplified %in% c("colon LP", "small intestine LP")) %>% pull(cellID), 
         group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.R", "CD4.S", "CD4.T", "CD4.U", "CD4.X", "CD4.Y") & organ_simplified %in% c("colon LP", "small intestine LP")) %>% pull(cellID),  
         title = "CD4.HI vs CD4.RSTUXY in gut"),

#5    
    list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.R", "CD4.S") & organ_simplified %in% c("spleen", "LN")) %>% pull(cellID), 
         group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.T", "CD4.U", "CD4.X", "CD4.Y", "CD4.H", "CD4.I") & organ_simplified %in% c("spleen", "LN")) %>% pull(cellID),  
         title = "CD4.RS vs CD4.TUXYHI in SLO"),
    
    list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.T", "CD4.U") & organ_simplified %in% c("spleen", "LN")) %>% pull(cellID), 
         group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.R", "CD4.S", "CD4.X", "CD4.Y", "CD4.H", "CD4.I") & organ_simplified %in% c("spleen", "LN")) %>% pull(cellID),  
         title = "CD4.TU vs CD4.RSXYHI in SLO"),
    
    list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.X", "CD4.Y") & organ_simplified %in% c("spleen", "LN")) %>% pull(cellID), 
         group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.R", "CD4.S", "CD4.T", "CD4.U", "CD4.H", "CD4.I") & organ_simplified %in% c("spleen", "LN")) %>% pull(cellID),  
         title = "CD4.XY vs CD4.RSTUHI in SLO"),
    
    list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.H", "CD4.I") & organ_simplified %in% c("spleen", "LN")) %>% pull(cellID), 
         group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.R", "CD4.S", "CD4.T", "CD4.U", "CD4.X", "CD4.Y") & organ_simplified %in% c("spleen", "LN")) %>% pull(cellID),  
         title = "CD4.HI vs CD4.RSTUXY in SLO"),

#6   
    list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.G")) %>% pull(cellID), 
         group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.R", "CD4.S", "CD4.T", "CD4.U", "CD4.X", "CD4.Y", "CD4.H", "CD4.I") ) %>% pull(cellID),  
         title = "CD4.G vs CD4.RSTUXYHI"),
    
    list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.Q")) %>% pull(cellID), 
         group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.R", "CD4.S", "CD4.T", "CD4.U", "CD4.X", "CD4.Y", "CD4.H", "CD4.I")) %>% pull(cellID),  
         title = "CD4.Q vs CD4.RSTUXYHI"),

    list(group1 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.Q")) %>% pull(cellID), 
         group2 = mdata_orig %>% filter(annotation_level1 %in% c("CD4") & annotation_level2 %in% c("CD4.R", "CD4.S")) %>% pull(cellID),  
         title = "CD4.Q vs CD4.RS")
    
)

results_list = list()

for (i in 1:length(group_list)) {
    print(group_list[[i]]$title)
    results_list[[group_list[[i]]$title]] = FlashierDGE(F1, L1, group_list[[i]]$group1, group_list[[i]]$group2, title_plot = group_list[[i]]$title)
}

results_list[["CD4.RS vs CD4.TUXY"]]

```

GP in tips
```{r}

pdf(sprintf("%s/FlashierDGE_volcano_GP_tips1.pdf", out_dir), 15,15, useDingbats = F)
results_list[[1]]$p1
results_list[[2]]$p1
results_list[[3]]$p1
results_list[[4]]$p1
dev.off()

pdf(sprintf("%s/FlashierDGE_volcano_GP_tips2.pdf", out_dir), 15,15, useDingbats = F)
results_list[[5]]$p1
results_list[[6]]$p1
results_list[[7]]$p1
dev.off()

pdf(sprintf("%s/FlashierDGE_volcano_GP_tips3.pdf", out_dir), 15,15, useDingbats = F)
results_list[[8]]$p1
results_list[[9]]$p1
results_list[[10]]$p1
dev.off()

pdf(sprintf("%s/FlashierDGE_volcano_GP_tipsGut.pdf", out_dir), 15,15, useDingbats = F)
results_list[[11]]$p1
results_list[[12]]$p1
results_list[[13]]$p1
results_list[[14]]$p1
dev.off()

pdf(sprintf("%s/FlashierDGE_volcano_GP_tipsSLO.pdf", out_dir), 15,15, useDingbats = F)
results_list[[15]]$p1
results_list[[16]]$p1
results_list[[17]]$p1
results_list[[18]]$p1
dev.off()

pdf(sprintf("%s/FlashierDGE_volcano_GP_Midland.pdf", out_dir), 15,15, useDingbats = F)
results_list[[19]]$p1
results_list[[20]]$p1
results_list[[21]]$p1
dev.off()

results_list[[11]]$diff_genes["Tbx21",]
results_list[[12]]$diff_genes["Gata3",]
results_list[[13]]$diff_genes["Rorc",]
results_list[[14]]$diff_genes["Bcl6",]

results_list[[11]]$diff_genes["Ifng",]
results_list[[12]]$diff_genes["Il5",]
results_list[[13]]$diff_genes["Il17a",]
results_list[[14]]$diff_genes["Il21",]
```



```{r}

tt = results_list[5:7]

results_list[[1]]$diff_genes["Tbx21",]
results_list[[2]]$diff_genes["Gata3",]
results_list[[3]]$diff_genes["Rorc",]
results_list[[4]]$diff_genes["Bcl6",]

results_list[[5]]$diff_genes["Tbx21",]
results_list[[6]]$diff_genes["Gata3",]
results_list[[7]]$diff_genes["Rorc",]

results_list[[5]]$p1
results_list[[6]]$p1
results_list[[7]]$p1

results_list[[8]]$p1
results_list[[9]]$p1
results_list[[10]]$p1

plot(results_list[[8]]$diff_genes$log2FC, results_list[[9]]$diff_genes$log2FC)

comp_filt = names(tt)
# comp_filt = comp_filt[grepl("_vs_others", comp_filt)]

# #filter gene names
# genes_to_label <- c()
# for (i in comp_filt) {
#   new_genes <- tt[[i]]$diff_genes %>%
#     arrange(desc(log2FC)) %>%
#     # filter(adj.P.Val < 0.05) %>%
#     pull(SYMBOL)
#   new_genes <- setdiff(new_genes, genes_to_label)
#   genes_to_label <- c(genes_to_label, head(new_genes, 20))
# }

genes_to_label <- c()
for (i in comp_filt) {
  df <- tt[[i]]$diff_genes #%>% filter(adj.P.Val < 0.05)
  # top up
  up_genes <- df %>% filter(log2FC > 0.249) %>%
    arrange(desc(log2FC)) %>%
    pull(SYMBOL)
  # top down
  # down_genes <- df %>%
  #   arrange(log2FC) %>%
  #   pull(SYMBOL)
  # remove genes already selected
  up_genes <- setdiff(up_genes, genes_to_label)
  # down_genes <- setdiff(down_genes, genes_to_label)

  # add top 20 each
  genes_to_label <- c(
    genes_to_label,
    head(up_genes, 100)#, head(down_genes, 10)
  )
}

genes_to_label <- genes_to_label[
  !grepl(
    "^Gm|^mt-|^Rpl|^Rps|^Snor|^Mir|^Linc|^Rp[sl]|-ps|Rik$|^Hbb",
    genes_to_label,
    ignore.case = TRUE
  )
]

#plot heatmap
# mat_coef <- sapply(tt, function(tt) tt$logFC)
mat_coef <- Reduce(cbind, lapply(tt, function(x) x$diff_genes$log2FC))
rownames(mat_coef) <- tt[[1]]$diff_genes$SYMBOL
colnames(mat_coef) = make.names(names(tt))
mat_plot = mat_coef[genes_to_label, ]
range(mat_plot)

library(ComplexHeatmap)
library(circlize)

col_fun <- circlize::colorRamp2(
  c(-3, 0, 3),
  c("#4575B4", "white", "#D73027")
)

# pdf(sprintf("%s/heatmap_memory_vs_acute_SLO_LCMV.pdf", out_dir), width = 3, 8, useDingbats = F)
# pdf(sprintf("%s/Heatmap_LCMV_SLO_memory_vs_acute_IGT363840.pdf", out_dir2), width = 3, 8, useDingbats = F)
Heatmap(
  mat_plot,
  name = "log2FC",
  col = col_fun,
  cluster_columns = FALSE,
  cluster_rows = T,
    row_names_gp = gpar(fontsize = 6),
      column_names_gp = gpar(fontsize = 6),
  column_title = "memory vs acute across clusters in LCMV SLO"
)
# dev.off()

# Access the plots and data
results$p1  # First plot (MA plot of factors)
results$p2  # Second plot (MA plot of genes)
results$diff_factors  # table for fold changes of gene programs
results$diff_genes %>% arrange(desc(log2FC)) # table for fold changes of genes


```



**T-RBI**
--> see CD4_TRBI.Rmd
**Memory Figure **
--> see CD4_TRBI.Rmd








