---
title: "CD4_misc"
output: html_document
date: "2025-05-28"
author: "David Zemmour"
editor_options: 
  chunk_output_type: console
---

**Initialize: load libraries and custom functions**

```{bash}
sinteractive --exclusive --time=12:00:00 --account=pi-zemmour --partition=jfkfloor2 --qos=jfkfloor2 --nodes=1 
sinteractive --exclusive --time=12:00:00 --account=pi-zemmour --partition=beagle3-bigmem --nodes=1 
sinteractive --exclusive --time=12:00:00 --account=pi-zemmour --partition=zemmour-hm --qos=zemmour --nodes=1 
sinteractive --time=03:00:00 --account=pi-zemmour --partition=caslake --nodes=1 --mem=96GB

module load python/anaconda-2022.05 
module load hdf5/1.12.0 #for BPCells but also others
module load openblas/0.3.13
source activate /project/zemmour/david/envs/scvi_20240315
#source activate /project/zemmour/david/envs/r443
module load openblas/0.3.13

cd /project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/CD4
```


```{r}
setwd("~/google_drive_uchicago/ImmgenT/analysis/data_integration/IGT1_96/CD4")
libs = c("ZemmourLib","forcats", "gridExtra","patchwork", "Matrix", "Seurat", "ggplot2", "dplyr", "reshape2", "ggrastr","scattermore", "RColorBrewer", "pals", "scales", "pheatmap", "rafalib") 
sapply(libs, function(x) suppressMessages(suppressWarnings(library(x, character.only = TRUE, quietly = T, warn.conflicts  = F))))
#source("/project/jfkfloor2/zemmourlab/david/immgent/immgent_integration_git/custom_functions_david.R")

ColorRamp = rev(colorRampPalette(brewer.pal(n = 7,name = "RdYlBu"))(100))
ColorRamp = rev(colorRampPalette(c('red','white','blue'))(20))
ColorRamp = rev(rainbow(10, end = 4/6))
ColorRamp = rev(viridis(100))
ColorRamp = brewer.pal(9, "Blues")

#mycol = c("black","blue","red","darkgreen", "orange","purple", "cyan", "greenyellow",    "salmon", "magenta","pink", "tan", "brown") # c("magenta", "red",  "darkgreen", "cyan", "blue", "blue", "blue", "black", "blue", "blue", "blue", "blue", "orange")

mypal_level2 = ZemmourLib::immgent_colors$level2
mypal_level2["not classified"] = "black"
# names(mypal_level2) = gsub("nonconv", replacement = "Tz",names(mypal_level2) )
# mypal_level2 = immgent_colors$level2_option2
# mypal_level2["not classified"] = "black"
mypal_level1 = immgent_colors$level1
mypal_level1["not classified"] = "black"
# names(mypal_level1) = gsub("nonconv", replacement = "Tz",names(mypal_level1) )
mypal_organ = immgent_colors$organ_simplified
mypal_level2group = c("resting" = "blue", "activated" = "red", "miniverse" = "darkgreen", "proliferating" = "black", "other" = "grey", "preT" = "grey" )

#Large Color palette
library("pals")
n = 70
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
mypal1 = unique(unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals))))
mypal1 = mypal1[-4]
mypal = c(glasbey(), polychrome(), mypal1)
names(mypal) = NULL
#pal.bands(mypal, main="Colormap suggestions")
# mypal_organ = setNames(mypal, unique(so$organ_simplified))
# mypal_organ = mypal_organ[!is.na(names(mypal_organ))]
# mypal_level1 = setNames(mypal, unique(so$annotation_level1))
# mypal_level1 = mypal_level1[!is.na(names(mypal_level1))]
# mypal_level2 = setNames(mypal, unique(so_orig$annotation_level2))
# mypal_level2 = mypal_level2[!is.na(names(mypal_level2))]
# mypal_level2["CD4_cl18"] = "blue"
mypal_igt = setNames(mypal, unique(so_orig$IGT))
mypal_igt = mypal_igt[!is.na(names(mypal_igt))]

```

```{r}
# fig_dir = "~/google_drive_uchicago/ImmgenT/20250109_freeze/paper_cosmology/figures/bits"
# fig_dir = "/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/paper_cosmology/figures/bits"

out_dir = "~/google_drive_uchicago/ImmgenT/analysis/data_integration/IGT1_96/CD4/misc_plots" 
out_dir = "/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/CD4/misc_plots"

```


**Most up-to-date data**
on Midway
```{r}
so_orig = readRDS("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/igt1_96_withtotalvi20260206_clean.Rds") 

so = so_orig[,so_orig@meta.data %>% filter(organ_simplified != "thymus") %>% row.names()]
so = NormalizeData(so, assay = "RNA", normalization.method = "LogNormalize")


```

on Desktop
```{r}
setwd("~/google_drive_uchicago/ImmgenT/analysis/data_integration/IGT1_96/")
so_orig = readRDS("igt1_96_withtotalvi20260206_clean_ADTonly.Rds")

so = so_orig[,so_orig@meta.data %>% filter(annotation_level1 == "CD4" & organ_simplified != "thymus") %>% row.names()]
# so = NormalizeData(so, assay = "RNA", normalization.method = "LogNormalize")
# so = NormalizeData(so, assay = "ADT", normalization.method = "LogNormalize")

```

**FCFC plots**
Tfh/Tex
```{r}
source('/Users/david/Library/CloudStorage/GoogleDrive-david.puyraimond@gmail.com/My Drive/ImmgenT/analysis/immgent_integration_git/custom_functions_david.R')

tt_list = readRDS('/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/analysis/data_integration/IGT1_96/CD4/DGE_limma/20250109/ttlist_OneVsAll.Rds')

names(tt_list)


# Create the vplot dataframe with fold changes
vplot = data.frame(fc1 = exp(tt_list$CD4_cl13_vs_All$logFC), fc2 = exp(tt_list$CD4_cl14_vs_All$logFC), SYMBOL = tt_list$CD4_cl13_vs_All$SYMBOL)

# Identify genes with large fold changes (>2)
genes_to_plot = abs(tt_list$CD4_cl13_vs_All$logFC) > 0.69 | abs(tt_list$CD4_cl14_vs_All$logFC) > 0.69
# genes_to_label = vplot[genes_to_plot, ]
genes_to_label = vplot$SYMBOL[genes_to_plot]

# Create the plot
FCFCplot(vplot, 
         xlab = "FC for CD4_cl13", 
         ylab = "FC for CD4_cl14", 
         main = "Fold Change vs Fold Change", 
         genes_to_label = genes_to_label)

FCFCplot(vplot, 
         xlab = "FC for CD4_cl13", 
         ylab = "FC for CD4_cl14", 
         main = "Fold Change vs Fold Change", 
         genes_to_label = NULL)

vplot = data.frame(fc1 = exp(tt_list$CD8_cl10_vs_All$logFC), fc2 = exp(tt_list$CD8_cl7_vs_All$logFC), SYMBOL = tt_list$CD8_cl10_vs_All$SYMBOL)
# Identify genes with large fold changes (>2)
genes_to_plot = abs(tt_list$CD8_cl10_vs_All$logFC) > 0.69 | abs(tt_list$CD8_cl7_vs_All$logFC) > 0.69
genes_to_label = vplot[genes_to_plot, ]
# Create the plot
FCFCplot(vplot, 
         xlab = "FC for CD8_cl10", 
         ylab = "FC for CD8_cl7", 
         main = "Fold Change vs Fold Change", 
         printgeomtext = T, 
         genes_to_label = genes_to_label)



```



**MDE plots**
```{r}
pdf("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/CD4/MDE_clusters_20250527.pdf", width = 10, height = 10, useDingbats = F)
p = DimPlot(so, reduction = "mde_incremental", group.by = "annotation_level2", split.by = NULL, raster = T, raster.dpi = c(512,512), label = T, pt.size = 1) + scale_color_manual(values = mypal_level2)
p
p + NoLegend()

p = DimPlot(so, reduction = "mde_incremental", group.by = "annotation_level2", split.by = NULL, raster = T, raster.dpi = c(512,512), label = F, pt.size = 1) + scale_color_manual(values = mypal_level2)
p + NoLegend()

p = MyDimPlotHighlight(seurat_object = so, umap_to_plot = "mde_incremental", cells_to_highlight = colnames(so)[which(so$annotation_level2 %in% c("CD4_cl1","CD4_cl2","CD4_cl3", "CD4_cl4", "CD4_cl5", "CD4_cl7", "CD4_cl12"))], highlight_column_name = "annotation_level2", pixels = c(512, 512), mycols = mypal_level2, title = "", highlight_size = 0.1, highlight_alpha = 0.1, print_plot1 = T, print_plot2 = T, labelclusters = T)
p

p = MyDimPlotHighlight(seurat_object = so, umap_to_plot = "mde_incremental", cells_to_highlight = colnames(so)[which(so$annotation_level2 %in% c("CD4_cl2"))], highlight_column_name = "annotation_level2", pixels = c(512, 512), mycols = mypal_level2, title = "", highlight_size = 0.1, highlight_alpha = 0.1, print_plot1 = T, print_plot2 = T, labelclusters = T)
p

p = MyDimPlotHighlight(seurat_object = so, umap_to_plot = "mde_incremental", cells_to_highlight = colnames(so)[which(so$annotation_level2 %in% c("CD4_cl5", "CD4_cl7"))], highlight_column_name = "annotation_level2", pixels = c(512, 512), mycols = mypal_level2, title = "", highlight_size = 0.1, highlight_alpha = 0.1, print_plot1 = T, print_plot2 = T, labelclusters = T)
p

p = MyDimPlotHighlight(seurat_object = so, umap_to_plot = "mde_incremental", cells_to_highlight = colnames(so)[which(so$annotation_level2 %in% c("CD4_cl9", "CD4_cl10", "CD4_cl13", "CD4_cl14"))], highlight_column_name = "annotation_level2", pixels = c(512, 512), mycols = mypal_level2, title = "", highlight_size = 0.1, highlight_alpha = 0.1, print_plot1 = T, print_plot2 = T, labelclusters = T)
p

p = MyDimPlotHighlight(seurat_object = so, umap_to_plot = "mde_incremental", cells_to_highlight = colnames(so)[which(so$annotation_level2 %in% c("CD4_cl9","CD4_cl21","CD4_cl10","CD4_cl13", "CD4_cl14", "CD4_cl15", "CD4_cl16", "CD4_cl17", "CD4_cl18", "CD4_cl22", "CD4_cl23"))], highlight_column_name = "annotation_level2", pixels = c(512, 512), mycols = mypal_level2, title = "", highlight_size = 0.1, highlight_alpha = 0.1, print_plot1 = T, print_plot2 = T, labelclusters = T)
p

p = MyDimPlotHighlight(seurat_object = so, umap_to_plot = "mde_incremental", cells_to_highlight = colnames(so)[which(so$annotation_level2 %in% c("CD4_cl15", "CD4_cl16", "CD4_cl17", "CD4_cl18"))], highlight_column_name = "annotation_level2", pixels = c(512, 512), mycols = mypal_level2, title = "", highlight_size = 0.1, highlight_alpha = 0.1, print_plot1 = T, print_plot2 = T, labelclusters = T)
p

p = MyDimPlotHighlight(seurat_object = so, umap_to_plot = "mde_incremental", cells_to_highlight = colnames(so)[which(so$annotation_level2 %in% c("CD4_cl21", "CD4_cl22", "CD4_cl23"))], highlight_column_name = "annotation_level2", pixels = c(512, 512), mycols = mypal_level2, title = "", highlight_size = 0.1, highlight_alpha = 0.1, print_plot1 = T, print_plot2 = T, labelclusters = T)
p


p = MyDimPlotHighlight(seurat_object = so, umap_to_plot = "mde_incremental", cells_to_highlight = colnames(so)[which(so$annotation_level2 %in% c("CD4_cl9","CD4_cl21"))], highlight_column_name = "annotation_level2", pixels = c(512, 512), mycols = mypal_level2, title = "", highlight_size = 0.1, highlight_alpha = 0.1, print_plot1 = T, print_plot2 = T, labelclusters = T)
p

p = MyDimPlotHighlight(seurat_object = so, umap_to_plot = "mde_incremental", cells_to_highlight = colnames(so)[which(so$annotation_level2 %in% c("CD4_cl10","CD4_cl13", "CD4_cl14", "CD4_cl15", "CD4_cl16", "CD4_cl17", "CD4_cl18", "CD4_cl22", "CD4_cl23"))], highlight_column_name = "annotation_level2", pixels = c(512, 512), mycols = mypal_level2, title = "", highlight_size = 0.1, highlight_alpha = 0.1, print_plot1 = T, print_plot2 = T, labelclusters = T)
p
dev.off()
```

**Feature plots**
```{r}
so = NormalizeData(so,assay = "RNA", normalization.method = "LogNormalize",verbose = T)

FeaturePlot(so, reduction = "mde_incremental", slot = "data", features = "Tox", order = T, raster = T, raster.dpi = c(1024,1024))
FeaturePlot(so, reduction = "mde_incremental", slot = "data", features = "Gata3", order = T, raster = T, raster.dpi = c(1024,1024))
FeaturePlot(so, reduction = "mde_incremental", slot = "data", features = "Tbx21", order = T, raster = T, raster.dpi = c(1024,1024))
FeaturePlot(so, reduction = "mde_incremental", slot = "data", features = "Rorc", order = T, raster = T, raster.dpi = c(1024,1024))
FeaturePlot(so, reduction = "mde_incremental", slot = "data", features = "Tox2", order = T, raster = T, raster.dpi = c(1024,1024))
```

**Downsample seurat object**
```{r}
sample_cells = so@meta.data %>%
  group_by(!!sym("IGTHT")) %>%
  sample_n(size = min(500, n()), replace = FALSE) %>% 
    pull(IGT_cellID)

so = so[,sample_cells]
saveRDS(so, "/Users/david/Desktop/Treg_App/igt1_96_CD4_20250212_downsampled.Rds")
```

**Figure Tip/BAS**

2a
```{r}
pdf(sprintf("%s/MDE_level2.pdf", out_dir), 5,5, useDingbats = F)
DimPlot(so, reduction = "mde_incremental", group.by = "annotation_level2", cols = mypal_level2, raster = T, raster.dpi = c(1024,1024), pt.size = 1, alpha = 1, label = T, label.size = 2, label.box = T) + NoGrid()  + NoLegend()
p = DimPlot(so, reduction = "mde_incremental", group.by = "annotation_level2", cols = mypal_level2, raster = T, raster.dpi = c(1024,1024), pt.size = 1, alpha = 1) + xlim(c(-2.5,4)) + ylim(c(-2.5,2.5))
p
p + theme_void() + NoLegend() + NoGrid()
dev.off()

MyFeatureScatter(so,feature1 = "CD62L", feature2 = "CD44")

```


**T-RBI**

Background CD4 immgenT
```{r}
immgent_colors = readRDS('/Users/david/Library/CloudStorage/GoogleDrive-david.puyraimond@gmail.com/My Drive/ImmgenT/analysis/immgent_integration_git/immgent_color_palettes/20250709/immgent_color_palettes_list_20250709.Rds')
mypal_level2 = immgent_colors$CD4

so_orig = readRDS("igt1_96_withtotalvi20250710_clean_ADTonly.Rds")
so_orig[["mde_incremental_level2"]] = so_orig[["mde_incremental"]] 
so_orig[["mde_incremental_allT"]] = so_orig[["mde2_totalvi_20241006"]] 

so = so_orig[,so_orig$annotation_level1 == "CD4"]
so@meta.data %>% count(annotation_level2)

# sub = so@meta.data %>% group_by(annotation_level2) %>% slice_sample(n = 10000) %>% ungroup() %>% pull(cellID)
# so = so[,sub]
so$sample = "immgent"

# DimPlot(so, reduction = "mde_incremental_level2", group.by = "annotation_level2", cols = mypal_level2)
# MyDimPlotHighlight(so, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "annotation_level2", mycols = mypal_level2, print_plot1 = T, print_plot2 = T,labelclusters = T)

bg <- AddEmbeddingRasterBackground("DP", interpolate = F, color = "grey", alpha = 0.5)
```

Background CD4 immgenT
```{r}
immgent_colors = readRDS('/Users/david/Library/CloudStorage/GoogleDrive-david.puyraimond@gmail.com/My Drive/ImmgenT/analysis/immgent_integration_git/immgent_color_palettes/20250709/immgent_color_palettes_list_20250709.Rds')
mypal_level2 = immgent_colors$CD4

bg <- AddEmbeddingRasterBackground("DP", interpolate = F, color = "grey", alpha = 0.5)
```


https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE281547 - Kevin/Jenkins - Listeria, spleen

```{r}
query_dir = '/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/20250109_freeze/odhran/CD4/GSE281547_Jenkins_b6_aLm/Trial_Final/'
so_query = readRDS('/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/20250109_freeze/odhran/CD4/GSE281547_Jenkins_b6_aLm/Trial_Final/Final_Seurat_object.Rds') 

so_query$sample_name[so_query$sample_name == "Sample-2_D06"] = "Lm_D06"
so_query$sample_name[so_query$sample_name == "Sample-3_D21"] = "Lm_D21"
so_query$sample_name[so_query$sample_name == "Sample-1_D60"] = "Lm_D60"

so_query$cellID = colnames(so_query)
sub = so_query@meta.data %>% group_by(sample_name) %>% slice_sample(n = 4500) %>% ungroup() %>% pull(cellID)
so_query = so_query[,sub]

mypal_annotation = mypal
mypal_annotation = setNames(mypal_annotation, unique(so_query$annotations))

# umap_mat <- as.matrix( so_query@meta.data[, c("umaprna_1", "umaprna_2")])
# colnames(umap_mat) <- c("umappaper_1", "umappaper_2")
# rownames(umap_mat) <- rownames(so_query@meta.data)
# so_query[["umap_paper"]] <- CreateDimReducObject(
#   embeddings = umap_mat,
#   key = "UMAPPAPER_",
#   assay = DefaultAssay(so_query)
# )

head(so_query@meta.data)
# table(so_query$cluster)
# table(so_query$level2_plot)
# table(so_query$batch)
# table(so_query$tissue)
# table(so_query$annotation, so_query$tissue)

# table(so_query$batch_sample)
# table(so_query$level2_plot, so_query$cluster)
# table(so_query$level2_plot, so_query$level1_final)
# so_query$annotation = paste("cl", so_query$annotation, sep = "_")

so_query@meta.data %>% count(sample_name)
so_query@meta.data %>% count(level2_final)

# so_query$sample = as.integer(sub('.*-(\\d+)$', '\\1', colnames(so_query)))

sub = so_query@meta.data %>% filter(!(level2_final %in% c("not classified"))) %>% row.names()

DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "annotations", cols = mypal_annotation)

p1 = DimPlot(so_query[,sub], reduction = "mde_incremental_level2", group.by = "annotations", cols = mypal_annotation) 
p$layers <- c(list(immgenTCD4MDE()), p1$layers)
p

pdf(sprintf("%s/GSE281547_umap_dataset.pdf",query_dir),14, 7, useDingbats = F)
p1 = DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "annotations", cols = mypal_annotation)
p2 = DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "level2_final", cols = mypal_level2)
p1 + p2
DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "annotations", cols = mypal_annotation, split.by = "sample_name")
DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "level2_final", cols = mypal_level2, split.by = "sample_name")
# FeaturePlot(so_query[,sub], reduction = "mde_incremental_level2", features = "Pdcd1")
dev.off()

so2 = merge(so, so_query[,sub], merge.dr = T)

# mypal_annotation = setNames(c("#5D7AA5", "#BED4EB", "#E59339", "#F5CA8F", "#6FA649", "#B9E185", "#B29A39", "#EFD374", "#649B9F", "#A9CBC2", "#D2635A", "#F3B0B0"), c("cl_0", "cl_1", "cl_2", "cl_3", "cl_4", "cl_5", "cl_6", "cl_7", "cl_8", "cl_9", "cl_10", "cl_11"))
# mypal_annotation = mypal

pdf(sprintf("%s/GSE281547_RBI.pdf",query_dir),7, 7, useDingbats = F)
MyDimPlotHighlight(so2, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "level2_final", mycols = mypal_level2, print_plot1 = T, print_plot2 = T,labelclusters = F, xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = T, highlight_size = 1)
MyDimPlotHighlight(so2, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "annotations", mycols = mypal_annotation, print_plot1 = T, print_plot2 = T,labelclusters = F, xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = T, highlight_size = 1)

# MyDimPlotHighlight(so2, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "annotation", mycols = mypal_annotation, print_plot1 = T, print_plot2 = T,labelclusters = F)
for (s in unique(so_query$sample_name)) {
    print(s)
    tmp = so2[,so2$sample_name %in% c(s) | so2$sample %in% c("immgent")]
    print(MyDimPlotHighlight(tmp, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "level2_final", mycols = mypal_level2, print_plot1 = T, print_plot2 = T, labelclusters = T, xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = F, highlight_size = 0.5, title = sprintf("sample %s", s)))
    print(MyDimPlotHighlight(tmp, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "annotations", mycols = mypal_annotation, print_plot1 = T, print_plot2 = T,labelclusters = T, , xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = F, highlight_size = 0.5, title = sprintf("sample %s", s)))
    # print(MyDimPlotHighlight(so2[,so2$sample %in% c("immgent", s)], umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "annotation", mycols = mypal_annotation, print_plot1 = T, print_plot2 = T,labelclusters = T, title = sprintf("sample: %s", s)))
}

# p1 = DimPlot(so_query[,sub], reduction = "mde_incremental_level2", group.by = "annotation", cols = mypal_annotation)
# p2 = DimPlot(so_query[,sub], reduction = "mde_incremental_level2", group.by = "level2_plot", cols = mypal_level2)
# p1 + p2
dev.off()


##DotPlot

library(dplyr)
library(ggplot2)
library(scales)

df_dot <- so_query@meta.data %>%
  count(annotations, level2_final, name = "n") %>%
  group_by(annotations) %>%
  mutate(
    prop = n / sum(n),
    logn = log10(n + 1)
  ) %>%
  ungroup() %>%
  mutate(
    # keep the order as they appear in the data
    annotations  = factor(annotations,  levels = unique(annotations)),
    level2_final = factor(level2_final, levels = unique(level2_final))
  )
df_dot$level2_final = factor(df_dot$level2_final, levels = c(levels(so$annotation_level2), "not classified"))

p = ggplot(df_dot, aes(x = level2_final, y = annotations)) +
  geom_point(aes(size = prop, color = n)) +
  scale_size_continuous(labels = percent_format(accuracy = 1)) +
  scale_color_viridis_c(option = "magma", name = "log10(n+1)", direction =  -1) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 60, hjust = 1),
    panel.grid.major = element_line(color = "grey90")
  ) +
  labs(
    x = "level2_final",
    y = "annotations",
    size = "% within annotation"
  )
pdf(sprintf("%s/DotPlot_PaperAnnot_immgenTAnnot.pdf",query_dir),7, 5, useDingbats = F)
p
dev.off()


```

https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE182320 - Andreatta/Carmona - LCMV arm and cl13, spleen

```{r}
query_dir = '/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/20250109_freeze/odhran/CD4/GSE182320_IAb_Tet_WT_Zbtb7bKO_LCMV/Trial_Final/'
so_query = readRDS('/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/20250109_freeze/odhran/CD4/GSE182320_IAb_Tet_WT_Zbtb7bKO_LCMV/Trial_Final/Final_Seurat_object.Rds') 

so_query@meta.data %>% colnames()
so_query@meta.data %>% count(Sample, sample_name)

so_query$sample_name = so_query$Sample
so_query$annotations = so_query$functional.cluster
so_query@meta.data %>% count(annotations)
so_query@meta.data %>% count(level2_final)

so_query$sample_type = NA
so_query$sample_type[grepl(x = so_query$sample_name, pattern = "Arm_d7")] = "Arm_d7"
so_query$sample_type[grepl(x = so_query$sample_name, pattern = "Arm_d21")] = "Arm_d21"
so_query$sample_type[grepl(x = so_query$sample_name, pattern = "Arm_d60")] = "Arm_d60"
so_query$sample_type[grepl(x = so_query$sample_name, pattern = "Cl13_d7")] = "Cl13_d7"
so_query$sample_type[grepl(x = so_query$sample_name, pattern = "Cl13_d21")] = "Cl13_d21"
so_query@meta.data %>% count(sample_name, sample_type)
so_query@meta.data %>% count(sample_type)

so_query$cellID = colnames(so_query)
sub = so_query@meta.data %>% group_by(sample_type) %>% slice_sample(n = 3913) %>% ungroup() %>% pull(cellID)
so_query = so_query[,sub]

mypal_annotation = mypal
mypal_annotation = setNames(mypal_annotation, unique(so_query$annotations))

# umap_mat <- as.matrix( so_query@meta.data[, c("umaprna_1", "umaprna_2")])
# colnames(umap_mat) <- c("umappaper_1", "umappaper_2")
# rownames(umap_mat) <- rownames(so_query@meta.data)
# so_query[["umap_paper"]] <- CreateDimReducObject(
#   embeddings = umap_mat,
#   key = "UMAPPAPER_",
#   assay = DefaultAssay(so_query)
# )

head(so_query@meta.data)
# table(so_query$cluster)
# table(so_query$level2_plot)
# table(so_query$batch)
# table(so_query$tissue)
# table(so_query$annotation, so_query$tissue)

# table(so_query$batch_sample)
# table(so_query$level2_plot, so_query$cluster)
# table(so_query$level2_plot, so_query$level1_final)
# so_query$annotation = paste("cl", so_query$annotation, sep = "_")

so_query@meta.data %>% count(sample_name)
so_query@meta.data %>% count(level2_final)

# so_query$sample = as.integer(sub('.*-(\\d+)$', '\\1', colnames(so_query)))

sub = so_query@meta.data %>% filter(!(level2_final %in% c("not classified"))) %>% row.names()

pdf(sprintf("%s/GSE182320_umap_dataset.pdf",query_dir),14, 7, useDingbats = F)
p1 = DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "annotations", cols = mypal_annotation)
p2 = DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "level2_final", cols = mypal_level2)
p3 = DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "sample_name", cols = mypal)
p1 + p2 + p3
DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "annotations", cols = mypal_annotation, split.by = "sample_type")
DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "level2_final", cols = mypal_level2, split.by = "sample_type")
# FeaturePlot(so_query[,sub], reduction = "mde_incremental_level2", features = "Pdcd1")
dev.off()

so2 = merge(so, so_query[,sub], merge.dr = T)

# mypal_annotation = setNames(c("#5D7AA5", "#BED4EB", "#E59339", "#F5CA8F", "#6FA649", "#B9E185", "#B29A39", "#EFD374", "#649B9F", "#A9CBC2", "#D2635A", "#F3B0B0"), c("cl_0", "cl_1", "cl_2", "cl_3", "cl_4", "cl_5", "cl_6", "cl_7", "cl_8", "cl_9", "cl_10", "cl_11"))
# mypal_annotation = mypal

pdf(sprintf("%s/GSE182320_RBI.pdf",query_dir),7, 7, useDingbats = F)
MyDimPlotHighlight(so2, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "level2_final", mycols = mypal_level2, print_plot1 = T, print_plot2 = T,labelclusters = F, xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = T, highlight_size = 1)
MyDimPlotHighlight(so2, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "annotations", mycols = mypal_annotation, print_plot1 = T, print_plot2 = T,labelclusters = F, xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = T, highlight_size = 1)
MyDimPlotHighlight(so2, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "sample_type", mycols = mypal, print_plot1 = T, print_plot2 = T,labelclusters = F, xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = T, highlight_size = 1)


# MyDimPlotHighlight(so2, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "annotation", mycols = mypal_annotation, print_plot1 = T, print_plot2 = T,labelclusters = F)
for (s in unique(so_query$sample_type)) {
    print(s)
    tmp = so2[,so2$sample_type %in% c(s) | so2$sample %in% c("immgent")]
    print(MyDimPlotHighlight(tmp, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "level2_final", mycols = mypal_level2, print_plot1 = T, print_plot2 = T, labelclusters = T, xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = F, highlight_size = 0.5, title = sprintf("sample %s", s)))
    print(MyDimPlotHighlight(tmp, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "annotations", mycols = mypal_annotation, print_plot1 = T, print_plot2 = T,labelclusters = T, , xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = F, highlight_size = 0.5, title = sprintf("sample %s", s)))
    # print(MyDimPlotHighlight(so2[,so2$sample %in% c("immgent", s)], umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "annotation", mycols = mypal_annotation, print_plot1 = T, print_plot2 = T,labelclusters = T, title = sprintf("sample: %s", s)))
}

# p1 = DimPlot(so_query[,sub], reduction = "mde_incremental_level2", group.by = "annotation", cols = mypal_annotation)
# p2 = DimPlot(so_query[,sub], reduction = "mde_incremental_level2", group.by = "level2_plot", cols = mypal_level2)
# p1 + p2
dev.off()

##DotPlot

library(dplyr)
library(ggplot2)
library(scales)

df_dot <- so_query@meta.data %>%
  count(annotations, level2_final, name = "n") %>%
  group_by(annotations) %>%
  mutate(
    prop = n / sum(n),
    logn = log10(n + 1)
  ) %>%
  ungroup() %>%
  mutate(
    # keep the order as they appear in the data
    annotations  = factor(annotations,  levels = unique(annotations)),
    level2_final = factor(level2_final, levels = unique(level2_final))
  )
df_dot$level2_final = factor(df_dot$level2_final, levels = c(levels(so$annotation_level2), "not classified"))

p = ggplot(df_dot, aes(x = level2_final, y = annotations)) +
  geom_point(aes(size = prop, color = n)) +
  scale_size_continuous(labels = percent_format(accuracy = 1)) +
  scale_color_viridis_c(option = "magma", name = "log10(n+1)", direction =  -1) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 60, hjust = 1),
    panel.grid.major = element_line(color = "grey90")
  ) +
  labs(
    x = "level2_final",
    y = "annotations",
    size = "% within annotation"
  )
pdf(sprintf("%s/DotPlot_PaperAnnot_immgenTAnnot.pdf",query_dir),7, 5, useDingbats = F)
p
dev.off()

```

https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE121002 - LCMV arm D7 and D30 (in WT and Zbtb7bKO), spleen - Ciucci/Bosselut paper
No paper annotation

```{r}
query_dir = '/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/20250109_freeze/odhran/CD4/GSE121002_IAb_Tet_binding_Listeria/Not_classified_Full_Trial/Not_classified_loop_fixed/'
so_query = readRDS('/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/20250109_freeze/odhran/CD4/GSE121002_IAb_Tet_binding_Listeria/Not_classified_Full_Trial/Not_classified_loop_fixed/Final_Seurat_object.Rds') 

so_query@meta.data %>% colnames()
head(so_query@meta.data)
so_query@meta.data %>% count(sample_name, level1_final)

so_query@meta.data %>% filter(sample_name %in% c("Batch1_GSM3423796_d7_GP66", "Batch1_GSM3423797_d7_WT", "Batch1_GSM3423799_d30_WT") & level1_final == "CD4") %>% count(sample_name)

so_query$sample_type = NA
so_query$sample_type[grepl(x = so_query$sample_name, pattern = "d7")] = "Arm_d7"
so_query$sample_type[grepl(x = so_query$sample_name, pattern = "d30")] = "Arm_d30"
so_query@meta.data %>% count(sample_name, sample_type)
so_query@meta.data %>% count(sample_type)

so_query$cellID = colnames(so_query)
sub = so_query@meta.data %>% filter(sample_name %in% c("Batch1_GSM3423796_d7_GP66", "Batch1_GSM3423797_d7_WT", "Batch1_GSM3423799_d30_WT") & level1_final == "CD4") %>% group_by(sample_type) %>% slice_sample(n = 471) %>% ungroup() %>% pull(cellID)
so_query = so_query[,sub]
so_query@meta.data %>% count(sample_name, sample_type)

# so_query$sample_name = so_query$Sample
# so_query$annotations = so_query$functional.cluster
# so_query@meta.data %>% count(annotations)
# so_query@meta.data %>% count(level2_final)

mypal_annotation = mypal
# mypal_annotation = setNames(mypal_annotation, unique(so_query$annotations))

# umap_mat <- as.matrix( so_query@meta.data[, c("umaprna_1", "umaprna_2")])
# colnames(umap_mat) <- c("umappaper_1", "umappaper_2")
# rownames(umap_mat) <- rownames(so_query@meta.data)
# so_query[["umap_paper"]] <- CreateDimReducObject(
#   embeddings = umap_mat,
#   key = "UMAPPAPER_",
#   assay = DefaultAssay(so_query)
# )

head(so_query@meta.data)
# table(so_query$cluster)
# table(so_query$level2_plot)
# table(so_query$batch)
# table(so_query$tissue)
# table(so_query$annotation, so_query$tissue)

# table(so_query$batch_sample)
# table(so_query$level2_plot, so_query$cluster)
# table(so_query$level2_plot, so_query$level1_final)
# so_query$annotation = paste("cl", so_query$annotation, sep = "_")

so_query@meta.data %>% count(sample_name)
so_query@meta.data %>% count(level2_final)

# so_query$sample = as.integer(sub('.*-(\\d+)$', '\\1', colnames(so_query)))

sub = so_query@meta.data %>% filter(!(level2_final %in% c("CD4"))) %>% row.names()

pdf(sprintf("%s/GSE121002_umap_dataset.pdf",query_dir),14, 7, useDingbats = F)
# p1 = DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "annotations", cols = mypal_annotation)
p2 = DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "level2_final", cols = mypal_level2)
p3 = DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "sample_name", cols = mypal)
# p1 + p2 + p3
p2 + p3
# DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "annotations", cols = mypal_annotation, split.by = "sample_type")
DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "level2_final", cols = mypal_level2, split.by = "sample_name")
# FeaturePlot(so_query[,sub], reduction = "mde_incremental_level2", features = "Pdcd1")
dev.off()

so2 = merge(so, so_query[,sub], merge.dr = T)

# mypal_annotation = setNames(c("#5D7AA5", "#BED4EB", "#E59339", "#F5CA8F", "#6FA649", "#B9E185", "#B29A39", "#EFD374", "#649B9F", "#A9CBC2", "#D2635A", "#F3B0B0"), c("cl_0", "cl_1", "cl_2", "cl_3", "cl_4", "cl_5", "cl_6", "cl_7", "cl_8", "cl_9", "cl_10", "cl_11"))
# mypal_annotation = mypal

pdf(sprintf("%s/GSE121002_RBI.pdf",query_dir),7, 7, useDingbats = F)
MyDimPlotHighlight(so2, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "level2_final", mycols = mypal_level2, print_plot1 = T, print_plot2 = T,labelclusters = F, xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = F, highlight_size = 1)
# MyDimPlotHighlight(so2, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "annotations", mycols = mypal_annotation, print_plot1 = T, print_plot2 = T,labelclusters = F, xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = T, highlight_size = 1)
MyDimPlotHighlight(so2, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "sample_type", mycols = mypal, print_plot1 = T, print_plot2 = T,labelclusters = F, xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = F, highlight_size = 1)


# MyDimPlotHighlight(so2, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "annotation", mycols = mypal_annotation, print_plot1 = T, print_plot2 = T,labelclusters = F)
for (s in unique(so_query$sample_type)) {
    print(s)
    tmp = so2[,so2$sample_type %in% c(s) | so2$sample %in% c("immgent")]
    print(MyDimPlotHighlight(tmp, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "level2_final", mycols = mypal_level2, print_plot1 = T, print_plot2 = T, labelclusters = T, xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = F, highlight_size = 0.5, title = sprintf("sample %s", s)))
    # print(MyDimPlotHighlight(tmp, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "annotations", mycols = mypal_annotation, print_plot1 = T, print_plot2 = T,labelclusters = T, , xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = F, highlight_size = 0.5, title = sprintf("sample %s", s)))
    # print(MyDimPlotHighlight(so2[,so2$sample %in% c("immgent", s)], umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "annotation", mycols = mypal_annotation, print_plot1 = T, print_plot2 = T,labelclusters = T, title = sprintf("sample: %s", s)))
}

# p1 = DimPlot(so_query[,sub], reduction = "mde_incremental_level2", group.by = "annotation", cols = mypal_annotation)
# p2 = DimPlot(so_query[,sub], reduction = "mde_incremental_level2", group.by = "level2_plot", cols = mypal_level2)
# p1 + p2
dev.off()



```

https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE158896 - LCMV arm D10 spleen, CD4+CD44+GP66+
No paper annotation

```{r}
query_dir = '/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/20250109_freeze/odhran/CD4/GSE158896_IAb_Tet_LCMV_d35/Trial_Final/'
so_query = readRDS('/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/20250109_freeze/odhran/CD4/GSE158896_IAb_Tet_LCMV_d35/Trial_Final/Final_Seurat_object.Rds') 

so_query@meta.data %>% colnames()
head(so_query@meta.data)
so_query@meta.data %>% count(sample_name, level1_final)

so_query$sample_name = so_query$sample
# so_query$annotations = so_query$functional.cluster
# so_query@meta.data %>% count(annotations)
# so_query@meta.data %>% count(level2_final)

mypal_annotation = mypal
# mypal_annotation = setNames(mypal_annotation, unique(so_query$annotations))head(so_query)

# umap_mat <- as.matrix( so_query@meta.data[, c("umaprna_1", "umaprna_2")])
# colnames(umap_mat) <- c("umappaper_1", "umappaper_2")
# rownames(umap_mat) <- rownames(so_query@meta.data)
# so_query[["umap_paper"]] <- CreateDimReducObject(
#   embeddings = umap_mat,
#   key = "UMAPPAPER_",
#   assay = DefaultAssay(so_query)
# )

head(so_query@meta.data)
# table(so_query$cluster)
# table(so_query$level2_plot)
# table(so_query$batch)
# table(so_query$tissue)
# table(so_query$annotation, so_query$tissue)

# table(so_query$batch_sample)
# table(so_query$level2_plot, so_query$cluster)
# table(so_query$level2_plot, so_query$level1_final)
# so_query$annotation = paste("cl", so_query$annotation, sep = "_")

so_query@meta.data %>% count(sample_name)
so_query@meta.data %>% count(level2_final)

# so_query$sample = as.integer(sub('.*-(\\d+)$', '\\1', colnames(so_query)))

sub = so_query@meta.data %>% filter(!(level2_final %in% c("not classified"))) %>% row.names()

pdf(sprintf("%s/GSE158896_umap_dataset.pdf",query_dir),14, 7, useDingbats = F)
# p1 = DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "annotations", cols = mypal_annotation)
p2 = DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "level2_final", cols = mypal_level2)
p3 = DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "sample_name", cols = mypal)
# p1 + p2 + p3
p2 + p3
# DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "annotations", cols = mypal_annotation, split.by = "sample_type")
DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "level2_final", cols = mypal_level2, split.by = "sample_name")
# FeaturePlot(so_query[,sub], reduction = "mde_incremental_level2", features = "Pdcd1")
dev.off()

so2 = merge(so, so_query[,sub], merge.dr = T)

# mypal_annotation = setNames(c("#5D7AA5", "#BED4EB", "#E59339", "#F5CA8F", "#6FA649", "#B9E185", "#B29A39", "#EFD374", "#649B9F", "#A9CBC2", "#D2635A", "#F3B0B0"), c("cl_0", "cl_1", "cl_2", "cl_3", "cl_4", "cl_5", "cl_6", "cl_7", "cl_8", "cl_9", "cl_10", "cl_11"))
# mypal_annotation = mypal

pdf(sprintf("%s/GSE121002_RBI.pdf",query_dir),7, 7, useDingbats = F)
MyDimPlotHighlight(so2, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "level2_final", mycols = mypal_level2, print_plot1 = T, print_plot2 = T,labelclusters = F, xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = T, highlight_size = 1)
# MyDimPlotHighlight(so2, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "annotations", mycols = mypal_annotation, print_plot1 = T, print_plot2 = T,labelclusters = F, xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = T, highlight_size = 1)
# MyDimPlotHighlight(so2, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "sample_type", mycols = mypal, print_plot1 = T, print_plot2 = T,labelclusters = F, xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = F, highlight_size = 1)


# MyDimPlotHighlight(so2, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "annotation", mycols = mypal_annotation, print_plot1 = T, print_plot2 = T,labelclusters = F)
# for (s in unique(so_query$sample_type)) {
#     print(s)
#     tmp = so2[,so2$sample_type %in% c(s) | so2$sample %in% c("immgent")]
#     print(MyDimPlotHighlight(tmp, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "level2_final", mycols = mypal_level2, print_plot1 = T, print_plot2 = T, labelclusters = T, xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = F, highlight_size = 0.5, title = sprintf("sample %s", s)))
    # print(MyDimPlotHighlight(tmp, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "annotations", mycols = mypal_annotation, print_plot1 = T, print_plot2 = T,labelclusters = T, , xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = F, highlight_size = 0.5, title = sprintf("sample %s", s)))
    # print(MyDimPlotHighlight(so2[,so2$sample %in% c("immgent", s)], umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "annotation", mycols = mypal_annotation, print_plot1 = T, print_plot2 = T,labelclusters = T, title = sprintf("sample: %s", s)))
# }

# p1 = DimPlot(so_query[,sub], reduction = "mde_incremental_level2", group.by = "annotation", cols = mypal_annotation)
# p2 = DimPlot(so_query[,sub], reduction = "mde_incremental_level2", group.by = "level2_plot", cols = mypal_level2)
# p1 + p2
dev.off()



```

https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE198466 - lung CD4+ tet+ against Blastomyces
No paper annotation
Mice were inoculated either intranasally (IN) or subcutaneously (SC) with Blastomyces dermatitidis endoglucanase-2 (Bl-Eng2) six weeks prior to pathogen challenge. Tetramer-positive CD4+ T cells were FACS sorted and sequenced 3 days after the pulmonary challenge with B. dermatitidis

```{r}
query_dir = '/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/20250109_freeze/odhran/CD4/GSE198466_Vaccinated/Trial_Final/'
so_query = readRDS('/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/20250109_freeze/odhran/CD4/GSE198466_Vaccinated/Trial_Final/Final_Seurat_object.Rds') 

so_query@meta.data %>% colnames()
head(so_query@meta.data)
so_query$sample_name = so_query$original
so_query@meta.data %>% count(sample_name)

# so_query$sample_type = NA
# so_query$sample_type[grepl(x = so_query$sample_name, pattern = "d7")] = "Arm_d7"
# so_query$sample_type[grepl(x = so_query$sample_name, pattern = "d30")] = "Arm_d30"
# so_query@meta.data %>% count(sample_name, sample_type)
# so_query@meta.data %>% count(sample_type)

so_query$cellID = colnames(so_query)
sub = so_query@meta.data %>% group_by(sample_name) %>% slice_sample(n = 10689) %>% ungroup() %>% pull(cellID)
so_query = so_query[,sub]
so_query@meta.data %>% count(sample_name, sample_type)

# so_query$sample_name = so_query$Sample
# so_query$annotations = so_query$functional.cluster
# so_query@meta.data %>% count(annotations)
# so_query@meta.data %>% count(level2_final)

mypal_annotation = mypal
# mypal_annotation = setNames(mypal_annotation, unique(so_query$annotations))

# umap_mat <- as.matrix( so_query@meta.data[, c("umaprna_1", "umaprna_2")])
# colnames(umap_mat) <- c("umappaper_1", "umappaper_2")
# rownames(umap_mat) <- rownames(so_query@meta.data)
# so_query[["umap_paper"]] <- CreateDimReducObject(
#   embeddings = umap_mat,
#   key = "UMAPPAPER_",
#   assay = DefaultAssay(so_query)
# )

head(so_query@meta.data)
# table(so_query$cluster)
# table(so_query$level2_plot)
# table(so_query$batch)
# table(so_query$tissue)
# table(so_query$annotation, so_query$tissue)

# table(so_query$batch_sample)
# table(so_query$level2_plot, so_query$cluster)
# table(so_query$level2_plot, so_query$level1_final)
# so_query$annotation = paste("cl", so_query$annotation, sep = "_")

so_query@meta.data %>% count(sample_name)
so_query@meta.data %>% count(level2_final)

# so_query$sample = as.integer(sub('.*-(\\d+)$', '\\1', colnames(so_query)))

sub = so_query@meta.data %>% filter(!(level2_final %in% c("not classified"))) %>% row.names()

pdf(sprintf("%s/umap_dataset.pdf",query_dir),14, 7, useDingbats = F)
# p1 = DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "annotations", cols = mypal_annotation)
p2 = DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "level2_final", cols = mypal_level2)
p3 = DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "sample_name", cols = mypal)
# p1 + p2 + p3
p2 + p3
# DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "annotations", cols = mypal_annotation, split.by = "sample_type")
DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "level2_final", cols = mypal_level2, split.by = "sample_name")
# FeaturePlot(so_query[,sub], reduction = "mde_incremental_level2", features = "Pdcd1")
dev.off()

so2 = merge(so, so_query[,sub], merge.dr = T)

# mypal_annotation = setNames(c("#5D7AA5", "#BED4EB", "#E59339", "#F5CA8F", "#6FA649", "#B9E185", "#B29A39", "#EFD374", "#649B9F", "#A9CBC2", "#D2635A", "#F3B0B0"), c("cl_0", "cl_1", "cl_2", "cl_3", "cl_4", "cl_5", "cl_6", "cl_7", "cl_8", "cl_9", "cl_10", "cl_11"))
# mypal_annotation = mypal

pdf(sprintf("%s/RBI.pdf",query_dir),7, 7, useDingbats = F)
MyDimPlotHighlight(so2, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "level2_final", mycols = mypal_level2, print_plot1 = T, print_plot2 = T,labelclusters = F, xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = F, highlight_size = 1)
# MyDimPlotHighlight(so2, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "annotations", mycols = mypal_annotation, print_plot1 = T, print_plot2 = T,labelclusters = F, xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = T, highlight_size = 1)
MyDimPlotHighlight(so2, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "sample_name", mycols = mypal, print_plot1 = T, print_plot2 = T,labelclusters = F, xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = F, highlight_size = 1)


# MyDimPlotHighlight(so2, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "annotation", mycols = mypal_annotation, print_plot1 = T, print_plot2 = T,labelclusters = F)
for (s in unique(so_query$sample_name)) {
    print(s)
    tmp = so2[,so2$sample_name %in% c(s) | so2$sample %in% c("immgent")]
    print(MyDimPlotHighlight(tmp, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "level2_final", mycols = mypal_level2, print_plot1 = T, print_plot2 = T, labelclusters = T, xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = F, highlight_size = 0.5, title = sprintf("sample %s", s)))
    # print(MyDimPlotHighlight(tmp, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "annotations", mycols = mypal_annotation, print_plot1 = T, print_plot2 = T,labelclusters = T, , xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = F, highlight_size = 0.5, title = sprintf("sample %s", s)))
    # print(MyDimPlotHighlight(so2[,so2$sample %in% c("immgent", s)], umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "annotation", mycols = mypal_annotation, print_plot1 = T, print_plot2 = T,labelclusters = T, title = sprintf("sample: %s", s)))
}

# p1 = DimPlot(so_query[,sub], reduction = "mde_incremental_level2", group.by = "annotation", cols = mypal_annotation)
# p2 = DimPlot(so_query[,sub], reduction = "mde_incremental_level2", group.by = "level2_plot", cols = mypal_level2)
# p1 + p2
dev.off()


```

https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE200635 - tet+ CD4+ from draining lymph node and tumor, after subcutaneous injection of MC38-OVA tumor cells
No paper annotation
GP66:I-Ab+ CD4+ from Draining lymph node and Tumor
we don't have the demultiplexed data..

```{r}
query_dir = '/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/20250109_freeze/odhran/CD4/GSE200635_IAb_Tet_LCMV/Trial_Final/'
so_query = readRDS('/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/20250109_freeze/odhran/CD4/GSE200635_IAb_Tet_LCMV/Trial_Final/Final_Seurat_object.Rds') 

so_query@meta.data %>% colnames()
head(so_query@meta.data)
so_query$sample_name = so_query$sample
so_query@meta.data %>% count(sample_name)

# so_query$sample_type = NA
# so_query$sample_type[grepl(x = so_query$sample_name, pattern = "d7")] = "Arm_d7"
# so_query$sample_type[grepl(x = so_query$sample_name, pattern = "d30")] = "Arm_d30"
# so_query@meta.data %>% count(sample_name, sample_type)
# so_query@meta.data %>% count(sample_type)

# so_query$cellID = colnames(so_query)
# sub = so_query@meta.data %>% group_by(sample_name) %>% slice_sample(n = 10689) %>% ungroup() %>% pull(cellID)
# so_query = so_query[,sub]
# so_query@meta.data %>% count(sample_name, sample_type)

# so_query$sample_name = so_query$Sample
# so_query$annotations = so_query$functional.cluster
# so_query@meta.data %>% count(annotations)
# so_query@meta.data %>% count(level2_final)

mypal_annotation = mypal
# mypal_annotation = setNames(mypal_annotation, unique(so_query$annotations))

# umap_mat <- as.matrix( so_query@meta.data[, c("umaprna_1", "umaprna_2")])
# colnames(umap_mat) <- c("umappaper_1", "umappaper_2")
# rownames(umap_mat) <- rownames(so_query@meta.data)
# so_query[["umap_paper"]] <- CreateDimReducObject(
#   embeddings = umap_mat,
#   key = "UMAPPAPER_",
#   assay = DefaultAssay(so_query)
# )

head(so_query@meta.data)
# table(so_query$cluster)
# table(so_query$level2_plot)
# table(so_query$batch)
# table(so_query$tissue)
# table(so_query$annotation, so_query$tissue)

# table(so_query$batch_sample)
# table(so_query$level2_plot, so_query$cluster)
# table(so_query$level2_plot, so_query$level1_final)
# so_query$annotation = paste("cl", so_query$annotation, sep = "_")

so_query@meta.data %>% count(sample_name)
so_query@meta.data %>% count(level2_final)

# so_query$sample = as.integer(sub('.*-(\\d+)$', '\\1', colnames(so_query)))

sub = so_query@meta.data %>% filter(!(level2_final %in% c("not classified"))) %>% row.names()

pdf(sprintf("%s/umap_dataset.pdf",query_dir),14, 7, useDingbats = F)
# p1 = DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "annotations", cols = mypal_annotation)
p2 = DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "level2_final", cols = mypal_level2)
p3 = DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "sample_name", cols = mypal)
# p1 + p2 + p3
p2 + p3
# DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "annotations", cols = mypal_annotation, split.by = "sample_type")
DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "level2_final", cols = mypal_level2, split.by = "sample_name")
# FeaturePlot(so_query[,sub], reduction = "mde_incremental_level2", features = "Pdcd1")
dev.off()

so2 = merge(so, so_query[,sub], merge.dr = T)

# mypal_annotation = setNames(c("#5D7AA5", "#BED4EB", "#E59339", "#F5CA8F", "#6FA649", "#B9E185", "#B29A39", "#EFD374", "#649B9F", "#A9CBC2", "#D2635A", "#F3B0B0"), c("cl_0", "cl_1", "cl_2", "cl_3", "cl_4", "cl_5", "cl_6", "cl_7", "cl_8", "cl_9", "cl_10", "cl_11"))
# mypal_annotation = mypal

pdf(sprintf("%s/RBI.pdf",query_dir),7, 7, useDingbats = F)
MyDimPlotHighlight(so2, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "level2_final", mycols = mypal_level2, print_plot1 = T, print_plot2 = T,labelclusters = F, xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = F, highlight_size = 1)
# MyDimPlotHighlight(so2, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "annotations", mycols = mypal_annotation, print_plot1 = T, print_plot2 = T,labelclusters = F, xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = T, highlight_size = 1)
MyDimPlotHighlight(so2, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "sample_name", mycols = mypal, print_plot1 = T, print_plot2 = T,labelclusters = F, xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = F, highlight_size = 1)


# MyDimPlotHighlight(so2, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "annotation", mycols = mypal_annotation, print_plot1 = T, print_plot2 = T,labelclusters = F)
# for (s in unique(so_query$sample_name)) {
#     print(s)
#     tmp = so2[,so2$sample_name %in% c(s) | so2$sample %in% c("immgent")]
#     print(MyDimPlotHighlight(tmp, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "level2_final", mycols = mypal_level2, print_plot1 = T, print_plot2 = T, labelclusters = T, xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = F, highlight_size = 0.5, title = sprintf("sample %s", s)))
#     # print(MyDimPlotHighlight(tmp, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "annotations", mycols = mypal_annotation, print_plot1 = T, print_plot2 = T,labelclusters = T, , xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = F, highlight_size = 0.5, title = sprintf("sample %s", s)))
#     # print(MyDimPlotHighlight(so2[,so2$sample %in% c("immgent", s)], umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "annotation", mycols = mypal_annotation, print_plot1 = T, print_plot2 = T,labelclusters = T, title = sprintf("sample: %s", s)))
# }

# p1 = DimPlot(so_query[,sub], reduction = "mde_incremental_level2", group.by = "annotation", cols = mypal_annotation)
# p2 = DimPlot(so_query[,sub], reduction = "mde_incremental_level2", group.by = "level2_plot", cols = mypal_level2)
# p1 + p2
dev.off()


```

https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE184259 - SMARTA CD4+ T cell subsets during LCMV infection D7 and D21, spleen, SI LP and IEL, WT and Bcl6 KO.
No paper annotation
```{r}
query_dir = '/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/20250109_freeze/odhran/CD4/GSE184259_LCMV_CD4_WT_Bcl6KO_D7_D21/Trial_Final/'
so_query = readRDS('/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/20250109_freeze/odhran/CD4/GSE184259_LCMV_CD4_WT_Bcl6KO_D7_D21/Trial_Final/Final_Seurat_object.Rds') 

so_query$cellID = colnames(so_query)
so_query@meta.data %>% count(sample_name, IGTHT)

##Update metadata
so_query$sample_name[grepl(x = so_query$IGTHT, pattern = "Batch1_GSM5582632_D21_SPL_1")] = "Arm_d21_SLO_GSE184259_1"
so_query$sample_name[grepl(x = so_query$IGTHT, pattern = "Batch1_GSM5582633_D21_LPL_1")] = "Arm_d21_SILP_GSE184259_1"
so_query$sample_name[grepl(x = so_query$IGTHT, pattern = "Batch1_GSM5582634_D21_SPL_2")] = "Arm_d21_SLO_GSE184259_2"
so_query$sample_name[grepl(x = so_query$IGTHT, pattern = "Batch1_GSM5582635_D21_LPL_2")] = "Arm_d21_SILP_GSE184259_2"
so_query$sample_name[grepl(x = so_query$IGTHT, pattern = "Batch1_GSM5582636_D7_SPL_1")] = "Arm_d7_SLO_GSE184259_1"
so_query$sample_name[grepl(x = so_query$IGTHT, pattern = "Batch1_GSM5582637_D7_IEL_1")] = "Arm_d7_SIIEL_GSE184259_1"
so_query$sample_name[grepl(x = so_query$IGTHT, pattern = "Batch1_GSM5582638_D7_LPL_1")] = "Arm_d7_SILP_GSE184259_1"
so_query$sample_name[grepl(x = so_query$IGTHT, pattern = "Batch1_GSM5582639_D7_SPL_2")] = "Arm_d7_SLO_GSE184259_2"
so_query$sample_name[grepl(x = so_query$IGTHT, pattern = "Batch1_GSM5582640_D7_IEL_2")] = "Arm_d7_SIIEL_GSE184259_2"
so_query$sample_name[grepl(x = so_query$IGTHT, pattern = "Batch1_GSM5582641_D7_LPL_2")] = "Arm_d7_SILP_GSE184259_2"
so_query$sample_name[grepl(x = so_query$IGTHT, pattern = "Batch1_GSM5582642_D7_Bcl6KO_SPL")] = "Arm_d7_SLO_Bcl6KO_GSE184259"
so_query$sample_name[grepl(x = so_query$IGTHT, pattern = "Batch1_GSM5582643_D7_Bcl6KO_IEL")] = "Arm_d7_SIIEL_Bcl6KO_GSE184259"
so_query$sample_name[grepl(x = so_query$IGTHT, pattern = "Batch1_GSM5582644_D7_Bcl6KO_LPL")] = "Arm_d7_SILP_Bcl6KO_GSE184259"

so_query$sample_type = NA
so_query$sample_type[grepl(x = so_query$sample_name, pattern = "Arm_d21_SLO_GSE184259")] = "Arm_d21_SLO"
so_query$sample_type[grepl(x = so_query$sample_name, pattern = "Arm_d21_SILP_GSE184259")] = "Arm_d21_SILP"
so_query$sample_type[grepl(x = so_query$sample_name, pattern = "Arm_d7_SLO_GSE184259")] = "Arm_d7_SLO"
so_query$sample_type[grepl(x = so_query$sample_name, pattern = "Arm_d7_SILP_GSE184259")] = "Arm_d7_SILP"
so_query$sample_type[grepl(x = so_query$sample_name, pattern = "Arm_d7_SIIEL_GSE184259")] = "Arm_d7_SIIEL"
so_query$sample_type[grepl(x = so_query$sample_name, pattern = "Arm_d7_SLO_Bcl6KO_GSE184259")] = "Arm_d7_SLO_Bcl6KO"
so_query$sample_type[grepl(x = so_query$sample_name, pattern = "Arm_d7_SILP_Bcl6KO_GSE184259")] = "Arm_d7_SILP_Bcl6KO"
so_query$sample_type[grepl(x = so_query$sample_name, pattern = "Arm_d7_SIIEL_Bcl6KO_GSE184259")] = "Arm_d7_SIIEL_Bcl6KO"

so_query$annotations = NA

so_query$organ_simplified = NA
so_query$organ_simplified[grepl(x = so_query$sample_name, pattern = "SLO")] = "SLO"
so_query$organ_simplified[grepl(x = so_query$sample_name, pattern = "SILP")] = "SILP"
so_query$organ_simplified[grepl(x = so_query$sample_name, pattern = "SIIEL")] = "SIIEL"

so_query$target_cells_simplified = "CD4pSmarta"
so_query$Ag_spe = "Smarta"
so_query$infection_agent = "LCMV Armstrong"
# so_query$infection_agent[grepl(x = so_query$sample_name, pattern = "Arm")] = "LCMV Armstrong"
so_query$infection_phase = NA
so_query$infection_phase[grepl(x = so_query$sample_name, pattern = "d7")] = "acute"
so_query$infection_phase[grepl(x = so_query$sample_name, pattern = "d21")] = "memory"

so_query@meta.data %>% count(IGTHT, sample_name)

so_query@meta.data %>% count(level2_final)

##Save a clean version
sub = so_query@meta.data %>% filter(!(level2_final %in% c("not classified"))) %>% row.names()
so_query = so_query[,sub]
# saveRDS(so_query, sprintf("%s/Final_Seurat_object_clean_david.Rds", query_dir))

#####Plots

# sub = so_query@meta.data %>% group_by(sample_name) %>% slice_sample(n = 307) %>% ungroup() %>% pull(cellID)
# so_query = so_query[,sub]
# so_query@meta.data %>% count(sample_name)

sub = so_query@meta.data %>% filter(!(level2_final %in% c("not classified"))) %>% row.names() 
so_query = so_query[,sub]

# DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "sample_name", cols = mypal)
# p1 = DimPlot(so_query[,sub], reduction = "mde_incremental_level2", group.by = "annotations", cols = mypal_annotation) 
# p$layers <- c(list(immgenTCD4MDE()), p1$layers)
# p

pdf(sprintf("%s/umap_dataset.pdf",query_dir),14, 10, useDingbats = F)
p1 = DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "annotations", cols = mypal_annotation)
p2 = DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "level2_final", cols = mypal_level2)
p1 + p2
DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "annotations", cols = mypal_annotation, split.by = "sample_name", ncol = 4)
DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "level2_final", cols = mypal_level2, split.by = "sample_name", ncol = 4)
# FeaturePlot(so_query[,sub], reduction = "mde_incremental_level2", features = "Pdcd1")
dev.off()


# mypal_annotation = setNames(c("#5D7AA5", "#BED4EB", "#E59339", "#F5CA8F", "#6FA649", "#B9E185", "#B29A39", "#EFD374", "#649B9F", "#A9CBC2", "#D2635A", "#F3B0B0"), c("cl_0", "cl_1", "cl_2", "cl_3", "cl_4", "cl_5", "cl_6", "cl_7", "cl_8", "cl_9", "cl_10", "cl_11"))
# mypal_annotation = mypal

pdf(sprintf("%s/RBI.pdf",query_dir),7, 7, useDingbats = F)
# MyDimPlotHighlight(so2, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "level2_final", mycols = mypal_level2, print_plot1 = T, print_plot2 = T,labelclusters = F, xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = T, highlight_size = 1)
# MyDimPlotHighlight(so2, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "annotations", mycols = mypal_annotation, print_plot1 = T, print_plot2 = T,labelclusters = F, xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = T, highlight_size = 1)

p1 = DimPlot(so_query[,sub], reduction = "mde_incremental_level2", group.by = "level2_final", cols = mypal_level2)
bg = AddEmbeddingRasterBackground("CD4")
p1$layers = c(list(bg), p1$layers)
p1 + xlim(c(-2.5,4)) + ylim(c(-2.5,2.5)) + theme(legend.position = "none") + labs(title = "level2_final")

p2 = DimPlot(so_query[,sub], reduction = "mde_incremental_level2", group.by = "annotations", cols = mypal)
bg = AddEmbeddingRasterBackground("CD4")
p2$layers = c(list(bg), p2$layers)
p2 + xlim(c(-2.5,4)) + ylim(c(-2.5,2.5)) + theme(legend.position = "none") + labs(title = "annotations")

# MyDimPlotHighlight(so2, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "annotation", mycols = mypal_annotation, print_plot1 = T, print_plot2 = T,labelclusters = F)
for (s in unique(so_query$sample_name)) {
    print(s)
    bg = AddEmbeddingRasterBackground("CD4")
    p1 = DimPlot(so_query[,so_query$sample_name %in% s], reduction = "mde_incremental_level2", group.by = "level2_final", cols = mypal_level2)
    p1$layers = c(list(bg), p1$layers)
    print(p1 + xlim(c(-2.5,4)) + ylim(c(-2.5,2.5)) + theme(legend.position = "none") + labs(title = sprintf("sample %s", s)))
    
    # p2 = DimPlot(so_query[,so_query$sample_name %in% s], reduction = "mde_incremental_level2", group.by = "annotations", cols = mypal_level2)
    # p2$layers = c(list(bg), p2$layers)
    # print(p2 + xlim(c(-2.5,4)) + ylim(c(-2.5,2.5)) + theme(legend.position = "none") + labs(title = "level2_final"))

}

for (s in unique(so_query$sample_type)) {
    print(s)
    bg = AddEmbeddingRasterBackground("CD4")
    p1 = DimPlot(so_query[,so_query$sample_type %in% s], reduction = "mde_incremental_level2", group.by = "level2_final", cols = mypal_level2)
    p1$layers = c(list(bg), p1$layers)
    print(p1 + xlim(c(-2.5,4)) + ylim(c(-2.5,2.5)) + theme(legend.position = "none") + labs(title = sprintf("sample %s", s)))
    
    # p2 = DimPlot(so_query[,so_query$sample_name %in% s], reduction = "mde_incremental_level2", group.by = "annotations", cols = mypal_level2)
    # p2$layers = c(list(bg), p2$layers)
    # print(p2 + xlim(c(-2.5,4)) + ylim(c(-2.5,2.5)) + theme(legend.position = "none") + labs(title = "level2_final"))

}

# p1 = DimPlot(so_query[,sub], reduction = "mde_incremental_level2", group.by = "annotation", cols = mypal_annotation)
# p2 = DimPlot(so_query[,sub], reduction = "mde_incremental_level2", group.by = "level2_plot", cols = mypal_level2)
# p1 + p2
dev.off()



```

https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE175742 - SMARTA CD4+ T cell subsets during LCMV infection D7 and D21 and D44, spleen

```{r}
query_dir = '/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/20250109_freeze/odhran/CD4/GSE175742_LCMV_CD4_D7_D21_D44/Trial_Final/'
so_query = readRDS('/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/20250109_freeze/odhran/CD4/GSE175742_LCMV_CD4_D7_D21_D44/Trial_Final/Final_Seurat_object.Rds') 

so_query$cellID = colnames(so_query)
so_query@meta.data %>% count(sample_name, IGTHT)

##Update metadata
so_query$sample_name[grepl(x = so_query$IGTHT, pattern = "Batch1_GSM5345277_D7_SPL_1")] = "Arm_d7_SLO_GSE175742_1"
so_query$sample_name[grepl(x = so_query$IGTHT, pattern = "Batch1_GSM5345278_D7_SPL_2")] = "Arm_d7_SLO_GSE175742_2"
so_query$sample_name[grepl(x = so_query$IGTHT, pattern = "Batch1_GSM5345279_D21_SPL_1")] = "Arm_d21_SLO_GSE175742_1"
so_query$sample_name[grepl(x = so_query$IGTHT, pattern = "Batch1_GSM5345280_D21_SPL_2")] = "Arm_d21_SLO_GSE175742_2"
so_query$sample_name[grepl(x = so_query$IGTHT, pattern = "Batch1_GSM5345281_D44_SPL_1")] = "Arm_d44_SLO_GSE175742_1"
so_query$sample_name[grepl(x = so_query$IGTHT, pattern = "Batch1_GSM5345282_D44_SPL_2")] = "Arm_d44_SLO_GSE175742_2"

so_query$sample_type = NA
so_query$sample_type[grepl(x = so_query$sample_name, pattern = "Arm_d7_SLO_GSE175742")] = "Arm_d7_SLO"
so_query$sample_type[grepl(x = so_query$sample_name, pattern = "Arm_d21_SLO_GSE175742")] = "Arm_d21_SLO"
so_query$sample_type[grepl(x = so_query$sample_name, pattern = "Arm_d44_SLO_GSE175742")] = "Arm_d44_SLO"

so_query$annotations = NA

so_query$organ_simplified = "SLO"
# so_query$organ_simplified[grepl(x = so_query$sample_name, pattern = "SLO")] = "SLO"

so_query$target_cells_simplified = "CD4pSmarta"
so_query$Ag_spe = "Smarta"
so_query$infection_agent = "LCMV Armstrong"
# so_query$infection_agent[grepl(x = so_query$sample_name, pattern = "Arm")] = "LCMV Armstrong"
so_query$infection_phase = NA
so_query$infection_phase[grepl(x = so_query$sample_name, pattern = "d7")] = "acute"
so_query$infection_phase[grepl(x = so_query$sample_name, pattern = "d21")] = "memory"
so_query$infection_phase[grepl(x = so_query$sample_name, pattern = "d44")] = "memory"

so_query@meta.data %>% count(sample_name, IGTHT)

so_query@meta.data %>% count(sample_name, sample_type, organ_simplified, target_cells_simplified, Ag_spe, infection_agent, infection_phase)

##Save a clean version
so_query@meta.data %>% count(level2_final)
sub = so_query@meta.data %>% filter(!(level2_final %in% c("not classified"))) %>% row.names()
so_query = so_query[,sub]
saveRDS(so_query, sprintf("%s/Final_Seurat_object_clean_david.Rds", query_dir))

#####Plots

# sub = so_query@meta.data %>% group_by(sample_name) %>% slice_sample(n = 307) %>% ungroup() %>% pull(cellID)
# so_query = so_query[,sub]
# so_query@meta.data %>% count(sample_name)

sub = so_query@meta.data %>% filter(!(level2_final %in% c("not classified"))) %>% row.names() 
so_query = so_query[,sub]

# DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "sample_name", cols = mypal)
# p1 = DimPlot(so_query[,sub], reduction = "mde_incremental_level2", group.by = "annotations", cols = mypal_annotation) 
# p$layers <- c(list(immgenTCD4MDE()), p1$layers)
# p

pdf(sprintf("%s/umap_dataset.pdf",query_dir),20, 10, useDingbats = F)
p1 = DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "sample_name", cols = mypal)
p2 = DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "level2_final", cols = mypal_level2)
p1 + p2
DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "annotations", cols = mypal_annotation, split.by = "sample_name", ncol = 4)
DimPlot(so_query[,sub], reduction = "umap_rna", group.by = "level2_final", cols = mypal_level2, split.by = "sample_name", ncol = 4)
# FeaturePlot(so_query[,sub], reduction = "mde_incremental_level2", features = "Pdcd1")
dev.off()


# mypal_annotation = setNames(c("#5D7AA5", "#BED4EB", "#E59339", "#F5CA8F", "#6FA649", "#B9E185", "#B29A39", "#EFD374", "#649B9F", "#A9CBC2", "#D2635A", "#F3B0B0"), c("cl_0", "cl_1", "cl_2", "cl_3", "cl_4", "cl_5", "cl_6", "cl_7", "cl_8", "cl_9", "cl_10", "cl_11"))
# mypal_annotation = mypal

pdf(sprintf("%s/RBI.pdf",query_dir),7, 7, useDingbats = F)
# MyDimPlotHighlight(so2, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "level2_final", mycols = mypal_level2, print_plot1 = T, print_plot2 = T,labelclusters = F, xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = T, highlight_size = 1)
# MyDimPlotHighlight(so2, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "annotations", mycols = mypal_annotation, print_plot1 = T, print_plot2 = T,labelclusters = F, xlim = c(-2.5,4), ylim = c(-2.5,2.5), highlight_raster = T, highlight_size = 1)

p1 = DimPlot(so_query[,sub], reduction = "mde_incremental_level2", group.by = "level2_final", cols = mypal_level2)
bg = AddEmbeddingRasterBackground("CD4")
p1$layers = c(list(bg), p1$layers)
p1 + xlim(c(-2.5,4)) + ylim(c(-2.5,2.5)) + theme(legend.position = "none") + labs(title = "level2_final")

p2 = DimPlot(so_query[,sub], reduction = "mde_incremental_level2", group.by = "annotations", cols = mypal)
bg = AddEmbeddingRasterBackground("CD4")
p2$layers = c(list(bg), p2$layers)
p2 + xlim(c(-2.5,4)) + ylim(c(-2.5,2.5)) + theme(legend.position = "none") + labs(title = "annotations")

# MyDimPlotHighlight(so2, umap_to_plot = "mde_incremental_level2", cells_to_highlight = sub, highlight_column_name =  "annotation", mycols = mypal_annotation, print_plot1 = T, print_plot2 = T,labelclusters = F)
for (s in unique(so_query$sample_name)) {
    print(s)
    bg = AddEmbeddingRasterBackground("CD4")
    p1 = DimPlot(so_query[,so_query$sample_name %in% s], reduction = "mde_incremental_level2", group.by = "level2_final", cols = mypal_level2)
    p1$layers = c(list(bg), p1$layers)
    print(p1 + xlim(c(-2.5,4)) + ylim(c(-2.5,2.5)) + theme(legend.position = "none") + labs(title = sprintf("sample %s", s)))
    
    # p2 = DimPlot(so_query[,so_query$sample_name %in% s], reduction = "mde_incremental_level2", group.by = "annotations", cols = mypal_level2)
    # p2$layers = c(list(bg), p2$layers)
    # print(p2 + xlim(c(-2.5,4)) + ylim(c(-2.5,2.5)) + theme(legend.position = "none") + labs(title = "level2_final"))

}

for (s in unique(so_query$sample_type)) {
    print(s)
    bg = AddEmbeddingRasterBackground("CD4")
    p1 = DimPlot(so_query[,so_query$sample_type %in% s], reduction = "mde_incremental_level2", group.by = "level2_final", cols = mypal_level2)
    p1$layers = c(list(bg), p1$layers)
    print(p1 + xlim(c(-2.5,4)) + ylim(c(-2.5,2.5)) + theme(legend.position = "none") + labs(title = sprintf("sample %s", s)))
    
    # p2 = DimPlot(so_query[,so_query$sample_name %in% s], reduction = "mde_incremental_level2", group.by = "annotations", cols = mypal_level2)
    # p2$layers = c(list(bg), p2$layers)
    # print(p2 + xlim(c(-2.5,4)) + ylim(c(-2.5,2.5)) + theme(legend.position = "none") + labs(title = "level2_final"))

}

# p1 = DimPlot(so_query[,sub], reduction = "mde_incremental_level2", group.by = "annotation", cols = mypal_annotation)
# p2 = DimPlot(so_query[,sub], reduction = "mde_incremental_level2", group.by = "level2_plot", cols = mypal_level2)
# p1 + p2
dev.off()



```



**Figure T-RBI**
see here for https://docs.google.com/spreadsheets/d/10icsN1kaUNOgMYi-oVTFx8AH9CI_52SlQb3p798MfQY/edit?usp=sharing

Background CD4 immgenT
```{r}
immgent_colors = readRDS('/Users/david/Library/CloudStorage/GoogleDrive-david.puyraimond@gmail.com/My Drive/ImmgenT/analysis/immgent_integration_git/immgent_color_palettes/20250709/immgent_color_palettes_list_20250709.Rds')
mypal_level2 = immgent_colors$CD4

bg <- AddEmbeddingRasterBackground("DP", interpolate = F, color = "grey", alpha = 0.5)
```

Merge seurat object
```{r}
queries = list()

so_query = readRDS('/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/20250109_freeze/odhran/CD4/GSE281547_Jenkins_b6_aLm/Trial_Final/Final_Seurat_object.Rds') 
so_query$cellID = colnames(so_query)
so_query@meta.data %>% count(sample_name, IGTHT)
so_query$sample_name[so_query$IGTHT == "Sample-2_D06"] = "Lm_D06_SLO_GSE281547"
so_query$sample_name[so_query$IGTHT == "Sample-3_D21"] = "Lm_D21_SLO_GSE281547"
so_query$sample_name[so_query$IGTHT == "Sample-1_D60"] = "Lm_D60_SLO_GSE281547"
so_query$sample_type = NA
so_query$sample_type[so_query$IGTHT == "Sample-2_D06"] = "Lm_D06_SLO"
so_query$sample_type[so_query$IGTHT == "Sample-3_D21"] = "Lm_D21_SLO"
so_query$sample_type[so_query$IGTHT == "Sample-1_D60"] = "Lm_D60_SLO"
so_query$annotations = so_query$annotations
so_query$organ_simplified = "SLO"
so_query$target_cells_simplified = "CD4p2WTetp"
so_query$Ag_spe = "2W:I-Ab"
so_query$infection_agent = NA
so_query$infection_agent[grepl(x = so_query$sample_name, pattern = "Lm")] = "Listeria monocytogenes"
so_query$infection_phase = NA
so_query$infection_phase[grepl(x = so_query$sample_name, pattern = "D06")] = "acute"
so_query$infection_phase[grepl(x = so_query$sample_name, pattern = "D21")] = "memory"
so_query$infection_phase[grepl(x = so_query$sample_name, pattern = "D60")] = "memory"
sub = so_query@meta.data %>% filter(!(level2_final %in% c("not classified"))) %>% row.names()
so_query = so_query[,sub]
so_query@meta.data %>% count(sample_name, sample_type, organ_simplified, target_cells_simplified, Ag_spe, infection_agent, infection_phase)
queries[["GSE281547_Osum_Lm"]] = so_query

so_query = readRDS('/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/20250109_freeze/odhran/CD4/GSE182320_IAb_Tet_WT_Zbtb7bKO_LCMV/Trial_Final/Final_Seurat_object.Rds')
so_query$cellID = colnames(so_query)
so_query$sample_name = sprintf("%s_%s_%s", so_query$Sample, "SLO", "GSE182320")
so_query@meta.data %>% count(sample_name, IGTHT)
so_query$sample_type = NA
so_query$sample_type[grepl(x = so_query$IGTHT, pattern = "Arm_d7")] = "Arm_d7_SLO"
so_query$sample_type[grepl(x = so_query$IGTHT, pattern = "Arm_d21")] = "Arm_d21_SLO"
so_query$sample_type[grepl(x = so_query$IGTHT, pattern = "Arm_d60")] = "Arm_d60_SLO"
so_query$sample_type[grepl(x = so_query$IGTHT, pattern = "Cl13_d7")] = "Cl13_d7_SLO"
so_query$sample_type[grepl(x = so_query$IGTHT, pattern = "Cl13_d21")] = "Cl13_d21_SLO"
so_query$annotations = so_query$functional.cluster
so_query$organ_simplified = "SLO"
so_query$target_cells_simplified = "CD4pGP66Tetp"
so_query$Ag_spe = "GP66:I-Ab"
so_query$infection_agent = NA
so_query$infection_agent[grepl(x = so_query$sample_name, pattern = "Arm")] = "LCMV Armstrong"
so_query$infection_agent[grepl(x = so_query$sample_name, pattern = "Cl13")] = "LCMV Clone 13"
so_query$infection_phase = NA
so_query$infection_phase[grepl(x = so_query$sample_name, pattern = "d7")] = "acute"
so_query$infection_phase[grepl(x = so_query$sample_name, pattern = "Arm_d21")] = "memory"
so_query$infection_phase[grepl(x = so_query$sample_name, pattern = "Cl13_d21")] = "chronic"
so_query$infection_phase[grepl(x = so_query$sample_name, pattern = "d60")] = "memory"
so_query@meta.data %>% count(sample_name, sample_type)
sub = so_query@meta.data %>% filter(!(level2_final %in% c("not classified"))) %>% row.names()
so_query = so_query[,sub]
so_query@meta.data %>% count(sample_name, sample_type, organ_simplified, target_cells_simplified, Ag_spe, Ag_spe)
queries[["GSE182320_Andreatta_LCMVArmCl13"]] = so_query

so_query = readRDS('/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/20250109_freeze/odhran/CD4/GSE121002_IAb_Tet_binding_Listeria/Not_classified_Full_Trial/Not_classified_loop_fixed/Final_Seurat_object.Rds') 
so_query$cellID = colnames(so_query)
so_query@meta.data %>% count(sample_name, IGTHT)
sub = so_query@meta.data %>% filter(sample_name %in% c("Batch1_GSM3423796_d7_GP66", "Batch1_GSM3423797_d7_WT", "Batch1_GSM3423799_d30_WT") & level1_final == "CD4") %>% pull(cellID)
so_query = so_query[,sub]
so_query$sample_name[grepl(x = so_query$sample_name, pattern = "d7")] = "Arm_d7_SLO_GSE121002"
so_query$sample_name[grepl(x = so_query$sample_name, pattern = "d30")] = "Arm_d30_SLO_GSE121002"
sub = so_query@meta.data %>% filter(!(level2_final %in% c("CD4"))) %>% row.names()
so_query$sample_type = NA
so_query$sample_type[grepl(x = so_query$sample_name, pattern = "d7")] = "Arm_d7_SLO"
so_query$sample_type[grepl(x = so_query$sample_name, pattern = "d30")] = "Arm_d30_SLO"
so_query$annotations = NA
so_query$organ_simplified = "SLO"
so_query$target_cells_simplified = "CD4pCD44pGP66Tetp"
so_query$Ag_spe = "GP66:I-Ab"
so_query$infection_agent = NA
so_query$infection_agent[grepl(x = so_query$sample_name, pattern = "Arm")] = "LCMV Armstrong"
so_query$infection_phase = NA
so_query$infection_phase[grepl(x = so_query$sample_name, pattern = "d7")] = "acute"
so_query$infection_phase[grepl(x = so_query$sample_name, pattern = "d30")] = "memory"
so_query = so_query[,sub]
so_query@meta.data %>% count(sample_name, sample_type, organ_simplified, target_cells_simplified, Ag_spe, Ag_spe)
queries[["GSE121002_Ciucci_LCMVArm"]] = so_query

so_query = readRDS('/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/20250109_freeze/odhran/CD4/GSE158896_IAb_Tet_LCMV_d35/Trial_Final/Final_Seurat_object.Rds') 
so_query$cellID = colnames(so_query)
so_query@meta.data %>% count(sample_name)
so_query$sample_name = so_query$sample
so_query$sample_name[so_query$IGTHT == "Batch1_GSM4815363_M1_GEX"] = "Arm_d10_SLO_GSE158896_1"
so_query$sample_name[so_query$IGTHT == "Batch1_GSM4815364_M2_GEX"] = "Arm_d10_SLO_GSE158896_2"
so_query$sample_name[so_query$IGTHT == "Batch1_GSM4815365_M3_GEX"] = "Arm_d10_SLO_GSE158896_3"
so_query$sample_name[so_query$IGTHT == "Batch1_GSM4815366_M4_GEX"] = "Arm_d10_SLO_GSE158896_4"
so_query$sample_name[so_query$IGTHT == "Batch1_GSM4815367_M5_GEX"] = "Arm_d10_SLO_GSE158896_5"
so_query$sample_type = "Arm_d10_SLO"
so_query$organ_simplified = "SLO"
so_query$target_cells_simplified = "CD4pCD44pGP66Tetp"
so_query$Ag_spe = "GP66:I-Ab"
so_query$infection_agent = NA
so_query$infection_agent[grepl(x = so_query$sample_name, pattern = "Arm")] = "LCMV Armstrong"
so_query$infection_phase = "acute"
sub = so_query@meta.data %>% filter(!(level2_final %in% c("not classified"))) %>% row.names()
so_query = so_query[,sub]
so_query@meta.data %>% count(sample_name, sample_type, organ_simplified, target_cells_simplified, Ag_spe, Ag_spe)
queries[["GSE158896_Khatun_LCMVArm"]] = so_query

so_query = readRDS('/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/20250109_freeze/odhran/CD4/GSE198466_Vaccinated/Trial_Final/Final_Seurat_object.Rds') 
so_query$cellID = colnames(so_query)
so_query@meta.data %>% count(sample_name)
so_query$sample_name[so_query$sample_name == "GSM5949203_IN_GSM5949203_IN"] = "Blasto_lung_secondaryD3_intranasalvaccine_GSE198466"
so_query$sample_name[so_query$sample_name == "GSM5949204_SC_GSM5949204_SC"] = "Blasto_lung_secondaryD3_subcutvaccine_GSE198466"
so_query$sample_type = NA
so_query$sample_type[so_query$sample_name == "Blasto_lung_secondaryD3_intranasalvaccine_GSE198466"] = "Blasto_lung_secondaryD3_intranasalvaccine"
so_query$sample_type[so_query$sample_name == "Blasto_lung_secondaryD3_subcutvaccine_GSE198466"] = "Blasto_lung_secondaryD3_subcutvaccine"
so_query$annotations = NA
so_query$organ_simplified = "lung"
so_query$target_cells_simplified = "CD4pEng2Tetp"
so_query$Ag_spe = "Bl-Eng2:classII"
so_query$infection_agent = NA
so_query$infection_agent[grepl(x = so_query$sample_name, pattern = "Blasto")] = "Blastomyces dermatitidis"
so_query$infection_phase = "recall"
sub = so_query@meta.data %>% filter(!(level2_final %in% c("not classified"))) %>% row.names()
so_query = so_query[,sub]
so_query@meta.data %>% count(sample_name, sample_type, organ_simplified, target_cells_simplified, Ag_spe)
queries[["GSE198466_Woodring_Blasto"]] = so_query

so_query = readRDS('/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/20250109_freeze/odhran/CD4/GSE200635_IAb_Tet_LCMV/Trial_Final/Final_Seurat_object.Rds') 
so_query$cellID = colnames(so_query)
so_query@meta.data %>% count(sample_name, IGTHT)
so_query$sample_name[so_query$IGTHT == "Batch1_GSM6040471"] = "MC38GP66_dLN_TIL_GSE200635"
so_query$sample_type = NA
so_query$sample_type[so_query$IGTHT == "Batch1_GSM6040471"] = "MC38GP66_dLN_TIL"
so_query$annotations = NA
so_query$organ_simplified = "dLN_Tumor"
so_query$target_cells_simplified = "CD4pGP66Tetp"
so_query$Ag_spe = "GP66:I-Ab"
so_query$infection_agent = NA
so_query$infection_phase = NA
sub = so_query@meta.data %>% filter(!(level2_final %in% c("not classified"))) %>% row.names()
so_query = so_query[,sub]
so_query@meta.data %>% count(sample_name, sample_type, organ_simplified, target_cells_simplified, Ag_spe, Ag_spe)
queries[["GSE200635_Andreatta_Tumor"]] = so_query

lapply(queries, function(x) x@meta.data %>% count(sample_name, sample_type, organ_simplified, target_cells_simplified, Ag_spe, infection_agent, infection_phase))

saveRDS(queries, file = "trbi_6studies_20260306.Rds")


```

Merge on Midway3 (rest of the code can be done on Laptop)
```{r}

#Merge and keep only overlapping genes
library(Seurat)
library(purrr)

so_list = readRDS("trbi_6studies_20260306.Rds")

common_features <- Reduce(
  intersect,
  lapply(so_list, rownames)
)
length(common_features)
head(common_features)

so_list_filt = list()
for (i in 1:length(so_list)) {
    print(i)
    so_list_filt[[i]] = subset(so_list[[i]], features = common_features)
    
    so_list_filt[[i]]$dataset = names(so_list)[i]
    
    message("Rename cells...")
    so_list_filt[[i]] = RenameCells(so_list_filt[[i]], new.names = paste0(names(so_list)[i], "_", colnames(so_list_filt[[i]])))
    
    
}

so_merged = merge(so_list_filt[[1]],so_list_filt[2:length(so_list_filt)],  collapse = T, merge.data = T, merge.dr = T)

so_merged[["mde_incremental"]] <- so_merged[["mde_incremental_level2"]]
so_merged[["mde_incremental_level2"]] <- NULL

so_merged[["mde2_totalvi_20241006"]] <- so_merged[["mde_incremental_allT"]]
so_merged[["mde_incremental_allT"]] <- NULL

newann = read.table('/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/20260206_freeze/annotation_table_20260201_conversion_forOdhran.csv', header = T, sep = ",")
so_merged$level2_final.v1 = so_merged$level2_final
so_merged$level2_final = newann$level2.new[match(so_merged$level2_final.v1, newann$level2)]
any(is.na(so_merged$level2_final))
so_merged@meta.data %>% count(level2_final, level2_final.v1)

saveRDS(so_merged, file = "trbi_6studies_20260306_merged.Rds")


## merge and Diet Seurat
so_list = readRDS("trbi_6studies_20260306.Rds")

so_list_slim = list()
for (i in 1:length(so_list)) {
    print(i)
    obj = so_list[[i]]
    dim_keep <- intersect(c("mde_incremental_allT", "mde_incremental_level2"), names(obj@reductions))
    features_keep = intersect(c("Cd4", "Cd8b1", "Cd8a", "Trdc", "Foxp3", "Zbtb16", "Trdc", 
                                "Ifng", "Il4", "Il13", "Il17a", "Il17f", "Il22", "Tbx21", "Rorc", "Gata3", "Tox", "Tox2", "Pdcd1", "Cxcr5", "Sell", "Cd44", "Ccr7", "Ly6c1", "Ly6c2"), rownames(obj[["RNA"]]@counts))
    
    message("DietSeurat...")
    obj_slim = DietSeurat(
        obj,
        assays = "RNA",
        layers = "counts",
        features = features_keep,
        dimreducs  = dim_keep,
        misc       = FALSE
    )
    
    obj_slim$dataset = names(so_list)[i]
    
    message("Rename cells...")
    obj_slim = RenameCells(obj_slim, new.names = paste0(names(so_list)[i], "_", colnames(obj_slim)))
    
    so_list_slim[[names(so_list)[i]]] = obj_slim
    
}

# Merge all slimmed objects
so_merged = merge(so_list_slim[[1]],so_list_slim[2:length(so_list_slim)],  collapse = T, merge.data = T, merge.dr = T)
# so_merged = Reduce(function(x, y) merge(x, y), so_list_slim)
so_merged

so_merged[["mde_incremental"]] <- so_merged[["mde_incremental_level2"]]
so_merged[["mde_incremental_level2"]] <- NULL

so_merged[["mde2_totalvi_20241006"]] <- so_merged[["mde_incremental_allT"]]
so_merged[["mde_incremental_allT"]] <- NULL

newann = read.table('/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/20260206_freeze/annotation_table_20260201_conversion_forOdhran.csv', header = T, sep = ",")
so_merged$level2_final.v1 = so_merged$level2_final
so_merged$level2_final = newann$level2.new[match(so_merged$level2_final.v1, newann$level2)]
any(is.na(so_merged$level2_final))
so_merged@meta.data %>% count(level2_final, level2_final.v1)


saveRDS(so_merged, file = "trbi_6studies_20260306_merged_diet.Rds")

```

Table with query datasets and cell counts: 
```{r}
so_merged_orig = readRDS("trbi_6studies_20260306_merged_diet.Rds")
so_merged_orig@meta.data %>% count(dataset,sample_name, name = "n_cells") %>% write.table(file = "trbi_6studies_20260306_table.txt", row.names = F, col.names = T, sep = "\t", quote = F)

so_merged = so_merged_orig[,!so_merged_orig$level1_final %in% c("nonT", "unclear")]

so_merged@meta.data %>% count(dataset, sample_name, organ_simplified, target_cells_simplified, Ag_spe, infection_phase) 

```

Convert to new annotation
```{r}
# newann = read.table('/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/20260206_freeze/annotation_table_20260201_conversion_forOdhran.csv', header = T, sep = ",")
# so_merged$level2_final.v1 = so_merged$level2_final
# so_merged$level2_final = newann$level2.new[match(so_merged$level2_final.v1, newann$level2)]
# any(is.na(so_merged$level2_final))

# so_merged@meta.data %>% count(sample_name)
# so_orig@meta.data %>% count(infection_agent)
# so_merged$infection_agent = NA
# so_merged$infection_agent[grepl(x = so_merged$sample_name, pattern = "Lm")] = "Listeria monocytogenes"
# so_merged$infection_agent[grepl(x = so_merged$sample_name, pattern = "Arm")] = "LCMV Armstrong"
# so_merged$infection_agent[grepl(x = so_merged$sample_name, pattern = "Cl13")] = "LCMV clone 13"
# so_merged$infection_agent[grepl(x = so_merged$sample_name, pattern = "Blasto")] = "Blastomyces dermatitidis"


```

Plot MDE
```{r}
pdf("trbi_6studies_mde.pdf", 14, 7, useDingbats = F)

so_merged@meta.data %>% count(sample_name, sample_type)
so_merged@meta.data %>% count(sample_type)
i = "Arm_d10"

pdf(sprintf("%s/TRBI_MDE sample_type.pdf",out_dir),7, 7, useDingbats = F)
for (i in unique(so_merged$sample_type)) {
    print(i)
    tmp = so_merged[,so_merged@meta.data %>% filter(sample_type == i) %>% rownames()]
    
    p = DimPlot(tmp, reduction = "mde_incremental", group.by = "level2_final", cols = mypal_level2) + ggtitle(i)
    bg = AddEmbeddingRasterBackground(lineage = "CD4")
    p$layers = c(list(bg), p$layers)
    p = p + ggplot2::coord_fixed() + xlim(c(-2.5,4)) + ylim(c(-2.5,2.5)) +theme_void() + NoLegend()
    print(p)
}
dev.off()
```

Plot % cluster by sample, grouped by infection_phase, one for each dataset
```{r}
library(dplyr)
library(ggplot2)
library(forcats)
library(scales)
library(stringr)

cluster_col <- "level2_final"   # change if needed

so_merged@meta.data %>% filter(!is.na(dataset), !is.na(sample_name), !is.na(infection_phase), infection_phase %in% c("acute", "memory")) %>% count(sample_name, infection_phase)
df_plot <- so_merged@meta.data %>%
  filter(!is.na(dataset), !is.na(sample_name), !is.na(infection_phase), infection_phase %in% c("acute", "memory")) %>%
  count(dataset, sample_name, infection_agent, infection_phase, .data[[cluster_col]], name = "n") %>%
  group_by(dataset, sample_name, infection_phase) %>%
  mutate(
    sample_total = sum(n),
    pct = 100 * n / sample_total
  ) %>%
  ungroup() %>%
  rename(cluster = all_of(cluster_col))

df_plot2 <- df_plot %>%
  filter(infection_phase %in% c("acute","memory")) %>%
  mutate(
    cluster_phase = paste(cluster, infection_phase)
  )
clusters <- levels(so$annotation_level2)[grepl("^CD4", levels(so$annotation_level2))]
x_levels <- as.vector(
  rbind(
    paste(clusters, "acute"),
    paste(clusters, "memory")
  )
)
df_plot2$cluster_phase <- factor(df_plot2$cluster_phase, levels = x_levels)

df_plot3 = df_plot2 %>% filter(!(cluster %in% c("CD4.wM", "CD4.wN", "CD4.wZ")), infection_agent %in% c("LCMV Armstrong"))
df_plot3$cluster = factor(df_plot3$cluster, levels = intersect(clusters, df_plot3$cluster))


p = ggplot(df_plot3, aes(cluster_phase, pct)) +
    # geom_point(
    #     aes(color = dataset),
    #     size = 1,
    #     alpha = 1
    # ) +
    geom_jitter(
        aes(color = dataset),
        width = 0.15,
        size = 1
    )+
    stat_summary(
        fun = mean,
        geom = "crossbar",
        width = 0.5,
        color = "black"
    )+
    facet_wrap(~ infection_phase, scales = "free_x") +
    scale_color_manual(values = mypal) +
    theme_bw() +
    theme(
        axis.text.x = element_text(angle = 60, hjust = 1),
        panel.grid.major.x = element_blank()
    ) +
    labs(
        x = "Cluster / infection phase",
        y = "% cells"
    )
p

# p2 = ggplot(df_plot3, aes(cluster_phase, pct)) +
#     # geom_point(
#     #     aes(color = infection_phase),
#     #     size = 1,
#     #     alpha = 1
#     # ) +
#     geom_jitter(
#         aes(color = infection_phase),
#         width = 0.15,
#         size = 1
#     )+
#     stat_summary(
#         fun = mean,
#         geom = "crossbar",
#         width = 0.5,
#         color = "black"
#     )+
#     facet_wrap(~ dataset, scales = "free_x") +
#     scale_color_manual(values = c("acute" = "blue", "memory" = "red")) +
#     theme_bw() +
#     theme(
#         axis.text.x = element_text(angle = 60, hjust = 1),
#         panel.grid.major.x = element_blank()
#     ) +
#     labs(
#         x = "Cluster / infection phase",
#         y = "% cells"
#     )
# p2

p2 = ggplot(df_plot3, aes(cluster, pct, color = infection_phase, shape = dataset)) +
    geom_jitter(
        position = position_jitterdodge(
            jitter.width = 0.1,
            dodge.width = 0.6
        ),
        size = 1
    ) +
    stat_summary(
        aes(group = infection_phase),
        fun = mean,
        geom = "crossbar",
        width = 0.5,
        position = position_dodge(width = 0.6),
        color = "black"
    ) +
    facet_wrap(~ dataset, scales = "free_x") +
    scale_color_manual(values = c("acute" = "blue", "memory" = "red")) +
    theme_bw() +
    theme(
        axis.text.x = element_text(angle = 60, hjust = 1),
        panel.grid.major.x = element_blank()
    ) +
    labs(
        x = "Cluster",
        y = "% cells"
    )
p2

p3 = ggplot(df_plot3, aes(cluster, pct, color = infection_phase, shape = dataset)) +
    geom_jitter(
        position = position_jitterdodge(
            jitter.width = 0.1,
            dodge.width = 0.6
        ),
        size = 2
    ) +
    stat_summary(
        aes(group = infection_phase),
        fun = mean,
        geom = "crossbar",
        width = 0.5,
        position = position_dodge(width = 0.6),
        color = "black"
    ) +
    scale_color_manual(values = c("acute" = "blue", "memory" = "red")) +
    theme_bw() +
    theme(
        axis.text.x = element_text(angle = 60, hjust = 1),
        panel.grid.major.x = element_blank()
    ) +
    labs(
        x = "Cluster",
        y = "% cells"
    )
p3

pdf(sprintf("%s/TRBI_memory_acute_ClusterProp_LCMV.pdf", out_dir), 10,5, useDingbats = F)
p + NoGrid()
p2 + NoGrid()
p3 + NoGrid()
dev.off()


##Listeria
df_plot3 = df_plot2 %>% filter(!(cluster %in% c("CD4.wM", "CD4.wN", "CD4.wZ")), infection_agent %in% c("Listeria monocytogenes"))

p3 = ggplot(df_plot3, aes(cluster, pct, color = infection_phase, shape = dataset)) +
    geom_jitter(
        position = position_jitterdodge(
            jitter.width = 0.1,
            dodge.width = 0.6
        ),
        size = 3
    ) +
    stat_summary(
        aes(group = infection_phase),
        fun = mean,
        geom = "crossbar",
        width = 0.5,
        position = position_dodge(width = 0.6),
        color = "black"
    ) +
    scale_color_manual(values = c("acute" = "blue", "memory" = "red")) +
    theme_bw() +
    theme(
        axis.text.x = element_text(angle = 60, hjust = 1),
        panel.grid.major.x = element_blank()
    ) +
    labs(
        x = "Cluster",
        y = "% cells"
    )
p3

pdf(sprintf("%s/TRBI_LCMVmemory_acute_ClusterProp_Listeria.pdf", out_dir), 7,5, useDingbats = F)
p3 + NoGrid()
dev.off()

```

Pseudobulk analysis
```{r}
so_merged = readRDS("trbi_6studies_20260306_merged.Rds")
head(so_merged@meta.data)

so_merged$pb_group <- paste(
    so_merged$sample_name,
    so_merged$level2_final,
    sep = "."
) %>% gsub("\\_", ".", .)


pb = AggregateExpression(
    object = so_merged,
    assay = "RNA",
    group.by = "pb_group",
    # features = tf,
    normalization.method = "LogNormalize",
    scale.factor = 10000,
    verbose = T,
    return.seurat = TRUE
)

so_merged@meta.data %>% colnames()
pb_meta <- so_merged@meta.data %>%
    dplyr::select(dataset, sample_name, sample_type, level2_final, organ_simplified, infection_agent, infection_phase, Ag_spe, pb_group) %>% group_by(pb_group) %>% mutate(ncells = n()) %>% ungroup() %>% distinct(pb_group, .keep_all = TRUE) %>% as.data.frame()

rownames(pb_meta) <- pb_meta$pb_group
pb_meta <- pb_meta[colnames(pb), ]

pb@meta.data = pb_meta

saveRDS(pb, file = "trbi_6studies_20260306_pseudobulk.Rds")

```

DGE limma - genes that are consistently different between memory and acute CD4 T cells in SLO, across all clusters, infections (LCMV arm and Listeria), and datasets.
```{r}
library(edgeR)

pb_orig = readRDS("trbi_6studies_20260306_pseudobulk.Rds")

pb = pb_orig[,pb_orig@meta.data %>% filter(infection_agent %in% c("LCMV Armstrong", "Listeria monocytogenes"), infection_phase %in% c("acute", "memory"), dataset %in% c("GSE121002_Ciucci_LCMVArm", "GSE182320_Andreatta_LCMVArmCl13", "GSE281547_Osum_Lm"), !(level2_final %in% c("CD4.wM", "CD4.wZ", "CD4.wN")), ncells > 10) %>% rownames()]
dim(pb)

#check that there are at least 2 replicates per cluster timepoint 
pb@meta.data %>% count(infection_phase, level2_final)

dge = DGEList(counts = pb@assays$RNA$counts)
dge = calcNormFactors(dge)

meta = pb@meta.data
meta %>% count(dataset, level2_final,  infection_agent, infection_phase)
meta$level2_final = factor(meta$level2_final, levels = unique(meta$level2_final))
meta$infection_phase = factor(meta$infection_phase, levels = c("acute", "memory"))
meta$infection_agent = make.names(meta$infection_agent)
meta$infection_agent = factor(meta$infection_agent, levels= c("LCMV.Armstrong", "Listeria.monocytogenes"))
meta$dataset = factor(meta$dataset)

design <- model.matrix(
  ~ 0 + level2_final*infection_phase + infection_agent + dataset,
  data = meta
)
colnames(design) = gsub(colnames(design), pattern = "level2_final", replacement = "")
colnames(design) = gsub(colnames(design), pattern = "dataset", replacement = "")
colnames(design) = gsub(colnames(design), pattern = "infection_phase", replacement = "")
colnames(design) = gsub(colnames(design), pattern = "infection_agent", replacement = "")
colnames(design) = make.names(colnames(design))
colnames(design) 

#Filter genes using the design-aware approach:
keep_genes <- filterByExpr(dge, design = design)
dge <- dge[keep_genes, , keep.lib.sizes = FALSE]
dge <- calcNormFactors(dge)

library(limma)
v = voom(dge, design, plot = TRUE)
fit = lmFit(v, design)

tt_list = list()

#acute vs memory across all clusters and both infection agents
contrast <- makeContrasts(
    memory,
    levels = design
)
fit2 <- contrasts.fit(fit, contrast)
fit2 <- eBayes(fit2)
tt = topTable(fit2, number = Inf, adjust.method = "BH", sort.by = "none")
head(tt)
i = "memory_vs_acute"
tt_list[[i]] = tt
tt_list[[i]]$SYMBOL = rownames(tt)

##acute vs memory in LCMV vs Listeria: mostly sex genes!
contrast <- makeContrasts(
    memory+Listeria.monocytogenes,
    levels = design
)
fit2 <- contrasts.fit(fit, contrast)
fit2 <- eBayes(fit2)
tt = topTable(fit2, number = Inf, adjust.method = "BH", sort.by = "none")
head(tt)
i = "memoryListeria_vs_memoryLCMV"
tt_list[[i]] = tt
tt_list[[i]]$SYMBOL = rownames(tt)

##acute vs memory in CD4.Q
contrast <- makeContrasts(
    CD4.Q.memory,
    levels = design
)
fit2 <- contrasts.fit(fit, contrast)
fit2 <- eBayes(fit2)
tt = topTable(fit2, number = Inf, adjust.method = "BH", sort.by = "none")
head(tt)
i = "CD4.Q.memory_vs_othermemory"
tt_list[[i]] = tt
tt_list[[i]]$SYMBOL = rownames(tt)




# vplot = data.frame(fc = 2^(tt$logFC), pval = tt$adj.P.Val, SYMBOL = rownames(tt))
# ZemmourLib::Vplot(vplot, ylimits = c(10^-50, 1), xlimits = NULL) + ggtitle(label = "memory_vs_acute_all")

# pdf(sprintf("%s/DiffExpression_volcano_%s.pdf", output_dir, prefix_file_name), width = 10, height = 10)
# i = "memory_vs_acute_all2"
# for (i in 1:length(tmp)) {
    print(i)
    vplot = na.omit(data.frame(SYMBOL = tt_list[[i]]$SYMBOL, fc = 2^tt_list[[i]]$logFC, pval = tt_list[[i]]$P.Value + 10^-300, AveExpr = 2^(tt_list[[i]]$AveExpr))) #dealing with NA
    # vplot = data.frame(SYMBOL = tmp[[i]]$SYMBOL, fc = 2^tmp[[i]]$logFC, pval = tmp[[i]]$P.Value+10^-300, AveExpr = 2^(tmp[[i]]$AveExpr))
    p = Vplot(vplot = vplot, xlab = names(tmp)[i], xlimits = c(min(vplot$fc[vplot$fc != 0]), max(vplot$fc)), ylimits = c(min(vplot$pval), 1)) #c(min(vplot$pval[vplot$pval != 0]),1)
    sub = vplot %>%
        filter(abs(log2(fc)) > 0.5 & pval < 0.05) %>%
        arrange(desc(log2(fc))) %>%
        pull(SYMBOL)
    highlight_genes = unique(c(head(sub, 50), tail(sub, 50)))
    # sub = (log2(vplot$fc) > 0.5 | log2(vplot$fc) < -0.5)  & vplot$pval < 0.05
    # sub = vplot[sub,] %>% arrange(desc(log2(fc)))
    # sub1 = sub %>% pull(SYMBOL) %>% head(25) #vplot$SYMBOL %in% as.character(sub2$SYMBOL)[1:200]
    # sub2 = sub %>% pull(SYMBOL) %>% tail(25)
    highlight_filter = vplot$SYMBOL %in% highlight_genes
    # sub = vplot$SYMBOL %in% c(sub1, sub2)#as.character(sub2$SYMBOL)[1:50]
    library(ggrepel)
    print(p + geom_text_repel(data = vplot[highlight_filter,], aes(x = fc, y = pval, label = SYMBOL)))
    # print(p + geom_text(data = vplot[highlight_filter,], aes(x = fc, y = pval, label = SYMBOL)))
# }
# dev.off()

#LCMV Arm SLO - acute vs memory
#Listeria SLO - acute vs memory
#LCMV Arm and Listeria - acute vs memory
#LCMV Arm SLO - acute vs memory in each cluster
#Listeria SLO - acute vs memory


```


DGE limma - in LCMV only
```{r}
library(edgeR)

pb_orig = readRDS("trbi_6studies_20260306_pseudobulk.Rds")

pb = pb_orig[,pb_orig@meta.data %>% filter(infection_agent %in% c("LCMV Armstrong"), infection_phase %in% c("acute", "memory"), dataset %in% c("GSE121002_Ciucci_LCMVArm", "GSE182320_Andreatta_LCMVArmCl13"), !(level2_final %in% c("CD4.wM", "CD4.wZ", "CD4.wN")), ncells > 50) %>% rownames()]
dim(pb)

#check that there are at least 2 replicates per cluster timepoint 
pb@meta.data %>% count(infection_phase, level2_final)

dge = DGEList(counts = pb@assays$RNA$counts)
dge = calcNormFactors(dge)

meta = pb@meta.data
meta %>% count(dataset, level2_final,  infection_agent, infection_phase)
meta$level2_final = factor(meta$level2_final, levels = unique(meta$level2_final))
meta$infection_phase = factor(meta$infection_phase, levels = c("acute", "memory"))
#meta$infection_agent = make.names(meta$infection_agent)
#meta$infection_agent = factor(meta$infection_agent, levels= c("LCMV.Armstrong", "Listeria.monocytogenes"))
meta$dataset = factor(meta$dataset)

design <- model.matrix(
  ~ 0 + level2_final*infection_phase + dataset,
  data = meta
)
colnames(design) = gsub(colnames(design), pattern = "level2_final", replacement = "")
colnames(design) = gsub(colnames(design), pattern = "dataset", replacement = "")
colnames(design) = gsub(colnames(design), pattern = "infection_phase", replacement = "")
colnames(design) = gsub(colnames(design), pattern = "infection_agent", replacement = "")
colnames(design) = make.names(colnames(design))
colnames(design) 

#Filter genes using the design-aware approach:
keep_genes <- filterByExpr(dge, design = design)
dge <- dge[keep_genes, , keep.lib.sizes = FALSE]
dge <- calcNormFactors(dge)

library(limma)
v = voom(dge, design, plot = TRUE)
fit = lmFit(v, design)

tt_list = list()

#acute vs memory across all clusters and both infection agents
contrast <- makeContrasts(
    memory,
    levels = design
)
fit2 <- contrasts.fit(fit, contrast)
fit2 <- eBayes(fit2)
tt = topTable(fit2, number = Inf, adjust.method = "BH", sort.by = "none")
head(tt)
i = "memory_vs_acute"
tt_list[[i]] = tt
tt_list[[i]]$SYMBOL = rownames(tt)

##acute vs memory in CD4.Q
contrast <- makeContrasts(
    CD4.Q.memory+memory,
    levels = design
)
fit2 <- contrasts.fit(fit, contrast)
fit2 <- eBayes(fit2)
tt = topTable(fit2, number = Inf, adjust.method = "BH", sort.by = "none")
head(tt)
i = "memory_vs_acute_inCD4Q"
tt_list[[i]] = tt
tt_list[[i]]$SYMBOL = rownames(tt)

##acute vs memory in CD4.B
contrast <- makeContrasts(
    CD4.B.memory,
    levels = design
)
fit2 <- contrasts.fit(fit, contrast)
fit2 <- eBayes(fit2)
tt = topTable(fit2, number = Inf, adjust.method = "BH", sort.by = "none")
head(tt)
i = "memory_vs_acute_inCD4B"
tt_list[[i]] = tt
tt_list[[i]]$SYMBOL = rownames(tt)

# vplot = data.frame(fc = 2^(tt$logFC), pval = tt$adj.P.Val, SYMBOL = rownames(tt))
# ZemmourLib::Vplot(vplot, ylimits = c(10^-50, 1), xlimits = NULL) + ggtitle(label = "memory_vs_acute_all")

# pdf(sprintf("%s/DiffExpression_volcano_%s.pdf", output_dir, prefix_file_name), width = 10, height = 10)
# i = "memory_vs_acute_all2"
# for (i in 1:length(tmp)) {
    print(i)
    vplot = na.omit(data.frame(SYMBOL = tt_list[[i]]$SYMBOL, fc = 2^tt_list[[i]]$logFC, pval = tt_list[[i]]$P.Value + 10^-300, AveExpr = 2^(tt_list[[i]]$AveExpr))) #dealing with NA
    # vplot = data.frame(SYMBOL = tmp[[i]]$SYMBOL, fc = 2^tmp[[i]]$logFC, pval = tmp[[i]]$P.Value+10^-300, AveExpr = 2^(tmp[[i]]$AveExpr))
    p = Vplot(vplot = vplot, xlab = names(tmp)[i], xlimits = c(min(vplot$fc[vplot$fc != 0]), max(vplot$fc)), ylimits = c(min(vplot$pval), 1)) #c(min(vplot$pval[vplot$pval != 0]),1)
    sub = vplot %>%
        filter(abs(log2(fc)) > 0.5 & pval < 0.05) %>%
        arrange(desc(log2(fc))) %>%
        pull(SYMBOL)
    highlight_genes = unique(c(head(sub, 50), tail(sub, 50)))
    # sub = (log2(vplot$fc) > 0.5 | log2(vplot$fc) < -0.5)  & vplot$pval < 0.05
    # sub = vplot[sub,] %>% arrange(desc(log2(fc)))
    # sub1 = sub %>% pull(SYMBOL) %>% head(25) #vplot$SYMBOL %in% as.character(sub2$SYMBOL)[1:200]
    # sub2 = sub %>% pull(SYMBOL) %>% tail(25)
    highlight_filter = vplot$SYMBOL %in% highlight_genes
    # sub = vplot$SYMBOL %in% c(sub1, sub2)#as.character(sub2$SYMBOL)[1:50]
    library(ggrepel)
    print(p + geom_text_repel(data = vplot[highlight_filter,], aes(x = fc, y = pval, label = SYMBOL)))
    # print(p + geom_text(data = vplot[highlight_filter,], aes(x = fc, y = pval, label = SYMBOL)))
# }
# dev.off()

#LCMV Arm SLO - acute vs memory
#LCMV Arm SLO - acute vs memory in each cluster
#Listeria SLO - acute vs memory
#LCMV Arm and Listeria - acute vs memory

#Listeria SLO - acute vs memory


```

**Memory Figure **

Plot MDE immgenT - SLO in LCMV arm and Listeria
```{r}
sub = so@meta.data %>% filter(infection_agent %in% c("ActA-deficient Listeria monocytogenes (DPL1942)-2W", "LCMV Armstrong"), organ_simplified %in% c("SLO", "LN", "spleen"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43")) %>% count(sample_code, sample_type, organ_simplified)
tmp = so[,so@meta.data %>% filter(infection_agent %in% c("ActA-deficient Listeria monocytogenes (DPL1942)-2W", "LCMV Armstrong"), organ_simplified %in% c("SLO", "LN", "spleen"), IGT %in% c("IGT36", "IGT38", "IGT40", "IGT43")) %>% rownames()]
tmp$organ_simplified0 = "SLO"

tmp@meta.data %>% count(infection_agent, timepoint, Ag_spe)

tmp@meta.data %>% filter(infection_agent == "LCMV Armstrong" & Ag_spe == "SMARTA") %>% count(timepoint)
sub = tmp@meta.data %>% filter(infection_agent == "LCMV Armstrong" & Ag_spe == "SMARTA") %>% group_by(timepoint) %>% slice_sample(n = 121) %>% ungroup() %>% pull(cellID)
tmp2 = tmp[,sub]
pdf(sprintf("%s/MDE_LCMV_timecourse.pdf",out_dir),7, 7, useDingbats = F)
for (i in unique(tmp2$timepoint)) {
    print(i)
    tmp3 = tmp2[,tmp2@meta.data %>% filter(timepoint == i) %>% rownames()]
    
    p = DimPlot(tmp3, reduction = "mde_incremental", group.by = "annotation_level2", cols = mypal_level2, pt.size = 2) + ggtitle(i)
    bg = AddEmbeddingRasterBackground(lineage = "CD4")
    p$layers = c(list(bg), p$layers)
    p = p + ggplot2::coord_fixed() + xlim(c(-2.5,4)) + ylim(c(-2.5,2.5)) +theme_void() + NoLegend()
    print(p)
}
dev.off()

tmp@meta.data %>% filter(infection_agent == "ActA-deficient Listeria monocytogenes (DPL1942)-2W") %>% count(timepoint)
sub = tmp@meta.data %>% filter(infection_agent == "ActA-deficient Listeria monocytogenes (DPL1942)-2W") %>% group_by(timepoint) %>% slice_sample(n = 500) %>% ungroup() %>% pull(cellID)
tmp2 = tmp[,sub]
pdf(sprintf("%s/MDE_Listeria_timecourse.pdf",out_dir),7, 7, useDingbats = F)
for (i in unique(tmp2$timepoint)) {
    print(i)
    tmp3 = tmp2[,tmp2@meta.data %>% filter(timepoint == i) %>% rownames()]
    
    p = DimPlot(tmp3, reduction = "mde_incremental", group.by = "annotation_level2", cols = mypal_level2, pt.size = 2) + ggtitle(i)
    bg = AddEmbeddingRasterBackground(lineage = "CD4")
    p$layers = c(list(bg), p$layers)
    p = p + ggplot2::coord_fixed() + xlim(c(-2.5,4)) + ylim(c(-2.5,2.5)) +theme_void() + NoLegend()
    print(p)
}
dev.off()
```

Plot MDE immgenT CD4 Trm
```{r}

so@meta.data %>% filter(condition_broad %in% c("bacteria", "virus"), !(organ_simplified %in% c("SLO", "LN", "spleen"))) %>% count(sample_code, sample_type, organ_simplified, Ag_spe)

so@meta.data %>% filter(infection_agent %in% c("LCMV Armstrong"), !(organ_simplified %in% c("SLO", "LN", "spleen")), Ag_spe == "SMARTA", IGT %in% c("IGT36", "IGT38", "IGT40")) %>% count(sample_code, sample_type, organ_simplified)
tmp = so[,so@meta.data %>% filter(infection_agent %in% c("LCMV Armstrong"), !(organ_simplified %in% c("SLO", "LN", "spleen")), Ag_spe == "SMARTA", IGT %in% c("IGT36", "IGT38", "IGT40"))  %>% rownames()]
tmp$organ_simplified0 = "NLT"

tmp@meta.data %>% count(infection_agent, timepoint, Ag_spe)

tmp@meta.data %>% filter(infection_agent == "LCMV Armstrong" & Ag_spe == "SMARTA") %>% count(timepoint)
sub = tmp@meta.data %>% filter(infection_agent == "LCMV Armstrong" & Ag_spe == "SMARTA") %>% group_by(timepoint) %>% slice_sample(n = 121) %>% ungroup() %>% pull(cellID)
tmp2 = tmp[,sub]
pdf(sprintf("%s/MDE_LCMV_timecourse_NLT.pdf",out_dir),7, 7, useDingbats = F)
for (i in unique(tmp2$timepoint)) {
    print(i)
    tmp3 = tmp2[,tmp2@meta.data %>% filter(timepoint == i) %>% rownames()]
    
    p = DimPlot(tmp3, reduction = "mde_incremental", group.by = "annotation_level2", cols = mypal_level2, pt.size = 2) + ggtitle(i)
    bg = AddEmbeddingRasterBackground(lineage = "CD4")
    p$layers = c(list(bg), p$layers)
    p = p + ggplot2::coord_fixed() + xlim(c(-2.5,4)) + ylim(c(-2.5,2.5)) +theme_void() + NoLegend()
    print(p)
}
dev.off()

```

For CB
```{r}
so@meta.data %>% count(Ag_spe) 
so@meta.data %>% filter(condition_broad != "healthy" & infection_phase %in% c("acute", "cleared")) %>% count(condition_detailed_simplified, organ_simplified, infection_phase, IGT, IGTHT) %>% write.table("~/Desktop/memory_figure_IGT.txt", sep = "\t", quote = F, row.names = F)

so@meta.data %>% filter(condition_detailed_simplified == "CtrachomatisD") %>% select(mouse_comments) %>% unique()
```








