---
title: "cluster_adjustments"
output: html_document
date: "2024-11-14"
editor_options: 
  chunk_output_type: console
---

#########Treg#########
Tregs are fine

#########Miniverse#########
Miniverse as is for now

#########unconvventional#########
Remove the “unconventional” category altogether (it is pretty sketchy), and re-arrange at annotation level1:
gdT: 2, 4, 5, 7, 9, 10, 11, 12, 23, 26 , 28, 32, 33, 35, 42, 43, 46
TCRb+ DN:  3,  15, 18, 29, 31, 34, 39, 41
TCRb+ CD8aa: 1, 16, 20, 40
NonConv: 8, 13, 17, 21, 24, 27, 45 and also include the cells in the attached file tcr.recog.nonconvs.csv (NKT, MAIT and others)
proliferating: 48
 
Remove these cells entirely: 
unconv_cl49                    (no CD3, looks like an ILC)
unconv_cl30                    (suspect, one experiment only)
unconv_cl38                    “”
unconv_cl36                    (tiny cluster, not clear TCR+)
 
Move to CD8ab:
unconv_cl6                       (8ab positive and live in that region)
unconv_cl22                    “”
unconv_cl44                    “”
 
Move to CD4:
unconv_cl19                       (4 positive close to 22)
 
Set aside for independent checking:
unconv_cl14, 25, 37, 47                       These are the B-T ones, need a closer look

#########CD8#########

Move to unconv: CD8_cl3_dp_barcode.csv --> not_CD4_CD8AB_cl51 --> abT_DP_cl51
Move to CD4: 8
Remove: 16 (crappy, ribo/rna, unstable)
Merge 12+17

#########CD4#########
remove: 6 (nonT),24 (non spe binding)
move to CD8: 11
move to unconv: 19

#########

**Start interactive session on Midway3**
```{bash}
ssh -Y zemmour@midway3.rcc.uchicago.edu
sinfo -s
screen -S david
sinteractive --exclusive --time=12:00:00 --account=pi-zemmour --partition=jfkfloor2 --qos=jfkfloor2 --nodes=1 
sinteractive --time=12:00:00 --account=pi-zemmour --partition=caslake --nodes=1 --mem=128GB
sinteractive --time=12:00:00 --account=pi-zemmour --partition=gpu --gres=gpu:1 --mem=128GB
sinteractive --time=12:00:00 --account=pi-zemmour --partition=beagle3 --gres=gpu:1 --mem=96GB

cd $LABHOME

module load python/anaconda-2022.05 
module load hdf5/1.12.0 #for BPCells but also others
module load openblas/0.3.13
source activate /project/zemmour/david/envs/scvi_20240315 #for R and old scvi
module load openblas/0.3.13 #load again or error

module load python
source activate /project/zemmour/david/envs/scvi120_20241008 #for scvi 1.2.0
cd /project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96

R
options(max.print=1000)
#options(Seurat.object.assay.version = 'v5')
setwd("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96")


```

**Initialize: load libraries and custom functions**

```{r}
options(max.print=1000)
setwd("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96")
source("/project/jfkfloor2/zemmourlab/david/immgent/immgent_integration_git/custom_functions_david.R")

libs = c("fgsea","fastTopics", "flashier", "Matrix", "Seurat","BPCells", "ggplot2", "dplyr", "reshape2", "ggrastr", "RColorBrewer", "pals", "scales", "pheatmap") 
#libs = c("matrixStats", "gplots","ggplot2", "reshape2", "scales", "gridExtra", "dplyr", "RColorBrewer", "grid", "Rtsne", "limma",   "RColorBrewer", "pheatmap", "Seurat",  "ggrastr", "ggbeeswarm") #pwr, preprocessCore, "readxl", vegan", "genefilter" "scde" "cellrangerRkit", "Rsamtools", "GenomicRanges", "GenomicAlignments", "VGAM", "WGCNA",rafalib, ‘UsingR’,MAST "ggrastr",
sapply(libs, function(x) suppressMessages(suppressWarnings(library(x, character.only = TRUE, quietly = T, warn.conflicts  = F))))


#Gradients
library(RColorBrewer)
ColorRamp = rev(colorRampPalette(brewer.pal(n = 7,name = "RdYlBu"))(100))
ColorRamp = rev(colorRampPalette(c('red','white','blue'))(20))
ColorRamp = rev(rainbow(10, end = 4/6))
library(viridis)
ColorRamp = rev(viridis(100))
ColorRamp = rev(cividis(100))
#image(1:100, 1, as.matrix(1:100), col = ColorRamp, main = "Color Palette", xaxt = 'n', yaxt = 'n', xlab = "", ylab = "")

#mycol = c("black","blue","red","darkgreen", "orange","purple", "cyan", "greenyellow",    "salmon", "magenta","pink", "tan", "brown") # c("magenta", "red",  "darkgreen", "cyan", "blue", "blue", "blue", "black", "blue", "blue", "blue", "blue", "orange")

#Large Color palette
library("pals")
n = 70
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
mypal1 = unique(unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals))))
mypal1 = mypal1[-4]
mypal = c(glasbey(), polychrome(), mypal1)
names(mypal) = NULL
#pal.bands(mypal, main="Colormap suggestions")

#parade = function(n) { return(Seurat::DiscretePalette(n, palette = "parade", shuffle = F)) }
```

**update annotation levels**

```{r}
#so = readRDS("igt1_96_withtotalvi20241028.Rds")

setwd("/Users/david/google_drive_uchicago/ImmgenT/analysis/data_integration/IGT1_96")

annotation_level2 = read.table("annotation_level2.csv", header = T, sep = ",", as.is = T, row.names = 1)
annotation_level2$cell_id = rownames(annotation_level2)
#write.table(x = names(table(annotation_level2$annotation_level2)), file = "annotation_level2_list.csv", row.names = F, quote = F)
#update by hand when possible

newannot = read.table("annotation_level2_list_20241113.csv", header = T, sep = ",", as.is = T)
#table(newannot$level2, newannot$level1)

annotation_table = merge(x = annotation_level2, y = newannot, by.x = "annotation_level2", by.y = "annotation_level2_v1", all.x = T)
apply(annotation_table, 2, function(x) length(which(is.na(x))))
annotation_table = annotation_table[match(annotation_level2$cell_id,annotation_table$cell_id),]
rownames(annotation_table) = annotation_table$cell_id
annotation_table$cell_id = NULL
all(annotation_table$annotation_level2 == annotation_level2$annotation_level2)

#add to nonconv: tcr.recog.nonconvs.csv --> not_CD4CD8_cl51
nonconv1 = read.table("tcr.recog.nonconvs.csv", header = F, sep = ",")[,1]
annotation_table$nonconv_tcr_recog = rownames(annotation_table) %in% nonconv1
#add to not_CD4CD8: CD8_cl3_dp_barcode.csv --> not_CD4CD8_cl51
nonconv2 = read.table("CD8_cl3_dp_barcode.csv", header = F, sep = ",")[,1]
annotation_table$level2[rownames(annotation_table) %in% nonconv2] = "not_CD4CD8_cl51"
annotation_table$level1[rownames(annotation_table) %in% nonconv2] = "not_CD4CD8"
annotation_table$level0[rownames(annotation_table) %in% nonconv2] = "not_CD4CD8"
head(annotation_table)
table(annotation_table$nonconv_tcr_recog, annotation_table$level1)

write.table(x = annotation_table, file = "annotation_table_20241113.csv", row.names = T, quote = F, sep = ",")

```

--> update annotation_table_20241113 in seurat object --> metadata_update.Rmd /project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/igt1_96_withtotalvi20241113.Rds
```{r}
so_orig = readRDS("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/igt1_96_withtotalvi20241113.Rds")
so = so_orig[,!so_orig$annotation_level0 %in% c("remove", "nonT")]
prefix = "totalvi_20241006"
umap_to_plot = "mde2_totalvi_20241006"

pdf(sprintf("%s/MDE_annotationV2_202141114.pdf", prefix), 20, 20, useDingbats = F)
DimPlot(so, reduction = umap_to_plot, group.by = "annotation_level2_oct24", raster = T, raster.dpi = c(1024,1024), label = F) + scale_color_manual(values = mypal)
DimPlot(so, reduction = umap_to_plot, group.by = "annotation_level0", raster = T, raster.dpi = c(1024,1024), label = F) + scale_color_manual(values = mypal)
DimPlot(so, reduction = umap_to_plot, group.by = "annotation_level1", raster = T, raster.dpi = c(1024,1024), label = F) + scale_color_manual(values = mypal)
DimPlot(so, reduction = umap_to_plot, group.by = "annotation_level2", raster = T, raster.dpi = c(1024,1024), label = T) + scale_color_manual(values = mypal)
DimPlot(so, reduction = umap_to_plot, group.by = "annotation_level2", raster = T, raster.dpi = c(1024,1024), label = T) + scale_color_manual(values = mypal) + NoLegend()
DimPlot(so, reduction = umap_to_plot, group.by = "nonconv_tcr_recog", raster = T, raster.dpi = c(1024,1024), label = F) + scale_color_manual(values = mypal)
DimPlot(so, reduction = umap_to_plot, group.by = "iNKT", raster = T, raster.dpi = c(1024,1024), label = F) + scale_color_manual(values = mypal)
DimPlot(so, reduction = umap_to_plot, group.by = "MAIT", raster = T, raster.dpi = c(1024,1024), label = F) + scale_color_manual(values = mypal)
# DimPlot(so, reduction = umap_to_plot, group.by = "annotation_level1", split.by = "organ_simplified", ncol = trunc(sqrt(length(unique(so$organ_simplified)))), raster = T, raster.dpi = c(1024,1024), label = F) + scale_color_manual(values = mypal)
dev.off()
```

add in mudata to calculate new MDE
```{python}
import warnings; warnings.simplefilter('ignore')
import scvi
import os
import sys
import scanpy as sc
import muon as mu
import mudata as mu

import numpy as np
import mplscience
import matplotlib.pyplot as plt
import pickle
import pandas as pd

working_dir='/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96'
path_to_mudata='/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/export_data/igt1_96_20241006.h5mu'

mdata = mu.read(path_to_mudata) 

annotationl1 = pd.read_csv(working_dir+"/annotation_level1.csv", index_col = 0, header = None)
mdata.mod['RNA'].obs['annotation_level1'] = annotationl1
mdata.mod['RNA'].obs['annotation_level1'].isna().sum()

mdata_treg = mdata[mdata.mod['RNA'].obs['annotation_level1'] == 'Treg'] 
mdata_treg.write(working_dir+'/export_data/igt1_96_20241014_Treg.h5mu')

mdata_cd4 = mdata[mdata.mod['RNA'].obs['annotation_level1'] == 'CD4'] 
mdata_cd4.write(working_dir+'/export_data/igt1_96_20241014_CD4.h5mu')

mdata_cd8 = mdata[mdata.mod['RNA'].obs['annotation_level1'] == 'CD8AB'] 
mdata_cd8.write(working_dir+'/export_data/igt1_96_20241014_CD8AB.h5mu')

mdata_unconv = mdata[mdata.mod['RNA'].obs['annotation_level1'] == 'unconventional'] 
mdata_unconv.write(working_dir+'/export_data/igt1_96_20241014_unconventional.h5mu')

mdata_miniverse = mdata[mdata.mod['RNA'].obs['annotation_level1'] == 'miniverse'] 
mdata_miniverse.write(working_dir+'/export_data/igt1_96_20241014_miniverse.h5mu')

mdata_treg.mod['RNA'].obs['annotation_level1'].unique()
mdata_cd4.mod['RNA'].obs['annotation_level1'].unique()
mdata_cd8.mod['RNA'].obs['annotation_level1'].unique()
mdata_unconv.mod['RNA'].obs['annotation_level1'].unique()
mdata_miniverse.mod['RNA'].obs['annotation_level1'].unique()

###Subset CD4:
annotation_cd4_l1 = pd.read_csv(working_dir+"/annotation_CD4_clusters_l1.csv", index_col = 0, header = None)
mdata.mod['RNA'].obs['annotation_cd4_l1'] = annotation_cd4_l1
mdata.mod['RNA'].obs['annotation_cd4_l1'].isna().sum()
mdata_cd4_act = mdata[mdata.mod['RNA'].obs['annotation_cd4_l1'] == 'CD4_activated']
mdata_cd4_act.mod['RNA'].obs['annotation_cd4_l1'].unique()
mdata_cd4_act.write(working_dir+'/export_data/igt1_96_20241014_CD4_activated.h5mu')

mdata_cd4_rest = mdata[mdata.mod['RNA'].obs['annotation_cd4_l1'] == 'CD4_resting'] 
mdata_cd4_rest.mod['RNA'].obs['annotation_cd4_l1'].unique()
mdata_cd4_rest.write(working_dir+'/export_data/igt1_96_20241014_CD4_resting.h5mu')

mdata_cd4_prolif = mdata[mdata.mod['RNA'].obs['annotation_cd4_l1'] == 'CD4_prolif'] 
mdata_cd4_prolif.mod['RNA'].obs['annotation_cd4_l1'].unique()
mdata_cd4_prolif.write(working_dir+'/export_data/igt1_96_20241014_CD4_prolif.h5mu')

```

**Individual MDE by cell type from the overall totalvi_20241008_rmIGTsample + subset mudata in separate files**

```{bash}
sinteractive --time=12:00:00 --account=pi-zemmour --partition=gpu --gres=gpu:1 --mem=128GB

cd $LABHOME

module load python
source activate /project/zemmour/david/envs/scvi120_20241008 #for scvi 1.2.0

#cd /project/jfkfloor2/zemmourlab/david/immgent/analysis/integration/IGT1_56
cd /project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96
```

```{python}
# module load python
# source activate /project/zemmour/david/envs/scvi120_20241008 #for scvi 1.2.0

# working_dir = '/project/jfkfloor2/zemmourlab/david/immgent/analysis/integration/IGT1_56/'
# path_to_mudata = '/project/jfkfloor2/zemmourlab/david/immgent/analysis/integration/IGT1_56/export_data/totalvi_igt1_56_20231030_allgenes_mdata.h5mu'
# prefix = 'totalvi_igt1_56_allgenes_20240526_igtsampleregressedout'
# batchkey = 'IGT'
# categorical_covariate_keys = ['sample_id']

working_dir='/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96'
path_to_mudata='/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/export_data/igt1_96_20241006.h5mu'
#path_to_mudata=/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/export_data/igt1_96_20241014_CD4.h5mu
path_to_annotation=working_dir+'/annotation_table_20241113.csv' #annotation_level1.csv
prefix = 'totalvi_20241008_rmIGTsample' #prefix='totalvi_20241006'
celltype='not_CD4CD8' #treg, CD4,CD8AB, not_CD4CD8
# batchkey='IGT'
# categorical_covariate_keys='IGTHT'
# corrected_counts=False
# denoised_data=False

print("Importing libraries")
import warnings; warnings.simplefilter('ignore')
import scvi
import os
import sys
import scanpy as sc
import muon as mu
import mudata as mu

import numpy as np
import mplscience
import matplotlib.pyplot as plt
import pickle
import pandas as pd
import seaborn as sns
import torch
import pymde #to run MDE
pymde.seed(0)

print("Global configurations")
pd.set_option('display.max_rows', 1000)
pd.set_option('display.max_columns', 1000)
# sc.set_figure_params(figsize=(6, 6), frameon=False)
scvi.settings.seed = 0  # optional: ensures reproducibility
pymde.seed(0)
#sns.set_theme()
if torch.cuda.is_available():
    print("CUDA is available")
    print("Using device:", torch.cuda.get_device_name())
    torch.set_float32_matmul_precision("high")

print("Reading mudata")
os.chdir(working_dir)
mdata = mu.read(path_to_mudata) #totalvi_igt1_56_allgenes_Treg_20240306/adata.h5mu") #mu.read("totalvi_igt1_56_allgenes_Treg_20240306/adata.h5mu
print(mdata)

print("Add latent variable")
latent_df = pd.read_csv(prefix+"/latent.csv", index_col = 0)
TOTALVI_LATENT_KEY = "X_totalVI"
mdata.obsm[TOTALVI_LATENT_KEY] = latent_df
# mdata.obsm[TOTALVI_LATENT_KEY] = mdata.obsm[TOTALVI_LATENT_KEY].values

annotationl1 = pd.read_csv(path_to_annotation, index_col = 0, header = 0)
print(mdata.mod['RNA'].obs.index.equals(annotationl1.index)) 
mdata.mod['RNA'].obs['annotation_level1'] = annotationl1['level0']
mdata.mod['RNA'].obs['annotation_level1'].isna().sum()
mdata.mod['RNA'].obs['annotation_level1'].unique()

mdata_orig = mdata

print(celltype)
mdata = mdata_orig[mdata_orig.mod['RNA'].obs['annotation_level1'] == celltype] 
mdata

mde = pymde.preserve_neighbors(
    mdata.obsm[TOTALVI_LATENT_KEY].values, #mdata.obsm[TOTALVI_LATENT_KEY]
    embedding_dim=2,
    constraint=pymde.Standardized(),
    repulsive_fraction=0.7,
    n_neighbors = 15,
    verbose=True,
    device = 'cuda'
)

import pickle
model_pkl_file = prefix+'/mde_'+prefix+'_in'+celltype+'_model.pkl' #prefix+"/mde_model.pkl"  
with open(model_pkl_file, 'wb') as file:  
    pickle.dump(mde, file)

# with open(model_pkl_file, 'rb') as file:  
#     mde = pickle.load(file)

embedding = mde.embed(verbose=True)
mde_df = pd.DataFrame(embedding.cpu(), index = mdata.mod['RNA'].obs.index) 
mde_df.to_csv(prefix+'/mde_'+prefix+'_in'+celltype+'.csv', index=True)

print("Save mde.csv")
# mdata.obsm["X_mde"] = scvi.model.utils.mde(mdata.obsm[TOTALVI_LATENT_KEY])
# mde_df = pd.DataFrame(mdata.obsm["X_mde"], index = mdata.mod['RNA'].obs.index)
# mde_df.to_csv(prefix+'/mde_'+prefix+'_'+celltype+'.csv', index=True)

print("save subset mudata for CD4, CD8, not_CD4CD8")
mdata = mdata_orig

mdata_cd4 = mdata[mdata.mod['RNA'].obs['annotation_level1'] == 'CD4']
mdata_cd4.mod['RNA'].obs['annotation_level1'].unique()
mdata_cd4.update()
mdata_cd4.write(working_dir+'/export_data/igt1_96_20241113_CD4.h5mu')

mdata_cd8 = mdata[mdata.mod['RNA'].obs['annotation_level1'] == 'CD8AB'] 
mdata_cd8.mod['RNA'].obs['annotation_level1'].unique()
mdata_cd8.update()
mdata_cd8.write(working_dir+'/export_data/igt1_96_20241113_CD8AB.h5mu')

mdata_unconv = mdata[mdata.mod['RNA'].obs['annotation_level1'] == 'not_CD4CD8'] 
mdata_unconv.mod['RNA'].obs['annotation_level1'].unique()
mdata_unconv.update()
mdata_unconv.write(working_dir+'/export_data/igt1_96_20241113_not_CD4CD8.h5mu')


# pairs, distorsions = mde.high_distortion_pairs()
# pairs_df = pd.DataFrame(pairs.cpu(), columns=['cell1', 'cell2'])
# pairs_df['cell1'] = mdata.mod['RNA'].obs.index[pairs_df['cell1']]
# pairs_df['cell2'] = mdata.mod['RNA'].obs.index[pairs_df['cell2']]
# distorsions_df = pd.DataFrame(distorsions.cpu(),columns=['distortion'])
# pairs_df['distortion'] = distorsions_df['distortion']
# #pairs_df.sample(n=1000000, replace=False).to_csv(prefix+"/mde_distorsions_sample.csv")
# pairs_df.to_csv(prefix+"/mde_distorsions.csv")
# #most_distorted_pairs = pairs[:500]
# mde.distortions_cdf()
# fig = plt.gcf() 
# fig.savefig(prefix+"/mde_distorsion_cdf.png")

# umap_mde_aligned = pymde.align(source=embedding.cpu(), target=mdata.obsm['X_umap'])
# umap_mde_aligned_df = pd.DataFrame(umap_mde_aligned, index = mdata.mod['RNA'].obs.index)
# umap_mde_aligned_df.to_csv(prefix+"/umap_mde_aligned.csv", index=True)

```


**New group integration and incremental MDE**

split mdata: see above
sbatch $SCRIPT_DIR/run_totalvi_wrapper_20241113_cd4_rmIGTsample.sh
sbatch $SCRIPT_DIR/run_totalvi_wrapper_20241113_cd8_rmIGTsample.sh
sbatch $SCRIPT_DIR/run_totalvi_wrapper_20241113_not_CD4CD8_rmIGTsample.sh

Anchor MDE from: https://github.com/cvxgrp/pymde/blob/main/examples/updating_an_existing_embedding.ipynb
sbatch $SCRIPT_DIR/run_mde_incremental_wrapper_20241121_notCD4CD8.sh
sbatch $SCRIPT_DIR/run_mde_incremental_wrapper_20241121_CD8AB.sh
sbatch $SCRIPT_DIR/run_mde_incremental_wrapper_20241121_CD4.sh

TEST 30000 CD4 + 30000 all other cells
1.subset the data
```{python}

working_dir='/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/'
path_to_mudata='/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/export_data/igt1_96_20241006.h5mu'
prefix='test_mde'
mde_ref_file='CD4/totalvi_20241014_CD4_rmIGTsample/mde2.csv'
totalvi_integrated_file='totalvi_20241006/latent.csv'

path_to_annotation=working_dir+'/annotation_table_20241113.csv' #annotation_level1.csv

print("Importing libraries")
import warnings; warnings.simplefilter('ignore')
import scvi
import os
import sys
import scanpy as sc
import muon as mu
import mudata as mu

import numpy as np
import mplscience
import matplotlib.pyplot as plt
import pickle
import pandas as pd
import seaborn as sns
import torch
import pymde #to run MDE
pymde.seed(0)

print("Global configurations")
pd.set_option('display.max_rows', 1000)
pd.set_option('display.max_columns', 1000)
# sc.set_figure_params(figsize=(6, 6), frameon=False)
scvi.settings.seed = 0  # optional: ensures reproducibility
pymde.seed(0)
#sns.set_theme()
if torch.cuda.is_available():
    print("CUDA is available")
    print("Using device:", torch.cuda.get_device_name())
    torch.set_float32_matmul_precision("high")

print("Reading mudata")
os.chdir(working_dir)
mdata = mu.read(path_to_mudata) #totalvi_igt1_56_allgenes_Treg_20240306/adata.h5mu") #mu.read("totalvi_igt1_56_allgenes_Treg_20240306/adata.h5mu
print(mdata)

annotationl1 = pd.read_csv(path_to_annotation, index_col = 0, header = 0)
print(mdata.mod['RNA'].obs.index.equals(annotationl1.index)) 
mdata.mod['RNA'].obs['annotation_level1'] = annotationl1['level0']
mdata.mod['RNA'].obs['annotation_level1'].isna().sum()
mdata.mod['RNA'].obs['annotation_level1'].unique()

cd4_indices = np.where(mdata.mod['RNA'].obs['annotation_level1'] == 'CD4')[0]
random_cd4_indices = np.random.choice(cd4_indices, size=30000, replace=False)
cd4_indices = mdata.mod['RNA'].obs.index[random_cd4_indices]
non_cd4_indices = np.where(mdata.mod['RNA'].obs['annotation_level1'] != 'CD4')[0]
random_non_cd4_indices = np.random.choice(non_cd4_indices, size=30000, replace=False)
random_non_cd4_indices = mdata.mod['RNA'].obs.index[random_non_cd4_indices]
combined_indices = np.concatenate((cd4_indices, random_non_cd4_indices))
cd4_indices.size, random_non_cd4_indices.size, combined_indices.size

#subset totalvi_integrated with combined_indices only
totalvi_integrated = pd.read_csv(totalvi_integrated_file, index_col = 0, header = 0)
totalvi_integrated.index.size
totalvi_integrated = totalvi_integrated.loc[totalvi_integrated.index.intersection(combined_indices)] 
totalvi_integrated.index.size

totalvi_integrated.to_csv(prefix+'/latent.csv' , index=True)


mde_ref_embedding = pd.read_csv(mde_ref_file, index_col = 0, header = 0)
mde_ref_embedding.index.size
mde_ref_embedding = mde_ref_embedding.loc[mde_ref_embedding.index.intersection(combined_indices)] #subset cells present in the 
mde_ref_embedding.index.size
mde_ref_embedding.to_csv(prefix+'/mde2.csv' , index=True)
```
2.sbatch $SCRIPT_DIR/run_mde_incremental_wrapper_20241121_test.sh
3.Plot
mde_totalvi_20241008_rmIGTsample_innot_CD4CD8: MDE from totalvi_20241008_rmIGT in not_CD4CD8 cells only
```{r}
AddLatentData = function(so, latent_file, prefix, calculate_umap = F) {
    totalvi = read.csv(file = latent_file, header = T, sep = ",", row.names = 1)
    # pattern = "(IGT\\S*)_(\\d+)"
    # rownames(totalvi) = gsub(pattern, "\\1", rownames(totalvi))
    colnames(totalvi) = as.character(seq(1:ncol(totalvi)))
    #all(rownames(totalvi) == colnames(so))
    totalvi = as.matrix(totalvi)
    pattern = "^IGT\\d{1,2}\\.[A-Z]{16}$"
    matches = grepl(pattern, rownames(totalvi))
    if (!all(matches)) {
        print("updating cell names in csv file, assuming same order as seurat object" )
        #rownames(totalvi) = gsub(pattern = "(\\S*)_(\\d+)", replacement = "\\1", rownames(totalvi))
        rownames(totalvi) = colnames(so)
    }
    totalvi = totalvi[colnames(so),]
    if (all(colnames(so) == rownames(totalvi))) {
        print("all cells ARE in the latent_file")
        so[[prefix]] = CreateDimReducObject(embeddings = totalvi, key = prefix, assay = "RNA") #sprintf("totalvi_%s", prefix2)
        if (calculate_umap == T) {
            so = RunUMAP(so, dims = 1:ncol(totalvi), reduction = prefix, n.components = 2, reduction.name = sprintf("umap_%s", prefix))
        }
        return(so)
    } else {
        print("all cells ARE NOT in the latent_file")
        return(so)
    }
}


so_orig = readRDS("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/igt1_96_withtotalvi20241113.Rds")
setwd("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/test_mde")

dir_with_file = "/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/test_mde"
tmp = read.table(sprintf("%s/mde_incremental.csv",dir_with_file), header = T, sep = ",", row.names = 1)
so = so_orig[,rownames(tmp)]

so = AddLatentData(so, latent_file = sprintf("%s/mde_incremental.csv",dir_with_file), prefix = "mde_incremental", calculate_umap = F)

pdf(sprintf("%s/MDE_clusters2.pdf", "."), 20, 20, useDingbats = F)
DimPlot(so[,so$annotation_level1 == "CD4"], reduction = "mde_incremental", group.by = "annotation_level1", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = "green")
DimPlot(so, reduction = "mde_incremental", group.by = "annotation_level1", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal)
DimPlot(so, reduction = "mde_incremental", group.by = "annotation_level2", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal)
DimPlot(so, reduction = "mde_incremental", group.by = "annotation_level2_oct24", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal)
dev.off()

```

**Split nonconv for new totalvi and update annotation table**
--> annotation_table_20241201.csv
abT_DP: unclear_cl14 --> abT_DP_cl14
not_CD4CD8_cl51 --> abT_DP_cl51
not_CD4CD8_cl50 --> gdT_cl50

```{r}
# setwd("/Users/david/google_drive_uchicago/ImmgenT/analysis/data_integration/IGT1_96/")
library(reshape2)
library(dplyr)
annot = read.table("annotation_table_20241113.csv", header = T, sep = ",")
table(annot$level2)
annot$level2[annot$level2 == "unclear_cl14"] = "abT_DP_cl14"
annot$level1[annot$level2 == "abT_DP_cl14"] = "abT_DP"
annot$level2[annot$level2 == "not_CD4CD8_cl51"]  = "abT_DP_cl51"
annot$level1[annot$level2 == "abT_DP_cl51"]  = "abT_DP"
annot$level2[annot$level2 == "not_CD4CD8_cl50"] = "gdT_cl50"
annot$level1[annot$level2 == "gdT_cl50"] = "gdT"

for (i in unique(annot$level1)) {
    print(i)
    print(annot %>% filter(level1 == i) %>% pull(level2) %>% unique())
}


annot$level0 = NULL
write.table(x = annot, file = "annotation_table_20241201.csv", row.names = F, quote = F, sep = ",")

annot_table = table(annot$level1, annot$level2) %>% as.data.frame() %>% melt() %>% filter(value != 0)
annot_table$variable = NULL
colnames(annot_table) = c("level1", "level2", "n_cells")

write.table(x = annot_table, file = "annotation_table_history_20241201.csv", row.names = F, quote = F, sep = ",")

```

add in mudata to calculate new TOTALVI for each subgroup of unconventional
```{python}
import warnings; warnings.simplefilter('ignore')
import scvi
import os
import sys
import scanpy as sc
import muon as mu
import mudata as mu

import numpy as np
import mplscience
import matplotlib.pyplot as plt
import pickle
import pandas as pd

working_dir='/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96'
path_to_mudata='/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/export_data/igt1_96_20241006.h5mu'

mdata = mu.read(path_to_mudata) 

annotationl1 = pd.read_csv(working_dir+"/annotation_table_20241201.csv", index_col = 0, header = 0)
set(annotationl1.index) == set(mdata.mod['RNA'].obs.index)
mdata.mod['RNA'].obs['level1'] = annotationl1['level1']
mdata.mod['RNA'].obs['level1'].isna().sum()

mdata.mod['RNA'].obs['level1'].unique()

mdata_gdT = mdata[mdata.mod['RNA'].obs['level1'] == 'gdT'] 
mdata_gdT.write(working_dir+'/export_data/igt1_96_20241201_gdT.h5mu')
mdata_gdT

mdata_abTCD8aa = mdata[mdata.mod['RNA'].obs['level1'] == 'abT_CD8aa'] 
mdata_abTCD8aa
mdata_abTCD8aa.write(working_dir+'/export_data/igt1_96_20241201_abT_CD8aa.h5mu')

mdata_nonconv = mdata[mdata.mod['RNA'].obs['level1'] == 'nonconv'] 
mdata_nonconv
mdata_nonconv.write(working_dir+'/export_data/igt1_96_20241201_nonconv.h5mu')

mdata_abT_DN = mdata[mdata.mod['RNA'].obs['level1'] == 'abT_DN'] 
mdata_abT_DN
mdata_abT_DN.write(working_dir+'/export_data/igt1_96_20241201_abT_DN.h5mu')


mdata_gdT.mod['RNA'].obs['level1'].unique()
mdata_abTCD8aa.mod['RNA'].obs['level1'].unique()
mdata_nonconv.mod['RNA'].obs['level1'].unique()
mdata_abT_DN.mod['RNA'].obs['level1'].unique()

```

sbatch $SCRIPT_DIR/run_totalvi_wrapper_20241201_gdT_rmIGTsample.sh
sbatch $SCRIPT_DIR/run_totalvi_wrapper_20241201_abT_CD8aa_rmIGTsample.sh
sbatch $SCRIPT_DIR/run_totalvi_wrapper_20241201_nonconv_rmIGTsample.sh
sbatch $SCRIPT_DIR/run_totalvi_wrapper_20241201_abT_DN_rmIGTsample.sh

--> see clustering_unconventional_igt_1_96.Rmd

**remove outliers in abT_DN and abT_CD8aa and update annotation table**
remove outliers in abT_DN and abT_CD8aa --> annotation_table_20241202.csv
```{r}
# setwd("/Users/david/google_drive_uchicago/ImmgenT/analysis/data_integration/IGT1_96/")
setwd("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96")
library(reshape2)
library(dplyr)
annot = read.table("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/annotation_table_20241201.csv", header = T, sep = ",")
dim(annot)
outliers1 = read.table("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/abT_DN/outliers_abT_DN.csv", header = F, sep = ",")[,1]
outliers2 = read.table("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/abT_CD8aa/outliers_abT_CD8aa.csv", header = F, sep = ",")[,1]
annot$level2[annot$cell_id %in% c(outliers1, outliers2)] = "remove"
annot$level1[annot$cell_id %in% c(outliers1, outliers2)] = "remove"

for (i in unique(annot$level1)) {
    print(i)
    print(annot %>% filter(level1 == i) %>% pull(level2) %>% unique())
}

annot$level0 = NULL
write.table(x = annot, file = "annotation_table_20241202.csv", row.names = F, quote = F, sep = ",")

annot_table = table(annot$level1, annot$level2) %>% as.data.frame() %>% melt() %>% filter(value != 0)
annot_table$variable = NULL
colnames(annot_table) = c("level1", "level2", "n_cells")

write.table(x = annot_table, file = "annotation_table_history_20241202.csv", row.names = F, quote = F, sep = ",")

```

remove in mudata and rerun totalvi
```{python}
import warnings; warnings.simplefilter('ignore')
import scvi
import os
import sys
import scanpy as sc
import muon as mu
import mudata as mu

import numpy as np
import mplscience
import matplotlib.pyplot as plt
import pickle
import pandas as pd

working_dir='/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96'
path_to_mudata='/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/export_data/igt1_96_20241006.h5mu'

mdata = mu.read(path_to_mudata) 

annotationl1 = pd.read_csv(working_dir+"/annotation_table_20241202.csv", index_col = 0, header = 0)
set(annotationl1.index) == set(mdata.mod['RNA'].obs.index)
mdata.mod['RNA'].obs['level1'] = annotationl1['level1']
mdata.mod['RNA'].obs['level1'].isna().sum()

mdata.mod['RNA'].obs['level1'].unique()

mdata_abTCD8aa = mdata[mdata.mod['RNA'].obs['level1'] == 'abT_CD8aa'] 
mdata_abTCD8aa
mdata_abTCD8aa.write(working_dir+'/export_data/igt1_96_20241202_abT_CD8aa.h5mu')


mdata_abT_DN = mdata[mdata.mod['RNA'].obs['level1'] == 'abT_DN'] 
mdata_abT_DN
mdata_abT_DN.write(working_dir+'/export_data/igt1_96_20241202_abT_DN.h5mu')

mdata_abTCD8aa.mod['RNA'].obs['level1'].unique()
mdata_abT_DN.mod['RNA'].obs['level1'].unique()

```

sbatch $SCRIPT_DIR/run_totalvi_wrapper_20241202_abT_CD8aa_rmIGTsample.sh
sbatch $SCRIPT_DIR/run_totalvi_wrapper_20241202_abT_DN_rmIGTsample.sh

**remove more outliers in abT_CD8aa and update annotation table** WASN'T USEFUL -- FORGET
remove outliers in abT_CD8aa --> annotation_table_20241203.csv
```{r}
# setwd("/Users/david/google_drive_uchicago/ImmgenT/analysis/data_integration/IGT1_96/")
setwd("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96")
library(reshape2)
library(dplyr)
annot = read.table("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/annotation_table_20241202.csv", header = T, sep = ",")
dim(annot)
outliers2 = read.table("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/abT_CD8aa/outliers_abT_CD8aa_20241202.csv", header = F, sep = ",")[,1]
annot$level2[annot$cell_id %in% c(outliers2)] = "remove"
annot$level1[annot$cell_id %in% c(outliers2)] = "remove"

for (i in unique(annot$level1)) {
    print(i)
    print(annot %>% filter(level1 == i) %>% pull(level2) %>% unique())
}

annot$level0 = NULL
write.table(x = annot, file = "annotation_table_20241203.csv", row.names = F, quote = F, sep = ",")

annot_table = table(annot$level1, annot$level2) %>% as.data.frame() %>% melt() %>% filter(value != 0)
annot_table$variable = NULL
colnames(annot_table) = c("level1", "level2", "n_cells")

write.table(x = annot_table, file = "annotation_table_history_20241203.csv", row.names = F, quote = F, sep = ",")

```

remove in mudata and rerun totalvi
```{python}
import warnings; warnings.simplefilter('ignore')
import scvi
import os
import sys
import scanpy as sc
import muon as mu
import mudata as mu

import numpy as np
import mplscience
import matplotlib.pyplot as plt
import pickle
import pandas as pd

working_dir='/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96'
path_to_mudata='/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/export_data/igt1_96_20241006.h5mu'

mdata = mu.read(path_to_mudata) 

annotationl1 = pd.read_csv(working_dir+"/annotation_table_20241203.csv", index_col = 0, header = 0)
set(annotationl1.index) == set(mdata.mod['RNA'].obs.index)
mdata.mod['RNA'].obs['level1'] = annotationl1['level1']
mdata.mod['RNA'].obs['level1'].isna().sum()

mdata.mod['RNA'].obs['level1'].unique()

mdata_abTCD8aa = mdata[mdata.mod['RNA'].obs['level1'] == 'abT_CD8aa'] 
mdata_abTCD8aa
mdata_abTCD8aa.write(working_dir+'/export_data/igt1_96_20241203_abT_CD8aa.h5mu')

mdata_abTCD8aa.mod['RNA'].obs['level1'].unique()

```
sbatch $SCRIPT_DIR/run_totalvi_wrapper_20241203_abT_CD8aa_rmIGTsample.sh

**Merge clusters in CD4 and CD8**
CD4: merge  cl21/26/27 as cl1
CD8: merge cl1/3 as cl1
     merge  cl9/10/21 as cl10
     merge  cl22/23 as 22

```{r}
# setwd("/Users/david/google_drive_uchicago/ImmgenT/analysis/data_integration/IGT1_96/")
setwd("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96")
library(reshape2)
library(dplyr)
annot = read.table("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/annotation_table_20241202.csv", header = T, sep = ",")
dim(annot)

annot$level2[annot$level2 == "CD4_cl26"] = "CD4_cl21"
annot$level2[annot$level2 == "CD4_cl27"] = "CD4_cl21"

annot$level2[annot$level2 == "CD8AB_cl3"] = "CD8AB_cl1"

annot$level2[annot$level2 == "CD8AB_cl9"] = "CD8AB_cl10"
annot$level2[annot$level2 == "CD8AB_cl21"] = "CD8AB_cl10"

annot$level2[annot$level2 == "CD8AB_cl23"] = "CD8AB_cl22"

for (i in unique(annot$level1)) {
    print(i)
    print(annot %>% filter(level1 == i) %>% pull(level2) %>% unique())
}

annot$level0 = NULL
annot$annotation_level2 = NULL

#Add resting, activated
annot$level2_parent1 = "not_applicable"
annot$level2_parent1[annot$level2 %in% c("CD4_cl10", "CD4_cl13", "CD4_cl14", "CD4_cl15", "CD4_cl16", "CD4_cl17", "CD4_cl18", "CD4_cl20", "CD4_cl21", "CD4_cl22", "CD4_cl23", "CD4_cl25", "CD4_cl9", "CD8AB_cl10", "CD8AB_cl11", "CD8AB_cl12", "CD8AB_cl13", "CD8AB_cl14", "CD8AB_cl15", "CD8AB_cl18", "CD8AB_cl20", "CD8AB_cl6", "CD8AB_cl7","treg_cl2", "treg_cl3", "treg_cl4", "treg_cl6", "treg_cl8")] = "activated"
annot$level2_parent1[annot$level2 %in% c("CD4_cl1", "CD4_cl12", "CD4_cl2", "CD4_cl3", "CD4_cl4", "CD4_cl5", "CD4_cl7", "CD8AB_cl1", "CD8AB_cl2", "CD8AB_cl22", "CD8AB_cl4", "CD8AB_cl5","treg_cl1", "treg_cl7"
)] = "resting"
annot$level2_parent1[annot$level2 %in% c("CD4_cl8", "CD8AB_cl19", "treg_cl5", "not_CD4CD8_cl48")] = "proliferating"
table(annot$level2_parent1, annot$level2)
table(annot$level2_parent1, annot$level1)

sum(rowSums(apply(annot,2,is.na)))

write.table(x = annot, file = "annotation_table_20241216.csv", row.names = F, quote = F, sep = ",")

annot_summary <- annot %>%
  group_by(level2, level1, level2_parent1) %>% 
  summarize(counts = n(), .groups = "drop") %>%
  arrange(
    sub("^.*_cl", "", level2) %>% as.numeric(), # Extract cluster number and convert to numeric
    level2                                     # Keep cluster prefix order (CD4, treg, etc.)
  ) %>% 
    arrange(level1) %>%
    as.data.frame()


write.table(x = annot_summary, file = "annotation_table_history_20241216.csv", row.names = F, quote = F, sep = ",")

```

**Level1 adjustment based on CITEseq CD4, CD8B, CD8A**
```{r}
mypal_level1 = setNames(mypal, unique(so$annotation_level1))
mypal_odhran = setNames(mypal, unique(so$odhran))

so_orig = readRDS("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/igt1_96_withtotalvi20241216.Rds")
so = so_orig[,!so_orig$annotation_level1 %in% c("nonT","remove")]
so$annotation_level1 = factor(so$annotation_level1)

discordantCD4 = read.table("DiscordantCD4.csv", header = T, sep = ",", row.names = 1)
discordantCD8ab = read.table("DiscordantCD8ab.csv", header = T, sep = ",", row.names = 1)
discordant = data.frame(odhran = c(discordantCD4$X3, discordantCD8ab$DBlabel))
rownames(discordant) = c(rownames(discordantCD4), rownames(discordantCD8ab))
so$odhran = discordant[colnames(so),"odhran"]
so$odhran[is.na(so$odhran)] = "same"

so = NormalizeData(so,assay = "ADT", normalization.method = "CLR", verbose = T)

pdf("Discordant_CD4CD8.pdf", 10,10, useDingbats = F)
FeatureScatter(so[,so$odhran != "same"], feature1 = "CD4", feature2 = "CD8B", group.by = "odhran", split.by = "annotation_level1", ncol = 2, pt.size = 2, raster = F, plot.cor = F) + scale_color_manual(values = mypal_level1)  + scale_color_manual(values = mypal_odhran) + 
    geom_hline(aes(yintercept = 2.2), linetype="dashed", color = "brown") +
    geom_vline(aes(xintercept = 1.5), linetype="dashed", color = "brown")
FeatureScatter(so[,so$odhran != "same"], feature1 = "CD4", feature2 = "CD8A", group.by = "odhran", split.by = "annotation_level1", ncol = 2, pt.size = 2, raster = F, plot.cor = F) + scale_color_manual(values = mypal_level1)  + scale_color_manual(values = mypal_odhran) + 
    geom_hline(aes(yintercept = 2.2), linetype="dashed", color = "brown") +
    geom_vline(aes(xintercept = 1.5), linetype="dashed", color = "brown")

for (i in c("CD4+", "CD8AA+", "CD8AB+")) {
    print(i)
    p = MyDimPlotHighlight(seurat_object = so, umap_to_plot = "mde2_totalvi_20241006", cells_to_highlight = colnames(so)[so$odhran %in% i], highlight_column_name = "annotation_level1", labelclusters = T,
                           title = i, pixels = c(512, 512), highlight_size = 1, 
                           highlight_alpha = 1, mycols = mypal,print_plot1 = F, 
                           print_plot2 = F)
    print(p)
}

dev.off()


cd8_inCD4 = names(which(so[["ADT"]]$data["CD8B",so$annotation_level1 == "CD4"] > 2 & so[["ADT"]]$data["CD4",so$annotation_level1 == "CD4"] < 1))
length(which(so[["ADT"]]$data["CD8B",so$annotation_level1 == "CD4"] > 2 & so[["ADT"]]$data["CD4",so$annotation_level1 == "CD4"] < 1))

so$cd8_inCD4 = colnames(so) %in% cd8_inCD4

# FeatureScatter(so, feature1 = "CD4", feature2 = "CD8B", split.by = NULL) + scale_color_manual(values = mypal_level1)  + scale_color_manual(values = mypal_odhran) + 
    geom_hline(aes(yintercept = 2.2), linetype="dashed", color = "brown") +
    geom_vline(aes(xintercept = 1.5), linetype="dashed", color = "brown")


# MyDimPlotHighlight(seurat_object = so, umap_to_plot = "mde2_totalvi_20241006", cells_to_highlight = colnames(so)[so$cd8_inCD4], highlight_column_name = "annotation_level1", labelclusters = T,
#                            title = "", pixels = c(512, 512), highlight_size = 1, 
#                            highlight_alpha = 1, mycols = mypal,print_plot1 = F, 
#                            print_plot2 = F)

```

add cd8_inCD4 in mudata to calculate new TOTALVI for each subgroup of unconventional
```{python}
import warnings; warnings.simplefilter('ignore')
import scvi
import os
import sys
import scanpy as sc
import muon as mu
import mudata as mu

import numpy as np
import mplscience
import matplotlib.pyplot as plt
import pickle
import pandas as pd

working_dir='/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96'
path_to_mudata='/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/export_data/igt1_96_20241006.h5mu'

mdata = mu.read(path_to_mudata) 

annotationl1 = pd.read_csv(working_dir+"/annotation_table_20241216.csv", index_col = 0, header = 0)
set(annotationl1.index) == set(mdata.mod['RNA'].obs.index)
mdata.mod['RNA'].obs['level1'] = annotationl1['level1']
mdata.mod['RNA'].obs['level1'].isna().sum()

discordant = pd.read_csv(working_dir+"/DiscordantCD4_david.csv", header = None)
discordant.columns = ['cell_id']
mdata.mod['RNA'].obs.loc[discordant.cell_id, 'level1'] = "CD8AB"
mdata.mod['RNA'].obs['level1'].unique()

mdata_CD8AB = mdata[mdata.mod['RNA'].obs['level1'] == 'CD8AB'] 
mdata_CD8AB
mdata_CD8AB.write(working_dir+'/export_data/igt1_96_20250103_CD8AB.h5mu')

```

sbatch $SCRIPT_DIR/run_totalvi_wrapper_20250103_cd8_rmIGTsample.sh



