---
title: "clustering_abT_DN"
output: html_document
date: "2024-12-01"
editor_options: 
  chunk_output_type: console
---

**Start interactive session on Midway3**
```{bash}
ssh -Y zemmour@midway3.rcc.uchicago.edu
sinfo -s
screen -S david
sinteractive --exclusive --time=12:00:00 --account=pi-zemmour --partition=jfkfloor2 --qos=jfkfloor2 --nodes=1 
sinteractive --time=12:00:00 --account=pi-zemmour --partition=caslake --nodes=1 --mem=128GB
sinteractive --time=12:00:00 --account=pi-zemmour --partition=gpu --gres=gpu:1 --mem=152GB
cd $LABHOME

module load python/anaconda-2022.05 
module load hdf5/1.12.0 #for BPCells but also others
module load openblas/0.3.13

#sinteractive --time=1:00:00 -p gpu --gres=gpu:1 --account=pi-zemmour 

#source activate /project/jfkfloor2/zemmourlab/david/envs/scvi
source activate /project/zemmour/david/envs/scvi_20240315
module load openblas/0.3.13 #load again or error

cd /project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/abT_DN
#cd ~/google_drive/ImmgenT/analysis/data_integration/IGT1_56/Treg

R
#options(max.print=1000)
#options(Seurat.object.assay.version = 'v5')
```

**Initialize: load libraries and custom functions**

```{r}
options(max.print=1000)
setwd("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/abT_DN")
source("/project/jfkfloor2/zemmourlab/david/immgent/immgent_integration_git/custom_functions_david.R")

libs = c("fgsea","fastTopics", "flashier", "Matrix", "Seurat","BPCells", "ggplot2", "dplyr", "reshape2", "ggrastr", "RColorBrewer", "pals", "scales", "pheatmap", "scattermore") 
#libs = c("matrixStats", "gplots","ggplot2", "reshape2", "scales", "gridExtra", "dplyr", "RColorBrewer", "grid", "Rtsne", "limma",   "RColorBrewer", "pheatmap", "Seurat",  "ggrastr", "ggbeeswarm") #pwr, preprocessCore, "readxl", vegan", "genefilter" "scde" "cellrangerRkit", "Rsamtools", "GenomicRanges", "GenomicAlignments", "VGAM", "WGCNA",rafalib, ‘UsingR’,MAST "ggrastr",
sapply(libs, function(x) suppressMessages(suppressWarnings(library(x, character.only = TRUE, quietly = T, warn.conflicts  = F))))


#Gradients
library(RColorBrewer)
ColorRamp = rev(colorRampPalette(brewer.pal(n = 7,name = "RdYlBu"))(100))
ColorRamp = rev(colorRampPalette(c('red','white','blue'))(20))
ColorRamp = rev(rainbow(10, end = 4/6))
library(viridis)
ColorRamp = rev(viridis(100))
ColorRamp = rev(cividis(100))
#image(1:100, 1, as.matrix(1:100), col = ColorRamp, main = "Color Palette", xaxt = 'n', yaxt = 'n', xlab = "", ylab = "")

#mycol = c("black","blue","red","darkgreen", "orange","purple", "cyan", "greenyellow",    "salmon", "magenta","pink", "tan", "brown") # c("magenta", "red",  "darkgreen", "cyan", "blue", "blue", "blue", "black", "blue", "blue", "blue", "blue", "orange")

#Large Color palette
library("pals")
n = 70
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
mypal1 = unique(unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals))))
mypal1 = mypal1[-4]
mypal = c(glasbey(), polychrome(), mypal1)
names(mypal) = NULL
#pal.bands(mypal, main="Colormap suggestions")

#parade = function(n) { return(Seurat::DiscretePalette(n, palette = "parade", shuffle = F)) }


```

**clustering V3 20241201**
 cluster_adjustments.Rmd: sbatch $SCRIPT_DIR/run_totalvi_wrapper_20241201_abT_DN_rmIGTsample.sh
 igt1_96_withtotalvi20241202.Rds
 annotation_table_20241202.csv
annotation_table_history_20241202.csv

```{r}
prefix = "totalvi_20241201_abT_DN_rmIGTsample"
umap_to_plot = "mde_totalvi_20241201_abT_DN_rmIGTsample"

so_orig = readRDS("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/igt1_96_withtotalvi20241201.Rds")
setwd("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/abT_DN")
so = so_orig[,so_orig$annotation_level1 == "abT_DN"]

dir_with_file = "/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/abT_DN/totalvi_20241201_abT_DN_rmIGTsample"
so = AddLatentData(so, latent_file = sprintf("%s/latent.csv",dir_with_file), prefix = "totalvi_20241201_abT_DN_rmIGTsample", calculate_umap = F)
so = AddLatentData(so, latent_file = sprintf("%s/mde2.csv",dir_with_file), prefix = "mde_totalvi_20241201_abT_DN_rmIGTsample", calculate_umap = F)


pdf(sprintf("%s/MDE_clusters.pdf", prefix), 20, 20, useDingbats = F)
DimPlot(so, reduction = umap_to_plot, group.by = "annotation_level1", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal)
DimPlot(so, reduction = umap_to_plot, group.by = "annotation_level2", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal)
MyDimPlotHighlight(seurat_object = so, umap_to_plot = umap_to_plot, cells_to_highlight = names(which(so$nonconv_tcr_recog == T)), highlight_column = "nonconv_tcr_recog", pixels = c(512, 512), mycols = "red")
MyDimPlotHighlight(seurat_object = so, umap_to_plot = umap_to_plot, cells_to_highlight = names(which(so$iNKT == T)), highlight_column = "iNKT", pixels = c(512, 512), mycols = "red")
MyDimPlotHighlight(seurat_object = so, umap_to_plot = umap_to_plot, cells_to_highlight = names(which(so$MAIT == T)), highlight_column = "MAIT", pixels = c(512, 512), mycols = "red")
dev.off()

outliers = names(which(so[[umap_to_plot]]@cell.embeddings[,2] >5))
write.table(outliers, file = "outliers_abT_DN.csv", sep = ",", quote = F, row.names = F, col.names = F)
# outliers = read.table("outliers_abT_DN.csv")[,1]

so$outliers_abT_DN = colnames(so) %in% outliers
so = so[,!colnames(so) %in% outliers]
saveRDS(so, file = "igt1_96_abT_DN_20241201.Rds")

pdf(sprintf("%s/MDE_clusters_nooutliers.pdf", prefix), 20, 20, useDingbats = F)
DimPlot(so, reduction = umap_to_plot, group.by = "annotation_level1", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal)
DimPlot(so, reduction = umap_to_plot, group.by = "annotation_level2", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal)
MyDimPlotHighlight(seurat_object = so, umap_to_plot = umap_to_plot, cells_to_highlight = names(which(so$nonconv_tcr_recog == T)), highlight_column = "nonconv_tcr_recog", pixels = c(512, 512), mycols = "red")
MyDimPlotHighlight(seurat_object = so, umap_to_plot = umap_to_plot, cells_to_highlight = names(which(so$iNKT == T)), highlight_column = "iNKT", pixels = c(512, 512), mycols = "red")
MyDimPlotHighlight(seurat_object = so, umap_to_plot = umap_to_plot, cells_to_highlight = names(which(so$MAIT == T)), highlight_column = "MAIT", pixels = c(512, 512), mycols = "red")
dev.off()

outliers = read.table("outliers_abT_DN.csv")[,1]
totalvi = read.csv(file = sprintf("%s/latent.csv",dir_with_file), header = T, sep = ",", row.names = 1) 
totalvi2 = totalvi[!rownames(totalvi) %in% outliers,]
write.table(totalvi2, file = sprintf("%s/latent_nooutliers.csv",dir_with_file), sep = ",", quote = F, row.names = T, col.names = T)

mde = read.csv(file = sprintf("%s/mde2.csv",dir_with_file), header = T, sep = ",", row.names = 1) 
mde2 = mde[!rownames(mde) %in% outliers,]
write.table(mde2, file = sprintf("%s/mde2_nooutliers.csv",dir_with_file), sep = ",", quote = F, row.names = T, col.names = T)


```


**Update Seurat object 20241216**
same clusters, adding level1, level2_parent1 etc

```{r}
AddLatentData = function(so, latent_file, prefix, calculate_umap = F) {
    totalvi = read.csv(file = latent_file, header = T, sep = ",", row.names = 1)
    # pattern = "(IGT\\S*)_(\\d+)"
    # rownames(totalvi) = gsub(pattern, "\\1", rownames(totalvi))
    colnames(totalvi) = as.character(seq(1:ncol(totalvi)))
    #all(rownames(totalvi) == colnames(so))
    totalvi = as.matrix(totalvi)
    pattern = "^IGT\\d{1,2}\\.[A-Z]{16}$"
    matches = grepl(pattern, rownames(totalvi))
    if (!all(matches)) {
        print("updating cell names in csv file, assuming same order as seurat object" )
        #rownames(totalvi) = gsub(pattern = "(\\S*)_(\\d+)", replacement = "\\1", rownames(totalvi))
        rownames(totalvi) = colnames(so)
    }
    totalvi = totalvi[colnames(so),]
    if (all(colnames(so) == rownames(totalvi))) {
        print("all cells ARE in the latent_file")
        so[[prefix]] = CreateDimReducObject(embeddings = totalvi, key = prefix, assay = "RNA") #sprintf("totalvi_%s", prefix2)
        if (calculate_umap == T) {
            so = RunUMAP(so, dims = 1:ncol(totalvi), reduction = prefix, n.components = 2, reduction.name = sprintf("umap_%s", prefix))
        }
        return(so)
    } else {
        print("all cells ARE NOT in the latent_file")
        return(so)
    }
}

so_orig = readRDS("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/igt1_96_withtotalvi20241216.Rds")
setwd("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/abT_DN")
so = so_orig[,so_orig$annotation_level1 == "abT_DN"] # so = so_orig[,so_orig$annotation_level0 == "CD4"]

table_data = table(so$annotation_level2, so$annotation_level1)
# pdf(sprintf("%s/ClustersOct24_Newclusters.pdf", "."), width = 10, height = 10)
# 
# # Convert the table to a data frame for better readability in the PDF
# table_df <- as.data.frame(table_data)
# 
# # Create the table in the PDF
# grid.table(table_df)
# 
# # Close the PDF device
# dev.off()

```

Add general totalvi
```{r}
dir_with_file = "/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/totalvi_20241006"
so = AddLatentData(so, latent_file = sprintf("%s/latent.csv",dir_with_file), prefix = "totalvi_20241006", calculate_umap = F)
so = AddLatentData(so, latent_file = sprintf("%s/mde2.csv",dir_with_file), prefix = "mde2_totalvi_20241006", calculate_umap = F)

```

Add not_CD4CD8 totalvi
```{r}
dir_with_file = "/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/unconventional/totalvi_20241113_not_CD4CD8_rmIGTsample"
so = AddLatentData(so, latent_file = sprintf("%s/latent.csv",dir_with_file), prefix = "totalvi_20241113_not_CD4CD8_rmIGTsample", calculate_umap = F)
so = AddLatentData(so, latent_file = sprintf("%s/mde_incremental.csv",dir_with_file), prefix = "mde_totalvi_20241113_not_CD4CD8_rmIGTsample", calculate_umap = F)

```

Add abT_DN specific MDE
```{r}
dir_with_file = "/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/abT_DN/totalvi_20241201_abT_DN_rmIGTsample"
so = AddLatentData(so, latent_file = sprintf("%s/latent.csv",dir_with_file), prefix = "totalvi_20241201_abT_DN_rmIGTsample", calculate_umap = F)
so = AddLatentData(so, latent_file = sprintf("%s/mde2.csv",dir_with_file), prefix = "mde_totalvi_20241201_abT_DN_rmIGTsample", calculate_umap = F)

```

Plots
```{r}

pdf(sprintf("%s/MDE_AnnotationLevel20241216.pdf", "."), 10, 10, useDingbats = F)
MyDimPlotHighlight(so_orig, umap_to_plot = "mde2_totalvi_20241006", cells_to_highlight = names(which(so_orig$annotation_level1 == "abT_DN")), highlight_column_name ="annotation_level2", highlight_size = 0.5, highlight_alpha = 0.5, pixels = c(512, 512), mycols = mypal, print_plot1 = , print_plot2 = F, labelclusters = T )
# MyDimPlotHighlight(so, umap_to_plot = "mde_totalvi_20241113_not_CD4CD8_rmIGTsample", cells_to_highlight = names(which(so$annotation_level1 == "abT_DN")), highlight_column_name ="annotation_level2", highlight_size = 0.5, highlight_alpha = 0.5, pixels = c(512, 512), mycols = mypal, print_plot1 = , print_plot2 = F, labelclusters = T )
DimPlot(so, reduction = "mde_totalvi_20241113_not_CD4CD8_rmIGTsample", group.by = "annotation_level2", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal)
DimPlot(so, reduction = "mde_totalvi_20241201_abT_DN_rmIGTsample", group.by = "annotation_level2", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal)
DimPlot(so, reduction = "mde_totalvi_20241201_abT_DN_rmIGTsample", group.by = "annotation_level2", raster = T, raster.dpi = c(512,512), label = F) + scale_color_manual(values = mypal) + theme_void() + NoLegend() 
# DimPlot(so, reduction = "mde_totalvi_20241201_abT_DN_rmIGTsample", group.by = "annotation_level2_parent1", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal)
# DimPlot(so, reduction = "mde_totalvi_20241201_abT_DN_rmIGTsample", group.by = "annotation_level2_parent1", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal) + theme_void() + NoLegend() 
dev.off()

saveRDS(so, file = "igt1_96_abT_DN_20241216.Rds")

```

remove unwanted cell embeddings
totalvi_20241006 - mde2_totalvi_20241006
totalvi_20241113_CD4_rmIGTsample - mde_incremental
```{r}
so[["umap_totalvi_20241006"]] = NULL
so[["mde_totalvi_20241006"]] = NULL
so[["umap2_totalvi_20241006"]] = NULL
so[["mdeumap_totalvi_20241006"]] = NULL
so[["umap_totalvi_20241008_rmIGTsample"]] = NULL
so[["mde_totalvi_20241008_rmIGTsample"]] = NULL
so[["mde2_totalvi_20241008_rmIGTsample"]] = NULL
so[["mdeumap_totalvi_20241008_rmIGTsample"]] = NULL
so[["mde_totalvi_20241029_fitleredGenesProt"]] = NULL
so[["umap_totalvi_20241029_fitleredGenesProt"]] = NULL

saveRDS(so, file = "igt1_96_abT_DN_20241216.Rds")
```



(**clustering V4 20241202** FORGET NOT USEFUL)
cluster_adjustments.Rmd: sbatch $SCRIPT_DIR/run_totalvi_wrapper_20241202_abT_CD8aa_rmIGTsample.sh
--> igt1_96_abT_CD8aa_20241202.Rds

```{r}
prefix = "totalvi_20241202_abT_DN_rmIGTsample"
umap_to_plot = "mde_totalvi_20241202_abT_DN_rmIGTsample"

so_orig = readRDS("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/igt1_96_withtotalvi20241202.Rds")
so = so_orig[,so_orig$annotation_level1 == "abT_DN"]

dir_with_file = "/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/abT_DN/totalvi_20241202_abT_DN_rmIGTsample"
so = AddLatentData(so, latent_file = sprintf("%s/latent.csv",dir_with_file), prefix = "totalvi_20241202_abT_DN_rmIGTsample", calculate_umap = F)
so = AddLatentData(so, latent_file = sprintf("%s/mde2.csv",dir_with_file), prefix = "mde_totalvi_20241202_abT_DN_rmIGTsample", calculate_umap = F)

pdf(sprintf("%s/MDE_clusters.pdf", prefix), 20, 20, useDingbats = F)
DimPlot(so, reduction = umap_to_plot, group.by = "annotation_level1", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal)
DimPlot(so, reduction = umap_to_plot, group.by = "annotation_level2", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal)
MyDimPlotHighlight(seurat_object = so, umap_to_plot = umap_to_plot, cells_to_highlight = names(which(so$nonconv_tcr_recog == T)), highlight_column = "nonconv_tcr_recog", pixels = c(512, 512), mycols = "red")
MyDimPlotHighlight(seurat_object = so, umap_to_plot = umap_to_plot, cells_to_highlight = names(which(so$iNKT == T)), highlight_column = "iNKT", pixels = c(512, 512), mycols = "red")
MyDimPlotHighlight(seurat_object = so, umap_to_plot = umap_to_plot, cells_to_highlight = names(which(so$MAIT == T)), highlight_column = "MAIT", pixels = c(512, 512), mycols = "red")
dev.off()

saveRDS(so, file = "igt1_96_abT_DN_20241202.Rds")

```

