---
title: "clustering_Tregs"
output: html_document
date: "2024-10-17"
editor_options: 
  chunk_output_type: console
---

**Start interactive session on Midway3**
```{bash}
ssh -Y zemmour@midway3.rcc.uchicago.edu
sinfo -s
screen -S david
sinteractive --exclusive --time=12:00:00 --account=pi-zemmour --partition=jfkfloor2 --qos=jfkfloor2 --nodes=1 
sinteractive --time=12:00:00 --account=pi-zemmour --partition=caslake --nodes=1 --mem=128GB
sinteractive --time=3:00:00 --account=pi-zemmour --partition=gpu --gres=gpu:1 --mem=152GB
cd $LABHOME

module load python/anaconda-2022.05 
module load hdf5/1.12.0 #for BPCells but also others
module load openblas/0.3.13

#sinteractive --time=1:00:00 -p gpu --gres=gpu:1 --account=pi-zemmour 

#source activate /project/jfkfloor2/zemmourlab/david/envs/scvi
source activate /project/zemmour/david/envs/scvi_20240315
module load openblas/0.3.13 #load again or error

cd /project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/Treg
#cd ~/google_drive/ImmgenT/analysis/data_integration/IGT1_56/Treg

R
#options(max.print=1000)
#options(Seurat.object.assay.version = 'v5')
```

**Initialize: load libraries and custom functions**

```{r}
options(max.print=1000)
setwd("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/Treg")
source("/project/jfkfloor2/zemmourlab/david/immgent/immgent_integration_git/custom_functions_david.R")

libs = c("fgsea","fastTopics", "flashier", "Matrix", "Seurat","BPCells", "ggplot2", "dplyr", "reshape2", "ggrastr", "RColorBrewer", "pals", "scales", "pheatmap", "scattermore") 
#libs = c("matrixStats", "gplots","ggplot2", "reshape2", "scales", "gridExtra", "dplyr", "RColorBrewer", "grid", "Rtsne", "limma",   "RColorBrewer", "pheatmap", "Seurat",  "ggrastr", "ggbeeswarm") #pwr, preprocessCore, "readxl", vegan", "genefilter" "scde" "cellrangerRkit", "Rsamtools", "GenomicRanges", "GenomicAlignments", "VGAM", "WGCNA",rafalib, ‘UsingR’,MAST "ggrastr",
sapply(libs, function(x) suppressMessages(suppressWarnings(library(x, character.only = TRUE, quietly = T, warn.conflicts  = F))))


#Gradients
library(RColorBrewer)
ColorRamp = rev(colorRampPalette(brewer.pal(n = 7,name = "RdYlBu"))(100))
ColorRamp = rev(colorRampPalette(c('red','white','blue'))(20))
ColorRamp = rev(rainbow(10, end = 4/6))
library(viridis)
ColorRamp = rev(viridis(100))
ColorRamp = rev(cividis(100))
#image(1:100, 1, as.matrix(1:100), col = ColorRamp, main = "Color Palette", xaxt = 'n', yaxt = 'n', xlab = "", ylab = "")

#mycol = c("black","blue","red","darkgreen", "orange","purple", "cyan", "greenyellow",    "salmon", "magenta","pink", "tan", "brown") # c("magenta", "red",  "darkgreen", "cyan", "blue", "blue", "blue", "black", "blue", "blue", "blue", "blue", "orange")

#Large Color palette
library("pals")
n = 70
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
mypal1 = unique(unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals))))
mypal1 = mypal1[-4]
mypal = c(glasbey(), polychrome(), mypal1)
names(mypal) = NULL
#pal.bands(mypal, main="Colormap suggestions")

mypal_level1 = setNames(mypal, unique(so_orig$annotation_level1))
mypal_level2 = setNames(mypal, unique(so_orig$annotation_level2))
mypal_level2_group = setNames(rev(mypal)[-3], unique(so_orig$annotation_level2_group))

#parade = function(n) { return(Seurat::DiscretePalette(n, palette = "parade", shuffle = F)) }


```

Seurat object with the results: igt1_96_treg_20241017.Rds
annotation: annotation_level2_treg.csv

```{r}

so = readRDS(file = "/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/Treg/igt1_96_treg_20241017.Rds")
so = NormalizeData(so,assay = "RNA", normalization.method = "LogNormalize",verbose = T)
so = NormalizeData(so,assay = "ADT", normalization.method = "CLR", verbose = T)

```

**Load seurat expression data**
--> igt1_96_treg_20241017.Rds
```{r}
so_orig = readRDS("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/igt1_96_withtotalvi20241006.Rds")

so = so_orig[,so_orig$annotation_level1 == "Treg"]
so[['mde_totalvi_20241006']] = NULL
so[['umap2_totalvi_20241006']] = NULL
so[['mdeumap_totalvi_20241006']] = NULL
so[['mde_totalvi_20241008_rmIGTsample']] = NULL
so[['mdeumap_totalvi_20241008_rmIGTsample']] = NULL

```

Add latent spaces, umaps
```{r}
# sprintf("%s_X_SCVI_%s.csv", prefix, prefix2)

AddLatentData = function(so, latent_file, prefix, calculate_umap = F) {
    totalvi = read.csv(file = latent_file, header = T, sep = ",", row.names = 1)
    # pattern = "(IGT\\S*)_(\\d+)"
    # rownames(totalvi) = gsub(pattern, "\\1", rownames(totalvi))
    colnames(totalvi) = as.character(seq(1:ncol(totalvi)))
    #all(rownames(totalvi) == colnames(so))
    totalvi = as.matrix(totalvi)
    pattern = "^IGT\\d{1,2}\\.[A-Z]{16}$"
    matches = grepl(pattern, rownames(totalvi))
    if (!all(matches)) {
        print("updating cell names in csv file, assuming same order as seurat object" )
        #rownames(totalvi) = gsub(pattern = "(\\S*)_(\\d+)", replacement = "\\1", rownames(totalvi))
        rownames(totalvi) = colnames(so)
    }
    totalvi = totalvi[colnames(so),]
    if (all(colnames(so) == rownames(totalvi))) {
        print("all cells ARE in the latent_file")
        so[[prefix]] = CreateDimReducObject(embeddings = totalvi, key = prefix, assay = "RNA") #sprintf("totalvi_%s", prefix2)
        if (calculate_umap == T) {
            so = RunUMAP(so, dims = 1:ncol(totalvi), reduction = prefix, n.components = 2, reduction.name = sprintf("umap_%s", prefix))
        }
        return(so)
    } else {
        print("all cells ARE NOT in the latent_file")
        return(so)
    }
}

dir_with_file = "/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/Treg/totalvi_20241014_Treg"
so = AddLatentData(so, latent_file = sprintf("%s/latent.csv",dir_with_file), prefix = "totalvi_20241014_Treg", calculate_umap = F)
so = AddLatentData(so, latent_file = sprintf("%s/umap_python.csv",dir_with_file), prefix = "umap_totalvi_20241014_Treg", calculate_umap = F)
so = AddLatentData(so, latent_file = sprintf("%s/mde2.csv",dir_with_file), prefix = "mde_totalvi_20241014_Treg", calculate_umap = F)
#write.table(so[['umap_totalvi_20241006']]@cell.embeddings, file = sprintf("%s/umap.csv",dir_with_file), quote = F, sep = ",", row.names = T, col.names = T)

dir_with_file = "/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/Treg/totalvi_20241014_Treg_rmIGTsample"
so = AddLatentData(so, latent_file = sprintf("%s/latent.csv",dir_with_file), prefix = "totalvi_20241014_Treg_rmIGTsample", calculate_umap = F)
so = AddLatentData(so, latent_file = sprintf("%s/umap_python.csv",dir_with_file), prefix = "umap_totalvi_20241014_Treg_rmIGTsample", calculate_umap = F)
so = AddLatentData(so, latent_file = sprintf("%s/mde2.csv",dir_with_file), prefix = "mde_totalvi_20241014_Treg_rmIGTsample", calculate_umap = F)

dir_with_file = "/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/totalvi_20241008_rmIGTsample"
so = AddLatentData(so, latent_file = sprintf("%s/mde_totalvi_20241008_rmIGTsample_inTreg.csv",dir_with_file), prefix = "mde_totalvi_20241008_rmIGTsample_inTreg", calculate_umap = F)


```

```{r}
prefix = "totalvi_20241014_Treg_rmIGTsample"
#setwd(prefix)
reduc = "umap_totalvi_20241014_Treg_rmIGTsample"
reduc = "mde_totalvi_20241014_Treg_rmIGTsample"
```

```{r}
RenameCluster = function(cluster.vector = so[[cluster.name]][,1], cluster.prefix = "cl") {
    cl = as.numeric(cluster.vector)
    ordered = order(unique(cl))
    cl = sprintf("%s%s", cluster.prefix, cl )
    renamed.vector = factor(cl, levels = unique(cl)[ordered])
    return(renamed.vector)
}

n_latent = ncol(so[[prefix]]@cell.embeddings)
so = FindNeighbors(so, dims = 1:n_latent, reduction = prefix)
so = FindClusters(so, resolution = 0.25, cluster.name = sprintf("Cluster_%s_Res0.25",prefix))
so = FindClusters(so, resolution = 0.5, cluster.name = sprintf("Cluster_%s_Res0.5",prefix))
so = FindClusters(so, resolution = 0.75, cluster.name = sprintf("Cluster_%s_Res0.75",prefix))
so = FindClusters(so, resolution = 1, cluster.name = sprintf("Cluster_%s_Res1",prefix))
so = FindClusters(so, resolution = 1.5, cluster.name = sprintf("Cluster_%s_Res1.5",prefix))
so = FindClusters(so, resolution = 2, cluster.name = sprintf("Cluster_%s_Res2",prefix))
so = FindClusters(so, resolution = 3, cluster.name = sprintf("Cluster_%s_Res3",prefix))
so = FindClusters(so, resolution = 4, cluster.name = sprintf("Cluster_%s_Res4",prefix))

cluster.name = sprintf("Cluster_%s_Res0.25",prefix)
so[[cluster.name]][,1] = RenameCluster(cluster.vector = so[[cluster.name]][,1], cluster.prefix = "cl")
cluster.name = sprintf("Cluster_%s_Res0.5",prefix)
so[[cluster.name]][,1] = RenameCluster(cluster.vector = so[[cluster.name]][,1], cluster.prefix = "cl")
cluster.name = sprintf("Cluster_%s_Res0.75",prefix)
so[[cluster.name]][,1] = RenameCluster(cluster.vector = so[[cluster.name]][,1], cluster.prefix = "cl")
cluster.name = sprintf("Cluster_%s_Res1",prefix)
so[[cluster.name]][,1] = RenameCluster(cluster.vector = so[[cluster.name]][,1], cluster.prefix = "cl")
cluster.name = sprintf("Cluster_%s_Res1.5",prefix)
so[[cluster.name]][,1] = RenameCluster(cluster.vector = so[[cluster.name]][,1], cluster.prefix = "cl")
cluster.name = sprintf("Cluster_%s_Res2",prefix)
so[[cluster.name]][,1] = RenameCluster(cluster.vector = so[[cluster.name]][,1], cluster.prefix = "cl")
cluster.name = sprintf("Cluster_%s_Res3",prefix)
so[[cluster.name]][,1] = RenameCluster(cluster.vector = so[[cluster.name]][,1], cluster.prefix = "cl")
cluster.name = sprintf("Cluster_%s_Res4",prefix)
so[[cluster.name]][,1] = RenameCluster(cluster.vector = so[[cluster.name]][,1], cluster.prefix = "cl")

```

UMAP/MDE with clusters
```{r}
# pdf(sprintf("%s/UMAP_Clusters.pdf", prefix), 20, 20, useDingbats = F)
# for (i in c(0.25,0.5,0.75,1,1.5,2,3,4)) {
#     MyPlots(seurat_object = so, 
#         dim1 = so[[reduc]]@cell.embeddings[,1], 
#         dim2 = so[[reduc]]@cell.embeddings[,2], 
#         color_by = sprintf("Cluster_%s_Res%s",prefix,i), 
#         split_by1 = "organ_simplified",
#         split_by2 =  NULL, 
#         genes = NULL, 
#         cluster_key = NULL, mypal = mypal) #glasbey()
# }
# dev.off()

pdf(sprintf("%s/UMAP_Clusters.pdf", prefix), 20, 20, useDingbats = F)
# pdf(sprintf("%s/MDE_Clusters.pdf", prefix), 20, 20, useDingbats = F)
for (i in c(0.25,0.5,0.75,1,1.5,2,3,4)) {
    print(i)
    print(DimPlot(so, reduction = reduc, group.by = sprintf("Cluster_%s_Res%s",prefix,i), raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal))
    print(DimPlot(so, reduction = reduc, group.by = sprintf("Cluster_%s_Res%s",prefix,i), split.by = "organ_simplified", ncol = trunc(sqrt(length(unique(so$organ_simplified)))), raster = T, raster.dpi = c(512,512), label = F, pt.size = 2) + scale_color_manual(values = mypal))
}
dev.off()

```

UMAP/MDE with genes
```{r}
# pdf(sprintf("%s/UMAP_genes.pdf", prefix), 20,20, useDingbats = F)
pdf(sprintf("%s/MDE_genes.pdf", prefix), 20,20, useDingbats = F)
DefaultAssay(so) = "RNA"
FeaturePlot(so, slot = "data", reduction = reduc, features = c("Cd4", "Cd8a", "Cd8b1","Foxp3", "Ccr7","Rorc", "Tbx21","Il1rl1", "Nr4a1", "Mki67", "Il10", "Il2ra", "Ctla4", "Ly6c1"), ncol = 4, order = T, raster = F, raster.dpi = c(512,512))
dev.off()

# pdf(sprintf("%s/UMAP_proteins.pdf", prefix), 20,20, useDingbats = F)
pdf(sprintf("%s/MDE_proteins.pdf", prefix), 20,20, useDingbats = F)
DefaultAssay(so) = "ADT"
FeaturePlot(so[,so$cite_seq == T], reduction = reduc,slot = "data", features = c("CD3","CD4","CD8A","CD8B","TCRB","IL2RA.CD25","CD62L","CD45RB","KLRG1","NEUROPILIN1.CD304", "SCA1", "FR4","CD49B", "CD69"), ncol = 4, order = T, raster = T, raster.dpi = c(512,512), pt.size = 2)
dev.off()
```

res 0.5 and 0.75 makes most sense biologically. Also the most robusts. 0.5 seems a bit more robust than 0.75 regarding skin mostly.
splitting Cl2 at 0.5 into 0.75_cl2 and 0.75_cl6
--> annotation_level2_treg.csv
```{r}
library(cluster)
cluster.name = sprintf("Cluster_%s_Res0.5",prefix)
cluster.names = sprintf("Cluster_%s_Res%s",prefix,c(0.25, 0.5, 0.75, 1,2,3,4))
#cluster.name = sprintf("Cluster_%s_Res0.75",prefix)

#1 in all the data
reduc.dist = dist(so[[prefix]]@cell.embeddings)

pdf(sprintf("%s/Cluster_totalvi_20241014_Treg_rmIGTsample_silhouettes.pdf", prefix), 10, 10, useDingbats = F)
for (cluster.name in cluster.names) {
    print(cluster.name)
    cl = as.integer(so[[cluster.name]][,1])
    sil = silhouette(cl, dist = reduc.dist)
    clust.col = mypal
    sil.cols = clust.col[ifelse(sil[,3] > 0, sil[,1], sil[,2])]
    sil.cols = sil.cols[order(-sil[,1], sil[,3])]
    plot(sil,
         main = sprintf("Silhouette in %s", cluster.name),
         border = sil.cols,
         col = sil.cols,
         do.col.sort = FALSE)
}
dev.off()

so$Treg_clusters = as.character(so$Cluster_totalvi_20241014_Treg_rmIGTsample_Res0.5)
so$Treg_clusters[so$Cluster_totalvi_20241014_Treg_rmIGTsample_Res0.5 == "cl2" & so$Cluster_totalvi_20241014_Treg_rmIGTsample_Res0.75 == "cl6"] = "cl8"
so$Treg_clusters = sprintf("treg_%s",so$Treg_clusters)
so$Treg_clusters = factor(so$Treg_clusters)

# pdf(sprintf("%s/UMAP_Clusters_Final.pdf", prefix), 20, 20, useDingbats = F)
pdf(sprintf("%s/MDE_Clusters_Final.pdf", prefix), 20, 20, useDingbats = F)
DimPlot(so, reduction = reduc, group.by = "Treg_clusters", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal)
DimPlot(so, reduction = reduc, group.by = "Treg_clusters", split.by = "organ_simplified", ncol = trunc(sqrt(length(unique(so$organ_simplified)))), raster = T, raster.dpi = c(512,512), label = F, pt.size = 2) + scale_color_manual(values = mypal)
dev.off()

cluster.name =  "Treg_clusters"
pdf(sprintf("%s/%s_silhouette.pdf", prefix, cluster.name), 10, 10, useDingbats = F)
cl = as.integer(so[[cluster.name]][,1])
sil = silhouette(cl, dist = reduc.dist)
clust.col = mypal
sil.cols = clust.col[ifelse(sil[,3] > 0, sil[,1], sil[,2])]
sil.cols = sil.cols[order(-sil[,1], sil[,3])]
plot(sil,
     main = sprintf("Silhouette in %s", cluster.name),
     border = sil.cols,
     col = sil.cols,
     do.col.sort = FALSE)
dev.off()

#write.table(data.frame(colnames(so), so$Treg_clusters), file = "../annotation_level2_treg.csv", sep = ",", quote = F, row.names = F, col.names = F)

```

Robustness of clusters with silhouette analysis for cluster.name at all resolution + deeper analysis for resolution 0.75 and 0.5 (the optimum)
in all dataset
in each organ
in each IGT
```{r}
SilhouetteSummaryPlots = function(sil_list = sil_list, column_name = column_name, categories = categories, cluster.name = cluster.name) {
    message("plot silhouettes")
    for (i in categories) {
        print(i)
        sil = sil_list[[i]]
        if (length(sil) >1) {
            # prepare colours:
            clust.col <- mypal#scater:::.get_palette("tableau10medium") # hidden scater colours
            sil.cols <- clust.col[ifelse(sil[,3] > 0, sil[,1], sil[,2])]
            sil.cols <- sil.cols[order(-sil[,1], sil[,3])]
            plot(sil,
                 main = sprintf("Silhouette in %s in %s:%s", cluster.name, i, column_name),
                 border = sil.cols,
                 col = sil.cols,
                 do.col.sort = FALSE)
        }
        
    }
    
    message("Plot Boxplot")
    sil_summary_list = list()
    for (i in categories) {
        sil = sil_list[[i]]
        if (length(sil) >1) {
            sil_summary_list[[i]] = as.data.frame(sil) %>% 
                group_by(cluster) %>% 
                summarise(mean_sil_width = mean(sil_width),
                          total = n(),  # Count total observations in each cluster
                          positive_count = sum(sil_width > 0),  # Count negative silhouette widths
                          percentage_positive = (positive_count / total) * 100)     %>%  # Calculate percentage) 
                as.data.frame()
            sil_summary_list[[i]][,column_name] = i
        }
    }
    sil_summary = Reduce(rbind,sil_summary_list)
    
    #sil_summary[sample(size = 20, x = 1:nrow(sil_summary),replace = F),]
    
    averaged_data <- sil_summary %>%
        group_by(cluster, column_name = sil_summary[,column_name]) %>%
        summarise(ncells = mean(total), mean_of_means = mean(mean_sil_width, na.rm = TRUE), mean_of_percentage_positive = mean(percentage_positive, na.rm = TRUE), .groups = 'drop') %>% data.frame()
    
    p = ggplot(data = averaged_data, aes(x = as.factor(cluster), y = mean_of_means)) +
        geom_boxplot(outlier.shape = NA, alpha = 0.7, width = 0.7, color = "black") +  # Draw boxplots
        geom_point(aes(color = column_name, size = ncells > 10), position = position_jitter(width = 0.2, height = 0), alpha = 0.6) +  # Add jittered points
        scale_color_manual(values=mypal) +  # Color scheme for points
        labs(title = "Boxplot of Average Mean Silhouette Width per Cluster",
             x = "Cluster",
             y = "Average of Mean Silhouette Width",
             color = column_name) +
        theme_minimal() +
        theme(axis.text.x = element_text(angle = 45, hjust = 1),  # Improve x-axis label readability
              legend.position = "right")  # Adjust the legend position
    print(p)
    
    message("Plot Individual")
    p = ggplot(data = averaged_data, aes(x = as.factor(cluster), y = mean_of_means,group = column_name, color = column_name)) +
        geom_line() +  # Add lines to connect points
        #geom_point(size = 3, alpha = 0.8) +  # Add points
        geom_point(alpha = 0.8, aes(size = ncells >10)) +
        scale_color_manual(values=mypal) +  # Use a distinguishable color palette
        labs(title = "Line Plot of Average Mean Silhouette Width per Cluster",
             x = "Cluster",
             y = "Average of Mean Silhouette Width",
             color = column_name) +
        theme_minimal() +
        theme(axis.text.x = element_text(angle = 45, hjust = 1),
              legend.position = "right") 
        
    print(p)
    print(p + facet_wrap(~column_name))
    
     p = ggplot(data = averaged_data, aes(x = as.factor(cluster), y = mean_of_percentage_positive)) +
        geom_boxplot(outlier.shape = NA, alpha = 0.7, width = 0.7, color = "black") +  # Draw boxplots
        geom_point(aes(color = column_name, size = ncells >10), position = position_jitter(width = 0.2, height = 0), alpha = 0.6) +  # Add jittered points
        scale_color_manual(values=mypal) +  # Color scheme for points
        labs(title = "Boxplot of Average Percentage of Cells with Positive Width per Cluster",
             x = "Cluster",
             y = "Average of Percentage of Cells with Positive Width",
             color = column_name) +
        theme_minimal() +
        theme(axis.text.x = element_text(angle = 45, hjust = 1),  # Improve x-axis label readability
              legend.position = "right")  # Adjust the legend position
    print(p)
    
    message("Plot Individual")
    p = ggplot(data = averaged_data, aes(x = as.factor(cluster), y = mean_of_percentage_positive, group = column_name, color = column_name)) +
        geom_line() +  # Add lines to connect points
        #geom_point(size = 3, alpha = 0.8) +  # Add points
        geom_point(alpha = 0.8, aes(size = ncells >10)) +
        scale_color_manual(values=mypal) +  # Use a distinguishable color palette
        labs(title = "Line Plot of Average Percentage of Cells with Positive per Cluster",
             x = "Cluster",
             y = "Average of Percentage of Cells with Positive Width",
             color = column_name) +
        theme_minimal() +
        theme(axis.text.x = element_text(angle = 45, hjust = 1),
              legend.position = "right") 
        
    print(p)
    print(p + facet_wrap(~column_name))
}

#2 in each organ_simplified: how much does it fit each organ?
so2 = so[,!is.na(so$organ_simplified)]

sil_list = list()
column_name = "organ_simplified"
categories = unique(so2[[column_name]][,1]) %>% na.omit() %>% as.character()
cluster.name =  "Treg_clusters"

sil_list = list()
for (i in categories) {
    print(i)
    reduc.dist = dist(so2[[prefix]]@cell.embeddings[so2[[column_name]] == i,])
    cl = as.integer(so2[[cluster.name]][so2[[column_name]] == i,1])
    #cl = as.integer(so2[[cluster.name]][so2[[column_name]] == i,1])
    sil_list[[i]] = silhouette(cl, dist = reduc.dist)
}

pdf(sprintf("%s/%s_silhouettes_by%s.pdf", prefix, cluster.name, column_name), 10, 10, useDingbats = F)
SilhouetteSummaryPlots(sil_list = sil_list, column_name = column_name, categories = categories, cluster.name = cluster.name)
dev.off()

#3. By IGT
sil_list = list()
column_name = "IGT"
categories = levels(so[[column_name]][,1])
cluster.name =  "Treg_clusters"

for (i in categories) {
    print(i)
    reduc.dist = dist(so[[prefix]]@cell.embeddings[so[[column_name]] == i,])
    cl = as.integer(so[[cluster.name]][so[[column_name]] == i,1])
    sil_list[[i]] = silhouette(cl, dist = reduc.dist)
}

pdf(sprintf("%s/%s_silhouettes_by%s.pdf", prefix, cluster.name, column_name), 10, 10, useDingbats = F)
SilhouetteSummaryPlots(sil_list = sil_list, column_name = column_name, categories = categories, cluster.name = cluster.name)
dev.off()

```

```{r}
DefaultAssay(so) = "ADT"
so[['RNA']]$data = NULL
DefaultAssay(so) = "RNA"
so[['ADT']]$data = NULL

# saveRDS(so, file = "/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/Treg/igt1_96_treg_20241017.Rds")

```

**Compare mde_totalvi_20241014_Treg_rmIGTsample and mde_totalvi_20241008_rmIGTsample_Treg (MDE from totalvi_20241008_rmIGTsample on all cells, then subset Tregs to calculate MDE)**
```{r}
so = readRDS("igt1_96_Treg_20241029_so.Rds")

pdf(sprintf("%s/MDE_Clusters_ComparingMDEs.pdf", "."), 20, 20, useDingbats = F)
DimPlot(so, reduction = "mde_totalvi_20241014_Treg_rmIGTsample", group.by = "annotation_level2_Treg", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal)
DimPlot(so, reduction = "mde_totalvi_20241008_rmIGTsample_inTreg", group.by = "annotation_level2_Treg", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal)
# print(DimPlot(so, reduction = "mde_totalvi_20241008_rmIGTsample", group.by = annotation_level2_Treg, split.by = "organ_simplified", ncol = trunc(sqrt(length(unique(so$organ_simplified)))), raster = T, raster.dpi = c(512,512), label = F, pt.size = 2) + scale_color_manual(values = mypal))
dev.off()

```

**Update Seurat object 20241216**
same clusters, adding level1, level2_parent1 etc

```{r}
AddLatentData = function(so, latent_file, prefix, calculate_umap = F) {
    totalvi = read.csv(file = latent_file, header = T, sep = ",", row.names = 1)
    # pattern = "(IGT\\S*)_(\\d+)"
    # rownames(totalvi) = gsub(pattern, "\\1", rownames(totalvi))
    colnames(totalvi) = as.character(seq(1:ncol(totalvi)))
    #all(rownames(totalvi) == colnames(so))
    totalvi = as.matrix(totalvi)
    pattern = "^IGT\\d{1,2}\\.[A-Z]{16}$"
    matches = grepl(pattern, rownames(totalvi))
    if (!all(matches)) {
        print("updating cell names in csv file, assuming same order as seurat object" )
        #rownames(totalvi) = gsub(pattern = "(\\S*)_(\\d+)", replacement = "\\1", rownames(totalvi))
        rownames(totalvi) = colnames(so)
    }
    totalvi = totalvi[colnames(so),]
    if (all(colnames(so) == rownames(totalvi))) {
        print("all cells ARE in the latent_file")
        so[[prefix]] = CreateDimReducObject(embeddings = totalvi, key = prefix, assay = "RNA") #sprintf("totalvi_%s", prefix2)
        if (calculate_umap == T) {
            so = RunUMAP(so, dims = 1:ncol(totalvi), reduction = prefix, n.components = 2, reduction.name = sprintf("umap_%s", prefix))
        }
        return(so)
    } else {
        print("all cells ARE NOT in the latent_file")
        return(so)
    }
}

so_orig = readRDS("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/igt1_96_withtotalvi20241216.Rds")
setwd("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/Treg")
so = so_orig[,so_orig$annotation_level1 == "treg"] # so = so_orig[,so_orig$annotation_level0 == "CD4"]

table_data = table(so$annotation_level2, so$annotation_level1)
# pdf(sprintf("%s/ClustersOct24_Newclusters.pdf", "."), width = 10, height = 10)
# 
# # Convert the table to a data frame for better readability in the PDF
# table_df <- as.data.frame(table_data)
# 
# # Create the table in the PDF
# grid.table(table_df)
# 
# # Close the PDF device
# dev.off()

```

Add general totalvi
```{r}
dir_with_file = "/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/totalvi_20241006"
so = AddLatentData(so, latent_file = sprintf("%s/latent.csv",dir_with_file), prefix = "totalvi_20241006", calculate_umap = F)
so = AddLatentData(so, latent_file = sprintf("%s/mde2.csv",dir_with_file), prefix = "mde2_totalvi_20241006", calculate_umap = F)

```

Add Treg specific MDE
```{r}
dir_with_file = "/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/Treg/totalvi_20241014_Treg_rmIGTsample"
so = AddLatentData(so, latent_file = sprintf("%s/latent.csv",dir_with_file), prefix = "totalvi_20241014_Treg_rmIGTsample", calculate_umap = F)
so = AddLatentData(so, latent_file = sprintf("%s/mde2.csv",dir_with_file), prefix = "mde_totalvi_20241014_Treg_rmIGTsample", calculate_umap = F)

```

Plots
```{r}

pdf(sprintf("%s/MDE_AnnotationLevel20241216.pdf", "."), 10, 10, useDingbats = F)
MyDimPlotHighlight(so_orig, umap_to_plot = "mde2_totalvi_20241006", cells_to_highlight = names(which(so_orig$annotation_level1 == "treg")), highlight_column_name ="annotation_level2", highlight_size = 0.5, highlight_alpha = 0.5, pixels = c(512, 512), mycols = mypal, print_plot1 = T, print_plot2 = T, labelclusters = T )
DimPlot(so, reduction = "mde_totalvi_20241014_Treg_rmIGTsample", group.by = "annotation_level2", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal)
DimPlot(so, reduction = "mde_totalvi_20241014_Treg_rmIGTsample", group.by = "annotation_level2", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal) + theme_void() + NoLegend() 
DimPlot(so, reduction = "mde_totalvi_20241014_Treg_rmIGTsample", group.by = "annotation_level2_parent1", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal)
DimPlot(so, reduction = "mde_totalvi_20241014_Treg_rmIGTsample", group.by = "annotation_level2_parent1", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal) + theme_void() + NoLegend() 
dev.off()

saveRDS(so, file = "igt1_96_Treg_20241216.Rds")

```

remove unwanted cell embeddings
totalvi_20241006 - mde2_totalvi_20241006
totalvi_20241113_CD4_rmIGTsample - mde_incremental
```{r}
so[["umap_totalvi_20241006"]] = NULL
so[["mde_totalvi_20241006"]] = NULL
so[["umap2_totalvi_20241006"]] = NULL
so[["mdeumap_totalvi_20241006"]] = NULL
so[["umap_totalvi_20241008_rmIGTsample"]] = NULL
so[["mde_totalvi_20241008_rmIGTsample"]] = NULL
so[["mde2_totalvi_20241008_rmIGTsample"]] = NULL
so[["mdeumap_totalvi_20241008_rmIGTsample"]] = NULL
so[["mde_totalvi_20241029_fitleredGenesProt"]] = NULL
so[["umap_totalvi_20241029_fitleredGenesProt"]] = NULL

saveRDS(so, file = "igt1_96_Treg_20241216.Rds")
```

**Annotation 20250109**

```{r}
so_orig = readRDS("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/igt1_96_withtotalvi20250109_level1only.Rds")
mypal_level1 = setNames(mypal, unique(so_orig$annotation_level1))
mypal_level2 = setNames(mypal, unique(so_orig$annotation_level2))
setwd("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/Treg")
so = so_orig[,so_orig$annotation_level1 == "treg"]

dir_with_file = "/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/totalvi_20241008_rmIGTsample"
so = AddLatentData(so, latent_file = sprintf("%s/latent.csv",dir_with_file), prefix = "totalvi_20241008_rmIGTsample", calculate_umap = F)
dir_with_file = "/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/Treg/20250109_Treg"
so = AddLatentData(so, latent_file = sprintf("%s/mde_incremental.csv",dir_with_file), prefix = "mde_incremental", calculate_umap = F)

```

No Transfer labels needed
rename Treg_miniverse, Treg_prolif in level2
FYI, no preT anyways


```{r}
MyTransferLabels = function(cell_embeddings_ref, cell_embeddings_query, labels_ref = as.character(so$annotation_level2_20241216[so$is_ref == "ref"]), n_neighbors = 30) {
    require(RcppAnnoy)
    # cell_embeddings_ref = so[["totalvi_20241008_rmIGTsample"]]@cell.embeddings[so$is_ref == "ref",]
    # cell_embeddings_query = so[["totalvi_20241008_rmIGTsample"]]@cell.embeddings[so$is_ref == "query",]
    # labels_ref = as.character(so$annotation_level2_20241216[so$is_ref == "ref"])
    
    # Use the Annoy algorithm to find nearest neighbors between query and ref datasets
    message(sprintf("Number of nearest neighbors to find betweeen query and ref: %s", n_neighbors))
    
    message("Create Annoy index for ref dataset")
    annoy_index = new(AnnoyAngular, ncol(cell_embeddings_ref)) ##use cosine distance Angular
    for (i in 1:nrow(cell_embeddings_ref)) {
        annoy_index$addItem(i - 1, cell_embeddings_ref[i, ])
    }
    annoy_index$build(10)  # Build the index with 10 trees
    
    message("Find nearest ref neighbors for each cell in query dataset")
    nn_indices <- t(sapply(1:nrow(cell_embeddings_query), function(i) {
        annoy_index$getNNsByVector(cell_embeddings_query[i, ], n_neighbors)
    }))
    
    # nn_indices gives you the indices of nearest neighbors in the ref dataset
    # the rows are cells from query dataset, columns are the 30 nearest cells in the ref dataset
    # dim(nn_indices)
    # head(nn_indices)
    
    message("Transfer labels based on majority vote from nearest neighbors")
    transfer_labels <- apply(nn_indices, 1, function(neighbors) {
        # Get labels for the nearest neighbors
        neighbor_labels <- labels_ref[neighbors + 1]  # Add 1 for R's 1-based index
        
        # Return the most common label (majority vote)
        most_common_label <- names(sort(table(neighbor_labels), decreasing = TRUE))[1]
        return(most_common_label)
    })
    
    names(transfer_labels) = rownames(cell_embeddings_query)
    
    return(transfer_labels)
}

so$level2_new = as.character(so$annotation_level2_20241216)
so$level2_new[so$annotation_level2_parent1 == "proliferating"] = "Treg_prolif"
so$level2_new[so$annotation_level2_parent1 == "miniverse"] = "Treg_miniverse"
# so$level2_new[so$annotation_level2_parent1 == "preT"] = "Treg_preT"
table(so$level2_new, so$annotation_level1)
table(so$level2_new, so$annotation_level2_parent1)

# so$is_ref = "other"
# so$is_ref[so$annotation_level1 == "treg" & so$annotation_level1_20241216 == "treg"] = "ref"
# so$is_ref[so$annotation_level1 == "treg" & so$annotation_level1_20241216 != "treg"] = "query"
# so$is_ref[so$annotation_level2_parent1 %in% c("miniverse")] = "other" #"proliferating", "preT"
# cell_embeddings_ref = so[["totalvi_20241008_rmIGTsample"]]@cell.embeddings[so$is_ref == "ref",]
# cell_embeddings_query = so[["totalvi_20241008_rmIGTsample"]]@cell.embeddings[so$is_ref == "query",]
# labels_ref = as.character(so$annotation_level2_20241216[so$is_ref == "ref"])
# transfer_labels = MyTransferLabels(cell_embeddings_ref, cell_embeddings_query, labels_ref, n_neighbors = 30)
# so$level2_new[so$is_ref == "query"] = transfer_labels
# head(transfer_labels)
# table(transfer_labels)


```

Check

```{r}
# prefix_latent = "totalvi_20241008_rmIGTsample"
# n_latent = ncol(so[[prefix_latent]]@cell.embeddings)
# so = FindNeighbors(so, dims = 1:n_latent, reduction = prefix_latent)
# so = FindClusters(so, resolution = 1, cluster.name = sprintf("Cluster_%s_Res1",prefix_latent))

pdf(sprintf("%s/MDE_annotation_20250109_TransferLabels_Reclustering_to_see.pdf", "20250109_Treg"), 20, 20, useDingbats = F)
DimPlot(so, reduction = "mde_incremental", group.by = "annotation_level1", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal_level1) + scale_x_continuous(limits = c(-2,4))
DimPlot(so, reduction = "mde_incremental", group.by = "annotation_level1_20241216", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal) + scale_x_continuous(limits = c(-2,4))
DimPlot(so, reduction = "mde_incremental", group.by = "level2_new", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal_level2) + ggtitle("transfer labels from annotation_level2_20241216") + scale_x_continuous(limits = c(-2,4))
DimPlot(so, reduction = "mde_incremental", group.by = "annotation_level2_20241216", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal_level2) + scale_x_continuous(limits = c(-2,4))
DimPlot(so, reduction = "mde_incremental", group.by = "annotation_level2_parent1", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal) + scale_x_continuous(limits = c(-2,4))
# DimPlot(so, reduction = "mde_incremental", group.by = sprintf("Cluster_%s_Res1",prefix_latent), raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal)
dev.off()

#remove one outlier cell that messes up the MDE plot so[["mde_incremental"]]@cell.embeddings[,1] < -5
so$annotation_level1["IGT70.TTGTAGGTCACGGTTA"] = "remove"
so$annotation_level2["IGT70.TTGTAGGTCACGGTTA"] = "remove"
so$annotation_level2_parent1["IGT70.TTGTAGGTCACGGTTA"] = "other"

pdf(sprintf("%s/MDE_annotation_20250109_level2.pdf", "20250109_Treg"), 20, 20, useDingbats = F)
DimPlot(so, reduction = "mde_incremental", group.by = "annotation_level1", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal_level1) 
DimPlot(so, reduction = "mde_incremental", group.by = "annotation_level1_20241216", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal)
DimPlot(so, reduction = "mde_incremental", group.by = "level2_new", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal) + ggtitle("After transfer labels and reclustering")
DimPlot(so, reduction = "mde_incremental", group.by = "annotation_level2_20241216", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal_level2)
DimPlot(so, reduction = "mde_incremental", group.by = "annotation_level2_parent1", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal)
 
DimPlot(so, reduction = "mde_incremental", group.by = "annotation_level1", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal_level1) + scale_x_continuous(limits = c(-2,4))
DimPlot(so, reduction = "mde_incremental", group.by = "annotation_level1_20241216", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal) + scale_x_continuous(limits = c(-2,4))
DimPlot(so, reduction = "mde_incremental", group.by = "level2_new", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal) + ggtitle("After transfer labels and reclustering") + scale_x_continuous(limits = c(-2,4))
DimPlot(so, reduction = "mde_incremental", group.by = "annotation_level2_20241216", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal_level2) + scale_x_continuous(limits = c(-2,4))
DimPlot(so, reduction = "mde_incremental", group.by = "annotation_level2_parent1", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal) + scale_x_continuous(limits = c(-2,4))
dev.off()

if (all(colnames(so_orig)[so_orig$annotation_level1 == "treg"] == colnames(so))) {
    print("True")
    so_orig$annotation_level2[so_orig$annotation_level1 == "treg"] = so$level2_new[colnames(so_orig)[so_orig$annotation_level1 == "treg"]]
    so_orig$annotation_level2_parent1[so_orig$annotation_level1 == "treg"] = so$annotation_level2_parent1[colnames(so_orig)[so_orig$annotation_level1 == "treg"]]
}
so_orig$annotation_level1["IGT70.TTGTAGGTCACGGTTA"] = "remove"
so_orig$annotation_level2["IGT70.TTGTAGGTCACGGTTA"] = "remove"
so_orig$annotation_level2_parent1["IGT70.TTGTAGGTCACGGTTA"] = "other"

table(so_orig$annotation_level2_parent1[so_orig$annotation_level1 == "treg"], so_orig$annotation_level2[so_orig$annotation_level1 == "treg"])
table(so_orig$annotation_level2[so_orig$annotation_level1 == "treg"], so_orig$annotation_level2_20241216[so_orig$annotation_level1 == "treg"])

table(so_orig$annotation_level2_parent1, so_orig$annotation_level2)

#Export to save awaiting the other annotations
annot = data.frame(cell_id = colnames(so_orig), so_orig@meta.data[,c("annotation_level1", "annotation_level2", "annotation_level2_parent1", "nonconv_tcr_recog")])
colnames(annot) = c("cell_id", "level1", "level2", "level2_parent1", "nonconv_tcr_recog")
table(annot$level2_parent1, annot$level1)
sum(rowSums(apply(annot,2,is.na)))
write.table(x = annot, file = "../annotation_table_20250109_withlevel2.csv", row.names = F, quote = F, sep = ",")

table(so_orig$organ[so_orig$organ == "thymus"], so_orig$annotation_level2_parent1[so_orig$organ == "thymus"])
table(so_orig$organ[so_orig$annotation_level2_parent1 == "preT"], so_orig$annotation_level2_parent1[so_orig$annotation_level2_parent1 == "preT"])
table(so_orig$organ[so_orig$annotation_level2_parent1 == "preT"], so_orig$annotation_level1[so_orig$annotation_level2_parent1 == "preT"])


```


Check Prolif clusters

```{r}
setwd("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/Treg")
so = so_orig[,so_orig$annotation_level1 == "treg"]
so = NormalizeData(so)
so$is_prolif = so$S.Score >0.07 | so$G2M.Score >0.2
dir_with_file = "/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/Treg/20250109_Treg"
so = AddLatentData(so, latent_file = sprintf("%s/mde_incremental.csv",dir_with_file), prefix = "mde_incremental", calculate_umap = F)


pdf(sprintf("%s/Prolif_scores_MDE.pdf","."), 20, 20, useDingbats = F)
FeaturePlot(so, reduction = "mde_incremental", slot = "data", features = "S.Score", order = F, raster = T, raster.dpi = c(1024,1024))
FeaturePlot(so, reduction = "mde_incremental", slot = "data", features = "G2M.Score", order = F, raster = T, raster.dpi = c(1024,1024))
DimPlot(so, reduction = "mde_incremental", group.by = "is_prolif", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = c("grey", "red"))
MyDimPlotHighlight(so, umap_to_plot = "mde_incremental", cells_to_highlight = names(which(so$is_prolif == "TRUE")), highlight_column_name ="is_prolif", highlight_size = 0.5, highlight_alpha = 0.5, pixels = c(512, 512), mycols = c("grey", "red"), label = F, print_plot1 = T, print_plot2 = F )
MyDimPlotHighlight(so, umap_to_plot = "mde_incremental", cells_to_highlight = names(which(so$annotation_level2_parent1 == "proliferating")), highlight_column_name ="annotation_level2_parent1", highlight_size = 0.5, highlight_alpha = 0.5, pixels = c(512, 512), mycols = mypal_level2_parent1, label = F, print_plot1 = T, print_plot2 = F )
dev.off()

```

export data
```{r}
# so_orig = readRDS("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/igt1_96_withtotalvi20250109.Rds")
so_orig = readRDS("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/igt1_96_withtotalvi20250212.Rds")

setwd("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/Treg")
so = so_orig[,so_orig$annotation_level1 == "Treg"]

# dir_with_file = "/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/totalvi_20241008_rmIGTsample"
# so = AddLatentData(so, latent_file = sprintf("%s/latent.csv",dir_with_file), prefix = "totalvi_20241008_rmIGTsample", calculate_umap = F)
dir_with_file = "/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/Treg/20250109_Treg"
so = AddLatentData(so, latent_file = sprintf("%s/mde_incremental.csv",dir_with_file), prefix = "mde_incremental", calculate_umap = F)

so[["umap_totalvi_20241006"]] = NULL
so[["mde_totalvi_20241006"]] = NULL
so[["umap2_totalvi_20241006"]] = NULL
so[["mdeumap_totalvi_20241006"]] = NULL
so[["umap_totalvi_20241008_rmIGTsample"]] = NULL
so[["mde_totalvi_20241008_rmIGTsample"]] = NULL
so[["mde2_totalvi_20241008_rmIGTsample"]] = NULL
so[["mdeumap_totalvi_20241008_rmIGTsample"]] = NULL
so[["mde_totalvi_20241029_fitleredGenesProt"]] = NULL
so[["umap_totalvi_20241029_fitleredGenesProt"]] = NULL


pdf(sprintf("%s/MDE_AnnotationLevel20250109.pdf", "."), 10, 10, useDingbats = F)
MyDimPlotHighlight(so_orig, umap_to_plot = "mde2_totalvi_20241006", cells_to_highlight = names(which(so_orig$annotation_level1 == "Treg")), highlight_column_name ="annotation_level2", highlight_size = 0.5, highlight_alpha = 0.5, pixels = c(512, 512), mycols = mypal_level2, print_plot1 = T, print_plot2 = T, labelclusters = T )
DimPlot(so, reduction = "mde_incremental", group.by = "annotation_level2", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal_level2)
DimPlot(so, reduction = "mde_incremental", group.by = "annotation_level2", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal_level2) + theme_void() + NoLegend() 
DimPlot(so, reduction = "mde_incremental", group.by = "annotation_level2_group", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal_level2_group)
DimPlot(so, reduction = "mde_incremental", group.by = "annotation_level2_group", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal_level2_group) + theme_void() + NoLegend() 
dev.off()

# saveRDS(so, file = "igt1_96_Treg_20250109.Rds")
saveRDS(so, file = "igt1_96_Treg_20250212.Rds")


```







(My silhouette code)
```{r}
#my silhouette code
# Load necessary library
library(stats)

# Simulated data
set.seed(123)
data <- rbind(matrix(rnorm(100, mean = 0), ncol = 2),
              matrix(rnorm(100, mean = 5), ncol = 2))

# K-means clustering
cl <- kmeans(data, 2)


custom_silhouette <- function(data, clusters) {
  # Number of clusters
  num_clusters <- length(unique(clusters$cluster))
  
  # Initialize the silhouette width vector
  s_widths <- numeric(nrow(data))
  
  # Calculate for each point
  for (i in 1:nrow(data)) {
    # Cluster of the current point
    own_cluster <- clusters$cluster[i]
    
    # Points in the same cluster
    own_cluster_points <- data[clusters$cluster == own_cluster, ]
    
    # a(i): Average distance to points in the same cluster
    a_i <- mean(sqrt(colSums((own_cluster_points - data[i, ])^2)))
    
    # Initialize b_i to a large number
    b_i <- Inf
    
    # Calculate b(i): Minimum average distance to any other cluster
    for (k in 1:num_clusters) {
      if (k != own_cluster) {
        other_cluster_points <- data[clusters$cluster == k, ]
        dist_to_other_cluster <- mean(sqrt(colSums((other_cluster_points - data[i, ])^2)))
        b_i <- min(b_i, dist_to_other_cluster)
      }
    }
    
    # Silhouette width calculation
    s_widths[i] <- (b_i - a_i) / max(a_i, b_i)
  }
  
  return(s_widths)
}

sil_widths <- custom_silhouette(data, cl)

# Plot the silhouette widths
plot(sil_widths, type = 'h', main = "Custom Silhouette Widths", xlab = "Index", ylab = "Silhouette Width")


```

