---
title: "cluster_adjustments"
output: html_document
date: "2024-11-14"
editor_options: 
  chunk_output_type: console
---

#########Treg#########
Tregs are fine

#########Miniverse#########
Miniverse as is for now

#########unconvventional#########
Remove the “unconventional” category altogether (it is pretty sketchy), and re-arrange at annotation level1:
gdT: 2, 4, 5, 7, 9, 10, 11, 12, 23, 26 , 28, 32, 33, 35, 42, 43, 46
TCRb+ DN:  3,  15, 18, 29, 31, 34, 39, 41
TCRb+ CD8aa: 1, 16, 20, 40
NonConv: 8, 13, 17, 21, 24, 27, 45 and also include the cells in the attached file tcr.recog.nonconvs.csv (NKT, MAIT and others)
proliferating: 48
 
Remove these cells entirely: 
unconv_cl49                    (no CD3, looks like an ILC)
unconv_cl30                    (suspect, one experiment only)
unconv_cl38                    “”
unconv_cl36                    (tiny cluster, not clear TCR+)
 
Move to CD8ab:
unconv_cl6                       (8ab positive and live in that region)
unconv_cl22                    “”
unconv_cl44                    “”
 
Move to CD4:
unconv_cl19                       (4 positive close to 22)
 
Set aside for independent checking:
unconv_cl14, 25, 37, 47                       These are the B-T ones, need a closer look

#########CD8#########

Move to unconv: CD8_cl3_dp_barcode.csv --> not_CD4_CD8AB_cl51 --> abT_DP_cl51
Move to CD4: 8
Remove: 16 (crappy, ribo/rna, unstable)
Merge 12+17

#########CD4#########
remove: 6 (nonT),24 (non spe binding)
move to CD8: 11
move to unconv: 19

#########

**Start interactive session on Midway3**
```{bash}
ssh -Y zemmour@midway3.rcc.uchicago.edu
sinfo -s
screen -S david
sinteractive --exclusive --time=12:00:00 --account=pi-zemmour --partition=jfkfloor2 --qos=jfkfloor2 --nodes=1 
sinteractive --time=12:00:00 --account=pi-zemmour --partition=caslake --nodes=1 --mem=128GB
sinteractive --time=12:00:00 --account=pi-zemmour --partition=gpu --gres=gpu:1 --mem=128GB
sinteractive --time=12:00:00 --account=pi-zemmour --partition=beagle3 --gres=gpu:1 --mem=96GB

cd $LABHOME

module load python/anaconda-2022.05 
module load hdf5/1.12.0 #for BPCells but also others
module load openblas/0.3.13
source activate /project/zemmour/david/envs/scvi_20240315 #for R and old scvi
module load openblas/0.3.13 #load again or error

module load python
source activate /project/zemmour/david/envs/scvi120_20241008 #for scvi 1.2.0
cd /project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96

R
options(max.print=1000)
#options(Seurat.object.assay.version = 'v5')
setwd("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96")


```

**Initialize: load libraries and custom functions**

```{r}
options(max.print=1000)
# setwd("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96")
source("/project/jfkfloor2/zemmourlab/david/immgent/immgent_integration_git/custom_functions_david.R")

libs = c("fgsea","fastTopics", "flashier", "Matrix", "Seurat","BPCells", "ggplot2", "dplyr","tidyr","scattermore", "reshape2", "ggrastr", "RColorBrewer", "pals", "scales", "pheatmap") 
#libs = c("matrixStats", "gplots","ggplot2", "reshape2", "scales", "gridExtra", "dplyr", "RColorBrewer", "grid", "Rtsne", "limma",   "RColorBrewer", "pheatmap", "Seurat",  "ggrastr", "ggbeeswarm") #pwr, preprocessCore, "readxl", vegan", "genefilter" "scde" "cellrangerRkit", "Rsamtools", "GenomicRanges", "GenomicAlignments", "VGAM", "WGCNA",rafalib, ‘UsingR’,MAST "ggrastr",
sapply(libs, function(x) suppressMessages(suppressWarnings(library(x, character.only = TRUE, quietly = T, warn.conflicts  = F))))


#Gradients
library(RColorBrewer)
ColorRamp = rev(colorRampPalette(brewer.pal(n = 7,name = "RdYlBu"))(100))
ColorRamp = rev(colorRampPalette(c('red','white','blue'))(20))
ColorRamp = rev(rainbow(10, end = 4/6))
library(viridis)
ColorRamp = rev(viridis(100))
ColorRamp = rev(cividis(100))
#image(1:100, 1, as.matrix(1:100), col = ColorRamp, main = "Color Palette", xaxt = 'n', yaxt = 'n', xlab = "", ylab = "")

#mycol = c("black","blue","red","darkgreen", "orange","purple", "cyan", "greenyellow",    "salmon", "magenta","pink", "tan", "brown") # c("magenta", "red",  "darkgreen", "cyan", "blue", "blue", "blue", "black", "blue", "blue", "blue", "blue", "orange")

#Large Color palette
library("pals")
n = 70
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
mypal1 = unique(unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals))))
mypal1 = mypal1[-4]
mypal = c(glasbey(), polychrome(), mypal1)
names(mypal) = NULL
#pal.bands(mypal, main="Colormap suggestions")

mypal_level1 = setNames(mypal, unique(so$annotation_level1))
mypal_level2 = setNames(mypal, unique(so$annotation_level2))
mypal_level2_parent1 = setNames(rev(mypal), unique(so_orig$annotation_level2_parent1))
# mypal_odhran = setNames(mypal, unique(so$odhran))

#parade = function(n) { return(Seurat::DiscretePalette(n, palette = "parade", shuffle = F)) }
```

**update annotation levels**
```{r}
#so = readRDS("igt1_96_withtotalvi20241028.Rds")

setwd("/Users/david/google_drive_uchicago/ImmgenT/analysis/data_integration/IGT1_96")

annotation_level2 = read.table("annotation_level2.csv", header = T, sep = ",", as.is = T, row.names = 1)
annotation_level2$cell_id = rownames(annotation_level2)
#write.table(x = names(table(annotation_level2$annotation_level2)), file = "annotation_level2_list.csv", row.names = F, quote = F)
#update by hand when possible

newannot = read.table("annotation_level2_list_20241113.csv", header = T, sep = ",", as.is = T)
#table(newannot$level2, newannot$level1)

annotation_table = merge(x = annotation_level2, y = newannot, by.x = "annotation_level2", by.y = "annotation_level2_v1", all.x = T)
apply(annotation_table, 2, function(x) length(which(is.na(x))))
annotation_table = annotation_table[match(annotation_level2$cell_id,annotation_table$cell_id),]
rownames(annotation_table) = annotation_table$cell_id
annotation_table$cell_id = NULL
all(annotation_table$annotation_level2 == annotation_level2$annotation_level2)

#add to nonconv: tcr.recog.nonconvs.csv --> not_CD4CD8_cl51
nonconv1 = read.table("tcr.recog.nonconvs.csv", header = F, sep = ",")[,1]
annotation_table$nonconv_tcr_recog = rownames(annotation_table) %in% nonconv1
#add to not_CD4CD8: CD8_cl3_dp_barcode.csv --> not_CD4CD8_cl51
nonconv2 = read.table("CD8_cl3_dp_barcode.csv", header = F, sep = ",")[,1]
annotation_table$level2[rownames(annotation_table) %in% nonconv2] = "not_CD4CD8_cl51"
annotation_table$level1[rownames(annotation_table) %in% nonconv2] = "not_CD4CD8"
annotation_table$level0[rownames(annotation_table) %in% nonconv2] = "not_CD4CD8"
head(annotation_table)
table(annotation_table$nonconv_tcr_recog, annotation_table$level1)

write.table(x = annotation_table, file = "annotation_table_20241113.csv", row.names = T, quote = F, sep = ",")

```

--> update annotation_table_20241113 in seurat object --> metadata_update.Rmd /project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/igt1_96_withtotalvi20241113.Rds
```{r}
so_orig = readRDS("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/igt1_96_withtotalvi20241113.Rds")
so = so_orig[,!so_orig$annotation_level0 %in% c("remove", "nonT")]
prefix = "totalvi_20241006"
umap_to_plot = "mde2_totalvi_20241006"

pdf(sprintf("%s/MDE_annotationV2_202141114.pdf", prefix), 20, 20, useDingbats = F)
DimPlot(so, reduction = umap_to_plot, group.by = "annotation_level2_oct24", raster = T, raster.dpi = c(1024,1024), label = F) + scale_color_manual(values = mypal)
DimPlot(so, reduction = umap_to_plot, group.by = "annotation_level0", raster = T, raster.dpi = c(1024,1024), label = F) + scale_color_manual(values = mypal)
DimPlot(so, reduction = umap_to_plot, group.by = "annotation_level1", raster = T, raster.dpi = c(1024,1024), label = F) + scale_color_manual(values = mypal)
DimPlot(so, reduction = umap_to_plot, group.by = "annotation_level2", raster = T, raster.dpi = c(1024,1024), label = T) + scale_color_manual(values = mypal)
DimPlot(so, reduction = umap_to_plot, group.by = "annotation_level2", raster = T, raster.dpi = c(1024,1024), label = T) + scale_color_manual(values = mypal) + NoLegend()
DimPlot(so, reduction = umap_to_plot, group.by = "nonconv_tcr_recog", raster = T, raster.dpi = c(1024,1024), label = F) + scale_color_manual(values = mypal)
DimPlot(so, reduction = umap_to_plot, group.by = "iNKT", raster = T, raster.dpi = c(1024,1024), label = F) + scale_color_manual(values = mypal)
DimPlot(so, reduction = umap_to_plot, group.by = "MAIT", raster = T, raster.dpi = c(1024,1024), label = F) + scale_color_manual(values = mypal)
# DimPlot(so, reduction = umap_to_plot, group.by = "annotation_level1", split.by = "organ_simplified", ncol = trunc(sqrt(length(unique(so$organ_simplified)))), raster = T, raster.dpi = c(1024,1024), label = F) + scale_color_manual(values = mypal)
dev.off()
```

add in mudata to calculate new MDE
```{python}
import warnings; warnings.simplefilter('ignore')
import scvi
import os
import sys
import scanpy as sc
import muon as mu
import mudata as mu

import numpy as np
import mplscience
import matplotlib.pyplot as plt
import pickle
import pandas as pd

working_dir='/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96'
path_to_mudata='/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/export_data/igt1_96_20241006.h5mu'

mdata = mu.read(path_to_mudata) 

annotationl1 = pd.read_csv(working_dir+"/annotation_level1.csv", index_col = 0, header = None)
mdata.mod['RNA'].obs['annotation_level1'] = annotationl1
mdata.mod['RNA'].obs['annotation_level1'].isna().sum()

mdata_treg = mdata[mdata.mod['RNA'].obs['annotation_level1'] == 'Treg'] 
mdata_treg.write(working_dir+'/export_data/igt1_96_20241014_Treg.h5mu')

mdata_cd4 = mdata[mdata.mod['RNA'].obs['annotation_level1'] == 'CD4'] 
mdata_cd4.write(working_dir+'/export_data/igt1_96_20241014_CD4.h5mu')

mdata_cd8 = mdata[mdata.mod['RNA'].obs['annotation_level1'] == 'CD8AB'] 
mdata_cd8.write(working_dir+'/export_data/igt1_96_20241014_CD8AB.h5mu')

mdata_unconv = mdata[mdata.mod['RNA'].obs['annotation_level1'] == 'unconventional'] 
mdata_unconv.write(working_dir+'/export_data/igt1_96_20241014_unconventional.h5mu')

mdata_miniverse = mdata[mdata.mod['RNA'].obs['annotation_level1'] == 'miniverse'] 
mdata_miniverse.write(working_dir+'/export_data/igt1_96_20241014_miniverse.h5mu')

mdata_treg.mod['RNA'].obs['annotation_level1'].unique()
mdata_cd4.mod['RNA'].obs['annotation_level1'].unique()
mdata_cd8.mod['RNA'].obs['annotation_level1'].unique()
mdata_unconv.mod['RNA'].obs['annotation_level1'].unique()
mdata_miniverse.mod['RNA'].obs['annotation_level1'].unique()

###Subset CD4:
annotation_cd4_l1 = pd.read_csv(working_dir+"/annotation_CD4_clusters_l1.csv", index_col = 0, header = None)
mdata.mod['RNA'].obs['annotation_cd4_l1'] = annotation_cd4_l1
mdata.mod['RNA'].obs['annotation_cd4_l1'].isna().sum()
mdata_cd4_act = mdata[mdata.mod['RNA'].obs['annotation_cd4_l1'] == 'CD4_activated']
mdata_cd4_act.mod['RNA'].obs['annotation_cd4_l1'].unique()
mdata_cd4_act.write(working_dir+'/export_data/igt1_96_20241014_CD4_activated.h5mu')

mdata_cd4_rest = mdata[mdata.mod['RNA'].obs['annotation_cd4_l1'] == 'CD4_resting'] 
mdata_cd4_rest.mod['RNA'].obs['annotation_cd4_l1'].unique()
mdata_cd4_rest.write(working_dir+'/export_data/igt1_96_20241014_CD4_resting.h5mu')

mdata_cd4_prolif = mdata[mdata.mod['RNA'].obs['annotation_cd4_l1'] == 'CD4_prolif'] 
mdata_cd4_prolif.mod['RNA'].obs['annotation_cd4_l1'].unique()
mdata_cd4_prolif.write(working_dir+'/export_data/igt1_96_20241014_CD4_prolif.h5mu')

```

**Individual MDE by cell type from the overall totalvi_20241008_rmIGTsample + subset mudata in separate files**

```{bash}
sinteractive --time=12:00:00 --account=pi-zemmour --partition=gpu --gres=gpu:1 --mem=128GB

cd $LABHOME

module load python
source activate /project/zemmour/david/envs/scvi120_20241008 #for scvi 1.2.0

#cd /project/jfkfloor2/zemmourlab/david/immgent/analysis/integration/IGT1_56
cd /project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96
```

```{python}
# module load python
# source activate /project/zemmour/david/envs/scvi120_20241008 #for scvi 1.2.0

# working_dir = '/project/jfkfloor2/zemmourlab/david/immgent/analysis/integration/IGT1_56/'
# path_to_mudata = '/project/jfkfloor2/zemmourlab/david/immgent/analysis/integration/IGT1_56/export_data/totalvi_igt1_56_20231030_allgenes_mdata.h5mu'
# prefix = 'totalvi_igt1_56_allgenes_20240526_igtsampleregressedout'
# batchkey = 'IGT'
# categorical_covariate_keys = ['sample_id']

working_dir='/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96'
path_to_mudata='/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/export_data/igt1_96_20241006.h5mu'
#path_to_mudata=/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/export_data/igt1_96_20241014_CD4.h5mu
path_to_annotation=working_dir+'/annotation_table_20241113.csv' #annotation_level1.csv
prefix = 'totalvi_20241008_rmIGTsample' #prefix='totalvi_20241006'
celltype='not_CD4CD8' #treg, CD4,CD8AB, not_CD4CD8
# batchkey='IGT'
# categorical_covariate_keys='IGTHT'
# corrected_counts=False
# denoised_data=False

print("Importing libraries")
import warnings; warnings.simplefilter('ignore')
import scvi
import os
import sys
import scanpy as sc
import muon as mu
import mudata as mu

import numpy as np
import mplscience
import matplotlib.pyplot as plt
import pickle
import pandas as pd
import seaborn as sns
import torch
import pymde #to run MDE
pymde.seed(0)

print("Global configurations")
pd.set_option('display.max_rows', 1000)
pd.set_option('display.max_columns', 1000)
# sc.set_figure_params(figsize=(6, 6), frameon=False)
scvi.settings.seed = 0  # optional: ensures reproducibility
pymde.seed(0)
#sns.set_theme()
if torch.cuda.is_available():
    print("CUDA is available")
    print("Using device:", torch.cuda.get_device_name())
    torch.set_float32_matmul_precision("high")

print("Reading mudata")
os.chdir(working_dir)
mdata = mu.read(path_to_mudata) #totalvi_igt1_56_allgenes_Treg_20240306/adata.h5mu") #mu.read("totalvi_igt1_56_allgenes_Treg_20240306/adata.h5mu
print(mdata)

print("Add latent variable")
latent_df = pd.read_csv(prefix+"/latent.csv", index_col = 0)
TOTALVI_LATENT_KEY = "X_totalVI"
mdata.obsm[TOTALVI_LATENT_KEY] = latent_df
# mdata.obsm[TOTALVI_LATENT_KEY] = mdata.obsm[TOTALVI_LATENT_KEY].values

annotationl1 = pd.read_csv(path_to_annotation, index_col = 0, header = 0)
print(mdata.mod['RNA'].obs.index.equals(annotationl1.index)) 
mdata.mod['RNA'].obs['annotation_level1'] = annotationl1['level0']
mdata.mod['RNA'].obs['annotation_level1'].isna().sum()
mdata.mod['RNA'].obs['annotation_level1'].unique()

mdata_orig = mdata

print(celltype)
mdata = mdata_orig[mdata_orig.mod['RNA'].obs['annotation_level1'] == celltype] 
mdata

mde = pymde.preserve_neighbors(
    mdata.obsm[TOTALVI_LATENT_KEY].values, #mdata.obsm[TOTALVI_LATENT_KEY]
    embedding_dim=2,
    constraint=pymde.Standardized(),
    repulsive_fraction=0.7,
    n_neighbors = 15,
    verbose=True,
    device = 'cuda'
)

import pickle
model_pkl_file = prefix+'/mde_'+prefix+'_in'+celltype+'_model.pkl' #prefix+"/mde_model.pkl"  
with open(model_pkl_file, 'wb') as file:  
    pickle.dump(mde, file)

# with open(model_pkl_file, 'rb') as file:  
#     mde = pickle.load(file)

embedding = mde.embed(verbose=True)
mde_df = pd.DataFrame(embedding.cpu(), index = mdata.mod['RNA'].obs.index) 
mde_df.to_csv(prefix+'/mde_'+prefix+'_in'+celltype+'.csv', index=True)

print("Save mde.csv")
# mdata.obsm["X_mde"] = scvi.model.utils.mde(mdata.obsm[TOTALVI_LATENT_KEY])
# mde_df = pd.DataFrame(mdata.obsm["X_mde"], index = mdata.mod['RNA'].obs.index)
# mde_df.to_csv(prefix+'/mde_'+prefix+'_'+celltype+'.csv', index=True)

print("save subset mudata for CD4, CD8, not_CD4CD8")
mdata = mdata_orig

mdata_cd4 = mdata[mdata.mod['RNA'].obs['annotation_level1'] == 'CD4']
mdata_cd4.mod['RNA'].obs['annotation_level1'].unique()
mdata_cd4.update()
mdata_cd4.write(working_dir+'/export_data/igt1_96_20241113_CD4.h5mu')

mdata_cd8 = mdata[mdata.mod['RNA'].obs['annotation_level1'] == 'CD8AB'] 
mdata_cd8.mod['RNA'].obs['annotation_level1'].unique()
mdata_cd8.update()
mdata_cd8.write(working_dir+'/export_data/igt1_96_20241113_CD8AB.h5mu')

mdata_unconv = mdata[mdata.mod['RNA'].obs['annotation_level1'] == 'not_CD4CD8'] 
mdata_unconv.mod['RNA'].obs['annotation_level1'].unique()
mdata_unconv.update()
mdata_unconv.write(working_dir+'/export_data/igt1_96_20241113_not_CD4CD8.h5mu')


# pairs, distorsions = mde.high_distortion_pairs()
# pairs_df = pd.DataFrame(pairs.cpu(), columns=['cell1', 'cell2'])
# pairs_df['cell1'] = mdata.mod['RNA'].obs.index[pairs_df['cell1']]
# pairs_df['cell2'] = mdata.mod['RNA'].obs.index[pairs_df['cell2']]
# distorsions_df = pd.DataFrame(distorsions.cpu(),columns=['distortion'])
# pairs_df['distortion'] = distorsions_df['distortion']
# #pairs_df.sample(n=1000000, replace=False).to_csv(prefix+"/mde_distorsions_sample.csv")
# pairs_df.to_csv(prefix+"/mde_distorsions.csv")
# #most_distorted_pairs = pairs[:500]
# mde.distortions_cdf()
# fig = plt.gcf() 
# fig.savefig(prefix+"/mde_distorsion_cdf.png")

# umap_mde_aligned = pymde.align(source=embedding.cpu(), target=mdata.obsm['X_umap'])
# umap_mde_aligned_df = pd.DataFrame(umap_mde_aligned, index = mdata.mod['RNA'].obs.index)
# umap_mde_aligned_df.to_csv(prefix+"/umap_mde_aligned.csv", index=True)

```


**New group integration and incremental MDE**

split mdata: see above
sbatch $SCRIPT_DIR/run_totalvi_wrapper_20241113_cd4_rmIGTsample.sh
sbatch $SCRIPT_DIR/run_totalvi_wrapper_20241113_cd8_rmIGTsample.sh
sbatch $SCRIPT_DIR/run_totalvi_wrapper_20241113_not_CD4CD8_rmIGTsample.sh

Anchor MDE from: https://github.com/cvxgrp/pymde/blob/main/examples/updating_an_existing_embedding.ipynb
sbatch $SCRIPT_DIR/run_mde_incremental_wrapper_20241121_notCD4CD8.sh
sbatch $SCRIPT_DIR/run_mde_incremental_wrapper_20241121_CD8AB.sh
sbatch $SCRIPT_DIR/run_mde_incremental_wrapper_20241121_CD4.sh

TEST 30000 CD4 + 30000 all other cells
1.subset the data
```{python}

working_dir='/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/'
path_to_mudata='/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/export_data/igt1_96_20241006.h5mu'
prefix='test_mde'
mde_ref_file='CD4/totalvi_20241014_CD4_rmIGTsample/mde2.csv'
totalvi_integrated_file='totalvi_20241006/latent.csv'

path_to_annotation=working_dir+'/annotation_table_20241113.csv' #annotation_level1.csv

print("Importing libraries")
import warnings; warnings.simplefilter('ignore')
import scvi
import os
import sys
import scanpy as sc
import muon as mu
import mudata as mu

import numpy as np
import mplscience
import matplotlib.pyplot as plt
import pickle
import pandas as pd
import seaborn as sns
import torch
import pymde #to run MDE
pymde.seed(0)

print("Global configurations")
pd.set_option('display.max_rows', 1000)
pd.set_option('display.max_columns', 1000)
# sc.set_figure_params(figsize=(6, 6), frameon=False)
scvi.settings.seed = 0  # optional: ensures reproducibility
pymde.seed(0)
#sns.set_theme()
if torch.cuda.is_available():
    print("CUDA is available")
    print("Using device:", torch.cuda.get_device_name())
    torch.set_float32_matmul_precision("high")

print("Reading mudata")
os.chdir(working_dir)
mdata = mu.read(path_to_mudata) #totalvi_igt1_56_allgenes_Treg_20240306/adata.h5mu") #mu.read("totalvi_igt1_56_allgenes_Treg_20240306/adata.h5mu
print(mdata)

annotationl1 = pd.read_csv(path_to_annotation, index_col = 0, header = 0)
print(mdata.mod['RNA'].obs.index.equals(annotationl1.index)) 
mdata.mod['RNA'].obs['annotation_level1'] = annotationl1['level0']
mdata.mod['RNA'].obs['annotation_level1'].isna().sum()
mdata.mod['RNA'].obs['annotation_level1'].unique()

cd4_indices = np.where(mdata.mod['RNA'].obs['annotation_level1'] == 'CD4')[0]
random_cd4_indices = np.random.choice(cd4_indices, size=30000, replace=False)
cd4_indices = mdata.mod['RNA'].obs.index[random_cd4_indices]
non_cd4_indices = np.where(mdata.mod['RNA'].obs['annotation_level1'] != 'CD4')[0]
random_non_cd4_indices = np.random.choice(non_cd4_indices, size=30000, replace=False)
random_non_cd4_indices = mdata.mod['RNA'].obs.index[random_non_cd4_indices]
combined_indices = np.concatenate((cd4_indices, random_non_cd4_indices))
cd4_indices.size, random_non_cd4_indices.size, combined_indices.size

#subset totalvi_integrated with combined_indices only
totalvi_integrated = pd.read_csv(totalvi_integrated_file, index_col = 0, header = 0)
totalvi_integrated.index.size
totalvi_integrated = totalvi_integrated.loc[totalvi_integrated.index.intersection(combined_indices)] 
totalvi_integrated.index.size

totalvi_integrated.to_csv(prefix+'/latent.csv' , index=True)


mde_ref_embedding = pd.read_csv(mde_ref_file, index_col = 0, header = 0)
mde_ref_embedding.index.size
mde_ref_embedding = mde_ref_embedding.loc[mde_ref_embedding.index.intersection(combined_indices)] #subset cells present in the 
mde_ref_embedding.index.size
mde_ref_embedding.to_csv(prefix+'/mde2.csv' , index=True)
```
2.sbatch $SCRIPT_DIR/run_mde_incremental_wrapper_20241121_test.sh
3.Plot
mde_totalvi_20241008_rmIGTsample_innot_CD4CD8: MDE from totalvi_20241008_rmIGT in not_CD4CD8 cells only
```{r}
AddLatentData = function(so, latent_file, prefix, calculate_umap = F) {
    totalvi = read.csv(file = latent_file, header = T, sep = ",", row.names = 1)
    # pattern = "(IGT\\S*)_(\\d+)"
    # rownames(totalvi) = gsub(pattern, "\\1", rownames(totalvi))
    colnames(totalvi) = as.character(seq(1:ncol(totalvi)))
    #all(rownames(totalvi) == colnames(so))
    totalvi = as.matrix(totalvi)
    pattern = "^IGT\\d{1,2}\\.[A-Z]{16}$"
    matches = grepl(pattern, rownames(totalvi))
    if (!all(matches)) {
        print("updating cell names in csv file, assuming same order as seurat object" )
        #rownames(totalvi) = gsub(pattern = "(\\S*)_(\\d+)", replacement = "\\1", rownames(totalvi))
        rownames(totalvi) = colnames(so)
    }
    totalvi = totalvi[colnames(so),]
    if (all(colnames(so) == rownames(totalvi))) {
        print("all cells ARE in the latent_file")
        so[[prefix]] = CreateDimReducObject(embeddings = totalvi, key = prefix, assay = "RNA") #sprintf("totalvi_%s", prefix2)
        if (calculate_umap == T) {
            so = RunUMAP(so, dims = 1:ncol(totalvi), reduction = prefix, n.components = 2, reduction.name = sprintf("umap_%s", prefix))
        }
        return(so)
    } else {
        print("all cells ARE NOT in the latent_file")
        return(so)
    }
}


so_orig = readRDS("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/igt1_96_withtotalvi20241113.Rds")
setwd("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/test_mde")

dir_with_file = "/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/test_mde"
tmp = read.table(sprintf("%s/mde_incremental.csv",dir_with_file), header = T, sep = ",", row.names = 1)
so = so_orig[,rownames(tmp)]

so = AddLatentData(so, latent_file = sprintf("%s/mde_incremental.csv",dir_with_file), prefix = "mde_incremental", calculate_umap = F)

pdf(sprintf("%s/MDE_clusters2.pdf", "."), 20, 20, useDingbats = F)
DimPlot(so[,so$annotation_level1 == "CD4"], reduction = "mde_incremental", group.by = "annotation_level1", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = "green")
DimPlot(so, reduction = "mde_incremental", group.by = "annotation_level1", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal)
DimPlot(so, reduction = "mde_incremental", group.by = "annotation_level2", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal)
DimPlot(so, reduction = "mde_incremental", group.by = "annotation_level2_oct24", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal)
dev.off()

```

**Split nonconv for new totalvi and update annotation table**
--> annotation_table_20241201.csv
abT_DP: unclear_cl14 --> abT_DP_cl14
not_CD4CD8_cl51 --> abT_DP_cl51
not_CD4CD8_cl50 --> gdT_cl50

```{r}
# setwd("/Users/david/google_drive_uchicago/ImmgenT/analysis/data_integration/IGT1_96/")
library(reshape2)
library(dplyr)
annot = read.table("annotation_table_20241113.csv", header = T, sep = ",")
table(annot$level2)
annot$level2[annot$level2 == "unclear_cl14"] = "abT_DP_cl14"
annot$level1[annot$level2 == "abT_DP_cl14"] = "abT_DP"
annot$level2[annot$level2 == "not_CD4CD8_cl51"]  = "abT_DP_cl51"
annot$level1[annot$level2 == "abT_DP_cl51"]  = "abT_DP"
annot$level2[annot$level2 == "not_CD4CD8_cl50"] = "gdT_cl50"
annot$level1[annot$level2 == "gdT_cl50"] = "gdT"

for (i in unique(annot$level1)) {
    print(i)
    print(annot %>% filter(level1 == i) %>% pull(level2) %>% unique())
}


annot$level0 = NULL
write.table(x = annot, file = "annotation_table_20241201.csv", row.names = F, quote = F, sep = ",")

annot_table = table(annot$level1, annot$level2) %>% as.data.frame() %>% melt() %>% filter(value != 0)
annot_table$variable = NULL
colnames(annot_table) = c("level1", "level2", "n_cells")

write.table(x = annot_table, file = "annotation_table_history_20241201.csv", row.names = F, quote = F, sep = ",")

```

add in mudata to calculate new TOTALVI for each subgroup of unconventional
```{python}
import warnings; warnings.simplefilter('ignore')
import scvi
import os
import sys
import scanpy as sc
import muon as mu
import mudata as mu

import numpy as np
import mplscience
import matplotlib.pyplot as plt
import pickle
import pandas as pd

working_dir='/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96'
path_to_mudata='/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/export_data/igt1_96_20241006.h5mu'

mdata = mu.read(path_to_mudata) 

annotationl1 = pd.read_csv(working_dir+"/annotation_table_20241201.csv", index_col = 0, header = 0)
set(annotationl1.index) == set(mdata.mod['RNA'].obs.index)
mdata.mod['RNA'].obs['level1'] = annotationl1['level1']
mdata.mod['RNA'].obs['level1'].isna().sum()

mdata.mod['RNA'].obs['level1'].unique()

mdata_gdT = mdata[mdata.mod['RNA'].obs['level1'] == 'gdT'] 
mdata_gdT.write(working_dir+'/export_data/igt1_96_20241201_gdT.h5mu')
mdata_gdT

mdata_abTCD8aa = mdata[mdata.mod['RNA'].obs['level1'] == 'abT_CD8aa'] 
mdata_abTCD8aa
mdata_abTCD8aa.write(working_dir+'/export_data/igt1_96_20241201_abT_CD8aa.h5mu')

mdata_nonconv = mdata[mdata.mod['RNA'].obs['level1'] == 'nonconv'] 
mdata_nonconv
mdata_nonconv.write(working_dir+'/export_data/igt1_96_20241201_nonconv.h5mu')

mdata_abT_DN = mdata[mdata.mod['RNA'].obs['level1'] == 'abT_DN'] 
mdata_abT_DN
mdata_abT_DN.write(working_dir+'/export_data/igt1_96_20241201_abT_DN.h5mu')


mdata_gdT.mod['RNA'].obs['level1'].unique()
mdata_abTCD8aa.mod['RNA'].obs['level1'].unique()
mdata_nonconv.mod['RNA'].obs['level1'].unique()
mdata_abT_DN.mod['RNA'].obs['level1'].unique()

```

sbatch $SCRIPT_DIR/run_totalvi_wrapper_20241201_gdT_rmIGTsample.sh
sbatch $SCRIPT_DIR/run_totalvi_wrapper_20241201_abT_CD8aa_rmIGTsample.sh
sbatch $SCRIPT_DIR/run_totalvi_wrapper_20241201_nonconv_rmIGTsample.sh
sbatch $SCRIPT_DIR/run_totalvi_wrapper_20241201_abT_DN_rmIGTsample.sh

--> see clustering_unconventional_igt_1_96.Rmd

**remove outliers in abT_DN and abT_CD8aa and update annotation table**
remove outliers in abT_DN and abT_CD8aa --> annotation_table_20241202.csv
```{r}
# setwd("/Users/david/google_drive_uchicago/ImmgenT/analysis/data_integration/IGT1_96/")
setwd("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96")
library(reshape2)
library(dplyr)
annot = read.table("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/annotation_table_20241201.csv", header = T, sep = ",")
dim(annot)
outliers1 = read.table("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/abT_DN/outliers_abT_DN.csv", header = F, sep = ",")[,1]
outliers2 = read.table("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/abT_CD8aa/outliers_abT_CD8aa.csv", header = F, sep = ",")[,1]
annot$level2[annot$cell_id %in% c(outliers1, outliers2)] = "remove"
annot$level1[annot$cell_id %in% c(outliers1, outliers2)] = "remove"

for (i in unique(annot$level1)) {
    print(i)
    print(annot %>% filter(level1 == i) %>% pull(level2) %>% unique())
}

annot$level0 = NULL
write.table(x = annot, file = "annotation_table_20241202.csv", row.names = F, quote = F, sep = ",")

annot_table = table(annot$level1, annot$level2) %>% as.data.frame() %>% melt() %>% filter(value != 0)
annot_table$variable = NULL
colnames(annot_table) = c("level1", "level2", "n_cells")

write.table(x = annot_table, file = "annotation_table_history_20241202.csv", row.names = F, quote = F, sep = ",")

```

remove in mudata and rerun totalvi
```{python}
import warnings; warnings.simplefilter('ignore')
import scvi
import os
import sys
import scanpy as sc
import muon as mu
import mudata as mu

import numpy as np
import mplscience
import matplotlib.pyplot as plt
import pickle
import pandas as pd

working_dir='/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96'
path_to_mudata='/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/export_data/igt1_96_20241006.h5mu'

mdata = mu.read(path_to_mudata) 

annotationl1 = pd.read_csv(working_dir+"/annotation_table_20241202.csv", index_col = 0, header = 0)
set(annotationl1.index) == set(mdata.mod['RNA'].obs.index)
mdata.mod['RNA'].obs['level1'] = annotationl1['level1']
mdata.mod['RNA'].obs['level1'].isna().sum()

mdata.mod['RNA'].obs['level1'].unique()

mdata_abTCD8aa = mdata[mdata.mod['RNA'].obs['level1'] == 'abT_CD8aa'] 
mdata_abTCD8aa
mdata_abTCD8aa.write(working_dir+'/export_data/igt1_96_20241202_abT_CD8aa.h5mu')


mdata_abT_DN = mdata[mdata.mod['RNA'].obs['level1'] == 'abT_DN'] 
mdata_abT_DN
mdata_abT_DN.write(working_dir+'/export_data/igt1_96_20241202_abT_DN.h5mu')

mdata_abTCD8aa.mod['RNA'].obs['level1'].unique()
mdata_abT_DN.mod['RNA'].obs['level1'].unique()

```

sbatch $SCRIPT_DIR/run_totalvi_wrapper_20241202_abT_CD8aa_rmIGTsample.sh
sbatch $SCRIPT_DIR/run_totalvi_wrapper_20241202_abT_DN_rmIGTsample.sh

**remove more outliers in abT_CD8aa and update annotation table** WASN'T USEFUL -- FORGET
remove outliers in abT_CD8aa --> annotation_table_20241203.csv
```{r}
# setwd("/Users/david/google_drive_uchicago/ImmgenT/analysis/data_integration/IGT1_96/")
setwd("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96")
library(reshape2)
library(dplyr)
annot = read.table("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/annotation_table_20241202.csv", header = T, sep = ",")
dim(annot)
outliers2 = read.table("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/abT_CD8aa/outliers_abT_CD8aa_20241202.csv", header = F, sep = ",")[,1]
annot$level2[annot$cell_id %in% c(outliers2)] = "remove"
annot$level1[annot$cell_id %in% c(outliers2)] = "remove"

for (i in unique(annot$level1)) {
    print(i)
    print(annot %>% filter(level1 == i) %>% pull(level2) %>% unique())
}

annot$level0 = NULL
write.table(x = annot, file = "annotation_table_20241203.csv", row.names = F, quote = F, sep = ",")

annot_table = table(annot$level1, annot$level2) %>% as.data.frame() %>% melt() %>% filter(value != 0)
annot_table$variable = NULL
colnames(annot_table) = c("level1", "level2", "n_cells")

write.table(x = annot_table, file = "annotation_table_history_20241203.csv", row.names = F, quote = F, sep = ",")

```

remove in mudata and rerun totalvi
```{python}
import warnings; warnings.simplefilter('ignore')
import scvi
import os
import sys
import scanpy as sc
import muon as mu
import mudata as mu

import numpy as np
import mplscience
import matplotlib.pyplot as plt
import pickle
import pandas as pd

working_dir='/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96'
path_to_mudata='/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/export_data/igt1_96_20241006.h5mu'

mdata = mu.read(path_to_mudata) 

annotationl1 = pd.read_csv(working_dir+"/annotation_table_20241203.csv", index_col = 0, header = 0)
set(annotationl1.index) == set(mdata.mod['RNA'].obs.index)
mdata.mod['RNA'].obs['level1'] = annotationl1['level1']
mdata.mod['RNA'].obs['level1'].isna().sum()

mdata.mod['RNA'].obs['level1'].unique()

mdata_abTCD8aa = mdata[mdata.mod['RNA'].obs['level1'] == 'abT_CD8aa'] 
mdata_abTCD8aa
mdata_abTCD8aa.write(working_dir+'/export_data/igt1_96_20241203_abT_CD8aa.h5mu')

mdata_abTCD8aa.mod['RNA'].obs['level1'].unique()

```
sbatch $SCRIPT_DIR/run_totalvi_wrapper_20241203_abT_CD8aa_rmIGTsample.sh

**Merge clusters in CD4 and CD8**
CD4: merge  cl21/26/27 as cl1
CD8: merge cl1/3 as cl1
     merge  cl9/10/21 as cl10
     merge  cl22/23 as 22

```{r}
# setwd("/Users/david/google_drive_uchicago/ImmgenT/analysis/data_integration/IGT1_96/")
setwd("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96")
library(reshape2)
library(dplyr)
annot = read.table("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/annotation_table_20241202.csv", header = T, sep = ",")
dim(annot)

annot$level2[annot$level2 == "CD4_cl26"] = "CD4_cl21"
annot$level2[annot$level2 == "CD4_cl27"] = "CD4_cl21"

annot$level2[annot$level2 == "CD8AB_cl3"] = "CD8AB_cl1"

annot$level2[annot$level2 == "CD8AB_cl9"] = "CD8AB_cl10"
annot$level2[annot$level2 == "CD8AB_cl21"] = "CD8AB_cl10"

annot$level2[annot$level2 == "CD8AB_cl23"] = "CD8AB_cl22"

for (i in unique(annot$level1)) {
    print(i)
    print(annot %>% filter(level1 == i) %>% pull(level2) %>% unique())
}

annot$level0 = NULL
annot$annotation_level2 = NULL

#Add resting, activated
annot$level2_parent1 = "not_applicable"
annot$level2_parent1[annot$level2 %in% c("CD4_cl10", "CD4_cl13", "CD4_cl14", "CD4_cl15", "CD4_cl16", "CD4_cl17", "CD4_cl18", "CD4_cl20", "CD4_cl21", "CD4_cl22", "CD4_cl23", "CD4_cl25", "CD4_cl9", "CD8AB_cl10", "CD8AB_cl11", "CD8AB_cl12", "CD8AB_cl13", "CD8AB_cl14", "CD8AB_cl15", "CD8AB_cl18", "CD8AB_cl20", "CD8AB_cl6", "CD8AB_cl7","treg_cl2", "treg_cl3", "treg_cl4", "treg_cl6", "treg_cl8")] = "activated"
annot$level2_parent1[annot$level2 %in% c("CD4_cl1", "CD4_cl12", "CD4_cl2", "CD4_cl3", "CD4_cl4", "CD4_cl5", "CD4_cl7", "CD8AB_cl1", "CD8AB_cl2", "CD8AB_cl22", "CD8AB_cl4", "CD8AB_cl5","treg_cl1", "treg_cl7"
)] = "resting"
annot$level2_parent1[annot$level2 %in% c("CD4_cl8", "CD8AB_cl19", "treg_cl5", "not_CD4CD8_cl48")] = "proliferating"
table(annot$level2_parent1, annot$level2)
table(annot$level2_parent1, annot$level1)

sum(rowSums(apply(annot,2,is.na)))

write.table(x = annot, file = "annotation_table_20241216.csv", row.names = F, quote = F, sep = ",")

annot_summary <- annot %>%
  group_by(level2, level1, level2_parent1) %>% 
  summarize(counts = n(), .groups = "drop") %>%
  arrange(
    sub("^.*_cl", "", level2) %>% as.numeric(), # Extract cluster number and convert to numeric
    level2                                     # Keep cluster prefix order (CD4, treg, etc.)
  ) %>% 
    arrange(level1) %>%
    as.data.frame()


write.table(x = annot_summary, file = "annotation_table_history_20241216.csv", row.names = F, quote = F, sep = ",")

```

**TEST with CB and CD8 TEST Level1 adjustment based on CITEseq discordant CD4**
```{r}
mypal_level1 = setNames(mypal, unique(so$annotation_level1))
mypal_odhran = setNames(mypal, unique(so$odhran))

so_orig = readRDS("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/igt1_96_withtotalvi20241216.Rds")
so = so_orig[,!so_orig$annotation_level1 %in% c("nonT","remove")]
so$annotation_level1 = factor(so$annotation_level1)

discordantCD4 = read.table("DiscordantCD4.csv", header = T, sep = ",", row.names = 1)
discordantCD8ab = read.table("DiscordantCD8ab.csv", header = T, sep = ",", row.names = 1)
discordant = data.frame(odhran = c(discordantCD4$X3, discordantCD8ab$DBlabel))
rownames(discordant) = c(rownames(discordantCD4), rownames(discordantCD8ab))
so$odhran = discordant[colnames(so),"odhran"]
so$odhran[is.na(so$odhran)] = "same"

so = NormalizeData(so,assay = "ADT", normalization.method = "CLR", verbose = T)

pdf("Discordant_CD4CD8.pdf", 10,10, useDingbats = F)
FeatureScatter(so[,so$odhran != "same"], feature1 = "CD4", feature2 = "CD8B", group.by = "odhran", split.by = "annotation_level1", ncol = 2, pt.size = 2, raster = F, plot.cor = F) + scale_color_manual(values = mypal_level1)  + scale_color_manual(values = mypal_odhran) + 
    geom_hline(aes(yintercept = 2.2), linetype="dashed", color = "brown") +
    geom_vline(aes(xintercept = 1.5), linetype="dashed", color = "brown")
FeatureScatter(so[,so$odhran != "same"], feature1 = "CD4", feature2 = "CD8A", group.by = "odhran", split.by = "annotation_level1", ncol = 2, pt.size = 2, raster = F, plot.cor = F) + scale_color_manual(values = mypal_level1)  + scale_color_manual(values = mypal_odhran) + 
    geom_hline(aes(yintercept = 2.2), linetype="dashed", color = "brown") +
    geom_vline(aes(xintercept = 1.5), linetype="dashed", color = "brown")

for (i in c("CD4+", "CD8AA+", "CD8AB+")) {
    print(i)
    p = MyDimPlotHighlight(seurat_object = so, umap_to_plot = "mde2_totalvi_20241006", cells_to_highlight = colnames(so)[so$odhran %in% i], highlight_column_name = "annotation_level1", labelclusters = T,
                           title = i, pixels = c(512, 512), highlight_size = 1, 
                           highlight_alpha = 1, mycols = mypal,print_plot1 = F, 
                           print_plot2 = F)
    print(p)
}

dev.off()


cd8_inCD4 = names(which(so[["ADT"]]$data["CD8B",so$annotation_level1 == "CD4"] > 2 & so[["ADT"]]$data["CD4",so$annotation_level1 == "CD4"] < 1))
length(which(so[["ADT"]]$data["CD8B",so$annotation_level1 == "CD4"] > 2 & so[["ADT"]]$data["CD4",so$annotation_level1 == "CD4"] < 1))

so$cd8_inCD4 = colnames(so) %in% cd8_inCD4

# FeatureScatter(so, feature1 = "CD4", feature2 = "CD8B", split.by = NULL) + scale_color_manual(values = mypal_level1)  + scale_color_manual(values = mypal_odhran) + 
    geom_hline(aes(yintercept = 2.2), linetype="dashed", color = "brown") +
    geom_vline(aes(xintercept = 1.5), linetype="dashed", color = "brown")


# MyDimPlotHighlight(seurat_object = so, umap_to_plot = "mde2_totalvi_20241006", cells_to_highlight = colnames(so)[so$cd8_inCD4], highlight_column_name = "annotation_level1", labelclusters = T,
#                            title = "", pixels = c(512, 512), highlight_size = 1, 
#                            highlight_alpha = 1, mycols = mypal,print_plot1 = F, 
#                            print_plot2 = F)

```

add cd8_inCD4 in mudata to calculate new TOTALVI
```{python}
import warnings; warnings.simplefilter('ignore')
import scvi
import os
import sys
import scanpy as sc
import muon as mu
import mudata as mu

import numpy as np
import mplscience
import matplotlib.pyplot as plt
import pickle
import pandas as pd

working_dir='/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96'
path_to_mudata='/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/export_data/igt1_96_20241006.h5mu'

mdata = mu.read(path_to_mudata) 

annotationl1 = pd.read_csv(working_dir+"/annotation_table_20241216.csv", index_col = 0, header = 0)
set(annotationl1.index) == set(mdata.mod['RNA'].obs.index)
mdata.mod['RNA'].obs['level1'] = annotationl1['level1']
mdata.mod['RNA'].obs['level1'].isna().sum()

discordant = pd.read_csv(working_dir+"/DiscordantCD4_david.csv", header = None)
discordant.columns = ['cell_id']
mdata.mod['RNA'].obs.loc[discordant.cell_id, 'level1'] = "CD8AB"
mdata.mod['RNA'].obs['level1'].unique()

mdata_CD8AB = mdata[mdata.mod['RNA'].obs['level1'] == 'CD8AB'] 
mdata_CD8AB
mdata_CD8AB.write(working_dir+'/export_data/igt1_96_20250103_CD8AB.h5mu')

```

sbatch $SCRIPT_DIR/run_totalvi_wrapper_20250103_cd8_rmIGTsample.sh 
--> check in clustering_CD8AB.Rmd

*TEST with CB and CD8: Level1 adjustment based on CITEseq CD4, CD8B, CD8A, TCRGD, TCRB*

```{r}
so_orig = readRDS("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/igt1_96_withtotalvi20241216.Rds")
# so = so_orig[,!so_orig$annotation_level1 %in% c("nonT","remove")]
so$annotation_level1 = factor(so$annotation_level1)

so = NormalizeData(so,assay = "ADT", normalization.method = "CLR", verbose = T)
```

Observation 1: Some annotation level 1 in other areas, e.g. “annotation level 1 CD4” cells in CD8AB, TCRGD, etc areas errors in clustering? Or interesting overlaps?
```{r}

pdf("MDE_annotation_20241216_level1.pdf", 10,10, useDingbats = F)
for (i in levels(so$annotation_level1)) {
    print(i)
    p = MyDimPlotHighlight(seurat_object = so, umap_to_plot = "mde2_totalvi_20241006", cells_to_highlight = colnames(so)[so$annotation_level1 == i], highlight_column_name = "annotation_level1", labelclusters = F,
                       title = i, pixels = c(512, 512), highlight_size = 0.5, 
                       highlight_alpha = 1, mycols = mypal_level1, print_plot1 = T, 
                       print_plot2 = T)
    print(p)
}
dev.off()



```

Observation 2: Discordances between annotation level 1 and CITEseq (CD4, CD8AB, CD8AA, TCRGD)
```{r}

pdf("FeatureScatter_annotation_20241216_level1.pdf", 20,20, useDingbats = F)
# for (i in levels(so$annotation_level1)) {
FeatureScatter(so, feature1 = "TCRB", feature2 = "TCRGD", group.by = "annotation_level1", split.by = "annotation_level1", ncol = 4, pt.size = 2, raster = T, plot.cor = F) + scale_color_manual(values = mypal_level1)
FeatureScatter(so, feature1 = "CD4", feature2 = "CD8B", group.by = "annotation_level1", split.by = "annotation_level1", ncol = 4, pt.size = 2, raster = T, plot.cor = F) + scale_color_manual(values = mypal_level1)
FeatureScatter(so, feature1 = "CD8B", feature2 = "CD8A", group.by = "annotation_level1", split.by = "annotation_level1", ncol = 4, pt.size = 2, raster = T, plot.cor = F) + scale_color_manual(values = mypal_level1)
# FeatureScatter(so[,so$gdTwithTCRab == T], feature1 = "CD4", feature2 = "CD8B", group.by = "gdT_CITEseq", split.by = NULL, ncol = 2, pt.size = 2, raster = F, plot.cor = F) 
# }
dev.off()

```

Plot iNKT, MAIT, gdT with TCRab
```{r}
iNKT = read.table("nonconv/inktInDec24Freeze.csv", header = T, sep = ",")[,1]
so$iNKT2 = colnames(so) %in% iNKT
MAIT = read.table("nonconv/maitInDec24Freeze.csv", header = T, sep = ",")[,1]
so$MAIT2 = colnames(so) %in% MAIT
gdTwithTCRab = read.table("gdT/gdtWithProductiveABtcr.csv", header = T, sep = ",")[,1]
so$gdTwithTCRab = colnames(so) %in% gdTwithTCRab
gdT_CITEseq = read.table("gating_Leon_20241116/TCRGD_manual_gating_byIGT.csv", header = T, sep = ",")[,1]
so$gdT_CITEseq = colnames(so) %in% gdT_CITEseq

pdf("MDE_NKT_MAIT.pdf", 10,10, useDingbats = F)
p = MyDimPlotHighlight(seurat_object = so, umap_to_plot = "mde2_totalvi_20241006", cells_to_highlight = colnames(so)[so$iNKT2 == T], highlight_column_name = "annotation_level1", labelclusters = T,
                       title = "iNKT", pixels = c(512, 512), highlight_size = 1, 
                       highlight_alpha = 1, mycols = mypal_level1,print_plot1 = T, 
                       print_plot2 = T)
print(p)

p = MyDimPlotHighlight(seurat_object = so, umap_to_plot = "mde2_totalvi_20241006", cells_to_highlight = colnames(so)[so$iNKT2 == T], highlight_column_name = "annotation_level2", labelclusters = T,
                       title = "iNKT", pixels = c(512, 512), highlight_size = 1, 
                       highlight_alpha = 1, mycols = mypal_level2,print_plot1 = T, 
                       print_plot2 = T)
print(p)

p = MyDimPlotHighlight(seurat_object = so, umap_to_plot = "mde2_totalvi_20241006", cells_to_highlight = colnames(so)[so$MAIT2 == T], highlight_column_name = "annotation_level1", labelclusters = T,
                       title = "MAIT", pixels = c(512, 512), highlight_size = 1, 
                       highlight_alpha = 1, mycols = mypal_level1,print_plot1 = T, 
                       print_plot2 = T)
print(p)

p = MyDimPlotHighlight(seurat_object = so, umap_to_plot = "mde2_totalvi_20241006", cells_to_highlight = colnames(so)[so$MAIT2 == T], highlight_column_name = "annotation_level2", labelclusters = T,
                       title = "MAIT", pixels = c(512, 512), highlight_size = 1, 
                       highlight_alpha = 1, mycols = mypal_level2,print_plot1 = T, 
                       print_plot2 = T)
print(p)

p = MyDimPlotHighlight(seurat_object = so, umap_to_plot = "mde2_totalvi_20241006", cells_to_highlight = colnames(so)[so$gdTwithTCRab == T], highlight_column_name = "annotation_level1", labelclusters = T,
                       title = "gdTwithTCRab", pixels = c(512, 512), highlight_size = 1, 
                       highlight_alpha = 1, mycols = mypal_level1,print_plot1 = T, 
                       print_plot2 = T)
print(p)

p = MyDimPlotHighlight(seurat_object = so, umap_to_plot = "mde2_totalvi_20241006", cells_to_highlight = colnames(so)[so$gdTwithTCRab == T], highlight_column_name = "annotation_level2", labelclusters = T,
                       title = "gdTwithTCRab", pixels = c(512, 512), highlight_size = 1, 
                       highlight_alpha = 1, mycols = mypal_level2,print_plot1 = T, 
                       print_plot2 = T)
print(p)

p = MyDimPlotHighlight(seurat_object = so, umap_to_plot = "mde2_totalvi_20241006", cells_to_highlight = colnames(so)[so$annotation_level1 == "gdT"], highlight_column_name = "annotation_level1", labelclusters = T,
                       title = "gdT in annotation level 1", pixels = c(512, 512), highlight_size = 1, 
                       highlight_alpha = 1, mycols = mypal_level2,print_plot1 = T, 
                       print_plot2 = T)
print(p)

p = MyDimPlotHighlight(seurat_object = so, umap_to_plot = "mde2_totalvi_20241006", cells_to_highlight = colnames(so)[so$gdT_CITEseq == T], highlight_column_name = "annotation_level1", labelclusters = T,
                       title = "gdT by CITEseq", pixels = c(512, 512), highlight_size = 1, 
                       highlight_alpha = 1, mycols = mypal_level2,print_plot1 = T, 
                       print_plot2 = T)
print(p)

dev.off()


# MyDimPlotHighlight(seurat_object = so, umap_to_plot = "mde2_totalvi_20241006", cells_to_highlight = colnames(so)[so$iNKT == T], highlight_column_name = "annotation_level1", labelclusters = T,
#                        title = "MAIT", pixels = c(512, 512), highlight_size = 1, 
#                        highlight_alpha = 1, mycols = mypal_level1,print_plot1 = T, 
#                        print_plot2 = T)

FeatureScatter(so[,so$gdTwithTCRab == T], feature1 = "TCRGD", feature2 = "TCRB", group.by = "annoat", split.by = NULL, ncol = 2, pt.size = 2, raster = F, plot.cor = F) + scale_color_manual(values = mypal_level1)
FeatureScatter(so[,so$gdTwithTCRab == T], feature1 = "CD4", feature2 = "CD8B", group.by = "gdT_CITEseq", split.by = NULL, ncol = 2, pt.size = 2, raster = F, plot.cor = F) 

```

Load gated cells and calculate the table inaccuracies by CITEseq
```{r}
tcrgd_gated = read.table("gating_Leon_20241116/TCRGD_manual_gating_byIGT.csv", header = F, sep = ",")[,1]
so$tcrgd_gated = colnames(so) %in% tcrgd_gated

cd4_gated = read.table("gating_Leon_20241116/IGT1-96_TCRB+CD4+_HardFilter.csv", header = F, sep = ",")[,1]
so$cd4_gated = colnames(so) %in% cd4_gated

cd8ab_gated = read.table("gating_Leon_20241116/IGT1-96_TCRB+CD8AA+CD8AB+_HardFilter.csv", header = F, sep = ",")[,1]
so$cd8ab_gated = colnames(so) %in% cd8ab_gated

cd8aa_gated = read.table("gating_Leon_20241116/IGT1-96_TCRB+CD8AA+CD8AB-_HardFilter.csv", header = F, sep = ",")[,1]
so$cd8aa_gated = colnames(so) %in% cd8aa_gated

so$gated_pop = NA
so$gated_pop[so$tcrgd_gated] = "gdT_gated"
so$gated_pop[so$cd4_gated] = "CD4_gated"
so$gated_pop[so$cd8ab_gated] = "CD8AB_gated"
so$gated_pop[so$cd8aa_gated] = "CD8AA_gated"

pdf("gating_Leon_20241116/FeaturePlots_GatedCells.pdf", width = 10, 10, useDingbats = F)
FeatureScatter(so, feature1 = "TCRB", feature2 = "TCRGD", group.by = "gated_pop", split.by = NULL, ncol = 2, pt.size = 2, raster = T, plot.cor = F) + scale_color_manual(values = mypal)
FeatureScatter(so, feature1 = "TCRB", feature2 = "TCRGD", group.by = "gated_pop", split.by = "gated_pop", ncol = 2, pt.size = 2, raster = T, plot.cor = F) + scale_color_manual(values = mypal)
FeatureScatter(so, feature1 = "CD4", feature2 = "CD8B", group.by = "gated_pop", split.by = "gated_pop", ncol = 2, pt.size = 2, raster = T, plot.cor = F) + scale_color_manual(values = mypal)
FeatureScatter(so, feature1 = "CD8B", feature2 = "CD8A", group.by = "gated_pop", split.by = "gated_pop", ncol = 2, pt.size = 2, raster = T, plot.cor = F) + scale_color_manual(values = mypal)
dev.off()

pdf("gating_Leon_20241116/FeaturePlots_GatedCells_level120241216.pdf", width = 20, 20, useDingbats = F)
FeatureScatter(so, feature1 = "TCRB", feature2 = "TCRGD", group.by = "gated_pop", split.by = "annotation_level1", ncol = 4, pt.size = 2, raster = T, plot.cor = F) + scale_color_manual(values = mypal[c(1:3,5)])
FeatureScatter(so, feature1 = "CD4", feature2 = "CD8B", group.by = "gated_pop", split.by = "annotation_level1", ncol = 4, pt.size = 2, raster = T, plot.cor = F) + scale_color_manual(values = mypal[c(1:3,5)])
FeatureScatter(so, feature1 = "CD8B", feature2 = "CD8A", group.by = "gated_pop", split.by = "annotation_level1", ncol = 4, pt.size = 2, raster = T, plot.cor = F) + scale_color_manual(values = mypal[c(1:3,5)])
dev.off()

df = data.frame(so@meta.data[,c("annotation_level1", "gated_pop")])
df_summary_level1 = df %>% group_by(annotation_level1, gated_pop) %>% summarize(ncells = n(), .groups = "drop") %>% group_by(annotation_level1) %>% mutate(total_cells = sum(ncells), prop = ncells/total_cells) %>% as.data.frame()
# write.table(df_summary_level1, file = "gating_Leon_20241116/annotation_20241216_level1_vs_gating.csv", quote = F, sep = ",", row.names = F, col.names = T)

pdf("gating_Leon_20241116/annotation_20241216_level1_vs_gating_barplots.pdf", width = 8, 6, useDingbats = F)
annotation_levels = unique(df_summary_level1$annotation_level1)
for (level in annotation_levels) {
    print(level)
    # Subset the data for the current annotation_level1
    data_subset <- df_summary_level1 %>%
        filter(annotation_level1 == level)
    
    # Create the barplot
    p <- ggplot(data_subset, aes(x = gated_pop, y = prop, fill = gated_pop)) +
        geom_bar(stat = "identity") +
        labs(
            title = paste("Proportion of Gated Populations in", level),
            x = "Gated Population",
            y = "Proportion"
        ) +
        scale_y_continuous(limits = c(0, 1)) +
        scale_color_manual(values = mypal_level1) +
    theme_minimal() +
        theme(
            axis.text.x = element_text(angle = 45, hjust = 1),
            legend.position = "none"
        )
    
    # Print the plot
    print(p)
}
dev.off()

df = data.frame(so@meta.data[,c("annotation_level2", "gated_pop")])
df_summary_level2 = df %>% group_by(annotation_level2, gated_pop) %>% summarize(ncells = n(), .groups = "drop") %>% group_by(annotation_level2) %>% mutate(total_cells = sum(ncells), prop = ncells/total_cells) %>% as.data.frame()
# write.table(df_summary_level2, file = "gating_Leon_20241116/annotation_20241216_level2_vs_gating.csv", quote = F, sep = ",", row.names = F, col.names = T)

pdf("gating_Leon_20241116/annotation_20241216_level2_vs_gating_barplots.pdf", width = 6, 6, useDingbats = F)
annotation_levels = unique(df_summary_level2$annotation_level2)
for (level in annotation_levels) {
    print(level)
    # Subset the data for the current annotation_level1
    data_subset <- df_summary_level2 %>%
        filter(annotation_level2 == level)
    
    # Create the barplot
    p <- ggplot(data_subset, aes(x = gated_pop, y = prop, fill = gated_pop)) +
        geom_bar(stat = "identity") +
        labs(
            title = paste("Proportion of Gated Populations in", level),
            x = "Gated Population",
            y = "Proportion"
        ) +
        scale_y_continuous(limits = c(0, 1)) +
        scale_color_manual(values = mypal_level1) +
    theme_minimal() +
        theme(
            axis.text.x = element_text(angle = 45, hjust = 1),
            legend.position = "none"
        )
    
    # Print the plot
    print(p)
}
dev.off()


```


**20250109: Level1 adjustment based on CITEseq CD4, CD8B, CD8A, TCRGD, TCRB using Leon's gating + additional manual cutoffs + also splitting prolif, miniverse**
-->annotation_table_20250109_withlevel220241216.csv, igt1_96_withtotalvi20250109.Rds

level1: gdT, CD4, CD8AB, abT_CD8aa, nonconv, abT_DN, abT_DP, Treg
level2_parent: resting, activated, prolif, miniverse, preT

To remember just in case:
- iNKT and MAIT list by TCR can be overrun by gating since they fall everywhere
- same with using cluster annotation (e.g. preT or miniverse here), can overrun
+++ not sure how well thymus is annotated properly here, because CD4 staining didn't work so well by CITEseq. would be to focus more on it. for now, remove thymus from analysis, or stay as thymocyte.

```{r}
so_orig = readRDS("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/igt1_96_withtotalvi20241216.Rds")
mypal_level1 = setNames(mypal, unique(so_orig$annotation_level1))
mypal_level2 = setNames(mypal, unique(so_orig$annotation_level2))

so = so_orig
so = NormalizeData(so,assay = "ADT", normalization.method = "CLR", verbose = T)
so$level1 = so$annotation_level1

tcrgd_gated = read.table("gating_Leon_20241116/TCRGD_manual_gating_byIGT.csv", header = F, sep = ",")[,1]
so$tcrgd_gated = colnames(so) %in% tcrgd_gated

cd4_gated = read.table("gating_Leon_20241116/IGT1-96_TCRB+CD4+_HardFilter.csv", header = F, sep = ",")[,1]
so$cd4_gated = colnames(so) %in% cd4_gated

cd8ab_gated = read.table("gating_Leon_20241116/IGT1-96_TCRB+CD8AA+CD8AB+_HardFilter.csv", header = F, sep = ",")[,1]
so$cd8ab_gated = colnames(so) %in% cd8ab_gated

cd8aa_gated = read.table("gating_Leon_20241116/IGT1-96_TCRB+CD8AA+CD8AB-_HardFilter.csv", header = F, sep = ",")[,1]
so$cd8aa_gated = colnames(so) %in% cd8aa_gated

so$gated_pop = NA
so$gated_pop[so$tcrgd_gated] = "gdT_gated"
so$gated_pop[so$cd4_gated] = "CD4_gated"
so$gated_pop[so$cd8ab_gated] = "CD8AB_gated"
so$gated_pop[so$cd8aa_gated] = "CD8AA_gated"

iNKT = read.table("nonconv/inktInDec24Freeze.csv", header = T, sep = ",")[,1]
# so$iNKT2 = colnames(so) %in% iNKT
MAIT = read.table("nonconv/maitInDec24Freeze.csv", header = T, sep = ",")[,1]
# so$MAIT2 = colnames(so) %in% MAIT

so$gated_pop[ colnames(so) %in% iNKT] = "iNKT"
so$gated_pop[ colnames(so) %in% MAIT] = "MAIT"

```

LEVEL2_PARENT
thymocyte --> preT
prolif, c("CD8AB_cl19", "CD4_cl8", "treg_cl5") --> prolif
miniverse
resting
activated
```{r}
so$level2_parent = "other"
so$level2_parent[so$annotation_level1 == "thymocyte"] = "preT"

so$level2_parent[so$annotation_level1 == "miniverse"] = "miniverse"

so$level2_parent[so$annotation_level1 == "prolif"] = "prolif"
so$level2_parent[so$annotation_level2 %in% c("CD8AB_cl19", "CD4_cl8", "treg_cl5", "miniverse_cl13")] = "prolif"

so$level2_parent[so$annotation_level2 %in% c("CD4_cl10", "CD4_cl13", "CD4_cl14", "CD4_cl15", "CD4_cl16", "CD4_cl17", "CD4_cl18", "CD4_cl20", "CD4_cl21", "CD4_cl22", "CD4_cl23", "CD4_cl25", "CD4_cl9", "CD8AB_cl10", "CD8AB_cl11", "CD8AB_cl12", "CD8AB_cl13", "CD8AB_cl14", "CD8AB_cl15", "CD8AB_cl18", "CD8AB_cl20", "CD8AB_cl6", "CD8AB_cl7","treg_cl2", "treg_cl3", "treg_cl4", "treg_cl6", "treg_cl8")] = "activated"
so$level2_parent[so$annotation_level2 %in% c("CD4_cl1", "CD4_cl12", "CD4_cl2", "CD4_cl3", "CD4_cl4", "CD4_cl5", "CD4_cl7", "CD8AB_cl1", "CD8AB_cl2", "CD8AB_cl22", "CD8AB_cl4", "CD8AB_cl5","treg_cl1", "treg_cl7")] = "resting"

table(so$annotation_level1, so$level2_parent)

```

LEVEL1
1. Cleaning and Defining nonconv: check if needs to be rerun at at the end before might be overrun by manual gating.
All iNKT and MAIT by TCR in abT_CD8aa,abT_DN,abT_DP,CD4,CD8AB,Treg,nonconv, gdT without TCRGD ADT signal --> nonconv in level1
nonconv without canonical iNKT or MAIT & gdT_gated --> gdT
```{r}
iNKT = read.table("nonconv/inktInDec24Freeze.csv", header = T, sep = ",")[,1]
so$iNKT2 = colnames(so) %in% iNKT
MAIT = read.table("nonconv/maitInDec24Freeze.csv", header = T, sep = ",")[,1]
so$MAIT2 = colnames(so) %in% MAIT

so$level1[so$iNKT2 & !so$annotation_level1 %in% c("remove", "nonT")] = "nonconv"
so$level1[so$MAIT2 & !so$annotation_level1 %in% c("remove", "nonT")] = "nonconv"
so$level1[so$annotation_level1 == "nonconv" & so$gated_pop == "gdT_gated" & !so$MAIT2 & !so$iNKT2] = "gdT"

table(so$level1, so$annotation_level1)

```


2. cleaning CD4
CD4_cl8 --> prolif
CD4 in annotation_level1 & CD8AB_gated --> CD8AB
CD4 in annotation_level1 & CD8AA_gated --> CD8AA
CD4 in annotation_level1 & gdT_gated --> gdT
```{r}
so$level1[so$annotation_level1 == "CD4" & so$gated_pop == "CD8AB_gated"] = "CD8AB"
so$level1[so$annotation_level1 == "CD4" & so$gated_pop == "CD8AA_gated"] = "abT_CD8aa"
so$level1[so$annotation_level1 == "CD4" & so$gated_pop == "gdT_gated"] = "gdT"
so$level1[so$level1 == "CD4" & so[["ADT"]]@data["CD8B",] > 2 & so[["ADT"]]@data["CD4",] < 1 ] = "CD8AB"
so$level1[so$level1 == "CD4" & so[["ADT"]]@data["CD8B",] > 2 & so[["ADT"]]@data["CD4",] > 1 ] = "abT_DP"
so$level1[so$level1 == "CD4" & so[["ADT"]]@data["CD8B",] < 2 & so[["ADT"]]@data["CD4",] < 1 & so[["ADT"]]@data["CD8A",] > 2] = "abT_CD8aa"

FeatureScatter(so[,so$annotation_level1 == "CD4"], feature1 = "CD4", feature2 = "CD8B", group.by = "level1", split.by = NULL, ncol = 2, pt.size = 2, raster = F, plot.cor = F) + scale_color_manual(values = mypal_level1)
MyDimPlotHighlight(seurat_object = so, umap_to_plot = "mde2_totalvi_20241006", cells_to_highlight = colnames(so)[so$annotation_level1 == "CD4"], highlight_column_name = "level1", labelclusters = T,
                   title = "", pixels = c(512, 512), highlight_size = 1, 
                   highlight_alpha = 1, mycols = mypal_level1,print_plot1 = F, 
                   print_plot2 = F)

table(so$level1, so$annotation_level1)

```

3. cleaning CD8AB
CD8AB_cl19 --> prolif
CD8AB in annotation_level1 & CD4_gated --> CD4
CD8AB in annotation_level1 & CD8AA_gated --> CD8AA
CD8AB in annotation_level1 & gdT_gated --> gdT
```{r}
so$level1[so$annotation_level1 == "CD8AB"] = "CD8AB"
so$level1[so$annotation_level1 == "CD8AB" & so$gated_pop == "CD4_gated"] = "CD4"
so$level1[so$annotation_level1 == "CD8AB" & so$gated_pop == "CD8AA_gated"] = "abT_CD8aa"
so$level1[so$annotation_level1 == "CD8AB" & so$gated_pop == "gdT_gated"] = "gdT"
so$level1[so$level1 == "CD8AB" & so[["ADT"]]@data["CD4",] > 2 & so[["ADT"]]@data["CD8B",] < 1 ] = "CD4"
so$level1[so$level1 == "CD8AB" & so[["ADT"]]@data["CD4",] > 2 & so[["ADT"]]@data["CD8B",] > 2 ] = "abT_DP"
# so$level1[so$level1 == "CD8AB" & so[["ADT"]]@data["CD8B",] < 1 & so[["ADT"]]@data["CD4",] < 1 & so[["ADT"]]@data["CD8A",] > 2] = "abT_CD8aa"
# so$level1[so$level1 == "CD8AB" & so[["ADT"]]@data["CD8B",] < 1 & so[["ADT"]]@data["CD4",] < 1 & so[["ADT"]]@data["CD8A",] < 1] = "abT_DN" #DOES NOT WORK CAN'T HAVE ALL NEG, JUST FALSE NEG HERE

FeatureScatter(so[,so$annotation_level1 == "CD8AB"], feature1 = "CD4", feature2 = "CD8B", group.by = "level1", split.by = "level1", ncol = 4, pt.size = 2, raster = F, plot.cor = F) + scale_color_manual(values = mypal_level1)
MyDimPlotHighlight(seurat_object = so, umap_to_plot = "mde2_totalvi_20241006", cells_to_highlight = colnames(so)[so$annotation_level1 == "CD8AB"], highlight_column_name = "level1", labelclusters = F,
                   title = "", pixels = c(512, 512), highlight_size = 1, 
                   highlight_alpha = 1, mycols = mypal_level1,print_plot1 = T, 
                   print_plot2 = F)

table(so$level1, so$annotation_level1)
```

5. Cleaning treg
treg in annotation_level1 & gdT_gated --> gdT
```{r}
so$level1[so$annotation_level1 == "treg"] = "treg"
so$level1[so$annotation_level1 == "treg" & so$gated_pop == "gdT_gated"] = "gdT"
so$level1[so$annotation_level1 == "treg" & so$gated_pop == "CD8AB_gated"] = "CD8AB"
so$level1[so$level1 == "treg" & so[["ADT"]]@data["CD4",] > 1 & so[["ADT"]]@data["CD8B",] > 2 ] = "abT_DP"
so$level1[so$level1 == "treg" & so[["ADT"]]@data["CD4",] < 1 & so[["ADT"]]@data["CD8B",] > 2 ] = "CD8AB"

FeatureScatter(so[,so$annotation_level1 == "treg" & so$organ != "thymus"], feature1 = "CD4", feature2 = "CD8B", group.by = "level1", split.by = NULL, ncol = 4, pt.size = 2, raster = F, plot.cor = F) + scale_color_manual(values = mypal_level1)
FeatureScatter(so[,so$annotation_level1 == "treg"], feature1 = "CD4", feature2 = "CD8B", group.by = "level1", split.by = NULL, ncol = 4, pt.size = 2, raster = F, plot.cor = F) + scale_color_manual(values = mypal_level1)
MyDimPlotHighlight(seurat_object = so, umap_to_plot = "mde2_totalvi_20241006", cells_to_highlight = colnames(so)[so$annotation_level1 == "treg"], highlight_column_name = "level1", labelclusters = F,
                   title = "", pixels = c(512, 512), highlight_size = 1, 
                   highlight_alpha = 1, mycols = mypal_level1,print_plot1 = T, 
                   print_plot2 = F)

table(so$level1, so$annotation_level1)
```

6. cleaning gdT
gdT in annotation_level1 & CD4_gated --> CD4
gdT in annotation_level1 & CD8AA_gated --> CD8AA
gdT in annotation_level1 & CD8AB_gated --> CD8AB
so$level1 == "gdT" & so[["ADT"]]@data["TCRB",] > 1 & so[["ADT"]]@data["TCRGD",] < 1 & !so$gated_pop %in% c("CD4_gated", "CD8AA_gated", "CD8AB_gated") --> abT_DN
```{r}
so$level1[so$annotation_level1 == "gdT" & so$gated_pop == "CD4_gated"] = "CD4"
so$level1[so$annotation_level1 == "gdT" & so$gated_pop == "CD8AA_gated"] = "abT_CD8aa"
so$level1[so$annotation_level1 == "gdT" & so$gated_pop == "CD8AB_gated"] = "CD8AB"
so$level1[so$level1 == "gdT" & so[["ADT"]]@data["TCRB",] > 1 & so[["ADT"]]@data["TCRGD",] < 1 & !so$gated_pop %in% c("CD4_gated", "CD8AA_gated", "CD8AB_gated")] = "abT_DN"

FeatureScatter(so[,so$level1 == "gdT"], feature1 = "TCRB", feature2 = "TCRGD", group.by = "gated_pop", split.by = NULL, ncol = 2, pt.size = 2, raster = T, plot.cor = F) + geom_vline(aes(xintercept = 1), linetype="dashed", color = "brown") 
MyDimPlotHighlight(seurat_object = so, umap_to_plot = "mde2_totalvi_20241006", cells_to_highlight = colnames(so)[so$annotation_level1 == "gdT"], highlight_column_name = "level1", labelclusters = F,
                   title = "", pixels = c(512, 512), highlight_size = 1, 
                   highlight_alpha = 1, mycols = mypal_level1,print_plot1 = T, 
                   print_plot2 = F)


table(so$level1, so$annotation_level1)

```

7. cleaning CD8AA
abT_CD8aa in annotation_level1 & CD4_gated --> CD4
abT_CD8aa in annotation_level1 & gdT_gated --> gdT
abT_CD8aa in annotation_level1 & CD8AB_gated --> CD8AB
```{r}
so$level1[so$annotation_level1 == "abT_CD8aa" & so$gated_pop == "CD4_gated"] = "CD4"
so$level1[so$annotation_level1 == "abT_CD8aa" & so$gated_pop == "gdT_gated"] = "gdT"
so$level1[so$annotation_level1 == "abT_CD8aa" & so$gated_pop == "CD8AB_gated"] = "CD8AB"
so$level1[so$level1 == "abT_CD8aa" & so[["ADT"]]@data["CD4",] > 2 & so[["ADT"]]@data["CD8B",] < 1 ] = "CD4"
so$level1[so$level1 == "abT_CD8aa" & so[["ADT"]]@data["CD4",] > 2 & so[["ADT"]]@data["CD8B",] > 2 ] = "abT_DP"

FeatureScatter(so[,so$annotation_level1 == "abT_CD8aa"], feature1 = "CD4", feature2 = "CD8B", group.by = "level1", split.by = NULL, ncol = 4, pt.size = 2, raster = F, plot.cor = F) + scale_color_manual(values = mypal_level1)
MyDimPlotHighlight(seurat_object = so, umap_to_plot = "mde2_totalvi_20241006", cells_to_highlight = colnames(so)[so$annotation_level1 == "abT_CD8aa"], highlight_column_name = "level1", labelclusters = F,
                   title = "", pixels = c(512, 512), highlight_size = 1, 
                   highlight_alpha = 1, mycols = mypal_level1,print_plot1 = T, 
                   print_plot2 = F)

table(so$level1, so$annotation_level1)
```

8. cleaning abT_DN
```{r}
so$level1[so$annotation_level1 == "abT_DN" & so$gated_pop == "CD4_gated"] = "CD4"
so$level1[so$annotation_level1 == "abT_DN" & so$gated_pop == "CD8AB_gated"] = "CD8AB"
so$level1[so$annotation_level1 == "abT_DN" & so$gated_pop == "CD8AA_gated"] = "abT_CD8aa"
so$level1[so$annotation_level1 == "abT_DN" & so$gated_pop == "gdT_gated"] = "gdT"
so$level1[so$level1 == "abT_DN" & so[["ADT"]]@data["CD4",] > 2 & so[["ADT"]]@data["CD8B",] < 1 ] = "CD4"
so$level1[so$level1 == "abT_DN" & so[["ADT"]]@data["CD4",] < 1 & so[["ADT"]]@data["CD8B",] > 2 ] = "CD8AB"
so$level1[so$level1 == "abT_DN" & so[["ADT"]]@data["CD4",] > 1 & so[["ADT"]]@data["CD8B",] > 2 ] = "abT_DP"

FeatureScatter(so[,so$annotation_level1 == "abT_DN"], feature1 = "CD4", feature2 = "CD8B", group.by = "level1", split.by = NULL, ncol = 4, pt.size = 2, raster = F, plot.cor = F) + scale_color_manual(values = mypal_level1)
MyDimPlotHighlight(seurat_object = so, umap_to_plot = "mde2_totalvi_20241006", cells_to_highlight = colnames(so)[so$annotation_level1 == "abT_DN"], highlight_column_name = "level1", labelclusters = F,
                   title = "", pixels = c(512, 512), highlight_size = 1, 
                   highlight_alpha = 1, mycols = mypal_level1,print_plot1 = T, 
                   print_plot2 = F)

table(so$level1, so$annotation_level1)

```

9. cleaning abT_DP
```{r}
so$level1[so$annotation_level1 == "abT_DP"] = "abT_DP"
so$level1[so$annotation_level1 == "abT_DP" & so$gated_pop == "CD4_gated"] = "CD4"
so$level1[so$annotation_level1 == "abT_DP" & so$gated_pop == "CD8AB_gated"] = "CD8AB"
so$level1[so$annotation_level1 == "abT_DP" & so$gated_pop == "CD8AA_gated"] = "abT_CD8aa"
so$level1[so$annotation_level1 == "abT_DP" & so$gated_pop == "gdT_gated"] = "gdT"
so$level1[so$level1 == "abT_DP" & so[["ADT"]]@data["CD4",] > 2 & so[["ADT"]]@data["CD8B",] < 1 ] = "CD4"
so$level1[so$level1 == "abT_DP" & so[["ADT"]]@data["CD4",] < 1 & so[["ADT"]]@data["CD8B",] > 2 ] = "CD8AB"
so$level1[so$level1 == "abT_DP" & so[["ADT"]]@data["CD4",] < 1 & so[["ADT"]]@data["CD8B",] < 1 & so[["ADT"]]@data["CD8A",] > 2] = "abT_CD8aa"
so$level1[so$level1 == "abT_DP" & so[["ADT"]]@data["TCRGD",] <2 & so[["ADT"]]@data["TCRB",] > 1 & so[["ADT"]]@data["CD4",] < 1 & so[["ADT"]]@data["CD8B",] < 1 & so[["ADT"]]@data["CD8A",] < 1] = "abT_DN"

FeatureScatter(so[,so$annotation_level1 == "abT_DP"], feature1 = "CD4", feature2 = "CD8B", group.by = "level1", split.by = "level1", ncol = 4, pt.size = 2, raster = F, plot.cor = F) + scale_color_manual(values = mypal_level1)
MyDimPlotHighlight(seurat_object = so, umap_to_plot = "mde2_totalvi_20241006", cells_to_highlight = colnames(so)[so$annotation_level1 == "abT_DP"], highlight_column_name = "level1", labelclusters = F,
                   title = "", pixels = c(512, 512), highlight_size = 1, 
                   highlight_alpha = 1, mycols = mypal_level1,print_plot1 = T, 
                   print_plot2 = F)



table(so$level1, so$annotation_level1)
table(so$level1, so$organ == "thymus")

```

10. splitting preT that are not in the thymus
```{r}
##1. some plots before reassigning to see

DimPlot(so, reduction = "mde2_totalvi_20241006", group.by = "is_thymus", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal)
DimPlot(so, reduction = "mde2_totalvi_20241006", group.by = "is_thymus", split = "is_thymus", raster = T, raster.dpi = c(512,512), label = F) + scale_color_manual(values = c("blue", "red"))

pdf("gating_Leon_20241116/FeatureScatter_preT_nonthymus_tofindgates.pdf", width = 7,7, useDingbats = F)
FeatureScatter(so[,so$annotation_level1 == "thymocyte" & so$organ != "thymus"], feature1 = "TCRB", feature2 = "TCRGD", group.by = "gated_pop", split.by = NULL, ncol = 2, pt.size = 2, raster = T, plot.cor = F) 
FeatureScatter(so[,so$annotation_level1 == "thymocyte" & so$organ != "thymus"], feature1 = "CD4", feature2 = "CD8B", group.by = "gated_pop", split.by = NULL, ncol = 2, pt.size = 2, raster = T, plot.cor = F) 
FeatureScatter(so[,so$annotation_level1 == "thymocyte" & so$organ != "thymus"], feature1 = "CD8B", feature2 = "CD8A", group.by = "gated_pop", split.by = NULL, ncol = 2, pt.size = 2, raster = T, plot.cor = F) 

FeatureScatter(so[,so$annotation_level1 == "thymocyte" & so$organ != "thymus" & !is.na(so$gated_pop)], feature1 = "TCRB", feature2 = "TCRGD", group.by = "gated_pop", split.by = NULL, ncol = 2, pt.size = 2, raster = T, plot.cor = F) 
FeatureScatter(so[,so$annotation_level1 == "thymocyte" & so$organ != "thymus" & !is.na(so$gated_pop)], feature1 = "CD4", feature2 = "CD8B", group.by = "gated_pop", split.by = NULL, ncol = 2, pt.size = 2, raster = T, plot.cor = F) 
FeatureScatter(so[,so$annotation_level1 == "thymocyte" & so$organ != "thymus" & !is.na(so$gated_pop)], feature1 = "CD8B", feature2 = "CD8A", group.by = "gated_pop", split.by = NULL, ncol = 2, pt.size = 2, raster = T, plot.cor = F) 

FeatureScatter(so[,so$annotation_level1 == "thymocyte" & so$organ != "thymus" & is.na(so$gated_pop)], feature1 = "TCRB", feature2 = "TCRGD", group.by = "annotation_level1", split.by = NULL, ncol = 2, pt.size = 2, raster = T, plot.cor = F) #+ geom_vline(aes(xintercept = 1
FeatureScatter(so[,so$annotation_level1 == "thymocyte" & so$organ != "thymus" & is.na(so$gated_pop)], feature1 = "CD4", feature2 = "CD8B", group.by = "annotation_level1", split.by = NULL, ncol = 2, pt.size = 2, raster = T, plot.cor = F) #+ geom_vline(aes(xintercept = 1
FeatureScatter(so[,so$annotation_level1 == "thymocyte" & so$organ != "thymus" & is.na(so$gated_pop)], feature1 = "CD8B", feature2 = "CD8A", group.by = "annotation_level1", split.by = NULL, ncol = 2, pt.size = 2, raster = T, plot.cor = F) #+ geom_vline(aes(xintercept = 1

dev.off()

##2. Reassign with Leon's gates
so$level1[so$annotation_level1 == "thymocyte" & so$organ != "thymus" & so$gated_pop == "CD4_gated"] = "CD4"
so$level1[so$annotation_level1 == "thymocyte" & so$organ != "thymus" & so$gated_pop == "CD8AB_gated"] = "CD8AB"
so$level1[so$annotation_level1 == "thymocyte" & so$organ != "thymus" & so$gated_pop == "CD8AA_gated"] = "abT_CD8aa"
so$level1[so$annotation_level1 == "thymocyte" & so$organ != "thymus" & so$gated_pop == "gdT_gated"] = "gdT"
# so$level1[so$annotation_level1 == "thymocyte" & so$organ != "thymus" & so$MAIT2] = "nonconv"
# so$level1[so$annotation_level1 == "thymocyte" & so$organ != "thymus" & so$iNKT2] = "nonconv"

##3. Manual gating of cells not gated before
so$level1[so$annotation_level1 == "thymocyte" & so$organ != "thymus" &  is.na(so$gated_pop) & so[["ADT"]]@data["TCRGD",] > 2 & so[["ADT"]]@data["TCRB",] < 0.6 ] = "gdT"
so$level1[so$annotation_level1 == "thymocyte" & so$organ != "thymus" &  is.na(so$gated_pop) & so[["ADT"]]@data["TCRGD",] < 2 & so[["ADT"]]@data["TCRB",] > 1 & so[["ADT"]]@data["CD4",] > 1.5 & so[["ADT"]]@data["CD8B",] < 2  ] = "CD4"
so$level1[so$annotation_level1 == "thymocyte" & so$organ != "thymus" &  is.na(so$gated_pop) & so[["ADT"]]@data["TCRGD",] < 2 & so[["ADT"]]@data["TCRB",] > 1 & so[["ADT"]]@data["CD4",] < 1 & so[["ADT"]]@data["CD8B",] > 2  ] = "CD8AB"
so$level1[so$annotation_level1 == "thymocyte" & so$organ != "thymus" &  is.na(so$gated_pop) & so[["ADT"]]@data["TCRGD",] < 2 & so[["ADT"]]@data["TCRB",] > 1 & so[["ADT"]]@data["CD8A",] > 2 & so[["ADT"]]@data["CD8B",] < 2  ] = "abT_CD8aa"
so$level1[so$annotation_level1 == "thymocyte" & so$organ != "thymus" &  is.na(so$gated_pop) & so[["ADT"]]@data["TCRGD",] < 2 & so[["ADT"]]@data["CD4",] > 1.5 & so[["ADT"]]@data["CD8B",] > 2  ] = "abT_DP"
so$level1[so$annotation_level1 == "thymocyte" & so$organ != "thymus" &  is.na(so$gated_pop) & so[["ADT"]]@data["TCRGD",] < 2 & so[["ADT"]]@data["TCRB",] > 0 & so[["ADT"]]@data["CD4",] < 2 & so[["ADT"]]@data["CD8B",] < 2  ] = "abT_DN" #need so[["ADT"]]@data["TCRB",] > 0 to make sure to not misgate IGT without CITEseq?
so$level1[so$annotation_level1 == "thymocyte" & so$organ != "thymus" &  is.na(so$gated_pop) & so[["ADT"]]@data["TCRGD",] < 2 & so[["ADT"]]@data["TCRB",] > 1 & so[["ADT"]]@data["CD4",] < 2 & so[["ADT"]]@data["CD8B",] < 2  ] = "abT_DN"
#540 cells without clear CITEseq signal to make a call
so$level1[so$level1 == "thymocyte" & so$organ != "thymus"] = "unclear"

##4. plot results
pdf("gating_Leon_20241116/FeatureScatter_preT_nonthymus_manual_gate.pdf", width = 7,7, useDingbats = F)
FeatureScatter(so[,so$annotation_level1 == "thymocyte" & so$organ != "thymus" &  is.na(so$gated_pop) ], feature1 = "TCRB", feature2 = "TCRGD", group.by = "level1", split.by = NULL, ncol = 2, pt.size = 2, raster = F, plot.cor = F) + scale_color_manual(values = mypal_level1)
FeatureScatter(so[,so$annotation_level1 == "thymocyte" & so$organ != "thymus" &  is.na(so$gated_pop) ], feature1 = "CD4", feature2 = "CD8B", group.by = "level1", split.by = NULL, ncol = 2, pt.size = 2, raster = F, plot.cor = F) + scale_color_manual(values = mypal_level1)
FeatureScatter(so[,so$annotation_level1 == "thymocyte" & so$organ != "thymus" &  is.na(so$gated_pop) ], feature1 = "CD8B", feature2 = "CD8A", group.by = "level1", split.by = NULL, ncol = 2, pt.size = 2, raster = F, plot.cor = F) + scale_color_manual(values = mypal_level1)
dev.off()

table(so$level1, so$annotation_level1)
table(so$level1, so$organ == "thymus")
table(so$level1, so$level2_parent)

pdf("gating_Leon_20241116/MDE_preT_nonthymus_newannotation.pdf", width = 7,7, useDingbats = F)
MyDimPlotHighlight(seurat_object = so, umap_to_plot = "mde2_totalvi_20241006", cells_to_highlight = colnames(so)[so$annotation_level1 == "thymocyte" & so$organ != "thymus"], highlight_column_name = "level1", labelclusters = T,
                   title = "", pixels = c(512, 512), highlight_size = 1, 
                   highlight_alpha = 1, mycols = mypal_level1,print_plot1 = T, 
                   print_plot2 = T)
dev.off()

pdf("gating_Leon_20241116/FeatureScatter_preT_nonthymus_newannotation.pdf", width = 7,7, useDingbats = F)
FeatureScatter(so[,so$annotation_level1 == "thymocyte" & so$organ != "thymus" ], feature1 = "TCRB", feature2 = "TCRGD", group.by = "level1", split.by = NULL, ncol = 2, pt.size = 2, raster = F, plot.cor = F) + scale_color_manual(values = mypal_level1)
FeatureScatter(so[,so$annotation_level1 == "thymocyte" & so$organ != "thymus" ], feature1 = "CD4", feature2 = "CD8B", group.by = "level1", split.by = NULL, ncol = 2, pt.size = 2, raster = F, plot.cor = F) + scale_color_manual(values = mypal_level1)
FeatureScatter(so[,so$annotation_level1 == "thymocyte" & so$organ != "thymus" ], feature1 = "CD8B", feature2 = "CD8A", group.by = "level1", split.by = NULL, ncol = 2, pt.size = 2, raster = F, plot.cor = F) + scale_color_manual(values = mypal_level1)
dev.off()

```

11. Miniverse
miniverse_cl13 to prolif!
```{r}
##1. initial assignement based on clusters
so$level1[so$annotation_level2 %in% "miniverse_cl1"] = "CD4"
so$level1[so$annotation_level2 %in% "miniverse_cl2"] = "CD4"
so$level1[so$annotation_level2 %in% "miniverse_cl3"] = "treg"
so$level1[so$annotation_level2 %in% "miniverse_cl4"] = "treg"
so$level1[so$annotation_level2 %in% "miniverse_cl5"] = "CD8AB"
so$level1[so$annotation_level2 %in% "miniverse_cl6"] = "CD8AB"
so$level1[so$annotation_level2 %in% "miniverse_cl7"] = "CD8AB"
so$level1[so$annotation_level2 %in% "miniverse_cl8"] = "CD8AB"
so$level1[so$annotation_level2 %in% "miniverse_cl9"] = "nonconv"
so$level1[so$annotation_level2 %in% "miniverse_cl10"] = "gdT"
so$level1[so$annotation_level2 %in% "miniverse_cl11"] = "abT_DN"
so$level1[so$annotation_level2 %in% "miniverse_cl12"] = "gdT"
so$level1[so$annotation_level2 %in% "miniverse_cl14"] = "abT_DP"
so$level1[so$annotation_level2 %in% "miniverse_cl15"] = "CD4" #really CD8AB and CD4 and will gate out CD8AB, easier to gate out
so$level1[so$annotation_level2 %in% "miniverse_cl16"] = "nonconv"
so$level1[so$annotation_level2 %in% "miniverse_cl17"] = "abT_DP"
so$level1[so$annotation_level2 %in% "miniverse_cl13"] = "prolif"

##2. Reassign with Leon's gates and TCR
so$level1[so$annotation_level1 == "miniverse" & so$gated_pop == "CD4_gated"] = "CD4"
so$level1[so$annotation_level1 == "miniverse" & so$gated_pop == "CD8AB_gated"] = "CD8AB"
so$level1[so$annotation_level1 == "miniverse" & so$gated_pop == "CD8AA_gated"] = "abT_CD8aa"
so$level1[so$annotation_level1 == "miniverse"& so$gated_pop == "gdT_gated"] = "gdT"
so$level1[so$annotation_level1 == "miniverse" & so$MAIT2] = "nonconv" # needs to be after since they fall in the previous miniverse_cl too!
so$level1[so$annotation_level1 == "miniverse" & so$iNKT2] = "nonconv" # needs to be after since they fall in the previous miniverse_cl too!

##3. Check
pdf("gating_Leon_20241116/FeatureScatter_miniverse_nonthymus_tofindgates.pdf", width = 7,7, useDingbats = F)
FeatureScatter(so[,so$annotation_level1 == "miniverse" & so$organ != "thymus" ], feature1 = "TCRB", feature2 = "TCRGD", group.by = "level1", split.by = "level1", ncol = 3, pt.size = 1, raster = F, plot.cor = F) + scale_color_manual(values = mypal_level1)
FeatureScatter(so[,so$annotation_level1 == "miniverse" & so$organ != "thymus" ], feature1 = "CD4", feature2 = "CD8B", group.by = "level1", split.by = "level1", ncol = 3, pt.size = 1, raster = F, plot.cor = F) + scale_color_manual(values = mypal_level1)
FeatureScatter(so[,so$annotation_level1 == "miniverse" & so$organ != "thymus" ], feature1 = "CD8B", feature2 = "CD8A", group.by = "level1", split.by = "level1", ncol = 3, pt.size = 1, raster = F, plot.cor = F) + scale_color_manual(values = mypal_level1)
FeatureScatter(so[,so$annotation_level1 == "miniverse" & so$organ != "thymus" & is.na(so$gated_pop)], feature1 = "TCRB", feature2 = "TCRGD", group.by = "level1", split.by = "level1", ncol = 3, pt.size = 1, raster = F, plot.cor = F) + scale_color_manual(values = mypal_level1)
FeatureScatter(so[,so$annotation_level1 == "miniverse" & so$organ != "thymus" & is.na(so$gated_pop)], feature1 = "CD4", feature2 = "CD8B", group.by = "level1", split.by = "level1", ncol = 3, pt.size = 1, raster = F, plot.cor = F) + scale_color_manual(values = mypal_level1)
FeatureScatter(so[,so$annotation_level1 == "miniverse" & so$organ != "thymus" & is.na(so$gated_pop)], feature1 = "CD8B", feature2 = "CD8A", group.by = "level1", split.by = "level1", ncol = 3, pt.size = 1, raster = F, plot.cor = F) + scale_color_manual(values = mypal_level1)
dev.off()

##4. Manual gating of cells not gated before
# NO NEED HERE: so$level1[so$annotation_level1 == "miniverse" & so$organ != "thymus" &  is.na(so$gated_pop) & so[["ADT"]]@data["TCRGD",] > 2 & so[["ADT"]]@data["TCRB",] < 0.6 ] = "gdT"
#CD4
so$level1[so$annotation_level1 == "miniverse" & so$level1 %in% c( "abt_DN", "CD4", "CD8AB","prolif", "treg") & so$organ != "thymus" &  is.na(so$gated_pop) & so[["ADT"]]@data["TCRGD",] < 2 & so[["ADT"]]@data["CD4",] > 1 & so[["ADT"]]@data["CD8B",] > 2  ] = "abT_DP"
so$level1[so$annotation_level1 == "miniverse" & so$organ != "thymus" & so$level1 %in% c("abt_DN", "abT_DP", "CD8AB", "prolif") &  is.na(so$gated_pop) & so[["ADT"]]@data["TCRGD",] < 2 & so[["ADT"]]@data["CD4",] > 1.5 & so[["ADT"]]@data["CD8B",] < 2  ] = "CD4"
so$level1[so$annotation_level1 == "miniverse" & so$organ != "thymus" & so$level1 %in% c("abt_DN", "abT_DP", "CD4", "prolif", "treg") &  is.na(so$gated_pop) & so[["ADT"]]@data["TCRGD",] < 2 & so[["ADT"]]@data["CD4",] < 1 & so[["ADT"]]@data["CD8B",] > 1.5  ] = "CD8AB"
so$level1[so$annotation_level1 == "miniverse" & so$organ != "thymus" & so$level1 %in% c("abt_DN", "abT_DP",  "CD4", "CD8AB", "prolif", "treg") &  is.na(so$gated_pop) & so[["ADT"]]@data["TCRGD",] < 2 & so[["ADT"]]@data["CD4",] < 1 & so[["ADT"]]@data["CD8B",] < 2 & so[["ADT"]]@data["CD8A",] > 2   ] = "abT_CD8aa"
so$level1[so$annotation_level1 == "miniverse" & so$organ != "thymus" & so$level1 %in% c("prolif") &  is.na(so$gated_pop) & so[["ADT"]]@data["TCRGD",] < 2 & so[["ADT"]]@data["TCRB",] > 1 & so[["ADT"]]@data["CD4",] < 1 & so[["ADT"]]@data["CD8B",] < 1 & so[["ADT"]]@data["CD8A",] < 1   ] = "abT_DN"

so$level1[so$annotation_level1 == "miniverse" & so$organ != "thymus" & so$level1 %in% c("gdT") &  is.na(so$gated_pop) & so[["ADT"]]@data["TCRGD",] < 1 & so[["ADT"]]@data["TCRB",] > 1 & so[["ADT"]]@data["CD4",] > 1 & so[["ADT"]]@data["CD8B",] > 2  ] = "abT_DP"
so$level1[so$annotation_level1 == "miniverse" & so$organ != "thymus" & so$level1 %in% c("gdT") &  is.na(so$gated_pop) & so[["ADT"]]@data["TCRGD",] < 1 & so[["ADT"]]@data["TCRB",] > 1& so[["ADT"]]@data["CD4",] > 1.5 & so[["ADT"]]@data["CD8B",] < 2  ] = "CD4"
so$level1[so$annotation_level1 == "miniverse" & so$organ != "thymus" & so$level1 %in% c("gdT") &  is.na(so$gated_pop) & so[["ADT"]]@data["TCRGD",] < 1 & so[["ADT"]]@data["TCRB",] > 1 & so[["ADT"]]@data["CD4",] < 1 & so[["ADT"]]@data["CD8B",] > 1.5  ] = "CD8AB"
so$level1[so$annotation_level1 == "miniverse" & so$organ != "thymus" & so$level1 %in% c("gdT") &  is.na(so$gated_pop) & so[["ADT"]]@data["TCRGD",] < 1 & so[["ADT"]]@data["TCRB",] > 1 & so[["ADT"]]@data["CD8A",] > 2 & so[["ADT"]]@data["CD8B",] < 2  ] = "abT_CD8aa"
so$level1[so$annotation_level1 == "miniverse" & so$organ != "thymus" & so$level1 %in% c("gdT", "abT_DP") &  is.na(so$gated_pop) & so[["ADT"]]@data["TCRGD",] < 1 & so[["ADT"]]@data["TCRB",] > 1 & so[["ADT"]]@data["CD4",] < 1 & so[["ADT"]]@data["CD8B",] < 1 & so[["ADT"]]@data["CD8A",] < 1    ] = "abT_DN"
#310 prolif cells that are unclear --> label as unclear
so$level1[so$level1 == "prolif"] = "unclear"

##5. plot results
table(so$level1, so$annotation_level1)
table(so$level1[so$annotation_level1 == "miniverse"], so$annotation_level2[so$annotation_level1 == "miniverse"])

pdf("gating_Leon_20241116/FeatureScatter_miniverse_nonthymus_manual_gate.pdf", width = 7,7, useDingbats = F)
FeatureScatter(so[,so$annotation_level1 == "miniverse" & so$organ != "thymus" ], feature1 = "TCRB", feature2 = "TCRGD", group.by = "level1", split.by = "level1", ncol = 3, pt.size = 2, raster = F, plot.cor = F) + scale_color_manual(values = mypal_level1)
FeatureScatter(so[,so$annotation_level1 == "miniverse" & so$organ != "thymus" ], feature1 = "CD4", feature2 = "CD8B", group.by = "level1", split.by = "level1", ncol = 3, pt.size = 2, raster = F, plot.cor = F) + scale_color_manual(values = mypal_level1)
FeatureScatter(so[,so$annotation_level1 == "miniverse" & so$organ != "thymus" ], feature1 = "CD8B", feature2 = "CD8A", group.by = "level1", split.by = "level1", ncol = 3, pt.size = 2, raster = F, plot.cor = F) + scale_color_manual(values = mypal_level1)
dev.off()

pdf("gating_Leon_20241116/MDE_miniverse_newannotation.pdf", width = 7,7, useDingbats = F)
MyDimPlotHighlight(seurat_object = so, umap_to_plot = "mde2_totalvi_20241006", cells_to_highlight = colnames(so)[so$annotation_level1 == "miniverse" & so$organ != "thymus"], highlight_column_name = "level1", labelclusters = ,
                   title = "", pixels = c(512, 512), highlight_size = 1, 
                   highlight_alpha = 1, mycols = mypal_level1,print_plot1 = T, 
                   print_plot2 = T)
dev.off()
```

12. Prolif of gdT, abT_CD8aa, abT_DN, abT_DP
```{r}

# MyDimPlotHighlight(seurat_object = so, umap_to_plot = "mde2_totalvi_20241006", cells_to_highlight = colnames(so)[so$annotation_level1 == "prolif"], highlight_column_name = "annotation_level1", labelclusters = ,
#                    title = "", pixels = c(512, 512), highlight_size = 1, 
#                    highlight_alpha = 1, mycols = mypal_level1,print_plot1 = T, 
#                    print_plot2 = T)

#1. Annotate from original totalvi_20241014_unconventional_rmIGTsample MDE
cl_unconv = read.table("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/unconventional/totalvi_20241014_unconventional_rmIGTsample/Cluster_totalvi_20241014_unconventional_rmIGTsample_Res.csv", header = F, sep = ",", row.names = 1)
colnames(cl_unconv) = c("res0.25", "res0.5", "res0.75","res1", "res1.5", "res2", "res3", "res4", "res5", "res6")
common_names = intersect(colnames(so), rownames(cl_unconv))
so$prolif_clusters = NA
so$prolif_clusters[common_names] = cl_unconv[common_names, "res4"]
# table(so$prolif_clusters, so$annotation_level1)

tmp = table(so$prolif_clusters, so$level1)
annot_cl = colnames(tmp)[apply(tmp,1, function(x) order(x, decreasing = T)[1])]
names(annot_cl)  = rownames(tmp)
annot_cl

so$prolif_clusters[so$annotation_level1 != "prolif"] = NA
prolif_clusters_annot = annot_cl[so$prolif_clusters]
names(prolif_clusters_annot) = colnames(so)
prolif_clusters_annot = prolif_clusters_annot[!is.na(prolif_clusters_annot)]
so$level1[intersect(colnames(so), names(prolif_clusters_annot))] = prolif_clusters_annot[intersect(colnames(so), names(prolif_clusters_annot))]
table(so$level1, so$annotation_level1)

so$level1[so$prolif_clusters %in% c("cl41","cl42", "cl47", "cl51", "cl64")] = "gdT" ##47 is nonconv or gdT
so$level1[so$prolif_clusters %in% c("cl38")] = "abT_DN"
so$level1[so$prolif_clusters %in% c("cl28","cl45")] = "CD8AB" #DN? nonconv?
so$level1[so$prolif_clusters %in% c("cl41")] = "CD8AB" 

table(so$level1, so$annotation_level1)

##2. Reassign based on Leon's gating and TCR
so$level1[so$annotation_level1 == "prolif" & so$gated_pop == "CD4_gated"] = "CD4"
so$level1[so$annotation_level1 == "prolif" & so$gated_pop == "CD8AB_gated"] = "CD8AB"
so$level1[so$annotation_level1 == "prolif" & so$gated_pop == "CD8AA_gated"] = "abT_CD8aa"
so$level1[so$annotation_level1 == "prolif"& so$gated_pop == "gdT_gated"] = "gdT"
so$level1[so$annotation_level1 == "prolif" & so$MAIT2] = "nonconv" # needs to be after since they fall in the previous cl_ too!
so$level1[so$annotation_level1 == "prolif" & so$iNKT2] = "nonconv" # needs to be after since they fall in the previous cl_ too!

table(so$level1, so$annotation_level1)

##3. check
pdf("gating_Leon_20241116/FeatureScatter_prolif_tofindgates.pdf", width = 7,7, useDingbats = F)
FeatureScatter(so[,so$annotation_level1 == "prolif" ], feature1 = "TCRB", feature2 = "TCRGD", group.by = "level1", split.by = "level1", ncol = 3, pt.size = 1, raster = F, plot.cor = F) + scale_color_manual(values = mypal_level1)
FeatureScatter(so[,so$annotation_level1 == "prolif"], feature1 = "CD4", feature2 = "CD8B", group.by = "level1", split.by = "level1", ncol = 3, pt.size = 1, raster = F, plot.cor = F) + scale_color_manual(values = mypal_level1)
FeatureScatter(so[,so$annotation_level1 == "prolif"], feature1 = "CD8B", feature2 = "CD8A", group.by = "level1", split.by = "level1", ncol = 3, pt.size = 1, raster = F, plot.cor = F) + scale_color_manual(values = mypal_level1)
FeatureScatter(so[,so$annotation_level1 == "prolif" & so$organ != "thymus" & is.na(so$gated_pop)], feature1 = "TCRB", feature2 = "TCRGD", group.by = "level1", split.by = "level1", ncol = 3, pt.size = 1, raster = F, plot.cor = F) + scale_color_manual(values = mypal_level1)
FeatureScatter(so[,so$annotation_level1 == "prolif" & so$organ != "thymus" & is.na(so$gated_pop)], feature1 = "CD4", feature2 = "CD8B", group.by = "level1", split.by = "level1", ncol = 3, pt.size = 1, raster = F, plot.cor = F) + scale_color_manual(values = mypal_level1)
FeatureScatter(so[,so$annotation_level1 == "prolif" & so$organ != "thymus" & is.na(so$gated_pop)], feature1 = "CD8B", feature2 = "CD8A", group.by = "level1", split.by = "level1", ncol = 3, pt.size = 1, raster = F, plot.cor = F) + scale_color_manual(values = mypal_level1)
dev.off()

##4. Manual gating of cells not gated before
# NO NEED HERE: so$level1[so$annotation_level1 == "miniverse" & so$organ != "thymus" &  is.na(so$gated_pop) & so[["ADT"]]@data["TCRGD",] > 2 & so[["ADT"]]@data["TCRB",] < 0.6 ] = "gdT"
so$level1[so$annotation_level1 == "prolif" & so$organ != "thymus" & so$level1 %in% c( "abt_DN", "CD4", "CD8AB", "treg") &  is.na(so$gated_pop) & so[["ADT"]]@data["TCRGD",] < 2 & so[["ADT"]]@data["CD4",] > 2 & so[["ADT"]]@data["CD8B",] > 2  ] = "abT_DP"
so$level1[so$annotation_level1 == "prolif" & so$organ != "thymus" & so$level1 %in% c("abt_DN", "abT_DP", "CD8AB") &  is.na(so$gated_pop) & so[["ADT"]]@data["TCRGD",] < 2 & so[["ADT"]]@data["CD4",] > 1.5 & so[["ADT"]]@data["CD8B",] < 2  ] = "CD4"
so$level1[so$annotation_level1 == "prolif" & so$organ != "thymus" & so$level1 %in% c("abt_DN", "abT_DP", "CD4", "treg") &  is.na(so$gated_pop) & so[["ADT"]]@data["TCRGD",] < 2 & so[["ADT"]]@data["CD4",] < 1 & so[["ADT"]]@data["CD8B",] > 2  ] = "CD8AB"
so$level1[so$annotation_level1 == "prolif" & so$organ != "thymus" & so$level1 %in% c("abt_DN", "abT_DP",  "CD4", "CD8AB", "treg") &  is.na(so$gated_pop) & so[["ADT"]]@data["TCRGD",] < 2 & so[["ADT"]]@data["CD4",] < 1 & so[["ADT"]]@data["CD8B",] < 2 & so[["ADT"]]@data["CD8A",] > 2   ] = "abT_CD8aa"

so$level1[so$annotation_level1 == "prolif" & so$organ != "thymus" & so$level1 %in% c("gdT") &  is.na(so$gated_pop) & so[["ADT"]]@data["TCRGD",] < 1 & so[["ADT"]]@data["TCRB",] > 1 & so[["ADT"]]@data["CD4",] > 1 & so[["ADT"]]@data["CD8B",] > 2  ] = "abT_DP"
so$level1[so$annotation_level1 == "prolif" & so$organ != "thymus" & so$level1 %in% c("gdT") &  is.na(so$gated_pop) & so[["ADT"]]@data["TCRGD",] < 1 & so[["ADT"]]@data["TCRB",] > 1 & so[["ADT"]]@data["CD4",] > 1.5 & so[["ADT"]]@data["CD8B",] < 2  ] = "CD4"
so$level1[so$annotation_level1 == "prolif" & so$organ != "thymus" & so$level1 %in% c("gdT") &  is.na(so$gated_pop) & so[["ADT"]]@data["TCRGD",] < 1 & so[["ADT"]]@data["TCRB",] > 1 & so[["ADT"]]@data["CD4",] < 1 & so[["ADT"]]@data["CD8B",] > 1.5  ] = "CD8AB"
so$level1[so$annotation_level1 == "prolif" & so$organ != "thymus" & so$level1 %in% c("gdT") &  is.na(so$gated_pop) & so[["ADT"]]@data["TCRGD",] < 1 & so[["ADT"]]@data["TCRB",] > 1 & so[["ADT"]]@data["CD8A",] > 2 & so[["ADT"]]@data["CD8B",] < 2  ] = "abT_CD8aa"
so$level1[so$annotation_level1 == "prolif" & so$organ != "thymus" & so$level1 %in% c("gdT", "abT_DP") &  is.na(so$gated_pop) & so[["ADT"]]@data["TCRGD",] < 1 & so[["ADT"]]@data["TCRB",] > 1 & so[["ADT"]]@data["CD4",] < 1 & so[["ADT"]]@data["CD8B",] < 1 & so[["ADT"]]@data["CD8A",] < 1    ] = "abT_DN"

##5. plot results
table(so$level1, so$annotation_level1)
table(so$level2_parent, so$level1)
table(so$level2_parent, so$annotation_level1)

pdf("gating_Leon_20241116/FeatureScatter_prolif_nonthymus_manual_gate.pdf", width = 7,7, useDingbats = F)
FeatureScatter(so[,so$annotation_level1 == "prolif" & so$organ != "thymus" ], feature1 = "TCRB", feature2 = "TCRGD", group.by = "level1", split.by = "level1", ncol = 3, pt.size = 2, raster = F, plot.cor = F) + scale_color_manual(values = mypal_level1)
FeatureScatter(so[,so$annotation_level1 == "prolif" & so$organ != "thymus" ], feature1 = "CD4", feature2 = "CD8B", group.by = "level1", split.by = "level1", ncol = 3, pt.size = 2, raster = F, plot.cor = F) + scale_color_manual(values = mypal_level1)
FeatureScatter(so[,so$annotation_level1 == "prolif" & so$organ != "thymus" ], feature1 = "CD8B", feature2 = "CD8A", group.by = "level1", split.by = "level1", ncol = 3, pt.size = 2, raster = F, plot.cor = F) + scale_color_manual(values = mypal_level1)

FeatureScatter(so[,so$level2_parent == "prolif" & so$organ != "thymus" ], feature1 = "TCRB", feature2 = "TCRGD", group.by = "level1", split.by = "level1", ncol = 3, pt.size = 2, raster = F, plot.cor = F) + scale_color_manual(values = mypal_level1)
FeatureScatter(so[,so$level2_parent == "prolif" & so$organ != "thymus" ], feature1 = "CD4", feature2 = "CD8B", group.by = "level1", split.by = "level1", ncol = 3, pt.size = 2, raster = F, plot.cor = F) + scale_color_manual(values = mypal_level1)
FeatureScatter(so[,so$level2_parent == "prolif" & so$organ != "thymus" ], feature1 = "CD8B", feature2 = "CD8A", group.by = "level1", split.by = "level1", ncol = 3, pt.size = 2, raster = F, plot.cor = F) + scale_color_manual(values = mypal_level1)
dev.off()

pdf("gating_Leon_20241116/MDE_prolif_newannotation.pdf", width = 7,7, useDingbats = F)
MyDimPlotHighlight(seurat_object = so, umap_to_plot = "mde2_totalvi_20241006", cells_to_highlight = colnames(so)[so$annotation_level1 == "prolif"], highlight_column_name = "level1", labelclusters = T, title = "only reannotated unconv prolif ", pixels = c(512, 512), highlight_size = 1,  highlight_alpha = 1, mycols = mypal_level1,print_plot1 = T,  print_plot2 = T)

MyDimPlotHighlight(seurat_object = so, umap_to_plot = "mde2_totalvi_20241006", cells_to_highlight = colnames(so)[so$level2_parent == "prolif"], highlight_column_name = "level1", labelclusters = T, title = "all level2 parent prolif cells", pixels = c(512, 512), highlight_size = 1, highlight_alpha = 1, mycols = mypal_level1,print_plot1 = T, print_plot2 = T)
dev.off()

```

WRITE ANNOTATION FILE
--> annotation_table_20250109_withlevel220241216.csv
igt1_96_withtotalvi20250109.Rds
```{r}
annot = data.frame(cell_id = colnames(so), so@meta.data[,c("level1", "level2_parent", "annotation_level2", "nonconv_tcr_recog")])
colnames(annot) = c("cell_id", "level1", "level2_parent1", "level2", "nonconv_tcr_recog")
annot$level2_parent1[annot$level2_parent1 == "prolif"] = "proliferating"
table(annot$level2_parent1, annot$level2)
table(annot$level2_parent1, annot$level1)

sum(rowSums(apply(annot,2,is.na)))

write.table(x = annot, file = "annotation_table_20250109_withlevel220241216.csv", row.names = F, quote = F, sep = ",")

# annot_summary <- annot %>%
#   group_by(level2, level1, level2_parent1) %>% 
#   summarize(counts = n(), .groups = "drop") %>%
#   arrange(
#     sub("^.*_cl", "", level2) %>% as.numeric(), # Extract cluster number and convert to numeric
#     level2                                     # Keep cluster prefix order (CD4, treg, etc.)
#   ) %>% 
#     arrange(level1) %>%
#     as.data.frame()
# write.table(x = annot_summary, file = "annotation_table_history_20250109.csv", row.names = F, quote = F, sep = ",")

annot_summary <- annot %>%
    select(level1,level2_parent1) %>%
    group_by(level1, level2_parent1) %>% 
    summarize(counts = n(), .groups = "drop") %>%
    arrange(level1) %>%
    as.data.frame()

write.table(x = annot_summary, file = "annotation_table_history_20250109_level1only.csv", row.names = F, quote = F, sep = ",")
```

Summary plots of changes: check files but should be ok:
```{r}
so_orig = readRDS("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/igt1_96_withtotalvi20250109.Rds")
so = so_orig

mypal_level1 = setNames(mypal, unique(so$annotation_level1))
mypal_level2 = setNames(mypal, unique(so$annotation_level2))
mypal_level2_parent1 = setNames(rev(mypal), unique(so$annotation_level2_parent1))

so = NormalizeData(so,assay = "ADT", normalization.method = "CLR", verbose = T)

#conversion to keep my previous code:
so$level1 = so$annotation_level1
so$level2_parent = so$annotation_level2_parent1
so$annotation_level1 = so$annotation_level1_20241216

table(so$level1, so$annotation_level1)
table(so$level2_parent, so$level1)
table(so$level2_parent, so$annotation_level1)
so2 = so[, !so$level1 %in% c("nonT", "remove", "unclear")]
so3 = so[, so$organ != "thymus" & !so$level1 %in% c("nonT", "remove", "unclear")]

pdf("gating_Leon_20241116/MDE_annotation_20250109_vs_20241216.pdf", 10,10, useDingbats = F)
for (i in unique(so$annotation_level1)) {
    print(i)
    p = MyDimPlotHighlight(seurat_object = so, umap_to_plot = "mde2_totalvi_20241006", cells_to_highlight = colnames(so)[so$annotation_level1 == i], highlight_column_name = "level1", labelclusters = F, title = sprintf("annotation_20241216_level1 %s",i), pixels = c(512, 512), highlight_size = 0.5, highlight_alpha = 0.5, mycols = mypal_level1, print_plot1 = T, print_plot2 = T)
    print(p)
}
dev.off()

pdf("gating_Leon_20241116/MDE_annotation_20250109_vs_20241216_nothymusnocrap.pdf", 10,10, useDingbats = F)
for (i in unique(so3$annotation_level1)) {
    print(i)
    p = MyDimPlotHighlight(seurat_object = so3, umap_to_plot = "mde2_totalvi_20241006", cells_to_highlight = colnames(so3)[so3$annotation_level1 == i], highlight_column_name = "level1", labelclusters = F, title = sprintf("annotation_20241216_level1 %s",i), pixels = c(512, 512), highlight_size = 0.5, highlight_alpha = 0.5, mycols = mypal_level1, print_plot1 = T, print_plot2 = T)
    print(p)
}
dev.off()

pdf("gating_Leon_20241116/FeatureScatter_annotation_20250109_vs_20241216_nothymusnocrap.pdf", 10,7, useDingbats = F)
for (i in unique(so3$annotation_level1)) {
    print(i)
    p = FeatureScatter(so3[,so3$annotation_level1 == i], feature1 = "TCRB", feature2 = "TCRGD", group.by = "level1", split.by = "level1", ncol = 3, pt.size = 2, raster = T, plot.cor = F) + scale_color_manual(values = mypal_level1) + ggtitle(sprintf("annotation_20241216_level1 %s",i)) + theme_dark()
    print(p)
    p =FeatureScatter(so3[,so3$annotation_level1 == i], feature1 = "CD4", feature2 = "CD8B", group.by = "level1", split.by = "level1", ncol = 3, pt.size = 2, raster = T, plot.cor = F) + scale_color_manual(values = mypal_level1) + ggtitle(sprintf("annotation_20241216_level1 %s",i)) + theme_dark()
    print(p)
    p =FeatureScatter(so3[,so3$annotation_level1 == i], feature1 = "CD8B", feature2 = "CD8A", group.by = "level1", split.by = "level1", ncol = 3, pt.size = 2, raster = T, plot.cor = F) + scale_color_manual(values = mypal_level1) + ggtitle(sprintf("annotation_20241216_level1 %s",i)) + theme_dark()
    print(p)
    }
dev.off()
```


Final MDE plots and FeatureScatter
```{r}
pdf("MDE_annotation_20250109_nothymusnocrap.pdf", 10,10, useDingbats = F)
DimPlot(so3, reduction = "mde2_totalvi_20241006", group.by = "level1", raster = T, raster.dpi = c(1024,1024), label = F) + scale_color_manual(values = mypal_level1)
DimPlot(so3, reduction = "mde2_totalvi_20241006", group.by = "level2_parent", raster = T, raster.dpi = c(1024,1024), label = F) + scale_color_manual(values = mypal_level2_parent1)
DimPlot(so3, reduction = "mde2_totalvi_20241006", group.by = "level1", split.by = "level2_parent", ncol = 2, raster = T, raster.dpi = c(1024,1024), label = F) + scale_color_manual(values = mypal_level1)
DimPlot(so3, reduction = "mde2_totalvi_20241006", group.by = "level2_parent", split.by = "level1", ncol = 2, raster = T, raster.dpi = c(1024,1024), label = F) + scale_color_manual(values = mypal_level2_parent1)
dev.off()

pdf("MDE_annotation_20250109_nothymusnocrap2.pdf", 10,10, useDingbats = F)
for (i in unique(so3$level1)) {
    print(i)
    p = MyDimPlotHighlight(seurat_object = so3, umap_to_plot = "mde2_totalvi_20241006", cells_to_highlight = colnames(so3)[so3$level1 == i], highlight_column_name = "level1", labelclusters = F, title = i, pixels = c(512, 512), highlight_size = 0.5, highlight_alpha = 0.5, mycols = mypal_level1, print_plot1 = T, print_plot2 = T)
    print(p)
    p = MyDimPlotHighlight(seurat_object = so3, umap_to_plot = "mde2_totalvi_20241006", cells_to_highlight = colnames(so3)[so3$level1 == i], highlight_column_name = "level2_parent", labelclusters = F, title = i, pixels = c(512, 512), highlight_size = 0.5, highlight_alpha = 0.5, mycols = mypal_level2_parent1, print_plot1 = T, print_plot2 = T)
    print(p)
}
for (i in unique(so3$level2_parent)) {
    print(i)
    p = MyDimPlotHighlight(seurat_object = so3, umap_to_plot = "mde2_totalvi_20241006", cells_to_highlight = colnames(so3)[so3$level2_parent == i], highlight_column_name = "level2_parent", labelclusters = T, title = i, pixels = c(512, 512), highlight_size = 0.5, highlight_alpha = 0.5, mycols = mypal_level2_parent1, print_plot1 = T, print_plot2 = T)
    print(p)
    p = MyDimPlotHighlight(seurat_object = so3, umap_to_plot = "mde2_totalvi_20241006", cells_to_highlight = colnames(so3)[so3$level2_parent == i], highlight_column_name = "level1", labelclusters = T, title = i, pixels = c(512, 512), highlight_size = 0.5, highlight_alpha = 0.5, mycols = mypal_level1, print_plot1 = T, print_plot2 = T)
    print(p)
}
dev.off()

pdf("FeatureScatter_annotation_20250109_level1_nothymusnocrap2.pdf", 7,7, useDingbats = F)
FeatureScatter(so3, feature1 = "TCRB", feature2 = "TCRGD", group.by = "level1", pt.size = 2, raster = T, plot.cor = F) + scale_color_manual(values = mypal_level1)
FeatureScatter(so3, feature1 = "CD4", feature2 = "CD8B", group.by = "level1", pt.size = 2, raster = T, plot.cor = F) + scale_color_manual(values = mypal_level1)
FeatureScatter(so3, feature1 = "CD8B", feature2 = "CD8A", group.by = "level1", pt.size = 2, raster = T, plot.cor = F) + scale_color_manual(values = mypal_level1)
dev.off()

pdf("FeatureScatter_annotation_20250109_level1_nothymusnocrap2.pdf", 40,5, useDingbats = F)
FeatureScatter(so3, feature1 = "TCRB", feature2 = "TCRGD", group.by = "level1", split.by = "level1", ncol = 8, pt.size = 2, raster = T, plot.cor = F) + scale_color_manual(values = mypal_level1)
FeatureScatter(so3, feature1 = "CD4", feature2 = "CD8B", group.by = "level1", split.by = "level1",  ncol = 8, pt.size = 2, raster = T, plot.cor = F) + scale_color_manual(values = mypal_level1)
FeatureScatter(so3, feature1 = "CD8B", feature2 = "CD8A", group.by = "level1", split.by = "level1",  ncol = 8, pt.size = 2, raster = T, plot.cor = F) + scale_color_manual(values = mypal_level1)
dev.off()

###Same but with thymus
pdf("MDE_annotation_20250109_nocrap.pdf", 10,10, useDingbats = F)
DimPlot(so2, reduction = "mde2_totalvi_20241006", group.by = "level1", raster = T, raster.dpi = c(1024,1024), label = F) + scale_color_manual(values = mypal_level1)
DimPlot(so2, reduction = "mde2_totalvi_20241006", group.by = "level2_parent", raster = T, raster.dpi = c(1024,1024), label = F) + scale_color_manual(values = mypal_level2_parent1)
DimPlot(so2, reduction = "mde2_totalvi_20241006", group.by = "level1", split.by = "level2_parent", ncol = 2, raster = T, raster.dpi = c(1024,1024), label = F) + scale_color_manual(values = mypal_level1)
DimPlot(so2, reduction = "mde2_totalvi_20241006", group.by = "level2_parent", split.by = "level1", ncol = 2, raster = T, raster.dpi = c(1024,1024), label = F) + scale_color_manual(values = mypal_level2_parent1)
dev.off()

pdf("MDE_annotation_20250109_nocrap2.pdf", 10,10, useDingbats = F)
for (i in unique(so2$level1)) {
    print(i)
    p = MyDimPlotHighlight(seurat_object = so2, umap_to_plot = "mde2_totalvi_20241006", cells_to_highlight = colnames(so2)[so2$level1 == i], highlight_column_name = "level1", labelclusters = F, title = i, pixels = c(512, 512), highlight_size = 0.5, highlight_alpha = 0.5, mycols = mypal_level1, print_plot1 = T, print_plot2 = T)
    print(p)
    p = MyDimPlotHighlight(seurat_object = so2, umap_to_plot = "mde2_totalvi_20241006", cells_to_highlight = colnames(so2)[so2$level1 == i], highlight_column_name = "level2_parent", labelclusters = F, title = i, pixels = c(512, 512), highlight_size = 0.5, highlight_alpha = 0.5, mycols = mypal_level2_parent1, print_plot1 = T, print_plot2 = T)
    print(p)
}
for (i in unique(so2$level2_parent)) {
    print(i)
    p = MyDimPlotHighlight(seurat_object = so2, umap_to_plot = "mde2_totalvi_20241006", cells_to_highlight = colnames(so2)[so2$level2_parent == i], highlight_column_name = "level2_parent", labelclusters = T, title = i, pixels = c(512, 512), highlight_size = 0.5, highlight_alpha = 0.5, mycols = mypal_level2_parent1, print_plot1 = T, print_plot2 = T)
    print(p)
    p = MyDimPlotHighlight(seurat_object = so2, umap_to_plot = "mde2_totalvi_20241006", cells_to_highlight = colnames(so2)[so2$level2_parent == i], highlight_column_name = "level1", labelclusters = T, title = i, pixels = c(512, 512), highlight_size = 0.5, highlight_alpha = 0.5, mycols = mypal_level1, print_plot1 = T, print_plot2 = T)
    print(p)
}
dev.off()

pdf("FeatureScatter_annotation_20250109_level1_nocrap2.pdf", 7,7, useDingbats = F)
FeatureScatter(so2, feature1 = "TCRB", feature2 = "TCRGD", group.by = "level1", pt.size = 2, raster = T, plot.cor = F) + scale_color_manual(values = mypal_level1)
FeatureScatter(so2, feature1 = "CD4", feature2 = "CD8B", group.by = "level1", pt.size = 2, raster = T, plot.cor = F) + scale_color_manual(values = mypal_level1)
FeatureScatter(so2, feature1 = "CD8B", feature2 = "CD8A", group.by = "level1", pt.size = 2, raster = T, plot.cor = F) + scale_color_manual(values = mypal_level1)
dev.off()

pdf("FeatureScatter_annotation_20250109_level1_nocrap2.pdf", 40,5, useDingbats = F)
FeatureScatter(so2, feature1 = "TCRB", feature2 = "TCRGD", group.by = "level1", split.by = "level1", ncol = 8, pt.size = 2, raster = T, plot.cor = F) + scale_color_manual(values = mypal_level1)
FeatureScatter(so2, feature1 = "CD4", feature2 = "CD8B", group.by = "level1", split.by = "level1",  ncol = 8, pt.size = 2, raster = T, plot.cor = F) + scale_color_manual(values = mypal_level1)
FeatureScatter(so2, feature1 = "CD8B", feature2 = "CD8A", group.by = "level1", split.by = "level1",  ncol = 8, pt.size = 2, raster = T, plot.cor = F) + scale_color_manual(values = mypal_level1)
dev.off()




```

Calculate inaccuracies with CITEseq
```{r}
df = data.frame(so@meta.data[,c("annotation_level1", "level1")])
df$level1 = factor(df$level1)
colnames(df) = c("annotation_level1_20241216", "level1")
df_summary_level1 = df %>% group_by(annotation_level1_20241216, level1) %>% summarize(ncells = n(), .groups = "drop") %>% complete(annotation_level1_20241216, level1, fill = list(ncells = 0)) %>%  group_by(annotation_level1_20241216) %>% mutate(total_cells = sum(ncells), prop = ncells/total_cells) %>% as.data.frame()
write.table(df_summary_level1, file = "gating_Leon_20241116/annotation_20250109_vs_20241216_level1.csv", quote = F, sep = ",", row.names = F, col.names = T)

pdf("gating_Leon_20241116/annotation_20250109_vs_20241216_barplots2.pdf", width = 8, 6, useDingbats = F)
annotation_levels = unique(df_summary_level1$annotation_level1_20241216)
df_summary_level1$level1 = factor(df$level1)
for (level in annotation_levels) {
    print(level)
    # Subset the data for the current annotation_level1
    data_subset <- df_summary_level1 %>%
        filter(annotation_level1_20241216 == level)
    
    # Create the barplot
    p <- ggplot(data_subset, aes(x = level1, y = prop, fill = level1)) +
        geom_bar(stat = "identity") +
        labs(
            title = paste("Proportion of New Annotation in", level),
            x = "New annotation",
            y = "Proportion"
        ) +
        scale_y_continuous(limits = c(0, 1)) +
        scale_color_manual(values = mypal_level1) +
    theme_minimal() +
        theme(
            axis.text.x = element_text(angle = 45, hjust = 1),
            legend.position = "none"
        )
    
    # Print the plot
    print(p)
}
dev.off()


```

Export mudata to calculate new incremental MDE
--> run_totalvi_wrapper_20250109_abT_CD8aa_rmIGTsample.sh etc
```{python}
import warnings; warnings.simplefilter('ignore')
import scvi
import os
import sys
import scanpy as sc
import muon as mu
import mudata as mu

import numpy as np
import mplscience
import matplotlib.pyplot as plt
import pickle
import pandas as pd

working_dir='/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96'
path_to_mudata='/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/export_data/igt1_96_20241006.h5mu'
path_to_latent='/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/totalvi_20241008_rmIGTsample/latent.csv'

mdata = mu.read(path_to_mudata) 

annotationl1 = pd.read_csv(working_dir+"/annotation_table_20250109_withlevel220241216.csv", index_col = 0, header = 0)
set(annotationl1.index) == set(mdata.mod['RNA'].obs.index)
mdata.mod['RNA'].obs['level1'] = annotationl1['level1']
mdata.mod['RNA'].obs['level1'].isna().sum()

latent = pd.read_csv(path_to_latent, index_col = 0, header = 0)

mdata_CD8AB = mdata[mdata.mod['RNA'].obs['level1'] == 'CD8AB'] 
mdata_CD8AB
# mdata_CD8AB.write(working_dir+'/export_data/igt1_96_20250109_CD8AB.h5mu')
latent_subset = latent.loc[mdata_CD8AB.mod['RNA'].obs.index]
latent_subset.to_csv('/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/totalvi_20241008_rmIGTsample/latent_CD8AB.csv', index=True)

mdata_CD4 = mdata[mdata.mod['RNA'].obs['level1'] == 'CD4'] 
mdata_CD4
# mdata_CD4.write(working_dir+'/export_data/igt1_96_20250109_CD4.h5mu')
latent_subset = latent.loc[mdata_CD4.mod['RNA'].obs.index]
latent_subset.to_csv('/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/totalvi_20241008_rmIGTsample/latent_CD4.csv', index=True)

mdata_Treg = mdata[mdata.mod['RNA'].obs['level1'] == 'treg'] 
mdata_Treg
# mdata_Treg.write(working_dir+'/export_data/igt1_96_20250109_Treg.h5mu')
latent_subset = latent.loc[mdata_Treg.mod['RNA'].obs.index]
latent_subset.to_csv('/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/totalvi_20241008_rmIGTsample/latent_Treg.csv', index=True)

mdata_gdT = mdata[mdata.mod['RNA'].obs['level1'] == 'gdT'] 
mdata_gdT
# mdata_gdT.write(working_dir+'/export_data/igt1_96_20250109_gdT.h5mu')
latent_subset = latent.loc[mdata_gdT.mod['RNA'].obs.index]
latent_subset.to_csv('/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/totalvi_20241008_rmIGTsample/latent_gdT.csv', index=True)

mdata_nonconv = mdata[mdata.mod['RNA'].obs['level1'] == 'nonconv'] 
mdata_nonconv
# mdata_nonconv.write(working_dir+'/export_data/igt1_96_20250109_nonconv.h5mu')
latent_subset = latent.loc[mdata_nonconv.mod['RNA'].obs.index]
latent_subset.to_csv('/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/totalvi_20241008_rmIGTsample/latent_nonconv.csv', index=True)

mdata_abT_CD8aa = mdata[mdata.mod['RNA'].obs['level1'] == 'abT_CD8aa'] 
mdata_abT_CD8aa
# mdata_abT_CD8aa.write(working_dir+'/export_data/igt1_96_20250109_abT_CD8aa.h5mu')
latent_subset = latent.loc[mdata_abT_CD8aa.mod['RNA'].obs.index]
latent_subset.to_csv('/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/totalvi_20241008_rmIGTsample/latent_abT_CD8aa.csv', index=True)

mdata_abT_DN = mdata[mdata.mod['RNA'].obs['level1'] == 'abT_DN'] 
mdata_abT_DN
# mdata_abT_DN.write(working_dir+'/export_data/igt1_96_20250109_abT_DN.h5mu')
latent_subset = latent.loc[mdata_abT_DN.mod['RNA'].obs.index]
latent_subset.to_csv('/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/totalvi_20241008_rmIGTsample/latent_abT_DN.csv', index=True)

mdata_abT_DP = mdata[mdata.mod['RNA'].obs['level1'] == 'abT_DP'] 
mdata_abT_DP
# mdata_abT_DP.write(working_dir+'/export_data/igt1_96_20250109_abT_DP.h5mu')
latent_subset = latent.loc[mdata_abT_DP.mod['RNA'].obs.index]
latent_subset.to_csv('/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/totalvi_20241008_rmIGTsample/latent_abT_DP.csv', index=True)


```


Transfer labels + add new clusters if needed +  Readjust level2 parent after transfer labels
--> see code in each clustering_XX.Rmd file
```{r}

```

WRITE ANNOTATION FILE
--> annotation_table_20250109.csv
igt1_96_withtotalvi20250109.Rds
```{r}
annot = data.frame(cell_id = colnames(so_orig), so_orig@meta.data[,c("annotation_level1", "annotation_level2_parent1", "annotation_level2", "nonconv_tcr_recog")])
colnames(annot) = c("cell_id", "level1", "level2_parent1", "level2", "nonconv_tcr_recog")
table(annot$level2_parent1, annot$level1)

sum(rowSums(apply(annot,2,is.na)))

write.table(x = annot, file = "annotation_table_20250109.csv", row.names = F, quote = F, sep = ",")

# annot_summary <- annot %>%
#   group_by(level2, level1, level2_parent1) %>% 
#   summarize(counts = n(), .groups = "drop") %>%
#   arrange(
#     sub("^.*_cl", "", level2) %>% as.numeric(), # Extract cluster number and convert to numeric
#     level2                                     # Keep cluster prefix order (CD4, treg, etc.)
#   ) %>% 
#     arrange(level1) %>%
#     as.data.frame()
# write.table(x = annot_summary, file = "annotation_table_history_20250109.csv", row.names = F, quote = F, sep = ",")

annot_summary <- annot %>%
    select(level1,level2_parent1) %>%
    group_by(level1, level2_parent1) %>% 
    summarize(counts = n(), .groups = "drop") %>%
    arrange(level1) %>%
    as.data.frame()

write.table(x = annot_summary, file = "annotation_table_history_20250109.csv", row.names = F, quote = F, sep = ",")
```



FYI not needed for Treg and DP
sbatch $SCRIPT_DIR/run_mde_incremental_wrapper_20250109_CD8AB.sh
sbatch $SCRIPT_DIR/run_mde_incremental_wrapper_20250109_CD4.sh
sbatch $SCRIPT_DIR/run_mde_incremental_wrapper_20250109_gdT.sh
sbatch $SCRIPT_DIR/run_mde_incremental_wrapper_20250109_abT_CD8aa.sh
sbatch $SCRIPT_DIR/run_mde_incremental_wrapper_20250109_abT_DN.sh
sbatch $SCRIPT_DIR/run_mde_incremental_wrapper_20250109_nonconv.sh
```{bash}
```




############
**Adjust clusters TEST - can just forget now**

```{r}
so = so_orig
so = NormalizeData(so,assay = "ADT", normalization.method = "CLR", verbose = T)
so$level1 = so$annotation_level1

tcrgd_gated = read.table("gating_Leon_20241116/TCRGD_manual_gating_byIGT.csv", header = F, sep = ",")[,1]
so$tcrgd_gated = colnames(so) %in% tcrgd_gated

cd4_gated = read.table("gating_Leon_20241116/IGT1-96_TCRB+CD4+_HardFilter.csv", header = F, sep = ",")[,1]
so$cd4_gated = colnames(so) %in% cd4_gated

cd8ab_gated = read.table("gating_Leon_20241116/IGT1-96_TCRB+CD8AA+CD8AB+_HardFilter.csv", header = F, sep = ",")[,1]
so$cd8ab_gated = colnames(so) %in% cd8ab_gated

cd8aa_gated = read.table("gating_Leon_20241116/IGT1-96_TCRB+CD8AA+CD8AB-_HardFilter.csv", header = F, sep = ",")[,1]
so$cd8aa_gated = colnames(so) %in% cd8aa_gated

so$gated_pop = NA
so$gated_pop[so$tcrgd_gated] = "gdT_gated"
so$gated_pop[so$cd4_gated] = "CD4_gated"
so$gated_pop[so$cd8ab_gated] = "CD8AB_gated"
so$gated_pop[so$cd8aa_gated] = "CD8AA_gated"

```

1. Thymocyte --> preT
```{r}
# so$level1[so$annotation_level2 %in% c("CD8AB_cl19", "CD4_cl8", "treg_cl5")] = "prolif"
so$level1[so$annotation_level1 == "thymocyte"] = "preT"

table(so$level1, so$annotation_level1)

```


2. Cleaning nonconv
All iNKT and MAIT by TCR in abT_CD8aa,abT_DN,abT_DP,CD4,CD8AB,Treg,nonconv, gdT without TCRGD ADT signal --> nonconv in level1
nonconv without canonical iNKT or MAIT & gdT_gated --> gdT
```{r}
iNKT = read.table("nonconv/inktInDec24Freeze.csv", header = T, sep = ",")[,1]
so$iNKT2 = colnames(so) %in% iNKT
MAIT = read.table("nonconv/maitInDec24Freeze.csv", header = T, sep = ",")[,1]
so$MAIT2 = colnames(so) %in% MAIT

so$level1[so$iNKT2 & so$annotation_level1 %in% c("CD4","CD8AB","Treg","abT_CD8aa","abT_DN","abT_DP","nonconv")] = "nonconv"
so$level1[so$MAIT2 & so$annotation_level1 %in% c("CD4","CD8AB","Treg","abT_CD8aa","abT_DN","abT_DP","nonconv")] = "nonconv"
so$level1[so$annotation_level1 == "nonconv" & so$gated_pop == "gdT_gated" & !so$MAIT2 & !so$iNKT2] = "gdT"
table(so$level1, so$annotation_level1)

```

3. cleaning CD4
CD4_cl8 --> prolif
CD4 in annotation_level1 & CD8AB_gated --> CD8AB
CD4 in annotation_level1 & CD8AA_gated --> CD8AA
CD4 in annotation_level1 & gdT_gated --> gdT
```{r}
so$level1[so$annotation_level2 %in% c("CD4_cl8")] = "prolif"
so$level1[so$annotation_level1 == "CD4" & so$gated_pop == "CD8AB_gated"] = "CD8AB"
so$level1[so$annotation_level1 == "CD4" & so$gated_pop == "CD8AA_gated"] = "abT_CD8aa"
so$level1[so$annotation_level1 == "CD4" & so$gated_pop == "gdT_gated"] = "gdT"

table(so$level1, so$annotation_level1)
```

4. cleaning CD8AB
CD8AB_cl19 --> prolif
CD8AB in annotation_level1 & CD4_gated --> CD4
CD8AB in annotation_level1 & CD8AA_gated --> CD8AA
CD8AB in annotation_level1 & gdT_gated --> gdT
```{r}
so$level1[so$annotation_level2 %in% c("CD8AB_cl19")] = "prolif"
so$level1[so$annotation_level1 == "CD8AB" & so$gated_pop == "CD4_gated"] = "CD4"
so$level1[so$annotation_level1 == "CD8AB" & so$gated_pop == "CD8AA_gated"] = "abT_CD8aa"
so$level1[so$annotation_level1 == "CD8AB" & so$gated_pop == "gdT_gated"] = "gdT"

table(so$level1, so$annotation_level1)
```

5. cleaning treg
treg in annotation_level1 & gdT_gated --> gdT
```{r}
so$level1[so$annotation_level2 %in% c("treg_cl5")] = "prolif"
so$level1[so$annotation_level1 == "treg" & so$gated_pop == "gdT_gated"] = "gdT"
table(so$level1, so$annotation_level1)
```

6. cleaning gdT
gdT in annotation_level1 & CD4_gated --> CD4
gdT in annotation_level1 & CD8AA_gated --> CD8AA
gdT in annotation_level1 & CD8AB_gated --> CD8AB
so$level1 == "gdT" & so[["ADT"]]@data["TCRB",] > 1 & so[["ADT"]]@data["TCRGD",] < 1 & !so$gated_pop %in% c("CD4_gated", "CD8AA_gated", "CD8AB_gated") --> abT_DN
```{r}
so$level1[so$annotation_level1 == "gdT" & so$gated_pop == "CD4_gated"] = "CD4"
so$level1[so$annotation_level1 == "gdT" & so$gated_pop == "CD8AA_gated"] = "abT_CD8aa"
so$level1[so$annotation_level1 == "gdT" & so$gated_pop == "CD8AB_gated"] = "CD8AB"


FeatureScatter(so[,so$level1 == "gdT"], feature1 = "TCRB", feature2 = "TCRGD", group.by = "gated_pop", split.by = NULL, ncol = 2, pt.size = 2, raster = T, plot.cor = F) + geom_vline(aes(xintercept = 1), linetype="dashed", color = "brown") 
so$level1[so$level1 == "gdT" & so[["ADT"]]@data["TCRB",] > 1 & so[["ADT"]]@data["TCRGD",] < 1 & !so$gated_pop %in% c("CD4_gated", "CD8AA_gated", "CD8AB_gated")] = "abT_DN"

table(so$level1, so$annotation_level1)

```

7. cleaning CD8AA
abT_CD8aa in annotation_level1 & CD4_gated --> CD4
abT_CD8aa in annotation_level1 & gdT_gated --> gdT
abT_CD8aa in annotation_level1 & CD8AB_gated --> CD8AB
```{r}
so$level1[so$annotation_level1 == "abT_CD8aa" & so$gated_pop == "CD4_gated"] = "CD4"
so$level1[so$annotation_level1 == "abT_CD8aa" & so$gated_pop == "gdT_gated"] = "gdT"
so$level1[so$annotation_level1 == "abT_CD8aa" & so$gated_pop == "CD8AB_gated"] = "CD8AB"

table(so$level1, so$annotation_level1)
```

8. cleaning abT_DN
```{r}

so$level1[so$annotation_level1 == "abT_DN" & so$gated_pop == "CD4_gated"] = "CD4"
so$level1[so$annotation_level1 == "abT_DN" & so$gated_pop == "CD8AB_gated"] = "CD8AB"
so$level1[so$annotation_level1 == "abT_DN" & so$gated_pop == "CD8AA_gated"] = "abT_CD8aa"
so$level1[so$annotation_level1 == "abT_DN" & so$gated_pop == "gdT_gated"] = "gdT"

table(so$level1, so$annotation_level1)

```

9. cleaning abT_DP
```{r}
so$level1[so$annotation_level1 == "abT_DP" & so$gated_pop == "CD4_gated"] = "CD4"
so$level1[so$annotation_level1 == "abT_DP" & so$gated_pop == "CD8AB_gated"] = "CD8AB"
so$level1[so$annotation_level1 == "abT_DP" & so$gated_pop == "CD8AA_gated"] = "abT_CD8aa"
so$level1[so$annotation_level1 == "abT_DP" & so$gated_pop == "gdT_gated"] = "gdT"

table(so$level1, so$annotation_level1)
table(so$level1, so$organ == "thymus")

```

10. cleaning preT
bring back preT that are not in the thymus
```{r}
##some plots before
so$is_thymus = as.character(so$organ %in% "thymus")

DimPlot(so, reduction = "mde2_totalvi_20241006", group.by = "is_thymus", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal)
DimPlot(so, reduction = "mde2_totalvi_20241006", group.by = "is_thymus", split = "is_thymus", raster = T, raster.dpi = c(512,512), label = F) + scale_color_manual(values = c("blue", "red"))

pdf("gating_Leon_20241116/FeatureScatter_preT_nonthymus_annotation.pdf", width = 7,7, useDingbats = F)
FeatureScatter(so[,so$annotation_level1 == "preT" & so$organ != "thymus"], feature1 = "TCRB", feature2 = "TCRGD", group.by = "gated_pop", split.by = NULL, ncol = 2, pt.size = 2, raster = T, plot.cor = F) 
FeatureScatter(so[,so$annotation_level1 == "preT" & so$organ != "thymus"], feature1 = "CD4", feature2 = "CD8B", group.by = "gated_pop", split.by = NULL, ncol = 2, pt.size = 2, raster = T, plot.cor = F) 
FeatureScatter(so[,so$annotation_level1 == "preT" & so$organ != "thymus"], feature1 = "CD8B", feature2 = "CD8A", group.by = "gated_pop", split.by = NULL, ncol = 2, pt.size = 2, raster = T, plot.cor = F) 

FeatureScatter(so[,so$annotation_level1 == "preT" & so$organ != "thymus" & !is.na(so$gated_pop)], feature1 = "TCRB", feature2 = "TCRGD", group.by = "gated_pop", split.by = NULL, ncol = 2, pt.size = 2, raster = T, plot.cor = F) 
FeatureScatter(so[,so$annotation_level1 == "preT" & so$organ != "thymus" & !is.na(so$gated_pop)], feature1 = "CD4", feature2 = "CD8B", group.by = "gated_pop", split.by = NULL, ncol = 2, pt.size = 2, raster = T, plot.cor = F) 
FeatureScatter(so[,so$annotation_level1 == "preT" & so$organ != "thymus" & !is.na(so$gated_pop)], feature1 = "CD8B", feature2 = "CD8A", group.by = "gated_pop", split.by = NULL, ncol = 2, pt.size = 2, raster = T, plot.cor = F) 

FeatureScatter(so[,so$annotation_level1 == "preT" & so$organ != "thymus" & is.na(so$gated_pop)], feature1 = "TCRB", feature2 = "TCRGD", group.by = "annotation_level1", split.by = NULL, ncol = 2, pt.size = 2, raster = T, plot.cor = F) #+ geom_vline(aes(xintercept = 1
FeatureScatter(so[,so$annotation_level1 == "preT" & so$organ != "thymus" & is.na(so$gated_pop)], feature1 = "CD4", feature2 = "CD8B", group.by = "annotation_level1", split.by = NULL, ncol = 2, pt.size = 2, raster = T, plot.cor = F) #+ geom_vline(aes(xintercept = 1
FeatureScatter(so[,so$annotation_level1 == "preT" & so$organ != "thymus" & is.na(so$gated_pop)], feature1 = "CD8B", feature2 = "CD8A", group.by = "annotation_level1", split.by = NULL, ncol = 2, pt.size = 2, raster = T, plot.cor = F) #+ geom_vline(aes(xintercept = 1

dev.off()


#Reassign
so$level1[so$annotation_level1 == "thymocyte" & so$organ != "thymus" & so$MAIT2] = "nonconv"
so$level1[so$annotation_level1 == "thymocyte" & so$organ != "thymus" & so$iNKT2] = "nonconv"
so$level1[so$annotation_level1 == "thymocyte" & so$organ != "thymus" & so$gated_pop == "CD4_gated"] = "CD4"
so$level1[so$annotation_level1 == "thymocyte" & so$organ != "thymus" & so$gated_pop == "CD8AB_gated"] = "CD8AB"
so$level1[so$annotation_level1 == "thymocyte" & so$organ != "thymus" & so$gated_pop == "CD8AA_gated"] = "abT_CD8aa"
so$level1[so$annotation_level1 == "thymocyte" & so$organ != "thymus" & so$gated_pop == "gdT_gated"] = "gdT"

#Manual gating of cells not gated before
so$level1[so$annotation_level1 == "thymocyte" & so$organ != "thymus" &  is.na(so$gated_pop) & so[["ADT"]]@data["TCRGD",] > 2 & so[["ADT"]]@data["TCRB",] < 0.6 ] = "gdT"
so$level1[so$annotation_level1 == "thymocyte" & so$organ != "thymus" &  is.na(so$gated_pop) & so[["ADT"]]@data["TCRGD",] < 2 & so[["ADT"]]@data["TCRB",] > 1 & so[["ADT"]]@data["CD4",] > 1.5 & so[["ADT"]]@data["CD8B",] < 2  ] = "CD4"
so$level1[so$annotation_level1 == "thymocyte" & so$organ != "thymus" &  is.na(so$gated_pop) & so[["ADT"]]@data["TCRGD",] < 2 & so[["ADT"]]@data["TCRB",] > 1 & so[["ADT"]]@data["CD4",] < 1 & so[["ADT"]]@data["CD8B",] > 2  ] = "CD8AB"
so$level1[so$annotation_level1 == "thymocyte" & so$organ != "thymus" &  is.na(so$gated_pop) & so[["ADT"]]@data["TCRGD",] < 2 & so[["ADT"]]@data["TCRB",] > 1 & so[["ADT"]]@data["CD8A",] > 2 & so[["ADT"]]@data["CD8B",] < 2  ] = "abT_CD8aa"
so$level1[so$annotation_level1 == "thymocyte" & so$organ != "thymus" &  is.na(so$gated_pop) & so[["ADT"]]@data["TCRGD",] < 2 & so[["ADT"]]@data["CD4",] > 1.5 & so[["ADT"]]@data["CD8B",] > 2  ] = "abT_DP"
so$level1[so$annotation_level1 == "thymocyte" & so$organ != "thymus" &  is.na(so$gated_pop) & so[["ADT"]]@data["TCRGD",] < 2 & so[["ADT"]]@data["CD4",] < 1.5 & so[["ADT"]]@data["CD8B",] < 2  ] = "abT_DN"

pdf("gating_Leon_20241116/FeatureScatter_preT_nonthymus_manual_gate.pdf", width = 7,7, useDingbats = F)
FeatureScatter(so[,so$annotation_level1 == "thymocyte" & so$organ != "thymus" &  is.na(so$gated_pop) ], feature1 = "TCRB", feature2 = "TCRGD", group.by = "level1", split.by = NULL, ncol = 2, pt.size = 2, raster = F, plot.cor = F) + scale_color_manual(values = mypal_level1)
FeatureScatter(so[,so$annotation_level1 == "thymocyte" & so$organ != "thymus" &  is.na(so$gated_pop) ], feature1 = "CD4", feature2 = "CD8B", group.by = "level1", split.by = NULL, ncol = 2, pt.size = 2, raster = F, plot.cor = F) + scale_color_manual(values = mypal_level1)
FeatureScatter(so[,so$annotation_level1 == "thymocyte" & so$organ != "thymus" &  is.na(so$gated_pop) ], feature1 = "CD8B", feature2 = "CD8A", group.by = "level1", split.by = NULL, ncol = 2, pt.size = 2, raster = F, plot.cor = F) + scale_color_manual(values = mypal_level1)
dev.off()

table(so$level1, so$annotation_level1)

pdf("gating_Leon_20241116/MDE_preT_nonthymus_newannotation.pdf", width = 7,7, useDingbats = F)
MyDimPlotHighlight(seurat_object = so, umap_to_plot = "mde2_totalvi_20241006", cells_to_highlight = colnames(so)[so$annotation_level1 == "thymocyte" & so$organ != "thymus"], highlight_column_name = "level1", labelclusters = T,
                       title = "", pixels = c(512, 512), highlight_size = 1, 
                       highlight_alpha = 1, mycols = mypal_level1,print_plot1 = T, 
                       print_plot2 = T)
dev.off()

pdf("gating_Leon_20241116/FeatureScatter_preT_nonthymus_newannotation.pdf", width = 7,7, useDingbats = F)
FeatureScatter(so[,so$annotation_level1 == "thymocyte" & so$organ != "thymus" ], feature1 = "TCRB", feature2 = "TCRGD", group.by = "level1", split.by = NULL, ncol = 2, pt.size = 2, raster = F, plot.cor = F) + scale_color_manual(values = mypal_level1)
FeatureScatter(so[,so$annotation_level1 == "thymocyte" & so$organ != "thymus" ], feature1 = "CD4", feature2 = "CD8B", group.by = "level1", split.by = NULL, ncol = 2, pt.size = 2, raster = F, plot.cor = F) + scale_color_manual(values = mypal_level1)
FeatureScatter(so[,so$annotation_level1 == "thymocyte" & so$organ != "thymus" ], feature1 = "CD8B", feature2 = "CD8A", group.by = "level1", split.by = NULL, ncol = 2, pt.size = 2, raster = F, plot.cor = F) + scale_color_manual(values = mypal_level1)
dev.off()

```


Final MDE and FeatureScatter
```{r}

pdf("MDE_annotation_20250107_level1.pdf", 10,10, useDingbats = F)
for (i in unique(so$level1)) {
    print(i)
    p = MyDimPlotHighlight(seurat_object = so, umap_to_plot = "mde2_totalvi_20241006", cells_to_highlight = colnames(so)[so$level1 == i], highlight_column_name = "level1", labelclusters = F,
                       title = i, pixels = c(512, 512), highlight_size = 0.5, 
                       highlight_alpha = 1, mycols = mypal_level1, print_plot1 = T, 
                       print_plot2 = T)
    print(p)
}
dev.off()



```

Observation 2: Discordances between annotation level 1 and CITEseq (CD4, CD8AB, CD8AA, TCRGD)
```{r}

pdf("FeatureScatter_annotation_20250107_level1.pdf", 20,20, useDingbats = F)
# for (i in levels(so$annotation_level1)) {
FeatureScatter(so, feature1 = "TCRB", feature2 = "TCRGD", group.by = "level1", split.by = "level1", ncol = 4, pt.size = 2, raster = T, plot.cor = F) + scale_color_manual(values = mypal_level1)
FeatureScatter(so, feature1 = "CD4", feature2 = "CD8B", group.by = "level1", split.by = "level1", ncol = 4, pt.size = 2, raster = T, plot.cor = F) + scale_color_manual(values = mypal_level1)
FeatureScatter(so, feature1 = "CD8B", feature2 = "CD8A", group.by = "level1", split.by = "level1", ncol = 4, pt.size = 2, raster = T, plot.cor = F) + scale_color_manual(values = mypal_level1)
# FeatureScatter(so[,so$gdTwithTCRab == T], feature1 = "CD4", feature2 = "CD8B", group.by = "gdT_CITEseq", split.by = NULL, ncol = 2, pt.size = 2, raster = F, plot.cor = F) 
# }
dev.off()

```




