---
title: "clustering_robustness"
output: html_document
date: "2026-01-13"
editor_options: 
  chunk_output_type: console
---

**Initialize: load libraries and custom functions**

```{bash}
sinteractive --exclusive --time=12:00:00 --account=pi-zemmour --partition=jfkfloor2 --qos=jfkfloor2 --nodes=1 
sinteractive --exclusive --time=12:00:00 --account=pi-zemmour --partition=beagle3-bigmem --nodes=1 
sinteractive --exclusive --time=12:00:00 --account=pi-zemmour --partition=zemmour-hm --qos=zemmour --nodes=1 
sinteractive --time=03:00:00 --account=pi-zemmour --partition=caslake --nodes=1 --mem=96GB

module load python/anaconda-2022.05 
module load hdf5/1.12.0 #for BPCells but also others
module load openblas/0.3.13
source activate /project/zemmour/david/envs/scvi_20240315
#source activate /project/zemmour/david/envs/r443
module load openblas/0.3.13

cd /project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96
R
```


```{r}
# setwd("~/google_drive_uchicago/ImmgenT/analysis/data_integration/IGT1_96/")
libs = c("ZemmourLib","forcats", "gridExtra","patchwork", "Matrix", "Seurat", "ggplot2", "dplyr", "reshape2", "ggrastr","scattermore", "RColorBrewer", "pals", "scales", "pheatmap", "rafalib") 
sapply(libs, function(x) suppressMessages(suppressWarnings(library(x, character.only = TRUE, quietly = T, warn.conflicts  = F))))
#source("/project/jfkfloor2/zemmourlab/david/immgent/immgent_integration_git/custom_functions_david.R")

ColorRamp = rev(colorRampPalette(brewer.pal(n = 7,name = "RdYlBu"))(100))
ColorRamp = rev(colorRampPalette(c('red','white','blue'))(20))
ColorRamp = rev(rainbow(10, end = 4/6))
ColorRamp = rev(viridis(100))
ColorRamp = brewer.pal(9, "Blues")

#mycol = c("black","blue","red","darkgreen", "orange","purple", "cyan", "greenyellow",    "salmon", "magenta","pink", "tan", "brown") # c("magenta", "red",  "darkgreen", "cyan", "blue", "blue", "blue", "black", "blue", "blue", "blue", "blue", "orange")

mypal_level2 = immgent_colors$level2
mypal_level2["not classified"] = "black"
# names(mypal_level2) = gsub("nonconv", replacement = "Tz",names(mypal_level2) )
# mypal_level2 = immgent_colors$level2_option2
# mypal_level2["not classified"] = "black"
mypal_level1 = immgent_colors$level1
mypal_level1["not classified"] = "black"
# names(mypal_level1) = gsub("nonconv", replacement = "Tz",names(mypal_level1) )
mypal_organ = immgent_colors$organ_simplified
mypal_level2group = c("resting" = "blue", "activated" = "red", "miniverse" = "darkgreen", "proliferating" = "black", "other" = "grey", "preT" = "grey" )

#Large Color palette
library("pals")
n = 70
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
mypal1 = unique(unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals))))
mypal1 = mypal1[-4]
mypal = c(glasbey(), polychrome(), mypal1)
names(mypal) = NULL
#pal.bands(mypal, main="Colormap suggestions")
# mypal_organ = setNames(mypal, unique(so$organ_simplified))
# mypal_organ = mypal_organ[!is.na(names(mypal_organ))]
# mypal_level1 = setNames(mypal, unique(so$annotation_level1))
# mypal_level1 = mypal_level1[!is.na(names(mypal_level1))]
# mypal_level2 = setNames(mypal, unique(so_orig$annotation_level2))
# mypal_level2 = mypal_level2[!is.na(names(mypal_level2))]
# mypal_level2["CD4_cl18"] = "blue"
mypal_igt = setNames(mypal, unique(so_orig$IGT))
mypal_igt = mypal_igt[!is.na(names(mypal_igt))]

```

```{r}
# fig_dir = "~/google_drive_uchicago/ImmgenT/20250109_freeze/paper_cosmology/figures/bits"
# fig_dir = "/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/paper_cosmology/figures/bits"

out_dir = "~/google_drive_uchicago/ImmgenT/analysis/data_integration/IGT1_96/misc_plots" 
out_dir = "/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/misc_plots"

```


**Most up-to-date data**
on Midway
```{r}
# so_orig = readRDS("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/igt1_96_withtotalvi20250513_clean.Rds") 
# so_orig = readRDS("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/igt1_96_withtotalvi20250710_clean.Rds") 
# so_orig$annotation_level2[so_orig$annotation_level2 == "Treg_cl7"] = "Treg_cl1"
# so_orig$annotation_level1 = factor(so_orig$annotation_level1, levels = c("CD8", "CD4", "Treg", "gdT", "CD8aa", "nonconv", "DN", "DP", "thymocyte"))

so_orig = readRDS("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/igt1_96_withtotalvi20260107_clean.Rds") 

so = so_orig[,so_orig@meta.data %>% filter(organ_simplified != "thymus") %>% row.names()]
so = NormalizeData(so, assay = "RNA", normalization.method = "LogNormalize")


```

on Desktop
```{r}
setwd("~/google_drive_uchicago/ImmgenT/analysis/data_integration/IGT1_96/")
# so_orig = readRDS("igt1_96_withtotalvi20250710_clean_ADTonly.Rds")
# so_orig$annotation_level2[so_orig$annotation_level2 == "Treg_cl7"] = "Treg_cl1"
# so_orig$annotation_level1 = factor(so_orig$annotation_level1, levels = c("CD8", "CD4", "Treg", "gdT", "CD8aa", "nonconv", "DN", "DP", "thymocyte"))
so_orig = readRDS("igt1_96_withtotalvi20260107_clean_ADTonly.Rds")

so = so_orig[,so_orig@meta.data %>% filter(organ_simplified != "thymus") %>% row.names()]
# so = NormalizeData(so, assay = "RNA", normalization.method = "LogNormalize")
# so = NormalizeData(so, assay = "ADT", normalization.method = "LogNormalize")

```

*S4A*
Clusters by IGT
```{r}
library(dplyr)
library(rlang)

CompositionOfEachCluster <- function(meta = so@meta.data, cluster_col = "annotation_level2", sample_col  = "IGT") {
    clust  <- sym(cluster_col)
    sample <- sym(sample_col)
    
    # 1) counts per cluster × sample
    comp <- meta %>%
        filter(!is.na(!!clust), !is.na(!!sample)) %>%
        count(!!clust, !!sample, name = "n") %>%
        # 2) normalize within sample (like your tmp2)
        group_by(!!sample) %>%
        mutate(p_in_sample = n / sum(n)) %>%
        ungroup() %>%
        # 3) normalize within cluster (like your tmp3)
        group_by(!!clust) %>%
        mutate(p_weighted = p_in_sample / sum(p_in_sample)) %>%
        ungroup()
    
    # 4) total n per cluster + label "cluster (n=XX)"
    cluster_sizes <- comp %>%
        group_by(!!clust) %>%
        summarise(n_total = sum(n), .groups = "drop") %>%
        mutate(
            clust_chr = as.character(!!clust),
            label = paste0(clust_chr, " (n=", n_total, ")")
        ) %>%
        select(-clust_chr)
    
    # 5) join labels back & make label a factor in a stable order
    comp_labeled <- comp %>%
        left_join(cluster_sizes, by = rlang::as_name(clust)) %>%
        mutate(label = factor(label, levels = cluster_sizes$label))
    
    return(comp_labeled)
}

comp_df = CompositionOfEachCluster(so@meta.data,
                                    cluster_col = "annotation_level2",
                                    sample_col  = "IGT")

head(comp_df)
comp_df$annotation_level1 = so_orig$annotation_level1[match(comp_df$annotation_level2, so_orig$annotation_level2)]
comp_df$annotation_level1 = factor(comp_df$annotation_level1, levels = levels(so$annotation_level1))

p <- ggplot(comp_df) +
    geom_bar(
        aes(x = label, y = p_weighted, fill = IGT),
        stat = "identity"
    ) +
    scale_fill_manual(values = mypal_igt) +
    theme_bw() +
    ggtitle("IGT composition in each cluster") +
    theme(
        axis.text.x = element_text(angle = 75, vjust = 0.5, size = 10),
        axis.title.x = element_blank()
    )


pdf(sprintf("%s/Barplot_Level2ByIGT.pdf", out_dir), 15,5, useDingbats = F)
p + facet_grid(
        . ~ annotation_level1,
        scales = "free_x",
        space  = "free_x"
    ) +
    labs(x = NULL, y = NULL) +
    theme_minimal(base_size = 11) +
    theme(
        axis.text.x = element_text(angle = 60, hjust = 1, vjust = 1, size = 12),
        axis.text.y = element_text(size = 12),
        panel.grid = element_blank(),
        strip.text = element_text(face = "bold")
    ) + NoGrid() + NoLegend()
dev.off()
```

*S4B*
Clusters distribution by IGT only in spleen standard --> see later in the code: "Boxplot_Level2ByIGT_spleencontrols.pdf"
```{r}
```

*S3B: Conservation of level2 in each dataset - RNA space*
classic silhouette, filter cells in each IGT with at least 10 cell per level2. 
```{r}
library(cluster)
so = so_orig[,so_orig@meta.data %>% filter(organ_simplified != "thymus") %>% row.names()]
# so = NormalizeData(so, assay = "RNA", normalization.method = "LogNormalize")
so$level2_int = as.integer(so[["annotation_level2"]][,1])
DefaultAssay(so) = "RNA"

dict_l2 = so@meta.data %>% count(annotation_level2, level2_int)

sil_summary_list = list()
for (i in unique(so$IGT)) {
    print(i) 
    tmp = so[,so$IGT == i]
    subcl = tmp@meta.data %>% count(annotation_level2) %>% filter(n>10) %>% pull(annotation_level2) %>% as.character()
    if (length(subcl) > 1) {
        tmp = tmp[,tmp$annotation_level2 %in% subcl]
        tmp = tmp %>% 
            NormalizeData(normalization.method = "LogNormalize", verbose = F) %>%
            FindVariableFeatures(selection.method = "vst", nfeatures = 2000, verbose = F) %>%
            ScaleData(features = VariableFeatures(.), verbose = F) %>%    
            RunPCA(features = VariableFeatures(.), npcs = 30, verbose = F) #%>% FindNeighbors(dims = 1:30, verbose = F) %>% RunUMAP(dims = 1:30, seed.use = 123, verbose = T)
        # DimPlot(tmp, reduction = "umap", group.by = "annotation_level2", label = T) + scale_color_manual(values = mypal_level2) + NoLegend()
        
        # message("Calculating silhouette post integration...")
        reduc.dist = dist(tmp[["pca"]]@cell.embeddings)
        cl = as.integer(tmp[["level2_int"]][,1])
        sil_post = silhouette(cl, dist = reduc.dist)
        sil_summary = as.data.frame(sil_post) %>% 
            group_by(cluster) %>% 
            summarise(mean_sil_width = mean(sil_width),
                      median_sil_width = median(sil_width),
                      total = n(),  # Count total observations in each cluster
                      positive_count = sum(sil_width > 0),  # Count negative silhouette widths
                      percentage_positive = (positive_count / total) * 100)     %>%  # Calculate percentage) 
            as.data.frame()
        # sil_summary$cluster == dict_l2$level2_int[match(sil_summary$cluster, dict_l2$level2_int)]
        sil_summary$annotation_level2 = dict_l2$annotation_level2[match(sil_summary$cluster, dict_l2$level2_int)]
        sil_summary$IGT = i
        sil_summary_list[[i]] = sil_summary 
    }
    
}
saveRDS(sil_summary_list, file = "RobustnessLevel2RNASpace_Silhouette_IGT.Rds")

sil_summary_merged = Reduce(rbind, sil_summary_list)

dict_l1 = so@meta.data %>% count(annotation_level1, annotation_level2)
sil_summary_merged$annotation_level1 = dict_l1$annotation_level1[match(sil_summary_merged$annotation_level2, dict_l1$annotation_level2)]
    

sil_summary_merged$annotation_level2 <- factor(
    sil_summary_merged$annotation_level2,
    levels = levels(so$annotation_level2)
)

p_list = list()
for(j in c("mean_sil_width", "median_sil_width", "percentage_positive")) {
    p_list[[j]] = ggplot(sil_summary_merged) + #%>% filter(total > 10
    aes(x = annotation_level2, y = !!sym(j)) +
    geom_jitter(aes(color = IGT),
                width = 0.15, height = 0,
                size = 1.8, alpha = 0.9) +
    geom_boxplot(outlier.shape = NA,
                 fill = "grey90",
                 colour = "black",
                 width = 0.6) +
    scale_color_manual(values = mypal_igt) +
    theme_bw() +
    # labs(
    #     x = "Cluster (annotation_level2)",
    #     y = "Median silhouette width",
    #     color = "IGT"
    # ) +
    facet_grid(
        . ~ annotation_level1,
        scales = "free_x",
        space  = "free_x"
    ) +
    labs(x = NULL, y = NULL) +
    ggtitle(j) +
    theme_minimal(base_size = 11) +
    theme(
        axis.text.x = element_text(angle = 60, hjust = 1, vjust = 1, size = 12),
        axis.text.y = element_text(size = 12),
        panel.grid = element_blank(),
        strip.text = element_text(face = "bold")
    )
}

pdf(sprintf("%s/Boxplot_Level2RNASpace.pdf", out_dir), 25,6, useDingbats = F)
print(p_list[[1]])
print(p_list[[1]] + NoLegend())
print(p_list[[2]] + NoLegend())
print(p_list[[3]] + NoLegend())
dev.off()
# pdf(sprintf("%s/Boxplot_Level2RNASpace.pdf", out_dir), 25,6, useDingbats = F)

```


*S3C: Conservation of level2 in each dataset - Protein/ADT space*
classic silhouette, filter cells in each IGT with at least 10 cell per level2. 
```{r}
library(cluster)
so = so_orig[,so_orig@meta.data %>% filter(organ_simplified != "thymus", cite_seq == T) %>% row.names()]
so$level2_int = as.integer(so[["annotation_level2"]][,1])
DefaultAssay(so) = "ADT"

dict_l2 = so@meta.data %>% count(annotation_level2, level2_int)
#Select only good ADT markers
markers = so[["ADT"]]@meta.features %>% filter(IGT1only == F & classification %in% c("good", "intermediate")) %>% row.names()

sil_summary_list = list()
for (i in unique(so$IGT)) {
    print(i)
    tmp = so[markers,so$IGT == i]
    subcl = tmp@meta.data %>% count(annotation_level2) %>% filter(n>10) %>% pull(annotation_level2) %>% as.character()
    if (length(subcl) > 1) {
        tmp = tmp[,tmp$annotation_level2 %in% subcl]
        tmp = tmp %>% 
            NormalizeData(normalization.method = "CLR", verbose = F) %>%
            FindVariableFeatures(selection.method = "vst", nfeatures = 50, verbose = F) %>%
            ScaleData(features = VariableFeatures(.), verbose = F) %>%    
            RunPCA(features = VariableFeatures(.), npcs = 10, verbose = F) #%>% FindNeighbors(dims = 1:10, verbose = F) %>% RunUMAP(dims = 1:10, seed.use = 123, verbose = T)
        # DimPlot(tmp, reduction = "umap", group.by = "annotation_level2", label = T) + scale_color_manual(values = mypal_level2) + NoLegend()
        
        # message("Calculating silhouette post integration...")
        reduc.dist = dist(tmp[["pca"]]@cell.embeddings)
        cl = as.integer(tmp[["level2_int"]][,1])
        sil_post = silhouette(cl, dist = reduc.dist)
        sil_summary = as.data.frame(sil_post) %>% 
            group_by(cluster) %>% 
            summarise(mean_sil_width = mean(sil_width),
                      median_sil_width = median(sil_width),
                      total = n(),  # Count total observations in each cluster
                      positive_count = sum(sil_width > 0),  # Count negative silhouette widths
                      percentage_positive = (positive_count / total) * 100)     %>%  # Calculate percentage) 
            as.data.frame()
        # sil_summary$cluster == dict_l2$level2_int[match(sil_summary$cluster, dict_l2$level2_int)]
        sil_summary$annotation_level2 = dict_l2$annotation_level2[match(sil_summary$cluster, dict_l2$level2_int)]
        sil_summary$IGT = i
        sil_summary_list[[i]] = sil_summary 
    }
    
}


sil_summary_merged = Reduce(rbind, sil_summary_list)

dict_l1 = so@meta.data %>% count(annotation_level1, annotation_level2)
sil_summary_merged$annotation_level1 = dict_l1$annotation_level1[match(sil_summary_merged$annotation_level2, dict_l1$annotation_level2)]
    

sil_summary_merged$annotation_level2 <- factor(
    sil_summary_merged$annotation_level2,
    levels = levels(so$annotation_level2)
)

p_list = list()
for(j in c("mean_sil_width", "median_sil_width", "percentage_positive")) {
    p_list[[j]] = ggplot(sil_summary_merged) + #%>% filter(total > 10
    aes(x = annotation_level2, y = !!sym(j)) +
    geom_jitter(aes(color = IGT),
                width = 0.15, height = 0,
                size = 1.8, alpha = 0.9) +
    geom_boxplot(outlier.shape = NA,
                 fill = "grey90",
                 colour = "black",
                 width = 0.6) +
    scale_color_manual(values = mypal_igt) +
    theme_bw() +
    # labs(
    #     x = "Cluster (annotation_level2)",
    #     y = "Median silhouette width",
    #     color = "IGT"
    # ) +
    facet_grid(
        . ~ annotation_level1,
        scales = "free_x",
        space  = "free_x"
    ) +
    labs(x = NULL, y = NULL) +
    ggtitle(j) +
    theme_minimal(base_size = 11) +
    theme(
        axis.text.x = element_text(angle = 60, hjust = 1, vjust = 1, size = 12),
        axis.text.y = element_text(size = 12),
        panel.grid = element_blank(),
        strip.text = element_text(face = "bold")
    )
}

pdf(sprintf("%s/Boxplot_Level2ProteinSpace.pdf", out_dir), 25,6, useDingbats = F)
print(p_list[[1]])
print(p_list[[1]] + NoLegend())
print(p_list[[2]])
print(p_list[[3]])
dev.off()
# pdf(sprintf("%s/Boxplot_Level2RNASpace.pdf", out_dir), 25,6, useDingbats = F)

```

colon IGT20, IGT24 and IGT27
```{r}
igt = readRDS('/Users/david/Library/CloudStorage/GoogleDrive-david.puyraimond@gmail.com/My Drive/ImmgenT/exp_id/20220812_exp_id_13_ColonSI/IGT20/dataset_clean.Rds')
tmp = so_orig
tmp$RNA_clusters = NA
tmp$RNA_clusters = as.character(igt$RNA_clusters[match(colnames(tmp), colnames(igt))])
pdf(sprintf("%s/IGT20_MDE_immgenTclusters.pdf", out_dir), 5,5, useDingbats = F)
mypal_batch = setNames(mypal, unique(igt$RNA_clusters))
mypal_batch = mypal_batch[!is.na(names(mypal_batch))]
igt@meta.data$annotation_level2 = so_orig$annotation_level2[match(colnames(igt), colnames(so_orig))]
igt@meta.data$annotation_level1 = so_orig$annotation_level1[match(colnames(igt), colnames(so_orig))]
DimPlot(igt[,grepl("Colon", igt$sample_name)], reduction = "umap_rna", group.by = "RNA_clusters") + scale_color_manual(values = mypal_batch) + NoGrid()
DimPlot(igt[,grepl("Colon", igt$sample_name)], reduction = "umap_rna", group.by = "annotation_level1") + scale_color_manual(values = mypal_level1) + NoGrid()
DimPlot(igt[,grepl("Colon", igt$sample_name)], reduction = "umap_rna", group.by = "annotation_level2") + scale_color_manual(values = mypal_level2) + NoGrid()
igt_colon = igt[,grepl("Colon", igt$sample_name)]
big_clusters = names(which(table(igt_colon$annotation_level2) >= 50))
p = DimPlot(igt_colon, reduction = "umap_rna", group.by = "annotation_level2", label = FALSE) + scale_color_manual(values = mypal_level2) + NoGrid()
p + NoLegend()
LabelClusters(p, id = "annotation_level2", clusters = big_clusters, size = 3, repel = TRUE, box = T) + NoLegend()

DimPlot(igt[,grepl("Colon", igt$sample_name)], reduction = "umap_adt", group.by = "RNA_clusters") + scale_color_manual(values = mypal_batch) + NoGrid()
DimPlot(igt[,grepl("Colon", igt$sample_name)], reduction = "umap_adt", group.by = "annotation_level1") + scale_color_manual(values = mypal_level1) + NoGrid()
DimPlot(igt[,grepl("Colon", igt$sample_name)], reduction = "umap_adt", group.by = "annotation_level2") + scale_color_manual(values = mypal_level2) + NoGrid()
igt_colon = igt[,grepl("Colon", igt$sample_name)]
big_clusters = names(which(table(igt_colon$annotation_level2) >= 50))
p = DimPlot(igt_colon, reduction = "umap_adt", group.by = "annotation_level2", label = FALSE) + scale_color_manual(values = mypal_level2) + NoGrid()
p + NoLegend()
LabelClusters(p, id = "annotation_level2", clusters = big_clusters, size = 3, repel = TRUE, box = T) + NoLegend()

# tmp2 = data.frame(dim1 = tmp[["mde2_totalvi_20241006"]]@cell.embeddings[, 1], dim2 = tmp[["mde2_totalvi_20241006"]]@cell.embeddings[, 2])
# bkrg = ggplot(tmp2) + scattermore::geom_scattermore(aes(dim1, dim2), color = "grey50", pointsize = 0.5, alpha = 0.5, pixels = c(512,512))
p = DimPlot(tmp[,colnames(igt_colon)], reduction = "mde2_totalvi_20241006", group.by = "annotation_level2", label = FALSE) + scale_color_manual(values = mypal_level2) + NoGrid()
p + NoLegend()
LabelClusters(p, id = "annotation_level2", clusters = big_clusters, size = 3, repel = TRUE, box = T) + NoLegend()

MyDimPlotHighlight(tmp, umap_to_plot = "mde2_totalvi_20241006", cells_to_highlight = colnames(igt)[grepl("Colon", igt$sample_name)], highlight_column_name = "annotation_level2",mycols = mypal_level2, print_plot1 = F, print_plot2 = T, labelclusters = T )
dev.off()

igt = readRDS('/Users/david/Library/CloudStorage/GoogleDrive-david.puyraimond@gmail.com/My Drive/ImmgenT/exp_id/20221019_exp_id_21_Iliev/IGT27/dataset_clean.Rds')
tmp = so_orig
tmp$RNA_clusters = NA
tmp$RNA_clusters = as.character(igt$RNA_clusters[match(colnames(tmp), colnames(igt))])
pdf(sprintf("%s/IGT27_MDE_immgenTclusters.pdf", out_dir), 5,5, useDingbats = F)
mypal_batch = setNames(mypal, unique(igt$RNA_clusters))
mypal_batch = mypal_batch[!is.na(names(mypal_batch))]
igt@meta.data$annotation_level2 = so_orig$annotation_level2[match(colnames(igt), colnames(so_orig))]
igt@meta.data$annotation_level1 = so_orig$annotation_level1[match(colnames(igt), colnames(so_orig))]
DimPlot(igt[,grepl("No_colon_M", igt$sample_name)], reduction = "umap_rna", group.by = "RNA_clusters") + scale_color_manual(values = mypal_batch) + NoGrid()
DimPlot(igt[,grepl("No_colon_M", igt$sample_name)], reduction = "umap_rna", group.by = "annotation_level1") + scale_color_manual(values = mypal_level1) + NoGrid()
DimPlot(igt[,grepl("No_colon_M", igt$sample_name)], reduction = "umap_rna", group.by = "annotation_level2") + scale_color_manual(values = mypal_level2) + NoGrid()
igt_colon = igt[,grepl("No_colon_M", igt$sample_name)]
big_clusters = names(which(table(igt_colon$annotation_level2) >= 10))
p = DimPlot(igt_colon, reduction = "umap_rna", group.by = "annotation_level2", label = FALSE) + scale_color_manual(values = mypal_level2) + NoGrid()
p + NoLegend()
LabelClusters(p, id = "annotation_level2", clusters = big_clusters, size = 3, repel = TRUE, box = T) + NoLegend()

DimPlot(igt[,grepl("No_colon_M", igt$sample_name)], reduction = "umap_adt", group.by = "RNA_clusters") + scale_color_manual(values = mypal_batch) + NoGrid()
DimPlot(igt[,grepl("No_colon_M", igt$sample_name)], reduction = "umap_adt", group.by = "annotation_level1") + scale_color_manual(values = mypal_level1) + NoGrid()
DimPlot(igt[,grepl("No_colon_M", igt$sample_name)], reduction = "umap_adt", group.by = "annotation_level2") + scale_color_manual(values = mypal_level2) + NoGrid()
igt_colon = igt[,grepl("No_colon_M", igt$sample_name)]
big_clusters = names(which(table(igt_colon$annotation_level2) >= 10))
p = DimPlot(igt_colon, reduction = "umap_adt", group.by = "annotation_level2", label = FALSE) + scale_color_manual(values = mypal_level2) + NoGrid()
p + NoLegend()
LabelClusters(p, id = "annotation_level2", clusters = big_clusters, size = 3, repel = TRUE, box = T) + NoLegend()

# tmp2 = data.frame(dim1 = tmp[["mde2_totalvi_20241006"]]@cell.embeddings[, 1], dim2 = tmp[["mde2_totalvi_20241006"]]@cell.embeddings[, 2])
# bkrg = ggplot(tmp2) + scattermore::geom_scattermore(aes(dim1, dim2), color = "grey50", pointsize = 0.5, alpha = 0.5, pixels = c(512,512))
p = DimPlot(tmp[,colnames(igt_colon)], reduction = "mde2_totalvi_20241006", group.by = "annotation_level2", label = FALSE) + scale_color_manual(values = mypal_level2) + NoGrid()
p + NoLegend()
LabelClusters(p, id = "annotation_level2", clusters = big_clusters, size = 3, repel = TRUE, box = T) + NoLegend()

MyDimPlotHighlight(tmp, umap_to_plot = "mde2_totalvi_20241006", cells_to_highlight = colnames(igt)[grepl("No_colon_M", igt$sample_name)], highlight_column_name = "annotation_level2",mycols = mypal_level2, print_plot1 = F, print_plot2 = T, labelclusters = T )
dev.off()

```

(silhouette_width where the distance of each cell to its cluster medoid is compared to the closest medioid of cluysters it doesn't belong to. Pretty much the same as normal silhouette, because similar definition... SO STICK TO THE SILHOUETTE!)
only focused on cells in an IGT with at least 10 cells in a level2 cluster
```{r}
silhouette_medoid_from_dist <- function(d, clusters) {
  # d: distance matrix (n x n) or 'dist' object
  # clusters: factor/character vector of cluster labels (length n)
  
  if (inherits(d, "dist")) {
    d <- as.matrix(d)
  } else {
    d <- as.matrix(d)
  }
  n <- nrow(d)
  if (n != ncol(d)) stop("d must be a square distance matrix")
  
  if (length(clusters) != n) stop("length(clusters) must match nrow(d)")
  clusters <- as.factor(clusters)
  k_levels <- levels(clusters)
  
  # 1. Find medoid index for each cluster
  medoids <- setNames(integer(length(k_levels)), k_levels)
  for (cl in k_levels) {
    idx <- which(clusters == cl)
    if (length(idx) == 1L) {
      medoids[cl] <- idx
    } else {
      # medoid = argmin_j sum_k d[j,k] within cluster
      subD <- d[idx, idx, drop = FALSE]
      tot <- rowSums(subD)
      medoids[cl] <- idx[which.min(tot)]
    }
  }
  
  a <- numeric(n)
  b <- numeric(n)
  s <- numeric(n)
  
  for (i in seq_len(n)) {
    ci <- clusters[i]
    # a(i): distance to own cluster medoid
    a[i] <- d[i, medoids[as.character(ci)]]
    
    # b(i): min distance to medoids of other clusters
    other_cls <- k_levels[k_levels != ci]
    if (length(other_cls) == 0L) {
      b[i] <- NA
      s[i] <- 0
      next
    }
    b[i] <- min(d[i, medoids[other_cls]])
    
    denom <- max(a[i], b[i])
    if (is.na(denom) || denom == 0) {
      s[i] <- 0
    } else {
      s[i] <- (b[i] - a[i]) / denom
    }
  }
  
  out <- data.frame(
    cluster = clusters,
    a = a,
    b = b,
    s = s,
    stringsAsFactors = FALSE
  )
  rownames(out) <- rownames(d)
  out
}


# set.seed(1)
# X <- rbind(
#   matrix(rnorm(50, mean = 0),  ncol = 2),
#   matrix(rnorm(50, mean = 3),  ncol = 2),
#   matrix(rnorm(50, mean = -3), ncol = 2)
# )
# colnames(X) <- c("x1", "x2")
# cl <- factor(rep(c("A", "B", "C"), each = 25))
# text(X[,1], X[,2], labels =  cl)
# sc <- silhouette_centroid(X, cl)
# head(sc)
# sc %>% group_by(cluster) %>% summarize(mean(s))


library(cluster)
so = so_orig[,so_orig@meta.data %>% filter(organ_simplified != "thymus", cite_seq == T) %>% row.names()]
so$level2_int = as.integer(so[["annotation_level2"]][,1])
DefaultAssay(so) = "ADT"

dict_l2 = so@meta.data %>% count(annotation_level2, level2_int)
#Select only good ADT markers
markers = so[["ADT"]]@meta.features %>% filter(IGT1only == F & classification %in% c("good", "intermediate")) %>% row.names()

sil_summary_list = list()
for (i in unique(so$IGT)) {
    print(i)
    tmp = so[markers,so$IGT == i]
    subcl = tmp@meta.data %>% count(annotation_level2) %>% filter(n>10) %>% pull(annotation_level2) %>% as.character()
    if (length(subcl) > 1) {
        tmp = tmp[,tmp$annotation_level2 %in% subcl]
        tmp = tmp %>% 
            NormalizeData(normalization.method = "CLR", verbose = F) %>%
            FindVariableFeatures(selection.method = "vst", nfeatures = 50, verbose = F) %>%
            ScaleData(features = VariableFeatures(.), verbose = F) %>%    
            RunPCA(features = VariableFeatures(.), npcs = 10, verbose = F) #%>% FindNeighbors(dims = 1:10, verbose = F) %>% RunUMAP(dims = 1:10, seed.use = 123, verbose = T)
        # DimPlot(tmp, reduction = "umap", group.by = "annotation_level2", label = T) + scale_color_manual(values = mypal_level2) + NoLegend()
        
        # message("Calculating silhouette post integration...")
        reduc.dist = dist(tmp[["pca"]]@cell.embeddings)
        cl = as.integer(tmp[["level2_int"]][,1])
        # sil_post = silhouette(cl, dist = reduc.dist)
        sil_post = silhouette_medoid_from_dist(d = reduc.dist, clusters = cl)
        sil_post$sil_width = sil_post$s
        sil_summary = as.data.frame(sil_post) %>% 
            group_by(cluster) %>% 
            summarise(mean_sil_width = mean(sil_width),
                      median_sil_width = median(sil_width),
                      total = n(),  # Count total observations in each cluster
                      positive_count = sum(sil_width > 0),  # Count negative silhouette widths
                      percentage_positive = (positive_count / total) * 100)     %>%  # Calculate percentage) 
            as.data.frame()
        # sil_summary$cluster == dict_l2$level2_int[match(sil_summary$cluster, dict_l2$level2_int)]
        sil_summary$annotation_level2 = dict_l2$annotation_level2[match(sil_summary$cluster, dict_l2$level2_int)]
        sil_summary$IGT = i
        sil_summary_list[[i]] = sil_summary 
    }
    
}


sil_summary_merged = Reduce(rbind, sil_summary_list)

dict_l1 = so@meta.data %>% count(annotation_level1, annotation_level2)
sil_summary_merged$annotation_level1 = dict_l1$annotation_level1[match(sil_summary_merged$annotation_level2, dict_l1$annotation_level2)]
    

sil_summary_merged$annotation_level2 <- factor(
    sil_summary_merged$annotation_level2,
    levels = levels(so$annotation_level2)
)

p = ggplot(sil_summary_merged %>% filter(total > 10),
       aes(x = annotation_level2, y = median_sil_width)) +
    geom_jitter(aes(color = IGT),
                width = 0.15, height = 0,
                size = 1.8, alpha = 0.9) +
    geom_boxplot(outlier.shape = NA,
                 fill = "grey90",
                 colour = "black",
                 width = 0.6) +
    scale_color_manual(values = mypal_igt) +
    theme_bw() +
    # labs(
    #     x = "Cluster (annotation_level2)",
    #     y = "Median silhouette width",
    #     color = "IGT"
    # ) +
    facet_grid(
        . ~ annotation_level1,
        scales = "free_x",
        space  = "free_x"
    ) +
    labs(x = NULL, y = NULL) +
    theme_minimal(base_size = 11) +
    theme(
        axis.text.x = element_text(angle = 60, hjust = 1, vjust = 1, size = 12),
        axis.text.y = element_text(size = 12),
        panel.grid = element_blank(),
        strip.text = element_text(face = "bold")
    )
p
# pdf(sprintf("%s/Boxplot_Level2RNASpace.pdf", out_dir), 25,6, useDingbats = F)

```

Cluster Separability in RNA Space
```{r}
cluster_separability_centroid <- function(emb, clusters, b_mode = c("mean", "min")) {
  # emb: matrix (n_cells x n_dims), e.g. PCA embeddings
  # clusters: vector length n_cells (cluster labels, numeric or character)
  # b_mode: "mean" compares to all other clusters (recommended)
  #         "min" compares to nearest other cluster (classic silhouette-like)
  b_mode <- match.arg(b_mode)

  emb <- as.matrix(emb)
  if (nrow(emb) != length(clusters)) stop("length(clusters) must match nrow(emb)")
  clusters <- as.factor(clusters)
  levs <- levels(clusters)

  # centroids per cluster
  centroids <- vapply(levs, function(cl) {
    colMeans(emb[clusters == cl, , drop = FALSE])
  }, FUN.VALUE = numeric(ncol(emb)))
  centroids <- t(centroids)  # K x D
  rownames(centroids) <- levs

  # a(C): mean distance of member cells to own centroid
  a <- vapply(levs, function(cl) {
    X <- emb[clusters == cl, , drop = FALSE]
    mu <- centroids[cl, , drop = FALSE]
    # d <- sqrt(rowSums((X - mu)^2))
    d <- sqrt(rowSums((t(apply(X, 1, function(r) {r - mu})))^2)) 
    mean(d)
  }, FUN.VALUE = numeric(1))
  names(a) <- levs

  # centroid-to-centroid distances
  Dcc <- as.matrix(dist(centroids))

  # b(C): mean (or min) distance from centroid to other centroids
  b <- vapply(levs, function(cl) {
    others <- setdiff(levs, cl)
    if (length(others) == 0) return(NA_real_)
    if (b_mode == "mean") mean(Dcc[cl, others])
    else min(Dcc[cl, others])
  }, FUN.VALUE = numeric(1))
  names(b) <- levs

  s <- (b - a) / pmax(a, b)
  data.frame(
    cluster = levs,
    a = unname(a),
    b = unname(b),
    s = unname(s),
    stringsAsFactors = FALSE
  )
}

library(cluster)
so = so_orig[,so_orig@meta.data %>% filter(organ_simplified != "thymus") %>% row.names()]
so$level2_int = as.integer(so[["annotation_level2"]][,1])
DefaultAssay(so) = "RNA"

dict_l2 = so@meta.data %>% count(annotation_level2, level2_int)

sil_summary_list <- list()

for (i in unique(so$IGT)) {
  message(i)
  tmp <- so[, so$IGT == i]

  subcl <- tmp@meta.data %>%
    dplyr::count(annotation_level2) %>%
    dplyr::filter(n > 10) %>%
    dplyr::pull(annotation_level2) %>%
    as.character()

  if (length(subcl) > 1) {
    tmp <- tmp[, tmp$annotation_level2 %in% subcl]

    tmp <- tmp %>%
        NormalizeData(normalization.method = "LogNormalize", scale.factor = 1e4, verbose = F) %>%
        FindVariableFeatures(selection.method = "vst", nfeatures = 2000, verbose = F) %>%
        ScaleData(features = VariableFeatures(.), verbose = F) %>%     # note the "."
        RunPCA(features = VariableFeatures(.), npcs = 50, verbose = F, reduction.name = "pca")

    # --- NEW: cluster-level separability on PCA embeddings ---
    emb <- Embeddings(tmp, "pca")[, 1:10, drop = FALSE]
    cl  <- as.integer(tmp[["level2_int"]][, 1])

    cl_sep <- cluster_separability_centroid(emb = emb, clusters = cl, b_mode = "mean")
    # cl_sep has one row per cluster, with s = cluster separability score

    sil_summary <- cl_sep %>% as.data.frame()

    sil_summary$annotation_level2 <- dict_l2$annotation_level2[match(sil_summary$cluster, dict_l2$level2_int)]
    sil_summary$IGT <- i
    sil_summary_list[[i]] <- sil_summary
  }
}

# saveRDS(sil_summary_list, file = "ClusterSeparation_RNA_IGT.Rds")

sil_summary_merged = Reduce(rbind, sil_summary_list)

dict_l1 = so@meta.data %>% count(annotation_level1, annotation_level2)
sil_summary_merged$annotation_level1 = dict_l1$annotation_level1[match(sil_summary_merged$annotation_level2, dict_l1$annotation_level2)]
    

sil_summary_merged$annotation_level2 <- factor(
    sil_summary_merged$annotation_level2,
    levels = levels(so$annotation_level2)
)

p = ggplot(sil_summary_merged,
       aes(x = annotation_level2, y = s)) +
    geom_jitter(aes(color = IGT),
                width = 0.15, height = 0,
                size = 1.8, alpha = 0.9) +
    geom_boxplot(outlier.shape = NA,
                 fill = "grey90",
                 colour = "black",
                 width = 0.6) +
    scale_color_manual(values = mypal_igt) +
    theme_bw() +
    # labs(
    #     x = "Cluster (annotation_level2)",
    #     y = "Median silhouette width",
    #     color = "IGT"
    # ) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "black", linewidth = 0.3) +
    facet_grid(
        . ~ annotation_level1,
        scales = "free_x",
        space  = "free_x"
    ) +
    labs(x = NULL, y = NULL) +
    theme_minimal(base_size = 11) +
    theme(
        axis.text.x = element_text(angle = 60, hjust = 1, vjust = 1, size = 12),
        axis.text.y = element_text(size = 12),
        panel.grid = element_blank(),
        strip.text = element_text(face = "bold")
    )

pdf(sprintf("%s/Boxplot_ClusterSeparability_Level2RNASpace.pdf", out_dir), 25,6, useDingbats = F)
p
dev.off()

mean(sil_summary_merged$s <0)


```


Cluster Separability in Protein Space
```{r}
library(cluster)
so = so_orig[,so_orig@meta.data %>% filter(organ_simplified != "thymus", cite_seq == T) %>% row.names()]
so$level2_int = as.integer(so[["annotation_level2"]][,1])
DefaultAssay(so) = "ADT"

dict_l2 = so@meta.data %>% count(annotation_level2, level2_int)
#Select only good ADT markers
markers = so[["ADT"]]@meta.features %>% filter(IGT1only == F & classification %in% c("good", "intermediate")) %>% row.names()

sil_summary_list <- list()

for (i in unique(so$IGT)) {
  message(i)
  tmp <- so[markers, so$IGT == i]

  subcl <- tmp@meta.data %>%
    dplyr::count(annotation_level2) %>%
    dplyr::filter(n > 10) %>%
    dplyr::pull(annotation_level2) %>%
    as.character()

  if (length(subcl) > 1) {
    tmp <- tmp[, tmp$annotation_level2 %in% subcl]

    tmp <- tmp %>%
      NormalizeData(normalization.method = "CLR", verbose = FALSE) %>%
      FindVariableFeatures(selection.method = "vst", nfeatures = 50, verbose = FALSE) %>%
      ScaleData(features = VariableFeatures(.), verbose = FALSE) %>%
      RunPCA(features = VariableFeatures(.), npcs = 10, verbose = FALSE)

    # --- NEW: cluster-level separability on PCA embeddings ---
    emb <- Embeddings(tmp, "pca")[, 1:10, drop = FALSE]
    cl  <- as.integer(tmp[["level2_int"]][, 1])

    cl_sep <- cluster_separability_centroid(emb = emb, clusters = cl, b_mode = "mean")
    # cl_sep has one row per cluster, with s = cluster separability score

    sil_summary <- cl_sep %>% as.data.frame()
    # sil_summary <- cl_sep %>%
    #   dplyr::mutate(
    #     mean_sil_width = s,          # keep your naming convention
    #     median_sil_width = s,        # same value at cluster level
    #     total = as.integer(table(clusters = cl)[as.character(cluster)]), # cluster size
    #     positive_count = ifelse(s > 0, total, 0),
    #     percentage_positive = ifelse(total > 0, 100 * positive_count / total, NA_real_)
    #   ) %>%
    #   dplyr::select(cluster, mean_sil_width, median_sil_width, total, positive_count, percentage_positive) %>%
    #   as.data.frame()

    sil_summary$annotation_level2 <- dict_l2$annotation_level2[match(sil_summary$cluster, dict_l2$level2_int)]
    sil_summary$IGT <- i
    sil_summary_list[[i]] <- sil_summary
  }
}

# saveRDS(sil_summary_list, file = "ClusterSeparation_ADT_IGT.Rds")

sil_summary_merged = Reduce(rbind, sil_summary_list)

dict_l1 = so@meta.data %>% count(annotation_level1, annotation_level2)
sil_summary_merged$annotation_level1 = dict_l1$annotation_level1[match(sil_summary_merged$annotation_level2, dict_l1$annotation_level2)]
    

sil_summary_merged$annotation_level2 <- factor(
    sil_summary_merged$annotation_level2,
    levels = levels(so$annotation_level2)
)

p = ggplot(sil_summary_merged,
       aes(x = annotation_level2, y = s)) +
    geom_jitter(aes(color = IGT),
                width = 0.15, height = 0,
                size = 1.8, alpha = 0.9) +
    geom_boxplot(outlier.shape = NA,
                 fill = "grey90",
                 colour = "black",
                 width = 0.6) +
    scale_color_manual(values = mypal_igt) +
    theme_bw() +
    # labs(
    #     x = "Cluster (annotation_level2)",
    #     y = "Median silhouette width",
    #     color = "IGT"
    # ) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "black", linewidth = 0.3) +
    facet_grid(
        . ~ annotation_level1,
        scales = "free_x",
        space  = "free_x"
    ) +
    labs(x = NULL, y = NULL) +
    theme_minimal(base_size = 11) +
    theme(
        axis.text.x = element_text(angle = 60, hjust = 1, vjust = 1, size = 12),
        axis.text.y = element_text(size = 12),
        panel.grid = element_blank(),
        strip.text = element_text(face = "bold")
    )

pdf(sprintf("%s/Boxplot_ClusterSeparability_Level2ADTSpace.pdf", out_dir), 25,6, useDingbats = F)
p
dev.off()

mean(sil_summary_merged$s <0)



```




