---
title: "clustering_gdT"
output: html_document
date: "2024-12-01"
editor_options: 
  chunk_output_type: console
---

**Start interactive session on Midway3**
```{bash}
ssh -Y zemmour@midway3.rcc.uchicago.edu
sinfo -s
screen -S david
sinteractive --exclusive --time=12:00:00 --account=pi-zemmour --partition=jfkfloor2 --qos=jfkfloor2 --nodes=1 
sinteractive --time=12:00:00 --account=pi-zemmour --partition=caslake --nodes=1 --mem=128GB
sinteractive --time=12:00:00 --account=pi-zemmour --partition=gpu --gres=gpu:1 --mem=152GB
cd $LABHOME

module load python/anaconda-2022.05 
module load hdf5/1.12.0 #for BPCells but also others
module load openblas/0.3.13

#sinteractive --time=1:00:00 -p gpu --gres=gpu:1 --account=pi-zemmour 

#source activate /project/jfkfloor2/zemmourlab/david/envs/scvi
source activate /project/zemmour/david/envs/scvi_20240315
module load openblas/0.3.13 #load again or error

cd /project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/gdT
#cd ~/google_drive/ImmgenT/analysis/data_integration/IGT1_56/Treg

R
#options(max.print=1000)
#options(Seurat.object.assay.version = 'v5')
```

**Initialize: load libraries and custom functions**

```{r}
options(max.print=1000)
setwd("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/gdT")
source("/project/jfkfloor2/zemmourlab/david/immgent/immgent_integration_git/custom_functions_david.R")

libs = c("fgsea","fastTopics", "flashier", "Matrix", "Seurat","BPCells", "ggplot2", "dplyr", "reshape2", "ggrastr", "RColorBrewer", "pals", "scales", "pheatmap", "scattermore") 
#libs = c("matrixStats", "gplots","ggplot2", "reshape2", "scales", "gridExtra", "dplyr", "RColorBrewer", "grid", "Rtsne", "limma",   "RColorBrewer", "pheatmap", "Seurat",  "ggrastr", "ggbeeswarm") #pwr, preprocessCore, "readxl", vegan", "genefilter" "scde" "cellrangerRkit", "Rsamtools", "GenomicRanges", "GenomicAlignments", "VGAM", "WGCNA",rafalib, ‘UsingR’,MAST "ggrastr",
sapply(libs, function(x) suppressMessages(suppressWarnings(library(x, character.only = TRUE, quietly = T, warn.conflicts  = F))))


#Gradients
library(RColorBrewer)
ColorRamp = rev(colorRampPalette(brewer.pal(n = 7,name = "RdYlBu"))(100))
ColorRamp = rev(colorRampPalette(c('red','white','blue'))(20))
ColorRamp = rev(rainbow(10, end = 4/6))
library(viridis)
ColorRamp = rev(viridis(100))
ColorRamp = rev(cividis(100))
#image(1:100, 1, as.matrix(1:100), col = ColorRamp, main = "Color Palette", xaxt = 'n', yaxt = 'n', xlab = "", ylab = "")

#mycol = c("black","blue","red","darkgreen", "orange","purple", "cyan", "greenyellow",    "salmon", "magenta","pink", "tan", "brown") # c("magenta", "red",  "darkgreen", "cyan", "blue", "blue", "blue", "black", "blue", "blue", "blue", "blue", "orange")

#Large Color palette
library("pals")
n = 70
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
mypal1 = unique(unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals))))
mypal1 = mypal1[-4]
mypal = c(glasbey(), polychrome(), mypal1)
names(mypal) = NULL
#pal.bands(mypal, main="Colormap suggestions")

#parade = function(n) { return(Seurat::DiscretePalette(n, palette = "parade", shuffle = F)) }


```

**clustering V3 20241201**
 cluster_adjustments.Rmd: sbatch $SCRIPT_DIR/run_totalvi_wrapper_20241201_gdT_rmIGTsample.sh

```{r}
prefix = "totalvi_20241201_gdT_rmIGTsample"

so_orig = readRDS("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/igt1_96_withtotalvi20241201.Rds")
setwd("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/gdT")
so = so_orig[,so_orig$annotation_level1 == "gdT"]

dir_with_file = "/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/gdT/totalvi_20241201_gdT_rmIGTsample"
so = AddLatentData(so, latent_file = sprintf("%s/latent.csv",dir_with_file), prefix = "totalvi_20241201_gdT_rmIGTsample", calculate_umap = F)
so = AddLatentData(so, latent_file = sprintf("%s/mde2.csv",dir_with_file), prefix = "mde_totalvi_20241201_gdT_rmIGTsample", calculate_umap = F)

prefix = "totalvi_20241201_gdT_rmIGTsample"
umap_to_plot = "mde_totalvi_20241201_gdT_rmIGTsample"
pdf(sprintf("%s/MDE_clusters.pdf", prefix), 20, 20, useDingbats = F)
DimPlot(so, reduction = umap_to_plot, group.by = "annotation_level1", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal)
DimPlot(so, reduction = umap_to_plot, group.by = "annotation_level2", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal)
MyDimPlotHighlight(seurat_object = so, umap_to_plot = umap_to_plot, cells_to_highlight = names(which(so$nonconv_tcr_recog == T)), highlight_column = "nonconv_tcr_recog", pixels = c(512, 512), mycols = "red")
MyDimPlotHighlight(seurat_object = so, umap_to_plot = umap_to_plot, cells_to_highlight = names(which(so$iNKT == T)), highlight_column = "iNKT", pixels = c(512, 512), mycols = "red")
MyDimPlotHighlight(seurat_object = so, umap_to_plot = umap_to_plot, cells_to_highlight = names(which(so$MAIT == T)), highlight_column = "MAIT", pixels = c(512, 512), mycols = "red")
dev.off()

saveRDS(so, file = "igt1_96_gdT_20241201.Rds")

```

**Update Seurat object 20241216**
same clusters, adding level1, level2_parent1 etc

```{r}
AddLatentData = function(so, latent_file, prefix, calculate_umap = F) {
    totalvi = read.csv(file = latent_file, header = T, sep = ",", row.names = 1)
    # pattern = "(IGT\\S*)_(\\d+)"
    # rownames(totalvi) = gsub(pattern, "\\1", rownames(totalvi))
    colnames(totalvi) = as.character(seq(1:ncol(totalvi)))
    #all(rownames(totalvi) == colnames(so))
    totalvi = as.matrix(totalvi)
    pattern = "^IGT\\d{1,2}\\.[A-Z]{16}$"
    matches = grepl(pattern, rownames(totalvi))
    if (!all(matches)) {
        print("updating cell names in csv file, assuming same order as seurat object" )
        #rownames(totalvi) = gsub(pattern = "(\\S*)_(\\d+)", replacement = "\\1", rownames(totalvi))
        rownames(totalvi) = colnames(so)
    }
    totalvi = totalvi[colnames(so),]
    if (all(colnames(so) == rownames(totalvi))) {
        print("all cells ARE in the latent_file")
        so[[prefix]] = CreateDimReducObject(embeddings = totalvi, key = prefix, assay = "RNA") #sprintf("totalvi_%s", prefix2)
        if (calculate_umap == T) {
            so = RunUMAP(so, dims = 1:ncol(totalvi), reduction = prefix, n.components = 2, reduction.name = sprintf("umap_%s", prefix))
        }
        return(so)
    } else {
        print("all cells ARE NOT in the latent_file")
        return(so)
    }
}

so_orig = readRDS("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/igt1_96_withtotalvi20241216.Rds")
setwd("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/gdT")
so = so_orig[,so_orig$annotation_level1 == "gdT"] # so = so_orig[,so_orig$annotation_level0 == "CD4"]

table_data = table(so$annotation_level2, so$annotation_level1)
# pdf(sprintf("%s/ClustersOct24_Newclusters.pdf", "."), width = 10, height = 10)
# 
# # Convert the table to a data frame for better readability in the PDF
# table_df <- as.data.frame(table_data)
# 
# # Create the table in the PDF
# grid.table(table_df)
# 
# # Close the PDF device
# dev.off()

```

Add general totalvi
```{r}
dir_with_file = "/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/totalvi_20241006"
so = AddLatentData(so, latent_file = sprintf("%s/latent.csv",dir_with_file), prefix = "totalvi_20241006", calculate_umap = F)
so = AddLatentData(so, latent_file = sprintf("%s/mde2.csv",dir_with_file), prefix = "mde2_totalvi_20241006", calculate_umap = F)

```

Add not_CD4CD8 totalvi
```{r}
dir_with_file = "/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/unconventional/totalvi_20241113_not_CD4CD8_rmIGTsample"
so = AddLatentData(so, latent_file = sprintf("%s/latent.csv",dir_with_file), prefix = "totalvi_20241113_not_CD4CD8_rmIGTsample", calculate_umap = F)
so = AddLatentData(so, latent_file = sprintf("%s/mde_incremental.csv",dir_with_file), prefix = "mde_totalvi_20241113_not_CD4CD8_rmIGTsample", calculate_umap = F)

```

Add gdT specific MDE
```{r}
dir_with_file = "/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/gdT/totalvi_20241201_gdT_rmIGTsample"
so = AddLatentData(so, latent_file = sprintf("%s/latent.csv",dir_with_file), prefix = "totalvi_20241201_gdT_rmIGTsample", calculate_umap = F)
so = AddLatentData(so, latent_file = sprintf("%s/mde2.csv",dir_with_file), prefix = "mde_totalvi_20241201_gdT_rmIGTsample", calculate_umap = F)

```

Plots
```{r}

pdf(sprintf("%s/MDE_AnnotationLevel20241216.pdf", "."), 10, 10, useDingbats = F)
MyDimPlotHighlight(so_orig, umap_to_plot = "mde2_totalvi_20241006", cells_to_highlight = names(which(so_orig$annotation_level1 == "gdT")), highlight_column_name ="annotation_level2", highlight_size = 0.5, highlight_alpha = 0.5, pixels = c(512, 512), mycols = mypal, print_plot1 = , print_plot2 = F, labelclusters = T )
# MyDimPlotHighlight(so, umap_to_plot = "mde_totalvi_20241113_not_CD4CD8_rmIGTsample", cells_to_highlight = names(which(so$annotation_level1 == "gdT")), highlight_column_name ="annotation_level2", highlight_size = 0.5, highlight_alpha = 0.5, pixels = c(512, 512), mycols = mypal, print_plot1 = , print_plot2 = F, labelclusters = T )
DimPlot(so, reduction = "mde_totalvi_20241113_not_CD4CD8_rmIGTsample", group.by = "annotation_level2", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal)
DimPlot(so, reduction = "mde_totalvi_20241201_gdT_rmIGTsample", group.by = "annotation_level2", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal)
DimPlot(so, reduction = "mde_totalvi_20241201_gdT_rmIGTsample", group.by = "annotation_level2", raster = T, raster.dpi = c(512,512), label = F) + scale_color_manual(values = mypal) + theme_void() + NoLegend() 
# DimPlot(so, reduction = "mde_totalvi_20241201_gdT_rmIGTsample", group.by = "annotation_level2_parent1", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal)
# DimPlot(so, reduction = "mde_totalvi_20241201_gdT_rmIGTsample", group.by = "annotation_level2_parent1", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal) + theme_void() + NoLegend() 
dev.off()

saveRDS(so, file = "igt1_96_gdT_20241216.Rds")

```

remove unwanted cell embeddings
totalvi_20241006 - mde2_totalvi_20241006
totalvi_20241113_CD4_rmIGTsample - mde_incremental
```{r}
so[["umap_totalvi_20241006"]] = NULL
so[["mde_totalvi_20241006"]] = NULL
so[["umap2_totalvi_20241006"]] = NULL
so[["mdeumap_totalvi_20241006"]] = NULL
so[["umap_totalvi_20241008_rmIGTsample"]] = NULL
so[["mde_totalvi_20241008_rmIGTsample"]] = NULL
so[["mde2_totalvi_20241008_rmIGTsample"]] = NULL
so[["mdeumap_totalvi_20241008_rmIGTsample"]] = NULL
so[["mde_totalvi_20241029_fitleredGenesProt"]] = NULL
so[["umap_totalvi_20241029_fitleredGenesProt"]] = NULL

saveRDS(so, file = "igt1_96_gdT_20241216.Rds")
```

**Annotation 20250109**

```{r}
so_orig = readRDS("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/igt1_96_withtotalvi20250109_level1only.Rds")
mypal_level1 = setNames(mypal, unique(so_orig$annotation_level1))
mypal_level2 = setNames(mypal, unique(so_orig$annotation_level2))
setwd("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/gdT")
so = so_orig[,so_orig$annotation_level1 == "gdT"]

dir_with_file = "/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/totalvi_20241008_rmIGTsample"
so = AddLatentData(so, latent_file = sprintf("%s/latent.csv",dir_with_file), prefix = "totalvi_20241008_rmIGTsample", calculate_umap = F)
dir_with_file = "/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/gdT/20250109_gdT"
so = AddLatentData(so, latent_file = sprintf("%s/mde_incremental.csv",dir_with_file), prefix = "mde_incremental", calculate_umap = F)

```

Transfer labels

```{r}
MyTransferLabels = function(cell_embeddings_ref, cell_embeddings_query, labels_ref = as.character(so$annotation_level2_20241216[so$is_ref == "ref"]), n_neighbors = 30) {
    require(RcppAnnoy)
    # cell_embeddings_ref = so[["totalvi_20241008_rmIGTsample"]]@cell.embeddings[so$is_ref == "ref",]
    # cell_embeddings_query = so[["totalvi_20241008_rmIGTsample"]]@cell.embeddings[so$is_ref == "query",]
    # labels_ref = as.character(so$annotation_level2_20241216[so$is_ref == "ref"])
    
    # Use the Annoy algorithm to find nearest neighbors between query and ref datasets
    message(sprintf("Number of nearest neighbors to find betweeen query and ref: %s", n_neighbors))
    
    message("Create Annoy index for ref dataset")
    annoy_index = new(AnnoyAngular, ncol(cell_embeddings_ref)) ##use cosine distance Angular
    for (i in 1:nrow(cell_embeddings_ref)) {
        annoy_index$addItem(i - 1, cell_embeddings_ref[i, ])
    }
    annoy_index$build(10)  # Build the index with 10 trees
    
    message("Find nearest ref neighbors for each cell in query dataset")
    nn_indices <- t(sapply(1:nrow(cell_embeddings_query), function(i) {
        annoy_index$getNNsByVector(cell_embeddings_query[i, ], n_neighbors)
    }))
    
    # nn_indices gives you the indices of nearest neighbors in the ref dataset
    # the rows are cells from query dataset, columns are the 30 nearest cells in the ref dataset
    # dim(nn_indices)
    # head(nn_indices)
    
    message("Transfer labels based on majority vote from nearest neighbors")
    transfer_labels <- apply(nn_indices, 1, function(neighbors) {
        # Get labels for the nearest neighbors
        neighbor_labels <- labels_ref[neighbors + 1]  # Add 1 for R's 1-based index
        
        # Return the most common label (majority vote)
        most_common_label <- names(sort(table(neighbor_labels), decreasing = TRUE))[1]
        return(most_common_label)
    })
    
    names(transfer_labels) = rownames(cell_embeddings_query)
    
    return(transfer_labels)
}

so$level2_new = as.character(so$annotation_level2_20241216)
so$level2_new[so$annotation_level2_parent1 == "proliferating"] = "gdT_prolif"
so$level2_new[so$annotation_level2_parent1 == "miniverse"] = "gdT_miniverse"
# so$level2_new[so$annotation_level2_parent1 == "preT"] = "CD4_preT"
table(so$level2_new, so$annotation_level1)
table(so$level2_new, so$annotation_level2_parent1)

so$is_ref = "other"
so$is_ref[so$annotation_level1 == "gdT" & so$annotation_level1_20241216 == "gdT"] = "ref"
so$is_ref[so$annotation_level1 == "gdT" & so$annotation_level1_20241216 != "gdT"] = "query"
so$is_ref[so$annotation_level2_parent1 %in% c("miniverse")] = "other" #"proliferating", "preT"
cell_embeddings_ref = so[["totalvi_20241008_rmIGTsample"]]@cell.embeddings[so$is_ref == "ref",]
cell_embeddings_query = so[["totalvi_20241008_rmIGTsample"]]@cell.embeddings[so$is_ref == "query",]
labels_ref = as.character(so$annotation_level2_20241216[so$is_ref == "ref"])

transfer_labels = MyTransferLabels(cell_embeddings_ref, cell_embeddings_query, labels_ref, n_neighbors = 30)

# Now, transfer_labels contains the predicted labels for the 3k PBMC dataset
so$level2_new[so$is_ref == "query"] = transfer_labels
table(transfer_labels)

table(so$annotation_level2_parent1, so$level2_new)
table(so$annotation_level1, so$level2_new)
table(so$annotation_level2_20241216, so$level2_new)

```

Check transfer and readjust clusters if needed, i.e. add new clusters
CD4_cl28: used to be abT DP

```{r}
# FeaturePlot(so, reduction = "mde_incremental",slot = "count", features = "Stmn1", order = T, raster = T, raster.dpi = c(1024,1024))

prefix_latent = "totalvi_20241008_rmIGTsample"
n_latent = ncol(so[[prefix_latent]]@cell.embeddings)
so = FindNeighbors(so, dims = 1:n_latent, reduction = prefix_latent)
so = FindClusters(so, resolution = 1, cluster.name = sprintf("Cluster_%s_Res1",prefix_latent))

pdf(sprintf("%s/MDE_annotation_20250109_TransferLabels_Reclustering_to_see.pdf", "20250109_gdT"), 20, 20, useDingbats = F)
DimPlot(so, reduction = "mde_incremental", group.by = "annotation_level1", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal_level1)
DimPlot(so, reduction = "mde_incremental", group.by = "annotation_level1_20241216", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal)
DimPlot(so, reduction = "mde_incremental", group.by = "level2_new", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal_level2) + ggtitle("transfer labels from annotation_level2_20241216")
DimPlot(so, reduction = "mde_incremental", group.by = "annotation_level2_20241216", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal_level2)
DimPlot(so, reduction = "mde_incremental", group.by = "annotation_level2_parent1", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal)
DimPlot(so, reduction = "mde_incremental", group.by = sprintf("Cluster_%s_Res1",prefix_latent), raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal)
dev.off()

so$level2_new[so$Cluster_totalvi_20241008_rmIGTsample_Res1 %in% "9" ] = "gdT_52"
so$level2_new[so$Cluster_totalvi_20241008_rmIGTsample_Res1 %in% "15" & so[["mde_incremental"]]@cell.embeddings[,1] > 1.7 ] = "gdT_53"
so$level2_new[so$Cluster_totalvi_20241008_rmIGTsample_Res1 %in% "15" & so[["mde_incremental"]]@cell.embeddings[,1] < 1.7 ] = "gdT_54"
so$level2_new[so$level2_new  %in% c("not_CD4CD8_cl48")] = "gdT_prolif"
so$level2_new[so$annotation_level2_parent1  %in% c("proliferating")] = "gdT_prolif"

# DimPlot(so, reduction = "mde_incremental", cells = names(so$level2_new[so$Cluster_totalvi_20241008_rmIGTsample_Res1 %in% "25" & so[["mde_incremental"]]@cell.embeddings[,1] < 1 ]), raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal_level2)

# so$annotation_level2_parent1 = "not_applicable"
# annot$level2_parent1[annot$level2 %in% c("CD4_cl10", "CD4_cl13", "CD4_cl14", "CD4_cl15", "CD4_cl16", "CD4_cl17", "CD4_cl18", "CD4_cl20", "CD4_cl21", "CD4_cl22", "CD4_cl23", "CD4_cl25", "CD4_cl9", "CD8AB_cl10", "CD8AB_cl11", "CD8AB_cl12", "CD8AB_cl13", "CD8AB_cl14", "CD8AB_cl15", "CD8AB_cl18", "CD8AB_cl20", "CD8AB_cl6", "CD8AB_cl7","treg_cl2", "treg_cl3", "treg_cl4", "treg_cl6", "treg_cl8")] = "activated"
# so$annotation_level2_parent1[so$annotation_level2 %in% c("CD4_cl1", "CD4_cl12", "CD4_cl2", "CD4_cl3", "CD4_cl4", "CD4_cl5", "CD4_cl7", "CD8AB_cl1", "CD8AB_cl2", "CD8AB_cl22", "CD8AB_cl4", "CD8AB_cl5","treg_cl1", "treg_cl7")] = "resting"
# so$annotation_level2_parent1[so$annotation_level2 %in% c("CD4_cl8", "CD8AB_cl19", "treg_cl5","not_CD4CD8_cl48")] = "proliferating"
# table(so$annotation_level2_parent1, so$annotation_level2)

#rerun annotation_level2_parent1 since new annotated cells from transfer labels etc
so$annotation_level2_parent1[!so$level2_new %in% c("gdT_prolif", "gdT_miniverse", "not_CD4CD8_cl48")] = "other"
table(so$annotation_level2_parent1, so$level2_new)

pdf(sprintf("%s/MDE_annotation_20250109_level2.pdf", "20250109_gdT"), 20, 20, useDingbats = F)
DimPlot(so, reduction = "mde_incremental", group.by = "annotation_level1", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal_level1)
DimPlot(so, reduction = "mde_incremental", group.by = "annotation_level1_20241216", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal)
DimPlot(so, reduction = "mde_incremental", group.by = "level2_new", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal) + ggtitle("After transfer labels and reclustering")
DimPlot(so, reduction = "mde_incremental", group.by = "annotation_level2_20241216", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal_level2)
DimPlot(so, reduction = "mde_incremental", group.by = "annotation_level2_parent1", raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal)
dev.off()

if (all(colnames(so_orig)[so_orig$annotation_level1 == "gdT"] == colnames(so))) {
    print("True")
    so_orig$annotation_level2[so_orig$annotation_level1 == "gdT"] = so$level2_new[colnames(so_orig)[so_orig$annotation_level1 == "gdT"]]
    so_orig$annotation_level2_parent1[so_orig$annotation_level1 == "gdT"] = so$annotation_level2_parent1[colnames(so_orig)[so_orig$annotation_level1 == "gdT"]]
}

table(so_orig$annotation_level2_parent1[so_orig$annotation_level1 == "gdT"], so_orig$annotation_level2[so_orig$annotation_level1 == "gdT"])
table(so_orig$annotation_level2[so_orig$annotation_level1 == "gdT"], so_orig$annotation_level2_20241216[so_orig$annotation_level1 == "gdT"])

#Export to save awaiting the other annotations
annot = data.frame(cell_id = colnames(so_orig), so_orig@meta.data[,c("annotation_level1", "annotation_level2", "annotation_level2_parent1", "nonconv_tcr_recog")])
colnames(annot) = c("cell_id", "level1", "level2_parent1", "level2", "nonconv_tcr_recog")
table(annot$level2_parent1, annot$level1)
sum(rowSums(apply(annot,2,is.na)))
write.table(x = annot, file = "../annotation_table_20250109_withlevel2CD8AB_CD4.csv", row.names = F, quote = F, sep = ",")


```
