---
title: "clustering_CD8"
output: html_document
date: "2024-10-17"
editor_options: 
  chunk_output_type: console
---

**Start interactive session on Midway3**
```{bash}
ssh -Y zemmour@midway3.rcc.uchicago.edu
sinfo -s
screen -S david
sinteractive --exclusive --time=12:00:00 --account=pi-zemmour --partition=jfkfloor2 --qos=jfkfloor2 --nodes=1 
sinteractive --time=12:00:00 --account=pi-zemmour --partition=caslake --nodes=1 --mem=128GB
sinteractive --time=3:00:00 --account=pi-zemmour --partition=gpu --gres=gpu:1 --mem=152GB
cd $LABHOME

module load python/anaconda-2022.05 
module load hdf5/1.12.0 #for BPCells but also others
module load openblas/0.3.13

#sinteractive --time=1:00:00 -p gpu --gres=gpu:1 --account=pi-zemmour 

#source activate /project/jfkfloor2/zemmourlab/david/envs/scvi
source activate /project/zemmour/david/envs/scvi_20240315
module load openblas/0.3.13 #load again or error

cd /project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/miniverse
#cd ~/google_drive/ImmgenT/analysis/data_integration/IGT1_56/Treg

R
#options(max.print=1000)
#options(Seurat.object.assay.version = 'v5')
```

**Initialize: load libraries and custom functions**

```{r}
options(max.print=1000)
setwd("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/miniverse")
source("/project/jfkfloor2/zemmourlab/david/immgent/immgent_integration_git/custom_functions_david.R")

libs = c("fgsea","fastTopics", "flashier", "Matrix", "Seurat","BPCells", "ggplot2", "dplyr", "reshape2", "ggrastr", "RColorBrewer", "pals", "scales", "pheatmap") 
#libs = c("matrixStats", "gplots","ggplot2", "reshape2", "scales", "gridExtra", "dplyr", "RColorBrewer", "grid", "Rtsne", "limma",   "RColorBrewer", "pheatmap", "Seurat",  "ggrastr", "ggbeeswarm") #pwr, preprocessCore, "readxl", vegan", "genefilter" "scde" "cellrangerRkit", "Rsamtools", "GenomicRanges", "GenomicAlignments", "VGAM", "WGCNA",rafalib, ‘UsingR’,MAST "ggrastr",
sapply(libs, function(x) suppressMessages(suppressWarnings(library(x, character.only = TRUE, quietly = T, warn.conflicts  = F))))


#Gradients
library(RColorBrewer)
ColorRamp = rev(colorRampPalette(brewer.pal(n = 7,name = "RdYlBu"))(100))
ColorRamp = rev(colorRampPalette(c('red','white','blue'))(20))
ColorRamp = rev(rainbow(10, end = 4/6))
library(viridis)
ColorRamp = rev(viridis(100))
ColorRamp = rev(cividis(100))
#image(1:100, 1, as.matrix(1:100), col = ColorRamp, main = "Color Palette", xaxt = 'n', yaxt = 'n', xlab = "", ylab = "")

#mycol = c("black","blue","red","darkgreen", "orange","purple", "cyan", "greenyellow",    "salmon", "magenta","pink", "tan", "brown") # c("magenta", "red",  "darkgreen", "cyan", "blue", "blue", "blue", "black", "blue", "blue", "blue", "blue", "orange")

#Large Color palette
library("pals")
n = 70
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
mypal1 = unique(unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals))))
mypal1 = mypal1[-4]
mypal = c(glasbey(), polychrome(), mypal1)
names(mypal) = NULL
#pal.bands(mypal, main="Colormap suggestions")

#parade = function(n) { return(Seurat::DiscretePalette(n, palette = "parade", shuffle = F)) }


```

Seurat object with the results: igt1_96_miniverse_20241017.Rds
annotation: annotation_miniverse_clusters_l1.csv

```{r}

so = readRDS(file = "/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/miniverse/igt1_96_miniverse_20241017.Rds")
so = NormalizeData(so,assay = "RNA", normalization.method = "LogNormalize",verbose = T)
so = NormalizeData(so,assay = "ADT", normalization.method = "CLR", verbose = T)

```

**Load seurat expression data**
--> igt1_96_treg_20241017.Rds
```{r}
so_orig = readRDS("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/igt1_96_withtotalvi20241021.Rds")

so = so_orig[,so_orig$annotation_level1 == "miniverse"]
so[['mde_totalvi_20241006']] = NULL
so[['umap2_totalvi_20241006']] = NULL
so[['mdeumap_totalvi_20241006']] = NULL
so[['mde_totalvi_20241008_rmIGTsample']] = NULL
so[['mdeumap_totalvi_20241008_rmIGTsample']] = NULL

```

Add latent spaces, umaps
```{r}
# sprintf("%s_X_SCVI_%s.csv", prefix, prefix2)

AddLatentData = function(so, latent_file, prefix, calculate_umap = F) {
    totalvi = read.csv(file = latent_file, header = T, sep = ",", row.names = 1)
    # pattern = "(IGT\\S*)_(\\d+)"
    # rownames(totalvi) = gsub(pattern, "\\1", rownames(totalvi))
    colnames(totalvi) = as.character(seq(1:ncol(totalvi)))
    #all(rownames(totalvi) == colnames(so))
    totalvi = as.matrix(totalvi)
    pattern = "^IGT\\d{1,2}\\.[A-Z]{16}$"
    matches = grepl(pattern, rownames(totalvi))
    if (!all(matches)) {
        print("updating cell names in csv file, assuming same order as seurat object" )
        #rownames(totalvi) = gsub(pattern = "(\\S*)_(\\d+)", replacement = "\\1", rownames(totalvi))
        rownames(totalvi) = colnames(so)
    }
    totalvi = totalvi[colnames(so),]
    if (all(colnames(so) == rownames(totalvi))) {
        print("all cells ARE in the latent_file")
        so[[prefix]] = CreateDimReducObject(embeddings = totalvi, key = prefix, assay = "RNA") #sprintf("totalvi_%s", prefix2)
        if (calculate_umap == T) {
            so = RunUMAP(so, dims = 1:ncol(totalvi), reduction = prefix, n.components = 2, reduction.name = sprintf("umap_%s", prefix))
        }
        return(so)
    } else {
        print("all cells ARE NOT in the latent_file")
        return(so)
    }
}

dir_with_file = "/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/miniverse/totalvi_20241014_miniverse"
so = AddLatentData(so, latent_file = sprintf("%s/latent.csv",dir_with_file), prefix = "totalvi_20241014_miniverse", calculate_umap = F)
so = AddLatentData(so, latent_file = sprintf("%s/umap_python.csv",dir_with_file), prefix = "umap_totalvi_20241014_miniverse", calculate_umap = F)
so = AddLatentData(so, latent_file = sprintf("%s/mde2.csv",dir_with_file), prefix = "mde_totalvi_20241014_miniverse", calculate_umap = F)

dir_with_file = "/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/miniverse/totalvi_20241014_miniverse_rmIGTsample"
so = AddLatentData(so, latent_file = sprintf("%s/latent.csv",dir_with_file), prefix = "totalvi_20241014_miniverse_rmIGTsample", calculate_umap = F)
so = AddLatentData(so, latent_file = sprintf("%s/umap_python.csv",dir_with_file), prefix = "umap_totalvi_20241014_miniverse_rmIGTsample", calculate_umap = F)
so = AddLatentData(so, latent_file = sprintf("%s/mde2.csv",dir_with_file), prefix = "mde_totalvi_20241014_miniverse_rmIGTsample", calculate_umap = F)


```

```{r}
prefix = "totalvi_20241014_miniverse_rmIGTsample"
#setwd(prefix)
reduc = "mde_totalvi_20241014_miniverse_rmIGTsample"
reduc = "umap_totalvi_20241014_miniverse_rmIGTsample"
```

```{r}
RenameCluster = function(cluster.vector = so[[cluster.name]][,1], cluster.prefix = "cl") {
    cl = as.numeric(cluster.vector)
    ordered = order(unique(cl))
    cl = sprintf("%s%s", cluster.prefix, cl )
    renamed.vector = factor(cl, levels = unique(cl)[ordered])
    return(renamed.vector)
}

n_latent = ncol(so[[prefix]]@cell.embeddings)
so = FindNeighbors(so, dims = 1:n_latent, reduction = prefix)
so = FindClusters(so, resolution = 0.25, cluster.name = sprintf("Cluster_%s_Res0.25",prefix))
so = FindClusters(so, resolution = 0.5, cluster.name = sprintf("Cluster_%s_Res0.5",prefix))
so = FindClusters(so, resolution = 0.75, cluster.name = sprintf("Cluster_%s_Res0.75",prefix))
so = FindClusters(so, resolution = 1, cluster.name = sprintf("Cluster_%s_Res1",prefix))
so = FindClusters(so, resolution = 1.5, cluster.name = sprintf("Cluster_%s_Res1.5",prefix))
so = FindClusters(so, resolution = 2, cluster.name = sprintf("Cluster_%s_Res2",prefix))
so = FindClusters(so, resolution = 3, cluster.name = sprintf("Cluster_%s_Res3",prefix))
so = FindClusters(so, resolution = 4, cluster.name = sprintf("Cluster_%s_Res4",prefix))

cluster.name = sprintf("Cluster_%s_Res0.25",prefix)
so[[cluster.name]][,1] = RenameCluster(cluster.vector = so[[cluster.name]][,1], cluster.prefix = "cl")
cluster.name = sprintf("Cluster_%s_Res0.5",prefix)
so[[cluster.name]][,1] = RenameCluster(cluster.vector = so[[cluster.name]][,1], cluster.prefix = "cl")
cluster.name = sprintf("Cluster_%s_Res0.75",prefix)
so[[cluster.name]][,1] = RenameCluster(cluster.vector = so[[cluster.name]][,1], cluster.prefix = "cl")
cluster.name = sprintf("Cluster_%s_Res1",prefix)
so[[cluster.name]][,1] = RenameCluster(cluster.vector = so[[cluster.name]][,1], cluster.prefix = "cl")
cluster.name = sprintf("Cluster_%s_Res1.5",prefix)
so[[cluster.name]][,1] = RenameCluster(cluster.vector = so[[cluster.name]][,1], cluster.prefix = "cl")
cluster.name = sprintf("Cluster_%s_Res2",prefix)
so[[cluster.name]][,1] = RenameCluster(cluster.vector = so[[cluster.name]][,1], cluster.prefix = "cl")
cluster.name = sprintf("Cluster_%s_Res3",prefix)
so[[cluster.name]][,1] = RenameCluster(cluster.vector = so[[cluster.name]][,1], cluster.prefix = "cl")
cluster.name = sprintf("Cluster_%s_Res4",prefix)
so[[cluster.name]][,1] = RenameCluster(cluster.vector = so[[cluster.name]][,1], cluster.prefix = "cl")

# write.table(data.frame(colnames(so), so@meta.data[,grepl("Cluster_totalvi_20241014_miniverse_rmIGTsample_Res", colnames(so@meta.data))]), file = sprintf("%s/Cluster_totalvi_20241014_miniverse_rmIGTsample_Res.csv", prefix), sep = ",", quote = F, row.names = F, col.names = F)

```

UMAP/MDE with clusters
```{r}
# pdf(sprintf("%s/UMAP_Clusters.pdf", prefix), 20, 20, useDingbats = F)
# for (i in c(0.25,0.5,0.75,1,1.5,2,3,4)) {
#     MyPlots(seurat_object = so, 
#         dim1 = so[[reduc]]@cell.embeddings[,1], 
#         dim2 = so[[reduc]]@cell.embeddings[,2], 
#         color_by = sprintf("Cluster_%s_Res%s",prefix,i), 
#         split_by1 = "organ_simplified",
#         split_by2 =  NULL, 
#         genes = NULL, 
#         cluster_key = NULL, mypal = mypal) #glasbey()
# }
# dev.off()

# pdf(sprintf("%s/UMAP_Clusters.pdf", prefix), 20, 20, useDingbats = F)
pdf(sprintf("%s/MDE_Clusters.pdf", prefix), 20, 20, useDingbats = F)
for (i in c(0.25,0.5,0.75,1,1.5,2,3,4)) {
    print(i)
    print(DimPlot(so, reduction = reduc, group.by = sprintf("Cluster_%s_Res%s",prefix,i), raster = T, raster.dpi = c(1024,1024), label = T) + scale_color_manual(values = mypal))
    print(DimPlot(so, reduction = reduc, group.by = sprintf("Cluster_%s_Res%s",prefix,i), split.by = "organ_simplified", ncol = trunc(sqrt(length(unique(so$organ_simplified)))), raster = T, raster.dpi = c(1024,1024), label = F, pt.size = 2) + scale_color_manual(values = mypal))
}
dev.off()

```

UMAP/MDE with genes
```{r}
# pdf(sprintf("%s/UMAP_genes.pdf", prefix), 20,20, useDingbats = F)
pdf(sprintf("%s/MDE_genes.pdf", prefix), 40,40, useDingbats = F)
DefaultAssay(so) = "RNA"
FeaturePlot(so, slot = "data", reduction = reduc, features = c("Trbc1", "Trbc2", "Trac", "Trgc1", "Trdc", "Cd4", "Cd8a", "Cd8b1", "Foxp3","Tbx21", "Rorc", "Nr4a1", "Mki67","Tox", "Pdcd1", "Havcr2", "Ctla4", "Tox", "Lag3", "Cd69", "Itgae", "Klrb1c", "Klrg1", "Cd44", "Cd69", "Ifng","Il4", "Il17a", "Il10","Il22", "Gzmb", "Prf1", "Areg", "Zbtb16"), ncol = 7, order = T, raster = T, raster.dpi = c(1024,1024))
dev.off()

# pdf(sprintf("%s/UMAP_proteins.pdf", prefix), 20,20, useDingbats = F)
pdf(sprintf("%s/MDE_proteins.pdf", prefix), 30,30, useDingbats = F)
DefaultAssay(so) = "ADT"
FeaturePlot(so[,so$cite_seq == T], reduction = reduc,slot = "data", features = c("CD3", "THY1.2", "TCRB", "TCRGD", "TCRVG1.1", "TCRVG2", "TCRVG3", "CD4","CD8A","CD8B", "CD62L", "CD44", "IL2RA.CD25", "IL7RA.CD127", "IL2RB.CD122", "CD69", "CD103", "GITR.CD357", "LAG3.CD223", "HAVCR2", "TIGIT"), ncol = 5, order = T, raster = T, raster.dpi = c(1024, 1024))
dev.off()

pdf(sprintf("%s/iNKT_MAIT.pdf", prefix), 20, 20, useDingbats = F)
DimPlot(so, reduction = reduc, group.by = "iNKT", raster = T, raster.dpi = c(1024,1024), label = T) + scale_color_manual(values = c("grey", "red"))
DimPlot(so, reduction = reduc, group.by = "MAIT", raster = T, raster.dpi = c(1024,1024), label = T) + scale_color_manual(values = c("grey", "red"))
dev.off()


```

Transfer labels from annotated clusters from the non miniverse
```{r}
set.seed(0)
ref = so_orig[,sample(colnames(so_orig), size = 50000, replace = F)]
ref = ref[, ref$annotation_level1 != "miniverse"]
# so_orig2$cell_groups = ifelse(so_orig2$annotation_level1 == "miniverse", yes = "miniverse", no = "other")
# 
# so_orig2[["RNA"]] = split(so_orig2[["RNA"]], f = so_orig2$cell_groups)

# pre-process dataset (without integration)
so_orig2 = NormalizeData(so_orig2)
so_orig2 = FindVariableFeatures(so_orig2)
so_orig2 = ScaleData(so_orig2)
so_orig2 = RunPCA(so_orig2)
so_orig2 = FindNeighbors(so_orig2, dims = 1:30)
so_orig2 = FindClusters(so_orig2)

so_orig2 = RunUMAP(so_orig2, dims = 1:30)
DimPlot(so_orig2, group.by = c("cell_groups"))

so_orig2 = IntegrateLayers(object = so_orig2, method = CCAIntegration, orig.reduction = "pca",
    new.reduction = "integrated.cca", verbose = FALSE)
# so_orig2 = FindNeighbors(so_orig2, reduction = "integrated.cca", dims = 1:30)
# so_orig2 = FindClusters(so_orig2)

so_orig2 = RunUMAP(so_orig2, reduction = "integrated.cca", dims = 1:30)
DimPlot(so_orig2, group.by = c("annotation_level1"))

# select two technologies for the query datasets
pancreas.query <- subset(panc8, tech %in% c("fluidigmc1", "celseq"))
pancreas.query <- NormalizeData(pancreas.query)
pancreas.anchors <- FindTransferAnchors(reference = pancreas.ref, query = pancreas.query, dims = 1:30,
    reference.reduction = "pca")
predictions <- TransferData(anchorset = pancreas.anchors, refdata = pancreas.ref$celltype, dims = 1:30)
pancreas.query <- AddMetaData(pancreas.query, metadata = predictions)

```


--> annotation_level2_miniverse.csv
```{r}
library(cluster)
cluster.names = sprintf("Cluster_%s_Res%s",prefix,c(0.25, 0.5, 0.75, 1,2,3,4))
#cluster.name = sprintf("Cluster_%s_Res0.75",prefix)

#1 in all the data
so_sub = so#so[,sample(colnames(so), 50000,replace = F)]
reduc.dist = dist(so_sub[[prefix]]@cell.embeddings)

pdf(sprintf("%s/Cluster_silhouettes.pdf", prefix), 10, 10, useDingbats = F)
for (cluster.name in cluster.names) {
    print(cluster.name)
    cl = as.integer(so_sub[[cluster.name]][,1])
    sil = silhouette(cl, dist = reduc.dist)
    clust.col = mypal
    sil.cols = clust.col[ifelse(sil[,3] > 0, sil[,1], sil[,2])]
    sil.cols = sil.cols[order(-sil[,1], sil[,3])]
    plot(sil,
         main = sprintf("Silhouette in %s", cluster.name),
         border = sil.cols,
         col = sil.cols,
         do.col.sort = FALSE)
}
dev.off()

so$miniverse_clusters = "unknown"
so$miniverse_clusters[so$Cluster_totalvi_20241014_miniverse_rmIGTsample_Res0.25 == "cl1"] = "miniverse_cl1" #CD4 resting
so$miniverse_clusters[so$Cluster_totalvi_20241014_miniverse_rmIGTsample_Res0.25 %in% c("cl3") & so[["mde_totalvi_20241014_miniverse_rmIGTsample"]]@cell.embeddings[,1]<1] = "miniverse_cl2" #CD4 act 
so$miniverse_clusters[so$Cluster_totalvi_20241014_miniverse_rmIGTsample_Res0.25 %in% c("cl3") & so[["mde_totalvi_20241014_miniverse_rmIGTsample"]]@cell.embeddings[,1]>1] = "miniverse_cl3" #"Treg" 
so$miniverse_clusters[so$Cluster_totalvi_20241014_miniverse_rmIGTsample_Res0.25 %in% c("cl4") & so[["mde_totalvi_20241014_miniverse_rmIGTsample"]]@cell.embeddings[,1]>0] = "miniverse_cl4" #"Treg" 
so$miniverse_clusters[so$Cluster_totalvi_20241014_miniverse_rmIGTsample_Res0.25 == "cl2"] = "miniverse_cl5" #CD8AB resting
so$miniverse_clusters[so$Cluster_totalvi_20241014_miniverse_rmIGTsample_Res0.5 == "cl14"] = "miniverse_cl6" #CD8AB act
so$miniverse_clusters[so$Cluster_totalvi_20241014_miniverse_rmIGTsample_Res1.5 == "cl13"] = "miniverse_cl7" #CD8AB act
so$miniverse_clusters[so$Cluster_totalvi_20241014_miniverse_rmIGTsample_Res0.25 == "cl6"] = "miniverse_cl8" #CD8AB act or rather unconventional CD8AB?
so$miniverse_clusters[so$Cluster_totalvi_20241014_miniverse_rmIGTsample_Res0.25 == "cl10"] = "miniverse_cl9" #Tgd DN
so$miniverse_clusters[so$Cluster_totalvi_20241014_miniverse_rmIGTsample_Res0.25 == "cl5" & so[["mde_totalvi_20241014_miniverse_rmIGTsample"]]@cell.embeddings[,1] < -1] = "miniverse_cl10" #Tgd DN
so$miniverse_clusters[so$Cluster_totalvi_20241014_miniverse_rmIGTsample_Res0.25 == "cl5" & so[["mde_totalvi_20241014_miniverse_rmIGTsample"]]@cell.embeddings[,1] > -1] = "miniverse_cl11" #Tab CD8AA
so$miniverse_clusters[so$Cluster_totalvi_20241014_miniverse_rmIGTsample_Res0.5 == "cl6"] = "miniverse_cl12" #Tgd CD8aa
so$miniverse_clusters[so$Cluster_totalvi_20241014_miniverse_rmIGTsample_Res0.25 == "cl8" & so[["mde_totalvi_20241014_miniverse_rmIGTsample"]]@cell.embeddings[,2] < 0] = "miniverse_cl13" #DP prolif
so$miniverse_clusters[so$Cluster_totalvi_20241014_miniverse_rmIGTsample_Res0.25 == "cl8" & so[["mde_totalvi_20241014_miniverse_rmIGTsample"]]@cell.embeddings[,2] > 0] = "miniverse_cl14" #DP
so$miniverse_clusters[so$Cluster_totalvi_20241014_miniverse_rmIGTsample_Res0.25 == "cl4" & so[["mde_totalvi_20241014_miniverse_rmIGTsample"]]@cell.embeddings[,1] < 0] = "miniverse_cl15" #other unconventional
so$miniverse_clusters[so$Cluster_totalvi_20241014_miniverse_rmIGTsample_Res0.25 == "cl9"] = "miniverse_cl16" #iNKT
so$miniverse_clusters[so$Cluster_totalvi_20241014_miniverse_rmIGTsample_Res0.25 == "cl7"] = "miniverse_cl17" #thymocytes

extract_numeric = function(x) {
  as.numeric(gsub("\\D", "", x))
}
ordered_vector = unique(so$miniverse_clusters)[order(extract_numeric(unique(so$miniverse_clusters)))] %>% as.character()
new_cluster_names = paste0("miniverse_cl", seq_along(ordered_vector))
names(new_cluster_names) = ordered_vector
so$miniverse_clusters_renamed = as.character(new_cluster_names[so$miniverse_clusters])

ordered_vector = unique(so$miniverse_clusters_renamed)[order(extract_numeric(unique(so$miniverse_clusters_renamed)))] %>% as.character()
so$miniverse_clusters = factor(so$miniverse_clusters_renamed, levels = ordered_vector)
so$miniverse_clusters_renamed = NULL

# so$miniverse_clusters = sprintf("miniverse_%s",so$miniverse_clusters)
so$miniverse_clusters = factor(so$miniverse_clusters)
any(is.na(so$miniverse_clusters))

table(so$miniverse_clusters)

# so$miniverse_clusters[so$Cluster_totalvi_20241014_miniverse_rmIGTsample_Res0.25 == "cl5" & so$Cluster_totalvi_20241014_miniverse_rmIGTsample_Res0.75 == "cl18"] = "miniverse_cl6" #Trm
# so$test = so$miniverse_clusters %in% "unknown"
# DimPlot(so, reduction = reduc, group.by = "test", raster = F, order = T) + scale_color_manual(values = mypal)

so$miniverse_clusters_annot = "unknown"
so$miniverse_clusters_annot[so$miniverse_clusters %in% "miniverse_cl1"] = "CD4_resting"
so$miniverse_clusters_annot[so$miniverse_clusters %in% "miniverse_cl2"] = "CD4_activated"
so$miniverse_clusters_annot[so$miniverse_clusters %in% "miniverse_cl3"] = "Treg"
so$miniverse_clusters_annot[so$miniverse_clusters %in% "miniverse_cl4"] = "Treg"
so$miniverse_clusters_annot[so$miniverse_clusters %in% "miniverse_cl5"] = "CD8AB_resting"
so$miniverse_clusters_annot[so$miniverse_clusters %in% "miniverse_cl6"] = "CD8AB_activated"
so$miniverse_clusters_annot[so$miniverse_clusters %in% "miniverse_cl7"] = "CD8AB_activated"
so$miniverse_clusters_annot[so$miniverse_clusters %in% "miniverse_cl8"] = "CD8AB_activated_or_unconv"
so$miniverse_clusters_annot[so$miniverse_clusters %in% "miniverse_cl9"] = "unconv_Tgd_DN"
so$miniverse_clusters_annot[so$miniverse_clusters %in% "miniverse_cl10"] = "unconv_Tgd_DN"
so$miniverse_clusters_annot[so$miniverse_clusters %in% "miniverse_cl11"] = "unconv_Tab_CD8AA"
so$miniverse_clusters_annot[so$miniverse_clusters %in% "miniverse_cl12"] = "unconv_Tgd_CD8AA"
so$miniverse_clusters_annot[so$miniverse_clusters %in% "miniverse_cl13"] = "unconv_DP_prolif"
so$miniverse_clusters_annot[so$miniverse_clusters %in% "miniverse_cl14"] = "unconv_DP"
so$miniverse_clusters_annot[so$miniverse_clusters %in% "miniverse_cl15"] = "unconv"
so$miniverse_clusters_annot[so$miniverse_clusters %in% "miniverse_cl16"] = "unconv_iNKT"
so$miniverse_clusters_annot[so$miniverse_clusters %in% "miniverse_cl17"] = "unconv_thymocytes"


# pdf(sprintf("%s/UMAP_Clusters_Final.pdf", prefix), 20, 20, useDingbats = F)
pdf(sprintf("%s/MDE_Clusters_Final.pdf", prefix), 20, 20, useDingbats = F)
DimPlot(so, reduction = reduc, group.by = "miniverse_clusters", raster = T, raster.dpi = c(1024,1024), label = T) + scale_color_manual(values = mypal)
DimPlot(so, reduction = reduc, group.by = "miniverse_clusters_annot", raster = T, raster.dpi = c(1024,1024), label = T) + scale_color_manual(values = mypal)
DimPlot(so, reduction = reduc, group.by = "miniverse_clusters", split.by = "miniverse_clusters", ncol = 4, raster = T, raster.dpi = c(512,512), label = T) + scale_color_manual(values = mypal)
DimPlot(so, reduction = reduc, group.by = "miniverse_clusters", split.by = "organ_simplified", ncol = trunc(sqrt(length(unique(so$organ_simplified)))), raster = T, raster.dpi = c(512,512), label = F, pt.size = 2) + scale_color_manual(values = mypal)
DimPlot(so, reduction = "mde_totalvi_20241014_miniverse", group.by = "miniverse_clusters", raster = T, raster.dpi = c(1024,1024), label = T) + scale_color_manual(values = mypal)
dev.off()

write.table(data.frame(colnames(so), miniverse_clusters = so$miniverse_clusters, miniverse_clusters_annot = so$miniverse_clusters_annot), file = "../annotation_level2_miniverse.csv", sep = ",", quote = F, row.names = F, col.names = F)

so_sub = so[,colnames(so_sub)]
cluster.name =  "miniverse_clusters"
pdf(sprintf("%s/%s_silhouette.pdf", prefix, cluster.name), 10, 10, useDingbats = F)
cl = as.integer(so_sub[[cluster.name]][,1])
sil = silhouette(cl, dist = reduc.dist)
clust.col = mypal
sil.cols = clust.col[ifelse(sil[,3] > 0, sil[,1], sil[,2])]
sil.cols = sil.cols[order(-sil[,1], sil[,3])]
plot(sil,
     main = sprintf("Silhouette in %s", cluster.name),
     border = sil.cols,
     col = sil.cols,
     do.col.sort = FALSE)
dev.off()

write.table(data.frame(colnames(so), miniverse_clusters = so$miniverse_clusters, miniverse_clusters_annot = so$miniverse_clusters_annot), file = "../annotation_level2_miniverse.csv", sep = ",", quote = F, row.names = F, col.names = F)

```

Robustness of clusters with silhouette analysis for cluster.name at all resolution + deeper analysis for resolution 0.75 and 0.5 (the optimum)
in all dataset
in each organ
in each IGT
```{r}
SilhouetteSummaryPlots = function(sil_list = sil_list, column_name = column_name, categories = categories, cluster.name = cluster.name) {
    message("plot silhouettes")
    for (i in categories) {
        print(i)
        sil = sil_list[[i]]
        if (length(sil) >1) {
            # prepare colours:
            clust.col <- mypal#scater:::.get_palette("tableau10medium") # hidden scater colours
            sil.cols <- clust.col[ifelse(sil[,3] > 0, sil[,1], sil[,2])]
            sil.cols <- sil.cols[order(-sil[,1], sil[,3])]
            plot(sil,
                 main = sprintf("Silhouette in %s in %s:%s", cluster.name, i, column_name),
                 border = sil.cols,
                 col = sil.cols,
                 do.col.sort = FALSE)
        }
        
    }
    
    message("Plot Boxplot")
    sil_summary_list = list()
    for (i in categories) {
        sil = sil_list[[i]]
        if (length(sil) >1) {
            sil_summary_list[[i]] = as.data.frame(sil) %>% 
                group_by(cluster) %>% 
                summarise(mean_sil_width = mean(sil_width),
                          total = n(),  # Count total observations in each cluster
                          positive_count = sum(sil_width > 0),  # Count negative silhouette widths
                          percentage_positive = (positive_count / total) * 100)     %>%  # Calculate percentage) 
                as.data.frame()
            sil_summary_list[[i]][,column_name] = i
        }
    }
    sil_summary = Reduce(rbind,sil_summary_list)
    
    #sil_summary[sample(size = 20, x = 1:nrow(sil_summary),replace = F),]
    
    averaged_data <- sil_summary %>%
        group_by(cluster, column_name = sil_summary[,column_name]) %>%
        summarise(ncells = mean(total), mean_of_means = mean(mean_sil_width, na.rm = TRUE), mean_of_percentage_positive = mean(percentage_positive, na.rm = TRUE), .groups = 'drop') %>% data.frame()
    
    p = ggplot(data = averaged_data, aes(x = as.factor(cluster), y = mean_of_means)) +
        geom_boxplot(outlier.shape = NA, alpha = 0.7, width = 0.7, color = "black") +  # Draw boxplots
        geom_point(aes(color = column_name, size = ncells > 10), position = position_jitter(width = 0.2, height = 0), alpha = 0.6) +  # Add jittered points
        scale_color_manual(values=mypal) +  # Color scheme for points
        labs(title = "Boxplot of Average Mean Silhouette Width per Cluster",
             x = "Cluster",
             y = "Average of Mean Silhouette Width",
             color = column_name) +
        theme_minimal() +
        theme(axis.text.x = element_text(angle = 45, hjust = 1),  # Improve x-axis label readability
              legend.position = "right")  # Adjust the legend position
    print(p)
    
    message("Plot Individual")
    p = ggplot(data = averaged_data, aes(x = as.factor(cluster), y = mean_of_means,group = column_name, color = column_name)) +
        geom_line() +  # Add lines to connect points
        #geom_point(size = 3, alpha = 0.8) +  # Add points
        geom_point(alpha = 0.8, aes(size = ncells >10)) +
        scale_color_manual(values=mypal) +  # Use a distinguishable color palette
        labs(title = "Line Plot of Average Mean Silhouette Width per Cluster",
             x = "Cluster",
             y = "Average of Mean Silhouette Width",
             color = column_name) +
        theme_minimal() +
        theme(axis.text.x = element_text(angle = 45, hjust = 1),
              legend.position = "right") 
        
    print(p)
    print(p + facet_wrap(~column_name))
    
     p = ggplot(data = averaged_data, aes(x = as.factor(cluster), y = mean_of_percentage_positive)) +
        geom_boxplot(outlier.shape = NA, alpha = 0.7, width = 0.7, color = "black") +  # Draw boxplots
        geom_point(aes(color = column_name, size = ncells >10), position = position_jitter(width = 0.2, height = 0), alpha = 0.6) +  # Add jittered points
        scale_color_manual(values=mypal) +  # Color scheme for points
        labs(title = "Boxplot of Average Percentage of Cells with Positive Width per Cluster",
             x = "Cluster",
             y = "Average of Percentage of Cells with Positive Width",
             color = column_name) +
        theme_minimal() +
        theme(axis.text.x = element_text(angle = 45, hjust = 1),  # Improve x-axis label readability
              legend.position = "right")  # Adjust the legend position
    print(p)
    
    message("Plot Individual")
    p = ggplot(data = averaged_data, aes(x = as.factor(cluster), y = mean_of_percentage_positive, group = column_name, color = column_name)) +
        geom_line() +  # Add lines to connect points
        #geom_point(size = 3, alpha = 0.8) +  # Add points
        geom_point(alpha = 0.8, aes(size = ncells >10)) +
        scale_color_manual(values=mypal) +  # Use a distinguishable color palette
        labs(title = "Line Plot of Average Percentage of Cells with Positive per Cluster",
             x = "Cluster",
             y = "Average of Percentage of Cells with Positive Width",
             color = column_name) +
        theme_minimal() +
        theme(axis.text.x = element_text(angle = 45, hjust = 1),
              legend.position = "right") 
        
    print(p)
    print(p + facet_wrap(~column_name))
}

#2 in each organ_simplified: how much does it fit each organ?
so2 = so[,!is.na(so$organ_simplified)]

sil_list = list()
column_name = "organ_simplified"
categories = unique(so2[[column_name]][,1]) %>% na.omit() %>% as.character()
cluster.name =  "miniverse_clusters"

sil_list = list()
for (i in categories) {
    print(i)
    reduc.dist = dist(so2[[prefix]]@cell.embeddings[so2[[column_name]] == i,])
    cl = as.integer(so2[[cluster.name]][so2[[column_name]] == i,1])
    #cl = as.integer(so2[[cluster.name]][so2[[column_name]] == i,1])
    sil_list[[i]] = silhouette(cl, dist = reduc.dist)
}

pdf(sprintf("%s/%s_silhouettes_by%s.pdf", prefix, cluster.name, column_name), 20, 20, useDingbats = F)
SilhouetteSummaryPlots(sil_list = sil_list, column_name = column_name, categories = categories, cluster.name = cluster.name)
dev.off()

#3. By IGT
sil_list = list()
column_name = "IGT"
categories = levels(so[[column_name]][,1])
cluster.name =  "miniverse_clusters"

for (i in categories) {
    print(i)
    reduc.dist = dist(so[[prefix]]@cell.embeddings[so[[column_name]] == i,])
    cl = as.integer(so[[cluster.name]][so[[column_name]] == i,1])
    sil_list[[i]] = silhouette(cl, dist = reduc.dist)
}

pdf(sprintf("%s/%s_silhouettes_by%s.pdf", prefix, cluster.name, column_name), 40, 40, useDingbats = F)
SilhouetteSummaryPlots(sil_list = sil_list, column_name = column_name, categories = categories, cluster.name = cluster.name)
dev.off()

```

```{r}
DefaultAssay(so) = "ADT"
so[['RNA']]$data = NULL
DefaultAssay(so) = "RNA"
so[['ADT']]$data = NULL

saveRDS(so, file = "/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/miniverse/igt1_96_miniverse_20241017.Rds")

```
