---
title: "nonconv_misc"
output: html_document
date: "2025-05-28"
author: "David Zemmour"
editor_options: 
  chunk_output_type: console
---

**Initialize: load libraries and custom functions**

```{bash}
sinteractive --exclusive --time=12:00:00 --account=pi-zemmour --partition=jfkfloor2 --qos=jfkfloor2 --nodes=1 
sinteractive --exclusive --time=12:00:00 --account=pi-zemmour --partition=beagle3-bigmem --nodes=1 
sinteractive --exclusive --time=12:00:00 --account=pi-zemmour --partition=zemmour-hm --qos=zemmour --nodes=1 
sinteractive --time=03:00:00 --account=pi-zemmour --partition=caslake --nodes=1 --mem=96GB

module load python/anaconda-2022.05 
module load hdf5/1.12.0 #for BPCells but also others
module load openblas/0.3.13
module load ImageMagick/7.1.1-7
source activate /project/zemmour/david/envs/scvi_20240315
#source activate /project/zemmour/david/envs/r443
module load openblas/0.3.13

cd /project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/
R
```


```{r}
setwd("~/google_drive_uchicago/ImmgenT/analysis/data_integration/IGT1_96/")
libs = c("ZemmourLib","tidyr", "forcats", "gridExtra","patchwork", "Matrix", "Seurat", "ggplot2", "dplyr", "reshape2", "ggrastr","scattermore", "RColorBrewer", "pals", "scales", "pheatmap", "rafalib") 
sapply(libs, function(x) suppressMessages(suppressWarnings(library(x, character.only = TRUE, quietly = T, warn.conflicts  = F))))
#source("/project/jfkfloor2/zemmourlab/david/immgent/immgent_integration_git/custom_functions_david.R")

ColorRamp = rev(colorRampPalette(brewer.pal(n = 7,name = "RdYlBu"))(100))
ColorRamp = rev(colorRampPalette(c('red','white','blue'))(20))
ColorRamp = rev(rainbow(10, end = 4/6))
ColorRamp = rev(viridis(100))
ColorRamp = brewer.pal(9, "Blues")

#mycol = c("black","blue","red","darkgreen", "orange","purple", "cyan", "greenyellow",    "salmon", "magenta","pink", "tan", "brown") # c("magenta", "red",  "darkgreen", "cyan", "blue", "blue", "blue", "black", "blue", "blue", "blue", "blue", "orange")

mypal_level2 = ZemmourLib::immgent_colors$level2
mypal_level2["not classified"] = "black"
# names(mypal_level2) = gsub("nonconv", replacement = "Tz",names(mypal_level2) )
# mypal_level2 = immgent_colors$level2_option2
# mypal_level2["not classified"] = "black"
mypal_level1 = immgent_colors$level1
mypal_level1["not classified"] = "black"
# names(mypal_level1) = gsub("nonconv", replacement = "Tz",names(mypal_level1) )
mypal_organ = immgent_colors$organ_simplified
mypal_level2group = c("resting" = "blue", "activated" = "red", "miniverse" = "darkgreen", "proliferating" = "black", "other" = "grey", "preT" = "grey" )

#Large Color palette
library("pals")
n = 70
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
mypal1 = unique(unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals))))
mypal1 = mypal1[-4]
mypal = c(glasbey(), polychrome(), mypal1)
names(mypal) = NULL
#pal.bands(mypal, main="Colormap suggestions")
# mypal_organ = setNames(mypal, unique(so$organ_simplified))
# mypal_organ = mypal_organ[!is.na(names(mypal_organ))]
# mypal_level1 = setNames(mypal, unique(so$annotation_level1))
# mypal_level1 = mypal_level1[!is.na(names(mypal_level1))]
# mypal_level2 = setNames(mypal, unique(so_orig$annotation_level2))
# mypal_level2 = mypal_level2[!is.na(names(mypal_level2))]
# mypal_level2["CD4_cl18"] = "blue"
mypal_igt = setNames(mypal, unique(so_orig$IGT))
mypal_igt = mypal_igt[!is.na(names(mypal_igt))]

```

```{r}
# fig_dir = "~/google_drive_uchicago/ImmgenT/20250109_freeze/paper_cosmology/figures/bits"
# fig_dir = "/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/paper_cosmology/figures/bits"

out_dir = "~/google_drive_uchicago/ImmgenT/analysis/data_integration/IGT1_96/nonconv/misc_plots" 
out_dir = "/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/nonconv/misc_plots"

```


**Most up-to-date data**
on Midway
```{r}
so_orig = readRDS("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/igt1_96_withtotalvi20260206_clean.Rds") 

so = so_orig[,so_orig@meta.data %>% filter(organ_simplified != "thymus") %>% row.names()]
so = NormalizeData(so, assay = "RNA", normalization.method = "LogNormalize")
```

on Desktop
```{r}
setwd("~/google_drive_uchicago/ImmgenT/analysis/data_integration/IGT1_96/")
so_orig = readRDS("igt1_96_withtotalvi20260206_clean_ADTonly.Rds")

so = so_orig[,so_orig@meta.data %>% filter(annotation_level1 %in% c("gdT", "CD8aa", "Tz", "DN", "DP") & organ_simplified != "thymus") %>% row.names()]
# so = NormalizeData(so, assay = "RNA", normalization.method = "LogNormalize")
# so = NormalizeData(so, assay = "ADT", normalization.method = "LogNormalize")

```

**Stats for the paper**

```{r}
so_orig@meta.data %>% filter(organ_simplified != "thymus") %>% count(annotation_level1) %>% mutate(tot = sum(n)) %>% filter(annotation_level1 %in% c("gdT", "CD8aa", "DN", "DP", "Tz")) %>% mutate(percent = n/tot*100) %>% summarize(sum(percent))

```

```{r}

so = so_orig[,so_orig@meta.data %>% filter(organ_simplified != "thymus") %>% row.names()]
mypal_level1b = mypal_level1
mypal_level1b[names(mypal_level1b) %in% c("gdT", "CD8aa", "Tz", "DN", "DP")] = c("grey", "grey", "grey", "grey", "grey")
pdf(sprintf("%s/MDE_alllevel1.pdf", out_dir), 5,5, useDingbats = F)
p = DimPlot(object = so, reduction = "mde2_totalvi_20241006", group.by = "annotation_level1", raster = T, raster.dpi = c(1024,1024)) + scale_color_manual(values = mypal_level1b) + NoGrid()
p + theme_void() + NoLegend() + ggtitle(label = NULL)
dev.off()
```


**FeatureScatter ADT**
```{r}
so = so_orig[,so_orig@meta.data %>% filter(cite_seq == T, organ_simplified != "thymus", annotation_level1 %in% c("gdT", "CD8aa", "Tz", "DN", "DP")) %>% row.names()]
so = NormalizeData(so, assay = "ADT", normalization.method = "LogNormalize")
features_list = list(c("TCRB", "TCRGD"),
                     c("CD4", "CD8B"),
                     c("CD8A", "CD8B")
                     )
thresholds = list("TCRB" = 3, "TCRGD" = 4,"CD4" = 4, "CD8B" = 5, "CD8A" = 4.5)

pdf(sprintf("%s/FeatureScatter_level1.pdf", out_dir), 3,3, useDingbats = F)
for (i in 1:length(features_list)) {
    feature1 = features_list[[i]][1]
    feature2 = features_list[[i]][2]
    th1 = thresholds[feature1][1] %>% as.numeric()
    th2 = thresholds[feature2][1] %>% as.numeric()
    p = FeatureScatter(so,slot = "data",feature1 = feature1, feature2 = feature2, group.by = "annotation_level1", raster = T, pt.size = 1, plot.cor = F) + scale_color_manual(values = mypal_level1) + geom_vline(aes(xintercept = th1), linetype="dashed", color = "brown") + geom_hline(aes(yintercept = th2), linetype="dashed", color = "brown") + NoLegend()
    print(p)
}
dev.off()

pdf(sprintf("%s/FeatureScatter_level1_split.pdf", out_dir), 10,3, useDingbats = F)
for (i in 1:length(features_list)) {
    print(i)
    feature1 = features_list[[i]][1]
    feature2 = features_list[[i]][2]
    th1 = thresholds[feature1][1] %>% as.numeric()
    th2 = thresholds[feature2][1] %>% as.numeric()
    p = FeatureScatter(so,slot = "data",feature1 = feature1, feature2 = feature2, group.by = "annotation_level1", split.by = "annotation_level1", raster = T, pt.size = 1) + scale_color_manual(values = mypal_level1) + geom_vline(aes(xintercept = th1), linetype="dashed", color = "brown") + geom_hline(aes(yintercept = th2), linetype="dashed", color = "brown") + NoLegend()
    print(p)
    
}
dev.off()

```

**FeaturePlot (Midway3) **
```{r}
so = so_orig[,so_orig@meta.data %>% filter(organ_simplified != "thymus") %>% row.names()]
so = NormalizeData(so, assay = "RNA", normalization.method = "LogNormalize")
pdf(sprintf("%s/FeaturePlot_MainMarkers_level1.pdf", out_dir), 12,15, useDingbats = F)
FeaturePlot(so, reduction = "mde2_totalvi_20241006", features = c("Foxp3", "Cd4", "Cd8b1", "Cd8a","Trdc", "Zbtb16"), order = T, cols = c("lightgrey", "blue"), raster = T,max.cutoff = "q95") + ggplot2::xlim(-2.5,2.5) + ggplot2::ylim(-2.5,2.5)#Trgc1, Trgv2
dev.off()

pdf(sprintf("%s/FeaturePlot_MainMarkers_level1_2.pdf", out_dir), 12,15, useDingbats = F)
FeaturePlot(so, reduction = "mde2_totalvi_20241006", features = c("Foxp3", "Zbtb16", "Mki67", "Cdkn2a", "Cdkn2b"), order = T, cols = c("lightgrey", "blue"), raster = T, max.cutoff = "q90",) + ggplot2::xlim(-2.5,2.5) + ggplot2::ylim(-2.5,2.5)#Trgc1, Trgv2
dev.off()

```

**TCRb x TCRgd feature plot showing cells expressing Trd Trg genes and no Tra or Trb genes (Midway3)**
```{r}
# library(Seurat)
# library(dplyr)
# library(Matrix)
# library(ggplot2)

so <- so_orig[, so_orig@meta.data %>%
  filter(
    cite_seq == TRUE,
    organ_simplified != "thymus",
  ) %>%
  rownames()
]
so <- NormalizeData(
  so,
  assay = "ADT",
  normalization.method = "LogNormalize"
)

#Find TCR genes
rna_genes <- rownames(so[["RNA"]])
tcr_patterns <- list(
  Trd = "^(Trdv|Trdd|Trdj|Trdc)",
  Trg = "^(Trgv|Trgj|Trgc)",
  Tra = "^(Trav|Traj|Trac)",
  Trb = "^(Trbv|Trbd|Trbj|Trbc)"
)
tcr_genes <- lapply(tcr_patterns, function(pat) {
  grep(pat, rna_genes, value = TRUE)
})
# print gene lists to double check
tcr_genes
sapply(tcr_genes, length)

#Define Trd/Trg RNA-positive, Tra/Trb RNA-negative cells
rna_counts <- LayerData(so, assay = "RNA", layer = "counts")

get_gene_sum <- function(g) {
  g <- intersect(g, rownames(rna_counts))
  
  if (length(g) == 0) {
    return(rep(0, ncol(rna_counts)))
  }
  
  Matrix::colSums(rna_counts[g, , drop = FALSE])
}

Trd_counts <- get_gene_sum(tcr_genes$Trd)
Trg_counts <- get_gene_sum(tcr_genes$Trg)
Tra_counts <- get_gene_sum(tcr_genes$Tra)
Trb_counts <- get_gene_sum(tcr_genes$Trb)

names(Trd_counts) <- colnames(so)
names(Trg_counts) <- colnames(so)
names(Tra_counts) <- colnames(so)
names(Trb_counts) <- colnames(so)

so$Trd_counts <- Trd_counts
so$Trg_counts <- Trg_counts
so$Tra_counts <- Tra_counts
so$Trb_counts <- Trb_counts

so$TrdTrg_pos_TraTrb_neg <- (
  (so$Trd_counts > 0 | so$Trg_counts > 0) &
    so$Tra_counts == 0 &
    so$Trb_counts == 0
)

so$TCR_RNA_class <- ifelse(
  so$TrdTrg_pos_TraTrb_neg,
  "Trd/Trg+ Tra/Trb-",
  "Other"
)
so@meta.data %>% count(annotation_level1, TCR_RNA_class)

```

Barplot % Trd/Trg+ Tra/Trb-
```{r}
library(dplyr)
library(ggplot2)

plot_df <- so@meta.data %>%
  as.data.frame() %>%
  count(annotation_level1, TCR_RNA_class) %>%
  group_by(annotation_level1) %>%
  mutate(
    total = sum(n),
    percent = 100 * n / total
  ) %>%
  ungroup() %>%
  filter(TCR_RNA_class == "Trd/Trg+ Tra/Trb-")

plot_df$annotation_level1 <- factor(
  plot_df$annotation_level1,
  levels = plot_df$annotation_level1[order(plot_df$percent, decreasing=T)]
  # levels = levels(so$annotation_level1)
)

p = ggplot(plot_df, aes(x = annotation_level1, y = percent, fill = annotation_level1)) +
  geom_col(color = "black", linewidth = 0.2) +
  geom_text(aes(label = sprintf("%.1f", percent)), vjust = -0.3, size = 3) +
  scale_fill_manual(values = mypal_level1) +
  labs(
    x = NULL,
    y = "% Trd/Trg+ Tra/Trb-"
  ) +
  theme_classic(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none"
  )

pdf(sprintf("%s/Barplot_TrdTrg_pos_TraTrb_neg_per_annotation_level1.pdf", out_dir), width = 4, height = 4, useDingbats = F)
print(p)
dev.off()

```

Feature Scatter of TCRB vs TCRGD CITE-seq with Trd/Trg+ Tra/Trb- cells
```{r}


DefaultAssay(so) <- "ADT"

# p <- FeatureScatter(
#   so,
#   slot = "data",
#   feature1 = "TCRB",
#   feature2 = "TCRGD",
#   group.by = "TCR_RNA_class",
#   raster = TRUE,
#   pt.size = 5,
#   plot.cor = FALSE
# ) +
#   scale_color_manual(
#     values = c(
#       "Other" = "grey80",
#       "Trd/Trg+ Tra/Trb-" = "red"
#     )
#   ) +
#   geom_vline(xintercept = 3, linetype = "dashed", color = "brown") +
#   geom_hline(yintercept = 4, linetype = "dashed", color = "brown") +
#   NoLegend() +
#   theme_classic()
# p

DefaultAssay(so) <- "ADT"
adt_df <- FetchData(
  so,
  vars = c("TCRB", "TCRGD", "TCR_RNA_class", "annotation_level1"),
  layer = "data"
)
adt_df$cell <- rownames(adt_df)
p1 <- ggplot() +
  geom_point_rast(
    data = adt_df %>% filter(TCR_RNA_class == "Other"),
    aes(x = TCRB, y = TCRGD),
    color = "grey80",
    size = 0.4,
    alpha = 0.5,
    raster.dpi = 300
  ) +
  geom_point_rast(
    data = adt_df %>% filter(TCR_RNA_class == "Trd/Trg+ Tra/Trb-"),
    aes(x = TCRB, y = TCRGD),
    color = "red",
    size = 0.6,
    alpha = 0.9,
    raster.dpi = 300
  ) +
  geom_vline(xintercept = 3, linetype = "dashed", color = "brown") +
  geom_hline(yintercept = 4, linetype = "dashed", color = "brown") +
  labs(
    x = "TCRB ADT",
    y = "TCRGD ADT",
    title = "TCRB vs TCRGD CITE-seq",
    subtitle = "Red: Trd/Trg RNA+ and Tra/Trb RNA-"
  ) +
  theme_classic(base_size = 14)

p2 <- ggplot() +
  geom_point_rast(
    data = adt_df %>% filter(TCR_RNA_class == "Other"),
    aes(x = TCRB, y = TCRGD),
    color = "grey80",
    size = 0.4,
    alpha = 0.5,
    raster.dpi = 300
  ) +
  geom_point_rast(
    data = adt_df %>% filter(TCR_RNA_class == "Trd/Trg+ Tra/Trb-"),
    aes(x = TCRB, y = TCRGD, color = annotation_level1),
    size = 0.6,
    alpha = 0.9,
    raster.dpi = 300
  ) +
  geom_vline(xintercept = 3, linetype = "dashed", color = "brown") +
  geom_hline(yintercept = 4, linetype = "dashed", color = "brown") +
      scale_color_manual(values = mypal_level1) +
  labs(
    x = "TCRB ADT",
    y = "TCRGD ADT",
    title = "TCRB vs TCRGD CITE-seq",
    subtitle = "Red: Trd/Trg RNA+ and Tra/Trb RNA-"
  ) +
  theme_classic(base_size = 14) + NoLegend()


pdf(sprintf("%s/FeatureScatter_TCRB_TCRGD_TrdTrg_pos_TraTrb_neg.pdf", out_dir), width = 8, height = 8, useDingbats = F)
p1
p2
dev.off()

```


**iNKT - MAIT**
```{r}
# DimPlot(so, reduction = "mde2_totalvi_20241006", group.by = "iNKT", order = T, raster = T) + scale_color_manual(values = c("grey", "red")) + ggplot2::xlim(-2.5,2.5) + ggplot2::ylim(-2.5,2.5)#Trgc1
so = so_orig[,so_orig@meta.data %>% filter(organ_simplified != "thymus") %>% row.names()]
pdf(sprintf("%s/DimPlot_iNKT_MAIT.pdf", out_dir), 10,10, useDingbats = F)
MyDimPlotHighlight(seurat_object = so, umap_to_plot = "mde2_totalvi_20241006", cells_to_highlight = so@meta.data %>% filter(iNKT) %>% row.names(),highlight_column_name = "iNKT", mycols = c("grey", "red"), print_plot1 = T, print_plot2 = F,labelclusters = F, highlight_raster = T,title = "TRAV11 
+ TRAJ18+ CDR3/alpha.junction = CVVGDRGSALGRLHF or CVVADRGSALGRLHF or CVVVDRGSALGRLHF ") #+ ggplot2::xlim(-2.5,2.5) + ggplot2::ylim(-2.5,2.5)#Trgc1
MyDimPlotHighlight(seurat_object = so, umap_to_plot = "mde2_totalvi_20241006", cells_to_highlight = so@meta.data %>% filter(MAIT) %>% row.names(),highlight_column_name = "MAIT", mycols = c("grey", "blue"), print_plot1 = T, print_plot2 = F,labelclusters = F, highlight_raster = T, title = "MAIT TRAV1 + TRAJ33 + alpha.cdr3 length=12 (beta chain indifferent)") #+ ggplot2::xlim(-2.5,2.5) + ggplot2::ylim(-2.5,2.5)#Trgc1
dev.off()

bg <- AddEmbeddingRasterBackground(
  lineage = "allT",
  interpolate = FALSE,
  color = "grey",
  alpha = 0.5
)

p1 <- Seurat::DimPlot(
  so[,so@meta.data %>% filter(iNKT) %>% row.names()],
  reduction = "mde2_totalvi_20241006",
  group.by = "iNKT",
    cols = "red",
    # alpha = 0.5,
  pt.size = 0.1,
  raster = FALSE
)
p1$layers <- c(list(bg), p1$layers)
p1 = p1 + ggplot2::coord_fixed() + 
  ggplot2::xlim(-2.5, 2.5) +
  ggplot2::ylim(-2.5, 2.5) +
  ggplot2::labs(title = "iNKT cells 
                TRAV11 + TRAJ18 + 
                CDR3/alpha.junction = CVVGDRGSALGRLHF 
                or CVVADRGSALGRLHF or CVVVDRGSALGRLHF") +
  ggplot2::theme(
    plot.title = element_text(hjust = 0.5, size = 8),
    # legend.position = "right"
  ) + NoLegend() + NoGrid() + NoAxes()

p2 <- Seurat::DimPlot(
  so[,so@meta.data %>% filter(MAIT) %>% row.names()],
  reduction = "mde2_totalvi_20241006",
  group.by = "MAIT",
    cols = "blue",
    # alpha = 0.5,
  pt.size = 0.1,
  raster = FALSE
)
p2$layers <- c(list(bg), p2$layers)
p2 = p2 + ggplot2::coord_fixed() + 
  ggplot2::xlim(-2.5, 2.5) +
  ggplot2::ylim(-2.5, 2.5) +
  ggplot2::labs(title = "MAIT cells
                TRAV1 + TRAJ33 + 
                alpha.cdr3 length=12 (beta chain indifferent)") +
  ggplot2::theme(
    plot.title = element_text(hjust = 0.5, size = 8),
    # legend.position = "right"
  ) + NoLegend() + NoGrid() + NoAxes()


pdf(sprintf("%s/DimPlot_iNKT_MAIT.pdf", out_dir), 10,10, useDingbats = F)
p1 
p2
dev.off()

```


**Stacked barplot with %  of gdT per condition_detailed_organ, CD8aa, DN, DP, descreasing order **

```{r}

prop = readRDS('/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/analysis/data_integration/IGT1_96/SampleComp/20260206/annotation_level1_PropPerSample_withMetadata.Rds')

# library(dplyr)
# library(tidyr)
# library(ggplot2)

clusters_keep <- c("gdT", "CD8aa", "DN", "DP", "Tz")

plot_df <- prop %>%
  mutate(total_minor = gdT.ncells + CD8aa.ncells + DN.ncells + DP.ncells + Tz.ncells) %>%
    filter(
    grepl("allT", target_cells_simplified),
    total_minor >= 50
  ) %>%
  group_by(condition_detailed_organ) %>%
  summarise(
    gdT = mean(gdT.prop, na.rm = TRUE),
    CD8aa = mean(CD8aa.prop, na.rm = TRUE),
    Tz = mean(Tz.prop, na.rm = TRUE),
    DN = mean(DN.prop, na.rm = TRUE),
    DP = mean(DP.prop, na.rm = TRUE),
    n_samples = n(),
    .groups = "drop"
  ) %>%
  mutate(total_prop = gdT + CD8aa + Tz + DN + DP) %>%
  arrange(desc(total_prop)) %>%
  pivot_longer(
    cols = all_of(clusters_keep),
    names_to = "annotation_level1",
    values_to = "prop"
  ) %>%
  mutate(
    annotation_level1 = factor(annotation_level1, levels = clusters_keep),
    condition_detailed_organ = factor(
      condition_detailed_organ,
      levels = unique(condition_detailed_organ)
    )
  )

p = ggplot(plot_df, aes(x = condition_detailed_organ, y = prop * 100, fill = annotation_level1)) +
  geom_col(color = "black", linewidth = 0.2) +
  scale_fill_manual(values = mypal_level1) +
  labs(
    x = NULL,
    y = "% of cells",
    fill = NULL
  ) +
  theme_classic(base_size = 6) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
p

pdf(sprintf("%s/stacked_barplot_gdT_CD8aa_DN_DP_per_condition_detailed_organ.pdf", out_dir ), width = 10, height = 6, useDingbats = F)
print(p)
dev.off()

#some stats
median(plot_df$total_prop)
max(plot_df$total_prop)


```

**Stacked barplot with %  of gdT, CD8aa, DN, DP clusters per organ, descreasing order **

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)

prop <- readRDS(
  "/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/analysis/data_integration/IGT1_96/SampleComp/20260206/annotation_level2_PropPerSample_withMetadata.Rds"
)

cl_order <- read.table(
  file = sprintf(
    "%s/Heatmap_correlation_clusters_top%dgenes_%dPCs_CLUSTERORDER.txt",
    out_dir,
    length(top_genes),
    npcs
  ),
  header = FALSE,
  stringsAsFactors = FALSE
)[, 1]

lineages_keep <- c("gdT", "CD8aa", "DN", "DP", "Tz")

cluster_cols <- grep(
  paste0("^(", paste(lineages_keep, collapse = "|"), ")\\..*\\.prop$"),
  colnames(prop),
  value = TRUE
)

cluster_names <- sub("\\.prop$", "", cluster_cols)

plot_df <- prop %>%
  filter(
    grepl("allT", target_cells_simplified),
    !is.na(organ_simplified),
    !is.na(sample_code)
  ) %>%
  select(
    sample_code,
    organ = organ_simplified,
    all_of(cluster_cols)
  ) %>%
  # Guarantees one value per sample
  distinct(sample_code, organ, .keep_all = TRUE) %>%
  pivot_longer(
    cols = all_of(cluster_cols),
    names_to = "cluster",
    values_to = "proportion"
  ) %>%
  mutate(
    cluster = sub("\\.prop$", "", cluster)
  ) %>%
  group_by(cluster, organ) %>%
  summarise(
    mean_prop = mean(proportion, na.rm = TRUE),
    n_samples = n_distinct(sample_code),
    .groups = "drop"
  ) %>%
  mutate(
    percent = 100 * mean_prop
  )

# cluster_order_use <- c(
#   cl_order[cl_order %in% unique(plot_df$cluster)],
#   setdiff(unique(plot_df$cluster), cl_order)
# )

plot_df <- plot_df %>%
    filter( cluster %in% cl_order) %>%
  mutate(
    cluster = factor(cluster, levels = rev(cl_order))
  )

organ_order <- plot_df %>%
  group_by(organ) %>%
  summarise(
    total_mean_prop = sum(mean_prop, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(desc(total_mean_prop)) %>%
  pull(organ)

plot_df <- plot_df %>%
  mutate(
    # Reverse so the most abundant organ is at the bottom of each stack
    organ = factor(organ, levels = rev(organ_order))
  )

p <- ggplot(
  plot_df,
  aes(
    x = cluster,
    y = percent,
    fill = organ
  )
) +
  geom_col(
    width = 0.85,
    color = "white",
    linewidth = 0.1
  ) +
  scale_fill_manual(
    values = mypal_organ,
    breaks = organ_order
  ) +
  scale_y_continuous(
    labels = label_number(suffix = "%"),
    expand = expansion(mult = c(0, 0.02))
  ) +
  labs(
    x = NULL,
    y = "Mean % of cells per sample",
    fill = "Organ"
  ) +
  theme_classic(base_size = 12) +
  theme(
    axis.text.x = element_text(
      angle = 90,
      hjust = 1,
      vjust = 0.5,
      size = 7
    ),
    legend.key.height = grid::unit(0.35, "cm")
  )

p + coord_flip()

pdf(sprintf("%s/stacked_barplot_gdT_CD8aa_DN_DP_per_organ.pdf", out_dir), width = 4, height = 6, useDingbats = FALSE)
p + coord_flip()
p + coord_flip() + NoLegend()

dev.off()
```





**Based on pseudobulk expression of the clusters in gdT, CD8aa, Tz, DN, DP, make a correlation heatmap**

```{r}
library(Seurat)
library(dplyr)
library(Matrix)
library(matrixStats)
library(ComplexHeatmap)
library(circlize)
library(grid)


pb = readRDS('/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/analysis/data_integration/IGT1_96/TF_analysis/20260319/igt1_96_pseudobulk_byclusterannotationlevel2_log1pCP10K.Rds')

pb@meta.data %>% count(annotation_level2)

lineages_keep <- c("gdT", "CD8aa", "Tz", "DN", "DP")

# Get log1pCP10K matrix
mat <- LayerData(pb, assay = "RNA", layer = "data")

# Identify pseudobulk columns to keep
if ("annotation_level1" %in% colnames(pb@meta.data)) {
  keep_cols <- rownames(pb@meta.data)[pb@meta.data$annotation_level1 %in% lineages_keep]
  lineage <- pb@meta.data[keep_cols, "annotation_level1"]
} else {
  keep_cols <- colnames(pb)[grepl("^(gdT|CD8aa|Tz|DN|DP)\\.", colnames(pb))]
  lineage <- sub("\\..*$", "", keep_cols)
}

mat_sub <- mat[, keep_cols, drop = FALSE]

# Convert to dense; this is fine here because there are only ~91 pseudobulk samples
mat_sub <- as.matrix(mat_sub)

# Removing TCR, mt, ribo, Gm, and Rik genes and genes that are not expressed
tcr_genes = grepl(x = rownames(mat_sub), pattern = "Trbv|Trbd|Trbj|Trbc|Trav|Traj|Trac|Trgv|Trgd|Trgj|Trgc|Trdv|Trdj|Trdc")
gm_rik_genes = grepl(x = rownames(mat_sub), pattern = "Gm|Rik$|\\-ps$")
ribo_genes = grepl(x = rownames(mat_sub), pattern = "Rpl|Rps|Mrpl|Mrps|Rsl")
mt_genes = grepl(x = rownames(mat_sub), pattern = "^mt-")
genes_not_expressed = rowSums(mat_sub) == 0
genes_to_keep = !(tcr_genes | gm_rik_genes | ribo_genes | mt_genes | genes_not_expressed)
cat("Number of genes to keep:", sum(genes_to_keep), "\n")

mat_sub <- mat_sub[genes_to_keep, , drop = FALSE]

# Keep genes expressed in at least 2 pseudobulk clusters
keep_genes <- rowSums(mat_sub > 0.1) >= 2
mat_sub <- mat_sub[keep_genes, , drop = FALSE]

# Select top variable genes
# library(dplyr)
# library(ggplot2)
# library(tibble)

gene_var <- matrixStats::rowVars(mat_sub + 0.1)
names(gene_var) <- rownames(mat_sub)

gene_mean <- matrixStats::rowMeans2(mat_sub + 0.1)
names(gene_mean) <- rownames(mat_sub)

dispersion <- gene_var / gene_mean
names(dispersion) <- rownames(mat_sub)
# remove bad values
dispersion <- dispersion[is.finite(dispersion) & dispersion > 0]

gene_stats <- data.frame(
  gene = rownames(mat_sub),
  gene_mean = gene_mean,
  gene_var = gene_var,
  dispersion = dispersion
) %>%
  filter(
    gene_mean > 0.15,
    dispersion > 0,
    is.finite(dispersion)
  )

top_n <- 1000

top_genes <- gene_stats %>%
  arrange(desc(dispersion)) %>%
  slice_head(n = top_n) %>%
  pull(gene)

plot_df <- tibble(
  gene = rownames(mat_sub),
  gene_mean = gene_mean[rownames(mat_sub)],
  gene_var = gene_var[rownames(mat_sub)],
  dispersion = gene_var / gene_mean,
  top_gene = gene %in% top_genes
) %>%
  filter(
    gene_mean > 0,
    dispersion > 0,
    is.finite(dispersion)
  )

ggplot(plot_df, aes(x = gene_mean, y = dispersion)) +
  geom_point(
    data = plot_df %>% filter(!top_gene),
    color = "grey70",
    size = 0.7,
    alpha = 0.5
  ) +
  geom_point(
    data = plot_df %>% filter(top_gene),
    color = "red",
    size = 0.8,
    alpha = 0.8
  ) +
  scale_x_log10() +
  scale_y_log10() +
  labs(
    x = "Mean expression",
    y = "Dispersion (variance / mean)",
    title = paste0("Top ", length(top_genes), " genes by dispersion")
  ) +
  theme_classic(base_size = 14)

mat_var <- mat_sub[top_genes, , drop = FALSE]

#Option 1: direct correlation on top variable genes / NOT USED
# cor_mat_direct <- cor(
#   mat_var,
#   method = "pearson",
#   use = "pairwise.complete.obs"
# )

#Option 2: PCA-denoised correlation
# PCA expects clusters as rows and genes as columns
X <- t(mat_var)

pca <- prcomp(
  X,
  center = TRUE,
  scale. = TRUE
)

var_explained <- pca$sdev^2 / sum(pca$sdev^2)

# choose PCs explaining ~80% variance, capped at 20
npcs <- which(cumsum(var_explained) >= 0.80)[1]
npcs <- min(npcs, 20)
npcs <- max(npcs, 3)

npcs
sum(var_explained[1:npcs])

# reconstruct denoised expression from selected PCs and do correlation / NOT USED
# Xhat <- pca$x[, 1:npcs, drop = FALSE] %*%
#   t(pca$rotation[, 1:npcs, drop = FALSE])
# cor_mat <- cor(
#   t(Xhat),
#   method = "pearson",
#   use = "pairwise.complete.obs"
# )
# rownames(cor_mat) <- rownames(X)
# colnames(cor_mat) <- rownames(X)

pc_scores <- pca$x[, 1:npcs, drop = FALSE]
cor_mat <- cor(
  t(pc_scores),
  method = "pearson",
  use = "pairwise.complete.obs"
)
rownames(cor_mat) <- rownames(pc_scores)
colnames(cor_mat) <- rownames(pc_scores)


#Plot correlation heatmap

library(ComplexHeatmap)
library(circlize)
library(grid)

col_fun <- circlize::colorRamp2(
  c(-1, 0, 1),
  c("#4575B4", "white", "#D73027")
)

hc <- hclust(as.dist(1 - cor_mat))
ord <- hc$order
ordered_names <- rownames(cor_mat)[ord]

cor_mat_ordered <- cor_mat[ordered_names, ordered_names]

Heatmap(
  cor_mat_ordered,
  name = "PC\ncorrelation",
  col = col_fun,
  cluster_rows = FALSE,
  cluster_columns = FALSE,
  show_row_names = TRUE,
  show_column_names = TRUE,
  row_names_gp = gpar(fontsize = 7),
  column_names_gp = gpar(fontsize = 7),
  column_names_rot = 45
)


pdf(sprintf("%s/Heatmap_correlation_clusters_top%dgenes_%dPCs.pdf", out_dir, length(top_genes), npcs), width = 7.5, height = 7, useDingbats = F)
ht = Heatmap(
  cor_mat,
  name = "PC\ncorrelation",
  col = col_fun,
  cluster_rows = TRUE,
  cluster_columns = TRUE,
  show_row_names = TRUE,
  show_column_names = TRUE,
  row_names_gp = gpar(fontsize = 7),
  column_names_gp = gpar(fontsize = 7),
  column_names_rot = 45,
  column_title = paste0(
    "Pseudobulk correlation based on top PCs\n",
    length(top_genes), " variable genes, ", npcs, " PCs"
  )
)
print(ht)
dev.off()

ht_drawn <- draw(ht)

cl_order = rownames(cor_mat)[row_order(ht_drawn)]
write.table(rownames(cor_mat)[row_order(ht_drawn)], file = sprintf("%s/Heatmap_correlation_clusters_top%dgenes_%dPCs_CLUSTERORDER.txt", out_dir, length(top_genes), npcs), quote = FALSE, row.names = FALSE, col.names = FALSE)

```


Make heatmap with clusters in c("gdT", "CD8aa", "Tz", "DN", "DP") 
Data from from igt1_96_pseudobulk_byclusterannotationlevel2_log1pCP10K.Rds
Use those genes: effmol = read.table(file = "/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/EffMol_analysis/effmol_GO0005615_And_CBList.txt")[,1]
keep order as saved in cl_order = read.table(file = sprintf("%s/Heatmap_correlation_clusters_top%dgenes_%dPCs_CLUSTERORDER.txt", out_dir, length(top_genes), npcs), header = F)[,1]

```{r}
Gene_AUC2 <- function(
        pseudobulk,
        genes,
        out_dir,
        mypal = NULL,
        order_groups = levels(so$annotation_level2),
        mypal_groups = mypal_level2,
        n_top = 20,
        expr_threshold = 0.5,
        cols_heatmap_gradient = rev(colorRampPalette(brewer.pal(n = 7, name = "RdYlBu"))(100)),
        pdf_width = 7,
        pdf_height = 7,
        ann_cols = c("annotation_level1", "annotation_level2_group"),
        ann_palettes = NULL,
        cl_order_file = NULL,
        cl_order = NULL
) {
    
    stopifnot("RNA" %in% names(pseudobulk))
    if (!dir.exists(out_dir)) dir.create(out_dir, recursive = TRUE)
    
    # 1. Extract matrix for genes of interest
    mat <- pseudobulk[["RNA"]]$data[genes, , drop = FALSE]
    
    # 2. Normalize each gene by its row sum
    row_sums <- rowSums(mat)
    row_sums[row_sums == 0] <- NA
    mat_norm <- mat / row_sums
    
    # 3. Filter genes: at least one cluster with expression > expr_threshold
    filt_expr <- rownames(mat)[rowSums(mat > expr_threshold, na.rm = TRUE) >= 1]
    mat_norm_filt <- mat_norm[filt_expr, , drop = FALSE]
    
    # 4. Long format + cumulative curves
    data_long <- mat_norm_filt |>
        as.data.frame() |>
        tibble::rownames_to_column(var = "Gene") |>
        tidyr::pivot_longer(cols = -Gene, names_to = "Group", values_to = "Expression") |>
        dplyr::group_by(Gene) |>
        dplyr::arrange(dplyr::desc(Expression), .by_group = TRUE) |>
        dplyr::mutate(CumulativeExpression = cumsum(Expression), Rank = dplyr::row_number()) |>
        dplyr::ungroup()
    
    # 5. AUC per gene
    gene_auc <- data_long |>
        dplyr::group_by(Gene) |>
        dplyr::summarise(auc = sum(CumulativeExpression) / dplyr::n(), .groups = "drop") |>
        dplyr::arrange(dplyr::desc(auc))
    
    utils::write.table(
        gene_auc,
        file = sprintf("%s/AUC.txt", out_dir),
        quote = FALSE, col.names = TRUE, row.names = FALSE, sep = "\t"
    )
    
    top_genes <- head(gene_auc$Gene, n_top)
    
    # =========================
    # 6. Cumulative curves PDF
    # =========================
    
    pdf(sprintf("%s/cum_curves.pdf", out_dir),
        7, 7, useDingbats = FALSE)
    
    p1 <- ggplot2::ggplot(
        data_long,
        ggplot2::aes(x = Rank, y = CumulativeExpression, group = Gene)
    ) +
        ggplot2::geom_line(size = 0.5, alpha = 0.1)
    
    p2 <- ggplot2::geom_line(
        data = subset(data_long, Gene %in% top_genes),
        mapping = ggplot2::aes(
            x = Rank, y = CumulativeExpression,
            group = Gene, color = Gene
        ),
        size = 1,
        alpha = 1
    )
    
    p3 <- ggrepel::geom_label_repel(
        data = data_long |>
            dplyr::filter(Gene %in% top_genes) |>
            dplyr::group_by(Gene) |>
            dplyr::slice_sample(n = 1) |>
            dplyr::ungroup(),
        ggplot2::aes(
            x = Rank, y = CumulativeExpression,
            label = Gene, color = Gene
        ),
        size = 4,
        segment.size = 0.2,
        segment.color = "grey",
        max.overlaps = 50
    )
    
    # Colors for lines
    if (is.null(mypal)) {
        cols <- grDevices::rainbow(length(top_genes))
        names(cols) <- top_genes
    } else {
        cols <- mypal
    }
    
    print(
        p1 + p2 + p3 +
            ggplot2::scale_color_manual(values = cols) +
            ggplot2::labs(
                title = "Cumulative Expression",
                x = "Ranked Groups by Expression",
                y = "Cumulative Expression"
            ) +
            ggplot2::theme_minimal() +
            ZemmourLib::NoLegend() +
            ZemmourLib::NoGrid()
    )
    
    dev.off()
    
    
    # =========================
    # 7. Barplots for top genes
    # =========================
    if (!is.null(so) && !is.null(mypal_groups)) {
        pdf(sprintf("%s/barplots_topgenes.pdf", out_dir),
            pdf_width, pdf_height, useDingbats = FALSE)
        
        for (g in top_genes) {
            gene_to_plot <- g
            data <- data.frame(
                Group    = colnames(pseudobulk[["RNA"]]$data),
                Expression = pseudobulk[["RNA"]]$data[gene_to_plot, ]
            )
            data$Group <- gsub("\\-", "_", data$Group)
            data$Group <- factor(data$Group, levels = order_groups)
            # data$level1  <- so$annotation_level1[match(data$Group, so$annotation_level1)]
            
            p <- ggplot2::ggplot(data, ggplot2::aes(x = Group, y = Expression, fill = Group)) +
                ggplot2::geom_bar(stat = "identity") +
                ggplot2::scale_fill_manual(values = mypal_groups) +
                ggplot2::labs(
                    title = sprintf("%s Expression in Groups", gene_to_plot),
                    x = "Group", y = "Expression"
                ) +
                ggplot2::theme_minimal() +
                ggplot2::theme(
                    axis.text.x  = ggplot2::element_text(size = 8, angle = 90,
                                                         hjust = 1, vjust = 0.5),
                    axis.title.x = ggplot2::element_text(size = 10),
                    axis.title.y = ggplot2::element_text(size = 10)
                )
            
            print(p + ZemmourLib::NoGrid())
            print(p + NoLegend() + ZemmourLib::NoGrid())
        }
        
        dev.off()
    }
    
    
    # =========================
    # 8. Heatmaps + annotations
    # =========================
    pdf(sprintf("%s/heatmap_topgenes.pdf", out_dir),
        pdf_width, pdf_height, useDingbats = FALSE)
    
    # --- Build annotation_col from pseudobulk@meta.data ---
    annotation_col <- NULL
    annotation_colors <- NULL
    
    if (!is.null(ann_cols) && length(ann_cols) > 0) {
        
        # Ensure meta.data rownames match matrix columns
        md <- pseudobulk@meta.data
        cols <- colnames(mat)
        
        # Try direct match first
        if (all(cols %in% rownames(md))) {
            md2 <- md[cols, , drop = FALSE]
        } else {
            # fallback: try swapping '-' and '_' (common mismatch)
            cols_alt1 <- gsub("-", "_", cols)
            cols_alt2 <- gsub("_", "-", cols)
            
            if (all(cols_alt1 %in% rownames(md))) {
                md2 <- md[cols_alt1, , drop = FALSE]
                rownames(md2) <- cols
            } else if (all(cols_alt2 %in% rownames(md))) {
                md2 <- md[cols_alt2, , drop = FALSE]
                rownames(md2) <- cols
            } else {
                stop("Could not match colnames(pseudobulk[['RNA']]$data) to rownames(pseudobulk@meta.data).")
            }
        }
        
        # keep only requested annotation columns (that exist)
        ann_cols_use <- intersect(ann_cols, colnames(md2))
        annotation_col <- md2[, ann_cols_use, drop = FALSE]
        
        # palettes: user-supplied or empty
        if (!is.null(ann_palettes)) {
            # keep only palettes for columns we actually use
            annotation_colors <- ann_palettes[names(ann_palettes) %in% ann_cols_use]
        }
    }
    
    # Heatmap 1: normalized rows sum to 1
    top_mat_norm <- mat_norm[top_genes, , drop = FALSE]
    ColorRamp <- colorRampPalette(RColorBrewer::brewer.pal(9, "Blues"))(50)
    
    pheatmap::pheatmap(
        top_mat_norm,
        color = ColorRamp,
        breaks = seq(0, 1, length.out = length(ColorRamp) + 1),
        legend_breaks = seq(0, 1, by = 0.1),
        legend_labels = seq(0, 1, by = 0.1),
        main = "rows sum to 1",
        annotation_col = annotation_col,
        annotation_colors = annotation_colors
    )
    
    # Heatmap 2: raw expression log1p CP10K
    pheatmap::pheatmap(
        mat[top_genes, , drop = FALSE],
        color = cols_heatmap_gradient,
        main = "expression log1p CP10K",
        annotation_col = annotation_col,
        annotation_colors = annotation_colors
    )
    
    # Heatmap 3: raw expression, no column clustering
    pheatmap::pheatmap(
        mat[top_genes, , drop = FALSE],
        color = cols_heatmap_gradient,
        cluster_cols = FALSE,
        main = "expression log1p CP10K (cols fixed)",
        annotation_col = annotation_col,
        annotation_colors = annotation_colors
    )
    
    # Heatmap 4: raw expression with externally supplied column order
    if (!is.null(cl_order_file) || !is.null(cl_order)) {
        
        if (is.null(cl_order)) {
            cl_order <- read.table(
                file = cl_order_file,
                header = FALSE,
                stringsAsFactors = FALSE
            )[, 1]
        }
        
        # Robust matching in case "-" vs "_" differs
        cl_order_match <- cl_order
        
        if (length(intersect(cl_order_match, colnames(mat))) == 0) {
            cl_order_alt1 <- gsub("-", "_", cl_order)
            cl_order_alt2 <- gsub("_", "-", cl_order)
            
            if (length(intersect(cl_order_alt1, colnames(mat))) > 0) {
                cl_order_match <- cl_order_alt1
            } else if (length(intersect(cl_order_alt2, colnames(mat))) > 0) {
                cl_order_match <- cl_order_alt2
            }
        }
        
        ordered_cols <- c(
            cl_order_match[cl_order_match %in% colnames(mat)],
            setdiff(colnames(mat), cl_order_match)
        )
        
        annotation_col_ordered <- annotation_col
        if (!is.null(annotation_col_ordered)) {
            annotation_col_ordered <- annotation_col_ordered[ordered_cols, , drop = FALSE]
        }
        
        pheatmap::pheatmap(
            mat[top_genes, ordered_cols, drop = FALSE],
            color = cols_heatmap_gradient,
            cluster_cols = FALSE,
            cluster_rows = TRUE,
            main = "expression log1p CP10K (external column order)",
            annotation_col = annotation_col_ordered,
            annotation_colors = annotation_colors
        )
    }
    
    dev.off()
    
}

pseudobulk_orig = readRDS('/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/analysis/data_integration/IGT1_96/TF_analysis/20260319/igt1_96_pseudobulk_byclusterannotationlevel2_log1pCP10K.Rds')
pseudobulk_orig@meta.data %>% count(annotation_level2)
pseudobulk_orig$annotation_level1 = sub("\\..*", "", pseudobulk_orig$annotation_level2) #so_orig$annotation_level1[match(pseudobulk_orig$annotation_level2,so_orig$annotation_level2)] %>% as.character()
pseudobulk_orig$annotation_level2_group = so_orig$annotation_level2_group[match(pseudobulk_orig$annotation_level2,so_orig$annotation_level2)] %>% as.character()

lineages_keep <- c("gdT", "CD8aa", "Tz", "DN", "DP")

ann_cols <- c("annotation_level1", "annotation_level2_group")
ann_palettes <- list(
  annotation_level1 = mypal_level1,
  annotation_level2_group = c(resting = "blue", activated = "red", other = "black")
)

df = data.frame(pseudobulk_orig$annotation_level1, pseudobulk_orig$annotation_level2)
df$mypal_level1b = mypal_level1[pseudobulk_orig$annotation_level1]
mypal_level1b = setNames(df$mypal_level1b, nm = pseudobulk_orig$annotation_level2)

pseudobulk = pseudobulk_orig[,pseudobulk_orig$annotation_level1 %in% lineages_keep]

curated_list = read.table('/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/analysis/data_integration/IGT1_96/EffMol_analysis/20260107/effmol_curated.txt', header = F, skip = 0, sep = "\t")[,1]
out_dir2 = sprintf("%s/%s", out_dir, "level2CuratedList")
cl_order = read.table(file = sprintf("%s/Heatmap_correlation_clusters_top%dgenes_%dPCs_CLUSTERORDER.txt", out_dir, length(top_genes), npcs), header = F)[,1]
Gene_AUC2(pseudobulk,
         genes = curated_list, # vector of genes
         out_dir = out_dir2,
         mypal = mypal,              # named vector of colors for TF lines
         order_groups = levels(so_orig$annotation_level2),
         mypal_groups = mypal_level1b,       # named vector of colors for groups in barplots
         n_top = length(curated_list),                # number of top TF to highlight
         expr_threshold = 0.5,       # threshold for filtering genes, in CK10K
         cols_heatmap_gradient = colorRampPalette(c('white','red','brown'))(100), #rev(colorRampPalette(brewer.pal(n = 7,name = "RdYlBu"))(100))  rev(viridis::rocket(100))
         ann_cols = ann_cols,
         ann_palettes = ann_palettes,
                 cl_order = cl_order,
         pdf_width = 15,
         pdf_height = 8
)


```

Plot correlation heatmap based on proportions in samples    
1. filter for  c("gdT", "CD8aa", "DN", "DP", "Tz") and grepl("allT", target_cells_simplified))
2. Calculate correlation in % between clusters across samples. Only keep correlation where there is at least 10 cells in each sample for comparison
3. Now make a correlation heatmap

```{r}

library(dplyr)
library(stringr)
library(ComplexHeatmap)
library(circlize)
library(grid)

prop <- readRDS(
  "/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My Drive/ImmgenT/analysis/data_integration/IGT1_96/SampleComp/20260206/annotation_level2_PropPerSample_withMetadata.Rds"
)

lineages_keep <- c("gdT", "CD8aa", "DN", "DP", "Tz")

# 1. Filter allT samples
prop_filt <- prop %>%
  filter(grepl("allT", target_cells_simplified))

# 2. Select relevant cluster prop/ncells columns
prop_cols <- grep(
  paste0("^(", paste(lineages_keep, collapse = "|"), ")\\..*\\.prop$"),
  colnames(prop_filt),
  value = TRUE
)

ncells_cols <- sub("\\.prop$", ".ncells", prop_cols)

# keep only columns where matching ncells column exists
keep <- ncells_cols %in% colnames(prop_filt)
prop_cols <- prop_cols[keep]
ncells_cols <- ncells_cols[keep]

cluster_names <- sub("\\.prop$", "", prop_cols)

# proportions as percent
prop_mat <- as.matrix(prop_filt[, prop_cols, drop = FALSE]) * 100
ncells_mat <- as.matrix(prop_filt[, ncells_cols, drop = FALSE])

colnames(prop_mat) <- cluster_names
colnames(ncells_mat) <- cluster_names

rownames(prop_mat) <- prop_filt$IGTHT
rownames(ncells_mat) <- prop_filt$IGTHT

##compute pairwise correlations using only samples where both clusters have at least 10 cells.
min_cells <- 1
min_samples <- 1

cor_mat <- matrix(
  NA_real_,
  nrow = ncol(prop_mat),
  ncol = ncol(prop_mat),
  dimnames = list(colnames(prop_mat), colnames(prop_mat))
)

n_used_mat <- matrix(
  0,
  nrow = ncol(prop_mat),
  ncol = ncol(prop_mat),
  dimnames = list(colnames(prop_mat), colnames(prop_mat))
)

for (i in seq_len(ncol(prop_mat))) {
  for (j in seq_len(ncol(prop_mat))) {
    
    keep_samples <- ncells_mat[, i] >= min_cells &
      ncells_mat[, j] >= min_cells &
      is.finite(prop_mat[, i]) &
      is.finite(prop_mat[, j])
    
    n_used <- sum(keep_samples)
    n_used_mat[i, j] <- n_used
    
    if (n_used >= min_samples) {
      cor_mat[i, j] <- cor(
        prop_mat[keep_samples, i],
        prop_mat[keep_samples, j],
        method = "pearson"
      )
    }
  }
}

##Add lineage and plot
col_fun <- circlize::colorRamp2(
  c(-1, 0, 1),
  c("#4575B4", "white", "#D73027")
)

cl_order = read.table(file = sprintf("%s/Heatmap_correlation_clusters_top%dgenes_%dPCs_CLUSTERORDER.txt", out_dir, length(top_genes), npcs), header = F)[,1]

lineage <- sub("\\..*$", "", cl_order)
lineage <- factor(lineage, levels = lineages_keep)

ha <- HeatmapAnnotation(
  lineage = lineage,
  col = list(
    lineage = mypal_level1[names(mypal_level1) %in% lineages_keep]
  )
)

pdf(sprintf("%s/Heatmap_correlation_clusters_AcrossSamples.pdf", out_dir, length(top_genes), npcs), width = 7.5, height = 7, useDingbats = F)
Heatmap(
  cor_mat[cl_order,cl_order],
  name = "Pearson\ncorrelation",
  col = col_fun,
  na_col = "grey90",
  top_annotation = ha,
  cluster_rows = F,
  cluster_columns = F,
  # row_split = lineage,
  # column_split = lineage,
  show_row_names = TRUE,
  show_column_names = TRUE,
  row_names_gp = gpar(fontsize = 7),
  column_names_gp = gpar(fontsize = 7),
  column_names_rot = 45,
  column_title = paste0(
    "Correlation of gdT, CD8aa, DN, DP, Tz clusters\n",
    "allT samples; pairwise samples require ≥", min_cells, " cells in both clusters"
  )
)
dev.off()

ht_drawn <- draw(ht)
```



