---
title: "worshop_20241028"
output: html_document
date: "2024-10-28"
editor_options: 
  chunk_output_type: console
---

**Start interactive session on Midway3**
```{bash}
ssh -Y zemmour@midway3.rcc.uchicago.edu
sinfo -s
screen -S david
sinteractive --exclusive --time=12:00:00 --account=pi-zemmour --partition=jfkfloor2 --qos=jfkfloor2 --nodes=1 
sinteractive --time=12:00:00 --account=pi-zemmour --partition=caslake --nodes=1 --mem=96GB
sinteractive --time=3:00:00 --account=pi-zemmour --partition=beagle3 --gres=gpu:1 --mem=128GB
sinteractive --time=12:00:00 --account=pi-zemmour --partition=beagle3 --gres=gpu:1 --mem=96GB

cd $LABHOME

module load python/anaconda-2022.05 
module load hdf5/1.12.0 #for BPCells but also others
module load openblas/0.3.13
#sinteractive --time=1:00:00 -p gpu --gres=gpu:1 --account=pi-zemmour 
#source activate /project/jfkfloor2/zemmourlab/david/envs/scvi
source activate /project/zemmour/david/envs/scvi_20240315 #for R and old scvi
module load openblas/0.3.13 #load again or error

module load python
source activate /project/zemmour/david/envs/scvi120_20241008 #for scvi 1.2.0

#cd /project/jfkfloor2/zemmourlab/david/immgent/analysis/integration/IGT1_56
cd /project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96

R
options(max.print=1000)
#options(Seurat.object.assay.version = 'v5')
setwd("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96")


```

**Initialize: load libraries and custom functions**

```{r}
options(max.print=1000)
setwd("/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96")
source("/project/jfkfloor2/zemmourlab/david/immgent/immgent_integration_git/custom_functions_david.R")

libs = c("fgsea","fastTopics", "flashier", "Matrix", "Seurat","BPCells", "ggplot2", "dplyr", "reshape2", "ggrastr", "RColorBrewer", "pals", "scales", "pheatmap") 
#libs = c("matrixStats", "gplots","ggplot2", "reshape2", "scales", "gridExtra", "dplyr", "RColorBrewer", "grid", "Rtsne", "limma",   "RColorBrewer", "pheatmap", "Seurat",  "ggrastr", "ggbeeswarm") #pwr, preprocessCore, "readxl", vegan", "genefilter" "scde" "cellrangerRkit", "Rsamtools", "GenomicRanges", "GenomicAlignments", "VGAM", "WGCNA",rafalib, ‘UsingR’,MAST "ggrastr",
sapply(libs, function(x) suppressMessages(suppressWarnings(library(x, character.only = TRUE, quietly = T, warn.conflicts  = F))))


#Gradients
library(RColorBrewer)
ColorRamp = rev(colorRampPalette(brewer.pal(n = 7,name = "RdYlBu"))(100))
ColorRamp = rev(colorRampPalette(c('red','white','blue'))(20))
ColorRamp = rev(rainbow(10, end = 4/6))
library(viridis)
ColorRamp = rev(viridis(100))
ColorRamp = rev(cividis(100))
#image(1:100, 1, as.matrix(1:100), col = ColorRamp, main = "Color Palette", xaxt = 'n', yaxt = 'n', xlab = "", ylab = "")

#mycol = c("black","blue","red","darkgreen", "orange","purple", "cyan", "greenyellow",    "salmon", "magenta","pink", "tan", "brown") # c("magenta", "red",  "darkgreen", "cyan", "blue", "blue", "blue", "black", "blue", "blue", "blue", "blue", "orange")

#Large Color palette
library("pals")
n = 70
qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
mypal1 = unique(unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals))))
mypal1 = mypal1[-4]
mypal = c(glasbey(), polychrome(), mypal1)
names(mypal) = NULL
#pal.bands(mypal, main="Colormap suggestions")

#parade = function(n) { return(Seurat::DiscretePalette(n, palette = "parade", shuffle = F)) }


ConvertS5toS3 = function(so, assay1 = "RNA", assay2 = "ADT") {
    so.v3 = CreateAssayObject(counts = so[[assay1]]$counts)
    so.v3 = CreateSeuratObject(so.v3)
    so.v3[[assay2]] = CreateAssayObject(counts = so[[assay2]]$counts) #samples@assays$ADT$counts
    #print(all(colnames(so.v3@assays$RNA$counts) == colnames(so.v3@assays$ADT$counts)))
    #all(colnames(so.v3) == colnames(so))
    so.v3@meta.data = so@meta.data
    return(so.v3)
}

PlotsAfterIntegration = function (seurat_object = so, dim1 = so@reductions$umap@cell.embeddings[,1], dim2 = so@reductions$umap@cell.embeddings[,2], split_by = "IGT", genes = c("Foxp3", "Il2ra"), cluster_key = "ClusterSCVI_Res") {
    
    so = seurat_object
    so@meta.data[,"split_by"] = so@meta.data[,split_by]
    
    alpha = 0.5
    sample_name_colors = mypal[1:length(unique(so@meta.data[,split_by]))]
    names(sample_name_colors) = levels(so$sample_name)
    sample_name_colors2 = sample_name_colors
    sample_name_colors2[grepl("WT", names(sample_name_colors))] = "grey"
    
    message("Plot 1: UMAP")
    p1 = ggplot(data.frame(so@meta.data, dim1 = dim1, dim2 = dim2)) + geom_point_rast(aes(dim1, dim2, color = split_by), alpha = I(alpha), raster.dpi = 100) + theme_bw() + scale_color_manual(values = mypal)
    print(p1)
    
    message("Plot 2: UMAP split")
    tmp = data.frame(so@meta.data, dim1 = dim1, dim2 = dim2)
    bkgrd = data.frame(dim1 = dim1, dim2 = dim2)
    
    p = ggplot(bkgrd) + geom_point_rast(aes(dim1, dim2), color = "grey", size = 0.1, alpha = 0.2, raster.dpi = 50)
    p2 = geom_point(data = tmp, aes(dim1, dim2, color = split_by), size = 1,  alpha = alpha) 
    
    q = p + p2 + scale_color_manual(values = sample_name_colors)  + theme_bw() + facet_wrap(~ split_by)
    print(q)
    
    message("Plot 3: UMAP genes")
    
    for (g in genes) {
        print(g)
        tmp = data.frame(so@meta.data, dim1 = dim1, dim2 = dim2, size = so@assays$RNA$counts[rownames(so@assays$RNA$counts) %in% g,])
        bkgrd = data.frame(dim1 = dim1, dim2 = dim2)
        
        p = ggplot(bkgrd) + geom_point_rast(aes(dim1, dim2), color = "grey", size = 0.1, raster.dpi = 50)
        p2 = geom_point(data = tmp, aes(dim1, dim2, color = size > 0, size = size,  alpha = size > 0)) 
        #p2 =  geom_point(data = tmp2, aes(dim1, dim2, color = size > 0, size = size), alpha = I(alpha))  + scale_color_manual
        
        p3 = p + p2  + scale_color_manual(values = c("black", "red")) + scale_alpha_manual(values = c(0.5,1))  + theme_bw() + facet_wrap(~split_by) + ggtitle(g)
        print(p3)
    }
    
    
    message("Plot 4: UMAP Clusters") 
    for (i in colnames(so@meta.data)[grepl("cluster_key", colnames(so@meta.data))]) {
        print(i)
        p = ggplot(data.frame(so@meta.data, dim1 = dim1, dim2 = dim2)) + 
            geom_point_rast(aes(dim1, dim2), color = so@meta.data[,i], alpha = I(alpha), raster.dpi = 100) + scale_color_manual(values = mypal) + ggtitle(i) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
        print(p + facet_wrap(~split_by))
    }
    
}


library(org.Hs.eg.db)
hs = org.Hs.eg.db
go2eg = as.list(org.Hs.egGO2ALLEGS)
symbols = select(hs, keys = unique(unlist(go2eg)), columns = "SYMBOL", keytype = "ENTREZID")
cc = sort(symbols$SYMBOL[symbols$ENTREZID %in% unique(c(go2eg[["GO:0007049"]]))])
cellmembrane = sort(symbols$SYMBOL[symbols$ENTREZID %in% unique(unlist(list(go2eg[["GO:0009986"]], go2eg[["GO:0009897"]], go2eg[["GO:0016494"]], go2eg[["GO:0098552"]])))]) #go2eg[["GO:0016020"]]
#is_cellmembrane = rownames(cor_genes) %in% is_cellmembrane
tf =  sort(symbols$SYMBOL[symbols$ENTREZID %in% unique(go2eg[["GO:0003700"]])])
secretedmol = sort(symbols$SYMBOL[symbols$ENTREZID %in% unique(go2eg[["GO:0005615"]])])
effmol = sort(symbols$SYMBOL[symbols$ENTREZID %in% unique(go2eg[["GO:0005615"]])])
chemreceptor = symbols$SYMBOL[symbols$ENTREZID %in% unique(c(go2eg[["GO:0004950"]]))]

library(org.Mm.eg.db)
mm = org.Mm.eg.db
go2eg = as.list(org.Mm.egGO2ALLEGS)
symbols = select(mm, keys = unique(unlist(go2eg)), columns = "SYMBOL", keytype = "ENTREZID")
cc = sort(symbols$SYMBOL[symbols$ENTREZID %in% unique(c(go2eg[["GO:0007049"]]))])
cellmembrane = sort(symbols$SYMBOL[symbols$ENTREZID %in% unique(unlist(list(go2eg[["GO:0009986"]], go2eg[["GO:0009897"]], go2eg[["GO:0016494"]], go2eg[["GO:0098552"]])))]) #go2eg[["GO:0016020"]]
#is_cellmembrane = rownames(cor_genes) %in% is_cellmembrane
tf =  sort(symbols$SYMBOL[symbols$ENTREZID %in% unique(go2eg[["GO:0003700"]])])
effmol = sort(symbols$SYMBOL[symbols$ENTREZID %in% unique(go2eg[["GO:0005615"]])])
chemreceptor = symbols$SYMBOL[symbols$ENTREZID %in% unique(c(go2eg[["GO:0004950"]]))]

```

**Updated the metadata for all**
see metadta_update
--> /project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/igt1_96_withtotalvi20241028.Rds

**Add TCR to unconventional**
mistake with order of cells...

```{r}
so = readRDS("/project/jfkfloor2/zemmourlab/david/immgent/analysis/integration/IGT1_56/Treg/totalvi_igt1_56_allgenes_Treg_20240528_so.Rds")
unconv = so[,so$annotation_level1 == "unconventional"]

tcr = read.table(file = "/project/zemmour/david/ImmgenT/analysis/data_integration/IGT1_96/TCR/All.IGTs.Enhanced.Summary.Table.2024-10-11.csv", sep = ",", header = T,  as.is = T)

df = merge(x = unconv@meta.data, y = tcr[,grepl("alpha|beta|IGT.cellID", colnames(tcr))], by.x = "colnames", by.y = "IGT.cellID", all.x = T)
rownames(df) = df$colnames

unconv@meta.data = df

unconv$alpha.vgene[unconv$isNKT]

saveRDS(unconv, "igt1_96_unconventional_20241029_so.Rds")
```

**Revise organ column**

```{r}
m = read.table("/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My\ Drive/ImmgenT/20141024_workshop/CD8AB/table_clusters_proportion_in_samples.csv", header = T, sep = ",")

m$organ2 = do.call(rbind, strsplit(x = m$condition_detailed_organ, "_"))[,1]

m[is.na(m$sample_name),]
m = m[!is.na(m$sample_name),]

#organ_simplified
m$organ_simplified = m$organ2
m$organ_simplified[grep(pattern = "lymph node", m$organ2)] = "LN"
m$organ_simplified[grep(pattern = "skin|Skin", m$organ2)] = "skin"
m$organ_simplified[grep(pattern = "colon", m$organ2)] = "colon"
m$organ_simplified[grep(pattern = "duoden|ileal|ileum|jejun|small intestine", m$organ2)] = "small intestine"
m$organ_simplified[m$organ_simplified %in% c("CNS", "brain")] = "CNS"
m$organ_simplified[m$organ_simplified %in% c("lamina propria", "gut")] = "small intestine"
m$organ_simplified[grep(pattern = "salivary|submandibular", m$organ_simplified)] = "submandibular salivary gland"
m$organ_simplified = make.names(m$organ_simplified)

m[is.na(m$organ2),]
m[is.na(m$organ_simplified),]

m[m$organ2 == "NA",]
which(m$organ2 == "NA")
m[which(m$organ2 == "NA"),"organ_simplified"] = "skin"
m[which(m$organ2 == "NA"),"organ2"] = "skin"

table(m$organ2, m$organ_simplified)

m$organ = NULL
m$organ = m$organ2

write.table(m, file = "/Users/david/Library/CloudStorage/GoogleDrive-zemmour@uchicago.edu/My\ Drive/ImmgenT/20141024_workshop/CD8AB/table_clusters_proportion_in_samples_corrected.tsv", sep = "\t", row.names = F, col.names = T, quote = F)


```



**Reassign clusters**
Cluster 8 from CD8AB should go to the CD4 people
CD4: fakecd4.csv


```{r}



```

